### 2015-01-14 paper.tex
add paper

/Users/inari/Documents/Courses/Grammar Formalisms/paper.tex

### 2015-01-14 paper.tex
asdf

@@
-/Users/inari/Documents/Courses/Grammar Formalisms/paper.tex
+\documentclass[a4paper, 11pt]{article}
+\topmargin-2.0cm
+
+\usepackage{fancyhdr}
+\usepackage{lastpage}
+\usepackage{extramarks}
+\usepackage{pagecounting}
+\usepackage{color}
+\usepackage[round]{natbib}
+\usepackage{lipsum}
+\usepackage{enumerate}
+%\usepackage{tipa}
+%\usepackage{gb4e}
+\usepackage{graphicx}
+\usepackage{amsmath, amsthm}
+%\usepackage{colortbl}
+\usepackage{caption}
+\usepackage[a4paper]{geometry}
+\usepackage{courier}
+\usepackage{listings}
+\lstset{
+         basicstyle=\footnotesize\ttfamily, 
+         numberstyle=\tiny,          
+         numbersep=5pt,             
+         tabsize=2,                
+         extendedchars=true,      
+         breaklines=true,        
+         showspaces=false,      
+         showtabs=false,       
+         xleftmargin=17pt,
+         framexleftmargin=17pt,
+         framexrightmargin=5pt,
+         framexbottommargin=4pt,
+         showstringspaces=false 
+ }
+ \lstloadlanguages{
+         Haskell
+ }
+
+\usepackage{url}
+
+\advance\oddsidemargin-0.35in
+\advance\evensidemargin-0.65in
+\textheight9.5in
+\textwidth6.5in
+
+\newcommand{\hsp}{\hspace*{\parindent}}
+\definecolor{gray}{rgb}{0.4,0.4,0.4}
+
+\newcommand\blfootnote[1]{
+  \begingroup
+  \renewcommand\thefootnote{}\footnote{#1}
+  \addtocounter{footnote}{-1}
+  \endgroup
+}
+
+%\sectionfont{\large}
+
+\begin{document}
+\lstset{language=Haskell}
+\pagestyle{fancy}
+\lhead{\textcolor{gray}{Inari Listenmaa}}
+\chead{\textcolor{gray}{\bf Grammar Formalisms}}
+\rhead{\textcolor{gray}{LP2 2014}}
+\lfoot{\textcolor{gray}{}}
+\rfoot{\thepage}
+\renewcommand{\headrulewidth}{0.5pt} 
+\renewcommand{\footrulewidth}{0.5pt} 
+\fancyfoot[C]{\footnotesize \textcolor{gray}{}}
+
+\centerline{ {\Large \bf Constraint Grammar} }
+\vspace*{0.2cm}
+
+\section{Introduction}
+
+In this course paper, I describe the implementation of Constraint Grammar in Haskell, using a SAT solver to find a solution that fits all the constraints.
+
+First I describe the existing implementations, followed by my own SAT-based solution.
+
+\section{Constraint Grammar}
+
+Constraint grammar was first introduced by \cite{KarlssonTODO}. 
+It a tool for disambiguating output by morphological analyser.
+
+\section{Standard implementation(s)}
+
+The first implementation, CG-1, was introduced in \cite{KarlssonTODO}.
+Karlsson describes the implementation as ``ad hoc`` and ``without any deeper technical contributions to parsing theory``, and adds that ideally, CG should be implemented with finite state methods.
+CG-2, implemented by \cite{tapanainen1996}, used finite state technology. 
+The latest version, CG-3, was implemented by \cite{TODO} without finite state technology. The implementation introduced new properties, such as ordering of the rules.
+
+\begin{quote}
+[--] the Lisp, C, and C++ implementations of the Constraint Grammar Parser are all fairly ad hoc rule interpreters without any deeper technical contributions to parsing theory.
+These implementations are not based on the use of well understood and theoretically sound parsing algorithms.
+Rather, they could be characterized as situation-action parsing programs.
+\end{quote}
+
+\subsection{Workflow}
+
+All implementations (TODO check) go through the input many times and apply all possible rules each time. 
+After one pass, it may be possible to apply some rules that didn't fire the last time.
+
+\subsection{Resolving conflicts}
+
+For CG-1 and CG-2, ... sections? TODO
+
+In CG-3, the rules are ordered. In case of a conflict, the rule that is described first is given priority.
+
+\section{SAT based implementation}
+
+Unit propagation magic!
+
+\subsection{Data types}
+
+% \begin{verbatim}
+ \begin{lstlisting}
+data Rule = Remove [Tag] Condition | Select [Tag] Condition deriving (Show)
+
+-- | There is no special constructor for empty condition (ie. remove/select tag everywhere),
+-- | but `C _ []' is assumed to mean that.
+data Condition = C Position [Tag]
+               | NOT Condition
+               | AND Condition Condition
+               | OR Condition Condition deriving (Show)
+
+-- | Position can be exact or at least.
+-- | The meaning of numbers is 
+-- | *  0: word itself
+-- | * -n: to the left
+-- | *  n: to the right.
+data Position = Exactly Integer | AtLeast Integer deriving (Show,Eq,Read)
+ \end{lstlisting}
+%\end{verbatim}
+
+%If you want references on a separate page, uncomment the following command.
+%\clearpage
+
+%\begin{small}
+%\bibliographystyle{plainnat}
+%\bibliography{references}
+%\end{small}
+
+\end{document}

### 2015-01-14 paper.tex
more text

@@
  \begin{lstlisting}
-data Rule = Remove [Tag] Condition | Select [Tag] Condition deriving (Show)
+data Rule = Remove [Tag] Condition | Select [Tag] Condition
 
--- | There is no special constructor for empty condition (ie. remove/select tag everywhere),
--- | but `C _ []' is assumed to mean that.
 data Condition = C Position [Tag]
@@
                | AND Condition Condition
-               | OR Condition Condition deriving (Show)
+               | OR Condition Condition 
+\end{lstlisting}
 
--- | Position can be exact or at least.
--- | The meaning of numbers is 
--- | *  0: word itself
--- | * -n: to the left
--- | *  n: to the right.
-data Position = Exactly Integer | AtLeast Integer deriving (Show,Eq,Read)
+The data type for Rule is either remove or select a list of tags, with condition(s).
+For Condition, there are basic conditions (position and a list of context tags) and combined conditions, with constructors `NOT', `AND' and `OR'.
+For instance, the following rule is read as ``remove tag for any verb, if the lemma is \emph{bear} or there is a determiner or an article anywhere to the left''.
+
+\begin{lstlisting}
+verb = [V, V2, V3, VV, V2V, VS]
+Remove verb (OR (C (Exactly 0)  [Lem "bear"])
+                (C (AtLeast -1) [Det,Art])
+            )
 \end{lstlisting}
-%\end{verbatim}
+
+There is no special constructor for empty condition (ie. remove/select tag everywhere), but an empty tag list in a condition, i.e. \texttt{C \_ []} is assumed to mean that.
+
+The data type for position can be exact n places or at least n places.
+The number 0 means the word itself, negative number means \emph{n} positions to the left and positive \emph{n} positions to the right.
+
+\begin{lstlisting}
+data Position = Exactly Integer | AtLeast Integer
+ \end{lstlisting}
 

### 2015-01-21 paper.tex
major restructuring for applyRule

@@
 
-Unit propagation magic!
+SAT solving is based on a technique called unit propagation:
+a set of complex Boolean clauses is simplified, starting from single literals and working up to
+For instance, the following constraints describe rules for driving a car.
+
+We have a set of variables that can be either true or false:
+
+\begin{verbatim}
+can_drive
+has_driving_licence
+is_over_18
+is_drunk
+\end{verbatim}
+
+Next, we represent a possible situation in the world.
+In order to be allowed to drive, a person must have a driving licence and they must not be drunk.
+This is represented by the clause $can\_drive ∧\wedge has\_driving\_licence ∧\wedge \neg is\_drunk$.
+Furthermore, from the world knowledge, we know that in order to have a driving licence, one must be over 18.
+This is represented by the second clause.
+
+\begin{itemize}
+%\item $can\_drive \wedge has\_driving\_licence \wedge \neg is\_drunk$,
+%\item $has\_driving\_licence \wedge is\_over\_18$, 
+\item $\neg is\_over\_18 \vee is\_drunk$
+\item $is\_over\_18$
+\end{itemize}
+
+Here we have a possible situation in a world: someone is behaving in a way that makes you think of a hypothesis ``either they are underage or drunk''.
+This is represented by the clause $\neg is\_over\_18 \vee is\_drunk$.
+Just by this clause we can't solve the reason for the behaviour: it can be one of them or both.
+Then we get evidence that they are over 18, represented by the clause $is\_over\_18$.
+Thus we know that the part $\neg is\_over\_18$ of the is not true, so we can eliminate that from the clause.
+The solution is below: these variables must be true for the situation to be consistent.
+
+\begin{itemize}
+\item $is\_drunk$ = True
+\item $is\_over\_18$ = True
+\end{itemize}
 

### 2015-01-21 CG.hs

@@ -80,18 +80,19 @@ toLists :: Condition -> [[Condition]]
 --for OR, we make a list of lists, all in sequence.
 --for AND, we need to make them parallel and put in a form with OR as the first constructor
 --ie. AND (OR C1 C2) (C3) ---> OR (AND C1 C3) (AND C2 C3)
---for NOT, TODO
-toLists c = case c of
-    (C position tags)               -> [[c]]
+--just one level of nesting, should make this work for all inputs ... then again who writes millions of nested ands and ors
+--for NOT, rather make some rules to get a normal form? e.g.
+-- * NOT (OR a b)  = NOT a && NOT b
+-- * NOT (AND a b) = NOT a || NOT b
+toLists cond = case cond of
+    (C position tags)               -> [[cond]]
     (OR  c1           c2)           -> toLists c1 ++ toLists c2 
-    (NOT c1)                        -> toLists c1 --TODO
-    (AND c1@(AND _ _) c2@(AND _ _)) -> [toListsAnd c]
-    (AND c1@(NOT _ )  c2         )  -> error "do you even know what you're doing D:"
-    (AND c1           c2@(NOT _) )  -> error "do you even know what you're doing D:"
-    (AND c1           c2         )  -> map toListsAnd $ and' <$> simpleList c1 <*> simpleList c2  
-  where and' c1 c2 = AND c1 c2
-
-        toListsAnd (AND c1 c2) = concat (toLists c1 ++ toLists c2)
+    (NOT c1)                        -> toLists c1 --applyRule allows one NOT at top level
+    (AND c1@(AND _ _) c2@(AND _ _)) -> [toListsAnd cond]
+    (AND c1@(NOT _ )  c2         )  -> error "so complicated D:"
+    (AND c1           c2@(NOT _) )  -> error "do you even know what you're doing ;__;"
+    (AND c1           c2         )  -> map toListsAnd $ AND <$> simpleList c1 <*> simpleList c2  
+  where toListsAnd (AND c1 c2) = concat (toLists c1 ++ toLists c2)
         toListsAnd c@(C _ _)   = [c]
 
         simpleList (AND c1 c2) = simpleList c1 ++ simpleList c2
@@ -100,23 +101,14 @@ toLists c = case c of
[…]

### 2015-01-21 CG_SAT.hs

@@ -56,8 +56,8 @@ mkLiterals bs = go bs 1
 
 -- 1) If something is unambiguous to start with, anchor that
 anchor :: [Literal] -> [Boolean]
-anchor lits = [bool | ((int, tag), bool) <- lits, 
-                      isUniq int indices]
+anchor lits = [bool | ((ind,tag),bool) <- lits, 
+                      isUniq ind indices]
    where indices = map (\((int, _), _) -> int) lits
          isUniq x xs = length (findIndices (==x) xs) == 1
 
@@ -83,96 +83,85 @@ applyRule :: Rule -> [Literal] -> [Boolean]
 applyRule rule lits = 
   case rule of
 
--- empty condition, remove/select <tags> everywhere
-    (Remove tags c@(C _ [])) -> [Not bool | (_,bool) <- chosen tags]
-    (Select tags c@(C _ [])) -> [bool | (_,bool) <- chosen tags] ++ 
+-- a) empty condition, remove/select <tags> everywhere
+    (Remove tags c@(C _ [])) -> trace ("applyRule: " ++ show c) $ [Not bool | (_,bool) <- chosen tags]
+    (Select tags c@(C _ [])) -> trace ("applyRule: " ++ show c ++ "\nother  : " ++ show (other tags) ++ "\nchosen : " ++ (show $ chosen tags)) $ 
+                                [bool | (_,bool) <- chosen tags] ++ 
                                 [Not bool | (_,bool) <- other tags]
 
--- conditions must all hold
-    (Remove tags c@(AND c1 c2)) -> trace (show rule) $ applyRuleAnd rule (toCList c) lits
-    (Select tags c@(AND c1 c2)) -> trace (show rule) $ applyRuleAnd rule (toCList c) lits
+-- b) condition(s) must not hold 
+-- TODO nesting of NOTs with AND and OR -- is it even relevant? do linguists write horribly nested and confusing stuff?
+    (Remove tags c@(NOT c1)) -> map Not $ applyRules rule (toLists c) lits
+    (Select tags c@(NOT c1)) -> map Not $ applyRules rule (toLists c) lits
 
--- condition(s) must not hold
-    (Remove tags c@(NOT c1)) -> trace (show rule) $ applyRuleNot rule (toCList c) lits
-    (Select tags c@(NOT c1)) -> trace (show rule) $ applyRuleNot rule (toCList c) lits
+-- c) general case, there can be nested ANDs, ORs or just plain rules
[…]

### 2015-01-21 paper.tex
cleanup

--- 2015-01-14 paper.tex
@@
 \usepackage{caption}
+\usepackage[T1]{fontenc}
 \usepackage[a4paper]{geometry}
@@
 
-Unit propagation magic!
+SAT solving is based on a technique called unit propagation:
+a set of complex Boolean clauses is simplified, starting from unit
+clauses, which consist of just a single variable, and working up to a solution, where all variables in the set
+have a True or False value.
+
+Let us show this with a small example.
+We have a set of variables describing properties of people: \texttt{is\_under\_18},
+\texttt{has\_driving\_licence} and \texttt{is\_drunk}.
+
+% For instance, the following constraints describe rules for driving a car.
+% \begin{verbatim}
+% can_drive
+% has_driving_licence
+% is_under_18
+% is_drunk
+% \end{verbatim}
+
+% Next, we represent a possible situation in the world.
+In order to be allowed to drive, a person must have a driving licence and they must not be drunk.
+This is represented by the clause $can\_drive ∧\wedge has\_driving\_licence ∧\wedge \neg is\_drunk$.
+Furthermore, from the world knowledge, we know that in order to have a driving licence, one must be over 18.
+This is represented by the second clause.
+
+Clause 1 describes a possible situation in a world: someone is behaving in away that makes you think of a hypothesis ``either they are underage or
+drunk''. This is represented as \texttt{is\_under\_18} $\vee$
+\texttt{is\_drunk}. Just by this clause we can't solve the reason for the
+behaviour: it can be one of them or both.
+
+Clause 2 comes from the world knowledge that in order to have a
+driving licence, one must be over 18. The form of the clause is
+obtained by applying negation distributively over the more intuitive
+clause $\neg$ (\texttt{has\_driving\_licence} $\wedge$
+\texttt{is\_under\_18}): it is not possible to both have driving
+licence and be under 18.
+
+\begin{enumerate}
+\item \texttt{is\_under\_18} $\vee$ \texttt{is\_drunk}
+\item ($\neg$ \texttt{has\_driving\_licence}) $\vee$ ($\neg$ \texttt{is\_under\_18})
+\item \texttt{has\_driving\_licence}
+\end{enumerate}
+
+Clause 3 provides us some undeniable evidence that
+the person has a driving licence. By this clause, 
+we know that the variable \texttt{is\_under\_18} is not true.
+Knowing that, we can eliminate \texttt{is\_under\_18} from the first
+clause. 
+The solution is below: the three variables must have these values in order to
+satisfy the constraints.
+
+\begin{itemize}
+\item \texttt{is\_drunk} = True
+\item \texttt{is\_under\_18} = False
+\item \texttt{has\_driving\_licence} = True
+\end{itemize}
+
+SAT solving is being used for many applications where
 
@@
 
+\subsection{Complex conditions}
+Conditions constructed by OR are represented as a sequence, and AND as
+parallel. In Haskell, I simply use nested lists: everything inside the
+outer list is subject to OR, and within the lists AND. This is best
+demonstrated by an example:
+
+\begin{itemize}
+\item \texttt{OR (OR C1 C2) C3} becomes \texttt{[[C1], [C2], [C3]]} 
+\item \texttt{AND (AND C1 C2) C3} becomes \texttt{[[C1, C2, C3]]} 
+\item \texttt{AND (OR C1 C2) C3} becomes \texttt{[[C1,C3], [C2,C3]]}
+\end{itemize}
+
+\subsection{Conflicts}
+If you try to use conflicting rules, you get user error: mzero.
+
+\subsection{Example run}
+
+We show an example run for the sentence \emph{They are both happy and go to the house}. This is
+taken from a real output of a morphological analyser, and was
+initially ambiguous in the following ways:
+
+\begin{itemize}
+\item[] are
+\begin{itemize}
+\item[] \texttt{<''be'', V, Pres>}
+\item[] \texttt{<''are'', N, Sg>}
+\end{itemize}
+
+\item[] both
+\begin{itemize}
+\item[] \texttt{<''both'', CoordConj>}
+\item[] \texttt{<''both'',Det>}
+\item[] \texttt{<''both'',Pron>}
+\end{itemize}
+
+\item[] go
+\begin{itemize}
+\item[] \texttt{<''go'', V, Inf>}
+\item[] \texttt{<''go'', V, Pres>}
+\item[] \texttt{<''go'', N, Sg>}
+\end{itemize}
+
+\item[] to
+\begin{itemize}
+\item[] \texttt{<''to'', Prep>}
+\item[] \texttt{<''to'', Adv>}
+\end{itemize}
+
+\item[] house
+\begin{itemize}
+\item[] \texttt{<''house'', V, Inf>}
+\item[] \texttt{<''house'', V, Pres>}
+\item[] \texttt{<''house'', N, Sg>}
+\end{itemize}
+\end{itemize}
+
+\begin{verbatim}
+Tag sequence:
+1: [Lem "they",Pron,Subj,Nom,P3]
+2: [Lem "are",N,Sg]
+3: [Lem "both",Pron]
+4: [Lem "happy",Adj]
+5: [Lem "and",CoordConj]
+6: [Lem "go",V,Inf]
+6: [Lem "go",V,Pres]
+7: [Lem "to",Prep]
+8: [Lem "the",Det]
+9: [Lem "house",N,Sg]
+\end{verbatim}
+
 %If you want references on a separate page, uncomment the following command.

### 2015-01-21 CG_SAT.hs

@@ -83,25 +83,46 @@ applyRule :: Rule -> [Literal] -> [Boolean]
 applyRule rule lits = 
   case rule of
 
--- a) empty condition, remove/select <tags> everywhere
-    (Remove tags c@(C _ [])) -> trace ("applyRule: " ++ show c) $ [Not bool | (_,bool) <- chosen tags]
-    (Select tags c@(C _ [])) -> trace ("applyRule: " ++ show c ++ "\nother  : " ++ show (other tags) ++ "\nchosen : " ++ (show $ chosen tags)) $ 
-                                [bool | (_,bool) <- chosen tags] ++ 
-                                [Not bool | (_,bool) <- other tags]
-
--- b) condition(s) must not hold 
+-- a) condition(s) must not hold 
 -- TODO nesting of NOTs with AND and OR -- is it even relevant? do linguists write horribly nested and confusing stuff?
     (Remove tags c@(NOT c1)) -> map Not $ applyRules rule (toLists c) lits
     (Select tags c@(NOT c1)) -> map Not $ applyRules rule (toLists c) lits
 
--- c) general case, there can be nested ANDs, ORs or just plain rules
+-- b) general case, there can be nested ANDs, ORs or just plain rules
     (Remove tags c) -> applyRules rule (toLists c) lits
     (Select tags c) -> applyRules rule (toLists c) lits
 
-  where -- chosen is simple: just get tags' that are in tags
-        chosen tags = filter (\((ind,tags'),bool) -> tags' `multiElem` tags) lits
+  -- where -- chosen is simple: just get tags' that are in tags
+  --       chosen tags = filter (\((ind,tags'),bool) -> tags' `multiElem` tags) lits
 
         -- other must filter tags' that are not in tags, but not from all lits:
+        -- just from the words that have somewhere an analysis which is in tags
+  --       other tags  = filter (\((ind,tags'),bool) -> tags' `multiNotElem` tags) (allWithReading tags)
+
+  --       -- list of positions in the sentence, where one of the analyses is in tags. e.g.
+  --       --      lits = (1,[N,Pl]), (2,[V,Sg]), (2,[N,Pl], (3,[Pron]), tags = [N]
+  --       -- ====> [1,2] is returned.
+  --       chosenInds tags = [ind | ((ind,_),_) <- chosen tags]
+
+  --       -- all lemmas that have the desired reading in one of their analyses. e.g.
[…]

### 2015-01-21 paper.tex
more stuff

@@
 
-SAT solving is being used for many applications where
+SAT solving is used for many applications where constraints must be
+satisfied, such as software and hardware verification.
+
+\subsection{CG formulated in SAT}
+
+We make each analysis a variable. For instance, the fifth line denotes
+a hypothesis that the correct analysis of the 3rd item in the sentence
+is \texttt{<V,Sg>}. We call this \texttt{Var 5}.
+
+\begin{lstlisting}
+((1,[Lem "the",Det]),Var 1)
+((2,[Lem "bear",N,Sg]),Var 2)
+((2,[Lem "bear",V,Pl]),Var 3)
+((3,[Lem "sleep",N,Pl]),Var 4)
+((3,[Lem "sleep",V,Sg]),Var 5)
+((4,[Lem "in",Prep]),Var 6)
+((4,[Lem "in",Adv]),Var 7)
+((5,[Lem "the",Det]),Var 8)
+((6,[Lem "house",V,Inf]),Var 9)
+((6,[Lem "house",N,Sg]),Var 10)
+\end{lstlisting}
+
+We have defined our variables, but not yet any truth values. We can
+start by doing two things:
+\begin{enumerate}
+\item Make unambiguous tokens true by default: e.g. \texttt{Var 1}
+  shows that the reading \texttt{Det} for the first token must be true.
+\item Make hypotheses of possible n-grams. In the example below, we
+show bigrams: \texttt{(Var 1 $\wedge$ Var 2) $\vee$ (Var 1 $\wedge$ Var 3)}
+means that the first two tokens have analyses \texttt{Det,N} or
+\texttt{Det,V}.
+\end{enumerate}
+
+We could easily add a third rule: for ambiguous tokens, make sure that
+only one reading can be true (or a less strict version, where all
+readings can't be true). For instance, \texttt{Not (Var 2 $\wedge$ Var
+  3)} would be the constraint that the second token can't be both
+\texttt{N} and \texttt{V} at the same time. However,
+\cite{KarlssonTODO} argues that CG cannot remove genuine ambiguities;
+in a case where a sentence is genuinely ambiguous, all readings are
+returned.
+
+\begin{lstlisting}
+Var 1
+Var 8
+(Var 1 :&&: Var 2) :||: (Var 1 :&&: Var 3)
+(Var 2 :&&: Var 4) :||: ((Var 2 :&&: Var 5) :||: ((Var 3 :&&: Var 4) :||: (Var 3 :&&: Var 5)))
+(Var 4 :&&: Var 6) :||: ((Var 4 :&&: Var 7) :||: ((Var 5 :&&: Var 6) :||: (Var 5 :&&: Var 7)))
+(Var 6 :&&: Var 8) :||: (Var 7 :&&: Var 8)
+(Var 8 :&&: Var 9) :||: ((Var 8 :&&: Var 10))
+\end{lstlisting}
+% Not (Var 2 :&&: Var 3)
+% Not (Var 4 :&&: Var 5)
+% Not (Var 6 :&&: Var 7)
+% Not (Var 9 :&&: Var 10)
+
+Another property of CG is that cannot remove all readings: if all
+other analyses are eliminated, the remaining analysis must be
+returned, even if it were to be eliminated at the next pass. 
+However, this property isn't implemented in the SAT version.
 
@@
 
-% \begin{verbatim}
 \begin{lstlisting}

### 2015-01-21 CG_SAT.hs

@@ -11,6 +11,19 @@ import Debug.Trace
 
 type Literal = ((Integer, [Tag]), Boolean)
 
+--we want to often compare the indices
+sameInd :: Literal -> Literal -> Bool
+sameInd lit lit' = ind lit == ind lit'
+
+ind :: Literal -> Integer
+ind ((i,_),_) = i
+
+tag' :: Literal -> [Tag]
+tag' ((_,t),_) = t
+
+bool' :: Literal -> Boolean
+bool' (_,b) = b
+
 instance Eq Boolean where
   Var n      == Var m        = m==n
   Var n      == _            = False
@@ -45,44 +58,47 @@ mkLits = mkLiterals . mkSymbols
 
 mkSymbols :: [[a]] -> [(Integer,a)]
 mkSymbols as = concat $ go as 1
-    where go []     n = []
-          go (x:xs) n = map (\tag -> (n,tag)) x : go xs (n+1)
+   where go []     n = []
+         go (x:xs) n = map (\tag -> (n,tag)) x : go xs (n+1)
 
 mkLiterals :: [b] -> [(b,Boolean)]
 mkLiterals bs = go bs 1
-    where go []     n = []
-          go (x:xs) n = (x,Var n) : go xs (n+1)
+   where go []     n = []
+         go (x:xs) n = (x,Var n) : go xs (n+1)
 
 
 -- 1) If something is unambiguous to start with, anchor that
 anchor :: [Literal] -> [Boolean]
 anchor lits = [bool | ((ind,tag),bool) <- lits, 
                       isUniq ind indices]
-   where indices = map (\((int, _), _) -> int) lits
-         isUniq x xs = length (findIndices (==x) xs) == 1
+  where indices = map ind lits
+        isUniq x xs = length (findIndices (==x) xs) == 1
 
 -- 2) Take all possible bigrams
 mkBigrams :: [Literal] -> [Boolean]
 mkBigrams lits = zipWith mkCombs pres posts --[lits] [lits]
-    where len = length lits
-          pres = take (len-1) $ groupBy sameNumber lits
-          posts = tail $ groupBy sameNumber lits
-          sameNumber ((n1,_),_) ((n2,_),_) = n1==n2 
+   where len = length lits
+         pres = take (len-1) $ groupBy sameInd lits
+         posts = tail $ groupBy sameInd lits
+         
 
 mkCombs :: [Literal] -> [Literal] -> Boolean
[…]

### 2015-01-27 paper.tex
more stuff

@@
 
-Constraint grammar was first introduced by \cite{KarlssonTODO}. 
+Constraint Grammar (CG) was first introduced by \cite{KarlssonTODO}. 
 It a tool for disambiguating output by morphological analyser.
+
+Parsing is an action where an unstructured input is transformed into a
+structured output. In a sense, applying CG rules on a text shouldn't
+be considered as parsing: it doesn't really build a new structure,
+such as phrase structure tree or dependency tree.
+The grammatical information is already in the input, and applying the
+rules leaves (ideally) only the correct options. 
+It doesn't require that the text should be parsed as full units; it is
+neither top-down or bottom up. Rules can have a context of specified
+amount of words (e.g. 1 to the left; 2 to the right), unlimited amount
+of words (e.g. anywhere to the left; 2 to the right and further from
+that) or they can be limited by a barrier (e.g. anywhere to the right
+until we see an adverb).
+
+In the example below, we show possible tags for the sentence ``the bear sleeps''.
+
+\begin{lstlisting}
+<the>     Det
+<bear>    Noun, Sg
+          Verb, Pl, Pres
+          Verb, Inf
+<sleep>   Noun, Pl
+          Verb, Sg3, Pres
+\end{lstlisting}
+
+We could disambiguate this sentence with two rules: 
+\begin{enumerate}
+\item Select noun (or remove verb) after determiner;
+\item Select verb, if is preceded by noun and followed by a sentence boundary.
+\end{enumerate}
+
+We would go through the input twice. The first rule would fire at the
+first pass: we find a case where there is something tagged as
+\texttt{Det} followed by something with \texttt{Noun} among the
+analyses. We will apply the rule where we select the noun reading, and
+discard the two other options (present tense verb or nonfinite verb).
+
+Then, at the next pass, we have two unambiguous tokens,
+and we go through the input again, to see
+if the first pass has created chances for new rules to
+activate. Now we find a context, where there is an ambiguous word
+preceded by a noun (\emph{bear}) and followed by a sentence boundary,
+so we can apply rule 2. The finished analysis is as follows:
+
+\begin{lstlisting}
+<the>     Det
+<bear>    Noun, Sg
+<sleep>   Verb, Sg3, Pres
+\end{lstlisting}
+
+CG can also be applied to shallow syntax. After having disambiguated
+the text morphologically, we can apply possible syntactic labels. For
+instance, a noun can have a syntactic role of subject, object or
+predicative; and a verb in present tense 3rd person singular can be
+only the main verb (possibly combined with a conjunction, e.g. ``the
+bear eats and sleeps''). The next step is shown below; the analyses
+chosen in the previous step are shown in the angle brackets, and
+syntactic labels are fully capitalised and preceded by @.
+
+\begin{lstlisting}
+<the,  Det>                
+<bear, N, Sg>           @SUBJ
+                        @OBJ
+                        @PREDC
+<sleep, V, Sg3, Pres>   @MAINV
+\end{lstlisting}
 

### 2015-01-27 paper.tex
add ref to kimmo

@@
 \end{lstlisting}
+
+This set would be disambiguated by a syntactic constraint that would
+select subject if followed by a finite verb form.
+
+\cite{koskenniemi92} include also clause boundaries (TODO read about
+it), such as marking coordinated or subordinated clauses, or showing
+beginning and end of an embedded clause.
+%; so the parse would be complete if \emph{the} and \emph{bear} would belong to
+
+(c/p from koskenniemi paper)
+``Rules in the CG formalism are typically dedicated
+for one of the above tasks, and they are
+executed as successive groups.
+In finite-state syntax, rules are logically unordered.''
 

### 2015-01-28 paper.tex
added barrier rules

@@
 It a tool for disambiguating output by morphological analyser.
+CG can run on unrestricted text, given that the morphological analyser
+has large vocabulary.
 
@@
 \begin{lstlisting}
-<the>     Det
-<bear>    Noun, Sg
-          Verb, Pl, Pres
-          Verb, Inf
-<sleep>   Noun, Pl
-          Verb, Sg3, Pres
+the     <Det>
+bear    <Noun, Sg>
+        <Verb, Pl, Pres>
+        <Verb, Inf>
+sleep   <Noun, Pl>
+        <Verb, Sg3, Pres>
 \end{lstlisting}
@@
 \begin{lstlisting}
-<the>     Det
-<bear>    Noun, Sg
-<sleep>   Verb, Sg3, Pres
+the     <Det>
+bear    <Noun, Sg>
+sleep   <Verb, Sg3, Pres>
 \end{lstlisting}
+
+\subsubsection{Syntax}
+
+``Because syntax, just as morphological disambiguation, relies on
+elimination of alternatives, it is advantageous to introduce all
+simple syntactic instantiations at once. Readings not having syntactic
+labels after morphological analysis get them via morphosyntactic
+mapping.''
+
+e.g. it makes sense
+7: [Lem "they",Pron,Subj,Nom,P3]
+if we have a morph tag, like a pron, to label it already
+since we know that they can only be SUBJ
+and them can only be OBJ
 
@@
 bear eats and sleeps''). The next step is shown below; the analyses
-chosen in the previous step are shown in the angle brackets, and
-syntactic labels are fully capitalised and preceded by @.
+chosen in the previous step are shown in the angle brackets, and potential
+syntactic labels are fully capitalised and preceded by @. The sign >
+in the tag \texttt{@DN>} expresses head/modifier relation determiner of the next head nominal to the right
 
 \begin{lstlisting}
-<the,  Det>                
-<bear, N, Sg>           @SUBJ
+the     <Det>            @DN>    
+bear    <N, Sg>          @SUBJ
                          @OBJ
                          @PREDC
-<sleep, V, Sg3, Pres>   @MAINV
+sleep   <V, Sg3, Pres>   @MAINV
 \end{lstlisting}
@@
 
-\cite{koskenniemi92} include also clause boundaries (TODO read about
+\cite{koskenniemi92} include also 4 types of clause boundaries (TODO read about
 it), such as marking coordinated or subordinated clauses, or showing
@@
 
+\begin{quote}The whole finite-state grammar consists of a set of rules which
+constrain the possible choices of word Interpretations, tags and
+boundaries to only those which are considered grammatical.
+The entire grammar Is effectively equivalent to the (theoretical) intersection of all individual rule automata.
+However, such an intersection would be impractical to compute due to Its huge size.
+The logical task for any finite-state parser in the current approach is
+to compute the intersection of the unanalyzed sentence automaton and
+each rule automaton. Actual parsing can be done in several alternative
+ways which are guaranteed to yield the same result, but which vary in
+terms of efficiency.
+\end{quote}
+
 \subsection{Workflow}
@@
 
-Let us show this with a small example.
+Let us demonstrate this with a small example.
 We have a set of variables describing properties of people: \texttt{is\_under\_18},

### 2015-01-28 CG.hs

@@ -10,7 +10,7 @@ import Data.List
 -- | Lemma is a special type of tag: Lem String
 data Tag = 
    Art | Adj | Adv | Det | N | PN | V | V2 | VV 
- | Particle | Prep | Pron
+ | Particle | Prep | Pron | Punct
  | CoordConj | SubordConj
  | Sg | Pl | P1 | P2 | P3 
  | Subj | Imper | Cond | Inf | Pres
@@ -53,7 +53,10 @@ data Condition = C Position [Tag]
 -- | *  0: word itself
 -- | * -n: to the left
 -- | *  n: to the right.
-data Position = Exactly Integer | AtLeast Integer deriving (Show,Eq,Read)
+data Position = Exactly Integer 
+              | AtLeast Integer
+              | Fwd Integer [Tag]  --Barrier rules
+              | Bck Integer [Tag] deriving (Show,Eq,Read)
 
 
 
@@ -108,7 +111,7 @@ mkC str tags | last str == '*' = C (AtLeast $ (read . init) str) tags
 
 lemmaBear :: Condition
 lemmaBear = mkC "0" [Lem "bear"]
-always = (mkC "0" [])
+always = mkC "0" []
 
 
 -- Sets of tags
@@ -130,6 +133,9 @@ slPrepIfDet = Select [Prep] (mkC "1" det)
 andTest = Select verb (AND (mkC "-2" (Adj:verb)) (mkC "-1" conj) )
 notTest = Select verb (NOT (mkC "-1" [Prep]))
 notOrTest = Select verb (NOT (OR (mkC "-1" conj) (mkC "1" [Prep])))
+barTest = Select [CoordConj] (C (Fwd 1 [Punct]) [CoordConj])
+--barTest = Select [CoordConj] (C (Fwd 1 [Punct]) [Particle])
+
 
 
 

### 2015-01-28 CG_SAT.hs

@@ -147,16 +147,34 @@ getContext original chosen (c@(C p contextTags):cs) = trace ("getContext: " ++ s
                       (Exactly n) -> filter (hasContextTags . exactly n) chosen
                       (AtLeast n) -> filter (hasContextTags . atleast n) chosen
 
+                      (Fwd n bs)  -> filter (hasContextTags . barrier n bs) chosen
+
+                      --takes all analyses that are between n and barrier
+                      --this will become the context, and applyRules will check 
+                      --if there's anything in it that matches the select/remove tags
+                      --(Fwd n bs)  -> nub $ concatMap (barrier n bs) chosen
+                      (Bck n bs)  -> undefined
+
         --given word and n, returns list of words that are n places away in original sentence
         exactly :: Integer -> Literal -> [Literal]
-        exactly n ((ind,_),_) = [lit | lit@((ind',_),_) <- original, ind+n == ind']
+        exactly n ((ind,_),_) = [lit | lit@((ind',_),_) <- original, ind' == ind+n]
 
         --same but list of tags that are at least n places away
-        atleast n ((ind,_),_) = [lit | lit@((ind',_),_) <- original, ind+n >= ind']
+        atleast n ((ind,_),_) = [lit | lit@((ind',_),_) <- original, ind' >= ind+n ]
 
+        -- between m and n places away
+        between m n ((ind,_),_) = [lit | lit@((ind',_),_) <- original, ind+m <= ind' && ind' <= ind+n ]
 
         hasContextTag ((_,tags),_) = tags `multiElem` contextTags
         hasContextTags ts = or $ map hasContextTag ts
+
+        barrier n btags lit = --trace ("\nbarrier: " ++ show lit ++ "\nmindist: " ++ show mindist ++ "\n" ++ show (between n mindist lit)) $ 
+           if dists==[] then [] else between n mindist lit
+           where barinds = [ind | ((ind,tags),_) <- original, tags `multiElem` btags]
+                 dists   = map (\i -> i - getInd lit) barinds :: [Integer]
+                 mindist = minimum dists
+                 
+
            
 
 --True if any of the items in AS is in BS
@@ -170,18 +188,19 @@ multiNotElem as bs = and $ map (\a -> a `notElem` bs) as
 showTag :: (Show t, Num t) => ((t, [Tag]), Boolean) -> String
 showTag ((t,tags),_) = show t ++ ": " ++ show tags
 
[…]

### 2015-01-28 CG_data.hs
forgot to add this earlier >_>

@@ -1,0 +1,71 @@
+module CG_data where
+
+import CG
+
+-- Analyses
+the = [[Lem "the", Det]]
+bear = [Lem "bear", N,Sg] : [[Lem "bear", V,y] | y <- [Pl]]
+--bear = [Lem "bear", N,Sg] : [[Lem "bear", V,y,z,w] | y <- [Sg,Pl], z <- [P1,P2], w <- [Subj,Imper]]
+--bear = [Lem "bear", N,Sg] : [[Lem "bear", V,y,z] | y <- [Sg,Pl], z <- [P1,P2]]
+
+sleeps = [[Lem "sleep", N,Pl],
+          [Lem "sleep", V,Sg,P3]]
+
+sleep = [[Lem "sleep", N,Sg],
+         [Lem "sleep", V,Inf],
+         [Lem "sleep", V,Pres]]
+
+in_ = [[Lem "in", Prep],
+       [Lem "in", Adv]]
+
+they = [[Lem "they", Pron, Subj, Nom, P3]]
+
+are = [[Lem "be", V, Pres], [Lem "are", N, Sg]]
+
+both = [[Lem "both", CoordConj], [Lem "both", Det], [Lem "both", Pron]]
+
+happy = [[Lem "happy", Adj], [Lem "happy", Particle]] --just for test
+
+and_ = [[Lem "and", CoordConj]]
+
+go = [[Lem "go", V, Inf], [Lem "go", V, Pres], [Lem "go", N, Sg]]
+
+to = [[Lem "to", Prep],
+       [Lem "to", Adv]]
+
+house = [[Lem "house", V, Inf], [Lem "house", V, Pres], [Lem "house", N, Sg]]
+
+comma = [[Lem ",", Punct]]
+
+part = [[Lem "particletest", Particle]]
+
+-- Sentence
+ex0 = [ the --unambiguous
+      , bear
+      , sleeps 
+      , in_
+      , the
+      , house ]
+
+ex1 = [ they
+      , are
+      , both
+      , happy
+      , and_
+      , go
+      , to
+      , the
+      , house ]
+
+ex2 = [ both
+      , house
+      , and_
+      , bear
+      , sleep
+      , comma
+      , they
+      , go
+      , and_
+      , sleep ]
+
+exs = [ex2, ex1, ex0]

### 2015-01-29 paper.tex
some truly horrible code

@@
 
+\subsection{Workflow}
+
+Assume we have a fragment ``both houses and cars''.
+The word \emph{both} has analyses \texttt{CoordConj} and
+\texttt{Pron}, and \emph{houses} is ambiguous between \texttt{N} and \texttt{V}.
+We have in our rule set the following rules:
+
+\begin{enumerate}
+\item remove N after Pron
+\item select N after *Conj
+\item select CoordConj before another CoordConj
+\end{enumerate}
+
+In the previous implementations, neither of rules 1 and 2 would fire
+before we have disambiguated \emph{both} with rule 3.
+In this implementation, we are creating all clauses before giving
+them to the SAT solver. We express these constraints as implications:
+\begin{itemize}
+\item \emph{both} is \texttt{Pron} $\Rightarrow$ remove \texttt{N} for \emph{houses}
+\item \emph{both} is \texttt{CoordConj} $\Rightarrow$ select
+  \texttt{N} for \emph{houses}
+\end{itemize}
+
+For lemmas with only one reading, this leads to redundant clauses,
+such as \texttt{Not (Var 8) :||: Not (Var 8)}, but SAT solver will
+take care of them.
+
 \subsection{Data types}

### 2015-01-29 CG.hs

@@ -132,9 +132,10 @@ rmNounIfPron = Remove noun (mkC "-1" [Pron])
 slPrepIfDet = Select [Prep] (mkC "1" det)
 andTest = Select verb (AND (mkC "-2" (Adj:verb)) (mkC "-1" conj) )
 notTest = Select verb (NOT (mkC "-1" [Prep]))
+slNounAfterConj = Select noun ((mkC "-1" conj))
 notOrTest = Select verb (NOT (OR (mkC "-1" conj) (mkC "1" [Prep])))
 barTest = Select [CoordConj] (C (Fwd 1 [Punct]) [CoordConj])
---barTest = Select [CoordConj] (C (Fwd 1 [Punct]) [Particle])
+
 
 
 

### 2015-01-29 CG_SAT.hs

@@ -110,16 +110,29 @@ applyRule rule lits =
     (Select tags c) -> applyRules rule (toLists c) lits
 
 applyRules :: Rule -> [[Condition]] -> [Literal] -> [Boolean]
-applyRules rule [] lits       = []
+applyRules rule []         lits = []
 applyRules rule x@(xs:xxs) lits = trace ("\napplyRules: " ++ show x) $ applyRules rule xxs lits ++
   case rule of 
-    (Remove tags c) -> map (Not . getBool) (chosen tags)
-    (Select tags c) -> map getBool (chosen tags) ++ map (Not . getBool) (other tags)
+    -- (Remove tags c) -> map (Not . getBool) (chosen tags)
+    -- (Select tags c) -> map getBool (chosen tags) ++ map (Not . getBool) (other tags)
 
-  where contextN = getContext lits lits xs :: [Literal]
+     -- reason => consequence    translates into  Not reason | consequence.
+     -- for example,
+     -- prevIsPron => Not Noun  ===============>  Not prevIsPron | Not Noun
+    (Remove tags c) -> [(Not . getBool) reason :||: Not (head conseq)
+                          | reason <- chosen tags, 
+                            let (Just conseq) = lookup reason contextReasons ]       
+    (Select tags c) -> undefined
+
+
+  where 
+        origContext = getContext lits lits xs :: [Literal]
+        contextReasons = getReasonForContext origContext origContext xs :: [(Literal,[Boolean])]
+        (context,_) = unzip contextReasons 
 
         -- chosen is simple: just get tags' that are in tags
-        chosen tags = filter (\lit -> getTags lit `multiElem` tags) contextN :: [Literal]
+        chosen tags = filter (\lit -> getTags lit `multiElem` tags) context :: [Literal]
+        
 
         -- other must filter tags' that are not in tags, but not from all lits in contextN:
         -- just from the words that have somewhere an analysis which is in tags
@@ -131,7 +144,24 @@ applyRules rule x@(xs:xxs) lits = trace ("\napplyRules: " ++ show x) $ applyRule
         -- ====> chosen tags will return (2,[V,Sg)
         -- ====> (2,[V,Sg]) and (2,[N,Pl]) are returned
         allWithReading tags = intersectBy sameInd lits (chosen tags)
- 
+
+
+--TODO TODO TODO get rid of this horror and integrate it into getContext
+getReasonForContext :: [Literal] -> [Literal] -> [Condition] -> [(Literal,[Boolean])]
[…]

### 2015-01-29 CG.hs
getReason: Not (Var 1) :||: Var 1

@@ -134,7 +134,7 @@ andTest = Select verb (AND (mkC "-2" (Adj:verb)) (mkC "-1" conj) )
 notTest = Select verb (NOT (mkC "-1" [Prep]))
 slNounAfterConj = Select noun ((mkC "-1" conj))
 notOrTest = Select verb (NOT (OR (mkC "-1" conj) (mkC "1" [Prep])))
-barTest = Select [CoordConj] (C (Fwd 1 [Punct]) [CoordConj])
+slCCifCC = Select [CoordConj] (C (Fwd 1 [Punct]) [CoordConj])
 
 
 

### 2015-01-29 CG_SAT.hs

@@ -116,18 +116,24 @@ applyRules rule x@(xs:xxs) lits = trace ("\napplyRules: " ++ show x) $ applyRule
     -- (Remove tags c) -> map (Not . getBool) (chosen tags)
     -- (Select tags c) -> map getBool (chosen tags) ++ map (Not . getBool) (other tags)
 
-     -- reason => consequence    translates into  Not reason | consequence.
+     -- reason => consequence    translates into  Not reason || consequence.
      -- for example,
-     -- prevIsPron => Not Noun  ===============>  Not prevIsPron | Not Noun
-    (Remove tags c) -> [(Not . getBool) reason :||: Not (head conseq)
-                          | reason <- chosen tags, 
-                            let (Just conseq) = lookup reason contextReasons ]       
-    (Select tags c) -> undefined
+     -- -1Pron && +1Verb => Not Noun  =========>  Not (-1Pron && +1Verb) || Not Noun
+     -- SAT solver will transform it into conjunctive normal form
+    (Remove tags c) -> [Not (foldr1 (:&&:) reason) :||: (Not . getBool) conseq
+                          | conseq <- chosen tags, 
+                            let (Just reason) = lookup conseq contextReasons ]       
+    (Select tags c) -> [Not (foldr1 (:&&:) reason) :||: getBool conseq
+                          | conseq <- chosen tags, 
+                            let (Just reason) = lookup conseq contextReasons ]  
+                     ++ [Not (foldr1 (:&&:) reason) :||: (Not . getBool) conseq
+                          | conseq <- other tags, 
+                            let (Just reason) = lookup conseq contextReasons ]
 
 
   where 
         origContext = getContext lits lits xs :: [Literal]
-        contextReasons = getReasonForContext origContext origContext xs :: [(Literal,[Boolean])]
+        contextReasons = getReasonForContext lits origContext xs :: [(Literal,[Boolean])]
         (context,_) = unzip contextReasons 
 
         -- chosen is simple: just get tags' that are in tags
@@ -146,20 +152,30 @@ applyRules rule x@(xs:xxs) lits = trace ("\napplyRules: " ++ show x) $ applyRule
         allWithReading tags = intersectBy sameInd lits (chosen tags)
 
 
---TODO TODO TODO get rid of this horror and integrate it into getContext
+--TODO TODO TODO get rid of this horror
 getReasonForContext :: [Literal] -> [Literal] -> [Condition] -> [(Literal,[Boolean])]
[…]

### 2015-02-26 CG.hs
change CG_SAT to use MiniSat

@@ -160,7 +160,7 @@ slNounAfterConj = Select noun (mkT "-1" conj)
 
 slCCifCC = Select cnjcoo (POS (C (Barrier 1 [[Tag "punct"]]) (True,cnjcoo)))
 
-rmPlIfSg = Remove pl (POS (C (Exactly (-1)) (False,sg)))
+rmPlIfSg = Remove pl (POS (C (Exactly (-1)) (True,sg)))
 rmSgIfPl = Remove sg (mkT "-1" pl)
 
 negTest   = Select verb (neg (mkT "-1" prep))

### 2015-02-26 CG_SAT.hs

@@ -1,36 +1,36 @@
-{-# LANGUAGE ScopedTypeVariables #-}
----- ^ that's for the IO error
-
-module CG_SAT where 
+module Main where 
 
 
 import CG
 import CG_data
-import CG_parse
-import Data.Boolean.SatSolver
+import CG_parse (parseData, parseRules)
 import Data.List
+import Data.List.Split
 import Data.Maybe
 import Data.Set (Set, fromList, isSubsetOf)
 import Control.Monad
 import Control.Exception
+import MiniSat
+import SAT.SAT
 import System.Environment
+import System.IO.Unsafe
 import Debug.Trace
 
 
-type Literal = ((Integer, [Tag]), Boolean)
+type Token = ((Integer, [Tag]), Bit)
 
-getInd :: Literal -> Integer
+getInd :: Token -> Integer
 getInd ((i,_),_) = i
 
-getTags :: Literal -> [Tag]
+getTags :: Token -> [Tag]
 getTags ((_,t),_) = t
 
-getBool :: Literal -> Boolean
-getBool (_,b) = b
+getBit :: Token -> Bit
+getBit (_,b) = b
 
 --we want to often compare the indices
-sameInd :: Literal -> Literal -> Bool
-sameInd lit lit' = getInd lit == getInd lit'
+sameInd :: Token -> Token -> Bool
+sameInd tok tok' = getInd tok == getInd tok'
 
 atLeast :: Int -> [a] -> Bool
 atLeast 0 _      = True
@@ -41,69 +41,47 @@ atLeast n (_:xs) = atLeast (n-1) xs
 --Rule     has [[Tag]].
 --Analysis has [Tag].
 --At least one complete sublist in the rule must be found in the analysis.
-tagsMatchRule :: [[Tag]] -> Literal -> Bool
-tagsMatchRule tags lit = or $ map (\tagset -> isSubsetOf tagset tagsInAna) tagsInRule
-  where tagsInAna  = fromList $ getTags lit :: Set Tag
+tagsMatchRule :: [[Tag]] -> Token -> Bool
+tagsMatchRule tags tok = or $ map (\tagset -> isSubsetOf tagset tagsInAna) tagsInRule
+  where tagsInAna  = fromList $ getTags tok :: Set Tag
         tagsInRule = map fromList tags :: [Set Tag]
 
-instance Eq Boolean where
-  Var n      == Var m        = m==n
-  Var n      == _            = False
-  _          == Var n        = False
[…]

### 2015-02-26 CG_parse.hs

@@ -25,31 +25,31 @@ type Env = [(String,CG.TagSet)]
 
 
 
-getRules :: String -> IO [CG.Rule]
-getRules s = case pGrammar (ParCG.myLexer s) of
-            Bad err  -> do putStrLn "getRules: syntax error"
+parseRules :: String -> IO [CG.Rule]
+parseRules s = case pGrammar (ParCG.myLexer s) of
+            Bad err  -> do putStrLn "parseRules: syntax error"
                            putStrLn err
                            exitFailure 
-            Ok  tree -> do let rules = evalState (getCGRules tree) []
-                           mapM_ pr rules
+            Ok  tree -> do let rules = evalState (parseCGRules tree) []
+                           --mapM_ pr rules
                            return $ rights rules
   where pr (Right rule) = putStrLn $ show rule
         pr (Left string) = putStrLn string
 
-getData :: String -> IO [CG.Sentence]
-getData s = case pText (ParApertium.myLexer s) of
-            Bad err  -> do putStrLn "getData: syntax error"
+parseData :: String -> IO [CG.Sentence]
+parseData s = case pText (ParApertium.myLexer s) of
+            Bad err  -> do putStrLn "parseData: syntax error"
                            putStrLn err
                            exitFailure 
-            Ok text  -> do let anals = transText text
-                           mapM_ print anals
-                           return (split anals)
+            Ok text  -> do return $ (split . transText) text
+                           -- mapM_ print anas
+                           -- return (split anas)
 
 main :: IO ()
 main = do args <- getArgs
           case args of
-             [file1,file2] -> do readFile file1 >>= getRules
-                                 readFile file2 >>= getData
+             [file1,file2] -> do readFile file1 >>= parseRules
+                                 readFile file2 >>= parseData
                                  putStrLn "foo"
              _             -> do putStrLn "Usage: CG_parse <rules> <data>"
                                  exitFailure
@@ -57,17 +57,17 @@ main = do args <- getArgs
 
 ---
 
-getCGRules :: Grammar -> State Env [Either String CG.Rule]
-getCGRules (Defs defs) = do mapM updateEnv defs
-                            env <- get
-                            return $ map (getRules env) defs
+parseCGRules :: Grammar -> State Env [Either String CG.Rule]
+parseCGRules (Defs defs) = do mapM updateEnv defs
+                              env <- get
[…]

### 2015-02-26 SAT/SAT.hs

@@ -1,5 +1,6 @@
-module Main where
+module SAT.SAT where
 
+import CG (Tag(..))
 import MiniSat
 import Data.List ( nub )
 
@@ -22,6 +23,7 @@ addClauseBit :: Solver -> [Bit] -> IO ()
 addClauseBit s xs
   | true `elem` xs = do return ()
   | otherwise      = do addClause s [ x | Lit x <- xs ]
+                        --putStrLn ("addClauseBit: " ++show xs)
                         return ()
 
 solveBit :: Solver -> [Bit] -> IO Bool
@@ -133,7 +135,7 @@ maximizeFromTop s as rs =
 
 --------------------------------------------------------------------------------
 
-main =
+test =
   do s <- newSolver
      xs <- sequence [ newBit s | i <- [1..50] ]
      atMostOne s (take 20 xs)
@@ -146,3 +148,34 @@ main =
       else
        do return ()
 
+
+--------------------------------------------------------------------------------
+
+main' = 
+  do s <- newSolver
+     let names = [(1,Tag "det"),(2, Tag "noun"),(2, Tag "verb"),(3, Tag "noun"),(3,Tag "verb")]
+     lits <- sequence [ newBit s | i <- [1..5] ]
+     addClauseBit s [(lits !! 0)] --anchor unambiguous
+     addClauseBit s [nt (lits !! 0), nt (lits !! 2)] --Remove Verb IF -1 Det
+     b <- maximize s [] lits
+     print b
+     if b then
+       do bs <- sequence [ modelValueBit s x | x <- lits ]
+          putStrLn [ if b == Just True then '1' else '0' | b <- bs ]
+      else
+       do return ()
+     
+-- Literals:
+-- 0det, 1n, 1v, 2n, 2v
+
+-- Clauses:
+-- -- If something is unambiguous to start with, anchor that:
+
+-- 0det
+-- 0det & 1n | 0det & 1v
+-- 1n   & 2n | 1n   & 2v | 1v & 2n | 1v & 2v
+
+-- rules:
+--   Remove verb (C (-1) det)
+
+--   ~(x-1 det) | ~(x verb)

### 2015-02-26 paper/paper.tex
change directory structure

--- 2015-01-29 paper.tex
@@
 
+After adding this feature, the first rule will give the clause
+\texttt{Not (Var 7) :||: Not (Var 6)} and the second \texttt{Not (Var
+  7) :||: Var 6}.
+This is easily generalised to multiple contexts; \texttt{Not (Var
+  7 :&&: Var 8) :||: Var 6}.
+
 For lemmas with only one reading, this leads to redundant clauses,
@@
 take care of them.
+
+\subsubsection{Negation}
+
+Say we have a rule that inforces agreement: if token at n-1 is Sg,
+choose Sg for n as well.
+
+\begin{lstlisting}
+rmPlIfSg = Remove [Pl] (C (Exactly -1) [Sg])
+rmSgIfPl = Remove [Sg] (C (Exactly -1) [Pl])
+\end{lstlisting}
+
+We could also express the rules like this:
+\begin{lstlisting}
+rmPlIfSg = Remove [Pl] (NOT (C (Exactly (-1)) [Pl]))
+rmSgIfPl = Remove [Sg] (NOT (C (Exactly (-1)) [Sg]))
+\end{lstlisting}
+
+Although, this is bad: 
+\begin{lstlisting}
+((1,[Lem "the",Det]),Var 1)
+((2,[Lem "bear",N,Sg]),Var 2)
+((2,[Lem "bear",V,Pl]),Var 3)
+((3,[Lem "sleep",N,Pl]),Var 4)
+((3,[Lem "sleep",V,Sg,P3]),Var 5)
+\end{lstlisting}
+
+Negation removes both readings for token 2 ``bear'', because the first
+token doesn't have any number reading.
 

### 2015-02-26 CG_SAT.hs

@@ -3,7 +3,7 @@ module Main where
 
 import CG
 import CG_data
-import CG_parse (parseData, parseRules)
+import BNFC.CG_parse (parseData, parseRules)
 import Data.List
 import Data.List.Split
 import Data.Maybe

### 2015-02-26 bnfc/CG_parse.hs

@@ -1,4 +1,4 @@
-module CG_parse where
+module BNFC.CG_parse where
 
 import System.Environment (getArgs)
 import System.Exit (exitFailure)
@@ -10,7 +10,7 @@ import PrintApertium
 import AbsCG
 import LexCG
 import ParCG
-import qualified PrintCG as PrCG
+import PrintCG
 import ErrM
 
 import Control.Applicative
@@ -66,7 +66,7 @@ parseCGRules (Defs defs) = do mapM updateEnv defs
         updateEnv (RuleDef r) = return ()
 
         parseRules :: Env -> Def -> Either String CG.Rule
-        parseRules _ (SetDef  s) = Left $ PrCG.printTree s
+        parseRules _ (SetDef  s) = Left $ PrintCG.printTree s
         parseRules e (RuleDef r) = Right $ evalState (transRule r) e
 
 

### 2015-02-26 morpho.hs

@@ -1,47 +1,0 @@
-import PGF2
-import CG 
-import qualified Data.Map as Map
-import Data.List
-import Data.Either
-import System.Environment
-
--- | Just a test to mimic morphological analyser.
--- | Takes sentences, uses GF lookupMorpho and outputs all analyses.
-
-main = do
-  args <- getArgs
-  if length args /= 3 then putStrLn "Usage: ./morpho <sentences> <pgf> <lang>" else do
-      sentsFile <- readFile (args !! 0)
-      pgf <- readPGF (args !! 1)
-      cnc <- getConcr pgf (args !! 2)
-      let sents = lines sentsFile
-          ws = map words sents -- :: [[String]]
-          analyses = (map . map) (lookupMorpho cnc) ws
-          onlyS = (map . map) (filter (\(name, f:field, _) -> length (findIndices ('_'==) name) <= 1 && f == 's')) analyses -- :: [[[MorphoAnalysis]]]
-
-
-      mapM_ putStrLn (zipWith printWA (head ws) (head onlyS))
-
-
-
-{- Output:
-"<bear>"
-      "bear_V2" VInf
-      "bear_V" s VInf
-      "bear_N" s Sg Nom
--}
-printWA w as = "\"<" ++ w ++ ">\"\t" ++ (unwords $ zipWith hargle lemmas anals)
-   where anals = map snd3 as
-         lemmas = map fst3 as
-         hargle l a = "\n      " ++ '"':l ++ '"':' ':a
-
-fst3 (val,_,_) = val
-snd3 (_,val,_) = val
-
-
-getConcr :: (Monad m) => PGF -> String -> m Concr
-getConcr pgf lang =                                  --b is m Concr
-    maybe                         -- maybe :: b -> (a -> b) -> Maybe a -> b
-          (fail $ "Concrete syntax not found: "++show lang)  -- :: m Concr
-          return                                             -- :: a -> m Concr
-          (Map.lookup lang (languages pgf))                  -- :: Maybe Concr

### 2015-02-26 CG.hs
more changes

@@ -1,11 +1,9 @@
 module CG where
 
 import Control.Applicative
-import Data.Boolean.SatSolver
 import Data.List
 
 -- TODO:
--- * More complicated set operations for tag sets, not just conditions.
 -- * CG2/CG3 decisions
 
 

### 2015-02-26 CG_SAT.hs

@@ -3,7 +3,7 @@ module Main where
 
 import CG
 import CG_data
-import BNFC.CG_parse (parseData, parseRules)
+import CG_parse (parseData, parseRules)
 import Data.List
 import Data.List.Split
 import Data.Maybe

### 2015-02-26 CG_parse.hs

@@ -1,17 +1,16 @@
-module BNFC.CG_parse where
+module CG_parse where
 
 import System.Environment (getArgs)
 import System.Exit (exitFailure)
 
-import AbsApertium
-import LexApertium
-import ParApertium
-import PrintApertium
-import AbsCG
-import LexCG
-import ParCG
-import PrintCG
-import ErrM
+import BNFC.AbsApertium
+import BNFC.LexApertium
+import BNFC.ParApertium
+import BNFC.AbsCG
+import BNFC.LexCG
+import BNFC.ParCG
+import BNFC.PrintCG
+import BNFC.ErrM
 
 import Control.Applicative
 import Control.Monad.State.Lazy
@@ -24,26 +23,25 @@ import qualified CG
 type Env = [(String,CG.TagSet)]
 
 
+test = False
 
 parseRules :: String -> IO [CG.Rule]
-parseRules s = case pGrammar (ParCG.myLexer s) of
+parseRules s = case pGrammar (BNFC.ParCG.myLexer s) of
             Bad err  -> do putStrLn "parseRules: syntax error"
                            putStrLn err
                            exitFailure 
             Ok  tree -> do let rules = evalState (parseCGRules tree) []
-                           --mapM_ pr rules
+                           when test (mapM_ pr rules)
                            return $ rights rules
   where pr (Right rule) = putStrLn $ show rule
         pr (Left string) = putStrLn string
 
 parseData :: String -> IO [CG.Sentence]
-parseData s = case pText (ParApertium.myLexer s) of
+parseData s = case pText (BNFC.ParApertium.myLexer s) of
             Bad err  -> do putStrLn "parseData: syntax error"
                            putStrLn err
                            exitFailure 
             Ok text  -> do return $ (split . transText) text
-                           -- mapM_ print anas
-                           -- return (split anas)
 
 main :: IO ()
 main = do args <- getArgs
@@ -66,7 +64,7 @@ parseCGRules (Defs defs) = do mapM updateEnv defs
         updateEnv (RuleDef r) = return ()
 
         parseRules :: Env -> Def -> Either String CG.Rule
-        parseRules _ (SetDef  s) = Left $ PrintCG.printTree s
+        parseRules _ (SetDef  s) = Left $ BNFC.PrintCG.printTree s
         parseRules e (RuleDef r) = Right $ evalState (transRule r) e
 
 

### 2015-02-26 bnfc/AbsApertium.hs

@@ -1,0 +1,29 @@
+
+
+module BNFC.AbsApertium where
+
+-- Haskell module generated by the BNF converter
+
+
+
+
+newtype Punct = Punct String deriving (Eq,Ord,Show,Read)
+newtype Iden = Iden String deriving (Eq,Ord,Show,Read)
+data Text =
+   Lines [Line]
+  deriving (Eq,Ord,Show,Read)
+
+data Line =
+   Line Iden [Analysis]
+ | LinePunct Punct
+ | NoAnalysis Iden Iden
+  deriving (Eq,Ord,Show,Read)
+
+data Analysis =
+   Anal Iden [TagA]
+  deriving (Eq,Ord,Show,Read)
+
+data TagA =
+   TagA Iden
+  deriving (Eq,Ord,Show,Read)
+

### 2015-02-26 bnfc/ErrM.hs

@@ -1,0 +1,37 @@
+-- BNF Converter: Error Monad
+-- Copyright (C) 2004  Author:  Aarne Ranta
+
+-- This file comes with NO WARRANTY and may be used FOR ANY PURPOSE.
+module BNFC.ErrM where
+
+-- the Error monad: like Maybe type with error msgs
+
+import Control.Monad (MonadPlus(..), liftM)
+import Control.Applicative (Applicative(..), Alternative(..))
+
+data Err a = Ok a | Bad String
+  deriving (Read, Show, Eq, Ord)
+
+instance Monad Err where
+  return      = Ok
+  fail        = Bad
+  Ok a  >>= f = f a
+  Bad s >>= _ = Bad s
+
+instance Applicative Err where
+  pure = Ok
+  (Bad s) <*> _ = Bad s
+  (Ok f) <*> o  = liftM f o
+
+
+instance Functor Err where
+  fmap = liftM
+
+instance MonadPlus Err where
+  mzero = Bad "Err.mzero"
+  mplus (Bad _) y = y
+  mplus x       _ = x
+
+instance Alternative Err where
+  empty = mzero
+  (<|>) = mplus

### 2015-02-26 bnfc/LexApertium.hs

@@ -1,0 +1,367 @@
+{-# LANGUAGE CPP,MagicHash #-}
+{-# LINE 3 "LexApertium.x" #-}
+
+{-# OPTIONS -fno-warn-incomplete-patterns #-}
+{-# OPTIONS_GHC -w #-}
+module BNFC.LexApertium where
+
+
+
+import qualified Data.Bits
+import Data.Word (Word8)
+
+#if __GLASGOW_HASKELL__ >= 603
+#include "ghcconfig.h"
+#elif defined(__GLASGOW_HASKELL__)
+#include "config.h"
+#endif
+#if __GLASGOW_HASKELL__ >= 503
+import Data.Array
+import Data.Char (ord)
+import Data.Array.Base (unsafeAt)
+#else
+import Array
+import Char (ord)
+#endif
+#if __GLASGOW_HASKELL__ >= 503
+import GHC.Exts
+#else
+import GlaExts
+#endif
+alex_base :: AlexAddr
+alex_base = AlexA# "\xf8\xff\xff\xff\xd9\xff\xff\xff\xb4\xff\xff\xff\x00\x00\x00\x00\x2d\x00\x00\x00\x00\x00\x00\x00\xad\x00\x00\x00\x2d\x01\x00\x00\xad\x01\x00\x00\x6d\x02\x00\x00\x2d\x02\x00\x00\x00\x00\x00\x00\x23\x03\x00\x00\xfe\xff\xff\xff\x00\x00\x00\x00\x00\x00\x00\x00\xfc\x03\x00\x00"#
+
+alex_table :: AlexAddr
[…]

### 2015-02-26 bnfc/LexCG.hs

@@ -1,0 +1,371 @@
+{-# LANGUAGE CPP,MagicHash #-}
+{-# LINE 3 "LexCG.x" #-}
+
+{-# OPTIONS -fno-warn-incomplete-patterns #-}
+{-# OPTIONS_GHC -w #-}
+module BNFC.LexCG where
+
+
+
+import qualified Data.Bits
+import Data.Word (Word8)
+
+#if __GLASGOW_HASKELL__ >= 603
+#include "ghcconfig.h"
+#elif defined(__GLASGOW_HASKELL__)
+#include "config.h"
+#endif
+#if __GLASGOW_HASKELL__ >= 503
+import Data.Array
+import Data.Char (ord)
+import Data.Array.Base (unsafeAt)
+#else
+import Array
+import Char (ord)
+#endif
+#if __GLASGOW_HASKELL__ >= 503
+import GHC.Exts
+#else
+import GlaExts
+#endif
+alex_base :: AlexAddr
+alex_base = AlexA# "\xf8\xff\xff\xff\xdd\xff\xff\xff\x3c\x00\x00\x00\xbc\x00\x00\x00\x3c\x01\x00\x00\xbc\x01\x00\x00\x3c\x02\x00\x00\xbc\x02\x00\x00\x00\x00\x00\x00\x2d\x03\x00\x00\x00\x00\x00\x00\x9e\x03\x00\x00\x00\x00\x00\x00\x0f\x04\x00\x00\x0f\x05\x00\x00\x31\x00\x00\x00\xcf\x04\x00\x00\x00\x00\x00\x00\x10\x05\x00\x00\x00\x00\x00\x00\x51\x05\x00\x00\x00\x00\x00\x00\x92\x05\x00\x00\x92\x06\x00\x00\x93\x06\x00\x00\xd7\xff\xff\xff\xd6\x06\x00\x00\x16\x07\x00\x00\x56\x07\x00\x00\x96\x07\x00\x00\x56\x08\x00\x00\x16\x08\x00\x00\x00\x00\x00\x00\x16\x09\x00\x00\xd6\x08\x00\x00\x00\x00\x00\x00\xd6\x09\x00\x00\x96\x09\x00\x00\x00\x00\x00\x00\x8c\x0a\x00\x00\x82\x0b\x00\x00\x
[…]

### 2015-02-26 bnfc/ParApertium.hs

@@ -1,0 +1,565 @@
+{-# OPTIONS_GHC -w #-}
+{-# OPTIONS -fglasgow-exts -cpp #-}
+{-# OPTIONS_GHC -fno-warn-incomplete-patterns -fno-warn-overlapping-patterns #-}
+module BNFC.ParApertium where
+import BNFC.AbsApertium
+import BNFC.LexApertium
+import BNFC.ErrM
+import qualified Data.Array as Happy_Data_Array
+import qualified GHC.Exts as Happy_GHC_Exts
+
+-- parser produced by Happy Version 1.18.10
+
+newtype HappyAbsSyn  = HappyAbsSyn HappyAny
+#if __GLASGOW_HASKELL__ >= 607
+type HappyAny = Happy_GHC_Exts.Any
+#else
+type HappyAny = forall a . a
+#endif
+happyIn10 :: (Punct) -> (HappyAbsSyn )
+happyIn10 x = Happy_GHC_Exts.unsafeCoerce# x
+{-# INLINE happyIn10 #-}
+happyOut10 :: (HappyAbsSyn ) -> (Punct)
+happyOut10 x = Happy_GHC_Exts.unsafeCoerce# x
+{-# INLINE happyOut10 #-}
+happyIn11 :: (Iden) -> (HappyAbsSyn )
+happyIn11 x = Happy_GHC_Exts.unsafeCoerce# x
+{-# INLINE happyIn11 #-}
+happyOut11 :: (HappyAbsSyn ) -> (Iden)
+happyOut11 x = Happy_GHC_Exts.unsafeCoerce# x
+{-# INLINE happyOut11 #-}
+happyIn12 :: (Text) -> (HappyAbsSyn )
+happyIn12 x = Happy_GHC_Exts.unsafeCoerce# x
+{-# INLINE happyIn12 #-}
+happyOut12 :: (HappyAbsSyn ) -> (Text)
+happyOut12 x = Happy_GHC_Exts.unsafeCoerce# x
+{-# INLINE happyOut12 #-}
+happyIn13 :: ([Line]) -> (HappyAbsSyn )
+happyIn13 x = Happy_GHC_Exts.unsafeCoerce# x
+{-# INLINE happyIn13 #-}
+happyOut13 :: (HappyAbsSyn ) -> ([Line])
+happyOut13 x = Happy_GHC_Exts.unsafeCoerce# x
+{-# INLINE happyOut13 #-}
+happyIn14 :: (Line) -> (HappyAbsSyn )
[…]

### 2015-02-26 bnfc/ParCG.hs

@@ -1,0 +1,948 @@
+{-# OPTIONS_GHC -w #-}
+{-# OPTIONS -fglasgow-exts -cpp #-}
+{-# OPTIONS_GHC -fno-warn-incomplete-patterns -fno-warn-overlapping-patterns #-}
+module BNFC.ParCG where
+import BNFC.AbsCG
+import BNFC.LexCG
+import BNFC.ErrM
+import qualified Data.Array as Happy_Data_Array
+import qualified GHC.Exts as Happy_GHC_Exts
+
+-- parser produced by Happy Version 1.18.10
+
+newtype HappyAbsSyn  = HappyAbsSyn HappyAny
+#if __GLASGOW_HASKELL__ >= 607
+type HappyAny = Happy_GHC_Exts.Any
+#else
+type HappyAny = forall a . a
+#endif
+happyIn19 :: (String) -> (HappyAbsSyn )
+happyIn19 x = Happy_GHC_Exts.unsafeCoerce# x
+{-# INLINE happyIn19 #-}
+happyOut19 :: (HappyAbsSyn ) -> (String)
+happyOut19 x = Happy_GHC_Exts.unsafeCoerce# x
+{-# INLINE happyOut19 #-}
+happyIn20 :: (UIdent) -> (HappyAbsSyn )
+happyIn20 x = Happy_GHC_Exts.unsafeCoerce# x
+{-# INLINE happyIn20 #-}
+happyOut20 :: (HappyAbsSyn ) -> (UIdent)
+happyOut20 x = Happy_GHC_Exts.unsafeCoerce# x
+{-# INLINE happyOut20 #-}
+happyIn21 :: (Id) -> (HappyAbsSyn )
+happyIn21 x = Happy_GHC_Exts.unsafeCoerce# x
+{-# INLINE happyIn21 #-}
+happyOut21 :: (HappyAbsSyn ) -> (Id)
+happyOut21 x = Happy_GHC_Exts.unsafeCoerce# x
+{-# INLINE happyOut21 #-}
+happyIn22 :: (Signed) -> (HappyAbsSyn )
+happyIn22 x = Happy_GHC_Exts.unsafeCoerce# x
[…]

### 2015-02-26 bnfc/PrintApertium.hs

@@ -1,0 +1,123 @@
+{-# OPTIONS_GHC -fno-warn-incomplete-patterns #-}
+module PrintApertium where
+
+-- pretty-printer generated by the BNF converter
+
+import AbsApertium
+import Data.Char
+
+
+-- the top-level printing method
+printTree :: Print a => a -> String
+printTree = render . prt 0
+
+type Doc = [ShowS] -> [ShowS]
+
+doc :: ShowS -> Doc
+doc = (:)
+
+render :: Doc -> String
+render d = rend 0 (map ($ "") $ d []) "" where
+  rend i ss = case ss of
+    "["      :ts -> showChar '[' . rend i ts
+    "("      :ts -> showChar '(' . rend i ts
+    "{"      :ts -> showChar '{' . new (i+1) . rend (i+1) ts
+    "}" : ";":ts -> new (i-1) . space "}" . showChar ';' . new (i-1) . rend (i-1) ts
+    "}"      :ts -> new (i-1) . showChar '}' . new (i-1) . rend (i-1) ts
+    ";"      :ts -> showChar ';' . new i . rend i ts
+    t  : "," :ts -> showString t . space "," . rend i ts
+    t  : ")" :ts -> showString t . showChar ')' . rend i ts
+    t  : "]" :ts -> showString t . showChar ']' . rend i ts
+    t        :ts -> space t . rend i ts
+    _            -> id
+  new i   = showChar '\n' . replicateS (2*i) (showChar ' ') . dropWhile isSpace
+  space t = showString t . (\s -> if null s then "" else (' ':s))
+
+parenth :: Doc -> Doc
+parenth ss = doc (showChar '(') . ss . doc (showChar ')')
+
+concatS :: [ShowS] -> ShowS
+concatS = foldr (.) id
+
+concatD :: [Doc] -> Doc
+concatD = foldr (.) id
+
+replicateS :: Int -> ShowS -> ShowS
+replicateS n f = concatS (replicate n f)
+
+-- the printer class does the job
+class Print a where
+  prt :: Int -> a -> Doc
+  prtList :: [a] -> Doc
+  prtList = concatD . map (prt 0)
+
+instance Print a => Print [a] where
+  prt _ = prtList
+
+instance Print Char where
+  prt _ s = doc (showChar '\'' . mkEsc '\'' s . showChar '\'')
[…]

### 2015-02-26 bnfc/PrintCG.hs

@@ -1,0 +1,173 @@
+{-# OPTIONS_GHC -fno-warn-incomplete-patterns #-}
+module BNFC.PrintCG where
+
+-- pretty-printer generated by the BNF converter
+
+import BNFC.AbsCG
+import Data.Char
+
+
+-- the top-level printing method
+printTree :: Print a => a -> String
+printTree = render . prt 0
+
+type Doc = [ShowS] -> [ShowS]
+
+doc :: ShowS -> Doc
+doc = (:)
+
+render :: Doc -> String
+render d = rend 0 (map ($ "") $ d []) "" where
+  rend i ss = case ss of
+    "["      :ts -> showChar '[' . rend i ts
+    "("      :ts -> showChar '(' . rend i ts
+    "{"      :ts -> showChar '{' . new (i+1) . rend (i+1) ts
+    "}" : ";":ts -> new (i-1) . space "}" . showChar ';' . new (i-1) . rend (i-1) ts
+    "}"      :ts -> new (i-1) . showChar '}' . new (i-1) . rend (i-1) ts
+    ";"      :ts -> showChar ';' . new i . rend i ts
+    t  : "," :ts -> showString t . space "," . rend i ts
+    t  : ")" :ts -> showString t . showChar ')' . rend i ts
+    t  : "]" :ts -> showString t . showChar ']' . rend i ts
+    t        :ts -> space t . rend i ts
+    _            -> id
+  new i   = showChar '\n' . replicateS (2*i) (showChar ' ') . dropWhile isSpace
+  space t = showString t . (\s -> if null s then "" else (' ':s))
+
+parenth :: Doc -> Doc
+parenth ss = doc (showChar '(') . ss . doc (showChar ')')
+
+concatS :: [ShowS] -> ShowS
+concatS = foldr (.) id
+
+concatD :: [Doc] -> Doc
+concatD = foldr (.) id
+
+replicateS :: Int -> ShowS -> ShowS
+replicateS n f = concatS (replicate n f)
+
+-- the printer class does the job
+class Print a where
+  prt :: Int -> a -> Doc
+  prtList :: [a] -> Doc
+  prtList = concatD . map (prt 0)
+
+instance Print a => Print [a] where
+  prt _ = prtList
+
+instance Print Char where
+  prt _ s = doc (showChar '\'' . mkEsc '\'' s . showChar '\'')
[…]

### 2015-03-01 CG_SAT.hs
change behaviour of position 0; add support for <wordform> <rule>  syntax

@@ -4,7 +4,6 @@ module CG_SAT where
 import CG
 import CG_data
 import Data.List
-import Data.List.Split
 import Data.Maybe
 import Data.Set (Set, fromList, isSubsetOf)
 import Control.Monad
@@ -28,11 +27,6 @@ getBit (_,b) = b
 --we want to often compare the indices
 sameInd :: Token -> Token -> Bool
 sameInd tok tok' = getInd tok == getInd tok'
-
-atLeast :: Int -> [a] -> Bool
-atLeast 0 _      = True
-atLeast _ []     = False
-atLeast n (_:xs) = atLeast (n-1) xs
 
 
 --Rule     has [[Tag]].
@@ -62,12 +56,11 @@ addPosition as = concat $ go as 1
 
 --------------------------------------------------------------------------------
 
--- 1) If something is unambiguous to start with, anchor that
-anchor :: [Token] -> [Bit]
-anchor toks = [getBit tok | tok <- toks
-                          , isUniq $ filter (sameInd tok) toks]
-  where isUniq [x] = True
-        isUniq _   = False
+-- | For each position, at least one analysis must be true.
+--   Groups tokes by index and returns lists of variables 
+--   to be used in disjunction.
+anchor :: [Token] -> [[Bit]]
+anchor toks = (map.map) getBit (groupBy sameInd toks)
 
 {-
 -- 2) Take all possible bigrams
@@ -84,15 +77,8 @@ mkCombs pre post = foldr1 (:||:) combinations
                                              (_, bool2) <- post,
                                              bool1 /= bool2]
 
--- 3) Exclude hypotheses for same index so that not all can be true at the same time
--- (probably disregard, doesn't fit in with the original philosophy of CG)
-exclude :: [Token] -> [Bit]
-exclude toks = map Not combinations
-  where boolsByIndex = (map . map) getBit (groupBy sameInd toks) :: [[Bit]]
-        ambiguous = filter (atLeast 1) boolsByIndex               :: [[Bit]]
-        combinations = map (foldr1 (:&&:)) ambiguous              :: [Bit]
+--}
 
--}
 
 -- 4) Apply rules to tokens. 
 applyRule :: [Token] -> Rule -> [[Bit]]
@@ -179,10 +165,8 @@ getContext tok allToks []     = []
 getContext tok allToks ((C position (bool,ctags)):cs) = getContext tok allToks cs ++
   case ctags of
[…]

### 2015-03-01 CG_parse.hs

@@ -146,11 +146,11 @@ transRule rl = case rl of
   RemoveAlways tags   -> liftM2 CG.Remove (transTagSet tags) (return $ CG.POS CG.always)
   MatchLemma lem rule -> do cgrule <- transRule rule
                             case cgrule of
-                               CG.Select ts c -> return $ CG.Select ts (insert c lem)
-                               CG.Remove ts c -> return $ CG.Remove ts (insert c lem)
-  where insert :: CG.Test -> String -> CG.Test
-        insert (CG.POS cs) str = CG.POS $ CG.AND cs (CG.mkC "0" [[CG.Lem str]])
-        insert (CG.NEG cs) str = CG.NEG $ CG.AND cs (CG.mkC "0" [[CG.Lem str]])
+                               CG.Select ts c -> return $ CG.Select (cart ts lem) c
+                               CG.Remove ts c -> return $ CG.Remove (cart ts lem) c
+  where cart :: CG.TagSet -> String -> CG.TagSet
+        cart ts str = [[CG.Lem str,t] | t<-concat ts]
+
 
 transConds :: [Cond] -> State Env CG.Test
 transConds c = do conds <- mapM transCond c
@@ -194,13 +194,13 @@ transText x = case x of
 
 transLine :: Line -> CG.Analysis
 transLine x = case x of
-  Line (Iden wform) analyses  -> map transAnalysis analyses
-  LinePunct (Punct str)     -> [[CG.Lem str, CG.Tag "punct"]]
-  NoAnalysis (Iden wform) _ -> [[CG.Lem wform]]
+  Line (Iden wform) analyses -> map (transAnalysis wform) analyses
+  LinePunct (Punct str)     -> [[CG.WF str, CG.Lem str, CG.Tag "punct"]]
+  NoAnalysis (Iden wform) _ -> [[CG.WF wform]]
 
 
-transAnalysis :: Analysis -> [CG.Tag]
-transAnalysis (Anal (Iden id) tags) = CG.Lem id:(map transTagA tags)
+transAnalysis :: String -> Analysis -> [CG.Tag]
+transAnalysis wf (Anal (Iden id) tags) = CG.WF wf:CG.Lem id:(map transTagA tags)
 
 transTagA :: TagA -> CG.Tag
 transTagA (TagA (Iden id)) = CG.Tag id

### 2015-03-01 paper/paper.tex
stuff

@@
 a set of complex Boolean clauses is simplified, starting from unit
-clauses, which consist of just a single variable, and working up to a solution, where all variables in the set
-have a True or False value.
-
-Let us demonstrate this with a small example.
-We have a set of variables describing properties of people: \texttt{is\_under\_18},
-\texttt{has\_driving\_licence} and \texttt{is\_drunk}.
-
-% For instance, the following constraints describe rules for driving a car.
-% \begin{verbatim}
-% can_drive
-% has_driving_licence
-% is_under_18
-% is_drunk
-% \end{verbatim}
-
-% Next, we represent a possible situation in the world.
-In order to be allowed to drive, a person must have a driving licence and they must not be drunk.
-This is represented by the clause $can\_drive ∧\wedge has\_driving\_licence ∧\wedge \neg is\_drunk$.
-Furthermore, from the world knowledge, we know that in order to have a driving licence, one must be over 18.
-This is represented by the second clause.
-
-Clause 1 describes a possible situation in a world: someone is behaving in away that makes you think of a hypothesis ``either they are underage or
-drunk''. This is represented as \texttt{is\_under\_18} $\vee$
-\texttt{is\_drunk}. Just by this clause we can't solve the reason for the
-behaviour: it can be one of them or both.
-
-Clause 2 comes from the world knowledge that in order to have a
-driving licence, one must be over 18. The form of the clause is
-obtained by applying negation distributively over the more intuitive
-clause $\neg$ (\texttt{has\_driving\_licence} $\wedge$
-\texttt{is\_under\_18}): it is not possible to both have driving
-licence and be under 18.
-
-\begin{enumerate}
-\item \texttt{is\_under\_18} $\vee$ \texttt{is\_drunk}
-\item ($\neg$ \texttt{has\_driving\_licence}) $\vee$ ($\neg$ \texttt{is\_under\_18})
-\item \texttt{has\_driving\_licence}
-\end{enumerate}
-
-Clause 3 provides us some undeniable evidence that
-the person has a driving licence. By this clause, 
-we know that the variable \texttt{is\_under\_18} is not true.
-Knowing that, we can eliminate \texttt{is\_under\_18} from the first
-clause. 
-The solution is below: the three variables must have these values in order to
-satisfy the constraints.
-
-\begin{itemize}
-\item \texttt{is\_drunk} = True
-\item \texttt{is\_under\_18} = False
-\item \texttt{has\_driving\_licence} = True
-\end{itemize}
+clauses, which consist of just a single variable, and working up to a
+solution, where all variables in the set have a True or False value.
 
@@
 satisfied, such as software and hardware verification.
+Application of logic to POS tagging or shallow parsing isn't new;
+\cite{lager1998, lager2000} presents POS tagging rules as logic, and uses
+a theorem prover (DisLog, TODO cite) to implement small constraint rules.
+Inspired by the approach, my implementation tries to emulate the CG
+formalism in behaviour and features offered.
 
@@
 \begin{lstlisting}
-((1,[Lem "the",Det]),Var 1)
-((2,[Lem "bear",N,Sg]),Var 2)
-((2,[Lem "bear",V,Pl]),Var 3)
-((3,[Lem "sleep",N,Pl]),Var 4)
-((3,[Lem "sleep",V,Sg]),Var 5)
-((4,[Lem "in",Prep]),Var 6)
-((4,[Lem "in",Adv]),Var 7)
-((5,[Lem "the",Det]),Var 8)
-((6,[Lem "house",V,Inf]),Var 9)
-((6,[Lem "house",N,Sg]),Var 10)
+((1,["the",<det>]),Lit v0)
+((2,["bear",<n>,<sg>]),Lit v1)
+((2,["bear",<vblex>,<pl>]),Lit v2)
+((2,["bear",<vblex>,<fin>]),Lit v3)
+((3,["sleep",<n>,<pl>]),Lit v4)
+((3,["sleep",<vblex>,<sg>,<p3>]),Lit v5)
 \end{lstlisting}
@@
 returned, even if it were to be eliminated at the next pass. 
-However, this property isn't implemented in the SAT version.
+This is implemented as a clause that says that at least one literal
+of each position must be true.
 
@@
 \begin{lstlisting}
-rmPlIfSg = Remove [Pl] (NOT (C (Exactly (-1)) [Pl]))
-rmSgIfPl = Remove [Sg] (NOT (C (Exactly (-1)) [Sg]))
+rmPlIfSg = Remove [Pl] (C (Exactly (-1)) (False,[Pl]))
+rmSgIfPl = Remove [Sg] (C (Exactly (-1)) (False,[Sg]))
 \end{lstlisting}
@@
 
-data Condition = C Position [Tag]
-               | NOT Condition
+data Condition = C Position (Bool,[Tag])
                | AND Condition Condition
@@
 The data type for Rule is either remove or select a list of tags, with condition(s).
-For Condition, there are basic conditions (position and a list of context tags) and combined conditions, with constructors `NOT', `AND' and `OR'.
+For Condition, there are basic conditions (position and a list of context tags) and combined conditions, with constructors `AND' and `OR'.
 For instance, the following rule is read as ``remove tag for any verb, if the lemma is \emph{bear} or there is a determiner or an article anywhere to the left''.

### 2015-03-01 CG.hs

@@ -10,18 +10,20 @@ import Data.List
 -- | All kinds of morphological tags are in the same data type: e.g.  Prep, P1, Conditional.
 -- | We don't specify e.g. which tags can be part of an analysis for which word classes.
 -- | An analysis can contain an arbitrary amount of tags.
--- | Lemma is a special type of tag: Lem String
-data Tag = Tag String | Lem String deriving (Eq,Read)
+-- | Lemma and word form are also in tags.
+data Tag = Tag String | Lem String | WF String deriving (Eq,Read)
 
 
--- | Lemma should be first element in an analysis.
+-- | Wordform should be first element in an analysis.
 instance Ord Tag where
+  WF l `compare` WF l' = l `compare` l'
+  WF l `compare` _      = LT
+  _     `compare` WF l  = GT
   Lem l `compare` Lem l' = l `compare` l'
-  Lem l `compare` _      = LT
-  _     `compare` Lem l  = GT
   Tag t `compare` Tag t' = t `compare` t'
-
+  foo   `compare` bar    = show foo `compare` show bar
 instance Show Tag where
+  show (WF str) = "\"<" ++ str ++ ">\""
   show (Lem str) = "\"" ++ str ++ "\""
   show (Tag str) = "<" ++ str ++ ">"
 
@@ -37,7 +39,7 @@ instance Show Tag where
 type TagSet = [[Tag]]
 
 -- | Analysis is just list of tags: for instance the word form "alusta" would get
--- | [[Lem "alus", N, Sg, Part], [Lem "alustaa", V, Sg, Imperative]]
+-- | [[WF "alusta", Lem "alus", N, Sg, Part], [WF "alusta", Lem "alustaa", V, Sg, Imperative]]
 type Analysis = [[Tag]]
 
 -- | Sentence is just a list of analyses: e.g. "the bear sleeps"
@@ -136,7 +138,7 @@ andTest = AND lemmaBear always
 
 -- Sets of tags
 verb = (map . map) Tag [["vblex"],["vbser"],["vbmod"]]
-noun = (map . map) Tag [["n", "np"]]
+noun = (map . map) Tag [["n"], ["np"]]
 det  = [[Tag "det"]]
 adv  = [[Tag "adv"]]
 conj = (map . map) Tag [["cnjcoo"],["cnjsub"]]
@@ -148,7 +150,7 @@ cnjcoo  = [[Tag "cnjcoo"]]
 -- Rules
[…]

### 2015-03-01 SAT/SAT.hs

@@ -35,6 +35,7 @@ modelValueBit :: Solver -> Bit -> IO (Maybe Bool)
 modelValueBit s (Lit x)  = modelValue s x
 modelValueBit s (Bool b) = return (Just b)
 
+
 --------------------------------------------------------------------------------
 
 andl, orl :: Solver -> [Bit] -> IO Bit

### 2015-03-20 paper/paper.tex
asdf

@@
 a hypothesis that the correct analysis of the 3rd item in the sentence
-is \texttt{<V,Sg>}. We call this \texttt{Var 5}.
+is \texttt{<V,Sg>}. We call this \texttt{v5}.
 
 \begin{lstlisting}
-((1,["the",<det>]),Lit v0)
-((2,["bear",<n>,<sg>]),Lit v1)
-((2,["bear",<vblex>,<pl>]),Lit v2)
-((2,["bear",<vblex>,<fin>]),Lit v3)
-((3,["sleep",<n>,<pl>]),Lit v4)
-((3,["sleep",<vblex>,<sg>,<p3>]),Lit v5)
+((1,["the",<det>]),v0)
+((2,["bear",<n>,<sg>]),v1)
+((2,["bear",<vblex>,<pl>]),v2)
+((2,["bear",<vblex>,<fin>]),v3)
+((3,["sleep",<n>,<pl>]),v4)
+((3,["sleep",<vblex>,<sg>,<p3>]),v5)
 \end{lstlisting}
@@
 We have defined our variables, but not yet any truth values. We can
-start by doing two things:
+start by adding a clause that makes sure at least one variable in each
+index is true:
 \begin{enumerate}
-\item Make unambiguous tokens true by default: e.g. \texttt{Var 1}
-  shows that the reading \texttt{Det} for the first token must be true.
-\item Make hypotheses of possible n-grams. In the example below, we
-show bigrams: \texttt{(Var 1 $\wedge$ Var 2) $\vee$ (Var 1 $\wedge$ Var 3)}
-means that the first two tokens have analyses \texttt{Det,N} or
-\texttt{Det,V}.
+\item \texttt{[v0]}
+\item \texttt{[v1,v2,v3]}
+\item \texttt{[v4,v5]}
 \end{enumerate}
 
-We could easily add a third rule: for ambiguous tokens, make sure that
+We could easily add a second rule: for ambiguous tokens, make sure that
 only one reading can be true (or a less strict version, where all
@@
 returned.
-
-\begin{lstlisting}
-Var 1
-Var 8
-(Var 1 :&&: Var 2) :||: (Var 1 :&&: Var 3)
-(Var 2 :&&: Var 4) :||: ((Var 2 :&&: Var 5) :||: ((Var 3 :&&: Var 4) :||: (Var 3 :&&: Var 5)))
-(Var 4 :&&: Var 6) :||: ((Var 4 :&&: Var 7) :||: ((Var 5 :&&: Var 6) :||: (Var 5 :&&: Var 7)))
-(Var 6 :&&: Var 8) :||: (Var 7 :&&: Var 8)
-(Var 8 :&&: Var 9) :||: ((Var 8 :&&: Var 10))
-\end{lstlisting}
-% Not (Var 2 :&&: Var 3)
-% Not (Var 4 :&&: Var 5)
-% Not (Var 6 :&&: Var 7)
-% Not (Var 9 :&&: Var 10)
 
@@
 
-\subsection{Conflicts}
-If you try to use conflicting rules, you get user error: mzero.
+\subsection{Ordering}
+
+We had three strategies:
+maximise: try to make as many rules true as possible;
+maximise from top: apply rules from the beginning, discard rule if it
+conflicts with the ones added;
+discard from bottom: try to add all rules, if conflict, find which
+conflict, and try again with only the rules that are in the list
+before the first conflicting rule.
+
+\subsection{Weird things}
+
+mkVars: difference between options 1) and 2)
+
+        mkVars :: [(Token,[Token])] -> (Bit -> Bit) -> [[Bit]]
+1)        mkVars tctx f = [ f conseq:(map nt causes) | (t, ts) <- tctx
+2)        mkVars tctx f = [ [nt cause,f conseq] | (t, ts) <- tctx
+                                               , let conseq = getBit t
+                                               , let causes = map getBit ts
+                                               , cause <- causes ]
+
+% Remove [[vblex]] (POS (AND (C (Exactly 0) (True,[[vblex,p1],[vblex,p2],[vblex,p3]])) (C (Barrier 1 [[cnjcoo],[cnjsub],[cnjadv],[rel]]) (True,[[vblex,p1],[vblex,p2],[vblex,p3]]))))
+% [Lit ~v6,Lit ~v2]
+
+We have rules:
+REMOVE:r_como_1 Vblex (0 ("como") OR ("Como")) (*-1 Vblex BARRIER Cnj_Rel) ;
+REMOVE:r_verb_1 Vblex (0 Vblex + Pers) (*1C Vblex + Pers BARRIER
+Cnj_Rel) ;
+
+And tokens:
+((1,[">>>",>>>]),Lit v0)
+((2,["<Jared>","*Jared"]),Lit v1)
+((3,["<explica>","explicar",vblexfoo,pri,p3,sg]),Lit v2)
+((3,["<explica>","explicar",vblex,imp,p2,sg]),Lit v3)
+((4,["<como>","como",pr]),Lit v4)
+((4,["<como>","como",rel,adv]),Lit v5)
+((4,["<como>","comer",vblex,pri,p1,sg]),Lit v6)
+((5,["<pequeñas>","pequeño",adj,f,pl]),Lit v7)
+((6,["<diferencias>","diferencia",n,f,pl]),Lit v8)
+((6,["<diferencias>","diferenciar",vblex,pri,p2,sg]),Lit v9)
+
+In CG3, r_como_1 is applied first, and v6 is eliminated. r_verb_1
+never has the Vblex+Pers condition available and is never triggered.
+In CGSAT, neither is applied first; r_como_1 is is implication ``v2 =>
+not v6'', ``v3 => not v6'' and r_verb_1 is ``v6 => not v2'' ``v6 => not v3''.
+
+You can flip rules!
+
+\begin{verbatim}
+SELECT (det "<az>") IF (1 AdjOrN);
+SELECT AdjOrN IF (-1 (det "<az>"));
+
+REMOVE ("de" adv) IF (NOT 1C (adj));
+REMOVE AllButAdj IF (-1 ("de" adv));
+\end{verbatim}
 

### 2015-03-20 CG_SAT.hs
stuff stuff more stuff

@@ -116,7 +116,7 @@ applyRules rule (conds:cs) allToks = applyRules rule cs allToks ++
      -- needed because the word at -1 could have many tags, and they could conflict.
         mkVars :: [(Token,[Token])] -> (Bit -> Bit) -> [[Bit]]
 --      mkVars tctx f = [ f conseq:(map nt causes) | (t, ts) <- tctx
-        mkVars tctx f = [ [f conseq,nt cause] | (t, ts) <- tctx
+        mkVars tctx f = [ [nt cause,f conseq] | (t, ts) <- tctx
                                                , let conseq = getBit t
                                                , let causes = map getBit ts
                                                , cause <- causes ]
@@ -203,7 +203,7 @@ disambiguate verbose rules sentence = do
     let usedrules = [ show rule ++ "\n" ++ show btags
                       | (brl:btags) <- applied 
                       , (rule,bit) <- rlsBits
-                      , bit == nt brl ] -- brl is negated in the implication
+                      , bit == nt brl || bit == brl ] -- brl is negated in the implication
     
     putStrLn "\ntokens:"
     mapM_ print toks

### 2015-03-20 CG_base.hs

@@ -183,11 +183,11 @@ showSentence = concatMap showAnalysis
 --	"are" n sg
 -- so need some trickery
 showAnalysis :: Analysis -> String
+showAnalysis []     = []
 showAnalysis (a:as) = unlines $ showTags a : map showTags as'
   where as' = (map.filter) notWF as
         notWF (WF _) = False
-        notWF _      = True   
-        
+        notWF _      = True        
   
 showTags :: [Tag] -> String
 showTags ts@(wf:as) = 

### 2015-03-20 CG_parse.hs

@@ -154,7 +154,10 @@ transTagSet ts = case ts of
 
   ----TODO all set operations!
 
-  Diff All ts    -> error "TODO this should have effect on a higher level"
+  Diff All ts    -> do env <- get
+                       let allTags = concatMap snd env 
+                       rmTags <- transTagSet ts
+                       return $ allTags \\ rmTags
   Diff ts All    -> error "something except everything? are you a philosopher?"
   Diff ts1 ts2   -> do tags1 <- transTagSet ts1
                        tags2 <- transTagSet ts2
@@ -200,15 +203,14 @@ transCond c = case c of
   CLinked (c:cs)      -> do first@(CGB.C pos tags) <- transCond c
                             let base = getPos pos
                             conds <- mapM transCond cs
-                            let op = if isLink0 conds then CGB.OR else CGB.AND
-                            return $ foldr op first (fixPos base conds [])
+                            return $ foldr CGB.AND first (fixPos base conds [])
 
   where isLink0 []                 = True
         isLink0 (CGB.C pos _ts:cs) = getPos pos == 0 && isLink0 cs
 
         fixPos base []                  res = res
         fixPos base (CGB.C pos tags:cs) res = 
-          let newBase = getPos pos
+          let newBase = base + getPos pos
               newPos = changePos pos newBase
           in fixPos newBase cs ((CGB.C newPos tags):res)
 
@@ -259,6 +261,8 @@ transLine :: Line -> CGB.Analysis
 transLine x = case x of
   Line (Iden wform) anas    -> map (transAnalysis wform) anas
   LinePunct (Punct p) anas  -> map (transAnalysis p) anas
+  OnlyPunct (Punct ",")     -> [[CGB.WF ",", CGB.Lem ",", CGB.Tag "cm"]]
+  OnlyPunct (Punct ".")     -> [[CGB.WF ".", CGB.Lem ".", CGB.Tag "sent"]]
   OnlyPunct (Punct str)     -> [[CGB.WF str, CGB.Lem str, CGB.Tag "punct"]]
   NoAnalysis (Iden wform) _ -> [[CGB.WF wform]]
 

### 2015-03-23 paper/nodalida2015.tex
add paper draft for nodalida

%
% File nodalida2015.tex
%
% Contact beata.megyesi@lingfil.uu.se
%
% Based on the instruction file for EACL 2014
% which in turn was based on the instruction files for previous 
% ACL and EACL conferences.

\documentclass[11pt]{article}
\usepackage{nodalida2015}
\usepackage{times}
\usepackage{mathptmx}
%\usepackage{txfonts}
\usepackage{url}
\usepackage{latexsym}
\usepackage{listings}
\special{papersize=210mm,297mm} % to avoid having to use "-t a4" with dvips 
%\setlength\titlebox{6.5cm}  % You can expand the title box if you really have to

\title{Towards ....}

\author{Inari Listenmaa \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  Affiliation / Address line 3 \\
  {\tt inari@chalmers.se} \\\And
  Koen Claessen \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  Affiliation / Address line 3 \\
  {\tt koen@chalmers.se} \\}

\date{}

\begin{document}
\maketitle
\begin{abstract}
We implement Constraint Grammar using a SAT solver.

We experiment with different strategies to maximise the number of instances to apply rules;
this will also lead to different orderings of the rules and ways of solving conflicts.

\end{abstract}

\section{Introduction}
Constraint Grammar (CG) was first introduced by \cite{KarlssonTODO}. 
It is a tool for disambiguating output by morphological analyser.

\section{Related work}
\label{sect:related}

\cite{lager1998, lager2000}  CG and FSIG (Finite State Intersectional Grammar) inspired POS tagging using theorem prover.
% What we do better: implement something that takes existing CG rules and applies them to text, trying to mimic/outperform current CG implementations.

Transformation-Based Learning of Rules for CG tagging (the paper with rules where ordering didn't matter so much)

\section{Implementation using SAT}
\label{sect:pdf}

SAT solving is based on a technique called unit propagation:
a set of complex Boolean clauses is simplified, starting from unit
clauses, which consist of just a single variable, and working up to a
solution, where all variables in the set have a True or False value.

SAT solving is used for many applications where constraints must be
satisfied, such as software and hardware verification.
Application of logic to POS tagging or shallow parsing isn't new;
Lager~\shortcite{lager1998, lager2000} presents POS tagging rules as logic, and uses
a theorem prover (DisLog, TODO cite) to implement small constraint rules.
Inspired by the approach, our implementation tries to emulate the CG
formalism in behaviour and features offered.

We make each analysis a variable. For instance, the fifth line denotes
a hypothesis that the correct analysis of the 3rd item in the sentence
is \texttt{<V,Sg>}. We call this \texttt{v5}.

\begin{lstlisting}
((1,["<the>","the",det]),v0)
((2,["<bear>","bear",n,sg]),v1)
((2,["<bear>","bear",vblex,pl]),v2)
((2,["<bear>","bear",vblex,fin]),v3)
((3,["<sleeps>","sleep",n,pl]),v4)
((3,["<sleeps>","sleep",vblex,sg,p3]),v5)
\end{lstlisting}

\subsection{Rule ordering}
\label{ssec:ordering}

Ordering and conflicts : the most important differences here

\cite{koskenniemi92}
\begin{quote}Rules in the CG formalism are typically [--] executed as successive groups.
In finite-state syntax, rules are logically unordered.\end{quote}

% VISLCG ad hoc ordering:
% 1. SELECT before REMOVE.
% 2. SELECT rules targetting more preferred tags before 
%   rules targetting less preferred targets.
% 3. REMOVE rules targetting less preferred tags before
 %   rules targetting more preferred targets.
% 4. By order of appearance in rule file.

In CG3, the rules are applied in the order of appearance in the rule file.

% \section*{Acknowledgments}

% Do not number the acknowledgment section. Do not include this section
% when submitting your paper for review.

\bibliographystyle{acl}
\bibliography{cg}

\end{document}

### 2015-03-23 CG_SAT.hs
no idea what I've been changing here sorry

@@ -28,8 +28,8 @@ isBoundary :: Token -> Bool
 isBoundary tok = not $ null ([BOS,EOS] `intersect` getTags tok)
 
 -- we don't need to apply rules to tokens that are already unambiguous
-isUnique :: Token -> [Token] -> Bool
-isUnique tok toks = length (filter (sameInd tok) toks) == 1
+isAmbig :: Token -> [Token] -> Bool
+isAmbig tok toks = length (filter (sameInd tok) toks) > 1
 
 --we want to often compare the indices
 sameInd :: Token -> Token -> Bool
@@ -122,13 +122,16 @@ applyRules rule (conds:cs) allToks = applyRules rule cs allToks ++
         mkVars tctx f = [ [nt cause,f conseq] | (t, ts) <- tctx
                                                , let conseq = getBit t
                                                , let causes = map getBit ts
-                                               , cause <- causes ]
+                                               , cause <- causes 
+--                                               , cause /= conseq
+--                                               , cause /= Bool True --with this ~200 sentences get worse
+                                               ]
 
 
         -- chosen: analyses that have the wanted readings and context
         chosen :: TagSet -> [(Token,[Token])]
         chosen tags = [(tok, concat ctx) | tok <- allToks
-                                         -- , not $ isUnique tok allToks -- not sure if this has any effect
+                                          , isAmbig tok allToks --only apply rules to ambiguous tokens
                                           , tagsMatchRule tags tok
                                           , let ctx = getContext tok allToks conds
                                           , allCondsHold ctx]
@@ -159,16 +162,19 @@ getContext :: Token           -- ^ a single analysis
 getContext tok allToks []     = []
 getContext tok allToks ((C position (bool,ctags)):cs) = getContext tok allToks cs ++
   case ctags of
-    []     -> [[tok]] -- empty tags in condition = remove/select always
-    [[]]   -> [[tok]] -- since we need a context, give the word itself
-    -- []     -> [[]]
-    -- [[]]   -> [[]]
+    []     -> [[dummyTok]] --empty conds = holds always
+    [[]]   -> [[dummyTok]] -- if I replace dummyTok with tok, ~150 sentences get worse in Pride
     (t:ts) -> case position of
+                Exactly 0 -> if neg' $ tagsMatchRule ctags tok 
[…]

### 2015-03-23 test/Test.hs

@@ -27,7 +27,8 @@ main = do
                                        | (r,g,orig) <- zip3 result goldst data' ]
                                   --     , null $ intersect r g ]
                           when v $ (mapM_ . mapM_) prDiff diff
-                          print (length result, length goldst)
+                          putStr "Sentences in the original: "
+                          print (length goldst)
                           putStr "Sentences that differ from vislcg3: "
                           print $ length $ filter (not.null) diff
         prDiff (a1,a2,s) = do putStrLn "Original sentence:"

### 2015-03-24 paper/nodalida2015.tex
more text

@@
 \usepackage{latexsym}
-\usepackage{listings}
 \special{papersize=210mm,297mm} % to avoid having to use "-t a4" with dvips 
@@
 
-\author{Inari Listenmaa \\
-  Affiliation / Address line 1 \\
-  Affiliation / Address line 2 \\
-  Affiliation / Address line 3 \\
-  {\tt inari@chalmers.se} \\\And
-  Koen Claessen \\
-  Affiliation / Address line 1 \\
-  Affiliation / Address line 2 \\
-  Affiliation / Address line 3 \\
-  {\tt koen@chalmers.se} \\}
+\author{Inari Listenmaa \and Koen Claessen \\
+  Chalmers University of Technology, Gothenburg, Sweden \\
+  {\tt \{inari,koen\}@chalmers.se} }
 
@@
 
-\cite{lager1998, lager2000}  CG and FSIG (Finite State Intersectional Grammar) inspired POS tagging using theorem prover.
+Application of logic to POS tagging or shallow parsing isn't new;
+\cite{lager1998, lager2000} present constraint rules as 
+a disjunctive logic program.
+
+Our implementation is inspired by this approach,
+but unlike Lager's our implementation tries to emulate the CG
+formalism in behaviour and features offered.
 % What we do better: implement something that takes existing CG rules and applies them to text, trying to mimic/outperform current CG implementations.
@@
 
-\section{Implementation using SAT}
+\section{Implementation using SAT solver}
 \label{sect:pdf}
@@
 satisfied, such as software and hardware verification.
-Application of logic to POS tagging or shallow parsing isn't new;
-Lager~\shortcite{lager1998, lager2000} presents POS tagging rules as logic, and uses
-a theorem prover (DisLog, TODO cite) to implement small constraint rules.
-Inspired by the approach, our implementation tries to emulate the CG
-formalism in behaviour and features offered.
 
-We make each analysis a variable. For instance, the fifth line denotes
-a hypothesis that the correct analysis of the 3rd item in the sentence
-is \texttt{<V,Sg>}. We call this \texttt{v5}.
+Let us demonstrate the approach with the example sentence 
+\emph{the bear sleeps}. After applying morphological tagger, we get the following analyses:
 
-\begin{lstlisting}
-((1,["<the>","the",det]),v0)
-((2,["<bear>","bear",n,sg]),v1)
-((2,["<bear>","bear",vblex,pl]),v2)
-((2,["<bear>","bear",vblex,fin]),v3)
-((3,["<sleeps>","sleep",n,pl]),v4)
-((3,["<sleeps>","sleep",vblex,sg,p3]),v5)
-\end{lstlisting}
+\begin{verbatim}
+"<the>"
+        "the" det
+"<bear>"
+        "bear" n sg
+        "bear" v pl
+"<sleeps>"
+        "sleep" n pl
+        "sleep" v sg p3
+\end{verbatim}
+
+We make each analysis a variable \texttt{v1}--\texttt{v5}. 
+For instance, \texttt{v3} is a hypothesis that the correct analysis 
+for the word form ``bear'' is \texttt{"bear" n pl}.
+Before applying rules, we must make sure that in each position
+at least one of the analyses is true. We add our first disjunctions:
+
+\begin{itemize}
+\item [] $v1$ \\ $v2 \vee v3$ \\ $v4 \vee v5$
+\end{itemize}
+
+The first
 
@@
 
+A typical CG rule sequence starts with stronger rules
+
 % \section*{Acknowledgments}

### 2015-03-24 paper/nodalida2015.tex
more text and add refs

@@
 \usepackage{latexsym}
+\usepackage{cite}
 \special{papersize=210mm,297mm} % to avoid having to use "-t a4" with dvips 
@@
 \section{Introduction}
-Constraint Grammar (CG) was first introduced by \cite{KarlssonTODO}. 
+Constraint Grammar (CG) was first introduced by \cite{karlsson1995constraint}. 
 It is a tool for disambiguating output by morphological analyser.
@@
 Application of logic to POS tagging or shallow parsing isn't new;
-\cite{lager1998, lager2000} present constraint rules as 
+\cite{lager98, lager_nivre01} present constraint rules as 
 a disjunctive logic program.
@@
 
-Transformation-Based Learning of Rules for CG tagging (the paper with rules where ordering didn't matter so much)
+\cite{lager01transformation} showed rules where ordering didn't matter so much in the end.
 
@@
 
-The first
-
 \subsection{Rule ordering}

### 2015-03-24 paper/nodalida2015.tex
more text

@@
 
-\date{}
+\date{\today}
 
@@
 
-We make each analysis a variable \texttt{v1}--\texttt{v5}. 
-For instance, \texttt{v3} is a hypothesis that the correct analysis 
+We make each analysis a variable; let us call them \texttt{det1}, \texttt{n2}, \texttt{v2}, \texttt{n3} and \texttt{v3} by the proposed POS tag and the position in the sentence.
+For instance, \texttt{n2} is a hypothesis that the correct analysis 
 for the word form ``bear'' is \texttt{"bear" n pl}.
@@
 \begin{itemize}
-\item [] $v1$ \\ $v2 \vee v3$ \\ $v4 \vee v5$
+\item [] $det1$ \\ $n2 \vee v2$ \\ $n3 \vee v3$
 \end{itemize}
+
+Let us add two rules
+\begin{itemize}
+\item [] \texttt{REMOVE v IF -1 det} \\ \texttt{REMOVE n IF -1 n}
+\end{itemize}
+
+Instead of applying these rules
+
+The condition of being unambiguously tagged becomes less necessary as well.
+We add many clauses, knowing that some of them end up being false.
 
@@
 
+\subsection{Other experiments}
+
+Different strategies for maximising: ...
+
+Modifications to an existing CG to better disambiguate with our implementation
+
+Flipping rules: 
+\begin{itemize}
+\item [] \texttt{REMOVE v IF -1 det} \\ $\rightarrow$ \texttt{REMOVE det IF 1 verb}   
+\item [] \texttt{REMOVE adv IF NOT 1C adj} \\ $\rightarrow$ \texttt{REMOVE (*) - Adj  IF -1 adv}  \\
+\end{itemize}
+
+With this experiment, we move even further away from a direct ``if this then that'' model of CG.
+% Not really CG anymore , but
+
+\section{Results and evaluation}
+
 % \section*{Acknowledgments}
@@
 % when submitting your paper for review.
+% Francis Tyers
 

### 2015-03-25 paper/nodalida2015.tex
fix formatting of bibtex

@@
 \usepackage{cite}
+\usepackage{authordate1-4}
 \special{papersize=210mm,297mm} % to avoid having to use "-t a4" with dvips 
@@
 
+Our implementation is inspired by approaches like \cite{lager98,lager_nivre01},
+but we try to emulate the CG formalism more closely, and use the same rule files.
+
 \section{Related work}
@@
 Application of logic to POS tagging or shallow parsing isn't new;
-\cite{lager98, lager_nivre01} present constraint rules as 
+\cite{lager98,lager_nivre01} present constraint rules as 
 a disjunctive logic program.
 
-Our implementation is inspired by this approach,
-but unlike Lager's our implementation tries to emulate the CG
-formalism in behaviour and features offered.
+\cite{lindberg_eineborg98ilp,asfrent14}
+
 % What we do better: implement something that takes existing CG rules and applies them to text, trying to mimic/outperform current CG implementations.

### 2015-03-25 paper/nodalida2015.tex
long explanation about the imp/pri rule in section 3.2

@@
 \begin{abstract}
-We implement Constraint Grammar using a SAT solver.
+We implement Constraint Grammar as a Boolean satisfiability problem.
 
@@
 
-\section{Implementation using SAT solver}
+\section{CG as a SAT problem}
 \label{sect:pdf}
@@
 
-We make each analysis a variable; let us call them \texttt{det1}, \texttt{n2}, \texttt{v2}, \texttt{n3} and \texttt{v3} by the proposed POS tag and the position in the sentence.
+We make each analysis a variable; let us call them \texttt{d1}, \texttt{n2}, \texttt{v2}, \texttt{n3} and \texttt{v3} by the proposed POS tag and the position in the sentence.
 For instance, \texttt{n2} is a hypothesis that the correct analysis 
-for the word form ``bear'' is \texttt{"bear" n pl}.
+for the word form \emph{bear}' is \texttt{"bear" n pl}.
 Before applying rules, we must make sure that in each position
@@
 \begin{itemize}
-\item [] $det1$ \\ $n2 \vee v2$ \\ $n3 \vee v3$
+\item [] $d1$ \\ $n2 \vee v2$ \\ $n3 \vee v3$
 \end{itemize}
 
-Let us add two rules
+Let us add two rules:
 \begin{itemize}
-\item [] \texttt{REMOVE v IF -1 det} \\ \texttt{REMOVE n IF -1 n}
+\item [] \texttt{REMOVE v IF -1 det} \\ 
+            \texttt{REMOVE n IF -1 n} 
+%             $d1 \Rightarrow \neg v2$ 
+%\item [] \texttt{REMOVE n IF -1 n} \\
+%              $n2   \Rightarrow \neg n3$
+
 \end{itemize}
 
-Instead of applying these rules
+The rules are applied as implications: \emph{if ``the'' is a determiner, then ``bear'' cannot be a verb}. 
+The antecedent is
+
+%\begin{itemize}
+%\item [] $d1 \Rightarrow \neg v2$ \\
+%           $n2   \Rightarrow \neg n3$
+%\end{itemize}
 
@@
 
-\cite{koskenniemi92}
-\begin{quote}Rules in the CG formalism are typically [--] executed as successive groups.
-In finite-state syntax, rules are logically unordered.\end{quote}
+%\cite{koskenniemi92}
+%\begin{quote}Rules in the CG formalism are typically [--] executed as successive groups.
+%In finite-state syntax, rules are logically unordered.\end{quote}
 
@@
 
-A typical CG rule sequence starts with stronger rules
+A typical CG rule sequence starts with weaker rules, and leaves stronger rules towards the end, 
+to be applied if none of the weaker rules has fired.
 
@@
 
-Modifications to an existing CG to better disambiguate with our implementation
+\paragraph{Modifications to an existing CG to better disambiguate with our implementation.}
+With the new model, the semantics of SELECT change.
+In the traditional CG, SELECT means ``remove everything else''. 
+In SAT-based implementation, SELECT means ``this variable will have value true''
 
-Flipping rules: 
+Choosing SELECT is less permanent as it is in the traditional.
+The following example illustrates the interaction of rules to produce the desired outcome.
+We have two ambiguous words with the following analysis:
+
+\begin{verbatim}
+"<la>"
+        "el" det def f sg
+        "lo" prn pro p3 f sg
+"<casa>"
+        "casa" n f sg
+        "casar" vblex pri p3 sg
+        "casar" vblex imp p2 sg
+\end{verbatim}
+
+We apply a rule which removes the imperative in favour of the present indicative\footnote{\texttt{REMOVE imp IF (0 pri LINK 0 imp) ;}}.
+For both VISLCG3 and our system, the rule removes \texttt{"casar" vblex imp p2 sg}, 
+other analyses stay in place. If we replace that rule with a version that selects the present indicative 
+version\footnote{\texttt{SELECT pri IF (0 pri LINK 0 imp) ;}}, this leads to the same result as the REMOVE rule: only the imperative reading is removed, but the noun reading isn't touched.
+
+What happens here? We get three clauses:
+
+\begin{itemize}
+\item [] $casaImp \Rightarrow casaPri$ \\
+            $casaPri \Rightarrow \neg casaImp$ \\
+            $casaPri \vee casaImp \Rightarrow \neg casaN$
+\end{itemize}
+
+Changed into disjunctions, the clauses will look like
+\begin{itemize}
+\item [] $\neg casaImp \vee casaPri$ \\
+            $\neg casaPri \vee \neg casaImp$ \\
+            $\neg casaPri \vee \neg casaImp \vee \neg casaN$
+\end{itemize}
+
+The first clause simplifies to $casaPri$, and using that as a new unit clause, the second clause simplifies to $\neg casaImp$. 
+For the third clause, we don't have any rule that would set the value of $casaN$, so the general maximisation step ensures that $casaN$ will be true.
+
+The fully disambiguated sequence is ``la\texttt{:det} casa\texttt{:n}'', so we need some more rules.
+We add the following rules to disambiguate the word \emph{la}:
+\begin{itemize}
+\item [] \texttt{REMOVE prn IF (1 n) ;} \\
+             \texttt{REMOVE det IF (1 vblex) ;}
+\end{itemize}
+
+In the standard CG, the rule that removes the pronoun would apply first, because it comes first in the rule file.
+The second rule will not fire, because it would remove the last reading for the word \emph{la}, and that is not allowed.
+To be more cautious, we could require the word at position +1 be disambiguated fully (\texttt{1C} instead of \texttt{1}), but in that case, neither of them would fire.
+The ambiguity between the remaining readings of \emph{casa} is left untouched by these rules.
+In contrast, with the clauses created by the SAT-based implementation, we disambiguate \emph{casa} for free, and ordering of the rules doesn't have an effect.
+
+\paragraph{Flipping rules.}
+Since implications can be presented as disjunctions, we can 
+The representation of the implications as disjunctions 
 \begin{itemize}
@@
 With this experiment, we move even further away from a direct ``if this then that'' model of CG.
-% Not really CG anymore , but
 

### 2015-03-25 paper/nodalida2015.tex
table for results

@@
 
-\paragraph{Modifications to an existing CG to better disambiguate with our implementation.}
-With the new model, the semantics of SELECT change.
-In the traditional CG, SELECT means ``remove everything else''. 
-In SAT-based implementation, SELECT means ``this variable will have value true''
-
-Choosing SELECT is less permanent as it is in the traditional.
-The following example illustrates the interaction of rules to produce the desired outcome.
-We have two ambiguous words with the following analysis:
+\paragraph{Less rules needed to disambiguate.}
+The following example illustrates the interaction of rules to produce the desired outcome. We have an ambiguous passage in Spanish, with the following analyses:
 
@@
 
-We apply a rule which removes the imperative in favour of the present indicative\footnote{\texttt{REMOVE imp IF (0 pri LINK 0 imp) ;}}.
-For both VISLCG3 and our system, the rule removes \texttt{"casar" vblex imp p2 sg}, 
-other analyses stay in place. If we replace that rule with a version that selects the present indicative 
-version\footnote{\texttt{SELECT pri IF (0 pri LINK 0 imp) ;}}, this leads to the same result as the REMOVE rule: only the imperative reading is removed, but the noun reading isn't touched.
-
-What happens here? We get three clauses:
-
-\begin{itemize}
-\item [] $casaImp \Rightarrow casaPri$ \\
-            $casaPri \Rightarrow \neg casaImp$ \\
-            $casaPri \vee casaImp \Rightarrow \neg casaN$
-\end{itemize}
-
-Changed into disjunctions, the clauses will look like
-\begin{itemize}
-\item [] $\neg casaImp \vee casaPri$ \\
-            $\neg casaPri \vee \neg casaImp$ \\
-            $\neg casaPri \vee \neg casaImp \vee \neg casaN$
-\end{itemize}
-
-The first clause simplifies to $casaPri$, and using that as a new unit clause, the second clause simplifies to $\neg casaImp$. 
-For the third clause, we don't have any rule that would set the value of $casaN$, so the general maximisation step ensures that $casaN$ will be true.
-
-The fully disambiguated sequence is ``la\texttt{:det} casa\texttt{:n}'', so we need some more rules.
 We add the following rules to disambiguate the word \emph{la}:
@@
 
-In the standard CG, the rule that removes the pronoun would apply first, because it comes first in the rule file.
+Standard CG will apply first the rule that removes the pronoun reading, because it comes first in the rule file.
 The second rule will not fire, because it would remove the last reading for the word \emph{la}, and that is not allowed.
 To be more cautious, we could require the word at position +1 be disambiguated fully (\texttt{1C} instead of \texttt{1}), but in that case, neither of them would fire.
-The ambiguity between the remaining readings of \emph{casa} is left untouched by these rules.
-In contrast, with the clauses created by the SAT-based implementation, we disambiguate \emph{casa} for free, and ordering of the rules doesn't have an effect.
+The ambiguity between the readings of \emph{casa} is left untouched by these rules.
 
-\paragraph{Flipping rules.}
+With the SAT clauses, we get two solutions: ``la\texttt{:det} casa\texttt{:n}''  and ``la\texttt{:prn} casa\texttt{:v}''.
+The interaction of the rules regarding \emph{la}  disambiguates \emph{casa} for free, and ordering of the rules doesn't matter.
+Since we don't have other rules nor further context, we still get two solutions, but there are no illegal combinations (\texttt{det v} or \texttt{prn n}) in one solution.
+
+\paragraph{Rules can be flipped.}
 Since implications can be presented as disjunctions, we can 
@@
 
+We took small constraint grammars (<500 rules) for Spanish and English, and disambiguated a number of texts from Project Gutenberg.
+We compared the result of using VISLCG3 for the same task; table~\ref{table:results} shows the differences in results in both sentence and word level.
+
+\begin{table}
+  \centering
+  \begin{tabular}{|r|r|r|r|r|}
+     \hline
+    \textbf{Text} & \textbf{Diff words} & \textbf{Diff sentences} \\
+    \hline
+	Pride and Prejudice & todo & 5457/7667 \\  
+    \hline
+       Don Quijote & todo & 3k/6k  \\ 
+     \hline
+  \end{tabular}
+  \caption{Results for comparing SATCG and VISLCG3.}
+  \label{table:results}
+\end{table}
+
+These tendencies show also in a larger context with more complex rules.
+
 % \section*{Acknowledgments}

### 2015-03-25 paper/nodalida2015.tex
more text

@@
 
-\title{Towards ....}
+\title{Towards something :-D}
 
@@
 Constraint Grammar (CG) was first introduced by \cite{karlsson1995constraint}. 
-It is a tool for disambiguating output by morphological analyser.
+It disambiguates output by morphological analyser, 
+and has been used for many tasks in computational linguistics, such as POS tagging,
+surface syntax and machine translation.
+The current state-of-the-art parser VISL CG-3 is implemented (...);
+in addition there are a number of finite-state based implementations \cite{nemeskey14}.
 
+We represent CG as a Boolean satisfiability problem using the 
+open-source SAT solver MiniSat \cite{een04sat}.
 Our implementation is inspired by approaches like \cite{lager98,lager_nivre01},
-but we try to emulate the CG formalism more closely, and use the same rule files.
+but we try to emulate the CG formalism more closely than the previous logic-based works.
+
+% NP-complete -- overkill for CG? But gives other benefits.
 
@@
 a disjunctive logic program.
+\cite{lindberg_eineborg98ilp,asfrent14} use Inductive Logic Programming to learn CG rules from a tagged corpus.
+\cite{lager01transformation} discusses ordering of the CG rules.
 
-\cite{lindberg_eineborg98ilp,asfrent14}
-
-% What we do better: implement something that takes existing CG rules and applies them to text, trying to mimic/outperform current CG implementations.
-
-\cite{lager01transformation} showed rules where ordering didn't matter so much in the end.
+Works from the late 90s and early 00s mention that the method is very slow;
+however, advances in SAT solving technique
 
@@
 solution, where all variables in the set have a True or False value.
-
-SAT solving is used for many applications where constraints must be
-satisfied, such as software and hardware verification.
-
-Let us demonstrate the approach with the example sentence 
-\emph{the bear sleeps}. After applying morphological tagger, we get the following analyses:
-
-\begin{verbatim}
-"<the>"
-        "the" det
-"<bear>"
-        "bear" n sg
-        "bear" v pl
-"<sleeps>"
-        "sleep" n pl
-        "sleep" v sg p3
-\end{verbatim}
-
-We make each analysis a variable; let us call them \texttt{d1}, \texttt{n2}, \texttt{v2}, \texttt{n3} and \texttt{v3} by the proposed POS tag and the position in the sentence.
-For instance, \texttt{n2} is a hypothesis that the correct analysis 
-for the word form \emph{bear}' is \texttt{"bear" n pl}.
-Before applying rules, we must make sure that in each position
-at least one of the analyses is true. We add our first disjunctions:
+We model an ambiguous sentence as a set of clauses, where each analysis is a variable, 
+and constraint rules form clauses operating on those variables.
 
 \begin{itemize}
-\item [] $d1$ \\ $n2 \vee v2$ \\ $n3 \vee v3$
+\item The condition of being unambiguously tagged becomes less necessary.
+We add many clauses, knowing that some of them end up being false.
+\item Rules are unordered. 
+\item Rules interact with each other and disambiguate more than same rules in the standard implementation.
 \end{itemize}
 
-Let us add two rules:
-\begin{itemize}
-\item [] \texttt{REMOVE v IF -1 det} \\ 
-            \texttt{REMOVE n IF -1 n} 
-%             $d1 \Rightarrow \neg v2$ 
-%\item [] \texttt{REMOVE n IF -1 n} \\
-%              $n2   \Rightarrow \neg n3$
+% Let us demonstrate the approach with the example sentence 
+% \emph{the bear sleeps}. After applying morphological tagger, we get the following analyses:
 
-\end{itemize}
+% \begin{verbatim}
+% "<the>"
+%         "the" det
+% "<bear>"
+%         "bear" n sg
+%         "bear" v pl
+% "<sleeps>"
+%         "sleep" n pl
+%         "sleep" v sg p3
+% \end{verbatim}
 
-The rules are applied as implications: \emph{if ``the'' is a determiner, then ``bear'' cannot be a verb}. 
-The antecedent is
+%We make each analysis a variable; let us call them \texttt{d1}, \texttt{n2}, \texttt{v2}, \texttt{n3} and \texttt{v3} by the proposed POS tag and the position in the sentence.
+%For instance, \texttt{n2} is a hypothesis that the correct analysis 
+%for the word form \emph{bear}' is \texttt{"bear" n pl}.
+%Before applying rules, we must make sure that in each position
+%at least one of the analyses is true. We add our first disjunctions:
+
+%\begin{itemize}
+%\item [] $d1$ \\ $n2 \vee v2$ \\ $n3 \vee v3$
+%\end{itemize}
+
+%Let us add two rules:
+%\begin{itemize}
+%\item [] \texttt{REMOVE v IF -1 det} \\ 
+%            \texttt{REMOVE n IF -1 n} 
+%\end{itemize}
+
+%The rules are applied as implications: \emph{if ``the'' is a determiner, then ``bear'' cannot be a verb}.
 
@@
 
-The condition of being unambiguously tagged becomes less necessary as well.
-We add many clauses, knowing that some of them end up being false.
-
 \subsection{Rule ordering}
@@
 
-Ordering and conflicts : the most important differences here
+A typical CG rule sequence starts with weaker rules, and leaves stronger rules towards the end, 
+to be applied if none of the weaker rules has fired.
 
-%\cite{koskenniemi92}
-%\begin{quote}Rules in the CG formalism are typically [--] executed as successive groups.
-%In finite-state syntax, rules are logically unordered.\end{quote}
+Unordered rules are not new: finite-state implementations of CG \cite{koskenniemi92} also have logically unordered rules.
 
@@
 
-A typical CG rule sequence starts with weaker rules, and leaves stronger rules towards the end, 
-to be applied if none of the weaker rules has fired.
-
-\subsection{Other experiments}
-
 Different strategies for maximising: ...
 
-\paragraph{Less rules needed to disambiguate.}
+\subsection{Rule interaction}
+
 The following example illustrates the interaction of rules to produce the desired outcome. We have an ambiguous passage in Spanish, with the following analyses:
@@
 
-We took small constraint grammars (<500 rules) for Spanish and English, and disambiguated a number of texts from Project Gutenberg.
+We took small constraint grammars (\textless{}500 rules) for Spanish and English, and disambiguated a number of texts from Project Gutenberg.
 We compared the result of using VISLCG3 for the same task; table~\ref{table:results} shows the differences in results in both sentence and word level.
@@
     \hline
-	Pride and Prejudice & todo & 5457/7667 \\  
+	Pride and Prejudice & todo & 5k/7k \\  
     \hline
@@
   \end{tabular}
-  \caption{Results for comparing SATCG and VISLCG3.}
+  \caption{Results for comparing CGSAT and VISLCG3.}
   \label{table:results}
@@
 
-These tendencies show also in a larger context with more complex rules.
+The tendency of needing less rules to disambiguate shows also in a larger context with more complex rules.
+In many of the differences, CGSAT disambiguated more. (TODO counts)
 

### 2015-03-25 CG_SAT.hs
minor variable renaming

@@ -118,10 +118,10 @@ applyRules rule (conds:cs) allToks = applyRules rule cs allToks ++
      -- cause is e.g. "-1 is noun" and consequence is "remove verb"
      -- needed because the word at -1 could have many tags, and they could conflict.
         mkVars :: [(Token,[[Token]])] -> (Bit -> Bit) -> [[Bit]]
-        mkVars tctx nt' = [ nt' conseq:ants | (t, ctx) <- tctx 
+        mkVars tctx nt' = [ nt' conseq:ants | (t, ctx) <- tctx -- :: (Token,[[Token]])
+                                            , tCombs <- sequence ctx  -- :: [[Token]]
                                             , let conseq = getBit t
-                                            , antCombs <- sequence ctx  -- :: [[Token]]
-                                            , let ants = map (nt . getBit) antCombs ] 
+                                            , let ants = map (nt . getBit) tCombs ] 
        -- sequence: say we have rule REMOVE v IF (-1 det LINK 2 n)
        -- and we get [ [(1,det)], [(3,n pl), (3,n sg)] ]
        -- we can't just put all of them in the list of antecedents,
@@ -207,13 +207,13 @@ disambiguate verbose rules sentence = do
   bitsForRules <- sequence [ newBit s | _ <- rules ]
   let toks = zip chunkedSent bitsForTags
       rlsBits = zip rules bitsForRules
-      unambig = anchor toks :: [[Bit]]
-      isAmbig = any (\x -> length x > 1) unambig 
+      allNotFalse = anchor toks :: [[Bit]]
+      isAmbig = any (\x -> length x > 1) allNotFalse 
       applied = [ nt x:c | (r,x) <- rlsBits
                          , c     <- applyRule toks r
                          , (not.null) c  ]
   
-  let usedrules = [ show rule ++ "\n" ++ show btags
+  let addedClauses = [ show rule ++ "\n" ++ show btags
                       | (brl:btags) <- applied 
                       , (rule,bit) <- rlsBits
                       , bit == nt brl ] -- brl is negated in the implication
@@ -221,12 +221,12 @@ disambiguate verbose rules sentence = do
         putStrLn "\ntokens:"
         mapM_ print toks
         putStrLn "\nformulas:"
-        mapM_ print unambig
-        --mapM_ putStrLn usedrules
+        mapM_ print allNotFalse
+        --mapM_ putStrLn addedClauses
 
 
   
[…]

### 2015-03-25 paper/nodalida2015.tex
change CGSAT to SATCG (it's not a type of SAT, it's a type of CG :-P)
add numbers for p&p and don quijote

@@
 We took small constraint grammars (\textless{}500 rules) for Spanish and English, and disambiguated a number of texts from Project Gutenberg.
-We compared the result of using VISLCG3 for the same task; table~\ref{table:results} shows the differences in results in both sentence and word level.
+Table~\ref{table:results} shows how much the results agree with the result given by VISL CG3 parser with the same grammar and text.
 
@@
   \centering
-  \begin{tabular}{|r|r|r|r|r|}
+  \begin{tabular}{|l|c|c|}
      \hline
-    \textbf{Text} & \textbf{Diff words} & \textbf{Diff sentences} \\
+    \textbf{Text} & \textbf{Words} & \textbf{\% same analyses} \\
     \hline
-	Pride and Prejudice & todo & 5k/7k \\  
+	Pride and Prejudice & 153386 & 90.6 \\  
     \hline
-       Don Quijote & todo & 3k/6k  \\ 
+       Don Quijote & 186460 & 96.3  \\ 
      \hline
   \end{tabular}
-  \caption{Results for comparing CGSAT and VISLCG3.}
+  \caption{Results for comparing SATCG and VISLCG3.}
   \label{table:results}
@@
 The tendency of needing less rules to disambiguate shows also in a larger context with more complex rules.
-In many of the differences, CGSAT disambiguated more. (TODO counts)
+In many of the differences, SATCG disambiguated more. (TODO counts)
 

### 2015-03-25 CG_SAT.hs
small cleanup

@@ -106,20 +106,16 @@ applyRules rule (conds:cs) allToks = applyRules rule cs allToks ++
 
 
   where
-        --getContext must return something for mkVars to work;
-        -- if the TagSet part in Condition is null,
-        -- getContext returns just the word itself as the context.
+        -- check if getContext has returned non-empty context for each condition
         allCondsHold :: [[Token]] -> Bool
         allCondsHold ts | null conds = True 
                         | otherwise  = all (not.null) ts
 
 
-     -- cause => consequence    translates into     Not cause || consequence.
-     -- cause is e.g. "-1 is noun" and consequence is "remove verb"
-     -- needed because the word at -1 could have many tags, and they could conflict.
+     -- `foo => bar' translates into `nt foo || bar'
         mkVars :: [(Token,[[Token]])] -> (Bit -> Bit) -> [[Bit]]
         mkVars tctx nt' = [ nt' conseq:ants | (t, ctx) <- tctx -- :: (Token,[[Token]])
-                                            , tCombs <- sequence ctx  -- :: [[Token]]
+                                            , tCombs <- sequence ctx -- :: [[Token]]
                                             , let conseq = getBit t
                                             , let ants = map (nt . getBit) tCombs ] 
        -- sequence: say we have rule REMOVE v IF (-1 det) (1 n)
@@ -139,21 +135,16 @@ applyRules rule (conds:cs) allToks = applyRules rule cs allToks ++
  
         -- other: analyses that don't have the wanted readings,
         -- but some word in the same location does have the wanted reading(s)
-        other :: TagSet -> [(Token,[[Token]])]
-        other tags = [(tok, context) | tok <- allWithReading tags
-                                     , not (tagsMatchRule tags tok)
-                                     , let context = getContext tok allToks conds]
-                                     -- , allCondsHold context]
-                                --no need to check allCondsHold; it comes from allWithReadings
-
- 
-        -- all words that have the desired reading in one of their analyses. e.g.
         --      allToks = [(1,[N,Pl]), (2,[V,Sg]), (2,[N,Pl])]
         --      tags    = [V]
-        -- ====> `chosen tags' will return (2,[V,Sg)
-        -- ====> (2,[V,Sg]) and (2,[N,Pl]) are returned
-        allWithReading tags = intersectBy sameInd allToks wantedToks
-          where wantedToks = map fst $ chosen tags
+        -- ====> `chosen tags' will return (2,[V,Sg])
[…]

### 2015-03-25 test/Test.hs
add more detailed printouts

@@ -21,34 +21,63 @@ main = do
     [r,d,"-2","-v"] -> go r d True True
     [r,d,"-v2"] -> go r d True True
     [r,d,"-2v"] -> go r d True True
-    _        -> do putStrLn "usage: ./test <rules> <data>"
+    _           -> putStrLn "usage: ./test <rules> <data>"
   where go r d v is2 = do rules <- readRules r
-                          data' <- readData d
-                          result <- mapM (disambiguate False rules) data' -- :: [Sentence]
-                          goldst <- gold r d is2  -- :: [Sentence]
-                          let diff = [ (diffByWord r g orig)
-                                       | (r,g,orig) <- zip3 result goldst data' ]
-                                  --     , null $ intersect r g ]
-                          when v $ (mapM_ . mapM_) prDiff diff
-                          putStr "Sentences in the original: "
-                          print (length goldst)
-                          putStr "Sentences that differ from vislcg3: "
-                          print $ length $ filter (not.null) diff
-        prDiff (a1,a2,s) = do putStrLn "Original sentence:"
-                              putStrLn (showSentence s)
-                              putStrLn "\nDisambiguation by satcg"
-                              putStrLn $ showAnalysis a1
-                              putStrLn "\nDisambiguation by vislcg3"
-                              putStrLn $ showAnalysis a2
-                              putStrLn "---------------\n"
+                          text <- readData d
+                          resSAT <- mapM (disambiguate False rules) text -- :: [Sentence]
+                          resVISL <- vislcg3 r d is2  -- :: [Sentence]
+                          let diffBS = [ (orig, diff)
+                                          | (s,v,orig) <- zip3 resSAT resVISL text
+                                          , let diff = diffBySent s v
+                                          , (not.null) diff ]
+                              moreD = [ satDisMore
+                                          | (s,v) <- zip resSAT resVISL
+                                          , let satDisMore = moreDisamb s v
+                                          , (not.null) satDisMore ]
+                              
+                          when v $ do
+                            mapM_ prDiff diffBS
 
-diffByWord :: Sentence -> Sentence -> Sentence -> [(Analysis,Analysis,Sentence)]
-diffByWord s1 s2 orig = [ (a1, a2, orig) | (a1, a2) <- zip s1 s2
-                                         , sort a1 /= sort a2 ]
[…]

### 2015-03-26 paper/nodalida2015.tex
change structure

@@
 
-\title{Towards something :-D}
+\title{Exploring Constraint Grammar as a SAT problem}
 
@@
 \maketitle
-\begin{abstract}
-We implement Constraint Grammar as a Boolean satisfiability problem.
 
-We experiment with different strategies to maximise the number of instances to apply rules;
-this will also lead to different orderings of the rules and ways of solving conflicts.
-
-\end{abstract}
+%\begin{abstract}
+%We implement Constraint Grammar as a Boolean satisfiability problem.
+%We experiment with different strategies to maximise the number of instances to apply rules;
+%this will also lead to different orderings of the rules and ways of solving conflicts.
+%\end{abstract}
 
 \section{Introduction}
-Constraint Grammar (CG) was first introduced by \cite{karlsson1995constraint}. 
-It disambiguates output by morphological analyser, 
-and has been used for many tasks in computational linguistics, such as POS tagging,
-surface syntax and machine translation.
-The current state-of-the-art parser VISL CG-3 is implemented (...);
-in addition there are a number of finite-state based implementations \cite{nemeskey14}.
+%Constraint Grammar (CG) was first introduced by \cite{karlsson1995constraint}. 
+%It disambiguates output by morphological analyser, 
+%and has been used for many tasks in computational linguistics, such as POS tagging,
+%surface syntax and machine translation.
+%The current state-of-the-art parser VISL CG-3 is implemented (...);
+%in addition there are a number of finite-state based implementations \cite{nemeskey14}.
 
-We represent CG as a Boolean satisfiability problem using the 
-open-source SAT solver MiniSat \cite{een04sat}.
-Our implementation is inspired by approaches like \cite{lager98,lager_nivre01},
-but we try to emulate the CG formalism more closely than the previous logic-based works.
+We represent Constraint Grammar \cite{karlsson1995constraint} 
+as a Boolean satisfiability (SAT) problem.
+%This increases the computational complexity, 
+%but we are exploring ways in which it could simplify the grammar writing:
+%our experiments show that less rules are needed, and features like ordering of the rules
+%and requirement of being unambiguously tagged become irrelevant.
+% Our implementation is inspired by \cite{lager98,lager_nivre01},
+%but we try to emulate the CG formalism more closely than the previous logic-based works.
+Let us demonstrate our approach with the following example in Spanish.
 
-% NP-complete -- overkill for CG? But gives other benefits.
+\begin{verbatim}
+"<la>"
+        "el" det def f sg
+        "lo" prn pro p3 f sg
+"<casa>"
+        "casa" n f sg
+        "casar" vblex pri p3 sg
+\end{verbatim}
 
-\section{Related work}
-\label{sect:related}
-
-Application of logic to POS tagging or shallow parsing isn't new;
-\cite{lager98,lager_nivre01} present constraint rules as 
-a disjunctive logic program.
-\cite{lindberg_eineborg98ilp,asfrent14} use Inductive Logic Programming to learn CG rules from a tagged corpus.
-\cite{lager01transformation} discusses ordering of the CG rules.
-
-Works from the late 90s and early 00s mention that the method is very slow;
-however, advances in SAT solving technique
-
-\section{CG as a SAT problem}
-\label{sect:pdf}
-
-SAT solving is based on a technique called unit propagation:
-a set of complex Boolean clauses is simplified, starting from unit
-clauses, which consist of just a single variable, and working up to a
-solution, where all variables in the set have a True or False value.
-We model an ambiguous sentence as a set of clauses, where each analysis is a variable, 
-and constraint rules form clauses operating on those variables.
+The ambiguous passage can be either a noun phrase, \emph{la\texttt{:det} casa\texttt{:n}} 
+`the house'  or a verb phrase \emph{la\texttt{:prn} casa\texttt{:v}} `(he/she) marries her'. 
+We add the following rules:
 
 \begin{itemize}
-\item The condition of being unambiguously tagged becomes less necessary.
-We add many clauses, knowing that some of them end up being false.
+\item [] \texttt{REMOVE prn IF (1 n) ;} \\
+             \texttt{REMOVE det IF (1 vblex) ;}
+\end{itemize}
+
+Standard CG will apply one\footnote{To be more cautious, we could require the word at position +1 be disambiguated fully (\texttt{1C} instead of \texttt{1}), but in that case, 
+neither of the rules would be applied.} of the rules to the word \emph{la}; 
+either the one that comes first, or by some other heuristic. 
+The second rule will not fire, because it would remove the last reading. 
+Both readings of \emph{casa} are left untouched by these rules.
+
+SAT solver 
+performs a search, 
+and starts building possible models that satisfy both constraints. 
+In addition to the given constraints, we have default rules to emulate
+the CG principles: an analysis is true if no rule affects it,
+and at least one analysis for each word is true---the notion of ``last'' is not applicable.
+With these constraints, we get two solutions: \texttt{det n} and \texttt{prn v}. 
+The interaction of the rules regarding \emph{la}  disambiguates \emph{casa} 
+for free, and ordering of the rules doesn't matter. 
+%Without more context or additional rules we cannot fully disambiguate the passage,
+%but unlike with standard CG, we get only legal combinations in one solution.
+
+%If we add a word, \emph{la casa grande} `the big house' 
+%and a rule that removes verb if followed by an adjective
+%\texttt{REMOVE (vblex) IF (1 (adj)) ;}
+%we will get a unique solution with \texttt{det n adj}.
+%Standard CG will also remove the verb reading, 
+%but the choice of the word \emph{la} doesn't contribute to
+
+The most important differences between the traditional and the SAT-based approach are summarised below:
+
+\begin{itemize}
+\item The condition of being unambiguously tagged becomes irrelevant.
+Rather than waiting for a word to get disambiguated, the SAT solver starts by 
+making assumptions (e.g. ``\emph{casa} is a noun'') and working under them,
+discarding the assumption if it doesn't lead to a model that satisfies all constraints.
+
 \item Rules are unordered. 
@@
 
-% Let us demonstrate the approach with the example sentence 
-% \emph{the bear sleeps}. After applying morphological tagger, we get the following analyses:
+%\section{CG as a SAT problem}
+%\label{sect:satcg}
 
-% \begin{verbatim}
-% "<the>"
-%         "the" det
-% "<bear>"
-%         "bear" n sg
-%         "bear" v pl
-% "<sleeps>"
-%         "sleep" n pl
-%         "sleep" v sg p3
-% \end{verbatim}
+%SAT solving is based on a technique called unit propagation:
+%a set of complex Boolean clauses is simplified, starting from unit
+%clauses, which consist of just a single variable, and working up to a
+%solution, where all variables in the set have a True or False value.
+%We model an ambiguous sentence as a set of clauses, where each analysis is a variable, 
+%and constraint rules form clauses operating on those variables.
 
-%We make each analysis a variable; let us call them \texttt{d1}, \texttt{n2}, \texttt{v2}, \texttt{n3} and \texttt{v3} by the proposed POS tag and the position in the sentence.
-%For instance, \texttt{n2} is a hypothesis that the correct analysis 
-%for the word form \emph{bear}' is \texttt{"bear" n pl}.
-%Before applying rules, we must make sure that in each position
-%at least one of the analyses is true. We add our first disjunctions:
-
-%\begin{itemize}
-%\item [] $d1$ \\ $n2 \vee v2$ \\ $n3 \vee v3$
-%\end{itemize}
-
-%Let us add two rules:
-%\begin{itemize}
-%\item [] \texttt{REMOVE v IF -1 det} \\ 
-%            \texttt{REMOVE n IF -1 n} 
-%\end{itemize}
-
-%The rules are applied as implications: \emph{if ``the'' is a determiner, then ``bear'' cannot be a verb}.
-
-%\begin{itemize}
-%\item [] $d1 \Rightarrow \neg v2$ \\
-%           $n2   \Rightarrow \neg n3$
-%\end{itemize}
-
-\subsection{Rule ordering}
-\label{ssec:ordering}
+\section{Rule ordering}
+\label{sec:ordering}
 
@@
 
-\subsection{Rule interaction}
-
-The following example illustrates the interaction of rules to produce the desired outcome. We have an ambiguous passage in Spanish, with the following analyses:
-
-\begin{verbatim}
-"<la>"
-        "el" det def f sg
-        "lo" prn pro p3 f sg
-"<casa>"
-        "casa" n f sg
-        "casar" vblex pri p3 sg
-        "casar" vblex imp p2 sg
-\end{verbatim}
-
-We add the following rules to disambiguate the word \emph{la}:
-\begin{itemize}
-\item [] \texttt{REMOVE prn IF (1 n) ;} \\
-             \texttt{REMOVE det IF (1 vblex) ;}
-\end{itemize}
-
-Standard CG will apply first the rule that removes the pronoun reading, because it comes first in the rule file.
-The second rule will not fire, because it would remove the last reading for the word \emph{la}, and that is not allowed.
-To be more cautious, we could require the word at position +1 be disambiguated fully (\texttt{1C} instead of \texttt{1}), but in that case, neither of them would fire.
-The ambiguity between the readings of \emph{casa} is left untouched by these rules.
-
-With the SAT clauses, we get two solutions: ``la\texttt{:det} casa\texttt{:n}''  and ``la\texttt{:prn} casa\texttt{:v}''.
-The interaction of the rules regarding \emph{la}  disambiguates \emph{casa} for free, and ordering of the rules doesn't matter.
-Since we don't have other rules nor further context, we still get two solutions, but there are no illegal combinations (\texttt{det v} or \texttt{prn n}) in one solution.
-
 \paragraph{Rules can be flipped.}
@@
 
+\section{Related work}
+\label{sect:related}
+
+Application of logic to POS tagging or shallow parsing isn't new;
+\cite{lager98,lager_nivre01} present constraint rules as 
+a disjunctive logic program.
+\cite{lindberg_eineborg98ilp,asfrent14} use Inductive Logic Programming to learn CG rules from a tagged corpus.
+\cite{lager01transformation} discusses ordering of the CG rules.
+
 \section{Results and evaluation}
 
+\paragraph{Time}
+
+SAT is NP-complete, whereas the current implementations of CG are polynomial, 
+but with advances in SAT solving techniques, the performance isn't as catastrophically bad as in 90s-00s.
+We used the open-source SAT solver MiniSat \cite{een04sat}.
+We do foreach sentence: foreach rule; 
+%whereas VISL CG3 has foreach rule: foreach sentence.
+the size of the SAT problem is dependent of the sentence (???).
+TODO: time it comparing it to VISL CG3.
+Note also that this is a first proof-of-concept implementation and could be optimised.
+
+\paragraph{Performance against VISL CG3}
 We took small constraint grammars (\textless{}500 rules) for Spanish and English, and disambiguated a number of texts from Project Gutenberg.
@@
 The tendency of needing less rules to disambiguate shows also in a larger context with more complex rules.
-In many of the differences, SATCG disambiguated more. (TODO counts)
+SATCG disambiguated more 17 % of the cases for P&P,
+
+\paragraph{Removing rules}
+
+In this experiment, we test the performance of the two implementations by working with a smaller grammar and comparing the result to
 

### 2015-03-26 paper/nodalida2015.tex
more text; now at 831 words

@@
 %\setlength\titlebox{6.5cm}  % You can expand the title box if you really have to
+
+\usepackage{color}
+\newcommand{\todo}[1]{{\color{cyan}\textbf{[TODO: }#1\textbf{]}}}
 
@@
 %and requirement of being unambiguously tagged become irrelevant.
-% Our implementation is inspired by \cite{lager98,lager_nivre01},
-%but we try to emulate the CG formalism more closely than the previous logic-based works.
 Let us demonstrate our approach with the following example in Spanish.
@@
 
-\begin{itemize}
-\item The condition of being unambiguously tagged becomes irrelevant.
+\paragraph{Condition of being unambiguously tagged becomes irrelevant.}
 Rather than waiting for a word to get disambiguated, the SAT solver starts by 
@@
 
-\item Rules are unordered. 
-\item Rules interact with each other and disambiguate more than same rules in the standard implementation.
+\paragraph{Rules disambiguate more.}
+% Rules interact with each other and disambiguate more than same rules in the standard implementation. 
+In standard CG, the rules only tell to remove readings for the word which is tagged as 
+\texttt{prn} or \texttt{det}. The SAT-based implementation interprets the rules as ``noun and pronoun together are illegal'', and is free to take action that concerns also the following word.
+This means that the rules can be flipped: compare the rules below to the ones in the example.
+
+\begin{itemize}
+\item [] \texttt{REMOVE n IF (-1 prn) ;} \\
+             \texttt{REMOVE vblex IF (-1 det) ;}
 \end{itemize}
 
-%\section{CG as a SAT problem}
-%\label{sect:satcg}
+SAT-based approach gives identical results with both sets of rules,
+whereas the standard CG would remove one reading from \emph{casa} and leave \emph{la} ambiguous.
 
-%SAT solving is based on a technique called unit propagation:
-%a set of complex Boolean clauses is simplified, starting from unit
-%clauses, which consist of just a single variable, and working up to a
-%solution, where all variables in the set have a True or False value.
-%We model an ambiguous sentence as a set of clauses, where each analysis is a variable, 
-%and constraint rules form clauses operating on those variables.
+\todo{Testing this property with more text, more complex rules and a larger rule set. }
 
-\section{Rule ordering}
-\label{sec:ordering}
+\paragraph{Rules are unordered.}
+A typical CG rule sequence starts with rules that are either very common, or have a complex condition to fulfil.
+Stronger rules appear towards the end, to be applied if none of the weaker rules has fired. 
+For a rule set such as following, this order is crucial:
 
-A typical CG rule sequence starts with weaker rules, and leaves stronger rules towards the end, 
-to be applied if none of the weaker rules has fired.
-
-Unordered rules are not new: finite-state implementations of CG \cite{koskenniemi92} also have logically unordered rules.
-
-% VISLCG ad hoc ordering:
-% 1. SELECT before REMOVE.
-% 2. SELECT rules targetting more preferred tags before 
-%   rules targetting less preferred targets.
-% 3. REMOVE rules targetting less preferred tags before
- %   rules targetting more preferred targets.
-% 4. By order of appearance in rule file.
-
-In CG3, the rules are applied in the order of appearance in the rule file.
-
-Different strategies for maximising: ...
-
-\paragraph{Rules can be flipped.}
-Since implications can be presented as disjunctions, we can 
-The representation of the implications as disjunctions 
 \begin{itemize}
-\item [] \texttt{REMOVE v IF -1 det} \\ $\rightarrow$ \texttt{REMOVE det IF 1 verb}   
-\item [] \texttt{REMOVE adv IF NOT 1C adj} \\ $\rightarrow$ \texttt{REMOVE (*) - Adj  IF -1 adv}  \\
+\item [] \texttt{SELECT Year IF (-1 Seasons OR Months) ;} \\
+             \texttt{REMOVE Year ;}
 \end{itemize}
 
-With this experiment, we move even further away from a direct ``if this then that'' model of CG.
+Finite-state implementations of CG \cite{koskenniemi92} \footnote{\todo{Inari: check Tapanainen 1996 if there's anything about ordering}} also have logically unordered rules.
+
+To make up for the lack of sequential order, we have experimented with different strategies of maximising and solving conflicts:
+\begin{itemize}
+\item Number of rules used
+\item Number of instances when a rule is applied
+\item 
+\end{itemize}
 
@@
 
-Application of logic to POS tagging or shallow parsing isn't new;
-\cite{lager98,lager_nivre01} present constraint rules as 
-a disjunctive logic program.
+Our work is inspired by \cite{lager98}, which presents constraint rules as a disjunctive logic program,
+and \cite{lager_nivre01}, which reconstructs 4 different formalisms from a logical point of view.
 \cite{lindberg_eineborg98ilp,asfrent14} use Inductive Logic Programming to learn CG rules from a tagged corpus.
@@
 the size of the SAT problem is dependent of the sentence (???).
-TODO: time it comparing it to VISL CG3.
+
+The table below shows
 Note also that this is a first proof-of-concept implementation and could be optimised.
+
+\begin{table}
+  \centering
+  \begin{tabular}{|l|c|c|}
+     \hline
+	         & Quijote      & Pride \\ \hline
+  VISL CG3 & 0m10.745s & 0m4.811s \\ \hline
+  SATCG    & 2m53.708s & 0m55.705s \\ \hline
+  \end{tabular}
+  \caption{Comparing time between SATCG and VISL CG3.}
+  \label{table:time}
+\end{table}
 
@@
   \end{tabular}
-  \caption{Results for comparing SATCG and VISLCG3.}
+  \caption{Comparing analyses between SATCG and VISL CG3.}
   \label{table:results}
@@
 The tendency of needing less rules to disambiguate shows also in a larger context with more complex rules.
-SATCG disambiguated more 17 % of the cases for P&P,
+SATCG disambiguated more 17 \% of the cases for Pride and Prejudice, and 28 \% of the cases for Don Quijote.
 

### 2015-03-26 paper/nodalida2015.tex
removed some text, added some TODOs

@@
 A typical CG rule sequence starts with rules that are either very common, or have a complex condition to fulfil.
-Stronger rules appear towards the end, to be applied if none of the weaker rules has fired. 
+Stronger rules appear towards the end, to be applied if none of the weaker rules has. 
 For a rule set such as following, this order is crucial:
@@
 
-Finite-state implementations of CG \cite{koskenniemi92} \footnote{\todo{Inari: check Tapanainen 1996 if there's anything about ordering}} also have logically unordered rules.
+To replace the sequential order, we have experimented with different strategies:
+\begin{itemize}
+\item Maximise number of rules used---assume if someone has written a rule, it is worth applying. Didn't give the best results.
+\item Maximise number of instances when a rule is applied---\todo{comment about this too}
+\end{itemize}
 
-To make up for the lack of sequential order, we have experimented with different strategies of maximising and solving conflicts:
-\begin{itemize}
-\item Number of rules used
-\item Number of instances when a rule is applied
-\item 
-\end{itemize}
+% Finite-state implementations of CG \cite{koskenniemi92} also have logically unordered rules. 
+% \todo{Inari: check if Tapanainen 1996 discuss ordering}
+
+These three features are likely to change the way that rules are written. \todo{more speculation, examples}
 
@@
 SAT is NP-complete, whereas the current implementations of CG are polynomial, 
-but with advances in SAT solving techniques, the performance isn't as catastrophically bad as in 90s-00s.
+but with advances in SAT solving techniques, 
+the performance is much more feasible than in the previous works done in 90s--00s.
 We used the open-source SAT solver MiniSat \cite{een04sat}.
-We do foreach sentence: foreach rule; 
+%We do foreach sentence: foreach rule; 
 %whereas VISL CG3 has foreach rule: foreach sentence.
-the size of the SAT problem is dependent of the sentence (???).
+%the size of the SAT problem is dependent of the sentence (???)
 
-The table below shows
-Note also that this is a first proof-of-concept implementation and could be optimised.
+Table~\ref{table:time} shows the execution time compared to VISL CG3.
+SATCG is just a first proof-of-concept implementation and can be optimised.
 
@@
      \hline
-	         & Quijote      & Pride \\ \hline
+	         & \textbf{Don Quijote}  & \textbf{Pride and Prejudice} \\ \hline
   VISL CG3 & 0m10.745s & 0m4.811s \\ \hline
@@
 The tendency of needing less rules to disambiguate shows also in a larger context with more complex rules. 
-SATCG disambiguated more 17 \% of the cases for Pride and Prejudice, and 28 \% of the cases for Don Quijote.
+SATCG disambiguated more 17 \% of the cases for Pride and Prejudice, and 28 \% of the cases for Don Quijote. \todo{Analysis of those cases}
 
@@
 
-In this experiment, we test the performance of the two implementations by working with a smaller grammar and comparing the result to
+In this experiment, we test the performance of the two implementations by working with a smaller grammar.
+We compare how much removing the rules affects the result, gold standard being for each implementation the result given by the full rule set.
+\todo{Perform this experiment}
 

### 2015-03-26 CG_SAT.hs
trying to include bits for both rules and instances, doesn't work

@@ -74,8 +74,8 @@ anchor toks = (map.map) getBit (groupBy sameInd toks)
 
 
 -- | Apply rules to tokens. 
-applyRule :: [Token] -> Rule -> [[Bit]]
-applyRule toks rule = --trace (show rule) $
+applyRule :: Rule -> [Token] -> [[Bit]]
+applyRule rule toks = --trace (show rule) $
   case rule of
 -- a) condition(s) must not hold 
 --for each token, if getContext is empty, remove/select it
@@ -114,15 +114,15 @@ applyRules rule (conds:cs) allToks = applyRules rule cs allToks ++
 
      -- `foo => bar' translates into `nt foo || bar'
         mkVars :: [(Token,[[Token]])] -> (Bit -> Bit) -> [[Bit]]
-        mkVars tctx nt' = [ nt' conseq:ants | (t, ctx) <- tctx -- :: (Token,[[Token]])
-                                            , tCombs <- sequence ctx -- :: [[Token]]
-                                            , let conseq = getBit t
-                                            , let ants = map (nt . getBit) tCombs ] 
-       -- sequence: say we have rule REMOVE v IF (-1 det) (1 n)
-       -- and we get [ [(1,det)], [(3,n pl), (3,n sg)] ]
-       -- we can't just put all of them in the list of antecedents,
-       -- because that would require n pl and n sg be true at the same time.
-       -- Instead we make combinations [(1,det) , (3,n sg)] and [(1,det) , (3,n pl)]
+        mkVars tctx nt' = [ conseq:ants | (t, ctx) <- tctx -- (Token,[[Token]])
+                                        , tCombs <- sequence ctx -- :: [[Token]]
+                                        , let conseq = nt' (getBit t)
+                                        , let ants = map (nt . getBit) tCombs ] 
+     -- sequence: say we have rule REMOVE v IF (-1 det) (1 n)
+     -- and we get [ [(1,det)], [(3,n pl), (3,n sg)] ]
+     -- we can't just put all of them in the list of antecedents
+     -- because that would require n pl and n sg be true at the same time.
+     -- Instead we make combinations [(1,det), (3,n sg)] and [(1,det), (3,n pl)]
 
 
         -- chosen: analyses that have the wanted readings and context
@@ -155,10 +155,10 @@ getContext tok allToks []     = []
 getContext tok allToks ((C position (bool,ctags)):cs) = getContext tok allToks cs ++
   case ctags of
     []     -> [[dummyTok]] --empty conds = holds always
[…]

### 2015-03-26 paper/nodalida2015.tex
add comment about flipping rules

@@
 
-\paragraph{Condition of being unambiguously tagged becomes irrelevant.}
+\paragraph{Condition of being unambiguously tagged is irrelevant.}
 Rather than waiting for a word to get disambiguated, the SAT solver starts by 
@@
 
-\todo{Testing this property with more text, more complex rules and a larger rule set. }
+Rules with more conditions translate into many rules; rules with negation become complements (\texttt{(*)-X} for \texttt{NOT X}).
+Rules which require the condition to be unambiguously tagged, don't have an equivalent flip in the standard CG.
+% We did not expect much practical benefits, save for realising that some rules are bad by having to think further what it implies,
+Out of interest, we manually flipped a small grammar and tested it against the original.
+\todo{The results show (the English just shows it's crap in the first place, 84\% to 85\% for a random 100-sentence text) }
+%\begin{itemize}
+%\item []          \texttt{SL Inf IF (-2 V-MOD) (-1 "not");} \\
+%$\rightarrow$\texttt{SL V-MOD   IF (1 "not") (2 INF);} \\
+%$\rightarrow$\texttt{SL "not" IF (-1 V-MOD) (1 Inf);}
+%\item []           \texttt{SL Foo IF (NOT 1 Bar)} \\
+%$\rightarrow$\texttt{SL (*)-Foo if (-1 Bar)}
+%\end{itemize}
 

### 2015-03-27 paper/nodalida2015.tex
stuff

@@
         "casa" n f sg
-        "casar" vblex pri p3 sg
+        "casar" v pri p3 sg
+        "casar" v imp p2 sg
 \end{verbatim}
@@
 \item [] \texttt{REMOVE prn IF (1 n) ;} \\
-             \texttt{REMOVE det IF (1 vblex) ;}
+             \texttt{REMOVE det IF (1 v) ;}
 \end{itemize}
@@
 The second rule will not fire, because it would remove the last reading. 
-Both readings of \emph{casa} are left untouched by these rules.
+All readings of \emph{casa} are left untouched by these rules.
 
@@
 %and a rule that removes verb if followed by an adjective
-%\texttt{REMOVE (vblex) IF (1 (adj)) ;}
+%\texttt{REMOVE (v) IF (1 (adj)) ;}
 %we will get a unique solution with \texttt{det n adj}.
@@
 \item [] \texttt{REMOVE n IF (-1 prn) ;} \\
-             \texttt{REMOVE vblex IF (-1 det) ;}
+             \texttt{REMOVE v IF (-1 det) ;}
 \end{itemize}
@@
 \begin{itemize}
-\item Maximise number of rules used---assume if someone has written a rule, it is worth applying. Didn't give the best results.
-\item Maximise number of instances when a rule is applied---\todo{comment about this too}
+\item Maximise number of rules used---assume if someone has written a rule, it is worth applying. Downside is that if a rule leads to a conflict in one case, it won't be applied anywhere.
+\item Maximise number of instances when a rule is applied---This gave better results. Assume that whenever we can apply a rule, we should do it.
+\item Maximise number of true analyses---With this heuristic, the solution \texttt{prn v} would be chosen, because there happens to be two analyses with \texttt{v}, and choosing that, 3 analyses are true, since there is no rule to disambiguate between imperative and indicative.
+We want analyses to be true by default
 \end{itemize}
@@
 \paragraph{Performance against VISL CG3}
+
+\footnote{https://svn.code.sf.net/p/apertium/svn/branches/apertium-swpost/apertium-en-es/es-tagger-data/es.tagged}
 We took small constraint grammars (\textless{}500 rules) for Spanish and English, and disambiguated a number of texts from Project Gutenberg.

### 2015-03-27 CG_SAT.hs

@@ -77,24 +77,8 @@ anchor toks = (map.map) getBit (groupBy sameInd toks)
 applyRule :: Rule -> [Token] -> [[Bit]]
 applyRule rule toks = --trace (show rule) $
   case rule of
--- a) condition(s) must not hold 
---for each token, if getContext is empty, remove/select it
-    (Remove tags (NEG conds)) -> [nt (getBit tok) | tok <- toks
-                                                    , tagsMatchRule tags tok
-                                                    , null (ctxt tok conds)]:[]
-
-    (Select tags (NEG conds)) -> [getBit tok | tok <- toks
-                                              , tagsMatchRule tags tok
-                                              , null (ctxt tok conds)]:[]
-                              ++ [nt (getBit tok) | tok <- toks
-                                                    , tagsMatchRule tags tok
-                                                    , (not.null) (ctxt tok conds)]:[]
-
--- b) general case, there can be nested ANDs, ORs or just plain rules
-    (Remove tags (POS conds)) -> applyRules rule (toLists conds) toks
-    (Select tags (POS conds)) -> applyRules rule (toLists conds) toks
-
-  where ctxt tok conds = concatMap (getContext tok toks) (toLists conds)
+    (Remove tags conds) -> applyRules rule (toLists conds) toks
+    (Select tags conds) -> applyRules rule (toLists conds) toks
 
 
 applyRules :: Rule -> [[Condition]] -> [Token] -> [[Bit]]

### 2015-03-27 CG_base.hs

@@ -46,12 +46,9 @@ type Analysis = [[Tag]]
 type Sentence = [Analysis]
 
 
--- | Rule is either remove or select a list of tags, with contextual test
-data Rule = Remove TagSet Test | Select TagSet Test deriving (Show)
+-- | Rule is either remove or select a list of tags, with contextual tests
+data Rule = Remove TagSet Condition | Select TagSet Condition deriving (Show)
 
--- | Test is a condition with a possible NEG, to implement CG3's NEGATE
---   NEGATE negates the whole result, NOT just a single clause.
-data Test = NEG Condition | POS Condition deriving (Show)
 
 -- | There is no special constructor for empty condition (ie. remove/select tag everywhere),
 --   but `C _ (_,[])' is assumed to mean that.
@@ -87,9 +84,6 @@ data Position = Exactly Integer
 Everything inside the outer list is constructed by OR (or just C).
 Everything inside an inner list is constructed by AND.
 -}
-
-
-
 toLists :: Condition -> [[Condition]]
 
 --for OR, we make a list of lists, all in sequence.
@@ -125,24 +119,22 @@ toLists cond = case cond of
         simpleList c           = [c]
 
 
--- Shorthand for writing positive tests without barriers
--- TODO better parser
-mkT :: String -> TagSet -> Test
-mkT str tags = POS $ mkC str tags
-
+-- Shorthand for writing conditions without barriers
 mkC :: String -> TagSet -> Condition
 mkC str tags | last str == '*' = C (AtLeast $ (read . init) str) (True, tags)
              | otherwise       = C (Exactly $ read str)          (True, tags)
-
-neg :: Test -> Test
-neg (POS t) = (NEG t)
-neg (NEG t) = (POS t)
 
 lemmaBear :: Condition
 lemmaBear = mkC "0" [[Lem "bear"]]
 always = mkC "0" []
 andTest = AND lemmaBear always
 
+hasBoundary :: Rule -> Bool
+hasBoundary rule = case rule of
+  (Select _t c) -> findBoundary c
+  (Remove _t c) -> findBoundary c
[…]

### 2015-03-27 CG_parse.hs

@@ -172,8 +172,8 @@ transRule :: Rule -> State Env CGB.Rule
 transRule rl = case rl of
   SelectIf _sl tags _if conds -> liftM2 CGB.Select (transTagSet tags) (transCondSet conds)
   RemoveIf _sl tags _if conds -> liftM2 CGB.Remove (transTagSet tags) (transCondSet conds)
-  SelectAlways _sl tags   -> liftM2 CGB.Select (transTagSet tags) (return $ CGB.POS CGB.always)
-  RemoveAlways _sl tags   -> liftM2 CGB.Remove (transTagSet tags) (return $ CGB.POS CGB.always)
+  SelectAlways _sl tags   -> liftM2 CGB.Select (transTagSet tags) (return CGB.always)
+  RemoveAlways _sl tags   -> liftM2 CGB.Remove (transTagSet tags) (return CGB.always)
   MatchLemma (Str lem) rl -> do cgrule <- transRule rl
                                 case cgrule of
                                   CGB.Select ts c -> return $ CGB.Select (cart ts lem) c
@@ -182,13 +182,13 @@ transRule rl = case rl of
         cart ts str = case str of
            ('"':'<':_) -> [[CGB.WF  (strip 2 str), t] | t <- concat ts]
            ('"':_    ) -> [[CGB.Lem (strip 1 str), t] | t <- concat ts]
-           _           -> [[CGB.Lem str        , t] | t <- concat ts]
+           _           -> [[CGB.Lem str          , t] | t <- concat ts]
         
 
-transCondSet :: [Cond] -> State Env CGB.Test
+transCondSet :: [Cond] -> State Env CGB.Condition
 transCondSet cs = do
   conds <- mapM transCond cs
-  return $ CGB.POS $ foldr1 CGB.AND conds
+  return $ foldr1 CGB.AND conds
 
 transCond :: Cond -> State Env CGB.Condition
 transCond c = case c of

### 2015-03-27 test/Test.hs

@@ -10,11 +10,19 @@ import System.Environment
 import System.IO
 import System.Process
 
+ambiguous = "data/ambiguous/es.tagged.ambiguous"
+apertium = "data/apertium-spa.spa.rlx"
+flipped = "data/spa_cg3_flip.rlx"
+nonflipped = "data/spa_onlyinari.rlx"
+
 main :: IO ()
 main = do
   args <- getArgs
   case args of
-    ["gold"] -> gold "data/spa_cg3.rlx" "data/ambiguous/es.tagged.ambiguous" 
+    ["gold","flip"] -> gold flipped ambiguous
+    ["gold","mini"] -> gold "/tmp/jugarcg" ambiguous
+    ["gold","orig"] -> gold apertium ambiguous
+    ["gold"] -> gold nonflipped ambiguous
     [r,d,"-gold"] -> gold r d
     [r,d]      -> go r d False
     [r,d,"-2"] -> go r d True
@@ -39,14 +47,10 @@ main = do
                                           | (sat,visl,orig) <- zip3 s v tx
                                           , let diff = diffBySent sat visl
                                           , (not.null) diff ]
-                              moreD = [ satDisMore
-                                          | (sat,visl) <- zip s v 
-                                          , let satDisMore = moreDisamb sat visl
-                                          , (not.null) satDisMore ]
-                              lessD = [ satDisLess
-                                          | (sat,visl) <- zip s v 
-                                          , let satDisLess = lessDisamb sat visl
-                                          , (not.null) satDisLess ]
+                              moreD = length $ filter (not.null) $ zipWith moreDisamb s v
+                              lessD = length $ filter (not.null) $ zipWith lessDisamb s v
+                              diffD = length $ filter (not.null) $ zipWith noIntersect s v
+                              restD = length $ filter (not.null) $ zipWith diffBySent s v
                                                       
                               
                           when (length diffBS < 100) $ mapM_ prDiff diffBS
@@ -63,10 +67,8 @@ main = do
                           print $ 100 * ((origsents - diffsents) / origsents)
                           putStr "% same words: "
                           print $ 100 * ((origwords - diffwords) / origwords)
-                          putStr "(SAT disambiguates more/SAT disambiguates differently): "
[…]

### 2015-03-29 paper/nodalida2015.tex
more tables

@@
 \usepackage{authordate1-4}
+\usepackage{multirow}
 \special{papersize=210mm,297mm} % to avoid having to use "-t a4" with dvips 
@@
 \newcommand{\todo}[1]{{\color{cyan}\textbf{[TODO: }#1\textbf{]}}}
-
+\usepackage{tikz}
+\def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;}
 \title{Exploring Constraint Grammar as a SAT problem}
@@
 whereas the standard CG would remove one reading from \emph{casa} and leave \emph{la} ambiguous.
+Testing this property with more complex rules and larger rule sets than remains to be done.
 
-Rules with more conditions translate into many rules; rules with negation become complements (\texttt{(*)-X} for \texttt{NOT X}).
+%Rules with more conditions translate into many rules; rules with negation become complements (\texttt{(*)-X} for \texttt{NOT X}).
 Rules which require the condition to be unambiguously tagged, don't have an equivalent flip in the standard CG.
 % We did not expect much practical benefits, save for realising that some rules are bad by having to think further what it implies,
-Out of interest, we manually flipped a small grammar and tested it against the original.
-\todo{The results show (the English just shows it's crap in the first place, 84\% to 85\% for a random 100-sentence text) }
+%Out of interest, we manually flipped a small grammar and tested it against the original.
+% \todo{The results show  }
 %\begin{itemize}
@@
 Stronger rules appear towards the end, to be applied if none of the weaker rules has. 
-For a rule set such as following, this order is crucial:
+%For a rule set such as following, this order is crucial:
 
-\begin{itemize}
-\item [] \texttt{SELECT Year IF (-1 Seasons OR Months) ;} \\
-             \texttt{REMOVE Year ;}
-\end{itemize}
+%\begin{itemize}
+%\item [] \texttt{SELECT Year IF (-1 Seasons OR Months) ;} \\
+ %            \texttt{REMOVE Year ;}
+%\end{itemize}
 
@@
 \begin{itemize}
-\item Maximise number of rules used---assume if someone has written a rule, it is worth applying. Downside is that if a rule leads to a conflict in one case, it won't be applied anywhere.
-\item Maximise number of instances when a rule is applied---This gave better results. Assume that whenever we can apply a rule, we should do it.
-\item Maximise number of true analyses---With this heuristic, the solution \texttt{prn v} would be chosen, because there happens to be two analyses with \texttt{v}, and choosing that, 3 analyses are true, since there is no rule to disambiguate between imperative and indicative.
-We want analyses to be true by default
+\item Maximise number of rules used---assume if someone has written a rule, it is worth applying.
+We discarded this strategy, because it was too strong: if a rule leads to a conflict in one case, it was not be applied anywhere.
+\item [\checkmark] Maximise number of instances when a rule is applied---assume that whenever we can apply a rule, we should do it. 
+\item Maximise number of true analyses---Without this heuristic, both solutions to our example \emph{la casa} are as likely.
+With this heuristic,  the example would be disambiguated as \texttt{prn v}, because there happens to be two analyses with \texttt{v}, and choosing that, 3 analyses are true, since there is no rule to disambiguate between imperative and indicative.
+% We want analyses to be true by default, but still disambiguate as much as possible.
+%We have a rule to maximise the number of true analyses; then we take a solution that has the least. 
+\item [\checkmark] Maximise rules, minimise number of true analyses---To combat the previous, we want each solution to contain as many true analyses as possible, to ensure that analyses untouched by rules are kept true.
+After having multiple solutions which all ensure , we pick the one that disambiguates the most.
+
 \end{itemize}
+
+The heuristics worked with examples in the order of tens of rules, but moving to hundreds of rules, the rules interact more, the results are more mixed. 
+Rule sets in order of thousands of rules remains to be tested.
 
@@
   \centering
-  \begin{tabular}{|l|c|c|}
+  \begin{tabular}{|l|l|c|c|}
      \hline
-	         & \textbf{Don Quijote}  & \textbf{Pride and Prejudice} \\ \hline
-  VISL CG3 & 0m10.745s & 0m4.811s \\ \hline
-  SATCG    & 2m53.708s & 0m55.705s \\ \hline
+	        \textbf{Text} & \textbf{\# words, rules} & \textbf{VISLCG3} &  \textbf{SATCG} \\ \hline
+    \multirow{2}{*}{\textbf{Don Quijote}} & 186460 words  & 10.7s &  2m53.708s \\ 
+		                                           & 261 rules & & \\ \hline
+    \multirow{2}{*}{\textbf{Pride and}}  & 153386 words & 4.811s & 55.705s \\ 
+                         {\textbf{Prejudice}}                   & 99 rules & \\ \hline
   \end{tabular}
@@
 
-\footnote{https://svn.code.sf.net/p/apertium/svn/branches/apertium-swpost/apertium-en-es/es-tagger-data/es.tagged}
-We took small constraint grammars (\textless{}500 rules) for Spanish and English, and disambiguated a number of texts from Project Gutenberg.
+We took a tagged gold standard corpus\footnote{https://svn.code.sf.net/p/apertium/svn/branches/apertium-swpost/apertium-en-es/es-tagger-data/es.tagged} of 21865 words for Spanish, 
+and a small constraint grammar\footnote{https://svn.code.sf.net/p/apertium/svn/languages/apertium-spa/apertium-spa.spa.rlx}, produced independently of the authors.
+% We ignored substitute rules, but kept subreadings and unification, even though our implementation doesn't handle them.
+We only took select and remove rules (no substitute/iff), after which the constraint grammar had 261 rules.
+With this setup, we took the text of the tagged corpus with all ambiguities, and ran both VISLCG3 and our implementation with the same grammar.
+
+% small constraint grammars (\textless{}500 rules) for Spanish and English, and disambiguated a number of texts from Project Gutenberg.
 Table~\ref{table:results} shows how much the results agree with the result given by VISL CG3 parser with the same grammar and text.
 
+%\begin{table}
+% \centering
+%  \begin{tabular}{|l|c|c|}
+%     \hline
+%    \textbf{Text} & \textbf{Words} & \textbf{\% same analyses} \\
+%    \hline
+%	Pride and Prejudice & 153386 & 90.6 \\  
+%    \hline
+%       Don Quijote & 186460 & 96.3  \\ 
+%     \hline
+%  \end{tabular}
+%  \caption{Comparing analyses between SATCG and VISLCG3.}
+%  \label{table:results}
+%\end{table}
+% The tendency of needing less rules to disambiguate shows also in a larger context with more complex rules. 
+% SATCG disambiguated more 17 \% of the cases for Pride and Prejudice, and 28 \% of the cases for Don Quijote. \todo{Analysis of those cases}
+
+\paragraph{Removing rules}
 \begin{table}
@@
      \hline
-    \textbf{Text} & \textbf{Words} & \textbf{\% same analyses} \\
+    \textbf{\# rules} & \textbf{SATCG} & \textbf{VISLCG3} \\
     \hline 
-	Pride and Prejudice & 153386 & 90.6 \\  
-    \hline
-       Don Quijote & 186460 & 96.3  \\ 
-     \hline
+	19 &  83.15 & 83.33 \\ \hline  % 19 out of 23 preselected by authors
+	18 &  83.19 & 83.04 \\ \hline  % 18 out of 19
+	17 &  83.201 & 83.03 \\ \hline   % 17 out of 18/19, no diff
+	16 &  83.197 & 83.030 \\ \hline  % 16 out of 17/18, no diff
+	15 &  83.188 & 83.022\\ \hline  % 15 out of 16/17. Tie with SATCG, chose to retain the rule that gave better results for VISLCG3.
+	14 &  83.180 & 83.005 \\ \hline  
+	13 &  & \\ \hline  
+	12 &  & \\ \hline  
+	11 &  & \\ \hline  
+	10 &  & \\ \hline  
+	2   & 80.23 & 79.97 \\ \hline
+	1   & 76.69 & 76.62 \\ \hline
+	0   & &  \\ \hline
   \end{tabular}
-  \caption{Comparing analyses between SATCG and VISL CG3.}
+  \caption{}
   \label{table:results}
 \end{table}
+In this experiment, we took 19 rules written by the authors, which give 83.15 \% correct results with SATCG and 83.33 \% with VISLCG3.
+Then we removed rules one by one, keeping rules with most significant gain
 
-The tendency of needing less rules to disambiguate shows also in a larger context with more complex rules. 
-SATCG disambiguated more 17 \% of the cases for Pride and Prejudice, and 28 \% of the cases for Don Quijote. \todo{Analysis of those cases}
-
-\paragraph{Removing rules}
-
-In this experiment, we test the performance of the two implementations by working with a smaller grammar.
+test the performance of the two implementations by working with a smaller grammar.
 We compare how much removing the rules affects the result, gold standard being for each implementation the result given by the full rule set.
 \todo{Perform this experiment}
+
+%5195:
+%Select [[det],[detnt],[rel,aa]] (C (Exactly 1) (False,[[pri],[prs],[imp],[pii],[cni],[pis],[ifi],[fti]]))
+%Select [[prn]] (AND (C (Exactly 0) (True,[[adj]])) (C (Exactly 1) (True,[[vblex],[vbmod],[vbhaver],[vbser],[adv],[prn,pro]])))
+%Select [[prn]] (AND (C (Exactly 0) (True,[[adj]])) (C (Exactly (-1)) (True,[[pr]])))
+%Select [[adv]] (C (Exactly 1) (True,[[],["sent"],[cm],[lpar],[rpar],["<;>"],[cnjcoo],[cnjsub],[cnjadv],[rel],[pr],[det],[detnt],[rel,aa]]))
+%Select [[preadv]] (C (Exactly 1) (True,[[pp]]))
+%Remove [[prs,p1],[pis,p1],[pii,p1],[cni,p1]] (C (Exactly (-1)) (True,[[n],[p3]]))
+%Remove [[vblex],[vbmod],[vbhaver],[vbser]] (C (Exactly (-1)) (True,[[vblex]]))
+%Remove [[pri],[prs],[imp],[pii],[cni],[pis],[ifi],[fti]] (AND (C (Exactly 0) (True,[[pr]])) (C (Exactly 1) (True,[[inf]])))
+%Remove [[pri],[prs],[imp],[pii],[cni],[pis],[ifi],[fti]] (C (Exactly (-1)) (True,[[pr]]))
+%6432:
+%Select [[prn,pro],[prn,itg],[prn,tn]] (C (Exactly 1) (True,[[pri],[prs],[imp],[pii],[cni],[pis],[ifi],[fti]]))
+%Select [[prn]] (AND (C (Exactly 0) (True,[[adj]])) (C (Exactly 1) (True,[[vblex],[vbmod],[vbhaver],[vbser],[adv],[prn,pro]])))
+%Select [[prn]] (AND (C (Exactly 0) (True,[[adj]])) (C (Exactly (-1)) (True,[[pr]])))
+%Select [[adv]] (C (Exactly 1) (True,[[],["sent"],[cm],[lpar],[rpar],["<;>"],[cnjcoo],[cnjsub],[cnjadv],[rel],[pr],[det],[detnt],[rel,aa]]))
+%Select [[preadv]] (C (Exactly 1) (True,[[pp]]))
+%Remove [[prs,p1],[pis,p1],[pii,p1],[cni,p1]] (C (Exactly (-1)) (True,[[n],[p3]]))
+%Remove [[vblex],[vbmod],[vbhaver],[vbser]] (C (Exactly (-1)) (True,[[vblex]]))
+%Remove [[pri],[prs],[imp],[pii],[cni],[pis],[ifi],[fti]] (AND (C (Exactly 0) (True,[[pr]])) (C (Exactly 1) (True,[[inf]])))
+%Remove [[pri],[prs],[imp],[pii],[cni],[pis],[ifi],[fti]] (C (Exactly (-1)) (True,[[pr]]))
 

### 2015-03-29 paper/nodalida2015.tex
more stuff
more text

@@
         "el" det def f sg
-        "lo" prn pro p3 f sg
+        "lo" prn p3 f sg
 "<casa>"
@@
 The ambiguous passage can be either a noun phrase, \emph{la\texttt{:det} casa\texttt{:n}} 
-`the house'  or a verb phrase \emph{la\texttt{:prn} casa\texttt{:v}} `(he/she) marries her'. 
+`the house'  or a verb phrase \emph{la\texttt{:prn} casa\texttt{:v pri p3}} `(he/she) marries her'. 
 We add the following rules:
@@
 and at least one analysis for each word is true---the notion of ``last'' is not applicable.
-With these constraints, we get two solutions: \texttt{det n} and \texttt{prn v}. 
-The interaction of the rules regarding \emph{la}  disambiguates \emph{casa} 
+With these constraints, we get two solutions: \texttt{det, n} and \texttt{prn, v pri p3, v imp p2}. 
+The interaction of the rules regarding \emph{la}  disambiguates the POS of \emph{casa} 
 for free, and ordering of the rules doesn't matter. 
@@
 discarding the assumption if it doesn't lead to a model that satisfies all constraints.
+
+\paragraph{Rules are unordered.}
+A typical CG rule sequence starts with rules that are either very common, or have a complex condition to fulfil.
+Stronger rules appear towards the end, to be applied if none of the weaker rules has. 
+%For a rule set such as following, this order is crucial:
+
+%\begin{itemize}
+%\item [] \texttt{SELECT Year IF (-1 Seasons OR Months) ;} \\
+ %            \texttt{REMOVE Year ;}
+%\end{itemize}
+
+To replace the sequential order, we have experimented with different strategies:
+\begin{itemize}
+\item Maximise number of rules used---We discarded this strategy, because it was too strong: if a rule leads to a conflict in one case, it was not be applied anywhere.
+\item Emulate sequential order---Enter clauses produced by each rule one by one, and assume the solver state reached so far is correct; if a conflict is introduced by new clauses, discard them and move on to next.
+\item [\checkmark] Maximise number of rule applications---If there is a conflict, find the smallest number of rule applications to discard so that the conflict is solved.
+\end{itemize}
+
+%combined with the following
+%\begin{itemize}
+%\item Maximise number of true analyses---Without this heuristic, both solutions to our example %\emph{la casa} are as likely.
+With this heuristic,  the example would be disambiguated as \texttt{prn v}, because there happens to be two analyses with \texttt{v}, and choosing that, 3 analyses are true, since there is no rule to disambiguate between imperative and indicative.
+% We want analyses to be true by default, but still disambiguate as much as possible.
+%We have a rule to maximise the number of true analyses; then we take a solution that has the least. 
+%\item Minimise number of true analyses---To combat the previous, we want each solution to contain as many true analyses as possible, to ensure that analyses untouched by rules are kept true.
+%After having multiple solutions which all ensure , we pick the one that disambiguates the most.
+%\end{itemize}
 
@@
 
-\paragraph{Rules are unordered.}
-A typical CG rule sequence starts with rules that are either very common, or have a complex condition to fulfil.
-Stronger rules appear towards the end, to be applied if none of the weaker rules has. 
-%For a rule set such as following, this order is crucial:
-
-%\begin{itemize}
-%\item [] \texttt{SELECT Year IF (-1 Seasons OR Months) ;} \\
- %            \texttt{REMOVE Year ;}
-%\end{itemize}
-
-To replace the sequential order, we have experimented with different strategies:
-\begin{itemize}
-\item Maximise number of rules used---assume if someone has written a rule, it is worth applying.
-We discarded this strategy, because it was too strong: if a rule leads to a conflict in one case, it was not be applied anywhere.
-\item [\checkmark] Maximise number of instances when a rule is applied---assume that whenever we can apply a rule, we should do it. 
-\item Maximise number of true analyses---Without this heuristic, both solutions to our example \emph{la casa} are as likely.
-With this heuristic,  the example would be disambiguated as \texttt{prn v}, because there happens to be two analyses with \texttt{v}, and choosing that, 3 analyses are true, since there is no rule to disambiguate between imperative and indicative.
-% We want analyses to be true by default, but still disambiguate as much as possible.
-%We have a rule to maximise the number of true analyses; then we take a solution that has the least. 
-\item [\checkmark] Maximise rules, minimise number of true analyses---To combat the previous, we want each solution to contain as many true analyses as possible, to ensure that analyses untouched by rules are kept true.
-After having multiple solutions which all ensure , we pick the one that disambiguates the most.
-
-\end{itemize}
-
 The heuristics worked with examples in the order of tens of rules, but moving to hundreds of rules, the rules interact more, the results are more mixed. 
@@
 
-SAT is NP-complete, whereas the current implementations of CG are polynomial, 
+SAT is NP-complete, whereas the standard implementations of CG are polynomial, 
 but with advances in SAT solving techniques, 
@@
      \hline
-	        \textbf{Text} & \textbf{\# words, rules} & \textbf{VISLCG3} &  \textbf{SATCG} \\ \hline
-    \multirow{2}{*}{\textbf{Don Quijote}} & 186460 words  & 10.7s &  2m53.708s \\ 
+	        \textbf{Text} & \textbf{\# words, rules} &  \textbf{SATCG} & \textbf{VISLCG3} \\ \hline
+    \multirow{2}{*}{\textbf{Don Quijote}} & 186460 words  &  2m53.7s & 10.7s \\ 
 		                                           & 261 rules & & \\ \hline
-    \multirow{2}{*}{\textbf{Pride and}}  & 153386 words & 4.811s & 55.705s \\ 
+    \multirow{2}{*}{\textbf{Pride and}}  & 153386 words& 55.7s  & 4.8s \\ 
                          {\textbf{Prejudice}}                   & 99 rules & \\ \hline
   \end{tabular}
-  \caption{Comparing time between SATCG and VISL CG3.}
+  \caption{Comparing execution times between SATCG and VISLCG3.}
   \label{table:time}
@@
 	16 &  83.197 & 83.030 \\ \hline  % 16 out of 17/18, no diff
-	15 &  83.188 & 83.022\\ \hline  % 15 out of 16/17. Tie with SATCG, chose to retain the rule that gave better results for VISLCG3.
-	14 &  83.180 & 83.005 \\ \hline  
-	13 &  & \\ \hline  
-	12 &  & \\ \hline  
-	11 &  & \\ \hline  
-	10 &  & \\ \hline  
-	2   & 80.23 & 79.97 \\ \hline
+	15 &  83.188 & 83.022\\ \hline  % 15 out of 16/17. Tie with 14 for SATCG, chose to retain the rule that gave better results for VISLCG3.
+	14 &  83.180 & 83.005 \\ \hline  % Tie with 15 for SATCG.
+	13 &  83.159 & 82.951 \\ \hline  % 13 out of 15 
+	12 &  83.134 & 82.884\\ \hline  % 12 out of 14
+	11 &  83.089 & 82.847\\ \hline  
+	10 &  83.051 & 82.802\\ \hline  
+	9  & 82.997 & 82.802 \\ \hline
+	8  & 82.906 & 82.681 \\ \hline  % REMOVE:r_imp Imp IF (0 Subj) (-1* CnjSub) ; 
+	7  & 82.743 & 82.594 \\ \hline  % SELECT:s_pro PrnIndep IF (1C VerbFin) ;
+	6  & 82.556 & 82.398\\ \hline  %  REMOVE:r_v_v Verb IF (-1C Vblex) ;
+	5  & 82.327 & 82.028 \\ \hline % SELECT:s_adv Adv IF (1 Cnj_Rel_End OR Prep OR Det)
+	4  & 82.065 & 82.028 \\ \hline % REMOVE:r_vfin_prep VerbFin IF (0 Prep) (1 Inf) 
+	3  & 81.387 & 81.233 \\ \hline % REMOVE:r_noun N    IF (-1 N OR Prn)
+	2  & 80.23 & 79.97 \\ \hline % SELECT:s_pri_1 (pri p3) IF (0 (vblex imp p2 sg))
 	1  & 76.69 & 76.62 \\ \hline 
-	0   & &  \\ \hline
+	0  & 70.70 & 70.70 \\ \hline
   \end{tabular}
-  \caption{}
-  \label{table:results}
+  \caption{Result of removing rules one by one, and how close to the gold standard they get.}
+  \label{table:results_remove}
 \end{table}
 In this experiment, we took 19 rules written by the authors, which give 83.15 \% correct results with SATCG and 83.33 \% with VISLCG3.
-Then we removed rules one by one, keeping rules with most significant gain
-
-test the performance of the two implementations by working with a smaller grammar.
-We compare how much removing the rules affects the result, gold standard being for each implementation the result given by the full rule set.
-\todo{Perform this experiment}
-
-%5195:
-%Select [[det],[detnt],[rel,aa]] (C (Exactly 1) (False,[[pri],[prs],[imp],[pii],[cni],[pis],[ifi],[fti]]))
-%Select [[prn]] (AND (C (Exactly 0) (True,[[adj]])) (C (Exactly 1) (True,[[vblex],[vbmod],[vbhaver],[vbser],[adv],[prn,pro]])))
-%Select [[prn]] (AND (C (Exactly 0) (True,[[adj]])) (C (Exactly (-1)) (True,[[pr]])))
-%Select [[adv]] (C (Exactly 1) (True,[[],["sent"],[cm],[lpar],[rpar],["<;>"],[cnjcoo],[cnjsub],[cnjadv],[rel],[pr],[det],[detnt],[rel,aa]]))
-%Select [[preadv]] (C (Exactly 1) (True,[[pp]]))
-%Remove [[prs,p1],[pis,p1],[pii,p1],[cni,p1]] (C (Exactly (-1)) (True,[[n],[p3]]))
-%Remove [[vblex],[vbmod],[vbhaver],[vbser]] (C (Exactly (-1)) (True,[[vblex]]))
-%Remove [[pri],[prs],[imp],[pii],[cni],[pis],[ifi],[fti]] (AND (C (Exactly 0) (True,[[pr]])) (C (Exactly 1) (True,[[inf]])))
-%Remove [[pri],[prs],[imp],[pii],[cni],[pis],[ifi],[fti]] (C (Exactly (-1)) (True,[[pr]]))
-%6432:
-%Select [[prn,pro],[prn,itg],[prn,tn]] (C (Exactly 1) (True,[[pri],[prs],[imp],[pii],[cni],[pis],[ifi],[fti]]))
-%Select [[prn]] (AND (C (Exactly 0) (True,[[adj]])) (C (Exactly 1) (True,[[vblex],[vbmod],[vbhaver],[vbser],[adv],[prn,pro]])))
-%Select [[prn]] (AND (C (Exactly 0) (True,[[adj]])) (C (Exactly (-1)) (True,[[pr]])))
-%Select [[adv]] (C (Exactly 1) (True,[[],["sent"],[cm],[lpar],[rpar],["<;>"],[cnjcoo],[cnjsub],[cnjadv],[rel],[pr],[det],[detnt],[rel,aa]]))
-%Select [[preadv]] (C (Exactly 1) (True,[[pp]]))
-%Remove [[prs,p1],[pis,p1],[pii,p1],[cni,p1]] (C (Exactly (-1)) (True,[[n],[p3]]))
-%Remove [[vblex],[vbmod],[vbhaver],[vbser]] (C (Exactly (-1)) (True,[[vblex]]))
-%Remove [[pri],[prs],[imp],[pii],[cni],[pis],[ifi],[fti]] (AND (C (Exactly 0) (True,[[pr]])) (C (Exactly 1) (True,[[inf]])))
-%Remove [[pri],[prs],[imp],[pii],[cni],[pis],[ifi],[fti]] (C (Exactly (-1)) (True,[[pr]]))
+Then we removed rules one by one; after getting a score for $n$ rules, we took the combination of $n-1$ rules for which SATCG gives the best score, and we ran the rule set with VISLCG3 for comparison.
+Table~\ref{table:results_remove} shows the percentages compared to the gold standard for both systems.
 

### 2015-03-29 test/Test.hs
small changes

@@ -65,7 +65,7 @@ optiComp rls dat n = do r <- readRules rls
                         res <- loop (filter (\x -> length x==n) $ subseq r) t g []
                         
                         putStrLn "optimal rule sequence:"
-                        mapM_ pr (take 3 (reverse (sortBy (\x y -> fst x `compare` fst y) res)))
+                        mapM_ pr (take 20 (reverse (sortBy (\x y -> fst x `compare` fst y) res)))
 
 loop :: [[Rule]] -> [Sentence] -> [Sentence] -> [(Float, [Rule])] -> IO [(Float, [Rule])]
 loop []     t g scores = return scores

### 2015-03-30 paper/nodalida2015.tex
squeezing a bit

@@
 \paragraph{Rules are unordered.}
-A typical CG rule sequence starts with rules that are either very common, or have a complex condition to fulfil.
-Stronger rules appear towards the end, to be applied if none of the weaker rules has. 
+%A typical CG rule sequence starts with rules that are either very common, or have a complex %condition to fulfil.
+%Stronger rules appear towards the end, to be applied if none of the weaker rules has. 
 %For a rule set such as following, this order is crucial:
@@
 
-To replace the sequential order, we have experimented with different strategies:
+%To replace the sequential order, 
+We have experimented with different strategies:
 \begin{itemize}
@@
 %\end{itemize}
+
+The chosen heuristic produced the best results with examples in the order of tens or hundreds of rules.
+Larger rule sets (thousands of rules) are yet to be tested.
+
+% Finite-state implementations of CG \cite{koskenniemi92} also have logically unordered rules. 
+% \todo{Inari: check if Tapanainen 1996 discuss ordering}
 
@@
 %\end{itemize}
-
-The heuristics worked with examples in the order of tens of rules, but moving to hundreds of rules, the rules interact more, the results are more mixed. 
-Rule sets in order of thousands of rules remains to be tested.
-
-% Finite-state implementations of CG \cite{koskenniemi92} also have logically unordered rules. 
-% \todo{Inari: check if Tapanainen 1996 discuss ordering}
 
@@
      \hline
-	        \textbf{Text} & \textbf{\# words, rules} &  \textbf{SATCG} & \textbf{VISLCG3} \\ \hline
-    \multirow{2}{*}{\textbf{Don Quijote}} & 186460 words  &  2m53.7s & 10.7s \\ 
-		                                           & 261 rules & & \\ \hline
-    \multirow{2}{*}{\textbf{Pride and}}  & 153386 words& 55.7s  & 4.8s \\ 
-                         {\textbf{Prejudice}}                   & 99 rules & \\ \hline
+   \textbf{\# words} & \textbf{\# rules} &  \textbf{SATCG} & \textbf{VISLCG3} \\ \hline
+              153,386    & 99    & 55.7s     & 4.8s \\  %\hline
+                186,460  & 261  &  2m53.7s & 10.7s \\\hline
   \end{tabular}
-  \caption{Comparing execution times between SATCG and VISLCG3.}
+  \caption{Execution times.}
   \label{table:time}

### 2015-03-30 paper/nodalida2015.tex
stuff

@@
 \def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;}
-\title{Exploring Constraint Grammar as a SAT problem}
+\title{Constraint Grammar as a SAT problem}
 
@@
 
-%\begin{abstract}
-%We implement Constraint Grammar as a Boolean satisfiability problem.
-%We experiment with different strategies to maximise the number of instances to apply rules;
-%this will also lead to different orderings of the rules and ways of solving conflicts.
-%\end{abstract}
+%\begin{abstract} \end{abstract}
 
@@
 as a Boolean satisfiability (SAT) problem.
+Applying logic to reductionist grammars has been explored earlier by 
+\cite{lager98,lager_nivre01}, but it was never adopted for use---logic programming was simply 
+too slow to be used for tagging or parsing. 
+However, modern SAT solvers are fast and used for many applications where constraints must be
+satisfied, such as software and hardware verification.
+%SAT solving is based on a technique called unit propagation:
+%a set of complex Boolean clauses is simplified, starting from unit
+%clauses, which consist of just a single variable, and working up to a
+%solution, where all variables in the set have a True or False value.
+
+It is attractive to model CG in terms of logic:
+these techniques are well-studied, and they capture more dependencies between the tags than the standard CG.
+
 %This increases the computational complexity, 
@@
 %and requirement of being unambiguously tagged become irrelevant.
+
+% Uncomment for full paper
+% \section{Related work}
+% \label{sect:related}
+
+% Our work is inspired by \cite{lager98}, which presents constraint rules as a disjunctive logic program,
+% and \cite{lager_nivre01}, which reconstructs 4 different formalisms from a logical point of view.
+% \cite{lindberg_eineborg98ilp,asfrent14} use Inductive Logic Programming to learn CG rules from a tagged corpus.
+% \cite{lager01transformation} discusses ordering of the CG rules.
+
+\section{CG as a SAT problem}
 Let us demonstrate our approach with the following example in Spanish.
@@
 
-SAT solver 
-performs a search, 
+SAT solver performs a search, 
 and starts building possible models that satisfy both constraints. 
@@
 %For a rule set such as following, this order is crucial:
-
 %\begin{itemize}
@@
 
-%To replace the sequential order, 
 We have experimented with different strategies:
 \begin{itemize}
-\item Maximise number of rules used---We discarded this strategy, because it was too strong: if a rule leads to a conflict in one case, it was not be applied anywhere.
+\item Maximise
+% \footnote{Something about local maximum}
+%The algorithm commits to a subset of clauses (e.g.
+\emph{all words must have more than 0 true analyses}), and tries to maximise the number of true variables without changing that subset.
+So it is a local maximum as regards to the given set of clauses.
+number of rules used---We discarded this strategy, because it was too strong: if a rule leads to a conflict in one case, it was not be applied anywhere.
 \item Emulate sequential order---Enter clauses produced by each rule one by one, and assume the solver state reached so far is correct; if a conflict is introduced by new clauses, discard them and move on to next.
@@
 \texttt{prn} or \texttt{det}. The SAT-based implementation interprets the rules as ``noun and pronoun together are illegal'', and is free to take action that concerns also the following word.
-This means that the rules can be flipped: compare the rules below to the ones in the example.
+This means that the rules are logically flipped: compare the rules below to the ones in the example.
 
@@
 whereas the standard CG would remove one reading from \emph{casa} and leave \emph{la} ambiguous.
-Testing this property with more complex rules and larger rule sets than remains to be done.
+Testing this property with more complex rules and larger rule sets remains to be done.
 
@@
 
-\section{Related work}
-\label{sect:related}
-
-Our work is inspired by \cite{lager98}, which presents constraint rules as a disjunctive logic program,
-and \cite{lager_nivre01}, which reconstructs 4 different formalisms from a logical point of view.
-\cite{lindberg_eineborg98ilp,asfrent14} use Inductive Logic Programming to learn CG rules from a tagged corpus.
-\cite{lager01transformation} discusses ordering of the CG rules.
-
 \section{Results and evaluation}
@@
 
-SAT is NP-complete, whereas the standard implementations of CG are polynomial, 
-but with advances in SAT solving techniques, 
-the performance is much more feasible than in the previous works done in 90s--00s.
+The worst-case complexity of SAT is exponential, whereas the standard implementations 
+of CG are polynomial, but with advances in SAT solving techniques, the performance 
+in the average case is much more feasible than in the previous works done in 90s--00s.
 We used the open-source SAT solver MiniSat \cite{een04sat}.
 %We do foreach sentence: foreach rule; 
-%whereas VISL CG3 has foreach rule: foreach sentence.
-%the size of the SAT problem is dependent of the sentence (???)
+%whereas VISL-CG3 has foreach rule: foreach sentence.
+%the size of the SAT problem is dependent of the number of rules
 
-Table~\ref{table:time} shows the execution time compared to VISL CG3.
+Table~\ref{table:time} shows the execution time compared to VISL-CG3. 
+The number of rules in the grammar is 
+Maximisation is the most costly operation
+Also, some of the time is explained simply by
 SATCG is just a first proof-of-concept implementation and can be optimised.
@@
      \hline
-   \textbf{\# words} & \textbf{\# rules} &  \textbf{SATCG} & \textbf{VISLCG3} \\ \hline
-              153,386    & 99    & 55.7s     & 4.8s \\  %\hline
-                186,460  & 261  &  2m53.7s & 10.7s \\\hline
+   \textbf{\# words} & \textbf{\# rules} &  \textbf{SATCG} & \textbf{VISL-CG3} \\ \hline
+              153,386    & 99    & 55.7s     & 4.8s \\  \hline	                         
+                186,460  & 19    &  13.7s      & 4.2s\\ %\hline
+                186,460  & 130  &  1m14.1s & 6.1s \\ %\hline
+                186,460  & 261  &  2m53.7s & 10.7s \\ \hline
   \end{tabular}
@@
 
-\paragraph{Performance against VISL CG3}
+\paragraph{Performance against VISL-CG3}
 
-We took a tagged gold standard corpus\footnote{https://svn.code.sf.net/p/apertium/svn/branches/apertium-swpost/apertium-en-es/es-tagger-data/es.tagged} of 21865 words for Spanish, 
+We took a tagged corpus\footnote{https://svn.code.sf.net/p/apertium/svn/branches/apertium-swpost/apertium-en-es/es-tagger-data/es.tagged} of 21865 words for Spanish, 
 and a small constraint grammar\footnote{https://svn.code.sf.net/p/apertium/svn/languages/apertium-spa/apertium-spa.spa.rlx}, produced independently of the authors.
@@
 We only took select and remove rules (no substitute/iff), after which the constraint grammar had 261 rules.
-With this setup, we took the text of the tagged corpus with all ambiguities, and ran both VISLCG3 and our implementation with the same grammar.
+With this setup, we took the text of the tagged corpus with all ambiguities, and ran both VISL-CG3 and our implementation with the same grammar. 
+Treating the chosen tagset as the gold standard, the disambiguation by SATCG achieves 75.09 \% correct result and VISL-CG3 75.80 \%.
 
-% small constraint grammars (\textless{}500 rules) for Spanish and English, and disambiguated a number of texts from Project Gutenberg.
-Table~\ref{table:results} shows how much the results agree with the result given by VISL CG3 parser with the same grammar and text.
-
-%\begin{table}
+% \paragraph{Removing rules}
+% \begin{table}
 %   \centering
@@
 %      \hline
-%    \textbf{Text} & \textbf{Words} & \textbf{\% same analyses} \\
+%     \textbf{\# rules} & \textbf{SATCG} & \textbf{VISL-CG3} \\
 %     \hline 
-%	Pride and Prejudice & 153386 & 90.6 \\  
-%    \hline
-%       Don Quijote & 186460 & 96.3  \\ 
-%     \hline
+% 	19 &  83.15 & 83.33 \\ \hline  % 19 out of 23 preselected by authors
+% 	18 &  83.19 & 83.04 \\ \hline  % 18 out of 19
+% 	17 &  83.201 & 83.03 \\ \hline   % 17 out of 18/19, no diff
+% 	16 &  83.197 & 83.030 \\ \hline  % 16 out of 17/18, no diff
+% 	15 &  83.188 & 83.022\\ \hline  % 15 out of 16/17. Tie with 14 for SATCG, chose to retain the rule that gave better results for VISL-CG3.
+% 	14 &  83.180 & 83.005 \\ \hline  % Tie with 15 for SATCG.
+% 	13 &  83.159 & 82.951 \\ \hline  % 13 out of 15 
+% 	12 &  83.134 & 82.884\\ \hline  % 12 out of 14
+% 	11 &  83.089 & 82.847\\ \hline  
+% 	10 &  83.051 & 82.802\\ \hline  
+% 	9  & 82.997 & 82.802 \\ \hline
+% 	8  & 82.906 & 82.681 \\ \hline  % REMOVE:r_imp Imp IF (0 Subj) (-1* CnjSub) ; 
+% 	7  & 82.743 & 82.594 \\ \hline  % SELECT:s_pro PrnIndep IF (1C VerbFin) ;
+% 	6  & 82.556 & 82.398\\ \hline  %  REMOVE:r_v_v Verb IF (-1C Vblex) ;
+% 	5  & 82.327 & 82.028 \\ \hline % SELECT:s_adv Adv IF (1 Cnj_Rel_End OR Prep OR Det)
+% 	4  & 82.065 & 82.028 \\ \hline % REMOVE:r_vfin_prep VerbFin IF (0 Prep) (1 Inf) 
+% 	3  & 81.387 & 81.233 \\ \hline % REMOVE:r_noun N    IF (-1 N OR Prn)
+% 	2  & 80.23 & 79.97 \\ \hline % SELECT:s_pri_1 (pri p3) IF (0 (vblex imp p2 sg))
+% 	1  & 76.69 & 76.62 \\ \hline 
+% 	0  & 70.70 & 70.70 \\ \hline
 %   \end{tabular}
-%  \caption{Comparing analyses between SATCG and VISLCG3.}
-%  \label{table:results}
-%\end{table}
-% The tendency of needing less rules to disambiguate shows also in a larger context with more complex rules. 
-% SATCG disambiguated more 17 \% of the cases for Pride and Prejudice, and 28 \% of the cases for Don Quijote. \todo{Analysis of those cases}
-
-\paragraph{Removing rules}
-\begin{table}
-  \centering
-  \begin{tabular}{|l|c|c|}
-     \hline
-    \textbf{\# rules} & \textbf{SATCG} & \textbf{VISLCG3} \\
-    \hline 
-	19 &  83.15 & 83.33 \\ \hline  % 19 out of 23 preselected by authors
-	18 &  83.19 & 83.04 \\ \hline  % 18 out of 19
-	17 &  83.201 & 83.03 \\ \hline   % 17 out of 18/19, no diff
-	16 &  83.197 & 83.030 \\ \hline  % 16 out of 17/18, no diff
-	15 &  83.188 & 83.022\\ \hline  % 15 out of 16/17. Tie with 14 for SATCG, chose to retain the rule that gave better results for VISLCG3.
-	14 &  83.180 & 83.005 \\ \hline  % Tie with 15 for SATCG.
-	13 &  83.159 & 82.951 \\ \hline  % 13 out of 15 
-	12 &  83.134 & 82.884\\ \hline  % 12 out of 14
-	11 &  83.089 & 82.847\\ \hline  
-	10 &  83.051 & 82.802\\ \hline  
-	9  & 82.997 & 82.802 \\ \hline
-	8  & 82.906 & 82.681 \\ \hline  % REMOVE:r_imp Imp IF (0 Subj) (-1* CnjSub) ; 
-	7  & 82.743 & 82.594 \\ \hline  % SELECT:s_pro PrnIndep IF (1C VerbFin) ;
-	6  & 82.556 & 82.398\\ \hline  %  REMOVE:r_v_v Verb IF (-1C Vblex) ;
-	5  & 82.327 & 82.028 \\ \hline % SELECT:s_adv Adv IF (1 Cnj_Rel_End OR Prep OR Det)
-	4  & 82.065 & 82.028 \\ \hline % REMOVE:r_vfin_prep VerbFin IF (0 Prep) (1 Inf) 
-	3  & 81.387 & 81.233 \\ \hline % REMOVE:r_noun N    IF (-1 N OR Prn)
-	2  & 80.23 & 79.97 \\ \hline % SELECT:s_pri_1 (pri p3) IF (0 (vblex imp p2 sg))
-	1  & 76.69 & 76.62 \\ \hline 
-	0  & 70.70 & 70.70 \\ \hline
-  \end{tabular}
-  \caption{Result of removing rules one by one, and how close to the gold standard they get.}
-  \label{table:results_remove}
-\end{table}
-In this experiment, we took 19 rules written by the authors, which give 83.15 \% correct results with SATCG and 83.33 \% with VISLCG3.
-Then we removed rules one by one; after getting a score for $n$ rules, we took the combination of $n-1$ rules for which SATCG gives the best score, and we ran the rule set with VISLCG3 for comparison.
+%   \caption{Result of removing rules one by one, and how close to the gold standard they get.}
+%   \label{table:results_remove}
+% \end{table}
+% In this experiment, we took 19 rules written by the authors, which give 83.15 \% correct results with SATCG and 83.33 \% with VISL-CG3.
+% Then we removed rules one by one; after getting a score for $n$ rules, we took the combination of $n-1$ rules for which SATCG gives the best score, and we ran the rule set with VISL-CG3 for comparison.
 Table~\ref{table:results_remove} shows the percentages compared to the gold standard for both systems.

### 2015-03-30 paper/nodalida2015.tex
more text

@@
 
-\author{Inari Listenmaa \and Koen Claessen \\
-  Chalmers University of Technology, Gothenburg, Sweden \\
-  {\tt \{inari,koen\}@chalmers.se} }
+% \author{Inari Listenmaa \and Koen Claessen \\
+%  Chalmers University of Technology, Gothenburg, Sweden \\
+%  {\tt \{inari,koen\}@chalmers.se} }
 
@@
 as a Boolean satisfiability (SAT) problem.
+
+Formal logic is well-studied and has an old tradition.
+The rules encoded in logic capture richer dependencies between the
+tags than normal reductionist grammars.
+
 Applying logic to reductionist grammars has been explored earlier by 
-\cite{lager98,lager_nivre01}, but it was never adopted for use---logic programming was simply 
-too slow to be used for tagging or parsing. 
-However, modern SAT solvers are fast and used for many applications where constraints must be
-satisfied, such as software and hardware verification.
+\cite{lager98,lager_nivre01}, but it was never adopted for use.
+% ; logic programming was too slow to be used for tagging or parsing. 
+Since those works, SAT solving techniques have improved significantly, and 
+they are used in domains such as microprocessor design and computational 
+biology---these problems easily match or exceed CG in complexity. 
+\todo{is this true for grammars with thousands of rules?}
+
+Due to these advances, we were able to revisit the idea and develop it further. 
+% Whereas \cite{lager98} provides a proof-of-concept description, 
+Missing from the previous logic-based works, 
+we solve eventual rule conflicts by finding a solution that 
+discards the least number of rules.
+% allows the maximal number of rules to be applied.
+We test our implementation by parsing texts in the order of 10,000s--100,000s
+words, using grammars with hundreds of rules.
+
 %SAT solving is based on a technique called unit propagation:
@@
 %solution, where all variables in the set have a True or False value.
-
-It is attractive to model CG in terms of logic:
-these techniques are well-studied, and they capture more dependencies between the tags than the standard CG.
 
@@
 and at least one analysis for each word is true---the notion of ``last'' is not applicable.
-With these constraints, we get two solutions: \texttt{det, n} and \texttt{prn, v pri p3, v imp p2}. 
+With these constraints, we get two solutions: 
+\begin{enumerate}
+\item []
+\begin{verbatim}
+"<la>"
+        "el" det def f sg
+"<casa>"
+        "casa" n f sg
+\end{verbatim}
+\item [and]
+\begin{verbatim}
+"<la>"
+        "lo" prn p3 f sg
+"<casa>"
+        "casar" v pri p3 sg
+        "casar" v imp p2 sg
+\end{verbatim}
+\end{enumerate} 
+% \texttt{det, n} and \texttt{prn, v pri p3, v imp p2}. 
 The interaction of the rules regarding \emph{la}  disambiguates the POS of \emph{casa} 
@@
 %We do foreach sentence: foreach rule; 
-%whereas VISL-CG3 has foreach rule: foreach sentence.
+%whereas VISL CG-3 has foreach rule: foreach sentence.
 %the size of the SAT problem is dependent of the number of rules
 
-Table~\ref{table:time} shows the execution time compared to VISL-CG3. 
+Table~\ref{table:time} shows the execution time compared to VISL CG-3. 
 The number of rules in the grammar is 
@@
   \centering
-  \begin{tabular}{|l|l|c|c|}
+  \begin{tabular}{|l|c|c|c|}
      \hline
-   \textbf{\# words} & \textbf{\# rules} &  \textbf{SATCG} & \textbf{VISL-CG3} \\ \hline
-              153,386    & 99    & 55.7s     & 4.8s \\  \hline	                         
-                186,460  & 19    &  13.7s      & 4.2s\\ %\hline
-                186,460  & 130  &  1m14.1s & 6.1s \\ %\hline
-                186,460  & 261  &  2m53.7s & 10.7s \\ \hline
+   \textbf{\# words} & \textbf{\# rules} &  \textbf{SATCG} & \textbf{VISL CG-3} \\ \hline
+               % 124,493  & 99    & 55.7s   & 4.8s \\  \hline	                         
+                384,155  & 19    &  13.7s  & 4.2s\\ %\hline
+                384,155  & 130  &  1m14.1s & 6.1s \\ %\hline
+                384,155  & 261  &  2m53.7s & 10.7s \\ \hline
   \end{tabular}
@@
 
-\paragraph{Performance against VISL-CG3}
+\paragraph{Performance against VISL CG-3}
 
@@
 We only took select and remove rules (no substitute/iff), after which the constraint grammar had 261 rules.
-With this setup, we took the text of the tagged corpus with all ambiguities, and ran both VISL-CG3 and our implementation with the same grammar. 
-Treating the chosen tagset as the gold standard, the disambiguation by SATCG achieves 75.09 \% correct result and VISL-CG3 75.80 \%.
+With this setup, we took the text of the tagged corpus with all ambiguities, and ran both VISL CG-3 and our implementation with the same grammar. 
+Treating the chosen tagset as the gold standard, the disambiguation by SATCG achieves 75.09 \% correct result and VISL CG-3 75.80 \%.
 
@@
 %      \hline
-%     \textbf{\# rules} & \textbf{SATCG} & \textbf{VISL-CG3} \\
+%     \textbf{\# rules} & \textbf{SATCG} & \textbf{VISL CG-3} \\
 %     \hline 
@@
 % 	16 &  83.197 & 83.030 \\ \hline  % 16 out of 17/18, no diff
-% 	15 &  83.188 & 83.022\\ \hline  % 15 out of 16/17. Tie with 14 for SATCG, chose to retain the rule that gave better results for VISL-CG3.
+% 	15 &  83.188 & 83.022\\ \hline  % 15 out of 16/17. Tie with 14 for SATCG, chose to retain the rule that gave better results for VISL CG-3.
 % 	14 &  83.180 & 83.005 \\ \hline  % Tie with 15 for SATCG.
@@
 % \end{table}
-% In this experiment, we took 19 rules written by the authors, which give 83.15 \% correct results with SATCG and 83.33 \% with VISL-CG3.
-% Then we removed rules one by one; after getting a score for $n$ rules, we took the combination of $n-1$ rules for which SATCG gives the best score, and we ran the rule set with VISL-CG3 for comparison.
+% In this experiment, we took 19 rules written by the authors, which give 83.15 \% correct results with SATCG and 83.33 \% with VISL CG-3.
+% Then we removed rules one by one; after getting a score for $n$ rules, we took the combination of $n-1$ rules for which SATCG gives the best score, and we ran the rule set with VISL CG-3 for comparison.
 Table~\ref{table:results_remove} shows the percentages compared to the gold standard for both systems.

### 2015-03-30 paper/nodalida2015.tex
more stuff

@@
 \section{Introduction}
-%Constraint Grammar (CG) was first introduced by \cite{karlsson1995constraint}. 
-%It disambiguates output by morphological analyser, 
-%and has been used for many tasks in computational linguistics, such as POS tagging,
-%surface syntax and machine translation.
-%The current state-of-the-art parser VISL CG-3 is implemented (...);
-%in addition there are a number of finite-state based implementations \cite{nemeskey14}.
 
@@
 biology---these problems easily match or exceed CG in complexity. 
-\todo{is this true for grammars with thousands of rules?}
+% Is this true for grammars with thousands of rules?
 
@@
 \paragraph{Rules are unordered.}
-%A typical CG rule sequence starts with rules that are either very common, or have a complex %condition to fulfil.
-%Stronger rules appear towards the end, to be applied if none of the weaker rules has. 
-%For a rule set such as following, this order is crucial:
-%\begin{itemize}
-%\item [] \texttt{SELECT Year IF (-1 Seasons OR Months) ;} \\
- %            \texttt{REMOVE Year ;}
-%\end{itemize}
 
@@
 \end{itemize}
-
-%combined with the following
-%\begin{itemize}
-%\item Maximise number of true analyses---Without this heuristic, both solutions to our example %\emph{la casa} are as likely.
-With this heuristic,  the example would be disambiguated as \texttt{prn v}, because there happens to be two analyses with \texttt{v}, and choosing that, 3 analyses are true, since there is no rule to disambiguate between imperative and indicative.
-% We want analyses to be true by default, but still disambiguate as much as possible.
-%We have a rule to maximise the number of true analyses; then we take a solution that has the least. 
-%\item Minimise number of true analyses---To combat the previous, we want each solution to contain as many true analyses as possible, to ensure that analyses untouched by rules are kept true.
-%After having multiple solutions which all ensure , we pick the one that disambiguates the most.
-%\end{itemize}
-
 The chosen heuristic produced the best results with examples in the order of tens or hundreds of rules.
-Larger rule sets (thousands of rules) are yet to be tested.
+Larger rule sets are yet to be tested.
 
@@
 \paragraph{Rules disambiguate more.}
-% Rules interact with each other and disambiguate more than same rules in the standard implementation. 
-In standard CG, the rules only tell to remove readings for the word which is tagged as 
-\texttt{prn} or \texttt{det}. The SAT-based implementation interprets the rules as ``noun and pronoun together are illegal'', and is free to take action that concerns also the following word.
-This means that the rules are logically flipped: compare the rules below to the ones in the example.
+Considering our example phrase and rules, standard CG implementation
+can only remove readings for the word which is tagged as \texttt{prn} or
+\texttt{det}. The SAT-based implementation interprets the rules as
+``determiner and verb together are illegal'', and is free to take action that concerns also the word in the condition \texttt{(1 n)} or \texttt{(1 v)}.
+This means that the rules are logically flipped: \texttt{REMOVE prn IF
+  (1 n)} translates into the same logical formula as  \texttt{REMOVE n
+  IF (-1 prn)}. A rule with more conditions can be written as many
+rules, each condition taking its turn to be in position to be removed
+or selected. \\
 
-\begin{itemize}
-\item [] \texttt{REMOVE n IF (-1 prn) ;} \\
-             \texttt{REMOVE v IF (-1 det) ;}
-\end{itemize}
+% \begin{itemize}
+% \item [] \texttt{REMOVE n IF (-1 prn) ;} \\
+%          \texttt{REMOVE v IF (-1 det) ;}
+% \end{itemize}
 
-SAT-based approach gives identical results with both sets of rules,
-whereas the standard CG would remove one reading from \emph{casa} and leave \emph{la} ambiguous.
-Testing this property with more complex rules and larger rule sets remains to be done.
+% SAT-based approach gives identical results with both sets of rules,
+% whereas the standard CG would remove one reading from \emph{casa} and leave \emph{la} ambiguous.
+% Testing this property with more complex rules and larger rule sets remains to be done.
 
@@
 
-These three features are likely to change the way that rules are written. \todo{more speculation, examples}
+\noindent These three features might change the way that rules are written. 
+We predict that less rules are needed; whether this holds in
+the order of thousands of rules remains to be tested.
+Getting rid of ordering could ease the task of the grammar writer,
+removing the responsibility to estimate the best sequence of rule applications.
+However, the implementation should not be too different from the
+existing ones---our preliminary evaluation against VISL CG-3 shows
+promising results in this regard.
 
@@
 
-The worst-case complexity of SAT is exponential, whereas the standard implementations 
-of CG are polynomial, but with advances in SAT solving techniques, the performance 
-in the average case is much more feasible than in the previous works done in 90s--00s.
+The worst-case complexity of SAT is exponential, whereas the standard
+implementations of CG are polynomial, but with advances in SAT solving
+techniques, the performance in the average case is much more feasible than in the previous works done in 90s--00s.
 We used the open-source SAT solver MiniSat \cite{een04sat}.
-%We do foreach sentence: foreach rule; 
-%whereas VISL CG-3 has foreach rule: foreach sentence.
-%the size of the SAT problem is dependent of the number of rules
 
 Table~\ref{table:time} shows the execution time compared to VISL CG-3. 
-The number of rules in the grammar is 
-Maximisation is the most costly operation
-Also, some of the time is explained simply by
-SATCG is just a first proof-of-concept implementation and can be optimised.
+The number of rules in the grammar is more significant than the word
+count; this holds also for the performance of VISL CG-3.
+From the SAT solving side, maximisation is the most costly operation.
+Parsing Don Quijote (384,155 words) with the rule set of 261 rules,
+the function was called 147,253 times, and with 19 rules, 132,255 times.
+%However, with both rule sets, half of the sentences in Don Quijote
+%needed less than 6 calls of maximise, and 75 \% of the sentences
+%needed less than 14.
+The difference in the execution times suggests that there are other
+reasons for the worse performance---this is to be expected,
+SATCG is just a first proof-of-concept implementation with no optimisations.
 
@@
    \textbf{\# words} & \textbf{\# rules} &  \textbf{SATCG} & \textbf{VISL CG-3} \\ \hline
-               % 124,493  & 99    & 55.7s   & 4.8s \\  \hline	                         
                 384,155  & 19   &  13.7s   & 4.2s\\ %\hline
+                124,493  & 99   &  55.7s   & 4.8s \\ % \hline
 	        384,155  & 130  &  1m14.1s & 6.1s \\ %\hline
@@
 With this setup, we took the text of the tagged corpus with all ambiguities, and ran both VISL CG-3 and our implementation with the same grammar. 
-Treating the chosen tagset as the gold standard, the disambiguation by SATCG achieves 75.09 \% correct result and VISL CG-3 75.80 \%.
+Treating the chosen tagset as the gold standard, the disambiguation by
+SATCG achieves 75.09 \% correct result and VISL CG-3 75.80 \%.
+Similar patterns were observed with small (\textless{}20) rule sets written by
+the authors; depending on the subset, SATCG and VISL CG-3 had
+a difference of at most 1.5 \%. Introducing rules one by one up to 19, the
+performance improved in a very similar rate, with less than 0.5 \%
+difference between implementations at each new rule.
+
+Additional tests could include plugging SATCG into Apertium
+translation pipeline, and comparing the translation quality.
 
@@
 % when submitting your paper for review.
-% Francis Tyers
 

### 2015-03-30 paper/nodalida2015.tex
more stuff

@@
 \usepackage{tikz}
-\def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;}
+\def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) --
+  (.25,.15) -- cycle;}
+\newcommand{\Mypm}{\mathbin{\tikz [x=1.4ex,y=1.4ex,line width=.1ex] \draw (0.0,0) -- (1.0,0) (0.5,0.08) -- (0.5,0.92) (0.0,0.5) -- (1.0,0.5);}}%
+
 \title{Constraint Grammar as a SAT problem}
@@
 promising results in this regard.
+\pagebreak
 
@@
 the authors; depending on the subset, SATCG and VISL CG-3 had
-a difference of at most 1.5 \%. Introducing rules one by one up to 19, the
+a difference of at most $\Mypm$ 1.5 \%. 
+Introducing rules one by one up to 19, the
 performance improved in a very similar rate, with less than 0.5 \%
-difference between implementations at each new rule.
+difference between the systems at each new rule.
 
-Additional tests could include plugging SATCG into Apertium
-translation pipeline, and comparing the translation quality.
+% Additional tests could include plugging SATCG into Apertium
+% translation pipeline, and comparing the translation quality.
 

### 2015-03-30 paper/nodalida2015.tex
polish intro

@@
 as a Boolean satisfiability (SAT) problem.
+This is attractive from several reasons: formal logic is
+well-studied, and serves as an abstract language to reason about the
+properties of CG. Constraint rules encoded in logic capture richer
+dependencies between the tags than standard CG.
 
-Formal logic is well-studied and has an old tradition.
-The rules encoded in logic capture richer dependencies between the
-tags than normal reductionist grammars.
-
-Applying logic to reductionist grammars has been explored earlier by 
-\cite{lager98,lager_nivre01}, but it was never adopted for use.
+Applying logic to reductionist grammars has been explored earlier by \cite{lager98,lager_nivre01}, but it was never adopted for use.
 % ; logic programming was too slow to be used for tagging or parsing. 

### 2015-03-30 paper/nodalida2015.tex
small changes here and there

@@
 
-The ambiguous passage can be either a noun phrase, \emph{la\texttt{:det} casa\texttt{:n}} 
-`the house'  or a verb phrase \emph{la\texttt{:prn} casa\texttt{:v pri p3}} `(he/she) marries her'. 
+The ambiguous passage can be either a noun phrase, \emph{la}\texttt{<det>} \emph{casa}\texttt{<n>} 
+`the house'  or a verb phrase \texttt{la<prn>}  \emph{casa}\texttt{<v><pri><p3>} `(he/she) marries her'. 
 We add the following rules:
@@
 the CG principles: an analysis is true if no rule affects it,
-and at least one analysis for each word is true---the notion of ``last'' is not applicable.
-With these constraints, we get two solutions: 
+and at least one analysis for each word is true---the notion of
+``last'' is not applicable.
+
+With these constraints, we get two solutions. The interaction of the rules regarding \emph{la}  disambiguates the POS of \emph{casa} for free, and ordering of the rules doesn't matter.
+
 \begin{enumerate}
-\item []
+\item [\texttt{1)}]
 \begin{verbatim}
@@
 \end{verbatim}
-\item [and]
+\item [\texttt{2)}]
 \begin{verbatim}
@@
 \end{enumerate}
-% \texttt{det, n} and \texttt{prn, v pri p3, v imp p2}. 
-The interaction of the rules regarding \emph{la}  disambiguates the POS of \emph{casa} 
-for free, and ordering of the rules doesn't matter. 
+
 %Without more context or additional rules we cannot fully disambiguate the passage,
@@
 
-The most important differences between the traditional and the SAT-based approach are summarised below:
+\noindent The most important differences between the traditional and the SAT-based approach are summarised below:
 
@@
 \texttt{det}. The SAT-based implementation interprets the rules as
-``determiner and verb together are illegal'', and is free to take action that concerns also the word in the condition \texttt{(1 n)} or \texttt{(1 v)}.
+``determiner and verb together are illegal'', and is free to take action that concerns also the word in the condition (\texttt{1 n} or \texttt{1 v}).
 This means that the rules are logically flipped: \texttt{REMOVE prn IF
@@
 Getting rid of ordering could ease the task of the grammar writer,
-removing the responsibility to estimate the best sequence of rule applications.
+since it removes the need to estimate the best sequence of rule
+applications.
+
 However, the implementation should not be too different from the
@@
 reasons for the worse performance---this is to be expected,
-SATCG is just a first proof-of-concept implementation with no optimisations.
+SATCG is currently just a naive proof-of-concept implementation with no optimisations.
 
@@
 We only took select and remove rules (no substitute/iff), after which the constraint grammar had 261 rules.
-With this setup, we took the text of the tagged corpus with all ambiguities, and ran both VISL CG-3 and our implementation with the same grammar. 
-Treating the chosen tagset as the gold standard, the disambiguation by
+With this setup, we took the text of the tagged corpus with all
+ambiguities, and ran both SATCG and VISL CG-3 with the same grammar. 
+Treating the corpus as the gold standard, the disambiguation by
 SATCG achieves 75.09 \% correct result and VISL CG-3 75.80 \%.
-Similar patterns were observed with small (\textless{}20) rule sets written by
-the authors; depending on the subset, SATCG and VISL CG-3 had
+Similar patterns were observed with small (\textless{}20) rule sets
+written by the authors; depending on the subset, SATCG and VISL CG-3 had
 a difference of at most $\Mypm$ 1.5 \%. 

### 2015-03-30 paper/satcg.tex
change name

--- 2015-03-30 paper/nodalida2015.tex
@@
 The ambiguous passage can be either a noun phrase, \emph{la}\texttt{<det>} \emph{casa}\texttt{<n>} 
-`the house'  or a verb phrase \texttt{la<prn>}  \emph{casa}\texttt{<v><pri><p3>} `(he/she) marries her'. 
+`the house'  or a verb phrase \emph{la}\texttt{<prn>}  \emph{casa}\texttt{<v><pri><p3>} `(he/she) marries her'. 
 We add the following rules:

### 2015-03-31 paper/nodalida2015.tex
asdf

--- 2015-03-30 paper/nodalida2015.tex
@@
-%
-% File nodalida2015.tex
-%
-% Contact beata.megyesi@lingfil.uu.se
-%
-% Based on the instruction file for EACL 2014
-% which in turn was based on the instruction files for previous 
-% ACL and EACL conferences.
+<!DOCTYPE html>
+<html lang="en" class="">
+  <head prefix="og: http://ogp.me/ns# fb: http://ogp.me/ns/fb# object: http://ogp.me/ns/object# article: http://ogp.me/ns/article# profile: http://ogp.me/ns/profile#">
+    <meta charset='utf-8'>
+    <meta http-equiv="X-UA-Compatible" content="IE=edge">
+    <meta http-equiv="Content-Language" content="en">
 
-\documentclass[11pt]{article}
-\usepackage{nodalida2015}
-\usepackage{times}
-\usepackage{mathptmx}
-%\usepackage{txfonts}
-\usepackage{url}
-\usepackage{latexsym}
-\usepackage{cite}
-\usepackage{authordate1-4}
-\usepackage{multirow}
-\special{papersize=210mm,297mm} % to avoid having to use "-t a4" with dvips 
-%\setlength\titlebox{6.5cm}  % You can expand the title box if you really have to
+    <title>cgsat/nodalida2015.tex at master · inariksit/cgsat · GitHub</title>
+    <link rel="search" type="application/opensearchdescription+xml" href="/opensearch.xml" title="GitHub">
+    <link rel="fluid-icon" href="https://github.com/fluidicon.png" title="GitHub">
+    <link rel="apple-touch-icon" sizes="57x57" href="/apple-touch-icon-114.png">
+    <link rel="apple-touch-icon" sizes="114x114" href="/apple-touch-icon-114.png">
+    <link rel="apple-touch-icon" sizes="72x72" href="/apple-touch-icon-144.png">
+    <link rel="apple-touch-icon" sizes="144x144" href="/apple-touch-icon-144.png">
+    <meta property="fb:app_id" content="1401488693436528">
 
-\usepackage{color}
-\newcommand{\todo}[1]{{\color{cyan}\textbf{[TODO: }#1\textbf{]}}}
-\usepackage{tikz}
-\def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) --
-  (.25,.15) -- cycle;}
-\newcommand{\Mypm}{\mathbin{\tikz [x=1.4ex,y=1.4ex,line width=.1ex] \draw (0.0,0) -- (1.0,0) (0.5,0.08) -- (0.5,0.92) (0.0,0.5) -- (1.0,0.5);}}%
+      <meta content="@github" name="twitter:site" /><meta content="summary" name="twitter:card" /><meta content="inariksit/cgsat" name="twitter:title" /><meta content="Contribute to cgsat development by creating an account on GitHub." name="twitter:description" /><meta content="https://avatars1.githubusercontent.com/u/1519665?v=3&amp;s=400" name="twitter:image:src" />
+      <meta content="GitHub" property="og:site_name" /><meta content="object" property="og:type" /><meta content="https://avatars1.githubusercontent.com/u/1519665?v=3&amp;s=400" property="og:image" /><meta content="inariksit/cgsat" property="og:title" /><meta content="https://github.com/inariksit/cgsat" property="og:url" /><meta content="Contribute to cgsat development by creating an account on GitHub." property="og:description" />
+      <meta name="browser-stats-url" content="/_stats">
+    <link rel="assets" href="https://assets-cdn.github.com/">
 
-\title{Constraint Grammar as a SAT problem}
+    <meta name="pjax-timeout" content="1000">
 
-% \author{Inari Listenmaa \and Koen Claessen \\
-%  Chalmers University of Technology, Gothenburg, Sweden \\
-%  {\tt \{inari,koen\}@chalmers.se} }
+    <meta name="msapplication-TileImage" content="/windows-tile.png">
+    <meta name="msapplication-TileColor" content="#ffffff">
+    <meta name="selected-link" value="repo_source" data-pjax-transient>
+      <meta name="google-analytics" content="UA-3769691-2">
 
-\date{\today}
+    <meta content="collector.githubapp.com" name="octolytics-host" /><meta content="collector-cdn.github.com" name="octolytics-script-host" /><meta content="github" name="octolytics-app-id" /><meta content="8110BB79:63FC:112422E:551A5EAC" name="octolytics-dimension-request_id" />
 
-\begin{document}
-\maketitle
+    <meta content="Rails, view, blob#show" name="analytics-event" />
+    <meta class="js-ga-set" name="dimension1" content="Logged Out">
+    <meta class="js-ga-set" name="dimension2" content="Header v3">
 
-%\begin{abstract} \end{abstract}
+    <link rel="icon" type="image/x-icon" href="https://assets-cdn.github.com/favicon.ico">
 
-\section{Introduction}
+    <meta content="authenticity_token" name="csrf-param" />
+<meta content="3xCYrA3acwWrQrZbkYV0AsS+NdsxtQ4QV0yxTkc3KiLfZTRKlUSdR7DLeyh7sjj/5THBcwXU6KhwedJfUtW/lg==" name="csrf-token" />
 
-We represent Constraint Grammar \cite{karlsson1995constraint} 
-as a Boolean satisfiability (SAT) problem.
+    <link href="https://assets-cdn.github.com/assets/github-391b32dc1132ef0837cee6b206cad9c2726c7f495e5047901a7afa551e81d695.css" media="all" rel="stylesheet" />
+    <link href="https://assets-cdn.github.com/assets/github2-14b629a12b237ffde72fb70f31de6a515754f0f686189a072cc00e54ee202fd1.css" media="all" rel="stylesheet" />
 
-Formal logic is well-studied and has an old tradition.
-The rules encoded in logic capture richer dependencies between the
-tags than normal reductionist grammars.
+    <meta http-equiv="x-pjax-version" content="0cdf73d6f5d84ba34af332a36c975c0a">
 
-Applying logic to reductionist grammars has been explored earlier by 
-\cite{lager98,lager_nivre01}, but it was never adopted for use.
-% ; logic programming was too slow to be used for tagging or parsing. 
-Since those works, SAT solving techniques have improved significantly, and 
-they are used in domains such as microprocessor design and computational 
-biology---these problems easily match or exceed CG in complexity. 
-% Is this true for grammars with thousands of rules?
+  <meta name="description" content="Contribute to cgsat development by creating an account on GitHub.">
+  <meta name="go-import" content="github.com/inariksit/cgsat git https://github.com/inariksit/cgsat.git">
 
-Due to these advances, we were able to revisit the idea and develop it further. 
-% Whereas \cite{lager98} provides a proof-of-concept description, 
-Missing from the previous logic-based works, 
-we solve eventual rule conflicts by finding a solution that 
-discards the least number of rules.
-% allows the maximal number of rules to be applied.
-We test our implementation by parsing texts in the order of 10,000s--100,000s
-words, using grammars with hundreds of rules.
+  <meta content="1519665" name="octolytics-dimension-user_id" /><meta content="inariksit" name="octolytics-dimension-user_login" /><meta content="28176249" name="octolytics-dimension-repository_id" /><meta content="inariksit/cgsat" name="octolytics-dimension-repository_nwo" /><meta content="true" name="octolytics-dimension-repository_public" /><meta content="false" name="octolytics-dimension-repository_is_fork" /><meta content="28176249" name="octolytics-dimension-repository_network_root_id" /><meta content="inariksit/cgsat" name="octolytics-dimension-repository_network_root_nwo" />
+  <link href="https://github.com/inariksit/cgsat/commits/master.atom" rel="alternate" title="Recent Commits to cgsat:master" type="application/atom+xml">
 
-%SAT solving is based on a technique called unit propagation:
-%a set of complex Boolean clauses is simplified, starting from unit
-%clauses, which consist of just a single variable, and working up to a
-%solution, where all variables in the set have a True or False value.
+  </head>
 
-%This increases the computational complexity, 
-%but we are exploring ways in which it could simplify the grammar writing:
-%our experiments show that less rules are needed, and features like ordering of the rules
-%and requirement of being unambiguously tagged become irrelevant.
+  <body class="logged_out  env-production  vis-public page-blob">
+    <a href="#start-of-content" tabindex="1" class="accessibility-aid js-skip-to-content">Skip to content</a>
+    <div class="wrapper">
 
-% Uncomment for full paper
-% \section{Related work}
-% \label{sect:related}
+        <div class="header header-logged-out" role="banner">
+  <div class="container clearfix">
 
-% Our work is inspired by \cite{lager98}, which presents constraint rules as a disjunctive logic program,
-% and \cite{lager_nivre01}, which reconstructs 4 different formalisms from a logical point of view.
-% \cite{lindberg_eineborg98ilp,asfrent14} use Inductive Logic Programming to learn CG rules from a tagged corpus.
-% \cite{lager01transformation} discusses ordering of the CG rules.
+    <a class="header-logo-wordmark" href="https://github.com/" data-ga-click="(Logged out) Header, go to homepage, icon:logo-wordmark">
+      <span class="mega-octicon octicon-logo-github"></span>
+    </a>
 
-\section{CG as a SAT problem}
-Let us demonstrate our approach with the following example in Spanish.
+    <div class="header-actions" role="navigation">
+        <a class="btn btn-primary" href="/join" data-ga-click="(Logged out) Header, clicked Sign up, text:sign-up">Sign up</a>
+      <a class="btn" href="/login?return_to=%2Finariksit%2Fcgsat%2Fblob%2Fmaster%2Fpaper%2Fnodalida2015.tex" data-ga-click="(Logged out) Header, clicked Sign in, text:sign-in">Sign in</a>
+    </div>
 
-\begin{verbatim}
-"<la>"
-        "el" det def f sg
-        "lo" prn p3 f sg
-"<casa>"
-        "casa" n f sg
-        "casar" v pri p3 sg
-        "casar" v imp p2 sg
-\end{verbatim}
+    <div class="site-search repo-scope js-site-search" role="search">
+      <form accept-charset="UTF-8" action="/inariksit/cgsat/search" class="js-site-search-form" data-global-search-url="/search" data-repo-search-url="/inariksit/cgsat/search" method="get"><div style="margin:0;padding:0;display:inline"><input name="utf8" type="hidden" value="&#x2713;" /></div>
+  <input type="text"
+    class="js-site-search-field is-clearable"
+    data-hotkey="s"
+    name="q"
+    placeholder="Search"
+    data-global-scope-placeholder="Search GitHub"
+    data-repo-scope-placeholder="Search"
+    tabindex="1"
+    autocapitalize="off">
+  <div class="scope-badge">This repository</div>
+</form>
+    </div>
 
-The ambiguous passage can be either a noun phrase, \emph{la\texttt{:det} casa\texttt{:n}} 
-`the house'  or a verb phrase \emph{la\texttt{:prn} casa\texttt{:v pri p3}} `(he/she) marries her'. 
-We add the following rules:
+      <ul class="header-nav left" role="navigation">
+          <li class="header-nav-item">
+            <a class="header-nav-link" href="/explore" data-ga-click="(Logged out) Header, go to explore, text:explore">Explore</a>
+          </li>
+          <li class="header-nav-item">
+            <a class="header-nav-link" href="/features" data-ga-click="(Logged out) Header, go to features, text:features">Features</a>
+          </li>
+          <li class="header-nav-item">
+            <a class="header-nav-link" href="https://enterprise.github.com/" data-ga-click="(Logged out) Header, go to enterprise, text:enterprise">Enterprise</a>
+          </li>
+          <li class="header-nav-item">
+            <a class="header-nav-link" href="/blog" data-ga-click="(Logged out) Header, go to blog, text:blog">Blog</a>
+          </li>
+      </ul>
 
-\begin{itemize}
-\item [] \texttt{REMOVE prn IF (1 n) ;} \\
-             \texttt{REMOVE det IF (1 v) ;}
-\end{itemize}
+  </div>
+</div>
 
-Standard CG will apply one\footnote{To be more cautious, we could require the word at position +1 be disambiguated fully (\texttt{1C} instead of \texttt{1}), but in that case, 
-neither of the rules would be applied.} of the rules to the word \emph{la}; 
-either the one that comes first, or by some other heuristic. 
-The second rule will not fire, because it would remove the last reading. 
-All readings of \emph{casa} are left untouched by these rules.
+      <div id="start-of-content" class="accessibility-aid"></div>
+          <div class="site" itemscope itemtype="http://schema.org/WebPage">
+    <div id="js-flash-container">
 
-SAT solver performs a search, 
-and starts building possible models that satisfy both constraints. 
-In addition to the given constraints, we have default rules to emulate
-the CG principles: an analysis is true if no rule affects it,
-and at least one analysis for each word is true---the notion of ``last'' is not applicable.
-With these constraints, we get two solutions: 
-\begin{enumerate}
-\item []
-\begin{verbatim}
-"<la>"
-        "el" det def f sg
-"<casa>"
-        "casa" n f sg
-\end{verbatim}
-\item [and]
-\begin{verbatim}
-"<la>"
-        "lo" prn p3 f sg
-"<casa>"
-        "casar" v pri p3 sg
-        "casar" v imp p2 sg
-\end{verbatim}
-\end{enumerate} 
-% \texttt{det, n} and \texttt{prn, v pri p3, v imp p2}. 
-The interaction of the rules regarding \emph{la}  disambiguates the POS of \emph{casa} 
-for free, and ordering of the rules doesn't matter. 
-%Without more context or additional rules we cannot fully disambiguate the passage,
-%but unlike with standard CG, we get only legal combinations in one solution.
+    </div>
+    <div class="pagehead repohead instapaper_ignore readability-menu">
+      <div class="container">
 
-%If we add a word, \emph{la casa grande} `the big house' 
-%and a rule that removes verb if followed by an adjective
-%\texttt{REMOVE (v) IF (1 (adj)) ;}
-%we will get a unique solution with \texttt{det n adj}.
-%Standard CG will also remove the verb reading, 
-%but the choice of the word \emph{la} doesn't contribute to
+<ul class="pagehead-actions">
 
-The most important differences between the traditional and the SAT-based approach are summarised below:
+  <li>
+      <a href="/login?return_to=%2Finariksit%2Fcgsat"
+    class="btn btn-sm btn-with-count tooltipped tooltipped-n"
+    aria-label="You must be signed in to watch a repository" rel="nofollow">
+    <span class="octicon octicon-eye"></span>
+    Watch
+  </a>
+  <a class="social-count" href="/inariksit/cgsat/watchers">
+    3
+  </a>
 
-\paragraph{Condition of being unambiguously tagged is irrelevant.}
-Rather than waiting for a word to get disambiguated, the SAT solver starts by 
-making assumptions (e.g. ``\emph{casa} is a noun'') and working under them,
-discarding the assumption if it doesn't lead to a model that satisfies all constraints.
+  </li>
 
-\paragraph{Rules are unordered.}
+  <li>
+      <a href="/login?return_to=%2Finariksit%2Fcgsat"
+    class="btn btn-sm btn-with-count tooltipped tooltipped-n"
+    aria-label="You must be signed in to star a repository" rel="nofollow">
+    <span class="octicon octicon-star"></span>
+    Star
+  </a>
 
-We have experimented with different strategies:
-\begin{itemize}
-\item Maximise
-% \footnote{Something about local maximum}
-%The algorithm commits to a subset of clauses (e.g.
+    <a class="social-count js-social-count" href="/inariksit/cgsat/stargazers">
+      0
+    </a>
+
+  </li>
+
+    <li>
+      <a href="/login?return_to=%2Finariksit%2Fcgsat"
+        class="btn btn-sm btn-with-count tooltipped tooltipped-n"
+        aria-label="You must be signed in to fork a repository" rel="nofollow">
+        <span class="octicon octicon-repo-forked"></span>
+        Fork
+      </a>
+      <a href="/inariksit/cgsat/network" class="social-count">
+        0
+      </a>
+    </li>
+</ul>
+
+        <h1 itemscope itemtype="http://data-vocabulary.org/Breadcrumb" class="entry-title public">
+          <span class="mega-octicon octicon-repo"></span>
+          <span class="author"><a href="/inariksit" class="url fn" itemprop="url" rel="author"><span itemprop="title">inariksit</span></a></span><!--
+       --><span class="path-divider">/</span><!--
+       --><strong><a href="/inariksit/cgsat" class="js-current-repository" data-pjax="#js-repo-pjax-container">cgsat</a></strong>
+
+          <span class="page-context-loader">
+            <img alt="" height="16" src="https://assets-cdn.github.com/assets/spinners/octocat-spinner-32-e513294efa576953719e4e2de888dd9cf929b7d62ed8d05f25e731d02452ab6c.gif" width="16" />
+          </span>
+
+        </h1>
+      </div><!-- /.container -->
+    </div><!-- /.repohead -->
+
+    <div class="container">
+      <div class="repository-with-sidebar repo-container new-discussion-timeline  ">
+        <div class="repository-sidebar clearfix">
+
+<nav class="sunken-menu repo-nav js-repo-nav js-sidenav-container-pjax js-octicon-loaders"
+     role="navigation"
+     data-pjax="#js-repo-pjax-container"
+     data-issue-count-url="/inariksit/cgsat/issues/counts">
+  <ul class="sunken-menu-group">
+    <li class="tooltipped tooltipped-w" aria-label="Code">
+      <a href="/inariksit/cgsat" aria-label="Code" class="selected js-selected-navigation-item sunken-menu-item" data-hotkey="g c" data-selected-links="repo_source repo_downloads repo_commits repo_releases repo_tags repo_branches /inariksit/cgsat">
+        <span class="octicon octicon-code"></span> <span class="full-word">Code</span>
+        <img alt="" class="mini-loader" height="16" src="https://assets-cdn.github.com/assets/spinners/octocat-spinner-32-e513294efa576953719e4e2de888dd9cf929b7d62ed8d05f25e731d02452ab6c.gif" width="16" />
+</a>    </li>
+
+      <li class="tooltipped tooltipped-w" aria-label="Issues">
+        <a href="/inariksit/cgsat/issues" aria-label="Issues" class="js-selected-navigation-item sunken-menu-item" data-hotkey="g i" data-selected-links="repo_issues repo_labels repo_milestones /inariksit/cgsat/issues">
+          <span class="octicon octicon-issue-opened"></span> <span class="full-word">Issues</span>
+          <span class="js-issue-replace-counter"></span>
+          <img alt="" class="mini-loader" height="16" src="https://assets-cdn.github.com/assets/spinners/octocat-spinner-32-e513294efa576953719e4e2de888dd9cf929b7d62ed8d05f25e731d02452ab6c.gif" width="16" />
+</a>      </li>
+
+    <li class="tooltipped tooltipped-w" aria-label="Pull requests">
+      <a href="/inariksit/cgsat/pulls" aria-label="Pull requests" class="js-selected-navigation-item sunken-menu-item" data-hotkey="g p" data-selected-links="repo_pulls /inariksit/cgsat/pulls">
+          <span class="octicon octicon-git-pull-request"></span> <span class="full-word">Pull requests</span>
+          <span class="js-pull-replace-counter"></span>
+          <img alt="" class="mini-loader" height="16" src="https://assets-cdn.github.com/assets/spinners/octocat-spinner-32-e513294efa576953719e4e2de888dd9cf929b7d62ed8d05f25e731d02452ab6c.gif" width="16" />
+</a>    </li>
+
+  </ul>
+  <div class="sunken-menu-separator"></div>
+  <ul class="sunken-menu-group">
+
+    <li class="tooltipped tooltipped-w" aria-label="Pulse">
+      <a href="/inariksit/cgsat/pulse" aria-label="Pulse" class="js-selected-navigation-item sunken-menu-item" data-selected-links="pulse /inariksit/cgsat/pulse">
+        <span class="octicon octicon-pulse"></span> <span class="full-word">Pulse</span>
+        <img alt="" class="mini-loader" height="16" src="https://assets-cdn.github.com/assets/spinners/octocat-spinner-32-e513294efa576953719e4e2de888dd9cf929b7d62ed8d05f25e731d02452ab6c.gif" width="16" />
+</a>    </li>
+
+    <li class="tooltipped tooltipped-w" aria-label="Graphs">
+      <a href="/inariksit/cgsat/graphs" aria-label="Graphs" class="js-selected-navigation-item sunken-menu-item" data-selected-links="repo_graphs repo_contributors /inariksit/cgsat/graphs">
+        <span class="octicon octicon-graph"></span> <span class="full-word">Graphs</span>
+        <img alt="" class="mini-loader" height="16" src="https://assets-cdn.github.com/assets/spinners/octocat-spinner-32-e513294efa576953719e4e2de888dd9cf929b7d62ed8d05f25e731d02452ab6c.gif" width="16" />
+</a>    </li>
+  </ul>
+
+</nav>
+
+              <div class="only-with-full-nav">
+
+<div class="clone-url open"
+  data-protocol-type="http"
+  data-url="/users/set_protocol?protocol_selector=http&amp;protocol_type=clone">
+  <h3><span class="text-emphasized">HTTPS</span> clone URL</h3>
+  <div class="input-group js-zeroclipboard-container">
+    <input type="text" class="input-mini input-monospace js-url-field js-zeroclipboard-target"
+           value="https://github.com/inariksit/cgsat.git" readonly="readonly">
+    <span class="input-group-button">
+      <button aria-label="Copy to clipboard" class="js-zeroclipboard btn btn-sm zeroclipboard-button" data-copied-hint="Copied!" type="button"><span class="octicon octicon-clippy"></span></button>
+    </span>
+  </div>
+</div>
+
+<div class="clone-url "
+  data-protocol-type="subversion"
+  data-url="/users/set_protocol?protocol_selector=subversion&amp;protocol_type=clone">
+  <h3><span class="text-emphasized">Subversion</span> checkout URL</h3>
+  <div class="input-group js-zeroclipboard-container">
+    <input type="text" class="input-mini input-monospace js-url-field js-zeroclipboard-target"
+           value="https://github.com/inariksit/cgsat" readonly="readonly">
+    <span class="input-group-button">
+      <button aria-label="Copy to clipboard" class="js-zeroclipboard btn btn-sm zeroclipboard-button" data-copied-hint="Copied!" type="button"><span class="octicon octicon-clippy"></span></button>
+    </span>
+  </div>
+</div>
+
+<p class="clone-options">You can clone with
+  <a href="#" class="js-clone-selector" data-protocol="http">HTTPS</a> or <a href="#" class="js-clone-selector" data-protocol="subversion">Subversion</a>.
+  <a href="https://help.github.com/articles/which-remote-url-should-i-use" class="help tooltipped tooltipped-n" aria-label="Get help on which URL is right for you.">
+    <span class="octicon octicon-question"></span>
+  </a>
+</p>
+
+                <a href="/inariksit/cgsat/archive/master.zip"
+                   class="btn btn-sm sidebar-button"
+                   aria-label="Download the contents of inariksit/cgsat as a zip file"
+                   title="Download the contents of inariksit/cgsat as a zip file"
+                   rel="nofollow">
+                  <span class="octicon octicon-cloud-download"></span>
+                  Download ZIP
+                </a>
+              </div>
+        </div><!-- /.repository-sidebar -->
+
+        <div id="js-repo-pjax-container" class="repository-content context-loader-container" data-pjax-container>
+
+<a href="/inariksit/cgsat/blob/b380387015c0c6f22e8f00ed3d7f272dcfd7aa3a/paper/nodalida2015.tex" class="hidden js-permalink-shortcut" data-hotkey="y">Permalink</a>
+
+<!-- blob contrib key: blob_contributors:v21:ba5b322387c6f6d683ca66ec1649ec78 -->
+
+<div class="file-navigation js-zeroclipboard-container">
+
+<div class="select-menu js-menu-container js-select-menu left">
+  <span class="btn btn-sm select-menu-button js-menu-target css-truncate" data-hotkey="w"
+    data-master-branch="master"
+    data-ref="master"
+    title="master"
+    role="button" aria-label="Switch branches or tags" tabindex="0" aria-haspopup="true">
+    <span class="octicon octicon-git-branch"></span>
+    <i>branch:</i>
+    <span class="js-select-button css-truncate-target">master</span>
+  </span>
+
+  <div class="select-menu-modal-holder js-menu-content js-navigation-container" data-pjax aria-hidden="true">
+
+    <div class="select-menu-modal">
+      <div class="select-menu-header">
+        <span class="select-menu-title">Switch branches/tags</span>
+        <span class="octicon octicon-x js-menu-close" role="button" aria-label="Close"></span>
+      </div>
+
+      <div class="select-menu-filters">
+        <div class="select-menu-text-filter">
+          <input type="text" aria-label="Filter branches/tags" id="context-commitish-filter-field" class="js-filterable-field js-navigation-enable" placeholder="Filter branches/tags">
+        </div>
+        <div class="select-menu-tabs">
+          <ul>
+            <li class="select-menu-tab">
+              <a href="#" data-tab-filter="branches" data-filter-placeholder="Filter branches/tags" class="js-select-menu-tab">Branches</a>
+            </li>
+            <li class="select-menu-tab">
+              <a href="#" data-tab-filter="tags" data-filter-placeholder="Find a tag…" class="js-select-menu-tab">Tags</a>
+            </li>
+          </ul>
+        </div>
+      </div>
+
+      <div class="select-menu-list select-menu-tab-bucket js-select-menu-tab-bucket" data-tab-filter="branches">
+
+        <div data-filterable-for="context-commitish-filter-field" data-filterable-type="substring">
+
+            <a class="select-menu-item js-navigation-item js-navigation-open selected"
+               href="/inariksit/cgsat/blob/master/paper/nodalida2015.tex"
+               data-name="master"
+               data-skip-pjax="true"
+               rel="nofollow">
+              <span class="select-menu-item-icon octicon octicon-check"></span>
+              <span class="select-menu-item-text css-truncate-target" title="master">
+                master
+              </span>
+            </a>
+        </div>
+
+          <div class="select-menu-no-results">Nothing to show</div>
+      </div>
+
+      <div class="select-menu-list select-menu-tab-bucket js-select-menu-tab-bucket" data-tab-filter="tags">
+        <div data-filterable-for="context-commitish-filter-field" data-filterable-type="substring">
+
+        </div>
+
+        <div class="select-menu-no-results">Nothing to show</div>
+      </div>
+
+    </div>
+  </div>
+</div>
+
+  <div class="btn-group right">
+    <a href="/inariksit/cgsat/find/master"
+          class="js-show-file-finder btn btn-sm empty-icon tooltipped tooltipped-s"
+          data-pjax
+          data-hotkey="t"
+          aria-label="Quickly jump between files">
+      <span class="octicon octicon-list-unordered"></span>
+    </a>
+    <button aria-label="Copy file path to clipboard" class="js-zeroclipboard btn btn-sm zeroclipboard-button" data-copied-hint="Copied!" type="button"><span class="octicon octicon-clippy"></span></button>
+  </div>
+
+  <div class="breadcrumb js-zeroclipboard-target">
+    <span class='repo-root js-repo-root'><span itemscope="" itemtype="http://data-vocabulary.org/Breadcrumb"><a href="/inariksit/cgsat" class="" data-branch="master" data-direction="back" data-pjax="true" itemscope="url"><span itemprop="title">cgsat</span></a></span></span><span class="separator">/</span><span itemscope="" itemtype="http://data-vocabulary.org/Breadcrumb"><a href="/inariksit/cgsat/tree/master/paper" class="" data-branch="master" data-direction="back" data-pjax="true" itemscope="url"><span itemprop="title">paper</span></a></span><span class="separator">/</span><strong class="final-path">nodalida2015.tex</strong>
+  </div>
+</div>
+
+<include-fragment class="commit commit-loader file-history-tease" src="/inariksit/cgsat/contributors/master/paper/nodalida2015.tex">
+  <div class="file-history-tease-header">
+    Fetching contributors&hellip;
+  </div>
+
+  <div class="participation">
+    <p class="loader-loading"><img alt="" height="16" src="https://assets-cdn.github.com/assets/spinners/octocat-spinner-32-EAF2F5-0bdc57d34b85c4a4de9d0d1db10cd70e8a95f33ff4f46c5a8c48b4bf4e5a9abe.gif" width="16" /></p>
+    <p class="loader-error">Cannot retrieve contributors at this time</p>
+  </div>
+</include-fragment>
+<div class="file">
+  <div class="file-header">
+    <div class="file-actions">
+
+      <div class="btn-group">
+        <a href="/inariksit/cgsat/raw/master/paper/nodalida2015.tex" class="btn btn-sm " id="raw-url">Raw</a>
+          <a href="/inariksit/cgsat/blame/master/paper/nodalida2015.tex" class="btn btn-sm js-update-url-with-hash">Blame</a>
+        <a href="/inariksit/cgsat/commits/master/paper/nodalida2015.tex" class="btn btn-sm " rel="nofollow">History</a>
+      </div>
+
+          <button type="button" class="octicon-btn disabled tooltipped tooltipped-n" aria-label="You must be signed in to make or propose changes">
+            <span class="octicon octicon-pencil"></span>
+          </button>
+
+        <button type="button" class="octicon-btn octicon-btn-danger disabled tooltipped tooltipped-n" aria-label="You must be signed in to make or propose changes">
+          <span class="octicon octicon-trashcan"></span>
+        </button>
+    </div>
+
+    <div class="file-info">
+        335 lines (271 sloc)
+        <span class="file-info-divider"></span>
+      15.875 kb
+    </div>
+  </div>
+
+  <div class="blob-wrapper data type-tex">
+      <table class="highlight tab-size-8 js-file-line-container">
+      <tr>
+        <td id="L1" class="blob-num js-line-number" data-line-number="1"></td>
+        <td id="LC1" class="blob-code js-file-line"><span class="pl-c">%</span></td>
+      </tr>
+      <tr>
+        <td id="L2" class="blob-num js-line-number" data-line-number="2"></td>
+        <td id="LC2" class="blob-code js-file-line"><span class="pl-c">% File nodalida2015.tex</span></td>
+      </tr>
+      <tr>
+        <td id="L3" class="blob-num js-line-number" data-line-number="3"></td>
+        <td id="LC3" class="blob-code js-file-line"><span class="pl-c">%</span></td>
+      </tr>
+      <tr>
+        <td id="L4" class="blob-num js-line-number" data-line-number="4"></td>
+        <td id="LC4" class="blob-code js-file-line"><span class="pl-c">% Contact beata.megyesi@lingfil.uu.se</span></td>
+      </tr>
+      <tr>
+        <td id="L5" class="blob-num js-line-number" data-line-number="5"></td>
+        <td id="LC5" class="blob-code js-file-line"><span class="pl-c">%</span></td>
+      </tr>
+      <tr>
+        <td id="L6" class="blob-num js-line-number" data-line-number="6"></td>
+        <td id="LC6" class="blob-code js-file-line"><span class="pl-c">% Based on the instruction file for EACL 2014</span></td>
+      </tr>
+      <tr>
+        <td id="L7" class="blob-num js-line-number" data-line-number="7"></td>
+        <td id="LC7" class="blob-code js-file-line"><span class="pl-c">% which in turn was based on the instruction files for previous </span></td>
+      </tr>
+      <tr>
+        <td id="L8" class="blob-num js-line-number" data-line-number="8"></td>
+        <td id="LC8" class="blob-code js-file-line"><span class="pl-c">% ACL and EACL conferences.</span></td>
+      </tr>
+      <tr>
+        <td id="L9" class="blob-num js-line-number" data-line-number="9"></td>
+        <td id="LC9" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L10" class="blob-num js-line-number" data-line-number="10"></td>
+        <td id="LC10" class="blob-code js-file-line"><span class="pl-c1">\documentclass</span>[11pt]{article}</td>
+      </tr>
+      <tr>
+        <td id="L11" class="blob-num js-line-number" data-line-number="11"></td>
+        <td id="LC11" class="blob-code js-file-line"><span class="pl-c1">\usepackage</span>{nodalida2015}</td>
+      </tr>
+      <tr>
+        <td id="L12" class="blob-num js-line-number" data-line-number="12"></td>
+        <td id="LC12" class="blob-code js-file-line"><span class="pl-c1">\usepackage</span>{times}</td>
+      </tr>
+      <tr>
+        <td id="L13" class="blob-num js-line-number" data-line-number="13"></td>
+        <td id="LC13" class="blob-code js-file-line"><span class="pl-c1">\usepackage</span>{mathptmx}</td>
+      </tr>
+      <tr>
+        <td id="L14" class="blob-num js-line-number" data-line-number="14"></td>
+        <td id="LC14" class="blob-code js-file-line"><span class="pl-c">%\usepackage{txfonts}</span></td>
+      </tr>
+      <tr>
+        <td id="L15" class="blob-num js-line-number" data-line-number="15"></td>
+        <td id="LC15" class="blob-code js-file-line"><span class="pl-c1">\usepackage</span>{url}</td>
+      </tr>
+      <tr>
+        <td id="L16" class="blob-num js-line-number" data-line-number="16"></td>
+        <td id="LC16" class="blob-code js-file-line"><span class="pl-c1">\usepackage</span>{latexsym}</td>
+      </tr>
+      <tr>
+        <td id="L17" class="blob-num js-line-number" data-line-number="17"></td>
+        <td id="LC17" class="blob-code js-file-line"><span class="pl-c1">\usepackage</span>{cite}</td>
+      </tr>
+      <tr>
+        <td id="L18" class="blob-num js-line-number" data-line-number="18"></td>
+        <td id="LC18" class="blob-code js-file-line"><span class="pl-c1">\usepackage</span>{authordate1-4}</td>
+      </tr>
+      <tr>
+        <td id="L19" class="blob-num js-line-number" data-line-number="19"></td>
+        <td id="LC19" class="blob-code js-file-line"><span class="pl-c1">\usepackage</span>{multirow}</td>
+      </tr>
+      <tr>
+        <td id="L20" class="blob-num js-line-number" data-line-number="20"></td>
+        <td id="LC20" class="blob-code js-file-line"><span class="pl-c1">\special</span>{papersize=210mm,297mm} <span class="pl-c">% to avoid having to use &quot;-t a4&quot; with dvips </span></td>
+      </tr>
+      <tr>
+        <td id="L21" class="blob-num js-line-number" data-line-number="21"></td>
+        <td id="LC21" class="blob-code js-file-line"><span class="pl-c">%\setlength\titlebox{6.5cm}  % You can expand the title box if you really have to</span></td>
+      </tr>
+      <tr>
+        <td id="L22" class="blob-num js-line-number" data-line-number="22"></td>
+        <td id="LC22" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L23" class="blob-num js-line-number" data-line-number="23"></td>
+        <td id="LC23" class="blob-code js-file-line"><span class="pl-c1">\usepackage</span>{color}</td>
+      </tr>
+      <tr>
+        <td id="L24" class="blob-num js-line-number" data-line-number="24"></td>
+        <td id="LC24" class="blob-code js-file-line"><span class="pl-c1">\newcommand</span>{<span class="pl-c1">\todo</span>}[1]{{<span class="pl-c1">\color</span>{cyan}<span class="pl-c1">\textbf</span>{[TODO: }#1<span class="pl-c1">\textbf</span>{]}}}</td>
+      </tr>
+      <tr>
+        <td id="L25" class="blob-num js-line-number" data-line-number="25"></td>
+        <td id="LC25" class="blob-code js-file-line"><span class="pl-c1">\usepackage</span>{tikz}</td>
+      </tr>
+      <tr>
+        <td id="L26" class="blob-num js-line-number" data-line-number="26"></td>
+        <td id="LC26" class="blob-code js-file-line"><span class="pl-c1">\def\checkmark</span>{<span class="pl-c1">\tikz\fill</span>[scale=0.4](0,.35) -- (.25,0) -- (1,.7) --</td>
+      </tr>
+      <tr>
+        <td id="L27" class="blob-num js-line-number" data-line-number="27"></td>
+        <td id="LC27" class="blob-code js-file-line">  (.25,.15) -- cycle;}</td>
+      </tr>
+      <tr>
+        <td id="L28" class="blob-num js-line-number" data-line-number="28"></td>
+        <td id="LC28" class="blob-code js-file-line"><span class="pl-c1">\newcommand</span>{<span class="pl-c1">\Mypm</span>}{<span class="pl-c1">\mathbin</span>{<span class="pl-c1">\tikz</span> [x=1.4ex,y=1.4ex,line width=.1ex] <span class="pl-c1">\draw</span> (0.0,0) -- (1.0,0) (0.5,0.08) -- (0.5,0.92) (0.0,0.5) -- (1.0,0.5);}}<span class="pl-c">%</span></td>
+      </tr>
+      <tr>
+        <td id="L29" class="blob-num js-line-number" data-line-number="29"></td>
+        <td id="LC29" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L30" class="blob-num js-line-number" data-line-number="30"></td>
+        <td id="LC30" class="blob-code js-file-line"><span class="pl-c1">\title</span>{Constraint Grammars as a SAT problem}</td>
+      </tr>
+      <tr>
+        <td id="L31" class="blob-num js-line-number" data-line-number="31"></td>
+        <td id="LC31" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L32" class="blob-num js-line-number" data-line-number="32"></td>
+        <td id="LC32" class="blob-code js-file-line"><span class="pl-c">% \author{Inari Listenmaa \and Koen Claessen \\</span></td>
+      </tr>
+      <tr>
+        <td id="L33" class="blob-num js-line-number" data-line-number="33"></td>
+        <td id="LC33" class="blob-code js-file-line"><span class="pl-c">%  Chalmers University of Technology, Gothenburg, Sweden \\</span></td>
+      </tr>
+      <tr>
+        <td id="L34" class="blob-num js-line-number" data-line-number="34"></td>
+        <td id="LC34" class="blob-code js-file-line"><span class="pl-c">%  {\tt \{inari,koen\}@chalmers.se} }</span></td>
+      </tr>
+      <tr>
+        <td id="L35" class="blob-num js-line-number" data-line-number="35"></td>
+        <td id="LC35" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L36" class="blob-num js-line-number" data-line-number="36"></td>
+        <td id="LC36" class="blob-code js-file-line"><span class="pl-c1">\date</span>{<span class="pl-c1">\today</span>}</td>
+      </tr>
+      <tr>
+        <td id="L37" class="blob-num js-line-number" data-line-number="37"></td>
+        <td id="LC37" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L38" class="blob-num js-line-number" data-line-number="38"></td>
+        <td id="LC38" class="blob-code js-file-line"><span class="pl-c1">\begin</span>{document}</td>
+      </tr>
+      <tr>
+        <td id="L39" class="blob-num js-line-number" data-line-number="39"></td>
+        <td id="LC39" class="blob-code js-file-line"><span class="pl-c1">\maketitle</span></td>
+      </tr>
+      <tr>
+        <td id="L40" class="blob-num js-line-number" data-line-number="40"></td>
+        <td id="LC40" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L41" class="blob-num js-line-number" data-line-number="41"></td>
+        <td id="LC41" class="blob-code js-file-line"><span class="pl-c">%\begin{abstract} \end{abstract}</span></td>
+      </tr>
+      <tr>
+        <td id="L42" class="blob-num js-line-number" data-line-number="42"></td>
+        <td id="LC42" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L43" class="blob-num js-line-number" data-line-number="43"></td>
+        <td id="LC43" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L44" class="blob-num js-line-number" data-line-number="44"></td>
+        <td id="LC44" class="blob-code js-file-line"><span class="pl-c1">\section</span>{Introduction}</td>
+      </tr>
+      <tr>
+        <td id="L45" class="blob-num js-line-number" data-line-number="45"></td>
+        <td id="LC45" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L46" class="blob-num js-line-number" data-line-number="46"></td>
+        <td id="LC46" class="blob-code js-file-line">We represent Constraint Grammars (CG) <span class="pl-c1">\cite</span>{karlsson1995constraint} </td>
+      </tr>
+      <tr>
+        <td id="L47" class="blob-num js-line-number" data-line-number="47"></td>
+        <td id="LC47" class="blob-code js-file-line">as a Boolean satisfiability (SAT) problem.</td>
+      </tr>
+      <tr>
+        <td id="L48" class="blob-num js-line-number" data-line-number="48"></td>
+        <td id="LC48" class="blob-code js-file-line">This is attractive from several reasons: formal logic is</td>
+      </tr>
+      <tr>
+        <td id="L49" class="blob-num js-line-number" data-line-number="49"></td>
+        <td id="LC49" class="blob-code js-file-line">well-studied, and serves as an abstract language to reason about the</td>
+      </tr>
+      <tr>
+        <td id="L50" class="blob-num js-line-number" data-line-number="50"></td>
+        <td id="LC50" class="blob-code js-file-line">properties of CG. Constraint rules encoded in logic capture richer</td>
+      </tr>
+      <tr>
+        <td id="L51" class="blob-num js-line-number" data-line-number="51"></td>
+        <td id="LC51" class="blob-code js-file-line">dependencies between the tags than standard CG.</td>
+      </tr>
+      <tr>
+        <td id="L52" class="blob-num js-line-number" data-line-number="52"></td>
+        <td id="LC52" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L53" class="blob-num js-line-number" data-line-number="53"></td>
+        <td id="LC53" class="blob-code js-file-line">Applying logic to reductionist grammars has been explored earlier by <span class="pl-c1">\cite</span>{lager98,lager_nivre01}, but it was never adopted for use.</td>
+      </tr>
+      <tr>
+        <td id="L54" class="blob-num js-line-number" data-line-number="54"></td>
+        <td id="LC54" class="blob-code js-file-line"><span class="pl-c">% ; logic programming was too slow to be used for tagging or parsing. </span></td>
+      </tr>
+      <tr>
+        <td id="L55" class="blob-num js-line-number" data-line-number="55"></td>
+        <td id="LC55" class="blob-code js-file-line">Since those works, SAT solving techniques have improved significantly, and </td>
+      </tr>
+      <tr>
+        <td id="L56" class="blob-num js-line-number" data-line-number="56"></td>
+        <td id="LC56" class="blob-code js-file-line">they are used in domains such as microprocessor design and computational </td>
+      </tr>
+      <tr>
+        <td id="L57" class="blob-num js-line-number" data-line-number="57"></td>
+        <td id="LC57" class="blob-code js-file-line">biology---these problems easily match or exceed CG in complexity. </td>
+      </tr>
+      <tr>
+        <td id="L58" class="blob-num js-line-number" data-line-number="58"></td>
+        <td id="LC58" class="blob-code js-file-line"><span class="pl-c">% Is this true for grammars with thousands of rules?</span></td>
+      </tr>
+      <tr>
+        <td id="L59" class="blob-num js-line-number" data-line-number="59"></td>
+        <td id="LC59" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L60" class="blob-num js-line-number" data-line-number="60"></td>
+        <td id="LC60" class="blob-code js-file-line">Due to these advances, we were able to revisit the idea and develop it further. </td>
+      </tr>
+      <tr>
+        <td id="L61" class="blob-num js-line-number" data-line-number="61"></td>
+        <td id="LC61" class="blob-code js-file-line"><span class="pl-c">% Whereas \cite{lager98} provides a proof-of-concept description, </span></td>
+      </tr>
+      <tr>
+        <td id="L62" class="blob-num js-line-number" data-line-number="62"></td>
+        <td id="LC62" class="blob-code js-file-line">Missing from the previous logic-based works, </td>
+      </tr>
+      <tr>
+        <td id="L63" class="blob-num js-line-number" data-line-number="63"></td>
+        <td id="LC63" class="blob-code js-file-line">we solve eventual rule conflicts by finding a solution that </td>
+      </tr>
+      <tr>
+        <td id="L64" class="blob-num js-line-number" data-line-number="64"></td>
+        <td id="LC64" class="blob-code js-file-line">discards the least number of rules.</td>
+      </tr>
+      <tr>
+        <td id="L65" class="blob-num js-line-number" data-line-number="65"></td>
+        <td id="LC65" class="blob-code js-file-line"><span class="pl-c">% allows the maximal number of rules to be applied.</span></td>
+      </tr>
+      <tr>
+        <td id="L66" class="blob-num js-line-number" data-line-number="66"></td>
+        <td id="LC66" class="blob-code js-file-line">We test our implementation by parsing texts in the order of 10,000s--100,000s</td>
+      </tr>
+      <tr>
+        <td id="L67" class="blob-num js-line-number" data-line-number="67"></td>
+        <td id="LC67" class="blob-code js-file-line">words, using grammars with hundreds of rules.</td>
+      </tr>
+      <tr>
+        <td id="L68" class="blob-num js-line-number" data-line-number="68"></td>
+        <td id="LC68" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L69" class="blob-num js-line-number" data-line-number="69"></td>
+        <td id="LC69" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L70" class="blob-num js-line-number" data-line-number="70"></td>
+        <td id="LC70" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L71" class="blob-num js-line-number" data-line-number="71"></td>
+        <td id="LC71" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L72" class="blob-num js-line-number" data-line-number="72"></td>
+        <td id="LC72" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L73" class="blob-num js-line-number" data-line-number="73"></td>
+        <td id="LC73" class="blob-code js-file-line"><span class="pl-c">%SAT solving is based on a technique called unit propagation:</span></td>
+      </tr>
+      <tr>
+        <td id="L74" class="blob-num js-line-number" data-line-number="74"></td>
+        <td id="LC74" class="blob-code js-file-line"><span class="pl-c">%a set of complex Boolean clauses is simplified, starting from unit</span></td>
+      </tr>
+      <tr>
+        <td id="L75" class="blob-num js-line-number" data-line-number="75"></td>
+        <td id="LC75" class="blob-code js-file-line"><span class="pl-c">%clauses, which consist of just a single variable, and working up to a</span></td>
+      </tr>
+      <tr>
+        <td id="L76" class="blob-num js-line-number" data-line-number="76"></td>
+        <td id="LC76" class="blob-code js-file-line"><span class="pl-c">%solution, where all variables in the set have a True or False value.</span></td>
+      </tr>
+      <tr>
+        <td id="L77" class="blob-num js-line-number" data-line-number="77"></td>
+        <td id="LC77" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L78" class="blob-num js-line-number" data-line-number="78"></td>
+        <td id="LC78" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L79" class="blob-num js-line-number" data-line-number="79"></td>
+        <td id="LC79" class="blob-code js-file-line"><span class="pl-c">%This increases the computational complexity, </span></td>
+      </tr>
+      <tr>
+        <td id="L80" class="blob-num js-line-number" data-line-number="80"></td>
+        <td id="LC80" class="blob-code js-file-line"><span class="pl-c">%but we are exploring ways in which it could simplify the grammar writing:</span></td>
+      </tr>
+      <tr>
+        <td id="L81" class="blob-num js-line-number" data-line-number="81"></td>
+        <td id="LC81" class="blob-code js-file-line"><span class="pl-c">%our experiments show that less rules are needed, and features like ordering of the rules</span></td>
+      </tr>
+      <tr>
+        <td id="L82" class="blob-num js-line-number" data-line-number="82"></td>
+        <td id="LC82" class="blob-code js-file-line"><span class="pl-c">%and requirement of being unambiguously tagged become irrelevant.</span></td>
+      </tr>
+      <tr>
+        <td id="L83" class="blob-num js-line-number" data-line-number="83"></td>
+        <td id="LC83" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L84" class="blob-num js-line-number" data-line-number="84"></td>
+        <td id="LC84" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L85" class="blob-num js-line-number" data-line-number="85"></td>
+        <td id="LC85" class="blob-code js-file-line"><span class="pl-c">% Uncomment for full paper</span></td>
+      </tr>
+      <tr>
+        <td id="L86" class="blob-num js-line-number" data-line-number="86"></td>
+        <td id="LC86" class="blob-code js-file-line"><span class="pl-c">% \section{Related work}</span></td>
+      </tr>
+      <tr>
+        <td id="L87" class="blob-num js-line-number" data-line-number="87"></td>
+        <td id="LC87" class="blob-code js-file-line"><span class="pl-c">% \label{sect:related}</span></td>
+      </tr>
+      <tr>
+        <td id="L88" class="blob-num js-line-number" data-line-number="88"></td>
+        <td id="LC88" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L89" class="blob-num js-line-number" data-line-number="89"></td>
+        <td id="LC89" class="blob-code js-file-line"><span class="pl-c">% Our work is inspired by \cite{lager98}, which presents constraint rules as a disjunctive logic program,</span></td>
+      </tr>
+      <tr>
+        <td id="L90" class="blob-num js-line-number" data-line-number="90"></td>
+        <td id="LC90" class="blob-code js-file-line"><span class="pl-c">% and \cite{lager_nivre01}, which reconstructs 4 different formalisms from a logical point of view.</span></td>
+      </tr>
+      <tr>
+        <td id="L91" class="blob-num js-line-number" data-line-number="91"></td>
+        <td id="LC91" class="blob-code js-file-line"><span class="pl-c">% \cite{lindberg_eineborg98ilp,asfrent14} use Inductive Logic Programming to learn CG rules from a tagged corpus.</span></td>
+      </tr>
+      <tr>
+        <td id="L92" class="blob-num js-line-number" data-line-number="92"></td>
+        <td id="LC92" class="blob-code js-file-line"><span class="pl-c">% \cite{lager01transformation} discusses ordering of the CG rules.</span></td>
+      </tr>
+      <tr>
+        <td id="L93" class="blob-num js-line-number" data-line-number="93"></td>
+        <td id="LC93" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L94" class="blob-num js-line-number" data-line-number="94"></td>
+        <td id="LC94" class="blob-code js-file-line"><span class="pl-c1">\section</span>{CG as a SAT problem}</td>
+      </tr>
+      <tr>
+        <td id="L95" class="blob-num js-line-number" data-line-number="95"></td>
+        <td id="LC95" class="blob-code js-file-line">Let us demonstrate our approach with the following example in Spanish.</td>
+      </tr>
+      <tr>
+        <td id="L96" class="blob-num js-line-number" data-line-number="96"></td>
+        <td id="LC96" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L97" class="blob-num js-line-number" data-line-number="97"></td>
+        <td id="LC97" class="blob-code js-file-line"><span class="pl-c1">\begin</span>{verbatim}</td>
+      </tr>
+      <tr>
+        <td id="L98" class="blob-num js-line-number" data-line-number="98"></td>
+        <td id="LC98" class="blob-code js-file-line">&quot;&lt;la&gt;&quot;</td>
+      </tr>
+      <tr>
+        <td id="L99" class="blob-num js-line-number" data-line-number="99"></td>
+        <td id="LC99" class="blob-code js-file-line">        &quot;el&quot; det def f sg</td>
+      </tr>
+      <tr>
+        <td id="L100" class="blob-num js-line-number" data-line-number="100"></td>
+        <td id="LC100" class="blob-code js-file-line">        &quot;lo&quot; prn p3 f sg</td>
+      </tr>
+      <tr>
+        <td id="L101" class="blob-num js-line-number" data-line-number="101"></td>
+        <td id="LC101" class="blob-code js-file-line">&quot;&lt;casa&gt;&quot;</td>
+      </tr>
+      <tr>
+        <td id="L102" class="blob-num js-line-number" data-line-number="102"></td>
+        <td id="LC102" class="blob-code js-file-line">        &quot;casa&quot; n f sg</td>
+      </tr>
+      <tr>
+        <td id="L103" class="blob-num js-line-number" data-line-number="103"></td>
+        <td id="LC103" class="blob-code js-file-line">        &quot;casar&quot; v pri p3 sg</td>
+      </tr>
+      <tr>
+        <td id="L104" class="blob-num js-line-number" data-line-number="104"></td>
+        <td id="LC104" class="blob-code js-file-line">        &quot;casar&quot; v imp p2 sg</td>
+      </tr>
+      <tr>
+        <td id="L105" class="blob-num js-line-number" data-line-number="105"></td>
+        <td id="LC105" class="blob-code js-file-line"><span class="pl-c1">\end</span>{verbatim}</td>
+      </tr>
+      <tr>
+        <td id="L106" class="blob-num js-line-number" data-line-number="106"></td>
+        <td id="LC106" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L107" class="blob-num js-line-number" data-line-number="107"></td>
+        <td id="LC107" class="blob-code js-file-line">The ambiguous passage can be either a noun phrase, <span class="pl-c1">\emph</span>{la}<span class="pl-c1">\texttt</span>{&lt;det&gt;} <span class="pl-c1">\emph</span>{casa}<span class="pl-c1">\texttt</span>{&lt;n&gt;} </td>
+      </tr>
+      <tr>
+        <td id="L108" class="blob-num js-line-number" data-line-number="108"></td>
+        <td id="LC108" class="blob-code js-file-line">`the house&#39;  or a verb phrase <span class="pl-c1">\emph</span>{la}<span class="pl-c1">\texttt</span>{&lt;prn&gt;}  <span class="pl-c1">\emph</span>{casa}<span class="pl-c1">\texttt</span>{&lt;v&gt;&lt;pri&gt;&lt;p3&gt;} `(he/she) marries her&#39;.
+</td>
+      </tr>
+      <tr>
+        <td id="L109" class="blob-num js-line-number" data-line-number="109"></td>
+        <td id="LC109" class="blob-code js-file-line">We add the following rules:</td>
+      </tr>
+      <tr>
+        <td id="L110" class="blob-num js-line-number" data-line-number="110"></td>
+        <td id="LC110" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L111" class="blob-num js-line-number" data-line-number="111"></td>
+        <td id="LC111" class="blob-code js-file-line"><span class="pl-c1">\begin</span>{itemize}</td>
+      </tr>
+      <tr>
+        <td id="L112" class="blob-num js-line-number" data-line-number="112"></td>
+        <td id="LC112" class="blob-code js-file-line"><span class="pl-c1">\item</span> [] <span class="pl-c1">\texttt</span>{REMOVE prn IF (1 n) ;} <span class="pl-c1">\\</span></td>
+      </tr>
+      <tr>
+        <td id="L113" class="blob-num js-line-number" data-line-number="113"></td>
+        <td id="LC113" class="blob-code js-file-line">             <span class="pl-c1">\texttt</span>{REMOVE det IF (1 v) ;}</td>
+      </tr>
+      <tr>
+        <td id="L114" class="blob-num js-line-number" data-line-number="114"></td>
+        <td id="LC114" class="blob-code js-file-line"><span class="pl-c1">\end</span>{itemize}</td>
+      </tr>
+      <tr>
+        <td id="L115" class="blob-num js-line-number" data-line-number="115"></td>
+        <td id="LC115" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L116" class="blob-num js-line-number" data-line-number="116"></td>
+        <td id="LC116" class="blob-code js-file-line">Standard CG will apply one<span class="pl-c1">\footnote</span>{To be more cautious, we could require the word at position +1 be disambiguated fully (<span class="pl-c1">\texttt</span>{1C} instead of <span class="pl-c1">\texttt</span>{1}), but in that case, </td>
+      </tr>
+      <tr>
+        <td id="L117" class="blob-num js-line-number" data-line-number="117"></td>
+        <td id="LC117" class="blob-code js-file-line">neither of the rules would be applied.} of the rules to the word <span class="pl-c1">\emph</span>{la}; </td>
+      </tr>
+      <tr>
+        <td id="L118" class="blob-num js-line-number" data-line-number="118"></td>
+        <td id="LC118" class="blob-code js-file-line">either the one that comes first, or by some other heuristic. </td>
+      </tr>
+      <tr>
+        <td id="L119" class="blob-num js-line-number" data-line-number="119"></td>
+        <td id="LC119" class="blob-code js-file-line">The second rule will not fire, because it would remove the last reading. </td>
+      </tr>
+      <tr>
+        <td id="L120" class="blob-num js-line-number" data-line-number="120"></td>
+        <td id="LC120" class="blob-code js-file-line">All readings of <span class="pl-c1">\emph</span>{casa} are left untouched by these rules.</td>
+      </tr>
+      <tr>
+        <td id="L121" class="blob-num js-line-number" data-line-number="121"></td>
+        <td id="LC121" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L122" class="blob-num js-line-number" data-line-number="122"></td>
+        <td id="LC122" class="blob-code js-file-line">The SAT solver performs a search, </td>
+      </tr>
+      <tr>
+        <td id="L123" class="blob-num js-line-number" data-line-number="123"></td>
+        <td id="LC123" class="blob-code js-file-line">and starts building possible models that satisfy both constraints. </td>
+      </tr>
+      <tr>
+        <td id="L124" class="blob-num js-line-number" data-line-number="124"></td>
+        <td id="LC124" class="blob-code js-file-line">In addition to the given constraints, we have default rules to emulate</td>
+      </tr>
+      <tr>
+        <td id="L125" class="blob-num js-line-number" data-line-number="125"></td>
+        <td id="LC125" class="blob-code js-file-line">the CG principles: an analysis is true if no rule affects it,</td>
+      </tr>
+      <tr>
+        <td id="L126" class="blob-num js-line-number" data-line-number="126"></td>
+        <td id="LC126" class="blob-code js-file-line">and at least one analysis for each word is true---the notion of</td>
+      </tr>
+      <tr>
+        <td id="L127" class="blob-num js-line-number" data-line-number="127"></td>
+        <td id="LC127" class="blob-code js-file-line">``last&#39;&#39; is not applicable.</td>
+      </tr>
+      <tr>
+        <td id="L128" class="blob-num js-line-number" data-line-number="128"></td>
+        <td id="LC128" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L129" class="blob-num js-line-number" data-line-number="129"></td>
+        <td id="LC129" class="blob-code js-file-line">With these constraints, we get two solutions.
+The interaction of the rules regarding <span class="pl-c1">\emph</span>{la}  disambiguates the POS of <span class="pl-c1">\emph</span>{casa} for free, and ordering of the rules doesn&#39;t matter.
+</td>
+      </tr>
+      <tr>
+        <td id="L130" class="blob-num js-line-number" data-line-number="130"></td>
+        <td id="LC130" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L131" class="blob-num js-line-number" data-line-number="131"></td>
+        <td id="LC131" class="blob-code js-file-line"><span class="pl-c1">\begin</span>{enumerate}</td>
+      </tr>
+      <tr>
+        <td id="L132" class="blob-num js-line-number" data-line-number="132"></td>
+        <td id="LC132" class="blob-code js-file-line"><span class="pl-c1">\item</span> [<span class="pl-c1">\texttt</span>{1)}]</td>
+      </tr>
+      <tr>
+        <td id="L133" class="blob-num js-line-number" data-line-number="133"></td>
+        <td id="LC133" class="blob-code js-file-line"><span class="pl-c1">\begin</span>{verbatim}</td>
+      </tr>
+      <tr>
+        <td id="L134" class="blob-num js-line-number" data-line-number="134"></td>
+        <td id="LC134" class="blob-code js-file-line">&quot;&lt;la&gt;&quot;</td>
+      </tr>
+      <tr>
+        <td id="L135" class="blob-num js-line-number" data-line-number="135"></td>
+        <td id="LC135" class="blob-code js-file-line">        &quot;el&quot; det def f sg</td>
+      </tr>
+      <tr>
+        <td id="L136" class="blob-num js-line-number" data-line-number="136"></td>
+        <td id="LC136" class="blob-code js-file-line">&quot;&lt;casa&gt;&quot;</td>
+      </tr>
+      <tr>
+        <td id="L137" class="blob-num js-line-number" data-line-number="137"></td>
+        <td id="LC137" class="blob-code js-file-line">        &quot;casa&quot; n f sg</td>
+      </tr>
+      <tr>
+        <td id="L138" class="blob-num js-line-number" data-line-number="138"></td>
+        <td id="LC138" class="blob-code js-file-line"><span class="pl-c1">\end</span>{verbatim}</td>
+      </tr>
+      <tr>
+        <td id="L139" class="blob-num js-line-number" data-line-number="139"></td>
+        <td id="LC139" class="blob-code js-file-line"><span class="pl-c1">\item</span> [<span class="pl-c1">\texttt</span>{2)}]</td>
+      </tr>
+      <tr>
+        <td id="L140" class="blob-num js-line-number" data-line-number="140"></td>
+        <td id="LC140" class="blob-code js-file-line"><span class="pl-c1">\begin</span>{verbatim}</td>
+      </tr>
+      <tr>
+        <td id="L141" class="blob-num js-line-number" data-line-number="141"></td>
+        <td id="LC141" class="blob-code js-file-line">&quot;&lt;la&gt;&quot;</td>
+      </tr>
+      <tr>
+        <td id="L142" class="blob-num js-line-number" data-line-number="142"></td>
+        <td id="LC142" class="blob-code js-file-line">        &quot;lo&quot; prn p3 f sg</td>
+      </tr>
+      <tr>
+        <td id="L143" class="blob-num js-line-number" data-line-number="143"></td>
+        <td id="LC143" class="blob-code js-file-line">&quot;&lt;casa&gt;&quot;</td>
+      </tr>
+      <tr>
+        <td id="L144" class="blob-num js-line-number" data-line-number="144"></td>
+        <td id="LC144" class="blob-code js-file-line">        &quot;casar&quot; v pri p3 sg</td>
+      </tr>
+      <tr>
+        <td id="L145" class="blob-num js-line-number" data-line-number="145"></td>
+        <td id="LC145" class="blob-code js-file-line">        &quot;casar&quot; v imp p2 sg</td>
+      </tr>
+      <tr>
+        <td id="L146" class="blob-num js-line-number" data-line-number="146"></td>
+        <td id="LC146" class="blob-code js-file-line"><span class="pl-c1">\end</span>{verbatim}</td>
+      </tr>
+      <tr>
+        <td id="L147" class="blob-num js-line-number" data-line-number="147"></td>
+        <td id="LC147" class="blob-code js-file-line"><span class="pl-c1">\end</span>{enumerate} </td>
+      </tr>
+      <tr>
+        <td id="L148" class="blob-num js-line-number" data-line-number="148"></td>
+        <td id="LC148" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L149" class="blob-num js-line-number" data-line-number="149"></td>
+        <td id="LC149" class="blob-code js-file-line"><span class="pl-c">%Without more context or additional rules we cannot fully disambiguate the passage,</span></td>
+      </tr>
+      <tr>
+        <td id="L150" class="blob-num js-line-number" data-line-number="150"></td>
+        <td id="LC150" class="blob-code js-file-line"><span class="pl-c">%but unlike with standard CG, we get only legal combinations in one solution.</span></td>
+      </tr>
+      <tr>
+        <td id="L151" class="blob-num js-line-number" data-line-number="151"></td>
+        <td id="LC151" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L152" class="blob-num js-line-number" data-line-number="152"></td>
+        <td id="LC152" class="blob-code js-file-line"><span class="pl-c">%If we add a word, \emph{la casa grande} `the big house&#39; </span></td>
+      </tr>
+      <tr>
+        <td id="L153" class="blob-num js-line-number" data-line-number="153"></td>
+        <td id="LC153" class="blob-code js-file-line"><span class="pl-c">%and a rule that removes verb if followed by an adjective</span></td>
+      </tr>
+      <tr>
+        <td id="L154" class="blob-num js-line-number" data-line-number="154"></td>
+        <td id="LC154" class="blob-code js-file-line"><span class="pl-c">%\texttt{REMOVE (v) IF (1 (adj)) ;}</span></td>
+      </tr>
+      <tr>
+        <td id="L155" class="blob-num js-line-number" data-line-number="155"></td>
+        <td id="LC155" class="blob-code js-file-line"><span class="pl-c">%we will get a unique solution with \texttt{det n adj}.</span></td>
+      </tr>
+      <tr>
+        <td id="L156" class="blob-num js-line-number" data-line-number="156"></td>
+        <td id="LC156" class="blob-code js-file-line"><span class="pl-c">%Standard CG will also remove the verb reading, </span></td>
+      </tr>
+      <tr>
+        <td id="L157" class="blob-num js-line-number" data-line-number="157"></td>
+        <td id="LC157" class="blob-code js-file-line"><span class="pl-c">%but the choice of the word \emph{la} doesn&#39;t contribute to</span></td>
+      </tr>
+      <tr>
+        <td id="L158" class="blob-num js-line-number" data-line-number="158"></td>
+        <td id="LC158" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L159" class="blob-num js-line-number" data-line-number="159"></td>
+        <td id="LC159" class="blob-code js-file-line"><span class="pl-c1">\noindent</span> The most important differences between the traditional and the SAT-based approach are summarised below:</td>
+      </tr>
+      <tr>
+        <td id="L160" class="blob-num js-line-number" data-line-number="160"></td>
+        <td id="LC160" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L161" class="blob-num js-line-number" data-line-number="161"></td>
+        <td id="LC161" class="blob-code js-file-line"><span class="pl-c1">\paragraph</span>{Condition of being unambiguously tagged is irrelevant.}</td>
+      </tr>
+      <tr>
+        <td id="L162" class="blob-num js-line-number" data-line-number="162"></td>
+        <td id="LC162" class="blob-code js-file-line">Rather than waiting for a word to get disambiguated, the SAT solver starts by </td>
+      </tr>
+      <tr>
+        <td id="L163" class="blob-num js-line-number" data-line-number="163"></td>
+        <td id="LC163" class="blob-code js-file-line">making assumptions (e.g. ``<span class="pl-c1">\emph</span>{casa} is a noun&#39;&#39;) and working under them,</td>
+      </tr>
+      <tr>
+        <td id="L164" class="blob-num js-line-number" data-line-number="164"></td>
+        <td id="LC164" class="blob-code js-file-line">discarding the assumption if it doesn&#39;t lead to a model that satisfies all constraints.</td>
+      </tr>
+      <tr>
+        <td id="L165" class="blob-num js-line-number" data-line-number="165"></td>
+        <td id="LC165" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L166" class="blob-num js-line-number" data-line-number="166"></td>
+        <td id="LC166" class="blob-code js-file-line"><span class="pl-c1">\paragraph</span>{Rules are unordered.}</td>
+      </tr>
+      <tr>
+        <td id="L167" class="blob-num js-line-number" data-line-number="167"></td>
+        <td id="LC167" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L168" class="blob-num js-line-number" data-line-number="168"></td>
+        <td id="LC168" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L169" class="blob-num js-line-number" data-line-number="169"></td>
+        <td id="LC169" class="blob-code js-file-line">We have experimented with different strategies:</td>
+      </tr>
+      <tr>
+        <td id="L170" class="blob-num js-line-number" data-line-number="170"></td>
+        <td id="LC170" class="blob-code js-file-line"><span class="pl-c1">\begin</span>{itemize}</td>
+      </tr>
+      <tr>
+        <td id="L171" class="blob-num js-line-number" data-line-number="171"></td>
+        <td id="LC171" class="blob-code js-file-line"><span class="pl-c1">\item</span> Maximise</td>
+      </tr>
+      <tr>
+        <td id="L172" class="blob-num js-line-number" data-line-number="172"></td>
+        <td id="LC172" class="blob-code js-file-line"><span class="pl-c">% \footnote{Something about local maximum}</span></td>
+      </tr>
+      <tr>
+        <td id="L173" class="blob-num js-line-number" data-line-number="173"></td>
+        <td id="LC173" class="blob-code js-file-line"><span class="pl-c">%The algorithm commits to a subset of clauses (e.g.
 \emph{all words must have more than 0 true analyses}), and tries to maximise the number of true variables without changing that subset.
 So it is a local maximum as regards to the given set of clauses.
-number of rules used---We discarded this strategy, because it was too strong: if a rule leads to a conflict in one case, it was not be applied anywhere.
-\item Emulate sequential order---Enter clauses produced by each rule one by one, and assume the solver state reached so far is correct; if a conflict is introduced by new clauses, discard them and move on to next.
-\item [\checkmark] Maximise number of rule applications---If there is a conflict, find the smallest number of rule applications to discard so that the conflict is solved.
-\end{itemize}
-The chosen heuristic produced the best results with examples in the order of tens or hundreds of rules.
-Larger rule sets are yet to be tested.
+</span></td>
+      </tr>
+      <tr>
+        <td id="L174" class="blob-num js-line-number" data-line-number="174"></td>
+        <td id="LC174" class="blob-code js-file-line">number of rules used---We discarded this strategy, because it was too strong: if a rule leads to a conflict in one case, it was not be applied anywhere.</td>
+      </tr>
+      <tr>
+        <td id="L175" class="blob-num js-line-number" data-line-number="175"></td>
+        <td id="LC175" class="blob-code js-file-line"><span class="pl-c1">\item</span> Emulate sequential order---Enter clauses produced by each rule one by one, and assume the solver state reached so far is correct; if a conflict is introduced by new clauses, discard them and move on to next.</td>
+      </tr>
+      <tr>
+        <td id="L176" class="blob-num js-line-number" data-line-number="176"></td>
+        <td id="LC176" class="blob-code js-file-line"><span class="pl-c1">\item</span> [<span class="pl-c1">\checkmark</span>] Maximise number of rule applications---If there is a conflict, find the smallest number of rule applications to discard so that the conflict is solved.</td>
+      </tr>
+      <tr>
+        <td id="L177" class="blob-num js-line-number" data-line-number="177"></td>
+        <td id="LC177" class="blob-code js-file-line"><span class="pl-c1">\end</span>{itemize}</td>
+      </tr>
+      <tr>
+        <td id="L178" class="blob-num js-line-number" data-line-number="178"></td>
+        <td id="LC178" class="blob-code js-file-line">The chosen heuristic produced the best results with examples in the order of tens or hundreds of rules.</td>
+      </tr>
+      <tr>
+        <td id="L179" class="blob-num js-line-number" data-line-number="179"></td>
+        <td id="LC179" class="blob-code js-file-line">Larger rule sets are yet to be tested.</td>
+      </tr>
+      <tr>
+        <td id="L180" class="blob-num js-line-number" data-line-number="180"></td>
+        <td id="LC180" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L181" class="blob-num js-line-number" data-line-number="181"></td>
+        <td id="LC181" class="blob-code js-file-line"><span class="pl-c">% Finite-state implementations of CG \cite{koskenniemi92} also have logically unordered rules. </span></td>
+      </tr>
+      <tr>
+        <td id="L182" class="blob-num js-line-number" data-line-number="182"></td>
+        <td id="LC182" class="blob-code js-file-line"><span class="pl-c">% \todo{Inari: check if Tapanainen 1996 discuss ordering}</span></td>
+      </tr>
+      <tr>
+        <td id="L183" class="blob-num js-line-number" data-line-number="183"></td>
+        <td id="LC183" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L184" class="blob-num js-line-number" data-line-number="184"></td>
+        <td id="LC184" class="blob-code js-file-line"><span class="pl-c1">\paragraph</span>{Rules disambiguate more.}</td>
+      </tr>
+      <tr>
+        <td id="L185" class="blob-num js-line-number" data-line-number="185"></td>
+        <td id="LC185" class="blob-code js-file-line">Considering our example phrase and rules, standard CG implementation</td>
+      </tr>
+      <tr>
+        <td id="L186" class="blob-num js-line-number" data-line-number="186"></td>
+        <td id="LC186" class="blob-code js-file-line">can only remove readings for the word which is tagged as <span class="pl-c1">\texttt</span>{prn} or</td>
+      </tr>
+      <tr>
+        <td id="L187" class="blob-num js-line-number" data-line-number="187"></td>
+        <td id="LC187" class="blob-code js-file-line"><span class="pl-c1">\texttt</span>{det}. The SAT-based implementation interprets the rules as</td>
+      </tr>
+      <tr>
+        <td id="L188" class="blob-num js-line-number" data-line-number="188"></td>
+        <td id="LC188" class="blob-code js-file-line">``determiner and verb together are illegal&#39;&#39;, and is free to take action that concerns also the word in the condition (<span class="pl-c1">\texttt</span>{1 n} or <span class="pl-c1">\texttt</span>{1 v}).</td>
+      </tr>
+      <tr>
+        <td id="L189" class="blob-num js-line-number" data-line-number="189"></td>
+        <td id="LC189" class="blob-code js-file-line">This means that the rules are logically flipped: <span class="pl-c1">\texttt</span>{REMOVE prn IF</td>
+      </tr>
+      <tr>
+        <td id="L190" class="blob-num js-line-number" data-line-number="190"></td>
+        <td id="LC190" class="blob-code js-file-line">  (1 n)} translates into the same logical formula as  <span class="pl-c1">\texttt</span>{REMOVE n</td>
+      </tr>
+      <tr>
+        <td id="L191" class="blob-num js-line-number" data-line-number="191"></td>
+        <td id="LC191" class="blob-code js-file-line">  IF (-1 prn)}. A rule with more conditions can be written as many</td>
+      </tr>
+      <tr>
+        <td id="L192" class="blob-num js-line-number" data-line-number="192"></td>
+        <td id="LC192" class="blob-code js-file-line">rules, each condition taking its turn to be in position to be removed</td>
+      </tr>
+      <tr>
+        <td id="L193" class="blob-num js-line-number" data-line-number="193"></td>
+        <td id="LC193" class="blob-code js-file-line">or selected. <span class="pl-c1">\\</span></td>
+      </tr>
+      <tr>
+        <td id="L194" class="blob-num js-line-number" data-line-number="194"></td>
+        <td id="LC194" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L195" class="blob-num js-line-number" data-line-number="195"></td>
+        <td id="LC195" class="blob-code js-file-line"><span class="pl-c">% \begin{itemize}</span></td>
+      </tr>
+      <tr>
+        <td id="L196" class="blob-num js-line-number" data-line-number="196"></td>
+        <td id="LC196" class="blob-code js-file-line"><span class="pl-c">% \item [] \texttt{REMOVE n IF (-1 prn) ;} \\</span></td>
+      </tr>
+      <tr>
+        <td id="L197" class="blob-num js-line-number" data-line-number="197"></td>
+        <td id="LC197" class="blob-code js-file-line"><span class="pl-c">%          \texttt{REMOVE v IF (-1 det) ;}</span></td>
+      </tr>
+      <tr>
+        <td id="L198" class="blob-num js-line-number" data-line-number="198"></td>
+        <td id="LC198" class="blob-code js-file-line"><span class="pl-c">% \end{itemize}</span></td>
+      </tr>
+      <tr>
+        <td id="L199" class="blob-num js-line-number" data-line-number="199"></td>
+        <td id="LC199" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L200" class="blob-num js-line-number" data-line-number="200"></td>
+        <td id="LC200" class="blob-code js-file-line"><span class="pl-c">% SAT-based approach gives identical results with both sets of rules,</span></td>
+      </tr>
+      <tr>
+        <td id="L201" class="blob-num js-line-number" data-line-number="201"></td>
+        <td id="LC201" class="blob-code js-file-line"><span class="pl-c">% whereas the standard CG would remove one reading from \emph{casa} and leave \emph{la} ambiguous.</span></td>
+      </tr>
+      <tr>
+        <td id="L202" class="blob-num js-line-number" data-line-number="202"></td>
+        <td id="LC202" class="blob-code js-file-line"><span class="pl-c">% Testing this property with more complex rules and larger rule sets remains to be done.</span></td>
+      </tr>
+      <tr>
+        <td id="L203" class="blob-num js-line-number" data-line-number="203"></td>
+        <td id="LC203" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L204" class="blob-num js-line-number" data-line-number="204"></td>
+        <td id="LC204" class="blob-code js-file-line"><span class="pl-c">%Rules with more conditions translate into many rules; rules with negation become complements (\texttt{(*)-X} for \texttt{NOT X}).
+Rules which require the condition to be unambiguously tagged, don&#39;t have an equivalent flip in the standard CG.</span></td>
+      </tr>
+      <tr>
+        <td id="L205" class="blob-num js-line-number" data-line-number="205"></td>
+        <td id="LC205" class="blob-code js-file-line"><span class="pl-c">% We did not expect much practical benefits, save for realising that some rules are bad by having to think further what it implies,</span></td>
+      </tr>
+      <tr>
+        <td id="L206" class="blob-num js-line-number" data-line-number="206"></td>
+        <td id="LC206" class="blob-code js-file-line"><span class="pl-c">%Out of interest, we manually flipped a small grammar and tested it against the original.</span></td>
+      </tr>
+      <tr>
+        <td id="L207" class="blob-num js-line-number" data-line-number="207"></td>
+        <td id="LC207" class="blob-code js-file-line"><span class="pl-c">% \todo{The results show  }</span></td>
+      </tr>
+      <tr>
+        <td id="L208" class="blob-num js-line-number" data-line-number="208"></td>
+        <td id="LC208" class="blob-code js-file-line"><span class="pl-c">%\begin{itemize}</span></td>
+      </tr>
+      <tr>
+        <td id="L209" class="blob-num js-line-number" data-line-number="209"></td>
+        <td id="LC209" class="blob-code js-file-line"><span class="pl-c">%\item []          \texttt{SL Inf IF (-2 V-MOD) (-1 &quot;not&quot;);} \\</span></td>
+      </tr>
+      <tr>
+        <td id="L210" class="blob-num js-line-number" data-line-number="210"></td>
+        <td id="LC210" class="blob-code js-file-line"><span class="pl-c">%$\rightarrow$\texttt{SL V-MOD   IF (1 &quot;not&quot;) (2 INF);} \\</span></td>
+      </tr>
+      <tr>
+        <td id="L211" class="blob-num js-line-number" data-line-number="211"></td>
+        <td id="LC211" class="blob-code js-file-line"><span class="pl-c">%$\rightarrow$\texttt{SL &quot;not&quot; IF (-1 V-MOD) (1 Inf);}</span></td>
+      </tr>
+      <tr>
+        <td id="L212" class="blob-num js-line-number" data-line-number="212"></td>
+        <td id="LC212" class="blob-code js-file-line"><span class="pl-c">%\item []           \texttt{SL Foo IF (NOT 1 Bar)} \\</span></td>
+      </tr>
+      <tr>
+        <td id="L213" class="blob-num js-line-number" data-line-number="213"></td>
+        <td id="LC213" class="blob-code js-file-line"><span class="pl-c">%$\rightarrow$\texttt{SL (*)-Foo if (-1 Bar)}</span></td>
+      </tr>
+      <tr>
+        <td id="L214" class="blob-num js-line-number" data-line-number="214"></td>
+        <td id="LC214" class="blob-code js-file-line"><span class="pl-c">%\end{itemize}</span></td>
+      </tr>
+      <tr>
+        <td id="L215" class="blob-num js-line-number" data-line-number="215"></td>
+        <td id="LC215" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L216" class="blob-num js-line-number" data-line-number="216"></td>
+        <td id="LC216" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L217" class="blob-num js-line-number" data-line-number="217"></td>
+        <td id="LC217" class="blob-code js-file-line"><span class="pl-c1">\noindent</span> These three features influence the way that rules are written. </td>
+      </tr>
+      <tr>
+        <td id="L218" class="blob-num js-line-number" data-line-number="218"></td>
+        <td id="LC218" class="blob-code js-file-line">We predict that less rules are needed; whether this holds in</td>
+      </tr>
+      <tr>
+        <td id="L219" class="blob-num js-line-number" data-line-number="219"></td>
+        <td id="LC219" class="blob-code js-file-line">the order of thousands of rules remains to be tested.</td>
+      </tr>
+      <tr>
+        <td id="L220" class="blob-num js-line-number" data-line-number="220"></td>
+        <td id="LC220" class="blob-code js-file-line">Getting rid of ordering could ease the task of the grammar writer,</td>
+      </tr>
+      <tr>
+        <td id="L221" class="blob-num js-line-number" data-line-number="221"></td>
+        <td id="LC221" class="blob-code js-file-line">since it removes the need to estimate the best sequence of rule</td>
+      </tr>
+      <tr>
+        <td id="L222" class="blob-num js-line-number" data-line-number="222"></td>
+        <td id="LC222" class="blob-code js-file-line">applications.</td>
+      </tr>
+      <tr>
+        <td id="L223" class="blob-num js-line-number" data-line-number="223"></td>
+        <td id="LC223" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L224" class="blob-num js-line-number" data-line-number="224"></td>
+        <td id="LC224" class="blob-code js-file-line">However, the results using our approach should not be too different from the</td>
+      </tr>
+      <tr>
+        <td id="L225" class="blob-num js-line-number" data-line-number="225"></td>
+        <td id="LC225" class="blob-code js-file-line">existing ones---our preliminary evaluation against VISL CG-3 shows</td>
+      </tr>
+      <tr>
+        <td id="L226" class="blob-num js-line-number" data-line-number="226"></td>
+        <td id="LC226" class="blob-code js-file-line">promising results in this regard.</td>
+      </tr>
+      <tr>
+        <td id="L227" class="blob-num js-line-number" data-line-number="227"></td>
+        <td id="LC227" class="blob-code js-file-line"><span class="pl-c1">\pagebreak</span></td>
+      </tr>
+      <tr>
+        <td id="L228" class="blob-num js-line-number" data-line-number="228"></td>
+        <td id="LC228" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L229" class="blob-num js-line-number" data-line-number="229"></td>
+        <td id="LC229" class="blob-code js-file-line"><span class="pl-c1">\section</span>{Results and evaluation}</td>
+      </tr>
+      <tr>
+        <td id="L230" class="blob-num js-line-number" data-line-number="230"></td>
+        <td id="LC230" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L231" class="blob-num js-line-number" data-line-number="231"></td>
+        <td id="LC231" class="blob-code js-file-line"><span class="pl-c1">\paragraph</span>{Time}</td>
+      </tr>
+      <tr>
+        <td id="L232" class="blob-num js-line-number" data-line-number="232"></td>
+        <td id="LC232" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L233" class="blob-num js-line-number" data-line-number="233"></td>
+        <td id="LC233" class="blob-code js-file-line">The worst-case complexity of SAT is exponential, whereas the standard</td>
+      </tr>
+      <tr>
+        <td id="L234" class="blob-num js-line-number" data-line-number="234"></td>
+        <td id="LC234" class="blob-code js-file-line">implementations of CG are polynomial, but with advances in SAT solving</td>
+      </tr>
+      <tr>
+        <td id="L235" class="blob-num js-line-number" data-line-number="235"></td>
+        <td id="LC235" class="blob-code js-file-line">techniques, the performance in the average case in practice is much more feasible than in the previous works done in 90s--00s.</td>
+      </tr>
+      <tr>
+        <td id="L236" class="blob-num js-line-number" data-line-number="236"></td>
+        <td id="LC236" class="blob-code js-file-line">We used the open-source SAT solver MiniSat <span class="pl-c1">\cite</span>{een04sat}.</td>
+      </tr>
+      <tr>
+        <td id="L237" class="blob-num js-line-number" data-line-number="237"></td>
+        <td id="LC237" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L238" class="blob-num js-line-number" data-line-number="238"></td>
+        <td id="LC238" class="blob-code js-file-line">Table~<span class="pl-c1">\ref</span>{table:time} shows the execution time compared to VISL CG-3. </td>
+      </tr>
+      <tr>
+        <td id="L239" class="blob-num js-line-number" data-line-number="239"></td>
+        <td id="LC239" class="blob-code js-file-line">The number of rules in the grammar is more significant than the word</td>
+      </tr>
+      <tr>
+        <td id="L240" class="blob-num js-line-number" data-line-number="240"></td>
+        <td id="LC240" class="blob-code js-file-line">count; this holds also for the performance of VISL CG-3.</td>
+      </tr>
+      <tr>
+        <td id="L241" class="blob-num js-line-number" data-line-number="241"></td>
+        <td id="LC241" class="blob-code js-file-line">From the SAT solving side, maximisation is the most costly operation.</td>
+      </tr>
+      <tr>
+        <td id="L242" class="blob-num js-line-number" data-line-number="242"></td>
+        <td id="LC242" class="blob-code js-file-line">Parsing Don Quijote (384,155 words) with the rule set of 261 rules,</td>
+      </tr>
+      <tr>
+        <td id="L243" class="blob-num js-line-number" data-line-number="243"></td>
+        <td id="LC243" class="blob-code js-file-line">the function was called 147,253 times, and with 19 rules, 132,255 times.</td>
+      </tr>
+      <tr>
+        <td id="L244" class="blob-num js-line-number" data-line-number="244"></td>
+        <td id="LC244" class="blob-code js-file-line"><span class="pl-c">%However, with both rule sets, half of the sentences in Don Quijote</span></td>
+      </tr>
+      <tr>
+        <td id="L245" class="blob-num js-line-number" data-line-number="245"></td>
+        <td id="LC245" class="blob-code js-file-line"><span class="pl-c">%needed less than 6 calls of maximise, and 75 \% of the sentences</span></td>
+      </tr>
+      <tr>
+        <td id="L246" class="blob-num js-line-number" data-line-number="246"></td>
+        <td id="LC246" class="blob-code js-file-line"><span class="pl-c">%needed less than 14.</span></td>
+      </tr>
+      <tr>
+        <td id="L247" class="blob-num js-line-number" data-line-number="247"></td>
+        <td id="LC247" class="blob-code js-file-line">The difference in the execution times suggests that there are other</td>
+      </tr>
+      <tr>
+        <td id="L248" class="blob-num js-line-number" data-line-number="248"></td>
+        <td id="LC248" class="blob-code js-file-line">reasons for the worse performance---this is to be expected,</td>
+      </tr>
+      <tr>
+        <td id="L249" class="blob-num js-line-number" data-line-number="249"></td>
+        <td id="LC249" class="blob-code js-file-line">SATCG is currently just a naive proof-of-concept implementation with no optimisations.</td>
+      </tr>
+      <tr>
+        <td id="L250" class="blob-num js-line-number" data-line-number="250"></td>
+        <td id="LC250" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L251" class="blob-num js-line-number" data-line-number="251"></td>
+        <td id="LC251" class="blob-code js-file-line"><span class="pl-c1">\begin</span>{table}</td>
+      </tr>
+      <tr>
+        <td id="L252" class="blob-num js-line-number" data-line-number="252"></td>
+        <td id="LC252" class="blob-code js-file-line">  <span class="pl-c1">\centering</span></td>
+      </tr>
+      <tr>
+        <td id="L253" class="blob-num js-line-number" data-line-number="253"></td>
+        <td id="LC253" class="blob-code js-file-line">  <span class="pl-c1">\begin</span>{tabular}{|l|c|c|c|}</td>
+      </tr>
+      <tr>
+        <td id="L254" class="blob-num js-line-number" data-line-number="254"></td>
+        <td id="LC254" class="blob-code js-file-line">     <span class="pl-c1">\hline</span></td>
+      </tr>
+      <tr>
+        <td id="L255" class="blob-num js-line-number" data-line-number="255"></td>
+        <td id="LC255" class="blob-code js-file-line">   <span class="pl-c1">\textbf</span>{<span class="pl-cce">\#</span> words} &amp; <span class="pl-c1">\textbf</span>{<span class="pl-cce">\#</span> rules} &amp;  <span class="pl-c1">\textbf</span>{SATCG} &amp; <span class="pl-c1">\textbf</span>{VISL CG-3} <span class="pl-c1">\\</span> <span class="pl-c1">\hline</span></td>
+      </tr>
+      <tr>
+        <td id="L256" class="blob-num js-line-number" data-line-number="256"></td>
+        <td id="LC256" class="blob-code js-file-line">                384,155  &amp; 19   &amp;  13.7s   &amp; 4.2s<span class="pl-c1">\\</span> <span class="pl-c">%\hline</span></td>
+      </tr>
+      <tr>
+        <td id="L257" class="blob-num js-line-number" data-line-number="257"></td>
+        <td id="LC257" class="blob-code js-file-line">                124,493  &amp; 99   &amp;  55.7s   &amp; 4.8s <span class="pl-c1">\\</span> <span class="pl-c">% \hline</span></td>
+      </tr>
+      <tr>
+        <td id="L258" class="blob-num js-line-number" data-line-number="258"></td>
+        <td id="LC258" class="blob-code js-file-line">	        384,155  &amp; 130  &amp;  1m14.1s &amp; 6.1s <span class="pl-c1">\\</span> <span class="pl-c">%\hline</span></td>
+      </tr>
+      <tr>
+        <td id="L259" class="blob-num js-line-number" data-line-number="259"></td>
+        <td id="LC259" class="blob-code js-file-line">                384,155  &amp; 261  &amp;  2m53.7s &amp; 10.7s <span class="pl-c1">\\</span> <span class="pl-c1">\hline</span></td>
+      </tr>
+      <tr>
+        <td id="L260" class="blob-num js-line-number" data-line-number="260"></td>
+        <td id="LC260" class="blob-code js-file-line">  <span class="pl-c1">\end</span>{tabular}</td>
+      </tr>
+      <tr>
+        <td id="L261" class="blob-num js-line-number" data-line-number="261"></td>
+        <td id="LC261" class="blob-code js-file-line">  <span class="pl-c1">\caption</span>{Execution times.}</td>
+      </tr>
+      <tr>
+        <td id="L262" class="blob-num js-line-number" data-line-number="262"></td>
+        <td id="LC262" class="blob-code js-file-line">  <span class="pl-c1">\label</span>{table:time}</td>
+      </tr>
+      <tr>
+        <td id="L263" class="blob-num js-line-number" data-line-number="263"></td>
+        <td id="LC263" class="blob-code js-file-line"><span class="pl-c1">\end</span>{table}</td>
+      </tr>
+      <tr>
+        <td id="L264" class="blob-num js-line-number" data-line-number="264"></td>
+        <td id="LC264" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L265" class="blob-num js-line-number" data-line-number="265"></td>
+        <td id="LC265" class="blob-code js-file-line"><span class="pl-c1">\paragraph</span>{Performance against VISL CG-3}</td>
+      </tr>
+      <tr>
+        <td id="L266" class="blob-num js-line-number" data-line-number="266"></td>
+        <td id="LC266" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L267" class="blob-num js-line-number" data-line-number="267"></td>
+        <td id="LC267" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L268" class="blob-num js-line-number" data-line-number="268"></td>
+        <td id="LC268" class="blob-code js-file-line">We took a tagged corpus<span class="pl-c1">\footnote</span>{https://svn.code.sf.net/p/apertium/svn/branches/apertium-swpost/apertium-en-es/es-tagger-data/es.tagged} of 21865 words for Spanish, </td>
+      </tr>
+      <tr>
+        <td id="L269" class="blob-num js-line-number" data-line-number="269"></td>
+        <td id="LC269" class="blob-code js-file-line">and a small constraint grammar<span class="pl-c1">\footnote</span>{https://svn.code.sf.net/p/apertium/svn/languages/apertium-spa/apertium-spa.spa.rlx}, produced independently of the authors.</td>
+      </tr>
+      <tr>
+        <td id="L270" class="blob-num js-line-number" data-line-number="270"></td>
+        <td id="LC270" class="blob-code js-file-line"><span class="pl-c">% We ignored substitute rules, but kept subreadings and unification, even though our implementation doesn&#39;t handle them.</span></td>
+      </tr>
+      <tr>
+        <td id="L271" class="blob-num js-line-number" data-line-number="271"></td>
+        <td id="LC271" class="blob-code js-file-line">We only took select and remove rules (no substitute/iff), after which the constraint grammar had 261 rules.</td>
+      </tr>
+      <tr>
+        <td id="L272" class="blob-num js-line-number" data-line-number="272"></td>
+        <td id="LC272" class="blob-code js-file-line">With this setup, we took the text of the tagged corpus with all</td>
+      </tr>
+      <tr>
+        <td id="L273" class="blob-num js-line-number" data-line-number="273"></td>
+        <td id="LC273" class="blob-code js-file-line">ambiguities, and ran both SATCG and VISL CG-3 with the same grammar. </td>
+      </tr>
+      <tr>
+        <td id="L274" class="blob-num js-line-number" data-line-number="274"></td>
+        <td id="LC274" class="blob-code js-file-line">Treating the corpus as the gold standard, the disambiguation by</td>
+      </tr>
+      <tr>
+        <td id="L275" class="blob-num js-line-number" data-line-number="275"></td>
+        <td id="LC275" class="blob-code js-file-line">SATCG achieves 75.09 <span class="pl-cce">\%</span> correct result and VISL CG-3 75.80 <span class="pl-cce">\%</span>.</td>
+      </tr>
+      <tr>
+        <td id="L276" class="blob-num js-line-number" data-line-number="276"></td>
+        <td id="LC276" class="blob-code js-file-line">Similar patterns were observed with small (<span class="pl-c1">\textless</span>{}20) rule sets</td>
+      </tr>
+      <tr>
+        <td id="L277" class="blob-num js-line-number" data-line-number="277"></td>
+        <td id="LC277" class="blob-code js-file-line">written by the authors; depending on the subset, SATCG and VISL CG-3 had</td>
+      </tr>
+      <tr>
+        <td id="L278" class="blob-num js-line-number" data-line-number="278"></td>
+        <td id="LC278" class="blob-code js-file-line">a difference of at most <span class="pl-s"><span class="pl-pds">$</span><span class="pl-c1">\Mypm</span><span class="pl-pds">$</span></span> 1.5 <span class="pl-cce">\%</span>.
+</td>
+      </tr>
+      <tr>
+        <td id="L279" class="blob-num js-line-number" data-line-number="279"></td>
+        <td id="LC279" class="blob-code js-file-line">Introducing rules one by one up to 19, the</td>
+      </tr>
+      <tr>
+        <td id="L280" class="blob-num js-line-number" data-line-number="280"></td>
+        <td id="LC280" class="blob-code js-file-line">performance improved in a very similar rate, with less than 0.5 <span class="pl-cce">\%</span></td>
+      </tr>
+      <tr>
+        <td id="L281" class="blob-num js-line-number" data-line-number="281"></td>
+        <td id="LC281" class="blob-code js-file-line">difference between the systems at each new rule.</td>
+      </tr>
+      <tr>
+        <td id="L282" class="blob-num js-line-number" data-line-number="282"></td>
+        <td id="LC282" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L283" class="blob-num js-line-number" data-line-number="283"></td>
+        <td id="LC283" class="blob-code js-file-line"><span class="pl-c">% Additional tests could include plugging SATCG into Apertium</span></td>
+      </tr>
+      <tr>
+        <td id="L284" class="blob-num js-line-number" data-line-number="284"></td>
+        <td id="LC284" class="blob-code js-file-line"><span class="pl-c">% translation pipeline, and comparing the translation quality.</span></td>
+      </tr>
+      <tr>
+        <td id="L285" class="blob-num js-line-number" data-line-number="285"></td>
+        <td id="LC285" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L286" class="blob-num js-line-number" data-line-number="286"></td>
+        <td id="LC286" class="blob-code js-file-line"><span class="pl-c">% \paragraph{Removing rules}</span></td>
+      </tr>
+      <tr>
+        <td id="L287" class="blob-num js-line-number" data-line-number="287"></td>
+        <td id="LC287" class="blob-code js-file-line"><span class="pl-c">% \begin{table}</span></td>
+      </tr>
+      <tr>
+        <td id="L288" class="blob-num js-line-number" data-line-number="288"></td>
+        <td id="LC288" class="blob-code js-file-line"><span class="pl-c">%   \centering</span></td>
+      </tr>
+      <tr>
+        <td id="L289" class="blob-num js-line-number" data-line-number="289"></td>
+        <td id="LC289" class="blob-code js-file-line"><span class="pl-c">%   \begin{tabular}{|l|c|c|}</span></td>
+      </tr>
+      <tr>
+        <td id="L290" class="blob-num js-line-number" data-line-number="290"></td>
+        <td id="LC290" class="blob-code js-file-line"><span class="pl-c">%      \hline</span></td>
+      </tr>
+      <tr>
+        <td id="L291" class="blob-num js-line-number" data-line-number="291"></td>
+        <td id="LC291" class="blob-code js-file-line"><span class="pl-c">%     \textbf{\# rules} &amp; \textbf{SATCG} &amp; \textbf{VISL CG-3} \\</span></td>
+      </tr>
+      <tr>
+        <td id="L292" class="blob-num js-line-number" data-line-number="292"></td>
+        <td id="LC292" class="blob-code js-file-line"><span class="pl-c">%     \hline </span></td>
+      </tr>
+      <tr>
+        <td id="L293" class="blob-num js-line-number" data-line-number="293"></td>
+        <td id="LC293" class="blob-code js-file-line"><span class="pl-c">% 	19 &amp;  83.15 &amp; 83.33 \\ \hline  % 19 out of 23 preselected by authors</span></td>
+      </tr>
+      <tr>
+        <td id="L294" class="blob-num js-line-number" data-line-number="294"></td>
+        <td id="LC294" class="blob-code js-file-line"><span class="pl-c">% 	18 &amp;  83.19 &amp; 83.04 \\ \hline  % 18 out of 19</span></td>
+      </tr>
+      <tr>
+        <td id="L295" class="blob-num js-line-number" data-line-number="295"></td>
+        <td id="LC295" class="blob-code js-file-line"><span class="pl-c">% 	17 &amp;  83.201 &amp; 83.03 \\ \hline   % 17 out of 18/19, no diff</span></td>
+      </tr>
+      <tr>
+        <td id="L296" class="blob-num js-line-number" data-line-number="296"></td>
+        <td id="LC296" class="blob-code js-file-line"><span class="pl-c">% 	16 &amp;  83.197 &amp; 83.030 \\ \hline  % 16 out of 17/18, no diff</span></td>
+      </tr>
+      <tr>
+        <td id="L297" class="blob-num js-line-number" data-line-number="297"></td>
+        <td id="LC297" class="blob-code js-file-line"><span class="pl-c">% 	15 &amp;  83.188 &amp; 83.022\\ \hline  % 15 out of 16/17.
+Tie with 14 for SATCG, chose to retain the rule that gave better results for VISL CG-3.</span></td>
+      </tr>
+      <tr>
+        <td id="L298" class="blob-num js-line-number" data-line-number="298"></td>
+        <td id="LC298" class="blob-code js-file-line"><span class="pl-c">% 	14 &amp;  83.180 &amp; 83.005 \\ \hline  % Tie with 15 for SATCG.</span></td>
+      </tr>
+      <tr>
+        <td id="L299" class="blob-num js-line-number" data-line-number="299"></td>
+        <td id="LC299" class="blob-code js-file-line"><span class="pl-c">% 	13 &amp;  83.159 &amp; 82.951 \\ \hline  % 13 out of 15 </span></td>
+      </tr>
+      <tr>
+        <td id="L300" class="blob-num js-line-number" data-line-number="300"></td>
+        <td id="LC300" class="blob-code js-file-line"><span class="pl-c">% 	12 &amp;  83.134 &amp; 82.884\\ \hline  % 12 out of 14</span></td>
+      </tr>
+      <tr>
+        <td id="L301" class="blob-num js-line-number" data-line-number="301"></td>
+        <td id="LC301" class="blob-code js-file-line"><span class="pl-c">% 	11 &amp;  83.089 &amp; 82.847\\ \hline  </span></td>
+      </tr>
+      <tr>
+        <td id="L302" class="blob-num js-line-number" data-line-number="302"></td>
+        <td id="LC302" class="blob-code js-file-line"><span class="pl-c">% 	10 &amp;  83.051 &amp; 82.802\\ \hline  </span></td>
+      </tr>
+      <tr>
+        <td id="L303" class="blob-num js-line-number" data-line-number="303"></td>
+        <td id="LC303" class="blob-code js-file-line"><span class="pl-c">% 	9  &amp; 82.997 &amp; 82.802 \\ \hline</span></td>
+      </tr>
+      <tr>
+        <td id="L304" class="blob-num js-line-number" data-line-number="304"></td>
+        <td id="LC304" class="blob-code js-file-line"><span class="pl-c">% 	8  &amp; 82.906 &amp; 82.681 \\ \hline  % REMOVE:r_imp Imp IF (0 Subj) (-1* CnjSub) ; </span></td>
+      </tr>
+      <tr>
+        <td id="L305" class="blob-num js-line-number" data-line-number="305"></td>
+        <td id="LC305" class="blob-code js-file-line"><span class="pl-c">% 	7  &amp; 82.743 &amp; 82.594 \\ \hline  % SELECT:s_pro PrnIndep IF (1C VerbFin) ;</span></td>
+      </tr>
+      <tr>
+        <td id="L306" class="blob-num js-line-number" data-line-number="306"></td>
+        <td id="LC306" class="blob-code js-file-line"><span class="pl-c">% 	6  &amp; 82.556 &amp; 82.398\\ \hline  %  REMOVE:r_v_v Verb IF (-1C Vblex) ;</span></td>
+      </tr>
+      <tr>
+        <td id="L307" class="blob-num js-line-number" data-line-number="307"></td>
+        <td id="LC307" class="blob-code js-file-line"><span class="pl-c">% 	5  &amp; 82.327 &amp; 82.028 \\ \hline % SELECT:s_adv Adv IF (1 Cnj_Rel_End OR Prep OR Det)</span></td>
+      </tr>
+      <tr>
+        <td id="L308" class="blob-num js-line-number" data-line-number="308"></td>
+        <td id="LC308" class="blob-code js-file-line"><span class="pl-c">% 	4  &amp; 82.065 &amp; 82.028 \\ \hline % REMOVE:r_vfin_prep VerbFin IF (0 Prep) (1 Inf) </span></td>
+      </tr>
+      <tr>
+        <td id="L309" class="blob-num js-line-number" data-line-number="309"></td>
+        <td id="LC309" class="blob-code js-file-line"><span class="pl-c">% 	3  &amp; 81.387 &amp; 81.233 \\ \hline % REMOVE:r_noun N    IF (-1 N OR Prn)</span></td>
+      </tr>
+      <tr>
+        <td id="L310" class="blob-num js-line-number" data-line-number="310"></td>
+        <td id="LC310" class="blob-code js-file-line"><span class="pl-c">% 	2  &amp; 80.23 &amp; 79.97 \\ \hline % SELECT:s_pri_1 (pri p3) IF (0 (vblex imp p2 sg))</span></td>
+      </tr>
+      <tr>
+        <td id="L311" class="blob-num js-line-number" data-line-number="311"></td>
+        <td id="LC311" class="blob-code js-file-line"><span class="pl-c">% 	1  &amp; 76.69 &amp; 76.62 \\ \hline </span></td>
+      </tr>
+      <tr>
+        <td id="L312" class="blob-num js-line-number" data-line-number="312"></td>
+        <td id="LC312" class="blob-code js-file-line"><span class="pl-c">% 	0  &amp; 70.70 &amp; 70.70 \\ \hline</span></td>
+      </tr>
+      <tr>
+        <td id="L313" class="blob-num js-line-number" data-line-number="313"></td>
+        <td id="LC313" class="blob-code js-file-line"><span class="pl-c">%   \end{tabular}</span></td>
+      </tr>
+      <tr>
+        <td id="L314" class="blob-num js-line-number" data-line-number="314"></td>
+        <td id="LC314" class="blob-code js-file-line"><span class="pl-c">%   \caption{Result of removing rules one by one, and how close to the gold standard they get.}</span></td>
+      </tr>
+      <tr>
+        <td id="L315" class="blob-num js-line-number" data-line-number="315"></td>
+        <td id="LC315" class="blob-code js-file-line"><span class="pl-c">%   \label{table:results_remove}</span></td>
+      </tr>
+      <tr>
+        <td id="L316" class="blob-num js-line-number" data-line-number="316"></td>
+        <td id="LC316" class="blob-code js-file-line"><span class="pl-c">% \end{table}</span></td>
+      </tr>
+      <tr>
+        <td id="L317" class="blob-num js-line-number" data-line-number="317"></td>
+        <td id="LC317" class="blob-code js-file-line"><span class="pl-c">% In this experiment, we took 19 rules written by the authors, which give 83.15 \% correct results with SATCG and 83.33 \% with VISL CG-3.</span></td>
+      </tr>
+      <tr>
+        <td id="L318" class="blob-num js-line-number" data-line-number="318"></td>
+        <td id="LC318" class="blob-code js-file-line"><span class="pl-c">% Then we removed rules one by one; after getting a score for $n$ rules, we took the combination of $n-1$ rules for which SATCG gives the best score, and we ran the rule set with VISL CG-3 for comparison.
+Table~\ref{table:results_remove} shows the percentages compared to the gold standard for both systems.</span></td>
+      </tr>
+      <tr>
+        <td id="L319" class="blob-num js-line-number" data-line-number="319"></td>
+        <td id="LC319" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L320" class="blob-num js-line-number" data-line-number="320"></td>
+        <td id="LC320" class="blob-code js-file-line"><span class="pl-c1">\section</span>{Conclusions}</td>
+      </tr>
+      <tr>
+        <td id="L321" class="blob-num js-line-number" data-line-number="321"></td>
+        <td id="LC321" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L322" class="blob-num js-line-number" data-line-number="322"></td>
+        <td id="LC322" class="blob-code js-file-line">SAT-solvers are nowadays powerful enough to be used for dealing with Constraint Grammars.
+A logic-based approach to CG has possible advantages over more traditional approaches; a SAT-solver may disambiguate more words, and may do so more precisely.
+Also, the SAT-solver requires less rules, and these rules are simpler.
+The ordering of rules is something we found was incompatible with a logic-based approach.
+We compensate this by maximising rule applications.
+Whether this is an advantage or disadvantage remains to be seen.
+Our initial experimental results are promising.</td>
+      </tr>
+      <tr>
+        <td id="L323" class="blob-num js-line-number" data-line-number="323"></td>
+        <td id="LC323" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L324" class="blob-num js-line-number" data-line-number="324"></td>
+        <td id="LC324" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L325" class="blob-num js-line-number" data-line-number="325"></td>
+        <td id="LC325" class="blob-code js-file-line"><span class="pl-c">% \section*{Acknowledgments}</span></td>
+      </tr>
+      <tr>
+        <td id="L326" class="blob-num js-line-number" data-line-number="326"></td>
+        <td id="LC326" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L327" class="blob-num js-line-number" data-line-number="327"></td>
+        <td id="LC327" class="blob-code js-file-line"><span class="pl-c">% Do not number the acknowledgment section. Do not include this section</span></td>
+      </tr>
+      <tr>
+        <td id="L328" class="blob-num js-line-number" data-line-number="328"></td>
+        <td id="LC328" class="blob-code js-file-line"><span class="pl-c">% when submitting your paper for review.</span></td>
+      </tr>
+      <tr>
+        <td id="L329" class="blob-num js-line-number" data-line-number="329"></td>
+        <td id="LC329" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L330" class="blob-num js-line-number" data-line-number="330"></td>
+        <td id="LC330" class="blob-code js-file-line"><span class="pl-c1">\bibliographystyle</span>{acl}</td>
+      </tr>
+      <tr>
+        <td id="L331" class="blob-num js-line-number" data-line-number="331"></td>
+        <td id="LC331" class="blob-code js-file-line"><span class="pl-c1">\bibliography</span>{cg}</td>
+      </tr>
+      <tr>
+        <td id="L332" class="blob-num js-line-number" data-line-number="332"></td>
+        <td id="LC332" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L333" class="blob-num js-line-number" data-line-number="333"></td>
+        <td id="LC333" class="blob-code js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L334" class="blob-num js-line-number" data-line-number="334"></td>
+        <td id="LC334" class="blob-code js-file-line"><span class="pl-c1">\end</span>{document}</td>
+      </tr>
+</table>
 
-% Finite-state implementations of CG \cite{koskenniemi92} also have logically unordered rules. 
-% \todo{Inari: check if Tapanainen 1996 discuss ordering}
+  </div>
 
-\paragraph{Rules disambiguate more.}
-Considering our example phrase and rules, standard CG implementation
-can only remove readings for the word which is tagged as \texttt{prn} or
-\texttt{det}. The SAT-based implementation interprets the rules as
-``determiner and verb together are illegal'', and is free to take action that concerns also the word in the condition \texttt{(1 n)} or \texttt{(1 v)}.
-This means that the rules are logically flipped: \texttt{REMOVE prn IF
-  (1 n)} translates into the same logical formula as  \texttt{REMOVE n
-  IF (-1 prn)}. A rule with more conditions can be written as many
-rules, each condition taking its turn to be in position to be removed
-or selected. \\
+</div>
 
-% \begin{itemize}
-% \item [] \texttt{REMOVE n IF (-1 prn) ;} \\
-%          \texttt{REMOVE v IF (-1 det) ;}
-% \end{itemize}
+<a href="#jump-to-line" rel="facebox[.linejump]" data-hotkey="l" style="display:none">Jump to Line</a>
+<div id="jump-to-line" style="display:none">
+  <form accept-charset="UTF-8" action="" class="js-jump-to-line-form" method="get"><div style="margin:0;padding:0;display:inline"><input name="utf8" type="hidden" value="&#x2713;" /></div>
+    <input class="linejump-input js-jump-to-line-field" type="text" placeholder="Jump to line&hellip;" autofocus>
+    <button type="submit" class="btn">Go</button>
+</form></div>
 
-% SAT-based approach gives identical results with both sets of rules,
-% whereas the standard CG would remove one reading from \emph{casa} and leave \emph{la} ambiguous.
-% Testing this property with more complex rules and larger rule sets remains to be done.
+        </div>
 
-%Rules with more conditions translate into many rules; rules with negation become complements (\texttt{(*)-X} for \texttt{NOT X}).
-Rules which require the condition to be unambiguously tagged, don't have an equivalent flip in the standard CG.
-% We did not expect much practical benefits, save for realising that some rules are bad by having to think further what it implies,
-%Out of interest, we manually flipped a small grammar and tested it against the original.
-% \todo{The results show  }
-%\begin{itemize}
-%\item []          \texttt{SL Inf IF (-2 V-MOD) (-1 "not");} \\
-%$\rightarrow$\texttt{SL V-MOD   IF (1 "not") (2 INF);} \\
-%$\rightarrow$\texttt{SL "not" IF (-1 V-MOD) (1 Inf);}
-%\item []           \texttt{SL Foo IF (NOT 1 Bar)} \\
-%$\rightarrow$\texttt{SL (*)-Foo if (-1 Bar)}
-%\end{itemize}
+      </div><!-- /.repo-container -->
+      <div class="modal-backdrop"></div>
+    </div><!-- /.container -->
+  </div><!-- /.site -->
 
-\noindent These three features might change the way that rules are written. 
-We predict that less rules are needed; whether this holds in
-the order of thousands of rules remains to be tested.
-Getting rid of ordering could ease the task of the grammar writer,
-removing the responsibility to estimate the best sequence of rule applications.
-However, the implementation should not be too different from the
-existing ones---our preliminary evaluation against VISL CG-3 shows
-promising results in this regard.
-\pagebreak
+    </div><!-- /.wrapper -->
 
-\section{Results and evaluation}
+      <div class="container">
+  <div class="site-footer" role="contentinfo">
+    <ul class="site-footer-links right">
+        <li><a href="https://status.github.com/" data-ga-click="Footer, go to status, text:status">Status</a></li>
+      <li><a href="https://developer.github.com" data-ga-click="Footer, go to api, text:api">API</a></li>
+      <li><a href="https://training.github.com" data-ga-click="Footer, go to training, text:training">Training</a></li>
+      <li><a href="https://shop.github.com" data-ga-click="Footer, go to shop, text:shop">Shop</a></li>
+        <li><a href="https://github.com/blog" data-ga-click="Footer, go to blog, text:blog">Blog</a></li>
+        <li><a href="https://github.com/about" data-ga-click="Footer, go to about, text:about">About</a></li>
 
-\paragraph{Time}
+    </ul>
 
-The worst-case complexity of SAT is exponential, whereas the standard
-implementations of CG are polynomial, but with advances in SAT solving
-techniques, the performance in the average case is much more feasible than in the previous works done in 90s--00s.
-We used the open-source SAT solver MiniSat \cite{een04sat}.
+    <a href="https://github.com" aria-label="Homepage">
+      <span class="mega-octicon octicon-mark-github" title="GitHub"></span>
+</a>
+    <ul class="site-footer-links">
+      <li>&copy; 2015 <span title="0.05098s from github-fe117-cp1-prd.iad.github.net">GitHub</span>, Inc.</li>
+        <li><a href="https://github.com/site/terms" data-ga-click="Footer, go to terms, text:terms">Terms</a></li>
+        <li><a href="https://github.com/site/privacy" data-ga-click="Footer, go to privacy, text:privacy">Privacy</a></li>
+        <li><a href="https://github.com/security" data-ga-click="Footer, go to security, text:security">Security</a></li>
+        <li><a href="https://github.com/contact" data-ga-click="Footer, go to contact, text:contact">Contact</a></li>
+    </ul>
+  </div>
+</div>
 
-Table~\ref{table:time} shows the execution time compared to VISL CG-3. 
-The number of rules in the grammar is more significant than the word
-count; this holds also for the performance of VISL CG-3.
-From the SAT solving side, maximisation is the most costly operation.
-Parsing Don Quijote (384,155 words) with the rule set of 261 rules,
-the function was called 147,253 times, and with 19 rules, 132,255 times.
-%However, with both rule sets, half of the sentences in Don Quijote
-%needed less than 6 calls of maximise, and 75 \% of the sentences
-%needed less than 14.
-The difference in the execution times suggests that there are other
-reasons for the worse performance---this is to be expected,
-SATCG is just a first proof-of-concept implementation with no optimisations.
+    <div class="fullscreen-overlay js-fullscreen-overlay" id="fullscreen_overlay">
+  <div class="fullscreen-container js-suggester-container">
+    <div class="textarea-wrap">
+      <textarea name="fullscreen-contents" id="fullscreen-contents" class="fullscreen-contents js-fullscreen-contents" placeholder=""></textarea>
+      <div class="suggester-container">
+        <div class="suggester fullscreen-suggester js-suggester js-navigation-container"></div>
+      </div>
+    </div>
+  </div>
+  <div class="fullscreen-sidebar">
+    <a href="#" class="exit-fullscreen js-exit-fullscreen tooltipped tooltipped-w" aria-label="Exit Zen Mode">
+      <span class="mega-octicon octicon-screen-normal"></span>
+    </a>
+    <a href="#" class="theme-switcher js-theme-switcher tooltipped tooltipped-w"
+      aria-label="Switch themes">
+      <span class="octicon octicon-color-mode"></span>
+    </a>
+  </div>
+</div>
 
-\begin{table}
-  \centering
-  \begin{tabular}{|l|c|c|c|}
-     \hline
-   \textbf{\# words} & \textbf{\# rules} &  \textbf{SATCG} & \textbf{VISL CG-3} \\ \hline
-                384,155  & 19   &  13.7s   & 4.2s\\ %\hline
-                124,493  & 99   &  55.7s   & 4.8s \\ % \hline
-	        384,155  & 130  &  1m14.1s & 6.1s \\ %\hline
-                384,155  & 261  &  2m53.7s & 10.7s \\ \hline
-  \end{tabular}
-  \caption{Execution times.}
-  \label{table:time}
-\end{table}
+    <div id="ajax-error-message" class="flash flash-error">
+      <span class="octicon octicon-alert"></span>
+      <a href="#" class="octicon octicon-x flash-close js-ajax-error-dismiss" aria-label="Dismiss error"></a>
+      Something went wrong with that request. Please try again.
+    </div>
 
-\paragraph{Performance against VISL CG-3}
+      <script crossorigin="anonymous" src="https://assets-cdn.github.com/assets/frameworks-d22b59d0085e83b7549ba4341ec9e68f80c2f29c8e49213ee182003dc8d568c6.js"></script>
+      <script async="async" crossorigin="anonymous" src="https://assets-cdn.github.com/assets/github-2c46bf7af744168851951e2fa4f404675a09c3d3f5b8ec92e4d28c3238de7936.js"></script>
 
-We took a tagged corpus\footnote{https://svn.code.sf.net/p/apertium/svn/branches/apertium-swpost/apertium-en-es/es-tagger-data/es.tagged} of 21865 words for Spanish, 
-and a small constraint grammar\footnote{https://svn.code.sf.net/p/apertium/svn/languages/apertium-spa/apertium-spa.spa.rlx}, produced independently of the authors.
-% We ignored substitute rules, but kept subreadings and unification, even though our implementation doesn't handle them.
-We only took select and remove rules (no substitute/iff), after which the constraint grammar had 261 rules.
-With this setup, we took the text of the tagged corpus with all ambiguities, and ran both VISL CG-3 and our implementation with the same grammar. 
-Treating the chosen tagset as the gold standard, the disambiguation by
-SATCG achieves 75.09 \% correct result and VISL CG-3 75.80 \%.
-Similar patterns were observed with small (\textless{}20) rule sets written by
-the authors; depending on the subset, SATCG and VISL CG-3 had
-a difference of at most $\Mypm$ 1.5 \%. 
-Introducing rules one by one up to 19, the
-performance improved in a very similar rate, with less than 0.5 \%
-difference between the systems at each new rule.
-
-% Additional tests could include plugging SATCG into Apertium
-% translation pipeline, and comparing the translation quality.
-
-% \paragraph{Removing rules}
-% \begin{table}
-%   \centering
-%   \begin{tabular}{|l|c|c|}
-%      \hline
-%     \textbf{\# rules} & \textbf{SATCG} & \textbf{VISL CG-3} \\
-%     \hline 
-% 	19 &  83.15 & 83.33 \\ \hline  % 19 out of 23 preselected by authors
-% 	18 &  83.19 & 83.04 \\ \hline  % 18 out of 19
-% 	17 &  83.201 & 83.03 \\ \hline   % 17 out of 18/19, no diff
-% 	16 &  83.197 & 83.030 \\ \hline  % 16 out of 17/18, no diff
-% 	15 &  83.188 & 83.022\\ \hline  % 15 out of 16/17. Tie with 14 for SATCG, chose to retain the rule that gave better results for VISL CG-3.
-% 	14 &  83.180 & 83.005 \\ \hline  % Tie with 15 for SATCG.
-% 	13 &  83.159 & 82.951 \\ \hline  % 13 out of 15 
-% 	12 &  83.134 & 82.884\\ \hline  % 12 out of 14
-% 	11 &  83.089 & 82.847\\ \hline  
-% 	10 &  83.051 & 82.802\\ \hline  
-% 	9  & 82.997 & 82.802 \\ \hline
-% 	8  & 82.906 & 82.681 \\ \hline  % REMOVE:r_imp Imp IF (0 Subj) (-1* CnjSub) ; 
-% 	7  & 82.743 & 82.594 \\ \hline  % SELECT:s_pro PrnIndep IF (1C VerbFin) ;
-% 	6  & 82.556 & 82.398\\ \hline  %  REMOVE:r_v_v Verb IF (-1C Vblex) ;
-% 	5  & 82.327 & 82.028 \\ \hline % SELECT:s_adv Adv IF (1 Cnj_Rel_End OR Prep OR Det)
-% 	4  & 82.065 & 82.028 \\ \hline % REMOVE:r_vfin_prep VerbFin IF (0 Prep) (1 Inf) 
-% 	3  & 81.387 & 81.233 \\ \hline % REMOVE:r_noun N    IF (-1 N OR Prn)
-% 	2  & 80.23 & 79.97 \\ \hline % SELECT:s_pri_1 (pri p3) IF (0 (vblex imp p2 sg))
-% 	1  & 76.69 & 76.62 \\ \hline 
-% 	0  & 70.70 & 70.70 \\ \hline
-%   \end{tabular}
-%   \caption{Result of removing rules one by one, and how close to the gold standard they get.}
-%   \label{table:results_remove}
-% \end{table}
-% In this experiment, we took 19 rules written by the authors, which give 83.15 \% correct results with SATCG and 83.33 \% with VISL CG-3.
-% Then we removed rules one by one; after getting a score for $n$ rules, we took the combination of $n-1$ rules for which SATCG gives the best score, and we ran the rule set with VISL CG-3 for comparison.
-Table~\ref{table:results_remove} shows the percentages compared to the gold standard for both systems.
-
-% \section*{Acknowledgments}
-
-% Do not number the acknowledgment section. Do not include this section
-% when submitting your paper for review.
-
-\bibliographystyle{acl}
-\bibliography{cg}
-
-\end{document}
+  </body>
+</html>

### 2015-03-31 paper/nodalida2015.tex

--- 2015-03-30 paper/nodalida2015.tex
@@
 
-\title{Constraint Grammar as a SAT problem}
+\title{Constraint Grammars as a SAT problem}
 
@@
 
-We represent Constraint Grammar \cite{karlsson1995constraint} 
+We represent Constraint Grammars (CG) \cite{karlsson1995constraint} 
 as a Boolean satisfiability (SAT) problem.
@@
 
-SAT solver performs a search, 
+The SAT solver performs a search, 
 and starts building possible models that satisfy both constraints. 
@@
 
-\noindent These three features might change the way that rules are written. 
+\noindent These three features influence the way that rules are written. 
 We predict that less rules are needed; whether this holds in
@@
 
-However, the implementation should not be too different from the
+However, the results using our approach should not be too different from the
 existing ones---our preliminary evaluation against VISL CG-3 shows
@@
 implementations of CG are polynomial, but with advances in SAT solving
-techniques, the performance in the average case is much more feasible than in the previous works done in 90s--00s.
+techniques, the performance in the average case in practice is much more feasible than in the previous works done in 90s--00s.
 We used the open-source SAT solver MiniSat \cite{een04sat}.
@@
 
+\section{Conclusions}
+
+SAT-solvers are nowadays powerful enough to be used for dealing with Constraint Grammars.
+A logic-based approach to CG has possible advantages over more traditional approaches; a SAT-solver may disambiguate more words, and may do so more precisely.
+Also, the SAT-solver requires less rules, and these rules are simpler.
+The ordering of rules is something we found was incompatible with a logic-based approach.
+We compensate this by maximising rule applications.
+Whether this is an advantage or disadvantage remains to be seen.
+Our initial experimental results are promising.
+
 % \section*{Acknowledgments}

### 2015-03-31 CG_SAT.hs
add named rules

@@ -77,16 +77,16 @@ anchor toks = (map.map) getBit (groupBy sameInd toks)
 applyRule :: Rule -> [Token] -> [[Bit]]
 applyRule rule toks = --trace (show rule) $
   case rule of
-    (Remove tags conds) -> applyRules rule (toLists conds) toks
-    (Select tags conds) -> applyRules rule (toLists conds) toks
+    (Remove _name tags conds) -> applyRules rule (toLists conds) toks
+    (Select _name tags conds) -> applyRules rule (toLists conds) toks
 
 
 applyRules :: Rule -> [[Condition]] -> [Token] -> [[Bit]]
 applyRules rule []         allToks = []
 applyRules rule (conds:cs) allToks = applyRules rule cs allToks ++
   case rule of 
-    (Remove tags _c) -> mkVars (chosen tags) nt 
-    (Select tags _c) -> mkVars (chosen tags) id ++ mkVars (other tags) nt
+    (Remove _n tags _c) -> mkVars (chosen tags) nt 
+    (Select _n tags _c) -> mkVars (chosen tags) id ++ mkVars (other tags) nt
 
 
   where
@@ -216,7 +216,7 @@ disambiguate verbose rules sentence = do
                               | (b, (rl,cl)) <- zip cs applied
                               , b /= Just True ]
                 putStrLn "These clauses were omitted due to conflicts:"
-                mapM_ putStrLn (take 2 conf) 
+                mapM_ putStrLn (conf) 
                 putStrLn $ "+ " ++ show ((length conf) - 2) ++ " others"
 
               let nonconf = [ c | (Just True, c) <- zip cs bitsForClauses ]

### 2015-03-31 CG_base.hs

@@ -47,8 +47,15 @@ type Sentence = [Analysis]
 
 
 -- | Rule is either remove or select a list of tags, with contextual tests
-data Rule = Remove TagSet Condition | Select TagSet Condition deriving (Show)
+data Rule = Remove Name TagSet Condition | Select Name TagSet Condition
+type Name = String
 
+instance Show Rule where
+  show (Remove name tags cond) = "REMOVE:" ++ name ++ " " ++
+                                  show tags ++ " " ++ show cond 
+  show (Select name tags cond) = "SELECT:" ++ name ++ " " ++
+                                  show tags ++ " " ++ show cond 
+  
 
 -- | There is no special constructor for empty condition (ie. remove/select tag everywhere),
 --   but `C _ (_,[])' is assumed to mean that.
@@ -131,8 +138,8 @@ andTest = AND lemmaBear always
 
 hasBoundary :: Rule -> Bool
 hasBoundary rule = case rule of
-  (Select _t c) -> findBoundary c
-  (Remove _t c) -> findBoundary c
+  (Select _n _t c) -> findBoundary c
+  (Remove _n _t c) -> findBoundary c
   where findBoundary c = or $ map hasB (concat (toLists c))
         hasB (C _pos (_b,tags)) = (not.null) $ [BOS,EOS] `intersect` concat tags
 
@@ -148,23 +155,23 @@ pl   = [[Tag "pl"]]
 cnjcoo  = [[Tag "cnjcoo"]]
 
 -- Rules
-rmParticle = Remove [[Tag "particle"]] always
-slVerbAlways = Select verb  always
-slNounIfBear = Select [[Lem "bear", n] | n <- concat noun]  always
+rmParticle = Remove "r_part" [[Tag "particle"]] always
+slVerbAlways = Select "s_verb" verb  always
+slNounIfBear = Select "s_bear_n" [[Lem "bear", n] | n <- concat noun]  always
 
-rmVerbIfDet = Remove verb (mkC "-1" det)
-rmAdvIfDet = Remove adv (mkC "1" det)
-rmNounIfPron = Remove noun (mkC "-1" [[Tag "pron"]])
-slPrepIfDet = Select prep (mkC "1" det)
-slNounAfterConj = Select noun (mkC "-1" conj)
[…]

### 2015-03-31 CG_parse.hs

@@ -170,19 +170,27 @@ transTagSet ts = case ts of
 
 transRule :: Rule -> State Env CGB.Rule
 transRule rl = case rl of
-  SelectIf _sl tags _if conds -> liftM2 CGB.Select (transTagSet tags) (transCondSet conds)
-  RemoveIf _sl tags _if conds -> liftM2 CGB.Remove (transTagSet tags) (transCondSet conds)
-  SelectAlways _sl tags   -> liftM2 CGB.Select (transTagSet tags) (return CGB.always)
-  RemoveAlways _sl tags   -> liftM2 CGB.Remove (transTagSet tags) (return CGB.always)
-  MatchLemma (Str lem) rl -> do cgrule <- transRule rl
-                                case cgrule of
-                                  CGB.Select ts c -> return $ CGB.Select (cart ts lem) c
-                                  CGB.Remove ts c -> return $ CGB.Remove (cart ts lem) c
+  SelectIf (SELECT_1 nm) tags _if conds ->
+    liftM2 (CGB.Select (getName nm)) (transTagSet tags) (transCondSet conds)
+  RemoveIf (REMOVE_1 nm) tags _if conds ->
+    liftM2 (CGB.Remove (getName nm)) (transTagSet tags) (transCondSet conds)
+  SelectAlways (SELECT_1 nm) tags ->
+    liftM2 (CGB.Select (getName nm)) (transTagSet tags) (return CGB.always)
+  RemoveAlways (REMOVE_1 nm) tags ->
+    liftM2 (CGB.Remove (getName nm)) (transTagSet tags) (return CGB.always)
+  MatchLemma (Str lem) rl -> 
+    do cgrule <- transRule rl
+       case cgrule of
+         CGB.Select n ts c -> return $ CGB.Select n (cart ts lem) c
+         CGB.Remove n ts c -> return $ CGB.Remove n (cart ts lem) c
   where cart :: CGB.TagSet -> String -> CGB.TagSet
         cart ts str = case str of
            ('"':'<':_) -> [[CGB.WF  (strip 2 str), t] | t <- concat ts]
            ('"':_    ) -> [[CGB.Lem (strip 1 str), t] | t <- concat ts]
            _           -> [[CGB.Lem str          , t] | t <- concat ts]
+
+        getName (MaybeName_1 (Id id)) = id
+        getName MaybeName_2           = ""
         
 
 transCondSet :: [Cond] -> State Env CGB.Condition

### 2015-03-31 SAT/SAT.hs

@@ -106,7 +106,7 @@ maximize s as rs =
        do return False
       else
        do a <- newBit s
-          let opti xs =
+          let opti n xs =
                 do bs <- sequence [ modelValueBit s x | x <- xs ]
                    sequence_
                      [ addClauseBit s [nt a, x]
@@ -117,10 +117,11 @@ maximize s as rs =
                    addClauseBit s (nt a : xs') 
                    b <- solveBit s (a:as)
                    if not b then
-                     do return True
+                     do putStr $ show n ++ ", "
+                        return True
                     else
-                     do opti xs'
-           in opti rs
+                     do opti (n+1) xs'
+           in opti 1 rs
 
 --------------------------------------------------------------------------------
 

### 2015-03-31 test/Test.hs

@@ -75,7 +75,6 @@ loop (r:rs) t g scores = do
                    , let dif = diffBySent sat gold
                    , (not.null) dif ]
       orig = fromIntegral $ length (concat t)
-     -- orig = fromIntegral $ wordCount t
       difbw = fromIntegral $ length (concat diff)
       perc = 100 * ((orig-difbw) / orig) :: Float
   when ((length scores) `mod` 100 == 0) $ print (length scores) 

### 2015-04-30 paper/nodalida2015.tex
add reference

@@
 % ; logic programming was too slow to be used for tagging or parsing. 
-Since those works, SAT solving techniques have improved significantly, and 
+Since those works, SAT solving techniques have improved significantly \cite{marques_silva2010}, and 
 they are used in domains such as microprocessor design and computational 

### 2015-04-30 CG_SAT.hs
bunch of new stuff.

@@ -231,7 +231,7 @@ disambiguate verbose debug rules sentence = do
         putStrLn "\ntokens:"
         mapM_ print toks
         putStrLn "\nfirst step, make sure all readings for a given word are not false:"
-        mapM_ print allNotFalse
+        print allNotFalse
         --putStrLn "\nclauses gotten by applying rules"
         --mapM_ print applied
 
@@ -307,6 +307,9 @@ disambiguateWithOrder verbose debug rules sentence = do
       --mapM_ print applied
       let onlyClauses = (map.map) snd $ groupBy (\x y -> fst x==fst y) applied :: [[[Lit]]]
 
+      sequence_ [ addClause s cls | cls <- concat onlyClauses ]
+
+
       when debug $ do
         putStrLn "\ntokens:"
         mapM_ print toks
@@ -318,34 +321,47 @@ disambiguateWithOrder verbose debug rules sentence = do
       bs <- sequence [ do k <- count s insts :: IO Unary
                           b <- solveMaximize s [] k
                           is <- sequence [ modelValue s x | x <- insts ] 
-                          sequence_ [ addClause s [cl] | (True, cl) <- zip is insts ]
+                          --putStrLn "\nmaximising"
+                          --putStrLn "true:"
+                          --mapM_ putStrLn [ show lit | (True, lit) <- zip is insts ]
+                          --putStrLn "\nfalse:"
+                          --mapM_ putStrLn [ show lit | (False, lit) <- zip is insts ]
+
+                          sequence_ [ addClause s [lit] -- >> addClause s [c]
+                                       | (True, lit, c) <- zip3 is insts cs]
                           return b
                   | cls <- onlyClauses
-                  , let insts = map (neg . head) cls ]
+                  , let insts = map (neg . head) cls
+                  , let cs = map (!! 1) cls ]
       when debug $ print bs
+
 
       when debug $ do 
        cs <- sequence [ modelValue s x | x <- litsForInsts ]
        putStrLn "\nUsed clauses:"
-       -- mapM_ putStrLn [ show rl ++ "\n* " ++ show cl 
-       --                    | (rl,cls) <- applied 
-       --                    , (True,cl) <- zip cs (concatMap snd applied) ] 
+       mapM_ putStrLn [ show rl ++ "\n* " ++ show cl 
+                          | (True, (rl,cl)) <- zip cs applied 
+                      ]
                          
-       -- putStrLn "\nUnused clauses:"
[…]

### 2015-04-30 test/Test.hs

@@ -9,52 +9,65 @@ import Data.Maybe
 import System.Environment
 import System.IO
 import System.Process
+import System.Random
+import System.Random.Shuffle
 import Text.Printf
 
-ambiguous = "data/es.tagged.ambiguous.bak"
+ambiguous = "data/es.tagged.ambiguous"
 apertium = "data/apertium-spa.spa.rlx"
 small = "data/spa_smallset.rlx"
-golddata = "data/es.tagged"
+esgold = "data/es.tagged"
+espre = "data/spa_pre.rlx"
 rusgr = undefined
 rustext = undefined
+engr = "data/eng_cg2.rlx"
+entext = "data/en.tagged.ambiguous"
+engold = "data/en.tagged"
+enpre = "data/en_pre.rlx"
 
 main :: IO ()
 main = do
   args <- getArgs
   case args of
-    ["shuffle"]     -> shuffle small ambiguous
+    ["shuffle"]     -> shuffle_ small ambiguous esgold espre
+    [r,"shuffle"]   -> shuffle_ r ambiguous esgold espre
+    ["enshuffle"]   -> shuffle_ engr entext engold enpre
+    ["reverse"]     -> rev small ambiguous esgold espre
+    [r,"reverse"]   -> rev r ambiguous esgold espre
+    ["enreverse"]   -> rev engr entext engold enpre
     ["opticomp",n]  -> optiComp small ambiguous (read n)
-    ["gold","orig"] -> gold apertium ambiguous
-    ["gold","opti"] -> opti small ambiguous
+    ["gold","orig"] -> gold apertium ambiguous esgold
+    ["gold","opti"] -> opti small ambiguous 
     ["gold","obs"]  -> optiBySz small ambiguous
-    ["gold"]        -> gold small ambiguous --using small as default grammar
-    ["goldrus"]     -> gold rusgr rustext
-
-    [r,"gold"]      -> gold r ambiguous --specify grammar
+    ["gold"]        -> gold small ambiguous esgold --using small as default grammar
+    ["goldrus"]     -> gold rusgr rustext undefined
+    ["engold"]     -> gold engr entext engold
+    [r,"gold"]      -> gold r ambiguous esgold --specify grammar
 
     (r:d:o) -> do rules <- readRules r
                   text <- readData d
                   let is2 = "2" `elem` o
                       verbose = "v" `elem` o
                       debug = "d" `elem` o
[…]

### 2015-05-07 CG_SAT.hs
add nice prettyprinting function to rules and make the code a mess.

@@ -95,11 +95,11 @@ anchor toks = (map.map) getLit (groupBy sameInd toks)
 applyRule :: Rule -> [Token] -> [[Lit]]
 applyRule rule toks = --trace (show rule) $
   case rule of 
-    (Remove _name tags conds) -> go False tags (toLists conds) toks
-    (Select _name tags conds) -> go True tags (toLists conds) toks
+    (Remove _name tags conds) -> go False (toTags tags) (toConds conds) toks
+    (Select _name tags conds) -> go True (toTags tags) (toConds conds) toks
 
     
-go :: Bool -> TagSet -> [[Condition]] -> [Token] -> [[Lit]]
+go :: Bool -> [[Tag]] -> [[Condition]] -> [Token] -> [[Lit]]
 go isSelect tags []         allToks = []
 go isSelect tags (conds:cs) allToks = --trace (show conds) $
   if isSelect 
@@ -125,14 +125,14 @@ go isSelect tags (conds:cs) allToks = --trace (show conds) $
      -- because that would require n pl and n sg be true at the same time.
      -- Instead we make combinations [(1,det), (3,n sg)] and [(1,det), (3,n pl)]
 
-        remove :: TagSet -> [(Token,[[Token]])]
+        remove :: [[Tag]] -> [(Token,[[Token]])]
         remove tags = [ (tok, ctx) | tok <- allToks
                                    , isAmbig tok allToks --only apply rules to ambiguous tokens
                                    , tagsMatchRule tags tok
                                    , let ctx = getContext tok allToks conds
                                    , allCondsHold ctx]
 
-        select :: TagSet -> ([(Token,[[Token]])], [(Token,[[Token]])])
+        select :: [[Tag]] -> ([(Token,[[Token]])], [(Token,[[Token]])])
         select tags = (chosen, other)
             -- chosen: analyses that have the wanted readings and context
  
@@ -158,11 +158,11 @@ getContext :: Token           -- ^ a single analysis
                -> [[Token]]   -- ^ context for the first arg. If all conditions match for a token, there will be as many non-empty Token lists as Conditions.
 getContext tok allToks []     = []
 getContext tok allToks ((C position (bool,ctags)):cs) = getContext tok allToks cs ++
-  case ctags of
+  case ctags' of
     []     -> [[dummyTok]] --empty conds = holds always
     [[]]   -> [[dummyTok]] 
     (t:ts) -> case position of
[…]

### 2015-05-07 CG_base.hs

@@ -3,6 +3,7 @@ module CG_base where
 import Control.Applicative
 import Data.List
 
+
 -- | All kinds of morphological tags are in the same data type: e.g.  Prep, P1, Conditional.
 -- | We don't specify e.g. which tags can be part of an analysis for which word classes.
 -- | An analysis can contain an arbitrary amount of tags.
@@ -27,8 +28,33 @@ instance Show Tag where
   show (Lem str) = "\"" ++ str ++ "\""
   show (Tag str) = str
   show BOS       = "" --">>>"
-  show EOS       = "" -- "<<<"
+  show EOS       = "" --"<<<"
 
+
+
+data TagSet =
+   TS [[Tag]]
+ | Or TagSet TagSet
+ | Diff TagSet TagSet
+ | Cart TagSet TagSet
+ | All
+  deriving (Eq,Ord,Read)
+
+instance Show TagSet where
+  show (TS tags) = showTagset tags
+  show (Or (TS ts1) ts2) = showTagset ts1 ++ " OR " ++ show ts2
+  show (Diff (TS ts1) ts2) = showTagset ts1 ++ " - " ++ show ts2
+  show (Cart (TS ts1) ts2) = showTagset ts1 ++ " + " ++ show ts2
+  show All = "(*)"
+  show x = "TODO"
+
+showTagset :: [[Tag]] -> String
+showTagset [[x]] = show x
+showTagset xs    = concatMap show' xs
+  where show' [y] = show y
+        show' ys  = "(" ++ unwords (map show ys) ++ ")"
+
+toTags :: TagSet -> [[Tag]]
 -- | TagSet translates to [[Tag]] : outer list is bound by OR, inner lists by AND
 --  For example, 
 --    LIST DefArt = (det def) ;
@@ -37,8 +63,13 @@ instance Show Tag where
 --  translate into
 --    defArt = [[Tag "det", Tag "def"]]
 --    dem    = [[Lem "az"],[Lem "ez"],[Lem "amaz"],[Lem "emez"]]
---  See the datatype for Test.
-type TagSet = [[Tag]]
+
+toTags (TS tags) = tags
+toTags (Or ts1 ts2) = toTags ts1 ++ toTags ts2
+toTags (Diff ts1 ts2) = toTags ts1 \\ toTags ts2 
+toTags (Cart ts1 ts2) = map concat $ sequence [(toTags ts1), (toTags ts2)]
+toTags All = [[]] --matches all
[…]

### 2015-05-07 CG_parse.hs

@@ -71,8 +71,8 @@ main = do args <- getArgs
 parseCGRules :: Grammar -> State Env [Either String CGB.Rule]
 parseCGRules (Defs defs) = do mapM updateEnv defs 
                               --in case the grammar doesn't specify boundaries 
-                              modify ((">>>",bos) :)
-                              modify (("<<<",eos) :)
+                              modify ((">>>", CGB.TS bos) :)
+                              modify (("<<<", CGB.TS eos) :)
                               env <- get
                               return $ map (parseRules' env) defs
   where updateEnv :: Def -> State Env ()
@@ -100,18 +100,19 @@ strip n = drop n . reverse . drop n . reverse
 transSetDecl :: SetDecl -> State Env (String, CGB.TagSet)
 transSetDecl (Set setname tagset) = 
   case setname of
-    BOS          -> return (">>>", bos)
-    EOS          -> return ("<<<", eos)
+    BOS          -> return (">>>", CGB.TS bos)
+    EOS          -> return ("<<<", CGB.TS eos)
     (SetName (UIdent name)) -> do 
       ts <- transTagSet tagset
       return (name, ts)
 transSetDecl (List setname tags) = 
   case setname of
-    BOS          -> return (">>>", bos)
-    EOS          -> return ("<<<", eos)
+    BOS          -> return (">>>", CGB.TS bos)
+    EOS          -> return ("<<<", CGB.TS eos)
     (SetName (UIdent name)) -> do 
       tl <- mapM transTag tags
-      return (name, concat tl)
+      let tl' = concatMap CGB.toTags tl
+      return (name, CGB.TS tl')
 
 
                                       
@@ -119,13 +120,14 @@ transSetDecl (List setname tags) =
 transTag :: Tag -> State Env CGB.TagSet
 transTag tag = case tag of
   Lemma (Str s) -> case s of
-                   ('"':'<':_) -> return [[CGB.WF (strip 2 s)]]
-                   ('"':    _) -> return [[CGB.Lem (strip 1 s)]]
-                   _           -> return [[CGB.Lem s]]
-  Tag (Id str) -> return [[CGB.Tag str]]
+                   ('"':'<':_) -> return $ CGB.TS [[CGB.WF (strip 2 s)]]
+                   ('"':    _) -> return $ CGB.TS [[CGB.Lem (strip 1 s)]]
+                   _           -> return $ CGB.TS [[CGB.Lem s]]
+  Tag (Id str) -> return $ CGB.TS [[CGB.Tag str]]
[…]

### 2015-05-07 paper/nodalida2015.tex
bunch of stuff

@@
 
-\title{Constraint Grammars as a SAT problem}
+\title{Constraint Grammar as a SAT problem}
 
-% \author{Inari Listenmaa \and Koen Claessen \\
-%  Chalmers University of Technology, Gothenburg, Sweden \\
-%  {\tt \{inari,koen\}@chalmers.se} }
+\author{Inari Listenmaa \and Koen Claessen \\
+ Chalmers University of Technology, Gothenburg, Sweden \\
+ {\tt \{inari,koen\}@chalmers.se} }
 
@@
 
-We represent Constraint Grammars (CG) \cite{karlsson1995constraint} 
+We represent Constraint Grammar (CG) \cite{karlsson1995constraint} 
 as a Boolean satisfiability (SAT) problem.
@@
    \textbf{\# words} & \textbf{\# rules} &  \textbf{SATCG} & \textbf{VISL CG-3} \\ \hline
-                384,155  & 19   &  13.7s   & 4.2s\\ %\hline
-                124,493  & 99   &  55.7s   & 4.8s \\ % \hline
-	        384,155  & 130  &  1m14.1s & 6.1s \\ %\hline
-                384,155  & 261  &  2m53.7s & 10.7s \\ \hline
+                384,155  & 19   &  23.1s   & 4.2s\\ %\hline
+	        384,155  & 99   &  1m40.6s & 6.1s \\ %\hline
+                384,155  & 261  &  2m36.7s & 10.7s \\ \hline
   \end{tabular}
@@
 
-SAT-solvers are nowadays powerful enough to be used for dealing with Constraint Grammars.
+SAT-solvers are nowadays powerful enough to be used for dealing with Constraint Grammar.
 A logic-based approach to CG has possible advantages over more traditional approaches; a SAT-solver may disambiguate more words, and may do so more precisely.

### 2015-05-07 executable/Main.hs

@@ -12,25 +12,28 @@ main = do
   args <- getArgs
   case args of
     ["v", f1]  -> do rules <- readRules f1
-                     loop (disambiguate False False rules)
+                     loop False (disambiguateWithOrder False False rules)
+    ["v", "d", f1]  -> do rules <- readRules f1
+                          loop True (disambiguateWithOrder True True rules)
     [f1, f2]   -> do rules <- readRules f1
                      data' <- readData f2 
                      mapM_ (disambiguate False False rules) data'
-    [f1, f2, "v"]   -> do rules <- readRules f1
-                          data' <- readData f2 
-                          mapM_ (disambiguate True True rules) data'
-    [f1, f2, "o"]   -> do rules <- readRules f1
-                          data' <- readData f2 
-                          mapM_ (disambiguateWithOrder True True rules) data'
-
+    (r:d:o) -> do rules <- readRules r
+                  text <- readData d
+                  let is2 = "2" `elem` o
+                      verbose = "v" `elem` o
+                      debug = "d" `elem` o
+                      disam = if "no" `elem` o then disambiguate
+                                               else disambiguateWithOrder
+                  mapM_ (disam verbose debug rules) text
     ("test":_) -> CG_SAT.test
     _          -> putStrLn "usage: ./Main (<rules> <data> | test) [v]"
 
 
 
-loop :: (Sentence -> IO Sentence) -> IO ()
-loop f = do
-    s <- parseData `fmap` getLine
+loop :: Bool -> (Sentence -> IO Sentence) -> IO ()
+loop debug f = do
+    s <- parseData `fmap` getLine    
     disam <- mapM f s
-    mapM_ (putStrLn . showSentence) disam
-    loop f
+    when (not debug) $ mapM_ (putStrLn . showSentence) disam --if debug, it will show already, no need to print it again
+    loop debug f

### 2015-05-07 test/Test.hs

@@ -23,6 +23,8 @@ rustext = undefined
 engr = "data/eng_cg2.rlx"
 entext = "data/en.tagged.ambiguous"
 engold = "data/en.tagged"
+-- entext = "data/vietnam.tagged.ambiguous"
+-- engold = "data/vietnam.tagged"
 enpre = "data/en_pre.rlx"
 
 main :: IO ()
@@ -49,7 +51,8 @@ main = do
                   let is2 = "2" `elem` o
                       verbose = "v" `elem` o
                       debug = "d" `elem` o
-                      disam = if "o" `elem` o then disambiguateWithOrder                                                         else disambiguate
+                      disam = if "no" `elem` o then disambiguate
+                                               else disambiguateWithOrder
                   resSAT <- mapM (disam verbose debug rules) text -- :: [Sentence]
                   resVISL <- vislcg3 r d is2  -- :: [Sentence] 
                   prAll "" resSAT resVISL text verbose

### 2015-05-18 paper/nodalida2015.tex
slightly more tofu

@@
 % ; logic programming was too slow to be used for tagging or parsing. 
-Since those works, SAT solving techniques have improved significantly \cite{marques_silva2010}, and 
-they are used in domains such as microprocessor design and computational 
+Since those works, SAT solving techniques have improved significantly \cite{marques_silva2010}, and they are used in domains such as microprocessor design and computational 
 biology---these problems easily match or exceed CG in complexity. 
-% Is this true for grammars with thousands of rules?
+Due to these advances, we were able to revisit the idea and develop it
+further.
 
-Due to these advances, we were able to revisit the idea and develop it further. 
-% Whereas \cite{lager98} provides a proof-of-concept description, 
-Missing from the previous logic-based works, 
-we solve eventual rule conflicts by finding a solution that 
-discards the least number of rules.
-% allows the maximal number of rules to be applied.
-We test our implementation by parsing texts in the order of 10,000s--100,000s
-words, using grammars with hundreds of rules.
+Encoding CG in logic increases computational complexity, 
+but as a tradeoff, it could simplify the grammar writing.
+Due to representing the rules as implications, the requirement of being unambiguously tagged becomes irrelevant.
+One rule corresponds to many logical formulas (each condition being
+the target), making the number of the rules potentially smaller.
+Ordering can be preserved or discarded; in the latter case, we solve
+eventual rule conflicts by finding a solution that discards the least
+number of rule applications.
+% our experiments show that it is possible to write a small grammar, 
+% which, when run with SAT-CG, disambiguates more than the same
+%grammar run with VISL CG-3.
+
+We test our implementation by parsing texts in the order of
+10,000s--100,000s words, using grammars with hundreds of rules.
 
@@
 
-%This increases the computational complexity, 
-%but we are exploring ways in which it could simplify the grammar writing:
-%our experiments show that less rules are needed, and features like ordering of the rules
-%and requirement of being unambiguously tagged become irrelevant.
+% Uncomment for full paper
+\section{Related work}
+\label{sect:related}
 
-% Uncomment for full paper
-% \section{Related work}
-% \label{sect:related}
-
-% Our work is inspired by \cite{lager98}, which presents constraint rules as a disjunctive logic program,
-% and \cite{lager_nivre01}, which reconstructs 4 different formalisms from a logical point of view.
-% \cite{lindberg_eineborg98ilp,asfrent14} use Inductive Logic Programming to learn CG rules from a tagged corpus.
-% \cite{lager01transformation} discusses ordering of the CG rules.
+Our work is inspired by \cite{lager98}, which presents constraint rules as a disjunctive logic program,
+and \cite{lager_nivre01}, which reconstructs 4 different formalisms from a logical point of view.
+\cite{lindberg_eineborg98ilp,asfrent14} use Inductive Logic Programming to learn CG rules from a tagged corpus.
+\cite{lager01transformation} discusses ordering of the CG rules.
 
@@
   IF (-1 prn)}. A rule with more conditions can be written as many
-rules, each condition taking its turn to be in position to be removed
-or selected. \\
+rules, each condition taking its turn to be the target of removal or selection. \\
 
@@
 promising results in this regard.
-\pagebreak
 

### 2015-05-19 test/QCTest.hs
works with TagSets now

@@ -36,10 +36,14 @@ arbCondList =
 
 --------------------------------------------------------------------------------
 allTags :: [Tag]
-allTags = concat $ verb ++ noun ++ det ++ adv ++ conj ++ prep ++ sg ++ pl ++ cnjcoo
+allTags = map Tag ["vblex", "vbser", "vbmod", "n", "np", "det", "adv", "cnjcoo", "cnjsub", "prep", "sg", "pl", "cnjcoo"]
 
 instance Arbitrary Tag where
   arbitrary = elements allTags
+
+instance Arbitrary TagSet where
+  arbitrary = do tags <- listOf arbitrary
+                 return $ TS [tags]
 
 instance Arbitrary SAT.Lit where
   arbitrary = elements [SAT.Lit (MiniSat.MkLit n) | n <- [1..50]]
@@ -47,13 +51,13 @@ instance Arbitrary SAT.Lit where
 instance Arbitrary Position where
   arbitrary = elements $ [Exactly n | n <- [-5..5]] ++
                          [AtLeast n | n <- [-5..5]] ++
-                         [Barrier n [[t]] | n <- [-5..5],
-                                            t <- allTags]
+                         [Barrier n (TS [[t]]) | n <- [-5..5],
+                                                 t <- allTags]
 
 instance Arbitrary Condition where
   arbitrary = do pos <- arbitrary
                  bool <- arbitrary
-                 tags <- listOf arbitrary
+                 tags <- arbitrary
                  return $ C pos (bool, tags)
 
 

### 2015-05-19 paper/nodalida2015.tex
stuff

@@
 but as a tradeoff, it could simplify the grammar writing.
-Due to representing the rules as implications, the requirement of being unambiguously tagged becomes irrelevant.
+Due to representing the rules as implications, cautious context
+condition becomes irrelevant.
 One rule corresponds to many logical formulas (each condition being
-the target), making the number of the rules potentially smaller.
+the target), which makes the number of the rules potentially smaller.
 Ordering can be preserved or discarded; in the latter case, we solve

### 2015-05-19 CG_SAT.hs

@@ -200,10 +200,11 @@ getContext tok allToks ((C position (bool,ctags)):cs) = getContext tok allToks c
                  
 
 ---- Main stuff
-
-
--- how to emulate ordering:
--- Maximise all instances of applying rule one; then commit to those; then maximise instances of rule 2, and so on.
+disamSection ::  ([Rule] -> Sentence -> IO Sentence) -> [[Rule]] -> Sentence -> IO Sentence
+disamSection disam []         sent = return sent
+disamSection disam [rs]       sent = disam rs sent
+disamSection disam (r1:r2:rs) sent = disam r1 sent
+                                      >>= \s -> (print s >> disamSection disam ((r1++r2):rs) s)
 
 disambiguate :: Bool -> Bool -> [Rule] -> Sentence -> IO Sentence
 disambiguate verbose debug rules sentence = do

### 2015-05-19 CG_parse.hs

@@ -31,7 +31,11 @@ parseRules :: Bool -> String -> [[CGB.Rule]] -- sections
 parseRules test s = case pGrammar (CG.Par.myLexer s) of
             CGErr.Bad err  -> error err
             CGErr.Ok  tree -> let rules = evalState (parseCGRules tree) []
-                              in  map rights rules
+                              in trace (if test then unwords $ pr rules else "") $
+                                 map rights rules
+  where pr = concatMap $ map pr'
+        pr' (Right rule)  = show rule
+        pr' (Left string) = string
 
 parseData :: String -> [CGB.Sentence]
 parseData s = case pText (Apertium.Par.myLexer s) of
@@ -47,8 +51,8 @@ parseData s = case pText (Apertium.Par.myLexer s) of
 
 
 --just because it's nice to use them  rules <- readRules foo
-readRules :: String -> IO [CGB.Rule]
-readRules fname = readFile fname >>= return . concat . parseRules False
+readRules :: String -> IO [[CGB.Rule]]
+readRules fname = readFile fname >>= return . parseRules False
 
 readData :: String -> IO [CGB.Sentence]
 readData fname = readFile fname >>= return . parseData

### 2015-05-19 executable/Main.hs

@@ -2,7 +2,7 @@ module Main where
 
 import CG_parse (readRules, readData, parseData)
 import CG_SAT
-import CG_base ( Sentence, showSentence )
+import CG_base ( Sentence, showSentence, Rule )
 import Control.Monad
 import System.Environment
 
@@ -11,13 +11,10 @@ main :: IO ()
 main = do
   args <- getArgs
   case args of
-    ["v", f1]  -> do rules <- readRules f1
+    ["v", f1]  -> do rules <- concat `fmap` readRules f1
                      loop False (disambiguateWithOrder False False rules)
-    ["v", "d", f1]  -> do rules <- readRules f1
+    ["v", "d", f1]  -> do rules <- concat `fmap` readRules f1
                           loop True (disambiguateWithOrder True True rules)
-    [f1, f2]   -> do rules <- readRules f1
-                     data' <- readData f2 
-                     mapM_ (disambiguate False False rules) data'
     (r:d:o) -> do rules <- readRules r
                   text <- readData d
                   let is2 = "2" `elem` o
@@ -25,11 +22,10 @@ main = do
                       debug = "d" `elem` o
                       disam = if "no" `elem` o then disambiguate
                                                else disambiguateWithOrder
-                  mapM_ (disam verbose debug rules) text
+                  mapM_ (disamSection (disam verbose debug) rules) text
     ("test":_) -> CG_SAT.test
     _          -> putStrLn "usage: ./Main (<rules> <data> | test) [v]"
-
-
+  
 
 loop :: Bool -> (Sentence -> IO Sentence) -> IO ()
 loop debug f = do

### 2015-05-19 test/Test.hs

@@ -51,9 +51,10 @@ main = do
                   let is2 = "2" `elem` o
                       verbose = "v" `elem` o
                       debug = "d" `elem` o
-                      disam = if "no" `elem` o then disambiguate
-                                               else disambiguateWithOrder
-                  resSAT <- mapM (disam verbose debug rules) text -- :: [Sentence]
+                      disam = if "no" `elem` o 
+                                then disambiguate verbose debug
+                                else disambiguateWithOrder verbose debug
+                  resSAT <- mapM (disamSection disam rules) text -- :: [Sentence]
                   resVISL <- vislcg3 r d is2  -- :: [Sentence] 
                   prAll "" resSAT resVISL text verbose
                   putStrLn ""
@@ -63,7 +64,7 @@ gold rl dt g = do rules <- readRules rl
                   text <- readData dt
                   gold <- readData g
 
-                  resSAT <- mapM (disambiguateWithOrder False False rules) text
+                  resSAT <- mapM (disamSection (disambiguateWithOrder False False) rules) text
                   resVISL <- vislcg3 rl dt True
                   putStrLn "SAT-CG in comparison to gold standard"
                   let verbose = length text < 100 --change if you want different output
@@ -186,7 +187,7 @@ wordCount s = length $ map (filter (\x -> null $ [BOS,EOS] `intersect` x)) $ con
 
 --shuffle :: [Rule] -> IO ()
 shuffle_ r d g pre = do
-  rls <- readRules r
+  rls <- concat `fmap` readRules r
   seed <- newStdGen
   sequence_ [ do mkRuleFile ps fname r pre
                  putStrLn (fname ++ "\n---------") 
@@ -198,7 +199,7 @@ shuffle_ r d g pre = do
                            in shuffle' xs (length xs) seed : shuffles (reverse xs) newSeed
 
 rev r d g pre = do
-  rls <- readRules r
+  rls <- concat `fmap` readRules r
   sequence_ [ do mkRuleFile ps fname r pre
                  putStrLn (fname ++ "\n---------") 
                  gold fname d g
@@ -220,7 +221,7 @@ mkRuleFile rules fp orig pre = do
 
 --------------------------------------------------------------------------------
 
-optiBySz rl dt = do r <- readRules rl
+optiBySz rl dt = do r <- concat `fmap` readRules rl
                     t <- readData dt
                     g <- readData esgold
[…]

### 2015-05-22 executable/Symbolic.hs
conflict detection

@@ -1,0 +1,85 @@
+import CG_base
+import CG_SAT hiding (chunk)
+import CG_parse
+import Data.List
+import SAT
+import SAT.Optimize
+import SAT.Unary hiding (modelValue)
+n = 2
+
+--t = map Tag ["n", "vblex", "pri", "p3", "sg", "pl"]
+tags = map Tag ["b","c","d","e"]
+
+
+rules = parseRules False  "REMOVE:r1 (d) IF (-1C (c)) ;\nREMOVE:r2 (b) ;\nREMOVE:r3 (e) IF (-1C (c)) ;"
+
+sentence = parseData "^first/f<b>/f<c>$ ^second/s<d>/s<e>/s<f>$" 
+
+
+{- 
+all subsequences of t should be considered for the position of one reading
+
+symbolic must have at least one word with >1 reading, because of the requirement that applying R must make a difference
+
+First rule: we must find some input so that it applies, no other requirements.
+Second rule: we must find some input so that it applies, and first rule doesn't apply.
+
+-----
+
+First task: find such input that a rule will have effect, or prove there is none
+1) Target    of the rule must be in the analysis
+2) Condition of the rule must be in the context
+-}
+
+
+main = mapM_ doSomething (concat rules)
+
+
+doSomething :: Rule -> IO ()
+doSomething rule = do
+
+  putStrLn "---------------------"
+  s <- newSolver
+  t <- sequence [ newLit s | _ <- tags
+                           , _ <- [1..n] ] -- # of words in sentence
+
+  let chunkedSymbolicSent = chunk $ replicate n (map (:[]) tags)
+      symbolicToks = zip chunkedSymbolicSent t :: [Token]
+      allNotFalse = anchor symbolicToks
+
+--  print chunkedSymbolicSent
+--  putStrLn $ showSentence (dechunk symbolicToks)
+  print allNotFalse
+
+  sequence_ [addClause s cl | cl <- applyRule rule symbolicToks ]
+
+  lt <- count s t
+  b <- solveMaximize s [] lt
+
+
+  as <- sequence [ modelValue s x | x <- t ]
+  let alltoks = [ ((i,(WF t:((Tag (sc++l)):ts))),lit) 
+                  | (b, ((i,(WF t:Tag l:ts)),lit)) <- zip as symbolicToks 
+                  , let sc = if b then "" else "; " ]
[…]

### 2015-05-22 paper/nodalida2015.tex
text about applications

@@
 
+\section{Applications}
+
+As described in Section~\ref{sec:eval}, our implementation is not
+competetive with the state of the art. It is more fruitful to focus on
+things where the SAT-based representation can make an impact.
+One such thing is grammar analysis.
+
+Traditionally, CG rules are applied deterministically and
+sequentially; at the execution of rule, there is no way to
+go back to earlier rules and undo them. Thus there is no way to know
+what part of the input comes from the morphological analyser and what
+from the CG.
+
+A conflict can be defined as 
+a set of two or more rules, such that applying the first makes the
+next rules impossible to apply, regardless of input.
+As a simple example, these rules could be
+
+\begin{itemize}
+\item [] \texttt{SELECT verb ;} \\
+         \texttt{REMOVE verb IF (-1 det) ;}
+\end{itemize}
+
+The first rule selects the verb always, removing all other readings,
+and the second rule removes verb reading, given certain condition.
+If the rules are introduced in a different order, there is no
+conflict: the REMOVE rule would not remove verb readings from all
+possible verb analyses, so there is a possibility for the SELECT rule
+to fire.
+
+The way we implement SAT-CG is ideal for this kind of grammar
+analysis. After applying each rule, it stores the clauses produced by
+it and commits to them -- this makes the execution slow, but it
+retains the dependencies.
+Given the design that each rule application (instead of rule) gets a variable,
+in case of a conflict, the program prints out the particular
+application that violates the previous clauses, with the sentence
+where it is applied. 
+This conflict solving can be done per section, or for the whole grammar. 
+In case of section-based, we run each section on the input, and give
+the output as an input to the new section, discarding all clauses.
+
+Maximisation-based conflict solving (conflicting rule set --> remove least
+possible amount to get a non-conflicting rule set) is not feasible for this task. 
+The whole definition of conflicts depends on ordering, and it would
+not differentiate between the case where \texttt{SELECT verb} comes
+before \texttt{REMOVE verb IF (-1 det)} or after.
+On the other hand, a formalism such as FSIG would benefit from the
+maximisation-based technique in conflict handling.
+
+% SELECT (a) IF (0 PALJON + Nom + Sg) (1 N + Nom + Sg) ; # PALJO VALITTAMINEN
+% SELECT (a) IF (0 PALJON + Par) (1 N + Par + Sg) ;
+% SELECT (a) IF (0 PALJON + LOC-CASE) (1 N + 1 LOC-CASE + Sg) ;
+% REMOVE (a) IF (0 PALJON) ; # "on asetettu paljon haltijaksi"
+
 \section{Results and evaluation}
-
+\label{sec:eval}
 \paragraph{Time}

### 2015-05-22 CG_SAT.hs
stuff

@@ -73,7 +73,7 @@ tagsDontMatchRule tagsInRule ((_, tagsInAna), _) = go tagsInRule tagsInAna
 -- ((2,["bear",<vblex>,<pl>]),Lit v2)
 -- ((3,["sleep",<n>,<pl>]),Lit v4)
 -- ((3,["sleep",<vblex>,<sg>,<p3>]),Lit v5)
-chunk :: Sentence -> [(Integer,[Tag])]
+--chunk :: Sentence -> [(Integer,[Tag])]
 chunk sent = concat $ go sent 1
    where go []    _n = []
          go (x:xs) n = map ((,) n) x : go xs (n+1)
@@ -203,8 +203,10 @@ getContext tok allToks ((C position (bool,ctags)):cs) = getContext tok allToks c
 disamSection ::  ([Rule] -> Sentence -> IO Sentence) -> [[Rule]] -> Sentence -> IO Sentence
 disamSection disam []         sent = return sent
 disamSection disam [rs]       sent = disam rs sent
-disamSection disam (r1:r2:rs) sent = disam r1 sent
-                                      >>= \s -> (print s >> disamSection disam ((r1++r2):rs) s)
+disamSection disam (r1:r2:rs) sent = disam (take 10 (reverse r1)) sent
+                                      >>= \s -> ({-print s >>-} disamSection disam 
+                                                ((r1++r2):rs) s)
+ 
 
 disambiguate :: Bool -> Bool -> [Rule] -> Sentence -> IO Sentence
 disambiguate verbose debug rules sentence = do
@@ -309,7 +311,8 @@ disambiguateWithOrder verbose debug rules sentence = do
       --mapM_ print applied
       let onlyClauses = (map.map) snd $ groupBy (\x y -> fst x==fst y) applied :: [[[Lit]]]
 
-      sequence_ [ addClause s cls | cls <- concat onlyClauses ]
+      --add clauses; in the "bs <- " loop when we add lits, some of those clauses will then be automatically removed
+      --sequence_ [ addClause s cls | cls <- concat onlyClauses ]
 
 
       when debug $ do

### 2015-05-22 CG_base.hs

@@ -49,9 +49,9 @@ instance Show TagSet where
   show x = "TODO"
 
 showTagset :: [[Tag]] -> String
-showTagset [[x]] = show x
+showTagset [[x]] = show x ++ " "
 showTagset xs    = concatMap show' xs
-  where show' [y] = show y
+  where show' [y] = show y ++ " "
         show' ys  = "(" ++ unwords (map show ys) ++ ")"
 
 toTags :: TagSet -> [[Tag]]

### 2015-05-22 test/Test.hs

@@ -51,10 +51,13 @@ main = do
                   let is2 = "2" `elem` o
                       verbose = "v" `elem` o
                       debug = "d" `elem` o
-                      disam = if "no" `elem` o 
+                      disam = if "noord" `elem` o 
                                 then disambiguate verbose debug
                                 else disambiguateWithOrder verbose debug
-                  resSAT <- mapM (disamSection disam rules) text -- :: [Sentence]
+                      disec = if "nosec" `elem` o
+                                then disam (concat rules)
+                                else disamSection disam rules
+                  resSAT <- mapM disec text -- :: [Sentence]
                   resVISL <- vislcg3 r d is2  -- :: [Sentence] 
                   prAll "" resSAT resVISL text verbose
                   putStrLn ""

### 2015-05-29 paper/nodalida2015.tex
more text for paper

@@
 \usepackage{mathptmx}
+\usepackage{fixltx2e}
 %\usepackage{txfonts}
@@
 
-% Uncomment for full paper
 \section{Related work}
@@
 \cite{lindberg_eineborg98ilp,asfrent14} use Inductive Logic Programming to learn CG rules from a tagged corpus.
-\cite{lager01transformation} discusses ordering of the CG rules.
-
+We explore different approaches to ordering. In the case of unordered
+rules, it resembles Finite-State Intersection Grammar \cite{koskenniemi90}.
+See also \cite{lager01transformation} for discussion of ordering of the CG rules.
 \section{CG as a SAT problem}
@@
 
-%Without more context or additional rules we cannot fully disambiguate the passage,
-%but unlike with standard CG, we get only legal combinations in one solution.
-
-%If we add a word, \emph{la casa grande} `the big house' 
-%and a rule that removes verb if followed by an adjective
-%\texttt{REMOVE (v) IF (1 (adj)) ;}
-%we will get a unique solution with \texttt{det n adj}.
-%Standard CG will also remove the verb reading, 
-%but the choice of the word \emph{la} doesn't contribute to
+% Without more context or additional rules we cannot fully disambiguate the passage,
+% but unlike with standard CG, we get only legal combinations in one solution.
 
@@
 making assumptions (e.g. ``\emph{casa} is a noun'') and working under them,
-discarding the assumption if it doesn't lead to a model that satisfies all constraints.
+discarding the assumption if it doesn't lead to a model that satisfies
+all constraints.
 
-\paragraph{Rules are unordered.}
+The traditional CG relies on that applying the rule set iteratively,
+some rules fire the first time and make way for new rules to fire for
+the second time.
+For SAT-based CG, we present the rules as implications:
+``if the \texttt{n} reading for \emph{casa} is true, then discard the
+\texttt{prn} reading''. Each rule generates a list these implications,
+and the SAT solver is left with the task of finding a model that will
+satisfy them.
 
-We have experimented with different strategies:
+Let us continue with the earlier example. 
+We can add a word to the sentence, \emph{la casa grande} `the big house',
+and a rule that removes verb if followed by an adjective.
 \begin{itemize}
-\item Maximise
-% \footnote{Something about local maximum}
-%The algorithm commits to a subset of clauses (e.g.
-\emph{all words must have more than 0 true analyses}), and tries to maximise the number of true variables without changing that subset.
-So it is a local maximum as regards to the given set of clauses.
-number of rules used---We discarded this strategy, because it was too strong: if a rule leads to a conflict in one case, it was not be applied anywhere.
-\item Emulate sequential order---Enter clauses produced by each rule one by one, and assume the solver state reached so far is correct; if a conflict is introduced by new clauses, discard them and move on to next.
-\item [\checkmark] Maximise number of rule applications---If there is a conflict, find the smallest number of rule applications to discard so that the conflict is solved.
+\item [] \texttt{REMOVE v IF (1 adj) ;}
 \end{itemize}
+Then the additional clauses will make it impossible for \emph{casa} to
+be a verb\footnote{If \texttt{adj} is the only reading for
+\emph{grande}, it must be true, because of the restriction that at
+least one analysis for each word is true. Then the implication
+\texttt{grande<adj> => $\neg$casa<v>} has a true antecedent, thus
+its consequent will hold.}, so we will get a unique solution with \texttt{det n adj}.
+Standard CG will also remove the verb reading, 
+but remember the beginning: if the \texttt{det} reading of \emph{la}
+was removed first, then we are stuck with that choice.
+
+% \todo{Unordered rules: also not applicable anymore to apply rules
+% iteratively; there's no ``this rule doesn't fire now but will after
+% applying X and Y'', it's all just implications ``this rule will fire
+% if this is true'' and let the SAT solver find if those rules can apply
+% peacefully to the same input.}
+
+\paragraph{Rules can be unordered.}
+
+As hinted by the previous property, SAT solver doesn't need a fixed
+order of the rules.
+Applying a rule to a sentence produces a number of clauses,
+and those clauses are fed into the SAT solver.
+However, in the unordered scheme, some information is lost: the
+following rule sets would be treated identically, whereas in the
+traditional CG, only the first would be considered as a bad order.
+
+\begin{itemize}
+\item [\texttt{1)}] \texttt{SELECT verb ;} \\
+         \texttt{REMOVE verb IF (-1 det) ;} \\
+         \texttt{REMOVE verb IF (-1 adj) ;} \\
+
+\end{itemize}
+
+\begin{itemize}
+\item [\texttt{2)}] \texttt{REMOVE verb IF (-1 det) ;} \\
+         \texttt{REMOVE verb IF (-1 adj) ;} \\
+         \texttt{SELECT verb ;}
+\end{itemize}
+
+Without order, both of these rule sets will conflict, if applied to an
+input that has sequence \texttt{det verb} or \texttt{adj verb}.
+The SAT solver is fed clauses that tell to select a verb and remove a
+verb, and it cannot build a model that satisfies all of those clauses.
+To solve this problem, we create a variable for every clause
+(\emph{instance of rule application}), and request a solution where maximally many of these variables are true.
+If there is no conflict, then the maximal solution is one where all of
+these variables are true; that is, all instances of rule application are kept.
+In case of a conflict, SAT solver makes it possible to discard only
+minimal amount of rule applications. For the example, assuming that
+there are instances for both contexts for the REMOVE rules, then just
+ignoring the SELECT rule will lead to maximal solution. 
+This corresponds loosely to the common design pattern in
+CGs, where there is a number of rules with the same target, ordered
+such that more secure rules come first,
+with a catch-all rule with no condition as the last resort, to be
+applied if none of the previous has fired.
+
+We can apply this for the whole grammar, or for one section at a time.
+
+The design of unordered rules can lead to new ways of writing CGs,
+potentially with less mental effort to the writer, but also less
+transparent. By testing against gold standard, we see that the
+unordered scheme not optimal for applying ready-made
+CGs, where the order is a crucial detail.
+We can also emulate order: enter clauses produced by each rule one by
+one, and assume the solver state reached so far is correct; if a
+conflict is introduced by new clauses, discard them and move on to
+next.
+This is also better for the performance, because the maximisation
+problems are smaller.
+
+%  that it is possible to write a small grammar, 
+% which, when run with the unordered scheme, disambiguates more than the
+% same grammar run with VISL CG-3. 
+% However, when tested with grammars that are written for the original
+% CG, SAT-CG loses with both strategies, but emulating order fares better.
+
 The chosen heuristic produced the best results with examples in the order of tens or hundreds of rules.
@@
 Considering our example phrase and rules, standard CG implementation
-can only remove readings for the word which is tagged as \texttt{prn} or
-\texttt{det}. The SAT-based implementation interprets the rules as
+can only remove readings for the target word (\texttt{prn} or
+\texttt{det}). The SAT-based implementation interprets the rules as
 ``determiner and verb together are illegal'', and is free to take action that concerns also the word in the condition (\texttt{1 n} or \texttt{1 v}).
@@
   (1 n)} translates into the same logical formula as  \texttt{REMOVE n
-  IF (-1 prn)}. A rule with more conditions can be written as many
+  IF (-1 prn)}. A rule with more conditions corresponds to many
 rules, each condition taking its turn to be the target of removal or selection. \\
@@
 
-However, the results using our approach should not be too different from the
-existing ones---our preliminary evaluation against VISL CG-3 shows
-promising results in this regard.
+When conducting our tests, we have been able to construct a small grammar 
+which, when run with the unordered scheme, disambiguates more than the
+same grammar run with VISL CG-3. However, when tested with grammars
+that are written for the original CG, SAT-CG loses with both ordering
+strategies, but emulating order fares better. 
+This suggests that our implementation should not compete with existing
+state-of-the-art, but rather it has value as a way of relating the CG
+formalism to wider context in the theory of computer science.
+In Section~\ref{sec:apps} we discuss more about possible applications.
 
-\section{Applications}
+\section{Results and evaluation}
+\label{sec:eval}
+\paragraph{Time}
 
+The worst-case complexity of SAT is exponential, whereas the standard
+implementations of CG are polynomial, but with advances in SAT solving
+techniques, the performance in the average case in practice is much more feasible than in the previous works done in 90s--00s.
+We used the open-source SAT solver MiniSat \cite{een04sat}.
+
+Table~\ref{table:time} shows the execution time compared to VISL CG-3. 
+The number of rules in the grammar is more significant than the word
+count; this holds also for the performance of VISL CG-3.
+From the SAT solving side, maximisation is the most costly operation;
+hence emulating ordering is faster, because maximisation problems are smaller.
+Parsing Don Quijote (384,155 words) with the rule set of 261 rules,
+the function was called 147,253 times, and with 19 rules, 132,255 times.
+%However, with both rule sets, half of the sentences in Don Quijote
+%needed less than 6 calls of maximise, and 75 \% of the sentences
+%needed less than 14.
+The difference in the execution times suggests that there are other
+reasons for the worse performance---this is to be expected,
+SATCG is currently just a naive proof-of-concept implementation with no optimisations.
+
+\begin{table}
+  \centering
+  \begin{tabular}{|l|c|c|c|c|}
+     \hline
+   \textbf{\# words} & \textbf{\# rules} &  \textbf{SATCG\textsubscript{u}} & \textbf{SATCG\textsubscript{o}} & \textbf{VISL CG-3} \\ \hline
+                384,155  & 19   & TODO &  23.1s   & 4.2s\\ %\hline
+	        384,155  & 99   & TODO & 1m40.6s & 6.1s \\ %\hline
+                384,155  & 261  & TODO & 2m36.7s & 10.7s \\ \hline
+  \end{tabular}
+  \caption{Execution times.}
+  \label{table:time}
+\end{table}
+
+\paragraph{Performance against VISL CG-3}
+
+We took a tagged corpus\footnote{https://svn.code.sf.net/p/apertium/svn/branches/apertium-swpost/apertium-en-es/es-tagger-data/es.tagged} of 21865 words for Spanish, 
+and a small constraint grammar\footnote{https://svn.code.sf.net/p/apertium/svn/languages/apertium-spa/apertium-spa.spa.rlx}, produced independently of the authors.
+% We ignored substitute rules, but kept subreadings and unification, even though our implementation doesn't handle them.
+We only took select and remove rules (no substitute/iff), after which the constraint grammar had 261 rules.
+With this setup, we took the text of the tagged corpus with all
+ambiguities, and ran both SATCG and VISL CG-3 with the same grammar. 
+Treating the corpus as the gold standard, the disambiguation by
+SATCG achieves 75.09 \% correct result and VISL CG-3 75.80 \%.
+Similar patterns were observed with small (\textless{}20) rule sets
+written by the authors; depending on the subset, SATCG and VISL CG-3 had
+a difference of at most $\Mypm$ 1.5 \%. 
+Introducing rules one by one up to 19, the
+performance improved in a very similar rate, with less than 0.5 \%
+difference between the systems at each new rule.
+
+% Additional tests could include plugging SATCG into Apertium
+% translation pipeline, and comparing the translation quality.
+
+\section{Applications and Future Work}
+\label{sec:apps}
 As described in Section~\ref{sec:eval}, our implementation is not
@@
 what part of the input comes from the morphological analyser and what
-from the CG.
+part from applying previous rules.
 
@@
 
-\section{Results and evaluation}
-\label{sec:eval}
-\paragraph{Time}
-
-The worst-case complexity of SAT is exponential, whereas the standard
-implementations of CG are polynomial, but with advances in SAT solving
-techniques, the performance in the average case in practice is much more feasible than in the previous works done in 90s--00s.
-We used the open-source SAT solver MiniSat \cite{een04sat}.
-
-Table~\ref{table:time} shows the execution time compared to VISL CG-3. 
-The number of rules in the grammar is more significant than the word
-count; this holds also for the performance of VISL CG-3.
-From the SAT solving side, maximisation is the most costly operation.
-Parsing Don Quijote (384,155 words) with the rule set of 261 rules,
-the function was called 147,253 times, and with 19 rules, 132,255 times.
-%However, with both rule sets, half of the sentences in Don Quijote
-%needed less than 6 calls of maximise, and 75 \% of the sentences
-%needed less than 14.
-The difference in the execution times suggests that there are other
-reasons for the worse performance---this is to be expected,
-SATCG is currently just a naive proof-of-concept implementation with no optimisations.
-
-\begin{table}
-  \centering
-  \begin{tabular}{|l|c|c|c|}
-     \hline
-   \textbf{\# words} & \textbf{\# rules} &  \textbf{SATCG} & \textbf{VISL CG-3} \\ \hline
-                384,155  & 19   &  23.1s   & 4.2s\\ %\hline
-	        384,155  & 99   &  1m40.6s & 6.1s \\ %\hline
-                384,155  & 261  &  2m36.7s & 10.7s \\ \hline
-  \end{tabular}
-  \caption{Execution times.}
-  \label{table:time}
-\end{table}
-
-\paragraph{Performance against VISL CG-3}
-
-We took a tagged corpus\footnote{https://svn.code.sf.net/p/apertium/svn/branches/apertium-swpost/apertium-en-es/es-tagger-data/es.tagged} of 21865 words for Spanish, 
-and a small constraint grammar\footnote{https://svn.code.sf.net/p/apertium/svn/languages/apertium-spa/apertium-spa.spa.rlx}, produced independently of the authors.
-% We ignored substitute rules, but kept subreadings and unification, even though our implementation doesn't handle them.
-We only took select and remove rules (no substitute/iff), after which the constraint grammar had 261 rules.
-With this setup, we took the text of the tagged corpus with all
-ambiguities, and ran both SATCG and VISL CG-3 with the same grammar. 
-Treating the corpus as the gold standard, the disambiguation by
-SATCG achieves 75.09 \% correct result and VISL CG-3 75.80 \%.
-Similar patterns were observed with small (\textless{}20) rule sets
-written by the authors; depending on the subset, SATCG and VISL CG-3 had
-a difference of at most $\Mypm$ 1.5 \%. 
-Introducing rules one by one up to 19, the
-performance improved in a very similar rate, with less than 0.5 \%
-difference between the systems at each new rule.
-
-% Additional tests could include plugging SATCG into Apertium
-% translation pipeline, and comparing the translation quality.
-
-% \paragraph{Removing rules}
-% \begin{table}
-%   \centering
-%   \begin{tabular}{|l|c|c|}
-%      \hline
-%     \textbf{\# rules} & \textbf{SATCG} & \textbf{VISL CG-3} \\
-%     \hline 
-% 	19 &  83.15 & 83.33 \\ \hline  % 19 out of 23 preselected by authors
-% 	18 &  83.19 & 83.04 \\ \hline  % 18 out of 19
-% 	17 &  83.201 & 83.03 \\ \hline   % 17 out of 18/19, no diff
-% 	16 &  83.197 & 83.030 \\ \hline  % 16 out of 17/18, no diff
-% 	15 &  83.188 & 83.022\\ \hline  % 15 out of 16/17. Tie with 14 for SATCG, chose to retain the rule that gave better results for VISL CG-3.
-% 	14 &  83.180 & 83.005 \\ \hline  % Tie with 15 for SATCG.
-% 	13 &  83.159 & 82.951 \\ \hline  % 13 out of 15 
-% 	12 &  83.134 & 82.884\\ \hline  % 12 out of 14
-% 	11 &  83.089 & 82.847\\ \hline  
-% 	10 &  83.051 & 82.802\\ \hline  
-% 	9  & 82.997 & 82.802 \\ \hline
-% 	8  & 82.906 & 82.681 \\ \hline  % REMOVE:r_imp Imp IF (0 Subj) (-1* CnjSub) ; 
-% 	7  & 82.743 & 82.594 \\ \hline  % SELECT:s_pro PrnIndep IF (1C VerbFin) ;
-% 	6  & 82.556 & 82.398\\ \hline  %  REMOVE:r_v_v Verb IF (-1C Vblex) ;
-% 	5  & 82.327 & 82.028 \\ \hline % SELECT:s_adv Adv IF (1 Cnj_Rel_End OR Prep OR Det)
-% 	4  & 82.065 & 82.028 \\ \hline % REMOVE:r_vfin_prep VerbFin IF (0 Prep) (1 Inf) 
-% 	3  & 81.387 & 81.233 \\ \hline % REMOVE:r_noun N    IF (-1 N OR Prn)
-% 	2  & 80.23 & 79.97 \\ \hline % SELECT:s_pri_1 (pri p3) IF (0 (vblex imp p2 sg))
-% 	1  & 76.69 & 76.62 \\ \hline 
-% 	0  & 70.70 & 70.70 \\ \hline
-%   \end{tabular}
-%   \caption{Result of removing rules one by one, and how close to the gold standard they get.}
-%   \label{table:results_remove}
-% \end{table}
-% In this experiment, we took 19 rules written by the authors, which give 83.15 \% correct results with SATCG and 83.33 \% with VISL CG-3.
-% Then we removed rules one by one; after getting a score for $n$ rules, we took the combination of $n-1$ rules for which SATCG gives the best score, and we ran the rule set with VISL CG-3 for comparison.
-Table~\ref{table:results_remove} shows the percentages compared to the gold standard for both systems.
-
 \section{Conclusions}

### 2015-05-31 paper/nodalida2015.tex
more text

@@
 
-\noindent The most important differences between the traditional and the SAT-based approach are summarised below:
+\noindent The most important differences between the traditional and
+the SAT-based approach are described in the following sections.
 
-\paragraph{Condition of being unambiguously tagged is irrelevant.}
-Rather than waiting for a word to get disambiguated, the SAT solver starts by 
-making assumptions (e.g. ``\emph{casa} is a noun'') and working under them,
-discarding the assumption if it doesn't lead to a model that satisfies
-all constraints.
+\subsection{Rules disambiguate more}
+Considering our example phrase and rules, the standard CG implementation
+can only remove readings from the target word (\texttt{prn} or
+\texttt{det}). The SAT-based implementation interprets the rules as
+``determiner and verb together are illegal'', and is free to take action that concerns also the word in the condition (\texttt{n} or \texttt{v}).
 
-The traditional CG relies on that applying the rule set iteratively,
-some rules fire the first time and make way for new rules to fire for
-the second time.
-For SAT-based CG, we present the rules as implications:
-``if the \texttt{n} reading for \emph{casa} is true, then discard the
-\texttt{prn} reading''. Each rule generates a list these implications,
-and the SAT solver is left with the task of finding a model that will
-satisfy them.
+This behaviour is explained by simple properties of logical formulae.
+When the rules are applied to the text, they are translated into
+implications: \texttt{REMOVE prn IF (1 n)} becomes
+ \emph{casa}\texttt{<n>} $\Rightarrow$ $\neg$\emph{la}\texttt{<prn>},
+ which reads ``if the \texttt{n} reading for \emph{casa} is true, then
+ discard the \texttt{prn} reading for \emph{la}''.
+Any implication $a\,\Rightarrow\,b$ can be represented as a disjunction
+$\neg a\,\vee\,b$; intuitively, either the antecedent is false
+and the consequent can be anything, or the consequent is true and the
+antecedent can be anything.
+Due to this property, our rule translates into the disjunction 
+$\neg$\emph{casa}\texttt{<n>} $\vee$ $\neg$\emph{la}\texttt{<prn>}.
+
+However, our rule is not the only way to reach this form: since
+disjunction doesn't care about order, we notice that also
+\emph{la}\texttt{<prn>} $\Rightarrow$ $\neg$\emph{casa}\texttt{<n>} 
+corresponds to the same disjunction.
+This means that the rules are logically flipped: \texttt{REMOVE prn IF
+  (1 n)} translates into the same logical formula as  \texttt{REMOVE n
+  IF (-1 prn)}. 
+A rule with more conditions corresponds to many
+rules, each condition taking its turn to be the target of removal or
+selection.
+
+% \begin{itemize}
+% \item [] \texttt{REMOVE n IF (-1 prn) ;} \\
+%          \texttt{REMOVE v IF (-1 det) ;}
+% \end{itemize}
+
+% SAT-based approach gives identical results with both sets of rules,
+% whereas the standard CG would remove one reading from \emph{casa} and leave \emph{la} ambiguous.
+% Testing this property with more complex rules and larger rule sets remains to be done.
+
+%Rules with more conditions translate into many rules; rules with negation become complements (\texttt{(*)-X} for \texttt{NOT X}).
+Rules which require the condition to be unambiguously tagged, don't have an equivalent flip in the standard CG.
+% We did not expect much practical benefits, save for realising that some rules are bad by having to think further what it implies,
+%Out of interest, we manually flipped a small grammar and tested it against the original.
+% \todo{The results show  }
+%\begin{itemize}
+%\item []          \texttt{SL Inf IF (-2 V-MOD) (-1 "not");} \\
+%$\rightarrow$\texttt{SL V-MOD   IF (1 "not") (2 INF);} \\
+%$\rightarrow$\texttt{SL "not" IF (-1 V-MOD) (1 Inf);}
+%\item []           \texttt{SL Foo IF (NOT 1 Bar)} \\
+%$\rightarrow$\texttt{SL (*)-Foo if (-1 Bar)}
+%\end{itemize}
+
+\subsection{Cautious context is irrelevant}
+% Rather than waiting for a word to get disambiguated, the SAT solver starts by 
+% making assumptions (e.g. ``\emph{casa} is a noun'') and working under them,
+% discarding the assumption if it doesn't lead to a model that satisfies
+% all constraints.
+
+The traditional CG relies on that applying the rule set iteratively:
+some rules fire during the first iteration, either because their
+conditions don't require cautious context, or because some words are
+unambiguous to start with. This makes some more words unambiguous, and
+new rules can fire during the second iteration.
+
+In SAT-CG, the notion of cautious context is irrelevant.
+When the rules are applied to the text, they are translated into
+implications, such as \\ 
+ \emph{casa}\texttt{<n>} $\Rightarrow$ $\neg$\emph{la}\texttt{<prn>},
+ which reads ``if the \texttt{n} reading for \emph{casa} is true, then
+ discard the \texttt{prn} reading for \emph{la}''. Each rule generates
+ a number of these implications, and the SAT solver tries to find a model that will satisfy them.
 
 Let us continue with the earlier example. 
-We can add a word to the sentence, \emph{la casa grande} `the big house',
-and a rule that removes verb if followed by an adjective.
+We can add a word to the input:
+\begin{itemize}
+\item [] \emph{la casa grande} `the big house'
+\end{itemize}
+and a rule that removes verb reading, if the word is followed by an adjective:
 \begin{itemize}
@@
 \end{itemize}
-Then the additional clauses will make it impossible for \emph{casa} to
-be a verb\footnote{If \texttt{adj} is the only reading for
-\emph{grande}, it must be true, because of the restriction that at
-least one analysis for each word is true. Then the implication
-\texttt{grande<adj> => $\neg$casa<v>} has a true antecedent, thus
-its consequent will hold.}, so we will get a unique solution with \texttt{det n adj}.
-Standard CG will also remove the verb reading, 
-but remember the beginning: if the \texttt{det} reading of \emph{la}
-was removed first, then we are stuck with that choice.
+
+% The SAT solver has already the clauses produced by the previous rules.
+% \emph{casa}\texttt{<n>} $\Rightarrow$ $\neg$\emph{la}\texttt{<prn>}
+% and   \emph{casa}\texttt{<v>} $\Rightarrow$ $\neg$\emph{la}\texttt{<det>}
+The new rule adds the implication
+ \emph{grande}\texttt{<adj>} $\Rightarrow$
+ $\neg$\emph{casa}\texttt{<v>}, which will
+disambiguate  \emph{casa} to a noun\footnote{Assuming that
+  \texttt{adj} is the only reading for \emph{grande}, it must be true,
+  because of the restriction that at least one analysis for each word
+  is true. Then the implication has a true antecedent (\emph{grande}\texttt{<adj>}), thus its
+  consequent ($\neg$\emph{casa}\texttt{<v>}) will hold.}. 
+With the status of \emph{casa} resolved, the SAT solver can now make
+decision between the two models that it would find with only \emph{la casa}.
+Now \emph{casa}\texttt{<n>} is true, so \emph{la}\texttt{<prn>} must
+be false. 
+Because \emph{la}\texttt{<prn>} is false,  \emph{casa} cannot
+ be a \texttt{v}, and we get a unique solution with \texttt{det n adj}.
+
+Contrast this with the behaviour of the standard CG.
+
+With the new rule, standard CG will also remove the verb reading from \emph{casa}, 
+but it is in no way connected to the choice for \emph{la}. It all
+depends of the order of the two rules; if the \texttt{det} reading of \emph{la} is removed first, then we are stuck with that choice.
 
@@
 
-\paragraph{Rules can be unordered.}
+\subsection{Rules can be unordered}
 
@@
 following rule sets would be treated identically, whereas in the
-traditional CG, only the first would be considered as a bad order.
+traditional CG, only the second would be considered as a bad order.
 
 \begin{itemize}
-\item [\texttt{1)}] \texttt{SELECT verb ;} \\
-         \texttt{REMOVE verb IF (-1 det) ;} \\
-         \texttt{REMOVE verb IF (-1 adj) ;} \\
-
-\end{itemize}
-
-\begin{itemize}
-\item [\texttt{2)}] \texttt{REMOVE verb IF (-1 det) ;} \\
+\item [\texttt{1)}] \texttt{REMOVE verb IF (-1 det) ;} \\
          \texttt{REMOVE verb IF (-1 adj) ;} \\
          \texttt{SELECT verb ;}
+\item [\texttt{2)}] \texttt{SELECT verb ;} \\
+         \texttt{REMOVE verb IF (-1 det) ;} \\
+         \texttt{REMOVE verb IF (-1 adj) ;} 
 \end{itemize}
@@
 % \todo{Inari: check if Tapanainen 1996 discuss ordering}
-
-\paragraph{Rules disambiguate more.}
-Considering our example phrase and rules, standard CG implementation
-can only remove readings for the target word (\texttt{prn} or
-\texttt{det}). The SAT-based implementation interprets the rules as
-``determiner and verb together are illegal'', and is free to take action that concerns also the word in the condition (\texttt{1 n} or \texttt{1 v}).
-This means that the rules are logically flipped: \texttt{REMOVE prn IF
-  (1 n)} translates into the same logical formula as  \texttt{REMOVE n
-  IF (-1 prn)}. A rule with more conditions corresponds to many
-rules, each condition taking its turn to be the target of removal or selection. \\
-
-% \begin{itemize}
-% \item [] \texttt{REMOVE n IF (-1 prn) ;} \\
-%          \texttt{REMOVE v IF (-1 det) ;}
-% \end{itemize}
-
-% SAT-based approach gives identical results with both sets of rules,
-% whereas the standard CG would remove one reading from \emph{casa} and leave \emph{la} ambiguous.
-% Testing this property with more complex rules and larger rule sets remains to be done.
-
-%Rules with more conditions translate into many rules; rules with negation become complements (\texttt{(*)-X} for \texttt{NOT X}).
-Rules which require the condition to be unambiguously tagged, don't have an equivalent flip in the standard CG.
-% We did not expect much practical benefits, save for realising that some rules are bad by having to think further what it implies,
-%Out of interest, we manually flipped a small grammar and tested it against the original.
-% \todo{The results show  }
-%\begin{itemize}
-%\item []          \texttt{SL Inf IF (-2 V-MOD) (-1 "not");} \\
-%$\rightarrow$\texttt{SL V-MOD   IF (1 "not") (2 INF);} \\
-%$\rightarrow$\texttt{SL "not" IF (-1 V-MOD) (1 Inf);}
-%\item []           \texttt{SL Foo IF (NOT 1 Bar)} \\
-%$\rightarrow$\texttt{SL (*)-Foo if (-1 Bar)}
-%\end{itemize}
 
@@
 
-The way we implement SAT-CG is ideal for this kind of grammar
+The SAT-CG implementation is ideal for this kind of corpus-based grammar
 analysis. After applying each rule, it stores the clauses produced by
@@
 
-Maximisation-based conflict solving (conflicting rule set --> remove least
-possible amount to get a non-conflicting rule set) is not feasible for this task. 
+Maximisation-based conflict solving (see Section 3) is not feasible for this task. 
 The whole definition of conflicts depends on ordering, and it would
@@
 maximisation-based technique in conflict handling.
+
+For future work, we intend to apply the SAT solver in testing for
+conflicts without a corpus. Let us illustrate the idea with the rule
+set shown in Section 3:
+
+\begin{itemize}
+\item [\texttt{1)}] \texttt{SELECT verb ;} \\
+         \texttt{REMOVE verb IF (-1 det) ;} \\
+
+\end{itemize}
+
+\begin{itemize}
+\item [\texttt{2)}] \texttt{REMOVE verb IF (-1 det) ;} \\
+         \texttt{SELECT verb ;}
+\end{itemize}
+
+Assume we have tag set \texttt{\{det, noun, verb, sg, pl\}}. Then we
+can ask the SAT solver a question if there exists an input which
+satisfies our constraints: both rules, applied in the order, remove
+something from the input.
+after applying the first rule, can be modified by the second rule.
+
+Implementing the design on a larger scale is left for future work.
 
@@
 
-% \section*{Acknowledgments}
+\section*{Acknowledgments}
+
+We thank Tino Didriksen, Eckhard Bick, Francis Tyers and Anssi
+Yli-Jyrä for comments and suggestions; also everyone at the CG
+workshop who participated in the discussion.
 

### 2015-05-31 paper/nodalida2015.tex
more and less and better and worse

@@
 $\neg$\emph{casa}\texttt{<n>} $\vee$ $\neg$\emph{la}\texttt{<prn>}.
-
-However, our rule is not the only way to reach this form: since
-disjunction doesn't care about order, we notice that also
+%Since disjunction doesn't care about the order, we notice that
+However, our rule is not the only way to reach this form---the implication
 \emph{la}\texttt{<prn>} $\Rightarrow$ $\neg$\emph{casa}\texttt{<n>} 
@@
 
-The traditional CG relies on that applying the rule set iteratively:
+Traditional CG relies on that applying the rule set iteratively:
 some rules fire during the first iteration, either because their
@@
 
-In SAT-CG, the notion of cautious context is irrelevant.
-When the rules are applied to the text, they are translated into
-implications, such as \\ 
- \emph{casa}\texttt{<n>} $\Rightarrow$ $\neg$\emph{la}\texttt{<prn>},
- which reads ``if the \texttt{n} reading for \emph{casa} is true, then
- discard the \texttt{prn} reading for \emph{la}''. Each rule generates
- a number of these implications, and the SAT solver tries to find a model that will satisfy them.
+In SAT-CG, the notion of cautious context is irrelevant. Instead of
+removing readings immediately,  each rule generates a  number of
+implications, and the SAT solver tries to find a model that will satisfy them.
 
@@
 Contrast this with the behaviour of the standard CG.
-
 With the new rule, standard CG will also remove the verb reading from \emph{casa}, 
 but it is in no way connected to the choice for \emph{la}. It all
-depends of the order of the two rules; if the \texttt{det} reading of \emph{la} is removed first, then we are stuck with that choice.
+depends of the order of the two rules; if the \texttt{det} reading of
+\emph{la} is removed first, then we are stuck with that choice.
+The SAT-based approach only removes readings after it has enough
+evidence to do that. From the grammar writer perspective, this removes
+a burden of having to decide whether the rule should be cautious or
+not---the SAT solver will only apply it if it has enough evidence from
+the surrounding context.
 
@@
 \begin{itemize}
-\item [\texttt{1)}] \texttt{REMOVE verb IF (-1 det) ;} \\
-         \texttt{REMOVE verb IF (-1 adj) ;} \\
-         \texttt{SELECT verb ;}
-\item [\texttt{2)}] \texttt{SELECT verb ;} \\
-         \texttt{REMOVE verb IF (-1 det) ;} \\
-         \texttt{REMOVE verb IF (-1 adj) ;} 
+\item [\texttt{1)}] \texttt{REMOVE v IF (-1 det) ;} \\
+         \texttt{REMOVE v IF (-1 adj) ;} \\
+         \texttt{SELECT v ;}
+\item [\texttt{2)}] \texttt{SELECT v ;} \\
+         \texttt{REMOVE v IF (-1 det) ;} \\
+         \texttt{REMOVE v IF (-1 adj) ;} 
 \end{itemize}
@@
 Without order, both of these rule sets will conflict, if applied to an
-input that has sequence \texttt{det verb} or \texttt{adj verb}.
+input that has sequence \texttt{det v} or \texttt{adj v}.
 The SAT solver is fed clauses that tell to select a verb and remove a
@@
 If there is no conflict, then the maximal solution is one where all of
-these variables are true; that is, all instances of rule application are kept.
+these variables are true; that is, all instances of rule application
+are kept.
+
 In case of a conflict, SAT solver makes it possible to discard only
-minimal amount of rule applications. For the example, assuming that
-there are instances for both contexts for the REMOVE rules, then just
-ignoring the SELECT rule will lead to maximal solution. 
+minimal amount of rule applications. Continuing with the example,
+assume that there are instances for both contexts for the REMOVE
+rules, then just ignoring the SELECT rule will lead to maximal solution. 
 This corresponds loosely to the common design pattern in
@@
 This is also better for the performance, because the maximisation
-problems are smaller.
+problems are smaller. \\
 
@@
 % CG, SAT-CG loses with both strategies, but emulating order fares better.
-
-The chosen heuristic produced the best results with examples in the order of tens or hundreds of rules.
-Larger rule sets are yet to be tested.
 
@@
 When conducting our tests, we have been able to construct a small grammar 
-which, when run with the unordered scheme, disambiguates more than the
-same grammar run with VISL CG-3. However, when tested with grammars
-that are written for the original CG, SAT-CG loses with both ordering
-strategies, but emulating order fares better. 
+which, when run with the unordered scheme, disambiguates more
+accurately than the same grammar run with VISL CG-3. However, when
+testing with grammars that are written for the original CG, SAT-CG
+loses with both ordering strategies, but emulating order fares better.
 This suggests that our implementation should not compete with existing
@@
 \label{sec:eval}
-\paragraph{Time}
+
+\subsection{Time}
 
@@
 count; this holds also for the performance of VISL CG-3.
-From the SAT solving side, maximisation is the most costly operation;
-hence emulating ordering is faster, because maximisation problems are smaller.
+From the SAT solving side, maximisation is the most costly operation; emulating ordering is faster, because maximisation problems are smaller.
 Parsing Don Quijote (384,155 words) with the rule set of 261 rules,
@@
 reasons for the worse performance---this is to be expected,
-SATCG is currently just a naive proof-of-concept implementation with no optimisations.
+SAT-CG is currently just a naive proof-of-concept implementation with no optimisations.
 
@@
   \centering
-  \begin{tabular}{|l|c|c|c|c|}
+  \begin{tabular}{|c|c|c|c|}
      \hline
-   \textbf{\# words} & \textbf{\# rules} &  \textbf{SATCG\textsubscript{u}} & \textbf{SATCG\textsubscript{o}} & \textbf{VISL CG-3} \\ \hline
-                384,155  & 19   & TODO &  23.1s   & 4.2s\\ %\hline
-	        384,155  & 99   & TODO & 1m40.6s & 6.1s \\ %\hline
-                384,155  & 261  & TODO & 2m36.7s & 10.7s \\ \hline
+     \textbf{\# rules} &  \textbf{SAT-CG\textsubscript{u}} & \textbf{SAT-CG\textsubscript{o}} & \textbf{VISL CG-3} \\ \hline
+      19   & TODO &  23.1s   & 4.2s\\ %\hline
+      99   & TODO & 1m40.6s & 6.1s \\ %\hline
+      261  & TODO & 2m36.7s & 10.7s \\ \hline
   \end{tabular}
@@
 
-\paragraph{Performance against VISL CG-3}
+\subsection{Performance against VISL CG-3}
 
@@
 With this setup, we took the text of the tagged corpus with all
-ambiguities, and ran both SATCG and VISL CG-3 with the same grammar. 
+ambiguities, and ran both SAT-CG and VISL CG-3 with the same grammar. 
 Treating the corpus as the gold standard, the disambiguation by
-SATCG achieves 75.09 \% correct result and VISL CG-3 75.80 \%.
+SAT-CG achieves F-score of 81.54 \%  and VISL CG-3 82.56 \%.
 Similar patterns were observed with small (\textless{}20) rule sets
-written by the authors; depending on the subset, SATCG and VISL CG-3 had
+written by the authors; depending on the subset, SAT-CG and VISL CG-3 had
 a difference of at most $\Mypm$ 1.5 \%. 
@@
 
-% Additional tests could include plugging SATCG into Apertium
+% Additional tests could include plugging SAT-CG into Apertium
 % translation pipeline, and comparing the translation quality.
@@
 One such thing is grammar analysis.
+There has been work on automatic tuning of hand-written CGs \todo{cite
+Eckhard}, but no tools which would search for inconsistencies or
+unoptimal design in order to point it out to the grammar writer.
 
-Traditionally, CG rules are applied deterministically and
+Traditional CG rules are applied deterministically and
 sequentially; at the execution of rule, there is no way to
@@
 
-The SAT-CG implementation is ideal for this kind of corpus-based grammar
-analysis. After applying each rule, it stores the clauses produced by
-it and commits to them -- this makes the execution slow, but it
-retains the dependencies.
+SAT-CG works for this kind of corpus-based grammar analysis as a side
+effect of its design. After applying each rule, it stores the clauses
+produced by it and commits to them---this meas that it retains the
+dependencies.
 Given the design that each rule application (instead of rule) gets a variable,
@@
 This conflict solving can be done per section, or for the whole grammar. 
-In case of section-based, we run each section on the input, and give
-the output as an input to the new section, discarding all clauses.
+In case of section-based application, we run each section on the
+input, and give the output as an input to the new section, discarding
+all clauses. Alternatively, we can let the rules interact with each
+other regardless of the section boundaries.
 
@@
 
-For future work, we intend to apply the SAT solver in testing for
-conflicts without a corpus. Let us illustrate the idea with the rule
+For future work, we intend to apply the SAT solver for corpus-free
+conflict testing. Let us illustrate the idea with the rule
 set shown in Section 3:
@@
 \begin{itemize}
-\item [\texttt{1)}] \texttt{SELECT verb ;} \\
-         \texttt{REMOVE verb IF (-1 det) ;} \\
+\item [\texttt{1)}] \texttt{REMOVE verb IF (-1 det) ;} \\
+         \texttt{SELECT verb ;}
 
-\end{itemize}
-
-\begin{itemize}
-\item [\texttt{2)}] \texttt{REMOVE verb IF (-1 det) ;} \\
-         \texttt{SELECT verb ;}
+\item [\texttt{2)}] \texttt{SELECT verb ;} \\
+         \texttt{REMOVE verb IF (-1 det) ;} 
 \end{itemize}
@@
 something from the input.
-after applying the first rule, can be modified by the second rule.
+SAT solver tells us that with sentence length 2 and maximum 4 readings
+for the whole sentence (1+3, 2+2 or 3+1), there are several inputs that
+satisfy the constraints for the first rule set, but none that
+satisfies the second rule set. Thus we can say that the first rule set
+is sane, but the second rule set is conflicting.
 
-Implementing the design on a larger scale is left for future work.
+Implementing and testing this on a larger scale is left for future work.
 
@@
 
-SAT-solvers are nowadays powerful enough to be used for dealing with Constraint Grammar.
-A logic-based approach to CG has possible advantages over more traditional approaches; a SAT-solver may disambiguate more words, and may do so more precisely.
-Also, the SAT-solver requires less rules, and these rules are simpler.
-The ordering of rules is something we found was incompatible with a logic-based approach.
+SAT-solvers are nowadays powerful enough to be used for dealing with
+Constraint Grammar. A logic-based approach to CG has possible
+advantages over more traditional approaches; a SAT-solver may
+disambiguate more words, and may do so more precisely. Also, the
+SAT-solver requires less rules, and these rules are simpler. 
+\todo{proper conclusions for the new stuff}
+% The ordering of rules is something we found was incompatible with a logic-based approach.
 We compensate this by maximising rule applications.
@@
 We thank Tino Didriksen, Eckhard Bick, Francis Tyers and Anssi
-Yli-Jyrä for comments and suggestions; also everyone at the CG
-workshop who participated in the discussion.
+Yli-Jyr{\"a} for comments and suggestions, samt the anonymous
+reviewers and everyone who participated in the discussion at the CG workshop.
 

### 2015-06-01 paper/nodalida2015.tex
update execution time table

@@
 decision between the two models that it would find with only \emph{la casa}.
-Now \emph{casa}\texttt{<n>} is true, so \emph{la}\texttt{<prn>} must
-be false. 
-Because \emph{la}\texttt{<prn>} is false,  \emph{casa} cannot
- be a \texttt{v}, and we get a unique solution with \texttt{det n adj}.
+As described in the previous section, the dependencies between tags
+are stronger than in standard CG, so we get a unique solution with \texttt{det n adj}.
+% Now \emph{casa}\texttt{<n>} is true, so \emph{la}\texttt{<prn>} must be false. 
+%Because \emph{la}\texttt{<prn>} is false,  \emph{casa} cannot
+% be a \texttt{v}, and we get a unique solution with \texttt{det n adj}.
 
@@
 a burden of having to decide whether the rule should be cautious or
-not---the SAT solver will only apply it if it has enough evidence from
-the surrounding context.
+not---the SAT solver will only take action the surrounding context
+supports the decision.
 
@@
      \textbf{\# rules} &  \textbf{SAT-CG\textsubscript{u}} & \textbf{SAT-CG\textsubscript{o}} & \textbf{VISL CG-3} \\ \hline
-      19   & TODO &  23.1s   & 4.2s\\ %\hline
-      99   & TODO & 1m40.6s & 6.1s \\ %\hline
-      261  & TODO & 2m36.7s & 10.7s \\ \hline
+      19   & 39.7s &  22.1s   & 4.2s\\ %\hline
+      99   & 1m34.1s & 1m14.9s & 6.1s \\ %\hline
+      261  & 2m54.1s & 2m31.6s & 10.7s \\ \hline
   \end{tabular}
-  \caption{Execution times.}
+  \caption{Execution times for 384,155 words.}
   \label{table:time}

### 2015-06-01 paper/nodalida2015.tex
small fixes

@@
 \emph{la} is removed first, then we are stuck with that choice.
-The SAT-based approach only removes readings after it has enough
-evidence to do that. From the grammar writer perspective, this removes
-a burden of having to decide whether the rule should be cautious or
-not---the SAT solver will only take action the surrounding context
-supports the decision.
+If we made the first rules cautious, that is, keeping the determiner
+open until \emph{casa} is disambiguated, then we get the same
+result as with the SAT solver.
+Ideally, both ways of grammar writing should yield good results;
+traditional CG is written in a more imperative way, whereas SAT-CG
+rules are more declarative.
+
+% The SAT-based approach only removes readings after it has enough
+% evidence to do that. From the grammar writer perspective, this removes
+% a burden of having to decide whether the rule should be cautious or
+% not---the SAT solver will only take action the surrounding context
+% supports the decision.
 
@@
 \item [\texttt{1)}] \texttt{REMOVE v IF (-1 det) ;} \\
-         \texttt{REMOVE v IF (-1 adj) ;} \\
          \texttt{SELECT v ;}
 \item [\texttt{2)}] \texttt{SELECT v ;} \\
-         \texttt{REMOVE v IF (-1 det) ;} \\
-         \texttt{REMOVE v IF (-1 adj) ;} 
+         \texttt{REMOVE v IF (-1 det) ;} 
 \end{itemize}
@@
 Without order, both of these rule sets will conflict, if applied to an
-input that has sequence \texttt{det v} or \texttt{adj v}.
+input that has sequence \texttt{det v}.
 The SAT solver is fed clauses that tell to select a verb and remove a
@@
 minimal amount of rule applications. Continuing with the example,
-assume that there are instances for both contexts for the REMOVE
-rules, then just ignoring the SELECT rule will lead to maximal solution. 
+it is not clear which instances would be discarded, but if the rules
+were part of a larger rule set, and in the context the REMOVE rule was
+the right one to choose, it is likely that the interaction between the desired
+rules would make a large set of clauses that fit together, and the
+SELECT rule would not fit in, hence it would be discarded.
+
 This corresponds loosely to the common design pattern in
@@
 applied if none of the previous has fired.
+The order-based heuristic in the traditional CG is replaced by a more
+holistic behaviour: if the rules conflict, discard the one that seems
+like an outlier.
 
-We can apply this for the whole grammar, or for one section at a time.
+%With the added interaction between rules, catch-all rules might even be less 
+%necessary. However, this is just speculation, we have not tested this effect.
 
-The design of unordered rules can lead to new ways of writing CGs,
-potentially with less mental effort to the writer, but also less
-transparent. By testing against gold standard, we see that the
-unordered scheme not optimal for applying ready-made
-CGs, where the order is a crucial detail.
-We can also emulate order: enter clauses produced by each rule one by
-one, and assume the solver state reached so far is correct; if a
-conflict is introduced by new clauses, discard them and move on to
-next.
-This is also better for the performance, because the maximisation
-problems are smaller. \\
+%We can apply this for the whole grammar, or for one section at a time.
+
+%The design of unordered rules can lead to new ways of writing CGs,
+%potentially with less mental effort to the writer, but also less transparent.
+
+%By testing against gold standard, we see that the
+%unordered scheme not optimal for applying ready-made
+%CGs, where the order is a crucial detail.
+
+We can also emulate order with SAT-CG. To do that, we enter clauses
+produced by each rule one by one, and assume the solver state reached
+so far is correct. If a conflict is introduced by new clauses, discard
+them and move on to the next rule.
+By testing against gold standard, we see that this scheme works better
+with ready-made CGs, which are written with ordering in mind.
+It is also slightly better for the performance, because the maximisation
+problems are smaller. 
+%We have also experimented with sections, and 
+\\
 
@@
 \noindent These three features influence the way that rules are written. 
-We predict that less rules are needed; whether this holds in
-the order of thousands of rules remains to be tested.
-Getting rid of ordering could ease the task of the grammar writer,
-since it removes the need to estimate the best sequence of rule
-applications.
+We predict that less rules are needed; whether this holds in the order
+of thousands of rules remains to be tested. On one hand, getting rid
+of ordering and cautious context could ease the task of the grammar
+writer, since it removes the need to estimate the best sequence of
+rule applications and whether to make the rules cautious. On the other
+hand, it is also less transparent, and might not scale up for larger grammars.
 

### 2015-06-01 paper/nodalida2015.tex
small fixes

@@
 
-With these constraints, we get two solutions. The interaction of the rules regarding \emph{la}  disambiguates the POS of \emph{casa} for free, and ordering of the rules doesn't matter.
+With these constraints, we get two solutions. The interaction of the rules regarding \emph{la}  disambiguates \emph{casa} for free, and ordering of the rules doesn't matter.
 
@@
 Due to this property, our rule translates into the disjunction 
-$\neg$\emph{casa}\texttt{<n>} $\vee$ $\neg$\emph{la}\texttt{<prn>}.
-%Since disjunction doesn't care about the order, we notice that
-However, our rule is not the only way to reach this form---the implication
-\emph{la}\texttt{<prn>} $\Rightarrow$ $\neg$\emph{casa}\texttt{<n>} 
-corresponds to the same disjunction.
+$\neg$\emph{casa}\texttt{<n>} $\vee$ $\neg$\emph{la}\texttt{<prn>},
+which is also equivalent to another implication, 
+ \emph{la}\texttt{<prn>} $\Rightarrow$ $\neg$\emph{casa}\texttt{<n>}.
 This means that the rules are logically flipped: \texttt{REMOVE prn IF
@@
   IF (-1 prn)}. 
-A rule with more conditions corresponds to many
-rules, each condition taking its turn to be the target of removal or
-selection.
+A rule with more conditions corresponds to many rules, each condition
+taking its turn to be the target. % of removal or selection.
 
@@
   consequent ($\neg$\emph{casa}\texttt{<v>}) will hold.}. 
-With the status of \emph{casa} resolved, the SAT solver can now make
-decision between the two models that it would find with only \emph{la casa}.
-As described in the previous section, the dependencies between tags
-are stronger than in standard CG, so we get a unique solution with \texttt{det n adj}.
+As the status of \emph{casa} is resolved, the SAT solver can now
+discard the model where \emph{casa} is a verb and \emph{la} is a
+pronoun---all these decisions are connected---and we get a unique solution with \texttt{det n adj}.
 % Now \emph{casa}\texttt{<n>} is true, so \emph{la}\texttt{<prn>} must be false. 
@@
 Ideally, both ways of grammar writing should yield good results;
-traditional CG is written in a more imperative way, whereas SAT-CG
+traditional CG are more imperative, whereas SAT-CG
 rules are more declarative.
@@
 input that has sequence \texttt{det v}.
-The SAT solver is fed clauses that tell to select a verb and remove a
+The SAT solver is given clauses that tell to select a verb and remove a
 verb, and it cannot build a model that satisfies all of those clauses.
-To solve this problem, we create a variable for every clause
-(\emph{instance of rule application}), and request a solution where maximally many of these variables are true.
+To solve this problem, we create a variable for every instance of rule application, and request a solution where maximally many of these variables are true.
 If there is no conflict, then the maximal solution is one where all of
-these variables are true; that is, all instances of rule application
-are kept.
+these variables are true; that is, all rules take action.
 
@@
 rule applications and whether to make the rules cautious. On the other
-hand, it is also less transparent, and might not scale up for larger grammars.
+hand, lack of order can make the rules less transparent, and might not scale up for larger grammars.
 

### 2015-06-01 paper/nodalida2015.tex
change order of time and performance

@@
 with ready-made CGs, which are written with ordering in mind.
-It is also slightly better for the performance, because the maximisation
-problems are smaller. 
+It also runs slightly faster than the unordered version.
+%the performance, because the maximisation problems are smaller. 
 %We have also experimented with sections, and 
@@
 of ordering and cautious context could ease the task of the grammar
-writer, since it removes the need to estimate the best sequence of
-rule applications and whether to make the rules cautious. On the other
+writer, since it removes the burden of estimating the best sequence of
+rules and whether to make the them cautious. On the other
 hand, lack of order can make the rules less transparent, and might not scale up for larger grammars.
-
-When conducting our tests, we have been able to construct a small grammar 
-which, when run with the unordered scheme, disambiguates more
-accurately than the same grammar run with VISL CG-3. However, when
-testing with grammars that are written for the original CG, SAT-CG
-loses with both ordering strategies, but emulating order fares better.
-This suggests that our implementation should not compete with existing
-state-of-the-art, but rather it has value as a way of relating the CG
-formalism to wider context in the theory of computer science.
-In Section~\ref{sec:apps} we discuss more about possible applications.
 
@@
 \label{sec:eval}
+
+When conducting our tests, we have been able to construct small
+grammars (\textless20 rules) which, when run with the unordered scheme,
+disambiguate more accurately than the same grammar run with VISL
+CG-3. However, when testing with larger grammars that are written with the
+traditional CG in mind, SAT-CG loses with both ordering strategies, but emulating
+order fares better.
+In the following sections, we report both execution time and
+performance against state of the art.
+
+%This suggests that our implementation should not compete with existing
+%state-of-the-art, but rather it has value as a way of relating the CG
+%formalism to wider context in the theory of computer science.
+%In Section~\ref{sec:apps} we discuss more about possible applications.
+
+\subsection{Performance against VISL CG-3}
+
+We took a manually tagged
+corpus\footnote{https://svn.code.sf.net/p/apertium/svn/branches/apertium-swpost/apertium-en-es/es-tagger-data/es.tagged}
+containing 21865 words of Spanish news text, 
+and a small constraint grammar\footnote{https://svn.code.sf.net/p/apertium/svn/languages/apertium-spa/apertium-spa.spa.rlx}, produced independently of the authors.
+% We ignored substitute rules, but kept subreadings and unification, even though our implementation doesn't handle them.
+We kept only select and remove rules, after which the grammar had 261 rules.
+With this setup, we took the text of the tagged corpus with all
+ambiguities, and ran both SAT-CG and VISL CG-3 with the same grammar. 
+Treating the corpus as the gold standard, the disambiguation by  VISL
+CG-3 achieves F-score of 82.6 \%, ordered
+SAT-CG 81.5 \%  and unordered SAT-CG 79.2 \%. 
+For comparison, with the best handcrafted grammar, unordered SAT-CG
+achieves 85.3 \% and VISL CG-3 85.9 \%.
+Similar patterns were observed with small (\textless{}20) rule sets
+written by the authors; depending on the subset, SAT-CG and VISL CG-3 had
+a difference of at most $\Mypm$ 1.5 \%. 
+Introducing rules one by one up to 19, the
+performance improved in a very similar rate, with less than 0.5 \%
+difference between the systems at each new rule.
+We did not evaluate on other languages or text genres, due to lack of suitable test data.
+
+% Additional tests could include plugging SAT-CG into Apertium
+% translation pipeline, and comparing the translation quality.
 
@@
 
-Table~\ref{table:time} shows the execution time compared to VISL CG-3. 
+We tested the performance by parsing Don Quijote (384,155 words) with
+the same Spanish grammar. Table~\ref{table:time} shows execution times compared to VISL
+CG-3 for the same text and rules. 
 The number of rules in the grammar is more significant than the word
 count; this holds also for the performance of VISL CG-3.
-From the SAT solving side, maximisation is the most costly operation; emulating ordering is faster, because maximisation problems are smaller.
-Parsing Don Quijote (384,155 words) with the rule set of 261 rules,
+From the SAT solving side, maximisation is the most costly operation; 
+emulating order (SAT-CG\textsubscript{o}) is slightly faster, because maximisation problems are smaller.
+SAT is not Parsing Don Quijote (384,155 words) with the rule set of 261 rules,
 the function was called 147,253 times, and with 19 rules, 132,255 times.
@@
 \end{table}
-
-\subsection{Performance against VISL CG-3}
-
-We took a tagged corpus\footnote{https://svn.code.sf.net/p/apertium/svn/branches/apertium-swpost/apertium-en-es/es-tagger-data/es.tagged} of 21865 words for Spanish, 
-and a small constraint grammar\footnote{https://svn.code.sf.net/p/apertium/svn/languages/apertium-spa/apertium-spa.spa.rlx}, produced independently of the authors.
-% We ignored substitute rules, but kept subreadings and unification, even though our implementation doesn't handle them.
-We only took select and remove rules (no substitute/iff), after which the constraint grammar had 261 rules.
-With this setup, we took the text of the tagged corpus with all
-ambiguities, and ran both SAT-CG and VISL CG-3 with the same grammar. 
-Treating the corpus as the gold standard, the disambiguation by
-SAT-CG achieves F-score of 81.54 \%  and VISL CG-3 82.56 \%.
-Similar patterns were observed with small (\textless{}20) rule sets
-written by the authors; depending on the subset, SAT-CG and VISL CG-3 had
-a difference of at most $\Mypm$ 1.5 \%. 
-Introducing rules one by one up to 19, the
-performance improved in a very similar rate, with less than 0.5 \%
-difference between the systems at each new rule.
-
-% Additional tests could include plugging SAT-CG into Apertium
-% translation pipeline, and comparing the translation quality.
 

### 2015-06-01 CG_SAT.hs
stuff

@@ -48,7 +48,6 @@ sameInd ((i,_),_) ((i',_),_) = i == i'
 --Rule     has [[Tag]].
 --Analysis has [Tag].
 --At least one complete sublist in the rule must be found in the analysis.
---If cautious mode, there must be only one sublist in analysis.
 tagsMatchRule :: [[Tag]] -> Token -> Bool
 tagsMatchRule tr ((_,ta),_) = go tr ta
   where go []       ta = False
@@ -119,7 +118,8 @@ go isSelect tags (conds:cs) allToks = --trace (show conds) $
         mkVars tctx nt = [ conseq:ants | (t, ctx) <- tctx -- (Token,[[Token]])
                                        , tCombs <- sequence ctx -- :: [[Token]]
                                        , let conseq = nt (getLit t)
-                                       , let ants = [] ] --map (neg . getLit) tCombs ] 
+--                                       , let ants = [] ] --for Cautious 
+                                       , let ants = map (neg . getLit) tCombs ] 
      -- tCombs <- sequence ctx: say we have rule REMOVE v IF (-1 det) (1 n)
      -- and we get [ [(1,det)], [(3,n pl), (3,n sg)] ]
      -- we can't just put all of them in the list of antecedents
@@ -164,12 +164,12 @@ getContext tok allToks ((C position (bool,ctags)):cs) = getContext tok allToks c
                                  else [filter match $ exactly 0 tok] --feature in  different reading
 
                 -- for cautious mode. ONLY WORKS WITH SYMBOLIC!
-                Exactly True n -> if length (exactly n tok) > 1 
-                                   then [[]]
-                                   else [filter match $ exactly n tok]
-                AtLeast True n -> if length (atleast n tok) > 1
-                                   then [[]]
-                                   else [filter match $ atleast n tok]
+                -- Exactly True n -> if length (exactly n tok) > 1 
+                --                    then [[]]
+                --                    else [filter match $ exactly n tok]
+                -- AtLeast True n -> if length (atleast n tok) > 1
+                --                    then [[]]
+                --                    else [filter match $ atleast n tok]
 
                 -- normal mode
                 Exactly _ n -> [filter match $ exactly n tok]
@@ -384,6 +384,7 @@ test :: IO ()
 test = mapM_ (disambiguateWithOrder True True rls) CG_data.exs
 
   where rls = [rmVerbIfDet
+             --, slVerbAlways --conflicts with anything that selects other than V 
              , rmNounIfPron
[…]

### 2015-06-01 executable/Main.hs

@@ -20,9 +20,13 @@ main = do
                   let is2 = "2" `elem` o
                       verbose = "v" `elem` o
                       debug = "d" `elem` o
-                      disam = if "no" `elem` o then disambiguate
-                                               else disambiguateWithOrder
-                  mapM_ (disamSection (disam verbose debug) rules) text
+                      disam = if "noord" `elem` o 
+                                then disambiguate verbose debug
+                                else disambiguateWithOrder verbose debug
+                      disec = if "nosec" `elem` o
+                                then disam (concat rules)
+                                else disamSection disam rules
+                  mapM_ disec text
     ("test":_) -> CG_SAT.test
     _          -> putStrLn "usage: ./Main (<rules> <data> | test) [v]"
   

### 2015-06-01 executable/Symbolic.hs

@@ -11,13 +11,14 @@ import System.Environment
 import System.IO.Unsafe
 
 
-n = 2
+n = 3
 
 --dummytags = map (Tag . (:[])) ['b'..'e']
 --dummyrules = parseRules False "REMOVE:r1 (d) IF (-1C (c) LINK 1 (e)) ;\nREMOVE:r2 (b) ;\nREMOVE:r3 (e) IF (-1C (c)) ;"
 
 --TODO: get tags from the grammar to be tested
-tags = map Tag ["det", "n", "v", "pri", "prs", "imp", "p3", "predet", "prn", "adj", "pr"]
+--tags = map Tag ["det", "n", "v", "pri", "prs", "imp", "p3", "predet", "prn", "adj", "pr"]
+tags = map Tag ["det", "v", "p3", "imp", "prs", "predet", "prn"]
 
 randomrules = parseRules False "REMOVE:r1 (v) IF (-1C (det)) ;\nREMOVE:r2 (prs) ;\nREMOVE:r3 (imp) IF (0 (p3)) ;"
 

### 2015-06-01 test/Test.hs

@@ -31,6 +31,7 @@ main :: IO ()
 main = do
   args <- getArgs
   case args of
+    ["subset",n]    -> subsetN (read n) apertium espre
     ["shuffle"]     -> shuffle_ small ambiguous esgold espre
     [r,"shuffle"]   -> shuffle_ r ambiguous esgold espre
     ["enshuffle"]   -> shuffle_ engr entext engold enpre
@@ -210,14 +211,20 @@ rev r d g pre = do
                | (n, ps) <- zip [0..] [rls, reverse rls]
                , let fname = "/tmp/rev" ++ show n ]
 
+subsetN n rules pre = do
+  rls <- concat `fmap` readRules rules
+  let fname = "/tmp/subset" ++ show n
+      ps = take n rls
+  mkRuleFile (take n rls) fname rules pre
+
 
 mkRuleFile :: [Rule] -> FilePath -> FilePath -> FilePath -> IO ()
-mkRuleFile rules fp orig pre = do
+mkRuleFile rules new orig pre = do
   lists <- readFile pre
-  writeFile fp lists
+  writeFile new lists
   rules <- sequence [ putStr name >> grep name | rl <- rules
                        , let name = head $ words $ show rl ]
-  appendFile fp (unwords rules)
+  appendFile new (unwords rules)
 
   where
     grep name = readProcess "egrep" [("\\<"++name++"\\>"), orig] []

### 2015-06-01 paper/nodalida2015.tex
small fixes

@@
 \subsection{Rules can be unordered}
+\label{sec:unord}
 
@@
 
-When conducting our tests, we have been able to construct small
-grammars (\textless20 rules) which, when run with the unordered scheme,
-disambiguate more accurately than the same grammar run with VISL
-CG-3. However, when testing with larger grammars that are written with the
-traditional CG in mind, SAT-CG loses with both ordering strategies, but emulating
-order fares better.
 In the following sections, we report both execution time and
@@
 corpus\footnote{https://svn.code.sf.net/p/apertium/svn/branches/apertium-swpost/apertium-en-es/es-tagger-data/es.tagged}
-containing 21865 words of Spanish news text, 
+containing approximately 22,000 words of Spanish news text, 
 and a small constraint grammar\footnote{https://svn.code.sf.net/p/apertium/svn/languages/apertium-spa/apertium-spa.spa.rlx}, produced independently of the authors.
@@
 We kept only select and remove rules, after which the grammar had 261 rules.
-With this setup, we took the text of the tagged corpus with all
-ambiguities, and ran both SAT-CG and VISL CG-3 with the same grammar. 
-Treating the corpus as the gold standard, the disambiguation by  VISL
-CG-3 achieves F-score of 82.6 \%, ordered
-SAT-CG 81.5 \%  and unordered SAT-CG 79.2 \%. 
-For comparison, with the best handcrafted grammar, unordered SAT-CG
-achieves 85.3 \% and VISL CG-3 85.9 \%.
-Similar patterns were observed with small (\textless{}20) rule sets
-written by the authors; depending on the subset, SAT-CG and VISL CG-3 had
-a difference of at most $\Mypm$ 1.5 \%. 
-Introducing rules one by one up to 19, the
-performance improved in a very similar rate, with less than 0.5 \%
-difference between the systems at each new rule.
-We did not evaluate on other languages or text genres, due to lack of suitable test data.
+With this setup, we produced an ambiguous version of the tagged
+corpus, and ran both SAT-CG and VISL CG-3 on it.
+Treating the original corpus as the gold standard, the disambiguation
+by  VISL CG-3 achieves F-score of 82.6 \%, ordered SAT-CG 81.5 \%  and
+unordered SAT-CG 79.2 \%. 
+We did not test with other languages or text genres due to lack of
+available gold standard.
+%This result suggests that that when running grammars that are written
+%with the traditional CG in mind, SAT-CG loses with both ordering
+%strategies, but emulating order fares better.
+
+We also wanted to test the hypothesis that SAT-CG needs different
+kinds of rules to perform well. We took the base grammar and started
+handcrafting rules; with our best performing and most concise
+grammar\footnote{https://github.com/inariksit/cgsat/blob/master/data/spa\_smallset.rlx}
+of just 19 rules, unordered SAT-CG achieves 85.3 \% and VISL CG-3 85.9
+\%. This experiment is naturally very small and might be explained by
+overfitting or mere chance, but it seems to indicate that rules that
+work well with SAT-CG, are also good for traditional CG.
+% It was possible to attain a rule set for which the unordered SAT-CG
+% works better (\textless0.5 \%), but it required trying
+% subsequences of the 19-rule set in a brute force manner---this hardly
+% presents any real life use case.
+
+% Similar patterns were observed with small (\textless{}20) rule sets
+% written by the authors; depending on the subset, SAT-CG and VISL CG-3 had
+% a difference of at most $\Mypm$ 1.5 \%. 
+% Introducing rules one by one up to 19, the
+% performance improved in a very similar rate, with less than 0.5 \%
+% difference between the systems at each new rule.
+% We did not evaluate on other languages or text genres, due to lack of suitable test data.
 
@@
 We tested the performance by parsing Don Quijote (384,155 words) with
-the same Spanish grammar. Table~\ref{table:time} shows execution times compared to VISL
+the same Spanish grammars as in previous experiment. Table~\ref{table:time} shows execution times compared to VISL
 CG-3 for the same text and rules. 
@@
 emulating order (SAT-CG\textsubscript{o}) is slightly faster, because maximisation problems are smaller.
-SAT is not Parsing Don Quijote (384,155 words) with the rule set of 261 rules,
-the function was called 147,253 times, and with 19 rules, 132,255 times.
+In any case, SAT does not seem to be the bottleneck: with 261 rules,
+the maximisation function was called 147,253 times, and with 19 rules,
+132,255 times. 
 %However, with both rule sets, half of the sentences in Don Quijote
@@
 things where the SAT-based representation can make an impact.
-One such thing is grammar analysis.
-There has been work on automatic tuning of hand-written CGs \todo{cite
-Eckhard}, but no tools which would search for inconsistencies or
-unoptimal design in order to point it out to the grammar writer.
+One such thing is grammar analysis\footnote{We thank Eckhard Bick for
+  the suggestion.}.
+There has been work on automatic tuning of hand-written CGs
+\cite{bick2013tuning}, but to our knowledge, no tools which would
+search for inconsistencies or unoptimal design in order to point it out to the grammar writer.
 
-Traditional CG rules are applied deterministically and
-sequentially; at the execution of rule, there is no way to
+The sequential application of traditional CG rules is good for
+performance and transparency. As soon as a rule applies, the analyses
+are removed from the input, and next rules get the modified sentence
+as their input. At the execution of rule, there is no way to
 go back to earlier rules and undo them. Thus there is no way to know
@@
 next rules impossible to apply, regardless of input.
-As a simple example, these rules could be
+We can reuse the example from Section~\ref{sec:unord}:
 
@@
 
-Maximisation-based conflict solving (see Section 3) is not feasible for this task. 
+Maximisation-based conflict solving (see Section~\ref{sec:unord}) is not feasible for this task. 
 The whole definition of conflicts depends on ordering, and it would

### 2015-06-01 paper/nodalida2015.tex
more fixes

@@
 result as with the SAT solver.
-Ideally, both ways of grammar writing should yield good results;
-traditional CG are more imperative, whereas SAT-CG
+Ideally, both ways of grammar writing should yield similar results;
+traditional CG rules are more imperative, and SAT-CG
 rules are more declarative.
@@
 unordered SAT-CG 79.2 \%. 
-We did not test with other languages or text genres due to lack of
+We did not test with other languages or text genres due to the lack of
 available gold standard.
@@
 We tested the performance by parsing Don Quijote (384,155 words) with
-the same Spanish grammars as in previous experiment. Table~\ref{table:time} shows execution times compared to VISL
-CG-3 for the same text and rules. 
-The number of rules in the grammar is more significant than the word
-count; this holds also for the performance of VISL CG-3.
-From the SAT solving side, maximisation is the most costly operation; 
-emulating order (SAT-CG\textsubscript{o}) is slightly faster, because maximisation problems are smaller.
+the same Spanish grammars as in the previous experiment. 
+Table~\ref{table:time} shows execution times compared to VISL CG-3;
+SAT-CG\textsubscript{U} is the unordered scheme and
+SAT-CG\textsubscript{O} is the ordered.
+% For both systems, the number of rules in the grammar affects the performance more than the word count.
+From the SAT solving side, maximisation is the most costly operation. 
+Emulating order is slightly faster, likely because the maximisation problems are smaller.
 In any case, SAT does not seem to be the bottleneck: with 261 rules,
 the maximisation function was called 147,253 times, and with 19 rules,
-132,255 times. 
+132,255 times, but 
 %However, with both rule sets, half of the sentences in Don Quijote
@@
 %needed less than 14.
-The difference in the execution times suggests that there are other
-reasons for the worse performance---this is to be expected,
-SAT-CG is currently just a naive proof-of-concept implementation with no optimisations.
+the differences in the execution times are much larger, which suggests
+that there are other reasons for the worse performance. 
+This is to be expected, SAT-CG is currently just a naive
+proof-of-concept implementation with no optimisations.
 
@@
 are removed from the input, and next rules get the modified sentence
-as their input. At the execution of rule, there is no way to
-go back to earlier rules and undo them. Thus there is no way to know
-what part of the input comes from the morphological analyser and what
-part from applying previous rules.
+as their input. 
+% At the execution of rule, there is no way to go back to earlier rules and undo them. 
+Thus there is no way to know what part of the input comes from the
+morphological analyser and what part from applying previous rules.
 
-A conflict can be defined as 
+A conflict in an ordered scheme can be defined as 
 a set of two or more rules, such that applying the first makes the
@@
 \begin{itemize}
-\item [] \texttt{SELECT verb ;} \\
-         \texttt{REMOVE verb IF (-1 det) ;}
+\item [] \texttt{SELECT v ;} \\
+         \texttt{REMOVE v IF (-1 det) ;}
 \end{itemize}
@@
 possible verb analyses, so there is a possibility for the SELECT rule
-to fire.
+to fire, as long as the input has verbs
 
-SAT-CG works for this kind of corpus-based grammar analysis as a side
-effect of its design. After applying each rule, it stores the clauses
-produced by it and commits to them---this meas that it retains the
+Ordered SAT-CG can be used to detect conflicts without any
+modifications, as a side effect of its design. 
+After applying each rule, it stores the clauses
+produced by it and commits to them---even if the new rules cannot
+change the earlier decisions, the SAT solver still retains the
 dependencies.
-Given the design that each rule application (instead of rule) gets a variable,
-in case of a conflict, the program prints out the particular
+Given that each application of a rule gets a variable,
+in case of a conflict, the program detects the particular
 application that violates the previous clauses, with the sentence
-where it is applied. 
+where it is applied. Thus we get feedback which rule fails, and on which
+particular word(s).
+
 This conflict solving can be done per section, or for the whole grammar. 
 In case of section-based application, we run each section on the
-input, and give the output as an input to the new section, discarding
+input, and give the output as a fresh input to the new section, discarding
 all clauses. Alternatively, we can let the rules interact with each
@@
 
-Maximisation-based conflict solving (see Section~\ref{sec:unord}) is not feasible for this task. 
+The unordered scheme with maximisation-based conflict solving (see Section~\ref{sec:unord}) is not feasible for this task. 
 The whole definition of conflicts depends on ordering, and it would
-not differentiate between the case where \texttt{SELECT verb} comes
-before \texttt{REMOVE verb IF (-1 det)} or after.
-On the other hand, a formalism such as FSIG would benefit from the
+not differentiate between the case where \texttt{SELECT v} comes
+before \texttt{REMOVE v IF (-1 det)} or after.
+On the other hand, an unordered formalism such as FSIG would benefit from the
 maximisation-based technique in conflict handling.
@@
 For future work, we intend to apply the SAT solver for corpus-free
-conflict testing. Let us illustrate the idea with the rule
-set shown in Section 3:
-
-\begin{itemize}
-\item [\texttt{1)}] \texttt{REMOVE verb IF (-1 det) ;} \\
-         \texttt{SELECT verb ;}
-
-\item [\texttt{2)}] \texttt{SELECT verb ;} \\
-         \texttt{REMOVE verb IF (-1 det) ;} 
-\end{itemize}
-
-Assume we have tag set \texttt{\{det, noun, verb, sg, pl\}}. Then we
+conflict testing. Let us illustrate the idea with the same two rules,
+\texttt{REMOVE v IF (-1 det)} and \texttt{SELECT v}
+in both orders.
+Assume we have tag set \texttt{\{det, n, v, sg, pl\}}. Then we
 can ask the SAT solver a question if there exists an input which
@@
 SAT solver tells us that with sentence length 2 and maximum 4 readings
-for the whole sentence (1+3, 2+2 or 3+1), there are several inputs that
+for the whole sentence, there are several inputs that
 satisfy the constraints for the first rule set, but none that
-satisfies the second rule set. Thus we can say that the first rule set
-is sane, but the second rule set is conflicting.
+satisfies the second rule set: for instance
 
+\begin{itemize}
+\item []
+\begin{verbatim}
+"<w1>"
+        det
+        v
+"<w2>"
+        n
+        v
+\end{verbatim}
+\end{itemize}
+Thus we can say that the first rule set is sane, but the second rule
+set is conflicting. 
 Implementing and testing this on a larger scale is left for future work.
@@
 We thank Tino Didriksen, Eckhard Bick, Francis Tyers and Anssi
-Yli-Jyr{\"a} for comments and suggestions, samt the anonymous
+Yli-Jyr{\"a} for comments and suggestions, as well as the anonymous
 reviewers and everyone who participated in the discussion at the CG workshop.

### 2015-06-01 paper/nodalida2015.tex
add abstract

@@
 
-%\begin{abstract} \end{abstract}
+\begin{abstract}
 
-\section{Introduction}
+We represent Constraint Grammar (CG) as a Boolean satisfiability (SAT) problem.
+Encoding CG in logic increases the computational complexity, 
+but brings some new features to the grammars.
+The rules are interpreted in a more declarative way, which makes
+cautious context and ordering irrelevant.
+One rule corresponds to many logical formulas,
+where each condition interpreted as a target---this makes the number of the rules
+potentially smaller. 
+Ordering can be preserved or discarded;
+in the latter case, we solve eventual rule conflicts by finding a
+solution that discards the least number of rule applications.
+We test our implementation by parsing texts in the order of
+10,000s--100,000s words, using grammars with hundreds of rules.
+\end{abstract}
 
-We represent Constraint Grammar (CG) \cite{karlsson1995constraint} 
-as a Boolean satisfiability (SAT) problem.
+\section{Introduction and previous research}
+
+\todo{make this nicer: Constraint Grammar (CG)
+  \cite{karlsson1995constraint} is a relatively young formalism.
+In this work, we implement CG using a SAT solver,
+thus relating CG to a wider context in the theory of computation.
 This is attractive from several reasons: formal logic is
@@
 properties of CG. Constraint rules encoded in logic capture richer
-dependencies between the tags than standard CG.
+dependencies between the tags than standard CG.}
 
@@
 
-Encoding CG in logic increases computational complexity, 
-but as a tradeoff, it could simplify the grammar writing.
-Due to representing the rules as implications, cautious context
-condition becomes irrelevant.
-One rule corresponds to many logical formulas (each condition being
-the target), which makes the number of the rules potentially smaller.
-Ordering can be preserved or discarded; in the latter case, we solve
-eventual rule conflicts by finding a solution that discards the least
-number of rule applications.
-% our experiments show that it is possible to write a small grammar, 
-% which, when run with SAT-CG, disambiguates more than the same
-%grammar run with VISL CG-3.
-
-We test our implementation by parsing texts in the order of
-10,000s--100,000s words, using grammars with hundreds of rules.
-
-%SAT solving is based on a technique called unit propagation:
-%a set of complex Boolean clauses is simplified, starting from unit
-%clauses, which consist of just a single variable, and working up to a
-%solution, where all variables in the set have a True or False value.
-
-\section{Related work}
-\label{sect:related}
-
-Our work is inspired by \cite{lager98}, which presents constraint rules as a disjunctive logic program,
+\todo{Our work is inspired by \cite{lager98}, which presents constraint rules as a disjunctive logic program,
 and \cite{lager_nivre01}, which reconstructs 4 different formalisms from a logical point of view.
@@
 rules, it resembles Finite-State Intersection Grammar \cite{koskenniemi90}.
-See also \cite{lager01transformation} for discussion of ordering of the CG rules.
+See also \cite{lager01transformation} for discussion of ordering of
+the CG rules.}
+
 \section{CG as a SAT problem}
@@
 
-\section{Results and evaluation}
+\section{Evaluation}
 \label{sec:eval}
@@
 Table~\ref{table:time} shows execution times compared to VISL CG-3;
-SAT-CG\textsubscript{U} is the unordered scheme and
-SAT-CG\textsubscript{O} is the ordered.
+SAT-CG\textsubscript{u} is the unordered scheme and
+SAT-CG\textsubscript{o} is the ordered.
 % For both systems, the number of rules in the grammar affects the performance more than the word count.
@@
 
-\section{Applications and Future Work}
+\section{Applications}
 \label{sec:apps}
@@
 The sequential application of traditional CG rules is good for
-performance and transparency. As soon as a rule applies, the analyses
+performance and transparency. As soon as a rule takes action, the analyses
 are removed from the input, and next rules get the modified sentence
@@
 % At the execution of rule, there is no way to go back to earlier rules and undo them. 
-Thus there is no way to know what part of the input comes from the
-morphological analyser and what part from applying previous rules.
+Thus there is no way to know which part comes directly from the
+raw input and which part from applying previous rules.
 

### 2015-06-01 paper/nodalida2015.tex
small fixes

@@
 
-\todo{make this nicer: Constraint Grammar (CG)
-  \cite{karlsson1995constraint} is a relatively young formalism.
-In this work, we implement CG using a SAT solver,
-thus relating CG to a wider context in the theory of computation.
+Constraint Grammar (CG) \cite{karlsson1995constraint} is a relatively
+young formalism. It belongs to the reductionist family of parsing
+strategies, along with Finite-State Intersection Grammar (FSIG) \cite{koskenniemi90}.
+In this work, we use a SAT solver to implement a version of CG that is
+roughly compatible with CG-2.
 This is attractive from several reasons: formal logic is
@@
 properties of CG. Constraint rules encoded in logic capture richer
-dependencies between the tags than standard CG.}
+dependencies between the tags than standard CG.
 
@@
 
-\todo{Our work is inspired by \cite{lager98}, which presents constraint rules as a disjunctive logic program,
-and \cite{lager_nivre01}, which reconstructs 4 different formalisms from a logical point of view.
-\cite{lindberg_eineborg98ilp,asfrent14} use Inductive Logic Programming to learn CG rules from a tagged corpus.
+Our work is inspired by \cite{lager98}, which presents constraint
+rules as a disjunctive logic program, and \cite{lager_nivre01}, which
+reconstructs four different formalisms in first-order logic.
+\cite{lindberg_eineborg98ilp} and \cite{asfrent14} use Inductive Logic Programming to learn CG rules from a tagged corpus.
 We explore different approaches to ordering. In the case of unordered
@@
 See also \cite{lager01transformation} for discussion of ordering of
-the CG rules.}
+the CG rules.
 
@@
 
-We thank Tino Didriksen, Eckhard Bick, Francis Tyers and Anssi
+We thank Eckhard Bick, Tino Didriksen, Francis Tyers and Anssi
 Yli-Jyr{\"a} for comments and suggestions, as well as the anonymous

### 2015-06-01 paper/nodalida2015.tex
yet more small fixes

@@
 Constraint Grammar (CG) \cite{karlsson1995constraint} is a relatively
-young formalism. It belongs to the reductionist family of parsing
-strategies, along with Finite-State Intersection Grammar (FSIG) \cite{koskenniemi90}.
-In this work, we use a SAT solver to implement a version of CG that is
-roughly compatible with CG-2.
-This is attractive from several reasons: formal logic is
+young formalism, born out of empirical linguistic approach. 
+It belongs to the reductionist family of parsing
+strategies, along with e.g. Finite-State Intersection Grammar (FSIG) \cite{koskenniemi90}.
+In this work, we implement CG\footnote{Roughly compatible with CG-2}
+using a SAT solver. This is attractive from several reasons: formal logic is
 well-studied, and serves as an abstract language to reason about the
@@
 
-Our work is inspired by \cite{lager98}, which presents constraint
+Our work is primarily inspired by \cite{lager98}, which presents constraint
 rules as a disjunctive logic program, and \cite{lager_nivre01}, which
 reconstructs four different formalisms in first-order logic.
-\cite{lindberg_eineborg98ilp} and \cite{asfrent14} use Inductive Logic Programming to learn CG rules from a tagged corpus.
-We explore different approaches to ordering. In the case of unordered
-rules, it resembles Finite-State Intersection Grammar \cite{koskenniemi90}.
-See also \cite{lager01transformation} for discussion of ordering of
-the CG rules.
+Other works combining logic to CG include
+\cite{lindberg_eineborg98ilp} and \cite{asfrent14}, both using Inductive Logic Programming to learn CG rules from a tagged corpus.
+%We explore different approaches to ordering. 
+%In the case of unordered rules, it resembles Finite-State Intersection Grammar \cite{koskenniemi90}.
+%See also \cite{lager01transformation} for discussion of ordering of the CG rules.
 

### 2015-06-01 paper/nodalida2015.tex
same

@@
 We represent Constraint Grammar (CG) as a Boolean satisfiability (SAT) problem.
-Encoding CG in logic increases the computational complexity, 
-but brings some new features to the grammars.
+Encoding CG in logic brings some new features to the grammars.
 The rules are interpreted in a more declarative way, which makes
@@
 Constraint Grammar (CG) \cite{karlsson1995constraint} is a relatively
-young formalism, born out of empirical linguistic approach. 
-It belongs to the reductionist family of parsing
-strategies, along with e.g. Finite-State Intersection Grammar (FSIG) \cite{koskenniemi90}.
-In this work, we implement CG\footnote{Roughly compatible with CG-2}
+young formalism, born out of practical need for a robust and
+language-independent method for part-of-speech tagging.
+%It belongs to the reductionist family of parsing
+%strategies, along with e.g. Finite-State Intersection Grammar (FSIG) \cite{koskenniemi90}.
+In this work, we present CG %\footnote{Roughly compatible with CG-2}
+as a Boolean satisfiability (SAT) problem, and describe an implementation
 using a SAT solver. This is attractive from several reasons: formal logic is
@@
 
-Standard CG will apply one\footnote{To be more cautious, we could require the word at position +1 be disambiguated fully (\texttt{1C} instead of \texttt{1}), but in that case, 
-neither of the rules would be applied.} of the rules to the word \emph{la}; 
+Standard CG will apply one  of the rules to the word \emph{la}; 
 either the one that comes first, or by some other heuristic. 
-The second rule will not fire, because it would remove the last reading. 
-All readings of \emph{casa} are left untouched by these rules.
+The second rule will not fire, because it would remove the last
+reading. 
+If we use the cautious mode (\texttt{1C n} or \texttt{1C v}), which
+ requires the word in the context to be fully disambiguated, 
+neither of the rules will be applied.
+In any case, all readings of \emph{casa} are left untouched by these rules.
 
@@
 discard the model where \emph{casa} is a verb and \emph{la} is a
-pronoun---all these decisions are connected---and we get a unique solution with \texttt{det n adj}.
+pronoun
+%---all these decisions are connected---
+and we get a unique solution with \texttt{det n adj}.
 % Now \emph{casa}\texttt{<n>} is true, so \emph{la}\texttt{<prn>} must be false. 
@@
 before \texttt{REMOVE v IF (-1 det)} or after.
-On the other hand, an unordered formalism such as FSIG would benefit from the
+On the other hand, an unordered formalism such as Finite-State
+Intersection Grammar \cite{koskenniemi90} would benefit from the
 maximisation-based technique in conflict handling.

### 2015-06-01 paper/nodalida2015.tex
stuff

@@
 
-In the following sections, we report both execution time and
-performance against state of the art.
+For evaluation, we measure the performance against the state of the
+art CG parser VISL CG-3.
+% both in terms of accuracy and execution time.
+SAT-CG fares slightly worse for accuracy, and significantly worse for execution time.
+The results are presented in more detail in the following sections.
 

### 2015-06-01 paper/nodalida2015.tex
small stuff

@@
 
-\subsection{Time}
+\subsection{Execution time}
 
@@
 \label{sec:apps}
-As described in Section~\ref{sec:eval}, our implementation is not
+As demonstrated in Section~\ref{sec:eval}, our implementation is not
 competetive with the state of the art. It is more fruitful to focus on
@@
 possible verb analyses, so there is a possibility for the SELECT rule
-to fire, as long as the input has verbs
+to fire.
 
-Ordered SAT-CG can be used to detect conflicts without any
+Ordered SAT-CG can be used to detect these conflicts without any
 modifications, as a side effect of its design. 
 After applying each rule, it stores the clauses
-produced by it and commits to them---even if the new rules cannot
-change the earlier decisions, the SAT solver still retains the
-dependencies.
-Given that each application of a rule gets a variable,
-in case of a conflict, the program detects the particular
-application that violates the previous clauses, with the sentence
+produced by it and commits to them
+% ---even if the new rules cannot
+% change the earlier decisions, the SAT solver still retains the
+% dependencies.
+% Given that each application of a rule 
+% is 
+% gets a variable,
+In case of a conflict, the program detects the particular rule that 
+violates the previous clauses, with the sentence
 where it is applied. Thus we get feedback which rule fails, and on which

### 2015-06-01 paper/nodalida2015.tex
stuff

@@
 The sequential application of traditional CG rules is good for
-performance and transparency. As soon as a rule takes action, the analyses
-are removed from the input, and next rules get the modified sentence
+performance and transparency. When a rule takes action, the analyses
+are removed from the input, and the next rules get the modified sentence
 as their input. 
 % At the execution of rule, there is no way to go back to earlier rules and undo them. 
-Thus there is no way to know which part comes directly from the
+As a downside, there is no way to know which part comes directly from the
 raw input and which part from applying previous rules.
@@
 a set of two or more rules, such that applying the first makes the
-next rules impossible to apply, regardless of input.
+next rules impossible to apply, regardless of the input.
 We can reuse the example from Section~\ref{sec:unord}:
@@
 
-The first rule selects the verb always, removing all other readings,
-and the second rule removes verb reading, given certain condition.
+The first rule selects the verb reading always, removing all other readings,
+and the second rule removes the verb reading conditionally.
 If the rules are introduced in a different order, there is no
@@
 After applying each rule, it stores the clauses
-produced by it and commits to them
-% ---even if the new rules cannot
-% change the earlier decisions, the SAT solver still retains the
-% dependencies.
-% Given that each application of a rule 
-% is 
-% gets a variable,
+produced by the rule and commits to them.
 In case of a conflict, the program detects the particular rule that 
@@
 
-This conflict solving can be done per section, or for the whole grammar. 
-In case of section-based application, we run each section on the
-input, and give the output as a fresh input to the new section, discarding
-all clauses. Alternatively, we can let the rules interact with each
-other regardless of the section boundaries.
+% This conflict solving can be done per section, or for the whole grammar. 
+% In case of section-based application, we run each section on the
+% input, and give the output as a fresh input to the new section, discarding
+% all clauses. Alternatively, we can let the rules interact with each
+% other regardless of the section boundaries.
 
-The unordered scheme with maximisation-based conflict solving (see Section~\ref{sec:unord}) is not feasible for this task. 
-The whole definition of conflicts depends on ordering, and it would
-not differentiate between the case where \texttt{SELECT v} comes
-before \texttt{REMOVE v IF (-1 det)} or after.
-On the other hand, an unordered formalism such as Finite-State
-Intersection Grammar \cite{koskenniemi90} would benefit from the
+Unordered SAT-CG with maximisation-based conflict solving is not
+suitable for this task: the whole definition of conflict depends on
+ordering, and the unordered scheme deliberately loses this information.
+On a more speculative note, an unordered formalism such as Finite-State
+Intersection Grammar \cite{koskenniemi90} might benefit from the
 maximisation-based technique in conflict handling.

### 2015-06-01 paper/nodalida2015.tex
stuff

@@
 as a Boolean satisfiability (SAT) problem, and describe an implementation
-using a SAT solver. This is attractive from several reasons: formal logic is
+using a SAT solver. This is attractive for several reasons: formal logic is
 well-studied, and serves as an abstract language to reason about the
@@
 following rule sets would be treated identically, whereas in the
-traditional CG, only the second would be considered as a bad order.
+traditional CG, only the first would be considered as a bad order.
 
 \begin{itemize}
-\item [\texttt{1)}] \texttt{REMOVE v IF (-1 det) ;} \\
+\item [\texttt{1)}] \texttt{SELECT v ;} \\
+         \texttt{REMOVE v IF (-1 det) ;} 
+\item [\texttt{2)}] \texttt{REMOVE v IF (-1 det) ;} \\
          \texttt{SELECT v ;}
-\item [\texttt{2)}] \texttt{SELECT v ;} \\
-         \texttt{REMOVE v IF (-1 det) ;} 
 \end{itemize}
@@
 produced by each rule one by one, and assume the solver state reached
-so far is correct. If a conflict is introduced by new clauses, discard
-them and move on to the next rule.
+so far is correct. If a new clause introduces a conflict with
+previous clauses, we discard it and move on to the next rule.
 By testing against gold standard, we see that this scheme works better
@@
 
-For future work, we intend to apply the SAT solver for corpus-free
-conflict testing. Let us illustrate the idea with the same two rules,
-\texttt{REMOVE v IF (-1 det)} and \texttt{SELECT v}
+As future work, we intend to test for conflicts
+without using a corpus.
+Let us illustrate the idea with the same two rules,
+\texttt{SELECT v} and \texttt{REMOVE v IF (-1 det)}
 in both orders.
-Assume we have tag set \texttt{\{det, n, v, sg, pl\}}. Then we
-can ask the SAT solver a question if there exists an input which
-satisfies our constraints: both rules, applied in the order, remove
-something from the input.
-SAT solver tells us that with sentence length 2 and maximum 4 readings
-for the whole sentence, there are several inputs that
-satisfy the constraints for the first rule set, but none that
-satisfies the second rule set: for instance
+Assume we have the tag set \texttt{\{det, n, v, sg, pl\}}, 
+and we want to find if there exists an input such that
+ both rules, applied in the given order, remove something from the input.
+% This is another constraint solving problem, hence we can model it in SAT.
+There are no inputs that satisfy the requirement with first order,
+but several that work with the second, for instance:
 
@@
 advantages over more traditional approaches; a SAT-solver may
-disambiguate more words, and may do so more precisely. Also, the
-SAT-solver requires less rules, and these rules are simpler. 
-\todo{proper conclusions for the new stuff}
+disambiguate more words, and may do so more precisely, capturing
+more dependencies between tags.
+We experimented with both ordered and unordered rules, and found the ordered
+scheme to work better with previously written grammars.
+
+% Also, the SAT-solver requires less rules, and these rules are simpler. 
 % The ordering of rules is something we found was incompatible with a logic-based approach.

### 2015-06-01 paper/nodalida2015.tex
stuff

@@
 
-We represent Constraint Grammar (CG) as a Boolean satisfiability (SAT) problem.
+We represent Constraint Grammar (CG) as a Boolean satisfiability 
+(SAT) problem.
 Encoding CG in logic brings some new features to the grammars.
-The rules are interpreted in a more declarative way, which makes
-cautious context and ordering irrelevant.
-One rule corresponds to many logical formulas,
-where each condition interpreted as a target---this makes the number of the rules
-potentially smaller. 
+The rules are interpreted in a more declarative way, which makes it
+possible to abstract away from details such as cautious context and
+ordering.
+The encoding in logic captures richer dependencies between the tags:
+a rule is allowed to take action also concerning the words in its
+context. This makes the number of the rules potentially smaller. 
 Ordering can be preserved or discarded;
@@
 There are no inputs that satisfy the requirement with first order,
-but several that work with the second, for instance:
+but several that work with the second, for instance
 
@@
 Constraint Grammar. A logic-based approach to CG has possible
-advantages over more traditional approaches; a SAT-solver may
+advantages over more traditional approaches; a SAT solver may
 disambiguate more words, and may do so more precisely, capturing
@@
 scheme to work better with previously written grammars.
+The evaluation shows that SAT-CG isn't a match for
+As future work, we 
+As a complementary approach, we intend to concentrate on grammar analysis
 

### 2015-06-01 paper/nodalida2015.tex
stuff

@@
 ordering.
-The encoding in logic captures richer dependencies between the tags:
-a rule is allowed to take action also concerning the words in its
-context. This makes the number of the rules potentially smaller. 
+%In addition to the target words, 
+A rule is allowed to affect its context words, which makes the number
+of the rules in a grammar potentially smaller. 
 Ordering can be preserved or discarded;
@@
 biology---these problems easily match or exceed CG in complexity. 
-Due to these advances, we were able to revisit the idea and develop it
+Thanks to these advances, we were able to revisit the idea and develop it
 further.
@@
 
-With these constraints, we get two solutions. The interaction of the rules regarding \emph{la}  disambiguates \emph{casa} for free, and ordering of the rules doesn't matter.
+With these constraints, we get two solutions. The interaction of the
+rules regarding \emph{la}  disambiguates the part of speech of
+\emph{casa} for free, and the order of the rules does not matter.
 
@@
 some rules fire during the first iteration, either because their
-conditions don't require cautious context, or because some words are
+conditions do not require cautious context, or because some words are
 unambiguous to start with. This makes some more words unambiguous, and
@@
 
-As hinted by the previous property, SAT solver doesn't need a fixed
+As hinted by the previous property, SAT solver does not need a fixed
 order of the rules.
@@
 
-In case of a conflict, SAT solver makes it possible to discard only
+In case of a conflict, the SAT solver makes it possible to discard only
 minimal amount of rule applications. Continuing with the example,
@@
 
-%With the added interaction between rules, catch-all rules might even be less 
-%necessary. However, this is just speculation, we have not tested this effect.
-
-%We can apply this for the whole grammar, or for one section at a time.
-
-%The design of unordered rules can lead to new ways of writing CGs,
-%potentially with less mental effort to the writer, but also less transparent.
-
-%By testing against gold standard, we see that the
-%unordered scheme not optimal for applying ready-made
-%CGs, where the order is a crucial detail.
-
 We can also emulate order with SAT-CG. To do that, we enter clauses
@@
 It also runs slightly faster than the unordered version.
-%the performance, because the maximisation problems are smaller. 
-%We have also experimented with sections, and 
 \\
 
-%  that it is possible to write a small grammar, 
-% which, when run with the unordered scheme, disambiguates more than the
-% same grammar run with VISL CG-3. 
-% However, when tested with grammars that are written for the original
-% CG, SAT-CG loses with both strategies, but emulating order fares better.
-
-% Finite-state implementations of CG \cite{koskenniemi92} also have logically unordered rules. 
-% \todo{Inari: check if Tapanainen 1996 discuss ordering}
-
-\noindent These three features influence the way that rules are written. 
+\noindent These three features influence the way rules are written. 
 We predict that less rules are needed; whether this holds in the order
-of thousands of rules remains to be tested. On one hand, getting rid
+of thousands of rules remains to be tested. On the one hand, getting rid
 of ordering and cautious context could ease the task of the grammar
 writer, since it removes the burden of estimating the best sequence of
-rules and whether to make the them cautious. On the other
+rules and whether to make them cautious. On the other
 hand, lack of order can make the rules less transparent, and might not scale up for larger grammars.
@@
 % We ignored substitute rules, but kept subreadings and unification, even though our implementation doesn't handle them.
-We kept only select and remove rules, after which the grammar had 261 rules.
+We kept only SELECT and REMOVE rules, which left us 261 rules.
 With this setup, we produced an ambiguous version of the tagged
@@
 
-We also wanted to test the hypothesis that SAT-CG needs different
-kinds of rules to perform well. We took the base grammar and started
+We also wanted to test the hypothesis that SAT-CG works better with
+small amount of rules. We took the base grammar and started
 handcrafting rules; with our best performing and most concise
 grammar\footnote{https://github.com/inariksit/cgsat/blob/master/data/spa\_smallset.rlx}
-of just 19 rules, unordered SAT-CG achieves 85.3 \% and VISL CG-3 85.9
+of only 19 rules, unordered SAT-CG achieves 85.3 \% and VISL CG-3 85.9
 \%. This experiment is naturally very small and might be explained by
@@
 implementations of CG are polynomial, but with advances in SAT solving
-techniques, the performance in the average case in practice is much more feasible than in the previous works done in 90s--00s.
+techniques, the performance in the average case in practice is more feasible than in the previous works done in 90s--00s.
 We used the open-source SAT solver MiniSat \cite{een04sat}.
@@
 that there are other reasons for the worse performance. 
-This is to be expected, SAT-CG is currently just a naive
+This is to be expected, as SAT-CG is currently just a naive
 proof-of-concept implementation with no optimisations.
@@
 performance and transparency. When a rule takes action, the analyses
-are removed from the input, and the next rules get the modified sentence
-as their input. 
+are removed from the sentence, and the next rules get the modified
+sentence as input. 
 % At the execution of rule, there is no way to go back to earlier rules and undo them. 
@@
 
-The first rule selects the verb reading always, removing all other readings,
-and the second rule removes the verb reading conditionally.
+The first rule selects the verb reading everywhere and removes
+all other readings, leaving no chance for the second rule to take action.
 If the rules are introduced in a different order, there is no
@@
 % This is another constraint solving problem, hence we can model it in SAT.
-There are no inputs that satisfy the requirement with first order,
-but several that work with the second, for instance
+There are no inputs that satisfy the requirement with the first order,
+but several that work with the second, such as the following:
 
@@
 \end{itemize}
-Thus we can say that the first rule set is sane, but the second rule
-set is conflicting. 
+Thus we can say that the first rule order is conflicting, but the second
+one is not. 
 Implementing and testing this on a larger scale is left for future work.
@@
 more dependencies between tags.
-We experimented with both ordered and unordered rules, and found the ordered
-scheme to work better with previously written grammars.
-The evaluation shows that SAT-CG isn't a match for
-As future work, we 
-As a complementary approach, we intend to concentrate on grammar analysis
+We experimented with both ordered and unordered rules, and found the 
+ordered scheme to work better with previously written grammars.
+For future direction, we intend to concentrate on grammar analysis,
+especially finding conflicts in grammars.
 

### 2015-06-01 paper/nodalida2015.tex
typos + using hyperref and smallcaps

@@
 \usepackage{cite}
+\usepackage{hyperref}
 \usepackage{authordate1-4}
@@
 
-Traditional CG relies on that applying the rule set iteratively:
+Traditional CG applies the rule set iteratively:
 some rules fire during the first iteration, either because their
@@
 it is not clear which instances would be discarded, but if the rules
-were part of a larger rule set, and in the context the REMOVE rule was
+were part of a larger rule set, and in the context the \textsc{remove} rule was
 the right one to choose, it is likely that the interaction between the desired
 rules would make a large set of clauses that fit together, and the
-SELECT rule would not fit in, hence it would be discarded.
+\textsc{select} rule would not fit in, hence it would be discarded.
 
@@
 We took a manually tagged
-corpus\footnote{https://svn.code.sf.net/p/apertium/svn/branches/apertium-swpost/apertium-en-es/es-tagger-data/es.tagged}
+corpus\footnote{\url{https://svn.code.sf.net/p/apertium/svn/branches/apertium-swpost/apertium-en-es/es-tagger-data/es.tagged}}
 containing approximately 22,000 words of Spanish news text, 
-and a small constraint grammar\footnote{https://svn.code.sf.net/p/apertium/svn/languages/apertium-spa/apertium-spa.spa.rlx}, produced independently of the authors.
+and a small constraint grammar\footnote{\url{https://svn.code.sf.net/p/apertium/svn/languages/apertium-spa/apertium-spa.spa.rlx}}, produced independently of the authors.
 % We ignored substitute rules, but kept subreadings and unification, even though our implementation doesn't handle them.
-We kept only SELECT and REMOVE rules, which left us 261 rules.
+We kept only \textsc{select} and \textsc{remove} rules, which left us 261 rules.
 With this setup, we produced an ambiguous version of the tagged
@@
 handcrafting rules; with our best performing and most concise
-grammar\footnote{https://github.com/inariksit/cgsat/blob/master/data/spa\_smallset.rlx}
+grammar\footnote{\url{https://github.com/inariksit/cgsat/blob/master/data/spa\_smallset.rlx}}
 of only 19 rules, unordered SAT-CG achieves 85.3 \% and VISL CG-3 85.9
@@
 As demonstrated in Section~\ref{sec:eval}, our implementation is not
-competetive with the state of the art. It is more fruitful to focus on
+competitive with the state of the art. It is more fruitful to focus on
 things where the SAT-based representation can make an impact.
@@
 If the rules are introduced in a different order, there is no
-conflict: the REMOVE rule would not remove verb readings from all
-possible verb analyses, so there is a possibility for the SELECT rule
+conflict: the \textsc{remove} rule would not remove verb readings from all
+possible verb analyses, so there is a possibility for the \textsc{select} rule
 to fire.
@@
 in both orders.
-Assume we have the tag set \texttt{\{det, n, v, sg, pl\}}, 
+Assume we have the tag set \texttt{\{det, n, v\}}, 
 and we want to find if there exists an input such that
@@
 For future direction, we intend to concentrate on grammar analysis,
-especially finding conflicts in grammars.
+especially finding conflicts in constraint grammars.
 

### 2015-06-01 paper/nodalida2015.tex
fix hyperref

@@
 \usepackage{cite}
-\usepackage{hyperref}
+\usepackage[colorlinks=true]{hyperref}
 \usepackage{authordate1-4}
@@
 
-\section{Applications}
+\section{Applications and future work}
 \label{sec:apps}
-As demonstrated in Section~\ref{sec:eval}, our implementation is not
-competitive with the state of the art. It is more fruitful to focus on
-things where the SAT-based representation can make an impact.
-One such thing is grammar analysis\footnote{We thank Eckhard Bick for
-  the suggestion.}.
+%As demonstrated in Section~\ref{sec:eval}, our implementation is not
+%competitive with the state of the art. 
+%VISL CG-3 is efficient, reliable and widely used, and
+Instead of trying to compete with the state of the art, we plan to use SAT-CG for grammar analysis\footnote{We thank Eckhard Bick for the idea.}.
 There has been work on automatic tuning of hand-written CGs
-\cite{bick2013tuning}, but to our knowledge, no tools which would
-search for inconsistencies or unoptimal design in order to point it out to the grammar writer.
+\cite{bick2013tuning}, but to our knowledge no tools which would
+search for inconsistencies or unoptimal design. %in order to point it out to the grammar writer.
 
@@
 
-As future work, we intend to test for conflicts
+Finally, we intend to test for conflicts
 without using a corpus.

### 2015-06-01 paper/nodalida2015.tex
change colours of links and other important stuff

@@
 \usepackage{cite}
-\usepackage[colorlinks=true]{hyperref}
+\usepackage[colorlinks=true,citecolor=blue,urlcolor=blue]{hyperref}
 \usepackage{authordate1-4}
@@
 either the one that comes first, or by some other heuristic. 
-The second rule will not fire, because it would remove the last
+The other rule will not fire, because it would remove the last
 reading. 
@@
 
-As hinted by the previous property, SAT solver does not need a fixed
+As hinted by the previous property, the SAT solver does not need a fixed
 order of the rules.
@@
 
-For evaluation, we measure the performance against the state of the
-art CG parser VISL CG-3.
+For evaluation, we measure the performance against the state-of-the-art CG parser VISL CG-3.
 % both in terms of accuracy and execution time.
@@
 
-We also wanted to test the hypothesis that SAT-CG works better with
-small amount of rules. We took the base grammar and started
-handcrafting rules; with our best performing and most concise
+We also wanted to test whether SAT-CG outperforms traditional CG with a
+small rule set. With our best performing and most concise
 grammar\footnote{\url{https://github.com/inariksit/cgsat/blob/master/data/spa\_smallset.rlx}}
-of only 19 rules, unordered SAT-CG achieves 85.3 \% and VISL CG-3 85.9
-\%. This experiment is naturally very small and might be explained by
+of only 19 rules, both SAT-CG and VISL CG-3  achieve a F-score of
+around 85 \%. This experiment is very small and might be explained by
 overfitting or mere chance, but it seems to indicate that rules that
-work well with SAT-CG, are also good for traditional CG.
+work well with SAT-CG are also good for traditional CG.
 % It was possible to attain a rule set for which the unordered SAT-CG
@@
 There has been work on automatic tuning of hand-written CGs
-\cite{bick2013tuning}, but to our knowledge no tools which would
-search for inconsistencies or unoptimal design. %in order to point it out to the grammar writer.
+\cite{bick2013tuning}, but to our knowledge no tools to
+search for inconsistencies or suboptimal design. %in order to point it out to the grammar writer.
 

### 2015-08-19 CG_SAT.hs
removed old crap & started rewriting Symbolic; works better than it used to, except cautious

@@ -99,7 +99,7 @@ chunk sent = concat $ go sent 1
 dechunk :: [Token] -> Sentence
 dechunk ts = (map.map) getTags (groupBy sameInd ts)
         
-  
+
 --------------------------------------------------------------------------------
 
 -- | For each position, at least one analysis must be true.
@@ -124,7 +124,7 @@ applyRule s toks rule = do
     (Select _name target conds) -> go s True False (toTags target) (toConds conds) toks
 
     
-go :: Solver -> Bool -> Bool -> [[Tag]] -> [[Condition]] -> [Token] -> IO [[Lit]]
+go :: Solver -> Bool -> Bool -> [[Tag]] -> [[Condition]] -> [Token] -> IO [Clause]
 go s isSelect isGrAna target []         allToks = return []
 go s isSelect isGrAna target (conds:cs) allToks = do
    let f = if isGrAna then mkVarsGrammarAnalysis else mkVars
@@ -171,7 +171,7 @@ go s isSelect isGrAna target (conds:cs) allToks = do
   --As before, we collapse multiple conditions into one SAT variable and create clauses.
   --In addition, stuff only for grammar analysis:
   --We want to find a sentence for which a rule /does not/ fire.
-  --It can be because of 
+  --It can be because of there is no target left, or because there is /only/ target left.
   mkVarsGrammarAnalysis :: [(Token,[[Token]])] -> [(Token,[[Token]])] -> [Condition] ->  IO [Clause]
   mkVarsGrammarAnalysis sl rm conds = do
     let onlyCtxs = [ ctx | tctx <- groupBy sndEq (sl++rm)
@@ -179,35 +179,66 @@ go s isSelect isGrAna target (conds:cs) allToks = do
     rs <- sequence [ help s ctx true conds | ctx <- onlyCtxs ]
     let ctxMap = zip onlyCtxs rs
 
-    rmLits <- sequence [ newLit s | _ <- rm ] --for the "only target left" case
-    slLits <- sequence [ newLit s | _ <- sl ]
 
-                        --r => ~tl || nl (which denotes the onlyTrg option)
-    let rmClauses = [ [neg r, neg tl, nl]:onlyTrg
-                        | (nl, (trg, ctx)) <- zip rmLits rm
-                        , let r = fromJust $ lookup ctx ctxMap
+    --should onlyTrgCls only apply to the original target of the rule that is analysed?
+    --no  : a rule cannot apply selectively if context is met
[…]

### 2015-08-19 executable/Symbolic.hs

@@ -29,14 +29,15 @@ lookup' tagcombs tagsInCond = --trace ("lookup': " ++ show tagsInCond) $
 testrules = concat $ snd $ parseRules False "SELECT:r1 (aa) OR (acr) IF (-1C (mf)) ;\nREMOVE:r2 (aa) IF (-1C (acr)) ;"
 
 main = do
-  ts <- filter (not.null) `fmap` map parse `fmap` words `fmap` readFile tagfile :: IO [[Tag]]
-  print ts
-  tc <- map parse `fmap` words `fmap` readFile tagcfile
+  tc <- (map parse . words) `fmap` readFile tagcfile
 
   args <- getArgs
   case args of
    []    -> do putStrLn "test"
-               let spl = init $ splits testrules
+               ts <- (filter (not.null) . map parse . words) `fmap` readFile "data/spa_tags.txt"
+  
+               print ts
+               let spl = splits testrules
                print spl
                (tsets, _) <- readRules' "data/spa_smallset.rlx"
                let tc = nub $ ts ++ concatMap toTags tsets
@@ -46,24 +47,42 @@ main = do
                mapM_ ((flip checkCorpus) corpus) tokens
 
 
-   (r:o) -> do let verbose = "v" `elem` o
+   ("tiny":_)
+         -> do let ts = map ((:[]) . Tag) ["det","v","n"] ++ [[Tag "adj", Tag "pred"], [Tag "adj", Tag "attr"]]
+               --let rules = concat $ snd $ parseRules False "REMOVE:r1 (v) IF (-1C (det)) ; REMOVE:r2 (adj) ; REMOVE:r3 (n) IF (-1C (det));"
+               let rules = concat $ snd $ parseRules False "REMOVE:r1 (v) IF (-1 (det)) (0 (n)) ; REMOVE:r2 (adj); REMOVE:r3 (n) IF (-1 (det));"
+               let spl = splits $ reverse rules
+               let tc = ts
+               tokens <- mapM (uncurry (testRule True ts tc)) spl
+               --mapM_ putStrLn (map (showSentence . dechunk) tokens)
+               putStrLn "\n---------\n"
 
-               (tsets, rls) <- readRules' r
+   ("nld":r)
+         -> do let verbose = "v" `elem` r
+               ts <- (filter (not.null) . map parse . words) `fmap` readFile "data/nld_tags.txt"
+  
+               (tsets, rls) <- readRules' "data/nld.rlx"
                let rules = concat rls
                mapM_ print rules
[…]

### 2015-08-19 CG_base.hs
stuff

@@ -43,7 +43,7 @@ data TagSet =
 instance Show TagSet where
   show (TS tags) = showTagset tags
   show (Or ts1 ts2) = show ts1 ++ " OR " ++ show ts2
-  show (Diff ts1 ts2) = show ts1 ++ " - " ++ show ts2
+  show (Diff ts1 ts2) = show ts1 ++ " - " ++ show ts2 
   show (Cart ts1 ts2) = show ts1 ++ " + " ++ show ts2
   show All = "(*)"
 
@@ -65,7 +65,7 @@ toTags :: TagSet -> [[Tag]]
 
 toTags (TS tags) = tags
 toTags (Or ts1 ts2) = toTags ts1 ++ toTags ts2
-toTags (Diff ts1 ts2) = toTags ts1 \\ toTags ts2 
+toTags (Diff ts1 ts2) = toTags ts1 \\ toTags ts2 --TODO: here or somewhere else, make sure that (prn) \\ (pers) doesn't match (prn pers)
 toTags (Cart ts1 ts2) = map concat $ sequence [(toTags ts1), (toTags ts2)]
 toTags All = [[]] --matches all
 

### 2015-08-19 test/Test.hs

@@ -27,6 +27,10 @@ engold = "data/en.tagged"
 -- engold = "data/vietnam.tagged"
 enpre = "data/en_pre.rlx"
 
+nldgr   = "data/nld.rlx"
+nldtext = "data/nld_story.txt"
+nldgold = "data/nld_story.gold"
+
 main :: IO ()
 main = do
   args <- getArgs
@@ -44,7 +48,8 @@ main = do
     ["gold","obs"]  -> optiBySz small ambiguous
     ["gold"]        -> gold small ambiguous esgold --using small as default grammar
     ["goldrus"]     -> gold rusgr rustext undefined
-    ["engold"]     -> gold engr entext engold
+    ["engold"]      -> gold engr entext engold
+    ["nldgold"]     -> gold nldgr nldtext nldgold
     [r,"gold"]      -> gold r ambiguous esgold --specify grammar
 
     (r:d:o) -> do rules <- readRules r
@@ -71,7 +76,7 @@ gold rl dt g = do rules <- readRules rl
                   resSAT <- mapM (disamSection (disambiguate False False) rules) text
                   resVISL <- vislcg3 rl dt True
                   putStrLn "SAT-CG in comparison to gold standard"
-                  let verbose = length text < 100 --change if you want different output
+                  let verbose = length text < 10 --change if you want different output
                   prAll "SAT" resSAT gold text verbose
                   putStrLn "\nVISLCG3 in comparison to gold standard"
                   prAll "VISL" resVISL gold text verbose

### 2015-08-19 CG_base.hs
restructure

@@ -65,7 +65,7 @@ toTags :: TagSet -> [[Tag]]
 
 toTags (TS tags) = tags
 toTags (Or ts1 ts2) = toTags ts1 ++ toTags ts2
-toTags (Diff ts1 ts2) = toTags ts1 \\ toTags ts2 --TODO: here or somewhere else, make sure that (prn) \\ (pers) doesn't match (prn pers)
+toTags (Diff ts1 ts2) = toTags ts1 \\ toTags ts2
 toTags (Cart ts1 ts2) = map concat $ sequence [(toTags ts1), (toTags ts2)]
 toTags All = [[]] --matches all
 

### 2015-08-19 executable/Symbolic.hs

@@ -65,7 +65,7 @@ main = do
                let rules = concat rls
                mapM_ print rules
                let tc = nub $ concatMap toTags tsets
-               let spl = take 2 $ drop 9 $ splits rules
+               let spl = splits rules
                mapM_ print spl
                tokens <- mapM (uncurry (testRule verbose ts tc)) spl
                corpus <- concat `fmap` readData "data/nld_story.txt"

### 2015-09-10 paper/nodalida/nodalida2015.tex
restructure
fix typo

--- 2015-06-01 paper/nodalida2015.tex
@@
 but it is in no way connected to the choice for \emph{la}. It all
-depends of the order of the two rules; if the \texttt{det} reading of
+depends on the order of the two rules; if the \texttt{det} reading of
 \emph{la} is removed first, then we are stuck with that choice.
@@
 
-We also wanted to test whether SAT-CG outperforms traditional CG with a
+We also tested whether SAT-CG outperforms traditional CG with a
 small rule set. With our best performing and most concise

### 2015-09-10 CG_SAT.hs
mkVarsGrammarAnalysis: fix creation of variables when a rule is triggered in multiple places

@@ -20,13 +20,14 @@ module CG_SAT
 where 
 
 
-import CG_base
+import CG_base hiding ( name )
 import CG_data
 import Data.List
 import Data.Maybe
 import Control.Monad ( liftM, liftM2, when )
-import SAT.Named -- Add name for variables, for nicer looking output
+import SAT.Named -- Added name for variables, for nicer looking output
 import SAT (Solver,newSolver,conflict,modelValueMaybe)
+import qualified SAT
 import SAT.Unary ( Unary )
 import Debug.Trace
 
@@ -53,7 +54,7 @@ prTags :: [Tag] -> String
 prTags = unwords . map show
 
 instance Show Token where
-  show = show . getLit --Token
+  show = show . getLit 
 
 type Context = (Condition, [Token])
 
@@ -73,7 +74,7 @@ mkNegative (T i tags lit c _) = T i tags lit c False
 dummyTok :: Token
 dummyTok = T { getInd  = 999
              , getTags = []
-             , getLit  = true
+             , getLit  = Lit "ALWAYS" SAT.true
              , isCautious = False
              , isPositive = True }
 
@@ -239,6 +240,8 @@ go s isSelect isGrAna trgs_diffs (conds:cs) allToks = do
 
     return $ (rmClauses ++ slClauses, concat helps) 
 
+  mapT :: (a -> b) -> (a, a) -> (b, b)
+  mapT f (a1, a2) = (f a1, f a2)
 
 
   mkVarsGrammarAnalysis :: [(Token,[Context])] -> [(Token,[Context])] -> IO ([Clause], [Lit])
@@ -251,9 +254,12 @@ go s isSelect isGrAna trgs_diffs (conds:cs) allToks = do
                            rm_trg <- implies s trgName if_ctx_holds (neg $ getLit trg)
                            return (if_ctx_holds, rm_trg)
                          | (trg, ctx) <- rm 
-                         , let condNames = unwords $ map (show.fst) ctx
+                         , let (conds,insts) = unzip ctx
+--                         , let instNames = map show insts
+                         , let condNames = unwords $ map show conds
                          , let ctxName = "IF ("++condNames++")" 
-                         , let trgName = "REMOVE " ++ show trg ]
+                         , let thisInstanceName = " if_(" ++ (intercalate "," (map (show.head) insts ++ ["etc."])) ++ ")"
+                         , let trgName = "rm_" ++ show trg ++ thisInstanceName ]
                          
 
[…]

### 2015-09-10 executable/Symbolic.hs

@@ -1,5 +1,5 @@
 import CG_base
-import CG_SAT hiding ( isCautious, isPositive, showToken )
+import CG_SAT hiding ( isCautious, isPositive )
 import CG_parse
 import SAT.Named
 import SAT ( Solver, newSolver, conflict )
@@ -10,20 +10,27 @@ import Debug.Trace
 
 import System.Environment
 import System.IO ( hFlush, stdout )
-import System.IO.Unsafe
 
 
-showToken = show . getLit
+ex_onlyTrgLeft = concat $ snd $ parseRules False
+    -- ( "REMOVE:r1 (adj OR det) IF (1 v) ;" ++
+     ( "REMOVE:r1 (adj OR det) ;" ++
+       "REMOVE:r2 v IF (-1 adj LINK 0 det ) ;" ) 
 
-testrules = concat $ snd $ parseRules False "SELECT:r1 (aa) OR (acr) IF (-1C (mf)) ;\nREMOVE:r2 (aa) IF (-1C (acr)) ;"
-tinyrules = concat $ snd $ parseRules False 
-          ( "SET DetNoAdj = (det) - (adj) ;" ++
-            "SET AdjNoDet = (adj) - (det) ;" ++
-           "REMOVE:r1 (v) IF (-1C (det)) (NOT 0 (n)) ;" ++
-           "REMOVE:r2 (det) IF (1 (v)) ;" ++
-           "REMOVE:r3 (v) IF (-1 det) (1 (det));" ) 
---           "REMOVE:r3 (n) IF ( (-1 DetNoAdj OR AdjNoDet) OR (-2 (v)) ) (-2 (adj)) ;" ++
---           "REMOVE:r3 (n) IF (-1C DetNoAdj OR AdjNoDet) ;" )
+ex_firstRuleStricter = concat $ snd $ parseRules False
+     ( "REMOVE:r1 v IF (-1C det) ;" ++
+       "REMOVE:r2 v IF (-1  det) ;" ) 
+
+ex_threerules = concat $ snd $ parseRules False 
+     ( "REMOVE:r1 (v) IF (-1C (det))  ;" ++
+       "REMOVE:r2 (det) IF (1 (v)) ;"    ++
+       "REMOVE:r3 (v) IF (-1 (det) ) ;" ) 
+
+ex_testDisj = concat $ snd $ parseRules False
[…]

### 2015-10-09 paper/lrec/lrec2016_listenmaa_claessen.tex
add stubs for paper

--- 2015-09-10 paper/nodalida/nodalida2015.tex
@@
 
-\title{Constraint Grammar as a SAT problem}
+\title{Analysing Constraint Grammar with a SAT solver}
 
@@
 
-\begin{abstract}
+% Introduction & problem description
+\input{problem_desc}
 
-We represent Constraint Grammar (CG) as a Boolean satisfiability 
-(SAT) problem.
-Encoding CG in logic brings some new features to the grammars.
-The rules are interpreted in a more declarative way, which makes it
-possible to abstract away from details such as cautious context and
-ordering.
-%In addition to the target words, 
-A rule is allowed to affect its context words, which makes the number
-of the rules in a grammar potentially smaller. 
-Ordering can be preserved or discarded;
-in the latter case, we solve eventual rule conflicts by finding a
-solution that discards the least number of rule applications.
-We test our implementation by parsing texts in the order of
-10,000s--100,000s words, using grammars with hundreds of rules.
-\end{abstract}
+% Implementation details; very brief for the abstract
+\input{implementation}
 
-\section{Introduction and previous research}
-
-Constraint Grammar (CG) \cite{karlsson1995constraint} is a relatively
-young formalism, born out of practical need for a robust and
-language-independent method for part-of-speech tagging.
-%It belongs to the reductionist family of parsing
-%strategies, along with e.g. Finite-State Intersection Grammar (FSIG) \cite{koskenniemi90}.
-In this work, we present CG %\footnote{Roughly compatible with CG-2}
-as a Boolean satisfiability (SAT) problem, and describe an implementation
-using a SAT solver. This is attractive for several reasons: formal logic is
-well-studied, and serves as an abstract language to reason about the
-properties of CG. Constraint rules encoded in logic capture richer
-dependencies between the tags than standard CG.
-
-Applying logic to reductionist grammars has been explored earlier by \cite{lager98,lager_nivre01}, but it was never adopted for use.
-% ; logic programming was too slow to be used for tagging or parsing. 
-Since those works, SAT solving techniques have improved significantly \cite{marques_silva2010}, and they are used in domains such as microprocessor design and computational 
-biology---these problems easily match or exceed CG in complexity. 
-Thanks to these advances, we were able to revisit the idea and develop it
-further.
-
-Our work is primarily inspired by \cite{lager98}, which presents constraint
-rules as a disjunctive logic program, and \cite{lager_nivre01}, which
-reconstructs four different formalisms in first-order logic.
-Other works combining logic to CG include
-\cite{lindberg_eineborg98ilp} and \cite{asfrent14}, both using Inductive Logic Programming to learn CG rules from a tagged corpus.
-%We explore different approaches to ordering. 
-%In the case of unordered rules, it resembles Finite-State Intersection Grammar \cite{koskenniemi90}.
-%See also \cite{lager01transformation} for discussion of ordering of the CG rules.
-
-\section{CG as a SAT problem}
-Let us demonstrate our approach with the following example in Spanish.
-
-\begin{verbatim}
-"<la>"
-        "el" det def f sg
-        "lo" prn p3 f sg
-"<casa>"
-        "casa" n f sg
-        "casar" v pri p3 sg
-        "casar" v imp p2 sg
-\end{verbatim}
-
-The ambiguous passage can be either a noun phrase, \emph{la}\texttt{<det>} \emph{casa}\texttt{<n>} 
-`the house'  or a verb phrase \emph{la}\texttt{<prn>}  \emph{casa}\texttt{<v><pri><p3>} `(he/she) marries her'. 
-We add the following rules:
-
-\begin{itemize}
-\item [] \texttt{REMOVE prn IF (1 n) ;} \\
-             \texttt{REMOVE det IF (1 v) ;}
-\end{itemize}
-
-Standard CG will apply one  of the rules to the word \emph{la}; 
-either the one that comes first, or by some other heuristic. 
-The other rule will not fire, because it would remove the last
-reading. 
-If we use the cautious mode (\texttt{1C n} or \texttt{1C v}), which
- requires the word in the context to be fully disambiguated, 
-neither of the rules will be applied.
-In any case, all readings of \emph{casa} are left untouched by these rules.
-
-The SAT solver performs a search, 
-and starts building possible models that satisfy both constraints. 
-In addition to the given constraints, we have default rules to emulate
-the CG principles: an analysis is true if no rule affects it,
-and at least one analysis for each word is true---the notion of
-``last'' is not applicable.
-
-With these constraints, we get two solutions. The interaction of the
-rules regarding \emph{la}  disambiguates the part of speech of
-\emph{casa} for free, and the order of the rules does not matter.
-
-\begin{enumerate}
-\item [\texttt{1)}]
-\begin{verbatim}
-"<la>"
-        "el" det def f sg
-"<casa>"
-        "casa" n f sg
-\end{verbatim}
-\item [\texttt{2)}]
-\begin{verbatim}
-"<la>"
-        "lo" prn p3 f sg
-"<casa>"
-        "casar" v pri p3 sg
-        "casar" v imp p2 sg
-\end{verbatim}
-\end{enumerate}
-
-% Without more context or additional rules we cannot fully disambiguate the passage,
-% but unlike with standard CG, we get only legal combinations in one solution.
-
-\noindent The most important differences between the traditional and
-the SAT-based approach are described in the following sections.
-
-\subsection{Rules disambiguate more}
-Considering our example phrase and rules, the standard CG implementation
-can only remove readings from the target word (\texttt{prn} or
-\texttt{det}). The SAT-based implementation interprets the rules as
-``determiner and verb together are illegal'', and is free to take action that concerns also the word in the condition (\texttt{n} or \texttt{v}).
-
-This behaviour is explained by simple properties of logical formulae.
-When the rules are applied to the text, they are translated into
-implications: \texttt{REMOVE prn IF (1 n)} becomes
- \emph{casa}\texttt{<n>} $\Rightarrow$ $\neg$\emph{la}\texttt{<prn>},
- which reads ``if the \texttt{n} reading for \emph{casa} is true, then
- discard the \texttt{prn} reading for \emph{la}''.
-Any implication $a\,\Rightarrow\,b$ can be represented as a disjunction
-$\neg a\,\vee\,b$; intuitively, either the antecedent is false
-and the consequent can be anything, or the consequent is true and the
-antecedent can be anything.
-Due to this property, our rule translates into the disjunction 
-$\neg$\emph{casa}\texttt{<n>} $\vee$ $\neg$\emph{la}\texttt{<prn>},
-which is also equivalent to another implication, 
- \emph{la}\texttt{<prn>} $\Rightarrow$ $\neg$\emph{casa}\texttt{<n>}.
-This means that the rules are logically flipped: \texttt{REMOVE prn IF
-  (1 n)} translates into the same logical formula as  \texttt{REMOVE n
-  IF (-1 prn)}. 
-A rule with more conditions corresponds to many rules, each condition
-taking its turn to be the target. % of removal or selection.
-
-% \begin{itemize}
-% \item [] \texttt{REMOVE n IF (-1 prn) ;} \\
-%          \texttt{REMOVE v IF (-1 det) ;}
-% \end{itemize}
-
-% SAT-based approach gives identical results with both sets of rules,
-% whereas the standard CG would remove one reading from \emph{casa} and leave \emph{la} ambiguous.
-% Testing this property with more complex rules and larger rule sets remains to be done.
-
-%Rules with more conditions translate into many rules; rules with negation become complements (\texttt{(*)-X} for \texttt{NOT X}).
-Rules which require the condition to be unambiguously tagged, don't have an equivalent flip in the standard CG.
-% We did not expect much practical benefits, save for realising that some rules are bad by having to think further what it implies,
-%Out of interest, we manually flipped a small grammar and tested it against the original.
-% \todo{The results show  }
-%\begin{itemize}
-%\item []          \texttt{SL Inf IF (-2 V-MOD) (-1 "not");} \\
-%$\rightarrow$\texttt{SL V-MOD   IF (1 "not") (2 INF);} \\
-%$\rightarrow$\texttt{SL "not" IF (-1 V-MOD) (1 Inf);}
-%\item []           \texttt{SL Foo IF (NOT 1 Bar)} \\
-%$\rightarrow$\texttt{SL (*)-Foo if (-1 Bar)}
-%\end{itemize}
-
-\subsection{Cautious context is irrelevant}
-% Rather than waiting for a word to get disambiguated, the SAT solver starts by 
-% making assumptions (e.g. ``\emph{casa} is a noun'') and working under them,
-% discarding the assumption if it doesn't lead to a model that satisfies
-% all constraints.
-
-Traditional CG applies the rule set iteratively:
-some rules fire during the first iteration, either because their
-conditions do not require cautious context, or because some words are
-unambiguous to start with. This makes some more words unambiguous, and
-new rules can fire during the second iteration.
-
-In SAT-CG, the notion of cautious context is irrelevant. Instead of
-removing readings immediately,  each rule generates a  number of
-implications, and the SAT solver tries to find a model that will satisfy them.
-
-Let us continue with the earlier example. 
-We can add a word to the input:
-\begin{itemize}
-\item [] \emph{la casa grande} `the big house'
-\end{itemize}
-and a rule that removes verb reading, if the word is followed by an adjective:
-\begin{itemize}
-\item [] \texttt{REMOVE v IF (1 adj) ;}
-\end{itemize}
-
-% The SAT solver has already the clauses produced by the previous rules.
-% \emph{casa}\texttt{<n>} $\Rightarrow$ $\neg$\emph{la}\texttt{<prn>}
-% and   \emph{casa}\texttt{<v>} $\Rightarrow$ $\neg$\emph{la}\texttt{<det>}
-The new rule adds the implication
- \emph{grande}\texttt{<adj>} $\Rightarrow$
- $\neg$\emph{casa}\texttt{<v>}, which will
-disambiguate  \emph{casa} to a noun\footnote{Assuming that
-  \texttt{adj} is the only reading for \emph{grande}, it must be true,
-  because of the restriction that at least one analysis for each word
-  is true. Then the implication has a true antecedent (\emph{grande}\texttt{<adj>}), thus its
-  consequent ($\neg$\emph{casa}\texttt{<v>}) will hold.}. 
-As the status of \emph{casa} is resolved, the SAT solver can now
-discard the model where \emph{casa} is a verb and \emph{la} is a
-pronoun
-%---all these decisions are connected---
-and we get a unique solution with \texttt{det n adj}.
-% Now \emph{casa}\texttt{<n>} is true, so \emph{la}\texttt{<prn>} must be false. 
-%Because \emph{la}\texttt{<prn>} is false,  \emph{casa} cannot
-% be a \texttt{v}, and we get a unique solution with \texttt{det n adj}.
-
-Contrast this with the behaviour of the standard CG.
-With the new rule, standard CG will also remove the verb reading from \emph{casa}, 
-but it is in no way connected to the choice for \emph{la}. It all
-depends on the order of the two rules; if the \texttt{det} reading of
-\emph{la} is removed first, then we are stuck with that choice.
-If we made the first rules cautious, that is, keeping the determiner
-open until \emph{casa} is disambiguated, then we get the same
-result as with the SAT solver.
-Ideally, both ways of grammar writing should yield similar results;
-traditional CG rules are more imperative, and SAT-CG
-rules are more declarative.
-
-% The SAT-based approach only removes readings after it has enough
-% evidence to do that. From the grammar writer perspective, this removes
-% a burden of having to decide whether the rule should be cautious or
-% not---the SAT solver will only take action the surrounding context
-% supports the decision.
-
-% \todo{Unordered rules: also not applicable anymore to apply rules
-% iteratively; there's no ``this rule doesn't fire now but will after
-% applying X and Y'', it's all just implications ``this rule will fire
-% if this is true'' and let the SAT solver find if those rules can apply
-% peacefully to the same input.}
-
-\subsection{Rules can be unordered}
-\label{sec:unord}
-
-As hinted by the previous property, the SAT solver does not need a fixed
-order of the rules.
-Applying a rule to a sentence produces a number of clauses,
-and those clauses are fed into the SAT solver.
-However, in the unordered scheme, some information is lost: the
-following rule sets would be treated identically, whereas in the
-traditional CG, only the first would be considered as a bad order.
-
-\begin{itemize}
-\item [\texttt{1)}] \texttt{SELECT v ;} \\
-         \texttt{REMOVE v IF (-1 det) ;} 
-\item [\texttt{2)}] \texttt{REMOVE v IF (-1 det) ;} \\
-         \texttt{SELECT v ;}
-\end{itemize}
-
-Without order, both of these rule sets will conflict, if applied to an
-input that has sequence \texttt{det v}.
-The SAT solver is given clauses that tell to select a verb and remove a
-verb, and it cannot build a model that satisfies all of those clauses.
-To solve this problem, we create a variable for every instance of rule application, and request a solution where maximally many of these variables are true.
-If there is no conflict, then the maximal solution is one where all of
-these variables are true; that is, all rules take action.
-
-In case of a conflict, the SAT solver makes it possible to discard only
-minimal amount of rule applications. Continuing with the example,
-it is not clear which instances would be discarded, but if the rules
-were part of a larger rule set, and in the context the \textsc{remove} rule was
-the right one to choose, it is likely that the interaction between the desired
-rules would make a large set of clauses that fit together, and the
-\textsc{select} rule would not fit in, hence it would be discarded.
-
-This corresponds loosely to the common design pattern in
-CGs, where there is a number of rules with the same target, ordered
-such that more secure rules come first,
-with a catch-all rule with no condition as the last resort, to be
-applied if none of the previous has fired.
-The order-based heuristic in the traditional CG is replaced by a more
-holistic behaviour: if the rules conflict, discard the one that seems
-like an outlier.
-
-We can also emulate order with SAT-CG. To do that, we enter clauses
-produced by each rule one by one, and assume the solver state reached
-so far is correct. If a new clause introduces a conflict with
-previous clauses, we discard it and move on to the next rule.
-By testing against gold standard, we see that this scheme works better
-with ready-made CGs, which are written with ordering in mind.
-It also runs slightly faster than the unordered version.
-\\
-
-\noindent These three features influence the way rules are written. 
-We predict that less rules are needed; whether this holds in the order
-of thousands of rules remains to be tested. On the one hand, getting rid
-of ordering and cautious context could ease the task of the grammar
-writer, since it removes the burden of estimating the best sequence of
-rules and whether to make them cautious. On the other
-hand, lack of order can make the rules less transparent, and might not scale up for larger grammars.
-
-\section{Evaluation}
-\label{sec:eval}
-
-For evaluation, we measure the performance against the state-of-the-art CG parser VISL CG-3.
-% both in terms of accuracy and execution time.
-SAT-CG fares slightly worse for accuracy, and significantly worse for execution time.
-The results are presented in more detail in the following sections.
-
-%This suggests that our implementation should not compete with existing
-%state-of-the-art, but rather it has value as a way of relating the CG
-%formalism to wider context in the theory of computer science.
-%In Section~\ref{sec:apps} we discuss more about possible applications.
-
-\subsection{Performance against VISL CG-3}
-
-We took a manually tagged
-corpus\footnote{\url{https://svn.code.sf.net/p/apertium/svn/branches/apertium-swpost/apertium-en-es/es-tagger-data/es.tagged}}
-containing approximately 22,000 words of Spanish news text, 
-and a small constraint grammar\footnote{\url{https://svn.code.sf.net/p/apertium/svn/languages/apertium-spa/apertium-spa.spa.rlx}}, produced independently of the authors.
-% We ignored substitute rules, but kept subreadings and unification, even though our implementation doesn't handle them.
-We kept only \textsc{select} and \textsc{remove} rules, which left us 261 rules.
-With this setup, we produced an ambiguous version of the tagged
-corpus, and ran both SAT-CG and VISL CG-3 on it.
-Treating the original corpus as the gold standard, the disambiguation
-by  VISL CG-3 achieves F-score of 82.6 \%, ordered SAT-CG 81.5 \%  and
-unordered SAT-CG 79.2 \%. 
-We did not test with other languages or text genres due to the lack of
-available gold standard.
-%This result suggests that that when running grammars that are written
-%with the traditional CG in mind, SAT-CG loses with both ordering
-%strategies, but emulating order fares better.
-
-We also tested whether SAT-CG outperforms traditional CG with a
-small rule set. With our best performing and most concise
-grammar\footnote{\url{https://github.com/inariksit/cgsat/blob/master/data/spa\_smallset.rlx}}
-of only 19 rules, both SAT-CG and VISL CG-3  achieve a F-score of
-around 85 \%. This experiment is very small and might be explained by
-overfitting or mere chance, but it seems to indicate that rules that
-work well with SAT-CG are also good for traditional CG.
-% It was possible to attain a rule set for which the unordered SAT-CG
-% works better (\textless0.5 \%), but it required trying
-% subsequences of the 19-rule set in a brute force manner---this hardly
-% presents any real life use case.
-
-% Similar patterns were observed with small (\textless{}20) rule sets
-% written by the authors; depending on the subset, SAT-CG and VISL CG-3 had
-% a difference of at most $\Mypm$ 1.5 \%. 
-% Introducing rules one by one up to 19, the
-% performance improved in a very similar rate, with less than 0.5 \%
-% difference between the systems at each new rule.
-% We did not evaluate on other languages or text genres, due to lack of suitable test data.
-
-% Additional tests could include plugging SAT-CG into Apertium
-% translation pipeline, and comparing the translation quality.
-
-\subsection{Execution time}
-
-The worst-case complexity of SAT is exponential, whereas the standard
-implementations of CG are polynomial, but with advances in SAT solving
-techniques, the performance in the average case in practice is more feasible than in the previous works done in 90s--00s.
-We used the open-source SAT solver MiniSat \cite{een04sat}.
-
-We tested the performance by parsing Don Quijote (384,155 words) with
-the same Spanish grammars as in the previous experiment. 
-Table~\ref{table:time} shows execution times compared to VISL CG-3;
-SAT-CG\textsubscript{u} is the unordered scheme and
-SAT-CG\textsubscript{o} is the ordered.
-% For both systems, the number of rules in the grammar affects the performance more than the word count.
-From the SAT solving side, maximisation is the most costly operation. 
-Emulating order is slightly faster, likely because the maximisation problems are smaller.
-In any case, SAT does not seem to be the bottleneck: with 261 rules,
-the maximisation function was called 147,253 times, and with 19 rules,
-132,255 times, but 
-%However, with both rule sets, half of the sentences in Don Quijote
-%needed less than 6 calls of maximise, and 75 \% of the sentences
-%needed less than 14.
-the differences in the execution times are much larger, which suggests
-that there are other reasons for the worse performance. 
-This is to be expected, as SAT-CG is currently just a naive
-proof-of-concept implementation with no optimisations.
-
-\begin{table}
-  \centering
-  \begin{tabular}{|c|c|c|c|}
-     \hline
-     \textbf{\# rules} &  \textbf{SAT-CG\textsubscript{u}} & \textbf{SAT-CG\textsubscript{o}} & \textbf{VISL CG-3} \\ \hline
-      19   & 39.7s &  22.1s   & 4.2s\\ %\hline
-      99   & 1m34.1s & 1m14.9s & 6.1s \\ %\hline
-      261  & 2m54.1s & 2m31.6s & 10.7s \\ \hline
-  \end{tabular}
-  \caption{Execution times for 384,155 words.}
-  \label{table:time}
-\end{table}
-
-\section{Applications and future work}
-\label{sec:apps}
-%As demonstrated in Section~\ref{sec:eval}, our implementation is not
-%competitive with the state of the art. 
-%VISL CG-3 is efficient, reliable and widely used, and
-Instead of trying to compete with the state of the art, we plan to use SAT-CG for grammar analysis\footnote{We thank Eckhard Bick for the idea.}.
-There has been work on automatic tuning of hand-written CGs
-\cite{bick2013tuning}, but to our knowledge no tools to
-search for inconsistencies or suboptimal design. %in order to point it out to the grammar writer.
-
-The sequential application of traditional CG rules is good for
-performance and transparency. When a rule takes action, the analyses
-are removed from the sentence, and the next rules get the modified
-sentence as input. 
-% At the execution of rule, there is no way to go back to earlier rules and undo them. 
-As a downside, there is no way to know which part comes directly from the
-raw input and which part from applying previous rules.
-
-A conflict in an ordered scheme can be defined as 
-a set of two or more rules, such that applying the first makes the
-next rules impossible to apply, regardless of the input.
-We can reuse the example from Section~\ref{sec:unord}:
-
-\begin{itemize}
-\item [] \texttt{SELECT v ;} \\
-         \texttt{REMOVE v IF (-1 det) ;}
-\end{itemize}
-
-The first rule selects the verb reading everywhere and removes
-all other readings, leaving no chance for the second rule to take action.
-If the rules are introduced in a different order, there is no
-conflict: the \textsc{remove} rule would not remove verb readings from all
-possible verb analyses, so there is a possibility for the \textsc{select} rule
-to fire.
-
-Ordered SAT-CG can be used to detect these conflicts without any
-modifications, as a side effect of its design. 
-After applying each rule, it stores the clauses
-produced by the rule and commits to them.
-In case of a conflict, the program detects the particular rule that 
-violates the previous clauses, with the sentence
-where it is applied. Thus we get feedback which rule fails, and on which
-particular word(s).
-
-% This conflict solving can be done per section, or for the whole grammar. 
-% In case of section-based application, we run each section on the
-% input, and give the output as a fresh input to the new section, discarding
-% all clauses. Alternatively, we can let the rules interact with each
-% other regardless of the section boundaries.
-
-Unordered SAT-CG with maximisation-based conflict solving is not
-suitable for this task: the whole definition of conflict depends on
-ordering, and the unordered scheme deliberately loses this information.
-On a more speculative note, an unordered formalism such as Finite-State
-Intersection Grammar \cite{koskenniemi90} might benefit from the
-maximisation-based technique in conflict handling.
-
-Finally, we intend to test for conflicts
-without using a corpus.
-Let us illustrate the idea with the same two rules,
-\texttt{SELECT v} and \texttt{REMOVE v IF (-1 det)}
-in both orders.
-Assume we have the tag set \texttt{\{det, n, v\}}, 
-and we want to find if there exists an input such that
- both rules, applied in the given order, remove something from the input.
-% This is another constraint solving problem, hence we can model it in SAT.
-There are no inputs that satisfy the requirement with the first order,
-but several that work with the second, such as the following:
-
-\begin{itemize}
-\item []
-\begin{verbatim}
-"<w1>"
-        det
-        v
-"<w2>"
-        n
-        v
-\end{verbatim}
-\end{itemize}
-Thus we can say that the first rule order is conflicting, but the second
-one is not. 
-Implementing and testing this on a larger scale is left for future work.
-
-% SELECT (a) IF (0 PALJON + Nom + Sg) (1 N + Nom + Sg) ; # PALJO VALITTAMINEN
-% SELECT (a) IF (0 PALJON + Par) (1 N + Par + Sg) ;
-% SELECT (a) IF (0 PALJON + LOC-CASE) (1 N + 1 LOC-CASE + Sg) ;
-% REMOVE (a) IF (0 PALJON) ; # "on asetettu paljon haltijaksi"
-
-\section{Conclusions}
-
-SAT-solvers are nowadays powerful enough to be used for dealing with
-Constraint Grammar. A logic-based approach to CG has possible
-advantages over more traditional approaches; a SAT solver may
-disambiguate more words, and may do so more precisely, capturing
-more dependencies between tags.
-We experimented with both ordered and unordered rules, and found the 
-ordered scheme to work better with previously written grammars.
-For future direction, we intend to concentrate on grammar analysis,
-especially finding conflicts in constraint grammars.
-
-% Also, the SAT-solver requires less rules, and these rules are simpler. 
-% The ordering of rules is something we found was incompatible with a logic-based approach.
-We compensate this by maximising rule applications.
-Whether this is an advantage or disadvantage remains to be seen.
-Our initial experimental results are promising.
+% Preliminary evaluation and future work
+\input{eval_futurework}
 
 \section*{Acknowledgments}
-
-We thank Eckhard Bick, Tino Didriksen, Francis Tyers and Anssi
-Yli-Jyr{\"a} for comments and suggestions, as well as the anonymous
-reviewers and everyone who participated in the discussion at the CG workshop.
-
-% Do not number the acknowledgment section. Do not include this section
-% when submitting your paper for review.
+We thank Eckhard Bick for the idea to apply SAT to CG analysis.
 
 \bibliographystyle{acl}
-\bibliography{cg}
+\bibliography{../cg}
 

### 2015-10-09 paper/lrec/problem_desc.tex
stuff

\section{Introduction}

Constraint Grammar (CG, \cite{karlsson1995constraint})
% \citep[CG][]{karlsson1995constraint} 
is a formalism for disambiguating morphologically analysed input.
CGs are valuable language resources for rule-based language processing, especially for lesser resourced languages. They are robust and require no extensive corpora or other language resources. The formalism is lightweight, and even small CGs are shown to be worthwile \cite{lene_trond2011}.

As CGs grow larger, .

\cite{bick2013tuning} presents efforts to optimise hand-written CGs using machine learning techniques.
While Bick's experiments show improvements in results, the grammar writer is none the wiser why is the grammar better.
Our toolset is designed to complement the existing efforts.
The ideal use case would be a grammar writer, who wants to know e.g.

\begin{itemize}
\item Is there a sentence that triggers rule(s) X but not rule(s) Y
\item Give a sentence that is ambiguous but doesn't trigger any rules
\item Restrict tag combinations within one word
\end{itemize}

The technique requires an existing morphological lexicon, but no corpus.

### 2015-10-09 paper/lrec/problem_desc.tex
stuff^2

@@
 \section{Introduction}
+\label{sec:intro}
 
@@
 The technique requires an existing morphological lexicon, but no corpus.
+This makes it suitable also for languages without large corpora.
+Authors writing grammars for well-resourced languages can also benefit from the technique, as an additional tool along with a gold standard corpus. 
+Hand-annotated corpora are commonly used in the development of CGs
+ \cite{voutilainen99}, because they give immediate feedback whether a new rule
+increases or decreases accuracy, and by how large margin.
+
+\todo{Relate to conf themes: "Methodologies and tools for LRs construction and annotation" and
+"Validation and quality assurance of LRs"}
+
+% * Describe problem: CGs are huge & prone to mistakes,
+%  * we are looking at conflicts such as ...
+%  * if you forgot a case
+%  * some evidence that big CGs have conflicts and this is a real problem
+%  * Eckhard's paper to explain why conflicting rules are a problem
+%  * Goal: help grammar writer while they are writing the grammar, to avoid these kinds of problems
+%  * Examples
+
+% * Describe the technique
+
+% * Preliminary results
+%  - dutch & spanish
+%  - mention scalability
+%  - talk about size of SAT problem -- give number of SAT clauses for the last rule in the biggest grammar I have
+
+% * Future work
+%  - analysing different grammar formalisms
+%  - asking different questions
+%  - restrict yourself to readings that are actually words

### 2015-10-11 paper/lrec/problem_desc.tex
more text

@@
 \item Give a sentence that is ambiguous but doesn't trigger any rules
-\item Restrict tag combinations within one word
+\item Does my grammar contain ``dead code'' -- rules that would not fire in any case?
 \end{itemize}
@@
 Authors writing grammars for well-resourced languages can also benefit from the technique, as an additional tool along with a gold standard corpus. 
-Hand-annotated corpora are commonly used in the development of CGs
- \cite{voutilainen99}, because they give immediate feedback whether a new rule
-increases or decreases accuracy, and by how large margin.
+Hand-annotated corpora are commonly used in the development of CGs, because they give immediate feedback whether a new rule increases or decreases accuracy, and by how large margin (\cite{voutilainen2004}).
+For a language with no free hand-tagged corpus available, \cite{tyers_reynolds2015} describe
+
+Things you can get from corpus:
+
+\begin{itemize}
+\item \# times the rule was applied
+\item \% times it was correct
+\end{itemize}
+Strengths: gives concrete feedback about the effect of rules. Good especially at the beginning stages of grammar writing.
+
+Things you can get from ML tuning:
+\begin{itemize}
+\item Variations in the place of the rule
+\item Variations in the conditions (careful, ...) of the rule
+\end{itemize}
+Good side: try all combinations to find the best result.
+
+Things you can get from SAT solver:
+\begin{itemize}
+\item Interaction of rules; 
+\item Create word sequences to test what is possible and what can never happen
+\end{itemize}
+
+Given a large morphological lexicon, it is relatively easy to restrict each individual word to contain only realistic ambiguities.
+We expand the lexicon, map each wordform to its possible analyses, and then ignore the actual wordforms.
+For instance, a Spanish morphological lexicon contains the entries \texttt{casa:<verb><sg><p3>} and \texttt{casa:<noun><sg>}, hence we know that a confusion between a 3rd person singular verb and a singular noun is possible.
+The lexicon is not likely to contain a wordform that can be analysed both as an adjective and punctuation, so that kind of ambiguous word is never created.
+However, since our method is corpus-free, it does not contain any restrictions as to what words can follow each other.
+\footnote{Encode morphological lexicon and CG rules in SAT + ask for a sentence that will trigger rules = natural language generation!
+\:D/}
 

### 2015-10-11 paper/lrec/problem_desc.tex
related work

@@
 % \citep[CG][]{karlsson1995constraint} 
-is a formalism for disambiguating morphologically analysed input.
-CGs are valuable language resources for rule-based language processing, especially for lesser resourced languages.
+is a rule-based method for disambiguating morphologically analysed text. 
+A grammar consists of hundreds to thousands of rules that target specific analyses for selection or removal, based on contextual tests.
+For example, the rule \texttt{REMOVE verb IF (-1C det)} removes a verb reading from a word which is preceded by an unambiguously (C) tagged determiner.
+CGs are valuable resources for rule-based NLP, especially for lesser resourced languages.
 They are robust and require no extensive corpora or other language resources.
-The formalism is lightweight, and even small CGs are shown to be worthwile \cite{lene_trond2011}.
+The formalism is lightweight, and even small CGs are shown to be effective \cite{lene_trond2011}.
 
@@
 \item Give a sentence that is ambiguous but doesn't trigger any rules
-\item Does my grammar contain ``dead code'' -- rules that would not fire in any case?
+\item Does my grammar contain rules that contradict each other?
+\item Does my grammar contain rules that will never fire?
 \end{itemize}
@@
 Hand-annotated corpora are commonly used in the development of CGs, because they give immediate feedback whether a new rule increases or decreases accuracy, and by how large margin (\cite{voutilainen2004}).
-For a language with no free hand-tagged corpus available, \cite{tyers_reynolds2015} describe
+For a language with no free hand-tagged corpus available, \cite{tyers_reynolds2015} describe a method where they apply their rules to Wikipedia texts and pick 100 examples at random for manual check.
+
+\section{Previous work}
+
+This work draws inspiration from various sources.
+
+\begin{itemize}
+\item Methods for grammar writing \cite{voutilainen2004},  \cite{tyers_reynolds2015}
+\item Optimising a hand-written CG \cite{bick2013tuning}
+\item Encoding CG as a logic program \cite{lager98}, \cite{lindberg_eineborg98ilp},  \cite{lager01transformation}, \cite{asfrent14}, \cite{listenmaa_claessen2015}
+\end{itemize}
 
@@
 \end{itemize}
-Strengths: gives concrete feedback about the effect of rules. Good especially at the beginning stages of grammar writing.
+Strengths: gives concrete feedback about the effect of rules. 
+Good especially at the beginning stages of grammar writing. 
+Sometimes you just want to know that a rule fired 500 times and was correct 490 times, rather than ``it is possible to construct a word sequence that will trigger this rule''.
+The connection to the reality has also weak sides.
+This method will not notice a missed opportunity.
+What if instead of firing 500 times, it could have fired 1000 times if it had been placed earlier in the rule sequence.
+ to even a better rule order, nor it will tell if a rule would have
+% \cite{voutilainen2004} state that the first 200 rules are probably enough to resolve 50--75 \% of ambiguities in the corpus used in the development.
 

### 2015-10-13 paper/lrec/implementation.tex
stuff

--- 2015-10-11 paper/lrec/problem_desc.tex
@@
-\section{Introduction}
-\label{sec:intro}
+\section{Implementation}
 
-Constraint Grammar (CG, \cite{karlsson1995constraint})
-% \citep[CG][]{karlsson1995constraint} 
-is a formalism for disambiguating morphologically analysed input.
-CGs are valuable language resources for rule-based language processing, especially for lesser resourced languages.
-They are robust and require no extensive corpora or other language resources.
-The formalism is lightweight, and even small CGs are shown to be worthwile \cite{lene_trond2011}.
+In this section, we describe the implementation of the toolkit.
 
-As CGs grow larger, .
+\paragraph{SAT encoding of CG}
 
-\cite{bick2013tuning} presents efforts to optimise hand-written CGs using machine learning techniques.
-While Bick's experiments show improvements in results, the grammar writer is none the wiser why is the grammar better.
-Our toolset is designed to complement the existing efforts.
-The ideal use case would be a grammar writer, who wants to know e.g.
+See \cite{listenmaa_claessen2015} for further description.
+We make
 
-\begin{itemize}
-\item Is there a sentence that triggers rule(s) X but not rule(s) Y
-\item Give a sentence that is ambiguous but doesn't trigger any rules
-\item Does my grammar contain ``dead code'' -- rules that would not fire in any case?
-\end{itemize}
+\begin{table*}[]
+\centering
+\caption{My caption}
+\label{my-label}
+\begin{tabular}{|l|p{1.9cm}|l|l|}
+\hline
+\textbf{Word} & \textbf{Variables} & \textbf{Default rules} & \textbf{Constraint rules}\\ \hline
+\multirow{2}{*} ``the''  & \texttt{the\_det}  & \texttt{the\_det} & {\texttt{the\_det} $\Rightarrow$ $\neg$\texttt{bear\_v\_pl}} \\
+                ``bear'' & \texttt{bear\_n\_sg}  \texttt{bear\_v\_pl} & \texttt{bear\_n\_sg} $\vee$ \texttt{bear\_v\_pl} & \\  \hline
+\end{tabular}
 
-The technique requires an existing morphological lexicon, but no corpus.
-This makes it suitable also for languages without large corpora.
-Authors writing grammars for well-resourced languages can also benefit from the technique, as an additional tool along with a gold standard corpus. 
-Hand-annotated corpora are commonly used in the development of CGs, because they give immediate feedback whether a new rule increases or decreases accuracy, and by how large margin (\cite{voutilainen2004}).
-For a language with no free hand-tagged corpus available, \cite{tyers_reynolds2015} describe
+\end{table*}
 
-Things you can get from corpus:
+\paragraph{Symbolic sentences}
 
-\begin{itemize}
-\item \# times the rule was applied
-\item \% times it was correct
-\end{itemize}
-Strengths: gives concrete feedback about the effect of rules. Good especially at the beginning stages of grammar writing.
-
-Things you can get from ML tuning:
-\begin{itemize}
-\item Variations in the place of the rule
-\item Variations in the conditions (careful, ...) of the rule
-\end{itemize}
-Good side: try all combinations to find the best result.
-
-Things you can get from SAT solver:
-\begin{itemize}
-\item Interaction of rules; 
-\item Create word sequences to test what is possible and what can never happen
-\end{itemize}
+\paragraph{Word-internal constraints}
 
@@
 \footnote{Encode morphological lexicon and CG rules in SAT + ask for a sentence that will trigger rules = natural language generation!
-\:D/}
-
-\todo{Relate to conf themes: "Methodologies and tools for LRs construction and annotation" and
-"Validation and quality assurance of LRs"}
-
-% * Describe problem: CGs are huge & prone to mistakes,
-%  * we are looking at conflicts such as ...
-%  * if you forgot a case
-%  * some evidence that big CGs have conflicts and this is a real problem
-%  * Eckhard's paper to explain why conflicting rules are a problem
-%  * Goal: help grammar writer while they are writing the grammar, to avoid these kinds of problems
-%  * Examples
-
-% * Describe the technique
-
-% * Preliminary results
-%  - dutch & spanish
-%  - mention scalability
-%  - talk about size of SAT problem -- give number of SAT clauses for the last rule in the biggest grammar I have
-
-% * Future work
-%  - analysing different grammar formalisms
-%  - asking different questions
-%  - restrict yourself to readings that are actually words
+:D}

### 2015-10-13 paper/lrec/problem_desc.tex
stuff

@@
 % \citep[CG][]{karlsson1995constraint} 
-is a rule-based method for disambiguating morphologically analysed text. 
-A grammar consists of hundreds to thousands of rules that target specific analyses for selection or removal, based on contextual tests.
-For example, the rule \texttt{REMOVE verb IF (-1C det)} removes a verb reading from a word which is preceded by an unambiguously (C) tagged determiner.
+is a method for disambiguating morphologically analysed text% and adding syntactic or dependency labels
+. 
+A grammar consists of rules that target specific analyses for selection or removal, based on contextual tests. For example, the following rule
+\begin{itemize}
+\item[] \texttt{REMOVE verb IF (-1C det) ;}
+\end{itemize}
+removes a verb reading from a word which is preceded by an unambiguously (C) tagged determiner.
+Given a text, such as
+
+\begin{verbatim}
+"<the>"
+        "the" det def
+"<bear>"
+        "bear" noun sg
+        "bear" verb pl
+\end{verbatim}
+
+the rule will apply to the word \emph{bear}, and remove the analysis \texttt{verb pl}.
+Rules are not allowed to remove the last remaining analysis: if  \texttt{verb pl} were the only analysis for \emph{bear}, then the rule would not apply, even if the condition is met.
+
 CGs are valuable resources for rule-based NLP, especially for lesser resourced languages.
-They are robust and require no extensive corpora or other language resources.
-The formalism is lightweight, and even small CGs are shown to be effective \cite{lene_trond2011}.
+They are robust and can be written without large corpora--only morphological analysis is needed.
+The formalism is lightweight and language-independent, and resources can be shared between related languages \cite{bick2006spanish}, \cite{lene_trond_linda2010}.
+Mature CGs contain some thousands of rules, but even small CGs are shown to be effective \cite{lene_trond2011}.
 
@@
 \begin{itemize}
-\item Is there a sentence that triggers rule(s) X but not rule(s) Y
+\item Is there a sentence that triggers rule(s) X but not rule(s) Y?
 \item Give a sentence that is ambiguous but doesn't trigger any rules
@@
 What if instead of firing 500 times, it could have fired 1000 times if it had been placed earlier in the rule sequence.
- to even a better rule order, nor it will tell if a rule would have
+Or the rule that was wrong in 60 \% of the time,
+
 % \cite{voutilainen2004} state that the first 200 rules are probably enough to resolve 50--75 \% of ambiguities in the corpus used in the development.
@@
 Good side: try all combinations to find the best result. 
+At a certain point, the grammar gets so big that it is hard to keep track of all the rules and their interactions.
+\cite{bick2013tuning} tries out combinations of moving rules in different sections or removing them in total, and in parallel, making their contexts stricter or less strict.
+This is a valuable tool, especially for grammars that are so big that it's hard to keep track of.
+Program can try all combinations whereas trying to make sense out of a huge set of rules would be hard for humans.
+As a downside, the grammar writer will likely not know why exactly is the tuned grammar performing better.
 
@@
 \begin{itemize}
-\item Interaction of rules; 
+\item Interaction of rules
 \item Create word sequences to test what is possible and what can never happen
 \end{itemize}
-
-Given a large morphological lexicon, it is relatively easy to restrict each individual word to contain only realistic ambiguities.
-We expand the lexicon, map each wordform to its possible analyses, and then ignore the actual wordforms.
-For instance, a Spanish morphological lexicon contains the entries \texttt{casa:<verb><sg><p3>} and \texttt{casa:<noun><sg>}, hence we know that a confusion between a 3rd person singular verb and a singular noun is possible.
-The lexicon is not likely to contain a wordform that can be analysed both as an adjective and punctuation, so that kind of ambiguous word is never created.
-However, since our method is corpus-free, it does not contain any restrictions as to what words can follow each other.
-\footnote{Encode morphological lexicon and CG rules in SAT + ask for a sentence that will trigger rules = natural language generation!
-\:D/}
 

### 2015-10-13 paper/lrec/implementation.tex
more text

@@
 \section{Implementation}
+\label{sec:implementation}
 
 In this section, we describe the implementation of the toolkit.
+We encode Constraint Grammar as a SAT problem, as in \cite{listenmaa_claessen2015}, and from the morphological lexicon, we get constraints for individual words.
+This allows us to create word sequences that will or will not trigger
+use a SAT solver to constrain the possibilities.
 
@@
 \centering
-\caption{My caption}
+\caption{SAT-encoding of CG}
 \label{my-label}
@@
 However, since our method is corpus-free, it does not contain any restrictions as to what words can follow each other.
-\footnote{Encode morphological lexicon and CG rules in SAT + ask for a sentence that will trigger rules = natural language generation!
-:D}
+% \footnote{Encode morphological lexicon and CG rules in SAT + ask for a sentence that will trigger rules = natural language generation! :D}

### 2015-10-13 paper/lrec/problem_desc.tex
more text

@@
 
-As CGs grow larger, .
-
+As CGs grow larger, it gets harder for the grammar writers and users to keep track of all the rules and their interaction.
 \cite{bick2013tuning} presents efforts to optimise hand-written CGs using machine learning techniques.
 While Bick's experiments show improvements in results, the grammar writer is none the wiser why is the grammar better.
-Our toolset is designed to complement the existing efforts.
+Our toolset is designed to complement the existing approaches in CG analysis and optimisation.
 The ideal use case would be a grammar writer, who wants to know e.g.
@@
 \begin{itemize}
-\item Is there a sentence that triggers rule(s) X but not rule(s) Y?
-\item Give a sentence that is ambiguous but doesn't trigger any rules
 \item Does my grammar contain rules that contradict each other?
 \item Does my grammar contain rules that will never fire?
+\item Generate a sequence that triggers rule(s) X but not rule(s) Y
+\item Generate a sequence that is ambiguous but doesn't trigger any rules
 \end{itemize}
@@
 The technique requires an existing morphological lexicon, but no corpus. 
-This makes it suitable also for languages without large corpora.
-Authors writing grammars for well-resourced languages can also benefit from the technique, as an additional tool along with a gold standard corpus. 
-Hand-annotated corpora are commonly used in the development of CGs, because they give immediate feedback whether a new rule increases or decreases accuracy, and by how large margin (\cite{voutilainen2004}).
-For a language with no free hand-tagged corpus available, \cite{tyers_reynolds2015} describe a method where they apply their rules to Wikipedia texts and pick 100 examples at random for manual check.
+The lexicon is needed in order to generate sequences that have realistic ambiguities: e.g. a word can be ambiguous between noun and verb, but not between adjective and punctuation. 
+Constraints for the structure of the sequences come from the CG rules themselves.
+We encode both of these constraints as a satisfiability problem, and use a SAT solver.
+
+The paper is structured as follows.
+Section~\ref{sec:prev} relates our work to the previous research.
+Section~\ref{sec:implementation} discusses the techniques we use, Section~X presents preliminary results and  Section~Y discusses future work.
+Section~Z concludes the paper.
 
 \section{Previous work}
+\label{sec:prev}
 
-This work draws inspiration from various sources.
+Our work draws inspiration from various sources.
+This section explains what our approach is and is not, and relates it to the previous research.
+We combine elements from the following aspects of CG research:
 
 \begin{itemize}
-\item Methods for grammar writing \cite{voutilainen2004},  \cite{tyers_reynolds2015}
+\item Methods to aid in hand-crafting rules \cite{voutilainen2004}, \cite{tyers_reynolds2015}
 \item Optimising a hand-written CG \cite{bick2013tuning}
-\item Encoding CG as a logic program \cite{lager98}, \cite{lindberg_eineborg98ilp},  \cite{lager01transformation}, \cite{asfrent14}, \cite{listenmaa_claessen2015}
+\item Encoding CG in logic \cite{lager98}, \cite{lager_nivre01}, \cite{lager01transformation}, \cite{listenmaa_claessen2015}
 \end{itemize}
 
-Things you can get from corpus:
+There is a large body of research on automatically inducing rules, e.g. \cite{inducing_cg1996}, \cite{lindberg_eineborg98ilp}%, \cite{asfrent14}
+. Our work is aimed to aid the process of hand-crafting rules.
 
+\paragraph{Corpus-based methods in manual grammar development}
+
+Hand-annotated corpora are commonly used in the development of CGs, because they give immediate feedback whether a new rule increases or decreases accuracy, and by how large margin \cite{voutilainen2004}.
+For a language with no free hand-tagged corpus available, \cite{tyers_reynolds2015} describe a method where they apply their rules to Wikipedia texts and pick 100 examples at random for manual check.
 \begin{itemize}
@@
 % \cite{voutilainen2004} state that the first 200 rules are probably enough to resolve 50--75 \% of ambiguities in the corpus used in the development. 
+Authors writing grammars for well-resourced languages can also benefit from the technique, as an additional tool along with a gold standard corpus.
 
-Things you can get from ML tuning:
+\paragraph{ML optimisation of hand-written grammars}
 \begin{itemize}
@@
 
-Things you can get from SAT solver:
+\paragraph{Things you can get from SAT solver}
+
 \begin{itemize}

### 2015-10-13 paper/lrec/problem_desc.tex
add quote by atro
move text around

@@
 As CGs grow larger, it gets harder for the grammar writers and users to keep track of all the rules and their interaction.
-\cite{bick2013tuning} presents efforts to optimise hand-written CGs using machine learning techniques.
+\cite{voutilainen2004} gives a detailed account about best practices of grammar writing, and efficient use of hand-annotated corpora to aid the grammar development.
+\cite{bick2013tuning} presents a method to optimise hand-written CGs using machine learning techniques.
+Our toolset is designed to complement the existing approaches in CG analysis and optimisation.
 While Bick's experiments show improvements in results, the grammar writer is none the wiser why is the grammar better.
-Our toolset is designed to complement the existing approaches in CG analysis and optimisation.
-The ideal use case would be a grammar writer, who wants to know e.g.
+Authors who have a large hand-annotated corpus can also benefit from the technique.
+With our method, it is possible to ask, for example, the following questions:
 
@@
 
-Our work draws inspiration from various sources.
-This section explains what our approach is and is not, and relates it to the previous research.
+\begin{quote}Another desirable facility in the grammar development environment would
+be a mechanism for identifying pairs of constraints that contradict each
+other.
+--Atro Voutilainen, 2004
+\end{quote}
+
+This section describes prior research and relates our contribution to the existing approaches.
 We combine elements from the following aspects of CG research:
@@
 \begin{itemize}
-\item Methods to aid in hand-crafting rules \cite{voutilainen2004}, \cite{tyers_reynolds2015}
-\item Optimising a hand-written CG \cite{bick2013tuning}
-\item Encoding CG in logic \cite{lager98}, \cite{lager_nivre01}, \cite{lager01transformation}, \cite{listenmaa_claessen2015}
+\item Corpus-based methods in manual grammar development \cite{voutilainen2004}, \cite{tyers_reynolds2015}
+\item Optimising hand-written grammars \cite{bick2013tuning}
+\item Encoding CG in logic \cite{lager98}, \cite{lager_nivre01}, \cite{listenmaa_claessen2015}
 \end{itemize}
 
-There is a large body of research on automatically inducing rules, e.g. \cite{inducing_cg1996}, \cite{lindberg_eineborg98ilp}%, \cite{asfrent14}
-. Our work is aimed to aid the process of hand-crafting rules.
+In addition, there is a large body of research on automatically inducing rules, e.g. \cite{inducing_cg1996}, \cite{lindberg_eineborg98ilp}%, \cite{lager01transformation}, \cite{asfrent14} 
+. However, since our work is aimed to aid the process of hand-crafting rules, we omit those works from our discussion.
 
@@
 Hand-annotated corpora are commonly used in the development of CGs, because they give immediate feedback whether a new rule increases or decreases accuracy, and by how large margin \cite{voutilainen2004}.
-For a language with no free hand-tagged corpus available, \cite{tyers_reynolds2015} describe a method where they apply their rules to Wikipedia texts and pick 100 examples at random for manual check.
-\begin{itemize}
-\item \# times the rule was applied
-\item \% times it was correct
-\end{itemize}
-Strengths: gives concrete feedback about the effect of rules. 
-Good especially at the beginning stages of grammar writing. 
-Sometimes you just want to know that a rule fired 500 times and was correct 490 times, rather than ``it is possible to construct a word sequence that will trigger this rule''.
-The connection to the reality has also weak sides.
-This method will not notice a missed opportunity.
-What if instead of firing 500 times, it could have fired 1000 times if it had been placed earlier in the rule sequence.
+For a language with no free or tagset-compatible corpus available, \cite{tyers_reynolds2015} describe a method where they apply their rules to unannotated Wikipedia texts and pick 100 examples at random for manual check.
+
+% \begin{itemize}
+% \item \# times the rule was applied
+% \item \% times it was correct
+% \end{itemize}
+CG rules are usually arranged in sections, from safest to least safe.
+Given a grammar that has three sections, the rules are applied to the text in the following manner.
+First apply rules from section 1, and repeat until nothing changes in the text.
+Then apply rules from sections 1--2, and finally all rules, from sections 1--3.
+The best strategy is to start with the most effective rules with low error rate, so that they make way for the following, more heuristic and less safe rules to act on.
+
+A representative corpus is arguably the best way to get concrete numbers: how many times a rule applied and how often it was correct.
+However, this method will not notice a missed opportunity, nor suggest a way to improve, for instance changing the place of a rule.
+
+% \cite{voutilainen2004} state that the around 200 rules are probably enough to resolve 50--75 \% of ambiguities in the corpus used in the development. 
+% This is invaluable, especially at the beginning stages of grammar writing. 
+% Sometimes you just want to know that a rule fired 500 times and was correct 490 times, rather than ``it is possible to construct a word sequence that will trigger this rule''.
+
+The connection to the reality has also its limitations. 
+% What if instead of firing 500 times, it could have fired 1000 times if it had been placed earlier in the rule sequence.
 Or the rule that was wrong in 60 \% of the time,
 
-% \cite{voutilainen2004} state that the first 200 rules are probably enough to resolve 50--75 \% of ambiguities in the corpus used in the development. 
-Authors writing grammars for well-resourced languages can also benefit from the technique, as an additional tool along with a gold standard corpus.
-
-\paragraph{ML optimisation of hand-written grammars}
+\paragraph{Optimising hand-written grammars}
 \begin{itemize}

### 2015-10-14 paper/lrec/problem_desc.tex
more text about the CG encoded in logic stuff

@@
 removes a verb reading from a word which is preceded by an unambiguously (C) tagged determiner.
-Given a text, such as
-
+Given the following text,
+\begin{itemize}
+\item[] 
 \begin{verbatim}
@@
 \end{verbatim}
-
+\end{itemize}
 the rule will apply to the word \emph{bear}, and remove the analysis \texttt{verb pl}.
-Rules are not allowed to remove the last remaining analysis: if  \texttt{verb pl} were the only analysis for \emph{bear}, then the rule would not apply, even if the condition is met.
+However, if a word has only one remaining analysis, then the rule will not apply, even if the condition is met.
+%Rules are not allowed to remove the last remaining analysis: if  \texttt{verb pl} were the only analysis for \emph{bear}, then the rule would not apply, even if the condition is met.
 
@@
 As CGs grow larger, it gets harder for the grammar writers and users to keep track of all the rules and their interaction.
-\cite{voutilainen2004} gives a detailed account about best practices of grammar writing, and efficient use of hand-annotated corpora to aid the grammar development.
+\cite{voutilainen2004} gives a detailed account about best practices of grammar writing and efficient use of hand-annotated corpora to aid the grammar development.
 \cite{bick2013tuning} presents a method to optimise hand-written CGs using machine learning techniques.
@@
 
+\todo{Relate to conf themes: "Methodologies and tools for LRs construction and annotation" and
+"Validation and quality assurance of LRs"}
+
 \section{Previous work}
@@
 \begin{itemize}
-\item Corpus-based methods in manual grammar development \cite{voutilainen2004}, \cite{tyers_reynolds2015}
+\item Corpus-based methods in manual grammar development \cite{voutilainen2004}
 \item Optimising hand-written grammars \cite{bick2013tuning}
@@
 
-The connection to the reality has also its limitations. 
+%The connection to the reality has also its limitations. 
 % What if instead of firing 500 times, it could have fired 1000 times if it had been placed earlier in the rule sequence.
-Or the rule that was wrong in 60 \% of the time,
+%Or the rule that was wrong in 60 \% of the time,
 
@@
 
-\paragraph{Things you can get from SAT solver}
+\paragraph{CG encoded in logic}
+
+\cite{lager98} presents a CG-like shallow parsing system encoded in logic, and \cite{lager_nivre2001} continues with a reconstruction of four different formalisms.
+\cite{listenmaa_claessen2015} presents a working CG implementation.
+
+%The earlier works on logical reconstruction don't envision grammar analysis as one of the use cases,
+Grammar analysis is a natural use case, due to some key features of the logical reconstruction.
+
+The straight-forward implementation of a CG compiler goes through the text, applies the rules, and discards analyses at each round, losing all information from the previous round when moving on to the next.
+A logic-based implementation models the rules as implications, where the consequent from rule $i$ becomes the antecedent of the rule $i+1$. 
+Composing all the clauses gives the final analysis of each word.
+What we added to this: at any time, we can stop and ask 
+``after applying these rules, can we build a sentence with properties \_\_\_?''
+
+% Application of rules is presented as a composition of clauses
+
+% They model CG as clauses o
+% The key elements in the ontology of CG are positions, words and sets of tags.
+% Rule order is denoted by predicate $pos^i$(word, [tag]), 
+% which denotes the part of speech of a given word after applying the $i^{th}$ rule.
+% The rules are modelled as implications, of the form below:
 
@@
 \end{itemize}
-
-\todo{Relate to conf themes: "Methodologies and tools for LRs construction and annotation" and
-"Validation and quality assurance of LRs"}
 

### 2015-10-14 paper/lrec/eval_futurework.tex
cleanup about logical stuff

--- 2015-10-13 paper/lrec/problem_desc.tex
@@
-\section{Introduction}
-\label{sec:intro}
-
-Constraint Grammar (CG, \cite{karlsson1995constraint})
-% \citep[CG][]{karlsson1995constraint} 
-is a method for disambiguating morphologically analysed text% and adding syntactic or dependency labels
-. 
-A grammar consists of rules that target specific analyses for selection or removal, based on contextual tests. For example, the following rule
-\begin{itemize}
-\item[] \texttt{REMOVE verb IF (-1C det) ;}
-\end{itemize}
-removes a verb reading from a word which is preceded by an unambiguously (C) tagged determiner.
-Given a text, such as
-
-\begin{verbatim}
-"<the>"
-        "the" det def
-"<bear>"
-        "bear" noun sg
-        "bear" verb pl
-\end{verbatim}
-
-the rule will apply to the word \emph{bear}, and remove the analysis \texttt{verb pl}.
-Rules are not allowed to remove the last remaining analysis: if  \texttt{verb pl} were the only analysis for \emph{bear}, then the rule would not apply, even if the condition is met.
-
-CGs are valuable resources for rule-based NLP, especially for lesser resourced languages.
-They are robust and can be written without large corpora--only morphological analysis is needed.
-The formalism is lightweight and language-independent, and resources can be shared between related languages \cite{bick2006spanish}, \cite{lene_trond_linda2010}.
-Mature CGs contain some thousands of rules, but even small CGs are shown to be effective \cite{lene_trond2011}.
-
-As CGs grow larger, it gets harder for the grammar writers and users to keep track of all the rules and their interaction.
-\cite{bick2013tuning} presents efforts to optimise hand-written CGs using machine learning techniques.
-While Bick's experiments show improvements in results, the grammar writer is none the wiser why is the grammar better.
-Our toolset is designed to complement the existing approaches in CG analysis and optimisation.
-The ideal use case would be a grammar writer, who wants to know e.g.
-
-\begin{itemize}
-\item Does my grammar contain rules that contradict each other?
-\item Does my grammar contain rules that will never fire?
-\item Generate a sequence that triggers rule(s) X but not rule(s) Y
-\item Generate a sequence that is ambiguous but doesn't trigger any rules
-\end{itemize}
-
-The technique requires an existing morphological lexicon, but no corpus. 
-The lexicon is needed in order to generate sequences that have realistic ambiguities: e.g. a word can be ambiguous between noun and verb, but not between adjective and punctuation. 
-Constraints for the structure of the sequences come from the CG rules themselves.
-We encode both of these constraints as a satisfiability problem, and use a SAT solver.
-
-The paper is structured as follows.
-Section~\ref{sec:prev} relates our work to the previous research.
-Section~\ref{sec:implementation} discusses the techniques we use, Section~X presents preliminary results and  Section~Y discusses future work.
-Section~Z concludes the paper.
-
-\section{Previous work}
-\label{sec:prev}
-
-Our work draws inspiration from various sources.
-This section explains what our approach is and is not, and relates it to the previous research.
-We combine elements from the following aspects of CG research:
-
-\begin{itemize}
-\item Methods to aid in hand-crafting rules \cite{voutilainen2004}, \cite{tyers_reynolds2015}
-\item Optimising a hand-written CG \cite{bick2013tuning}
-\item Encoding CG in logic \cite{lager98}, \cite{lager_nivre01}, \cite{lager01transformation}, \cite{listenmaa_claessen2015}
-\end{itemize}
-
-There is a large body of research on automatically inducing rules, e.g. \cite{inducing_cg1996}, \cite{lindberg_eineborg98ilp}%, \cite{asfrent14}
-. Our work is aimed to aid the process of hand-crafting rules.
-
-\paragraph{Corpus-based methods in manual grammar development}
-
-Hand-annotated corpora are commonly used in the development of CGs, because they give immediate feedback whether a new rule increases or decreases accuracy, and by how large margin \cite{voutilainen2004}.
-For a language with no free hand-tagged corpus available, \cite{tyers_reynolds2015} describe a method where they apply their rules to Wikipedia texts and pick 100 examples at random for manual check.
-\begin{itemize}
-\item \# times the rule was applied
-\item \% times it was correct
-\end{itemize}
-Strengths: gives concrete feedback about the effect of rules. 
-Good especially at the beginning stages of grammar writing. 
-Sometimes you just want to know that a rule fired 500 times and was correct 490 times, rather than ``it is possible to construct a word sequence that will trigger this rule''.
-The connection to the reality has also weak sides.
-This method will not notice a missed opportunity.
-What if instead of firing 500 times, it could have fired 1000 times if it had been placed earlier in the rule sequence.
-Or the rule that was wrong in 60 \% of the time,
-
-% \cite{voutilainen2004} state that the first 200 rules are probably enough to resolve 50--75 \% of ambiguities in the corpus used in the development. 
-Authors writing grammars for well-resourced languages can also benefit from the technique, as an additional tool along with a gold standard corpus.
-
-\paragraph{ML optimisation of hand-written grammars}
-\begin{itemize}
-\item Variations in the place of the rule
-\item Variations in the conditions (careful, ...) of the rule
-\end{itemize}
-Good side: try all combinations to find the best result. 
-At a certain point, the grammar gets so big that it is hard to keep track of all the rules and their interactions.
-\cite{bick2013tuning} tries out combinations of moving rules in different sections or removing them in total, and in parallel, making their contexts stricter or less strict.
-This is a valuable tool, especially for grammars that are so big that it's hard to keep track of.
-Program can try all combinations whereas trying to make sense out of a huge set of rules would be hard for humans.
-As a downside, the grammar writer will likely not know why exactly is the tuned grammar performing better.
-
-\paragraph{Things you can get from SAT solver}
-
-\begin{itemize}
-\item Interaction of rules
-\item Create word sequences to test what is possible and what can never happen
-\end{itemize}
-
-\todo{Relate to conf themes: "Methodologies and tools for LRs construction and annotation" and
-"Validation and quality assurance of LRs"}
-
-% * Describe problem: CGs are huge & prone to mistakes,
-%  * we are looking at conflicts such as ...
-%  * if you forgot a case
-%  * some evidence that big CGs have conflicts and this is a real problem
-%  * Eckhard's paper to explain why conflicting rules are a problem
-%  * Goal: help grammar writer while they are writing the grammar, to avoid these kinds of problems
-%  * Examples
-
-% * Describe the technique
+\section{Evaluation and Future Work}
 

### 2015-10-14 paper/lrec/problem_desc.tex
cleanup about logical stuff

@@
 
-\cite{lager98} presents a CG-like shallow parsing system encoded in logic, and \cite{lager_nivre2001} continues with a reconstruction of four different formalisms.
-\cite{listenmaa_claessen2015} presents a working CG implementation.
-
+\cite{lager98} presents a CG-like shallow parsing system encoded in logic, and \cite{lager_nivre01} continues with a reconstruction of four different formalisms.
 %The earlier works on logical reconstruction don't envision grammar analysis as one of the use cases,
@@
 
-The straight-forward implementation of a CG compiler goes through the text, applies the rules, and discards analyses at each round, losing all information from the previous round when moving on to the next.
-A logic-based implementation models the rules as implications, where the consequent from rule $i$ becomes the antecedent of the rule $i+1$. 
-Composing all the clauses gives the final analysis of each word.
-What we added to this: at any time, we can stop and ask 
-``after applying these rules, can we build a sentence with properties \_\_\_?''
+The/A(n) 
+\todo{imperative/straight-forward/efficient/...?}
+CG compiler
+%, such as VISL CG-3,
+ cannot capture any dependencies between rules.
+% It discards analyses immediately: the output of the $i^{th}$ rule becomes the input of the $i+1^{th}$ rule, but there is no information which rules have been applied before.
+In contrast, a logic-based CG compiler does that by default. 
+The rules are modelled as implications and composed in the order of the rule sequence, such that 
+the consequent from the $i^{th}$ rule becomes the antecedent of the $i+1^{th}$ rule.
+Given this design, we added on top a way to ask for solutions with certain properties.
+%, such as ``after applying rules $0-i$, is it possible for $i+1$ to apply''.
 
@@
 
-\begin{itemize}
-\item Interaction of rules
-\item Create word sequences to test what is possible and what can never happen
-\end{itemize}
+% \begin{itemize}
+% \item Interaction of rules
+% \item Create word sequences to test what is possible and what can never happen
+% \end{itemize}
 
@@
 % * Describe the technique
-
-% * Preliminary results
-%  - dutch & spanish
-%  - mention scalability
-%  - talk about size of SAT problem -- give number of SAT clauses for the last rule in the biggest grammar I have
-
-% * Future work
-%  - analysing different grammar formalisms
-%  - asking different questions
-%  - restrict yourself to readings that are actually words

### 2015-10-14 paper/lrec/implementation.tex
add more text on symbolic sentences and what kinds of questions we can ask

--- 2015-10-13 paper/lrec/implementation.tex
@@
 \section{Implementation}
+\label{sec:implementation}
 
 In this section, we describe the implementation of the toolkit.
+We encode CG rules as a SAT problem, as in \cite{listenmaa_claessen2015}, and from the morphological lexicon, we get constraints for individual words.
+This allows us to create word sequences that will or will not trigger a given rule or rule set.
 
@@
 
+Table~\ref{table:vars} shows the translation of the analyses to variables, and rules to clauses. The column ``Default rules'' refers to the rule that each word must have at least one analysis.
 See \cite{listenmaa_claessen2015} for further description.
-We make
 
@@
 \centering
-\caption{My caption}
-\label{my-label}
+\caption{SAT-encoding of CG}
+\label{table:vars}
 \begin{tabular}{|l|p{1.9cm}|l|l|}
@@
 \end{tabular}
-
 \end{table*}
-
-\paragraph{Symbolic sentences}
 
@@
 The lexicon is not likely to contain a wordform that can be analysed both as an adjective and punctuation, so that kind of ambiguous word is never created.
-However, since our method is corpus-free, it does not contain any restrictions as to what words can follow each other.
-\footnote{Encode morphological lexicon and CG rules in SAT + ask for a sentence that will trigger rules = natural language generation!
-:D}
+
+%However, since our method is corpus-free, we don't add any external restrictions as to what words can follow each other.
+
+\paragraph{Symbolic sentences}
+
+When we start applying rules, we need to have a sentence to which to apply.
+As this is a corpus-free method, we don't have one.
+Instead, we create a \emph{symbolic sentence}: sequence of words that can, in the beginning, have any tag combination.
+At the word level, these tag combinations are restrained as described in the previous paragraph.
+At the sentence level, applying rules will start shaping the sentence.
+An indeterminate blob of possible words undergoes rules that describe illegal sequences of analyses,
+and at the application of each rule, more and more possibilities are discarded.
+All rules are applied to all word, given that their context is in scope: for instance, the sentence-initial word doesn't meet any condition that requires something of previous words.
+
+The applications of these rules shape our symbolic sentence into containing only things that can survive the rules.
+This means that we can ask a question ``after applying rules $0-i$, is it possible for $i+1$ to apply?'', and the answer is reliable; we can not find a counterexample by just having a bit more data, because we start from literally all possibilities, and each rule
+restricts the set of possible sentences.
+
+Note that this method is not suitable for language generation: rules in any realistic CG are tailored for real life ambiguities, whereas our symbolic sentence starts with literally all possibilities of tag combinations.
+
+\paragraph{What kind of questions we can ask?}
+
+The example so far has been of type ``Can rule Z apply after A-Y?''
+
+We can construct all kinds of symbolic sentences. We can set the length, restrict individual words (e.g. ``3rd word must be a noun''), require that it triggers some rule but not other.
+In case of a conflict, we can dig into the conflict and refine where the problem is: e.g. is the condition not met, or is it the case that the target is already removed or the only remaining analysis.

### 2015-10-14 paper/lrec/implementation.tex
more text on symbolic sentences

@@
 Instead, we create a \emph{symbolic sentence}: sequence of words that can, in the beginning, have any tag combination.
-At the word level, these tag combinations are restrained as described in the previous paragraph.
+At the word level, these tag combinations are restricted to only contain realistic ambiguities, as described in the previous paragraph.
 At the sentence level, applying rules will start shaping the sentence.
@@
 and at the application of each rule, more and more possibilities are discarded.
+
 All rules are applied to all word, given that their context is in scope: for instance, the sentence-initial word doesn't meet any condition that requires something of previous words.
+However, being a condition for another target will also shape a word.
+At a level of real texts, a rule such as \texttt{REMOVE b IF (-1 a)} is an instruction to remove a \texttt{b} if we find an \texttt{a} before it.
+In case of a symbolic sentence, we start in a situation where every word can be an \texttt{a} or \texttt{b}.
+In that context, the rule behaves as a statement about that \texttt{b} may not follow \texttt{a}.
+This relation is encoded between all adjacent words, and each of them can follow the restriction in the way that is compliant with other rules.
+For instance, word 1 may be constrained by another rule to have no \texttt{a}, and this means that word 2 is free to have \texttt{b} or not have it.
+If word 3 is required to have an  \texttt{a}, then word 4 either has to have no \texttt{b}, or it has to have only \texttt{b}, by the rule that the last analysis of the word will not be removed.
 

### 2015-10-15 paper/lrec/problem_desc.tex
cut some crap from intro

@@
 They are robust and can be written without large corpora--only morphological analysis is needed.
-The formalism is lightweight and language-independent, and resources can be shared between related languages \cite{bick2006spanish}, \cite{lene_trond_linda2010}.
+The formalism is lightweight and language-independent, and resources can be shared between related languages \cite{bick2006spanish,lene_trond_linda2010}.
 Mature CGs contain some thousands of rules, but even small CGs are shown to be effective \cite{lene_trond2011}.
@@
 As CGs grow larger, it gets harder for the grammar writers and users to keep track of all the rules and their interaction.
-\cite{voutilainen2004} gives a detailed account about best practices of grammar writing and efficient use of hand-annotated corpora to aid the grammar development.
-\cite{bick2013tuning} presents a method to optimise hand-written CGs using machine learning techniques.
-Our toolset is designed to complement the existing approaches in CG analysis and optimisation.
-While Bick's experiments show improvements in results, the grammar writer is none the wiser why is the grammar better.
-Authors who have a large hand-annotated corpus can also benefit from the technique.
-With our method, it is possible to ask, for example, the following questions:
+%\cite{voutilainen2004} gives a detailed account about best practices of grammar writing and efficient use of hand-annotated corpora to aid the grammar development.
+%\cite{bick2013tuning} presents a method to optimise hand-written CGs using machine learning techniques.
+%Our tools complement the existing methods in CG analysis and optimisation.
+%While Bick's experiments show improvements in results, the grammar writer is none the wiser why is the grammar better.
+%Authors who have a large hand-annotated corpus can also benefit from our tools.
+With our tools, the grammar writer or user can ask specific questions about their grammar, such as:
 
@@
 Constraints for the structure of the sequences come from the CG rules themselves.
-We encode both of these constraints as a satisfiability problem, and use a SAT solver.
+We encode both of these constraints as a satisfiability problem, and use a SAT solver to generate answers to the previous questions.
 
@@
 \item Optimising hand-written grammars \cite{bick2013tuning}
-\item Encoding CG in logic \cite{lager98}, \cite{lager_nivre01}, \cite{listenmaa_claessen2015}
+\item Encoding CG in logic \cite{lager98,lager_nivre01,listenmaa_claessen2015}
 \end{itemize}
@@
 
-% \begin{itemize}
-% \item \# times the rule was applied
-% \item \% times it was correct
-% \end{itemize}
-CG rules are usually arranged in sections, from safest to least safe.
+% CG rules are usually arranged in sections, from safest to least safe.
 Given a grammar that has three sections, the rules are applied to the text in the following manner.
@@
 Then apply rules from sections 1--2, and finally all rules, from sections 1--3.
-The best strategy is to start with the most effective rules with low error rate, so that they make way for the following, more heuristic and less safe rules to act on.
+% The best strategy is to start with the most effective rules with low error rate, so that they make way for the following, more heuristic and less safe rules to act on.
 

### 2015-10-15 CG_parse.hs
add more  stuff to grammar to parse finnish

@@ -125,6 +125,13 @@ transSetDecl (Set setname tagset) =
     (SetName (UIdent name)) -> do 
       ts <- transTagSet tagset
       return (name, ts)
+    (SetMeta (UIdent name)) -> do 
+      ts <- transTagSet tagset
+      return ("<" ++ name ++ ">", ts)
+    (SetSynt (UIdent name)) -> do 
+      ts <- transTagSet tagset
+      return ("@" ++ name, ts)
+
 transSetDecl (List setname tags) = 
   case setname of
     BOS          -> return (">>>", CGB.TS bos)

### 2015-10-15 paper/lrec/problem_desc.tex
more stuff

@@
 \end{itemize}
-the rule will apply to the word \emph{bear}, and remove the analysis \texttt{verb pl}.
-However, if a word has only one remaining analysis, then the rule will not apply, even if the condition is met.
-%Rules are not allowed to remove the last remaining analysis: if  \texttt{verb pl} were the only analysis for \emph{bear}, then the rule would not apply, even if the condition is met.
+the rule will match to the word \emph{bear}, and remove the analysis \texttt{verb pl}.
+However, if the target word has only one remaining analysis, then the rule will not apply, even if the condition is met.
 
@@
 The formalism is lightweight and language-independent, and resources can be shared between related languages \cite{bick2006spanish,lene_trond_linda2010}.
-Mature CGs contain some thousands of rules, but even small CGs are shown to be effective \cite{lene_trond2011}.
+%Mature CGs contain some thousands of rules, but even small CGs are shown to be effective \cite{lene_trond2011}.
 
@@
 \begin{itemize}
-\item Does my grammar contain rules that contradict each other?
-\item Does my grammar contain rules that will never fire?
+\item Are there rules that contradict each other?
+\item Are there rules that will never fire?
 \item Generate a sequence that triggers rule(s) X but not rule(s) Y
@@
 
-The technique requires an existing morphological lexicon, but no corpus. 
-The lexicon is needed in order to generate sequences that have realistic ambiguities: e.g. a word can be ambiguous between noun and verb, but not between adjective and punctuation. 
+This technique requires an existing morphological lexicon, but no corpus. 
+The lexicon is needed in order to generate realistic ambiguities: e.g. a word can be ambiguous between noun and verb, but not between adjective and punctuation. 
 Constraints for the structure of the sequences come from the CG rules themselves.
@@
 Section~\ref{sec:prev} relates our work to the previous research.
-Section~\ref{sec:implementation} discusses the techniques we use, Section~X presents preliminary results and  Section~Y discusses future work.
+Section~\ref{sec:implementation} discusses the implementation, Section~X presents preliminary results and  Section~Y discusses future work.
 Section~Z concludes the paper.
 
-\todo{Relate to conf themes: "Methodologies and tools for LRs construction and annotation" and
-"Validation and quality assurance of LRs"}
+%\todo{Relate to conf themes: "Methodologies and tools for LRs construction and annotation" and
+%"Validation and quality assurance of LRs"}
 
-\section{Previous work}
+\section{Related work}
 \label{sec:prev}
 
-\begin{quote}Another desirable facility in the grammar development environment would
-be a mechanism for identifying pairs of constraints that contradict each
-other.
---Atro Voutilainen, 2004
-\end{quote}
+% \begin{quote}Another desirable facility in the grammar development environment would
+% be a mechanism for identifying pairs of constraints that contradict each
+% other.
+% --Atro Voutilainen, 2004
+% \end{quote}
 
-This section describes prior research and relates our contribution to the existing approaches.
+%This section describes prior research and relates our contribution to the existing approaches.
 We combine elements from the following aspects of CG research:
@@
 Hand-annotated corpora are commonly used in the development of CGs, because they give immediate feedback whether a new rule increases or decreases accuracy, and by how large margin \cite{voutilainen2004}.
-For a language with no free or tagset-compatible corpus available, \cite{tyers_reynolds2015} describe a method where they apply their rules to unannotated Wikipedia texts and pick 100 examples at random for manual check.
+%\cite{voutilainen2004} gives a detailed account about best practices of grammar writing and efficient use of corpora to aid the grammar development.
+%For a language with no free or tagset-compatible corpus available, \cite{tyers_reynolds2015} describe a method where they apply their rules to unannotated Wikipedia texts and pick 100 examples at random for manual check.
 
-% CG rules are usually arranged in sections, from safest to least safe.
-Given a grammar that has three sections, the rules are applied to the text in the following manner.
-First apply rules from section 1, and repeat until nothing changes in the text.
-Then apply rules from sections 1--2, and finally all rules, from sections 1--3.
-% The best strategy is to start with the most effective rules with low error rate, so that they make way for the following, more heuristic and less safe rules to act on.
+CG rules are usually arranged in sections, and run in the following manner. 
+First apply rules from section 1, and repeat until nothing changes in the text. Then apply rules from sections 1--2, then 1--3 and so on, until the set includes all rules.
+The best strategy is to place the safest and most effective rules in the first sections,
+so that they make way for the following, more heuristic and less safe rules to act on.
 
-A representative corpus is arguably the best way to get concrete numbers: how many times a rule applied and how often it was correct.
-However, this method will not notice a missed opportunity, nor suggest a way to improve, for instance changing the place of a rule.
+A representative corpus is arguably the best way to get concrete numbers---how many times a rule applied and how often it was correct---and to arrange the rules in sections based on that feedback.
+However, this method will not notice a missed opportunity or a grammar-internal conflict, nor suggest ways to improve.
+%, for instance changing the place of a rule.
 
@@
 \paragraph{Optimising hand-written grammars}
-\begin{itemize}
-\item Variations in the place of the rule
-\item Variations in the conditions (careful, ...) of the rule
-\end{itemize}
-Good side: try all combinations to find the best result. 
-At a certain point, the grammar gets so big that it is hard to keep track of all the rules and their interactions.
+The corpus-based method can tell the effect of each single rule at their place in the rule sequence, and it is up to the grammar writer to make changes in the grammar.
+\cite{bick2013tuning} automates the latter process: the rules are moved into different sections or removed in total based on their effect, and in parallel, their contexts are made stricter or less strict.
+After trying out several combinations, the version that produces the best results is kept.
+
+This is a valuable tool for handling grammars that are too big to keep track of, and Bick's evaluation shows that the tuned grammars perform better than the original.
+As a downside, the grammar writer will likely not know why exactly is the tuned grammar performing better.
+% At a certain point, the grammar gets so big that it is hard to keep track of all the rules and their interactions.
 \cite{bick2013tuning} tries out combinations of moving rules in different sections or removing them in total, and in parallel, making their contexts stricter or less strict.
-This is a valuable tool, especially for grammars that are so big that it's hard to keep track of.
+% This is a valuable tool, especially for grammars that are so big that it's hard to keep track of.
 Program can try all combinations whereas trying to make sense out of a huge set of rules would be hard for humans.
-As a downside, the grammar writer will likely not know why exactly is the tuned grammar performing better.
 
@@
 
-\cite{lager98} presents a CG-like shallow parsing system encoded in logic, and \cite{lager_nivre01} continues with a reconstruction of four different formalisms.
+%\cite{lager98} presents a CG-like shallow parsing system encoded in logic, and \cite{lager_nivre01} continues with a reconstruction of four different formalisms.
 %The earlier works on logical reconstruction don't envision grammar analysis as one of the use cases,
+\cite{lager98} and \cite{lager_nivre01} reconstruct the CG formalism in first-order predicate logic, and \cite{listenmaa_claessen2015} implement a CG compiler using a SAT solver.
 Grammar analysis is a natural use case, due to some key features of the logical reconstruction.
-
-The/A(n) 
-\todo{imperative/straight-forward/efficient/...?}
-CG compiler
+The traditional CG compiler 
 %, such as VISL CG-3,

### 2015-10-15 CG_parse.hs
hammer the finnish cg to comply with my crappy parser

@@ -297,7 +297,8 @@ transPosition pos = case pos of
   Exactly (Signed str)     -> return $ CGB.Exactly False $ read str
   AtLeastPre (Signed str)  -> return $ CGB.AtLeast False $ read str
   AtLeastPost (Signed str) -> return $ CGB.AtLeast False $ read str
-  AtLPostCaut (Signed str) -> return $ CGB.AtLeast True $ read str
+  AtLPostCaut1 (Signed str) -> return $ CGB.AtLeast True $ read str
+  AtLPostCaut2 (Signed str) -> return $ CGB.AtLeast True $ read str
   Cautious position        -> cautious `fmap` transPosition position
   where cautious (CGB.Exactly _b num) = CGB.Exactly True num
         cautious (CGB.AtLeast _b num) = CGB.AtLeast True num

### 2015-10-16 paper/lrec/implementation.tex
make table nicer

@@
 \centering
-\caption{SAT-encoding of CG}
+\caption{SAT-encoding of \texttt{REMOVE v IF (-1 det)} applied to \emph{the bear}}
 \label{table:vars}
-\begin{tabular}{|l|p{1.9cm}|l|l|}
+\begin{tabular}{|l|l|l|l|}
 \hline
-\textbf{Word} & \textbf{Variables} & \textbf{Default rules} & \textbf{Constraint rules}\\ \hline
-\multirow{2}{*} ``the''  & \texttt{the\_det}  & \texttt{the\_det} & {\texttt{the\_det} $\Rightarrow$ $\neg$\texttt{bear\_v\_pl}} \\
-                ``bear'' & \texttt{bear\_n\_sg}  \texttt{bear\_v\_pl} & \texttt{bear\_n\_sg} $\vee$ \texttt{bear\_v\_pl} & \\  \hline
+\textbf{Word} & \textbf{Variables}                                                          & \textbf{Default rules}               & \textbf{Constraint rules}                                  \\ \hline
+``the''         & \texttt{the\_det}                                                           & \texttt{the\_det}                    & \multirow{2}{*}{\texttt{the\_det}  $\Rightarrow$ $\neg$\texttt{bear\_v\_pl}} \\ \cline{1-3}
+``bear''        & \begin{tabular}[c]{@{}l@{}}\texttt{bear\_n\_sg} \\ \texttt{bear\_v\_pl}\end{tabular} & \texttt{bear\_n\_sg} $\vee$ \texttt{bear\_v\_pl} &                                                            \\ \hline
 \end{tabular}

### 2015-10-16 CG_parse.hs
parse subreadings properly (and ignore the subpart)

@@ -330,6 +330,7 @@ transPosition pos = case pos of
   AtLeastPost (Signed str) -> return $ CGB.AtLeast False $ read str
   AtLPostCaut1 (Signed str) -> return $ CGB.AtLeast True $ read str
   AtLPostCaut2 (Signed str) -> return $ CGB.AtLeast True $ read str
+  Subreading (Signed str) _ -> return $ CGB.Exactly False $ read str
   Cautious position        -> cautious `fmap` transPosition position
   where cautious (CGB.Exactly _b num) = CGB.Exactly True num
         cautious (CGB.AtLeast _b num) = CGB.AtLeast True num

### 2015-10-18 paper/lrec/eval_futurework.tex
more text
more text/pictures

@@
 \section{Evaluation and Future Work}
+
+We tested three grammars for rules that cannot apply. The smallest grammar was Dutch, with 68 rules; second was Spanish with 279 rules, and the largest was Finnish, with 1185 rules.
+For the smaller grammars, we were able to verify manually that the results hold.
+
+\begin{table}[]
+\centering
+\begin{tabular}{|l|l|l|l|}
+\hline
+                         & \textsc{nld}  & \textsc{spa}  & \textsc{fin}  \\ \hline
+Rules in grammar      & 68              & 279               & 1185              \\ \hline
+Redundant rules found & 1               & ???               & ???    \\ \hline
+Running time          & n sec           & ??? min           & ??? h    \\ \hline
+\end{tabular}
+\caption{Results}
+\label{table:res}
+\end{table}
 

### 2015-10-18 paper/lrec/implementation.tex
more text/pictures

@@
+\begin{table*}[]
+\centering
+\begin{tabular}{|l|l|l|l|}
+\hline
+\textbf{Word} & \textbf{Variables}                                                          & \textbf{Default rules}               & \textbf{Constraint rules}                                  \\ \hline
+``the''         & \texttt{the\_det}                                                           & \texttt{the\_det}                    & \multirow{2}{*}{\texttt{the\_det}  $\Rightarrow$ $\neg$\texttt{bear\_v\_pl}} \\ \cline{1-3}
+``bear''        & \begin{tabular}[c]{@{}l@{}}\texttt{bear\_n\_sg} \\ \texttt{bear\_v\_pl}\end{tabular} & \texttt{bear\_n\_sg} $\vee$ \texttt{bear\_v\_pl} &                                                            \\ \hline
+\end{tabular}
+\caption{SAT-encoding of \texttt{REMOVE v IF (-1 det)} applied to \emph{the bear}}
+\label{table:vars}
+\end{table*}
+
 \section{Implementation}
@@
 See \cite{listenmaa_claessen2015} for further description.
-
-\begin{table*}[]
-\centering
-\caption{SAT-encoding of \texttt{REMOVE v IF (-1 det)} applied to \emph{the bear}}
-\label{table:vars}
-\begin{tabular}{|l|l|l|l|}
-\hline
-\textbf{Word} & \textbf{Variables}                                                          & \textbf{Default rules}               & \textbf{Constraint rules}                                  \\ \hline
-``the''         & \texttt{the\_det}                                                           & \texttt{the\_det}                    & \multirow{2}{*}{\texttt{the\_det}  $\Rightarrow$ $\neg$\texttt{bear\_v\_pl}} \\ \cline{1-3}
-``bear''        & \begin{tabular}[c]{@{}l@{}}\texttt{bear\_n\_sg} \\ \texttt{bear\_v\_pl}\end{tabular} & \texttt{bear\_n\_sg} $\vee$ \texttt{bear\_v\_pl} &                                                            \\ \hline
-\end{tabular}
-\end{table*}
 
@@
 However, being a condition for another target will also shape a word.
-At a level of real texts, a rule such as \texttt{REMOVE b IF (-1 a)} is an instruction to remove a \texttt{b} if we find an \texttt{a} before it.
-In case of a symbolic sentence, we start in a situation where every word can be an \texttt{a} or \texttt{b}.
-In that context, the rule behaves as a statement about that \texttt{b} may not follow \texttt{a}.
+At a level of real texts, a rule such as \texttt{REMOVE verb IF (-1 det)} is an instruction to remove a \texttt{verb} if we find a \texttt{det} before it.
+In case of a symbolic sentence, we start in a situation where every word can be a \texttt{det} or a \texttt{verb}.
+In that context, the rule behaves as a broader constraint about every pair of two words \texttt{verb} may not follow \texttt{det}.
 This relation is encoded between all adjacent words, and each of them can follow the restriction in the way that is compliant with other rules.
-For instance, word 1 may be constrained by another rule to have no \texttt{a}, and this means that word 2 is free to have \texttt{b} or not have it.
-If word 3 is required to have an  \texttt{a}, then word 4 either has to have no \texttt{b}, or it has to have only \texttt{b}, by the rule that the last analysis of the word will not be removed.
+
+\begin{figure}[]
+
+\includegraphics[width=0.42\textwidth]{symbolic_sentence.png}
+
+\caption{Symbolic sentence beginning to take shape}
+\label{fig:example}
+\end{figure}
+
+% For instance, word 1 may be constrained by another rule to have no \texttt{det}, and this means that word 2 is free to have \texttt{verb} or not have it.
+If word 3 is required to have a  \texttt{det}, then word 4 either has to have no \texttt{verb}, or it has to have only \texttt{verb}, by the rule that the last analysis of the word will not be removed.
 

### 2015-10-19 paper/lrec/lrec2016_listenmaa_claessen.tex
stuff
add abstract stub
more text/pictures
add bunch of packages, move table from implementation here for positioning it to second page >_>

@@
 \usepackage{url}
+\usepackage{multirow}
 \usepackage{latexsym}
 \usepackage{cite}
+%\usepackage{graphics}
+\usepackage{graphicx}
+\usepackage{float}
 \usepackage[colorlinks=true,citecolor=blue,urlcolor=blue]{hyperref}
@@
 \usepackage{multirow}
-\special{papersize=210mm,297mm} % to avoid having to use "-t a4" with dvips 
+%\special{papersize=210mm,297mm} % to avoid having to use "-t a4" with dvips 
 %\setlength\titlebox{6.5cm}  % You can expand the title box if you really have to
+
+\usepackage{fontspec}
+
+\newcommand{\mainfontfamily}{Palatino}
+\setmainfont[
+  Ligatures=TeX,
+  Scale=0.95,
+]{\mainfontfamily}
+% \newfontfamily\scshape[Letters=SmallCaps, Numbers=Uppercase] % {Myriad Pro} % {TeX Gyre Pagella}
+
+\newcommand{\monofontfamily}{Monaco} % Courier
+\setmonofont[Scale=0.8]{\monofontfamily}
 
@@
 \newcommand{\todo}[1]{{\color{cyan}\textbf{[TODO: }#1\textbf{]}}}
-\usepackage{tikz}
-\def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) --
-  (.25,.15) -- cycle;}
-\newcommand{\Mypm}{\mathbin{\tikz [x=1.4ex,y=1.4ex,line width=.1ex] \draw (0.0,0) -- (1.0,0) (0.5,0.08) -- (0.5,0.92) (0.0,0.5) -- (1.0,0.5);}}%
+% \usepackage{tikz}
+% \def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) --
+%   (.25,.15) -- cycle;}
+% \newcommand{\Mypm}{\mathbin{\tikz [x=1.4ex,y=1.4ex,line width=.1ex] \draw (0.0,0) -- (1.0,0) (0.5,0.08) -- (0.5,0.92) (0.0,0.5) -- (1.0,0.5);}}%
 
@@
 \maketitle
+
+\begin{abstract}
+We describe a method for analysing Constraint Grammar.
+We encode a grammar as a SAT problem and find out
+whether there are conflicting or redundant rules.
+With a morphological lexicon, we can construct example sequences that
+have realistic ambiguities.
+We test our tools with grammars from 100-1000 rules.
+\end{abstract}
+
+\begin{table*}[]
+\centering
+\begin{tabular}{|l|l|l|l|}
+\hline
+\textbf{Word} & \textbf{Variables}                                                          & \textbf{Default rules}               & \textbf{Constraint rules}                                  \\ \hline
+``the''         & \texttt{the\_det}                                                           & \texttt{the\_det}                    & \multirow{2}{*}{\texttt{the\_det}  $\Rightarrow$ $\neg$\texttt{bear\_v\_pl}} \\ \cline{1-3}
+``bear''        & \begin{tabular}[c]{@{}l@{}}\texttt{bear\_n\_sg} \\ \texttt{bear\_v\_pl}\end{tabular} & \texttt{bear\_n\_sg} $\vee$ \texttt{bear\_v\_pl} &                                                            \\ \hline
+\end{tabular}
+\caption{SAT-encoding of \texttt{REMOVE v IF (-1 det)} applied to \emph{the bear}}
+\label{table:vars}
+\end{table*}
 
@@
 \bibliographystyle{acl}
-\bibliography{../cg}
+\bibliography{cg}
 

### 2015-10-19 paper/lrec/implementation.tex
text

@@
-\begin{table*}[]
-\centering
-\begin{tabular}{|l|l|l|l|}
-\hline
-\textbf{Word} & \textbf{Variables}                                                          & \textbf{Default rules}               & \textbf{Constraint rules}                                  \\ \hline
-``the''         & \texttt{the\_det}                                                           & \texttt{the\_det}                    & \multirow{2}{*}{\texttt{the\_det}  $\Rightarrow$ $\neg$\texttt{bear\_v\_pl}} \\ \cline{1-3}
-``bear''        & \begin{tabular}[c]{@{}l@{}}\texttt{bear\_n\_sg} \\ \texttt{bear\_v\_pl}\end{tabular} & \texttt{bear\_n\_sg} $\vee$ \texttt{bear\_v\_pl} &                                                            \\ \hline
-\end{tabular}
-\caption{SAT-encoding of \texttt{REMOVE v IF (-1 det)} applied to \emph{the bear}}
-\label{table:vars}
-\end{table*}
-
 \section{Implementation}
@@
 
-In this section, we describe the implementation of the toolkit.
+In this section, we describe the implementation of the toolset.
 We encode CG rules as a SAT problem, as in \cite{listenmaa_claessen2015}, and from the morphological lexicon, we get constraints for individual words.
@@
 
+% table:vars moved to main file for positioning
+
 \paragraph{Word-internal constraints}
 
-Given a large morphological lexicon, it is relatively easy to restrict each individual word to contain only realistic ambiguities.
+Given a large morphological lexicon, we can restrict each individual word to contain only realistic ambiguities.
 We expand the lexicon, map each wordform to its possible analyses, and then ignore the actual wordforms.
 For instance, a Spanish morphological lexicon contains the entries \texttt{casa:<verb><sg><p3>} and \texttt{casa:<noun><sg>}, hence we know that a confusion between a 3rd person singular verb and a singular noun is possible.
-The lexicon is not likely to contain a wordform that can be analysed both as an adjective and punctuation, so that kind of ambiguous word is never created.
+%The lexicon is not likely to contain a wordform that can be analysed both as an adjective and punctuation, so that kind of ambiguous word is never created.
 

### 2015-10-19 paper/lrec/lrec2016_listenmaa_claessen.tex
text

@@
 \usepackage{graphicx}
-\usepackage{float}
+%\usepackage{float}
 \usepackage[colorlinks=true,citecolor=blue,urlcolor=blue]{hyperref}
@@
 \begin{abstract}
-We describe a method for analysing Constraint Grammar.
-We encode a grammar as a SAT problem and find out
-whether there are conflicting or redundant rules.
-With a morphological lexicon, we can construct example sequences that
-have realistic ambiguities.
-We test our tools with grammars from 100-1000 rules.
+We describe a method for analysing Constraint Grammar by encoding it
+as a SAT problem.
+Our tools can detect internal conflicts or redundancies in a grammar,
+as well as generate example sequences to demonstrate the effect of some rule or rule set.
+No corpus is required, only a morphological lexicon.
+%We test our tools with grammars from 100-1000 rules.
 \end{abstract}

### 2015-10-19 paper/lrec/problem_desc.tex
more text
text

@@
 % \citep[CG][]{karlsson1995constraint} 
-is a method for disambiguating morphologically analysed text% and adding syntactic or dependency labels
+is a method for disambiguating morphologically analysed text
+% and adding syntactic or dependency labels
 . 
@@
 \item Generate a sequence that triggers rule(s) X but not rule(s) Y
-\item Generate a sequence that is ambiguous but doesn't trigger any rules
+%\item Generate a sequence that is ambiguous but doesn't trigger any rules
 \end{itemize}
@@
 
-In addition, there is a large body of research on automatically inducing rules, e.g. \cite{inducing_cg1996}, \cite{lindberg_eineborg98ilp}%, \cite{lager01transformation}, \cite{asfrent14} 
+In addition, there is a large body of research on automatically inducing rules, e.g. \cite{inducing_cg1996}, \cite{lindberg_eineborg98ilp}
+%, \cite{lager01transformation}, \cite{asfrent14} 
 . However, since our work is aimed to aid the process of hand-crafting rules, we omit those works from our discussion.
@@
 Hand-annotated corpora are commonly used in the development of CGs, because they give immediate feedback whether a new rule increases or decreases accuracy, and by how large margin \cite{voutilainen2004}.
+This helps the grammar writer to arrange the rules in appropriate sections, from safest and most effective to less certain and unusual. 
+However, this method will not notice a missed opportunity or a grammar-internal conflict, nor suggest ways to improve.
+
+% -- uncomment for full version ?
+
 %\cite{voutilainen2004} gives a detailed account about best practices of grammar writing and efficient use of corpora to aid the grammar development.
@@
 
-CG rules are usually arranged in sections, and run in the following manner. 
-First apply rules from section 1, and repeat until nothing changes in the text. Then apply rules from sections 1--2, then 1--3 and so on, until the set includes all rules.
-The best strategy is to place the safest and most effective rules in the first sections,
-so that they make way for the following, more heuristic and less safe rules to act on.
+% CG rules are usually arranged in sections, and run in the following manner. 
+% First apply rules from section 1, and repeat until nothing changes in the text. Then apply rules from sections 1--2, then 1--3 and so on, until the set includes all rules.
+% The best strategy is to place the safest and most effective rules in the first sections,
+% so that they make way for the following, more heuristic and less safe rules to act on.
 
-A representative corpus is arguably the best way to get concrete numbers---how many times a rule applied and how often it was correct---and to arrange the rules in sections based on that feedback.
-However, this method will not notice a missed opportunity or a grammar-internal conflict, nor suggest ways to improve.
-%, for instance changing the place of a rule.
+% A representative corpus is arguably the best way to get concrete numbers---how many times a rule applied and how often it was correct---and to arrange the rules in sections based on that feedback.
+% However, this method will not notice a missed opportunity or a grammar-internal conflict, nor suggest ways to improve.
 
 % \cite{voutilainen2004} state that the around 200 rules are probably enough to resolve 50--75 \% of ambiguities in the corpus used in the development.
-% This is invaluable, especially at the beginning stages of grammar writing. 
-% Sometimes you just want to know that a rule fired 500 times and was correct 490 times, rather than ``it is possible to construct a word sequence that will trigger this rule''.
 
-%The connection to the reality has also its limitations. 
-% What if instead of firing 500 times, it could have fired 1000 times if it had been placed earlier in the rule sequence.
-%Or the rule that was wrong in 60 \% of the time,
+\paragraph{Automatic optimisation of hand-written grammars }
+% The corpus-based method can tell the effect of each single rule at their place in the rule sequence, and leaves the grammar writer to make changes in the grammar.
 
-\paragraph{Optimising hand-written grammars}
-The corpus-based method can tell the effect of each single rule at their place in the rule sequence, and it is up to the grammar writer to make changes in the grammar.
-\cite{bick2013tuning} automates the latter process: the rules are moved into different sections or removed in total based on their effect, and in parallel, their contexts are made stricter or less strict.
-After trying out several combinations, the version that produces the best results is kept.
-
-This is a valuable tool for handling grammars that are too big to keep track of, and Bick's evaluation shows that the tuned grammars perform better than the original.
-As a downside, the grammar writer will likely not know why exactly is the tuned grammar performing better.
+\cite{bick2013tuning} modifies the grammar automatically, by trying out different orderings of the rules, and changing their contexts to stricter or less strict.
+Bick reports error reductions of 7--15\% compared to the original grammars.
+However, the grammar writer will likely not know why exactly is the tuned grammar performing better.
 % At a certain point, the grammar gets so big that it is hard to keep track of all the rules and their interactions.

### 2015-10-19 paper/lrec/lrec2016_listenmaa_claessen.tex
more stuff to table

@@
 ]{\mainfontfamily}
-% \newfontfamily\scshape[Letters=SmallCaps, Numbers=Uppercase] % {Myriad Pro} % {TeX Gyre Pagella}
+%\newfontfamily\scshape[Letters=SmallCaps, Numbers=Uppercase] % {Myriad Pro} % {TeX Gyre Pagella}
 
@@
 Our tools can detect internal conflicts or redundancies in a grammar,
-as well as generate example sequences to demonstrate the effect of some rule or rule set.
+as well as generate examples to demonstrate the effect of some rule or rule set.
 No corpus is required, only a morphological lexicon.
@@
 \hline
-\textbf{Word} & \textbf{Variables}                                                          & \textbf{Default rules}               & \textbf{Constraint rules}                                  \\ \hline
-``the''         & \texttt{the\_det}                                                           & \texttt{the\_det}                    & \multirow{2}{*}{\texttt{the\_det}  $\Rightarrow$ $\neg$\texttt{bear\_v\_pl}} \\ \cline{1-3}
-``bear''        & \begin{tabular}[c]{@{}l@{}}\texttt{bear\_n\_sg} \\ \texttt{bear\_v\_pl}\end{tabular} & \texttt{bear\_n\_sg} $\vee$ \texttt{bear\_v\_pl} &                                                            \\ \hline
+
+\textbf{Analyses}                                                                                                 & \textbf{Variables}                                                          & \textbf{Default rules}                                    & \textbf{Constraint rules}                                                          \\ \hline
+\multicolumn{1}{|l|}{\begin{tabular}[c]{@{}l@{}}``\textless{}the\textgreater''\\$\qquad\quad$"the" det\end{tabular}}     & \multicolumn{1}{l|}{\texttt{the\_det}}                                      & \multicolumn{1}{l|}{\texttt{the\_det}}                    & \multicolumn{1}{l|}{\multirow{2}{*}{\texttt{the\_det $\Rightarrow$ $\neg$bear\_v\_pl}}} \\ \cline{1-3}
+\begin{tabular}[c]{@{}l@{}}"\textless{}bear\textgreater"\\$\qquad\quad$"bear" n sg\\$\qquad\quad$"bear" v pl\end{tabular} & \texttt{\begin{tabular}[c]{@{}l@{}}bear\_n\_sg \\ bear\_v\_pl\end{tabular}} & \multicolumn{1}{l|}{\texttt{bear\_n\_sg $\vee$ bear\_v\_pl}} & \multicolumn{1}{l|}{}                                                              \\ \hline
 \end{tabular}

### 2015-10-19 paper/lrec/implementation.tex
more text

@@
 In this section, we describe the implementation of the toolset.
-We encode CG rules as a SAT problem, as in \cite{listenmaa_claessen2015}, and from the morphological lexicon, we get constraints for individual words.
-This allows us to create word sequences that will or will not trigger a given rule or rule set.
+\cite{listenmaa_claessen2015} presents the SAT encoding of the rules, when applied to an existing text.
+Since we want to test the rules in a way that doesn't depend on data, we create a \emph{symbolic sentence}: a word sequence which we can modify according to the properties we want to test.
+These words are combinations of tags
+In order to create words with real ambiguities, we use a morphological lexicon
+This sentence consists of symbolic words, which we c
+In addition, we use a morphological lexicon to create symbolic words that contain real ambiguities.
 
-\paragraph{SAT encoding of CG}
+\paragraph{SAT encoding of CG rules}
 
-Table~\ref{table:vars} shows the translation of the analyses to variables, and rules to clauses. The column ``Default rules'' refers to the rule that each word must have at least one analysis.
-See \cite{listenmaa_claessen2015} for further description.
+A single CG rule is a template which creates a concrete clause for each instance where it is applied.
+%Each potential analysis is represented as a variable, and the rules are translated into clauses that operate on those variables.
+Table~\ref{table:vars} shows the rule \texttt{REMOVE v IF (-1 det)} applied to the text \emph{the bear}.
+The analyses of the word forms are represented as variables, and
+the default rule ensures that each word has at least one analysis.
+%This makes \texttt{the\_det} true by default, therefore the implication must hold, and \texttt{bear\_v\_pl} will be removed.
+For more details about the encoding, see \cite{listenmaa_claessen2015}.
+
+%As shown in table~\ref{table:vars}, \texttt{bear\_n\_sg} represents the possibility that \emph{bear} is a noun, and \texttt{bear\_v\_pl} 
+%The rules are translated into clauses; one for each concrete application.
+%Table~\ref{table:vars} shows the translation of the analyses to variables, and rules to clauses. See \cite{listenmaa_claessen2015} for further description.
 
@@
 
+The symbolic sentence consists of symbolic words.
+They are combinations of tags, and they may be underspecified---similarly, a given CG rule may target any noun, a plural noun or the noun \emph{bear}.
 Given a large morphological lexicon, we can restrict each individual word to contain only realistic ambiguities.
-We expand the lexicon, map each wordform to its possible analyses, and then ignore the actual wordforms.
-For instance, a Spanish morphological lexicon contains the entries \texttt{casa:<verb><sg><p3>} and \texttt{casa:<noun><sg>}, hence we know that a confusion between a 3rd person singular verb and a singular noun is possible.
+We map each word form in the lexicon to its possible analyses, and then ignore the actual word forms.
+For instance, a Spanish morphological lexicon contains the entries \texttt{casa:casar<verb><sg><p3>} and \texttt{casa:casa<noun><sg>}, hence we know that a confusion between a 3rd person singular verb and a singular noun is possible.
 %The lexicon is not likely to contain a wordform that can be analysed both as an adjective and punctuation, so that kind of ambiguous word is never created.

### 2015-10-19 paper/lrec/lrec2016_listenmaa_claessen.tex
more text

@@
 
-\textbf{Analyses}                                                                                                 & \textbf{Variables}                                                          & \textbf{Default rules}                                    & \textbf{Constraint rules}                                                          \\ \hline
+\textbf{Analyses}
+& \textbf{Variables}
+& \textbf{Default rule: word has $\geq$1 analysis}                                    & \textbf{Constraint rules}                                                          \\ \hline
 \multicolumn{1}{|l|}{\begin{tabular}[c]{@{}l@{}}``\textless{}the\textgreater''\\$\qquad\quad$"the" det\end{tabular}}     & \multicolumn{1}{l|}{\texttt{the\_det}}                                      & \multicolumn{1}{l|}{\texttt{the\_det}}                    & \multicolumn{1}{l|}{\multirow{2}{*}{\texttt{the\_det $\Rightarrow$ $\neg$bear\_v\_pl}}} \\ \cline{1-3}
@@
 \end{tabular}
-\caption{SAT-encoding of \texttt{REMOVE v IF (-1 det)} applied to \emph{the bear}}
+\caption{\texttt{REMOVE v IF (-1 det)} applied to \emph{the bear}}
 \label{table:vars}

### 2015-10-19 paper/lrec/problem_desc.tex
more text

@@
 Hand-annotated corpora are commonly used in the development of CGs, because they give immediate feedback whether a new rule increases or decreases accuracy, and by how large margin \cite{voutilainen2004}.
-This helps the grammar writer to arrange the rules in appropriate sections, from safest and most effective to less certain and unusual. 
+This helps the grammar writer to arrange the rules in appropriate sections, with safest and most effective rules coming first.
 However, this method will not notice a missed opportunity or a grammar-internal conflict, nor suggest ways to improve.
@@
 \cite{bick2013tuning} modifies the grammar automatically, by trying out different orderings of the rules, and changing their contexts to stricter or less strict.
-Bick reports error reductions of 7--15\% compared to the original grammars.
-However, the grammar writer will likely not know why exactly is the tuned grammar performing better.
+Bick reports error reduction of 7--15\% compared to the original grammars.
+As a downside, the grammar writer will likely not know why exactly is the tuned grammar performing better.
 % At a certain point, the grammar gets so big that it is hard to keep track of all the rules and their interactions.

### 2015-10-19 paper/lrec/eval_futurework.tex
add running time of finnish
add time for dutch
more text

@@
 \section{Evaluation and Future Work}
+\label{sec:eval}
 
-We tested three grammars for rules that cannot apply. The smallest grammar was Dutch, with 68 rules; second was Spanish with 279 rules, and the largest was Finnish, with 1185 rules.
-For the smaller grammars, we were able to verify manually that the results hold.
+We tested three grammars to find rules that cannot apply. The smallest grammar was Dutch, with 68 rules; second was Spanish with 279 rules, and the largest was Finnish, with 1185 rules.
+For the smaller grammars, we were able to verify manually that the results are true.
 
@@
 Redundant rules found & 1               & ???               & ???    \\ \hline
-Running time          & n sec           & ??? min           & ??? h    \\ \hline
+Running time          & 7s              & 3min 22s          & 3h 26min    \\ \hline
 \end{tabular}
@@
 \end{table}
+
+\todo{Other things we found from the grammar:}
+With the Spanish grammar, the method pointed us to a nonsensical set definition.
+\texttt{w2<n><np><adj><det><preadv><adv><vblex><vbmod><vbhaver><vbser><prn><pr><cnjcoo><cnjsub><cnjadv><rel><ij>}
+
+For future work, we want to improve the performance and scale up to larger grammars, and handle the full expressivity of CG-3, with \textsc{map}, \textsc{add} and \textsc{substitute} rules.
+We want to also ask different questions, such as \todo{...}.
+Finally, we want to test the approach to other grammar formalisms.
 

### 2015-10-19 paper/lrec/implementation.tex
more text

@@
 In this section, we describe the implementation of the toolset.
-\cite{listenmaa_claessen2015} presents the SAT encoding of the rules, when applied to an existing text.
-Since we want to test the rules in a way that doesn't depend on data, we create a \emph{symbolic sentence}: a word sequence which we can modify according to the properties we want to test.
-These words are combinations of tags
-In order to create words with real ambiguities, we use a morphological lexicon
-This sentence consists of symbolic words, which we c
-In addition, we use a morphological lexicon to create symbolic words that contain real ambiguities.
+The SAT encoding we use is introduced in \cite{listenmaa_claessen2015}, along with a working CG compiler.
+Since we want to test the rules in a way that doesn't depend on data, we operate on a \emph{symbolic sentence}: a word sequence which we can modify according to the properties we want to test.
 
@@
 
-%As shown in table~\ref{table:vars}, \texttt{bear\_n\_sg} represents the possibility that \emph{bear} is a noun, and \texttt{bear\_v\_pl} 
-%The rules are translated into clauses; one for each concrete application.
-%Table~\ref{table:vars} shows the translation of the analyses to variables, and rules to clauses. See \cite{listenmaa_claessen2015} for further description.
-
 % table:vars moved to main file for positioning
@@
 The symbolic sentence consists of symbolic words.
-They are combinations of tags, and they may be underspecified---similarly, a given CG rule may target any noun, a plural noun or the noun \emph{bear}.
-Given a large morphological lexicon, we can restrict each individual word to contain only realistic ambiguities.
-We map each word form in the lexicon to its possible analyses, and then ignore the actual word forms.
+In the beginning, a symbolic word is a list of all possible tag combinations, and we don't know which ones are true.
+%These tag combinations may be underspecified for lemma or other features---similarly, a CG rule may target any noun, a plural noun or the noun \emph{bear}.
+Given a large morphological lexicon, we can restrict the individual words to contain only realistic ambiguities. 
 For instance, a Spanish morphological lexicon contains the entries \texttt{casa:casar<verb><sg><p3>} and \texttt{casa:casa<noun><sg>}, hence we know that a confusion between a 3rd person singular verb and a singular noun is possible.
@@
 
-When we start applying rules, we need to have a sentence to which to apply.
-As this is a corpus-free method, we don't have one.
-Instead, we create a \emph{symbolic sentence}: sequence of words that can, in the beginning, have any tag combination.
-At the word level, these tag combinations are restricted to only contain realistic ambiguities, as described in the previous paragraph.
-At the sentence level, applying rules will start shaping the sentence.
-An indeterminate blob of possible words undergoes rules that describe illegal sequences of analyses,
-and at the application of each rule, more and more possibilities are discarded.
+Figure~\ref{fig:example} shows a fragment of of a two-word symbolic sentence, before and after applying the rule \texttt{REMOVE verb IF (-1 det)}\footnote{For simplicity, we don't show the word-internal constraints that come from the lexicon, but only the rule application.}.
+As with real analyses, each tag combination is given a variable, and applying a rule means creating a concrete clause for each instance where a determiner is found preceding a verb.
+In a symbolic sentence, every word has potentially every tag combination, and applying the rules will shape the sentence into something that respects the constraints of the language.
 
-All rules are applied to all word, given that their context is in scope: for instance, the sentence-initial word doesn't meet any condition that requires something of previous words.
-However, being a condition for another target will also shape a word.
-At a level of real texts, a rule such as \texttt{REMOVE verb IF (-1 det)} is an instruction to remove a \texttt{verb} if we find a \texttt{det} before it.
-In case of a symbolic sentence, we start in a situation where every word can be a \texttt{det} or a \texttt{verb}.
-In that context, the rule behaves as a broader constraint about every pair of two words \texttt{verb} may not follow \texttt{det}.
-This relation is encoded between all adjacent words, and each of them can follow the restriction in the way that is compliant with other rules.
-
-\begin{figure}[]
-
+\begin{figure}[h]
 \includegraphics[width=0.42\textwidth]{symbolic_sentence.png}
-
-\caption{Symbolic sentence beginning to take shape}
+\caption{Symbolic sentence beginning to take shape \todo{make this less confusing?}}
 \label{fig:example}
@@
 
-% For instance, word 1 may be constrained by another rule to have no \texttt{det}, and this means that word 2 is free to have \texttt{verb} or not have it.
-If word 3 is required to have a  \texttt{det}, then word 4 either has to have no \texttt{verb}, or it has to have only \texttt{verb}, by the rule that the last analysis of the word will not be removed.
+The example rule states that the sentence may not have a verb after a determiner.
+The variable \texttt{w2<verb>} is the situation of the second word's verb status before applying the rule.
+We create a new variable \texttt{w2'<verb>}, whose value depends on whether the condition holds (if \texttt{w1<det>} is true); whether it was true on the last round (if \texttt{w2<verb>} is already removed by another rule, \texttt{w2'<verb>} cannot reappear); and whether it is the only analysis left (not depicted in the picture).
+All the analyses not targeted by this rule remain the same.
 
-The applications of these rules shape our symbolic sentence into containing only things that can survive the rules.
-This means that we can ask a question ``after applying rules $0-i$, is it possible for $i+1$ to apply?'', and the answer is reliable; we can not find a counterexample by just having a bit more data, because we start from literally all possibilities, and each rule
-restricts the set of possible sentences.
+% \begin{figure}[]
+% \includegraphics[width=0.42\textwidth]{symbolic_sentence.png}
+% \caption{Symbolic sentence beginning to take shape \todo{make this less confusing?}}
+% \label{fig:example}
+% \end{figure}
 
-Note that this method is not suitable for language generation: rules in any realistic CG are tailored for real life ambiguities, whereas our symbolic sentence starts with literally all possibilities of tag combinations.
+%The applications of these rules shape our symbolic sentence into containing only things that can survive the rules.
+
+The effect of the old rule applications is retained when a new rule is added.
+This means that we can ask a question ``after applying rules $0-i$, is it possible for $i+1$ to apply?'', and the answer is reliable; we cannot find a counterexample by just having a bit more data, because we start from literally all possibilities, and each rule restricts the set of possible sentences.
+%If rule \texttt{r100} would require the first word to be a noun, and causes a conflict because it is not a noun, we know that it is because there is no way for w1 to be a noun.
+
+%Note that this method is not suitable for language generation: rules in any realistic CG are tailored for real life ambiguities, whereas our symbolic sentence starts with literally all possibilities of tag combinations.
 
@@
 The example so far has been of type ``Can rule Z apply after A-Y?''
+In case a conflict is detected, we can dig into it and refine where the problem is: e.g. is the condition not met, or is it the case that the target is already removed or the only remaining analysis.
+We can find what other rule or rules prevent others from applying, or if the conflict is rule-internal, such as nonexisting tag set or contradicting requirements of a context word (e.g.
+a word must be unambiguously two different POS).
 
-We can construct all kinds of symbolic sentences. We can set the length, restrict individual words (e.g. ``3rd word must be a noun''), require that it triggers some rule but not other.
-In case of a conflict, we can dig into the conflict and refine where the problem is: e.g. is the condition not met, or is it the case that the target is already removed or the only remaining analysis.
+In addition to analysing a whole grammar, 
+we can construct all kinds of symbolic sentences to test out individual rules. 
+We can set the length, restrict individual words (e.g. ``3rd word must be a noun''), require that it triggers some rule but not other. 
+This functionality can aid the grammar writer in the process, to see if they have missed a case or defined the tagset correctly.

### 2015-10-19 paper/lrec/problem_desc.tex
more text

@@
 
-The paper is structured as follows.
-Section~\ref{sec:prev} relates our work to the previous research.
-Section~\ref{sec:implementation} discusses the implementation, Section~X presents preliminary results and  Section~Y discusses future work.
-Section~Z concludes the paper.
+%The paper is structured as follows. Section~\ref{sec:prev} relates our work to the previous research. 
+% Section~\ref{sec:implementation} discusses the implementation, 
+% Section~\ref{sec:eval} presents preliminary results and discusses future work. 
+%Section~\ref{sec:conclusion} concludes the paper.
 
@@
 
-Hand-annotated corpora are commonly used in the development of CGs, because they give immediate feedback whether a new rule increases or decreases accuracy, and by how large margin \cite{voutilainen2004}.
+Hand-annotated corpora are commonly used in the development of CGs, because they give immediate feedback whether a new rule increases or decreases accuracy \cite{voutilainen2004}.
 This helps the grammar writer to arrange the rules in appropriate sections, with safest and most effective rules coming first.
@@
 %The earlier works on logical reconstruction don't envision grammar analysis as one of the use cases,
-\cite{lager98} and \cite{lager_nivre01} reconstruct the CG formalism in first-order predicate logic, and \cite{listenmaa_claessen2015} implement a CG compiler using a SAT solver.
+\cite{lager98} and \cite{lager_nivre01} reconstruct the CG formalism in first-order predicate logic.
+%, and \cite{listenmaa_claessen2015} implement a CG compiler using a SAT solver.
 Grammar analysis is a natural use case, due to some key features of the logical reconstruction.
@@
 
-% Application of rules is presented as a composition of clauses
-
-% They model CG as clauses o
 % The key elements in the ontology of CG are positions, words and sets of tags.
@@
 % The rules are modelled as implications, of the form below:
-
-% \begin{itemize}
-% \item Interaction of rules
-% \item Create word sequences to test what is possible and what can never happen
-% \end{itemize}
 

### 2015-10-21 paper/lrec/eval_futurework.tex
stuff

@@
 
-\todo{Other things we found from the grammar:}
-With the Spanish grammar, the method pointed us to a nonsensical set definition.
-\texttt{w2<n><np><adj><det><preadv><adv><vblex><vbmod><vbhaver><vbser><prn><pr><cnjcoo><cnjsub><cnjadv><rel><ij>}
+% Instead of whole grammar, look at specific rules, test with different number of rules before
+
+\todo{Analyse the results.
+I didn't even touch the Finnish results yet.
+Maybe mention other things we found from the grammar, like the nonsensical set definition from Spanish grammar (confused and and or, requiring a word to be noun AND verb AND conj AND preposition ...).
+}
+
+Currently, the speed is not great. The Finnish grammar approaches a realistic size of grammars.
+The size of the SAT problem was \todo{some real numbers and analysis here \& promises that it can be fixed}.
+
+% w2<n><np><adj><det><preadv><adv><vblex><vbmod><vbhaver><vbser><prn><pr><cnjcoo><cnjsub><cnjadv><rel><ij>
 
@@
 We want to also ask different questions, such as \todo{...}.
+We can be smarter about SAT things.
+
+Compare with a corpus-based test: do we find the same problems?
+
 Finally, we want to test the approach to other grammar formalisms.

### 2015-10-22 paper/lrec/eval_futurework.tex
more text

--- 2015-10-19 paper/lrec/eval_futurework.tex
@@
 
-We tested three grammars to find rules that cannot apply. The smallest grammar was Dutch, with 68 rules; second was Spanish with 279 rules, and the largest was Finnish, with 1185 rules.
+We tested three grammars to find rules that cannot apply.
+The smallest grammar was Dutch\footnote{\url{https://svn.code.sf.net/p/apertium/svn/languages/apertium-nld/apertium-nld.nld.rlx}}, with 59 rules; second was Spanish\footnote{\url{https://svn.code.sf.net/p/apertium/svn/languages/apertium-spa/apertium-spa.spa.rlx}}with 279 rules, and the largest was Finnish\footnote{\url{https://github.com/flammie/omorfi/blob/master/src/vislcg3/omorfi.cg3}}, with 1185 rules.
 For the smaller grammars, we were able to verify manually that the results are true.
@@
                       & \textsc{nld}  & \textsc{spa}  & \textsc{fin}  \\ \hline
-Rules in grammar      & 68              & 279               & 1185              \\ \hline
-Redundant rules found & 1               & ???               & ???    \\ \hline
-Running time          & 7s              & 3min 22s          & 3h 26min    \\ \hline
+Rules in grammar      & 59              & 279               & 1185     \\ \hline
+Conflicts found       & 1               & ???               & ???    \\ \hline
+Time: all rules       & 7s              & 3min 22s          & 3h 26min    \\ \hline
+Time: last rule       & 4s              & 7s                & 3min 40s    \\ \hline
+Time: longest rule    & ?s              & ?s                & ?min     \\ \hline
+
 \end{tabular}
@@
 
-\todo{Other things we found from the grammar:}
-With the Spanish grammar, the method pointed us to a nonsensical set definition.
-\texttt{w2<n><np><adj><det><preadv><adv><vblex><vbmod><vbhaver><vbser><prn><pr><cnjcoo><cnjsub><cnjadv><rel><ij>}
+The first row shows the whole number of rules. For the Finnish grammar, we took only \textsc{remove} and \textsc{select} rules. The other two only contained those rules to start with.
+
+We ran for all three grammars the following thing.
+For each rule, run all the rules before it to the symbolic sentence, and then ask if that rule can fire.
+This test revealed problems in all grammars.
+The Dutch grammar had the following rule sequence, making the latter impossible to apply:
+
+\begin{itemize}
+\item[] \texttt{REMOVE adv IF (1 n|np) ;}
+\item[] \texttt{REMOVE adv IF (-1 det) (0 adj) (1 n|np) ;}
+\end{itemize}
+
+The Spanish grammar had an erroneous set definition, where a word was required to have almost all POS tags at the same time. After fixing the set definition, our test didn't reveal other problems.
+The program marked __ problem rules for the Finnish grammar, but we didn't have time to verify if they were problematic or if our program is buggy.
+
+As seen in table~\ref{table:res}, checking every rule in the Finnish grammar takes currently 3.5 hours.
+A rule takes longer to run if it appears later in the rule sequence: 1000 previous rules is a larger SAT problem than 10 previous rules.
+This can be seen from the time it took to compute the result for the last rule.
+For Finnish, the last rule took almost 4 minutes---had this been the average, the whole grammar would have taken 18 hours instead of 3.5.
+We also measured the running times for the longest rules in the grammar.
+\todo{find longest rule, change place of that rule, see what matters more. Promise: we can be smarter about SAT things.}
 
 For future work, we want to improve the performance and scale up to larger grammars, and handle the full expressivity of CG-3, with \textsc{map}, \textsc{add} and \textsc{substitute} rules.
-We want to also ask different questions, such as \todo{...}.
+We want to detect more kinds of conflicts.
+For instance, say we have the following rules:
+
+\begin{itemize}
+\item[] \texttt{REMOVE verb IF (-1C det) ;}
+\item[] \texttt{REMOVE noun IF (-1C det) ;}
+\end{itemize}
+
+With our symbolic sentence, this rule set will be no problem; after all, there can be something else in the target beside noun and a verb.
+However, if the grammar writer actually means the following
+
+\begin{itemize}
+\item[] \texttt{REMOVE verb IF (-1C det) (0 noun) ;}
+\item[] \texttt{REMOVE noun IF (-1C det) (0 verb) ;}
+\end{itemize}
+
+then those rules are contradictory, and we would like to find it out.
+
+\todo{Compare with a corpus-based test: do we find the same problems?}
+
 Finally, we want to test the approach to other grammar formalisms.

### 2015-10-23 paper/lrec/eval_futurework.tex
text

@@
 
-We ran for all three grammars the following thing.
-For each rule, run all the rules before it to the symbolic sentence, and then ask if that rule can fire.
+We tested all three grammars if they contain non-applicable rules:
+for each rule, apply all the preceding rules to the symbolic sentence, and then ask if that last rule can fire.
 This test revealed problems in all grammars.
-The Dutch grammar had the following rule sequence, making the latter impossible to apply:
+The Dutch grammar had the following rules in the given order, making the latter impossible to apply:
 
 \begin{itemize}
-\item[] \texttt{REMOVE adv IF (1 n|np) ;}
-\item[] \texttt{REMOVE adv IF (-1 det) (0 adj) (1 n|np) ;}
+\item[] 
+\begin{verbatim}REMOVE adv IF (1 n|np) ;
+REMOVE adv IF (-1 det) (0 adj) (1 n|np) ;
+\end{verbatim}
 \end{itemize}
@@
 The Spanish grammar had an erroneous set definition, where a word was required to have almost all POS tags at the same time. After fixing the set definition, our test didn't reveal other problems.
-The program marked __ problem rules for the Finnish grammar, but we didn't have time to verify if they were problematic or if our program is buggy.
+The program marked \todo{n} problem rules for the Finnish grammar, but we didn't have time to verify if they were genuinely problematic or if our program is at fault.
+We suspect that at least part of is related to the length of the sentences---for a condition that states $n$ or more words to the left or right, we only create a word exactly $n$ words away.
 
@@
 This can be seen from the time it took to compute the result for the last rule.
-For Finnish, the last rule took almost 4 minutes---had this been the average, the whole grammar would have taken 18 hours instead of 3.5.
+For Finnish, the last rule took almost 4 minutes---if this was the average speed, the whole grammar would have taken 18 hours instead of 3.5.
 We also measured the running times for the longest rules in the grammar.
@@
 
-For future work, we want to improve the performance and scale up to larger grammars, and handle the full expressivity of CG-3, with \textsc{map}, \textsc{add} and \textsc{substitute} rules.
-We want to detect more kinds of conflicts.
+For future work, we will continue improving the performance, and make sure that the implementation for the basic rules is correct; for instance, fixing the earlier mentioned issue with non-exact position.
+After this, we want to handle the full expressivity of CG-3, with \textsc{map}, \textsc{add} and \textsc{substitute} rules.
+
+Independently of 
+As another line of investigation, we want to detect more kinds of conflicts.
 For instance, say we have the following rules:
@@
 
-With our symbolic sentence, this rule set will be no problem; after all, there can be something else in the target beside noun and a verb.
-However, if the grammar writer actually means the following
+With our symbolic sentence, these rules will be no problem; after all, there can be something else in the target beside noun and a verb.
+However, it is likely that the grammar writer actually means the following:
 
@@
 
-then those rules are contradictory, and we would like to find it out.
+In that case, those rules are contradictory, and we would like to find it out.
 
-\todo{Compare with a corpus-based test: do we find the same problems?}
+Compare with a corpus-based test: do we find the same problems?
 

### 2015-10-23 paper/lrec/eval_futurework.tex
more future work

@@
 Rules in grammar      & 59              & 279               & 1185     \\ \hline
-Conflicts found       & 1               & ???               & ???    \\ \hline
+Conflicts found       & 1               & 1                 & ???    \\ \hline
 Time: all rules       & 7s              & 3min 22s          & 3h 26min    \\ \hline
@@
 
-The first row shows the whole number of rules. For the Finnish grammar, we took only \textsc{remove} and \textsc{select} rules. The other two only contained those rules to start with.
+The first row shows the whole number of rules. We included only \textsc{remove} and \textsc{select} rules.
 
@@
 This can be seen from the time it took to compute the result for the last rule.
-For Finnish, the last rule took almost 4 minutes---if this was the average speed, the whole grammar would have taken 18 hours instead of 3.5.
+For Finnish, the last rule took almost 4 minutes, but if we divide the time with the number of rules, the average time for one rule is just 10 seconds.
 We also measured the running times for the longest rules in the grammar.
@@
 
-For future work, we will continue improving the performance, and make sure that the implementation for the basic rules is correct; for instance, fixing the earlier mentioned issue with non-exact position.
+For future work, we will continue improving the performance, and make sure that the implementation for the basic rules is correct; for instance, fixing the earlier mentioned issue with a non-exact position.
 After this, we want to handle the full expressivity of CG-3, with \textsc{map}, \textsc{add} and \textsc{substitute} rules.
+At that stage, it would be interesting to test our program with actual grammar writers,
+in comparison with a corpus-based method or machine learning.
 
-Independently of 
 As another line of investigation, we want to detect more kinds of conflicts.
@@
 \begin{itemize}
-\item[] \texttt{REMOVE verb IF (-1C det) ;}
-\item[] \texttt{REMOVE noun IF (-1C det) ;}
+\item[] \begin{verbatim}
+REMOVE verb IF (-1C det) ;
+REMOVE noun IF (-1C det) ;
+\end{verbatim}
 \end{itemize}
 
-With our symbolic sentence, these rules will be no problem; after all, there can be something else in the target beside noun and a verb.
-However, it is likely that the grammar writer actually means the following:
+With our symbolic sentence, these rules will be no problem; to apply the latter, we only need to construct a target that has a realistic ambiguity with a noun; the verb is gone already.
+However, in the context of real life texts, the grammar writer might actually mean the following:
 
 \begin{itemize}
-\item[] \texttt{REMOVE verb IF (-1C det) (0 noun) ;}
-\item[] \texttt{REMOVE noun IF (-1C det) (0 verb) ;}
+\item[] \begin{verbatim}
+REMOVE verb IF (-1C det) (0 noun) ;
+REMOVE noun IF (-1C det) (0 verb) ;
+\end{verbatim}
 \end{itemize}
@@
 In that case, those rules are contradictory, and we would like to find it out.
-
-Compare with a corpus-based test: do we find the same problems?
 

### 2015-10-23 paper/lrec/problem_desc.tex
add more reasons why this is interesting

@@
 % \citep[CG][]{karlsson1995constraint} 
-is a method for disambiguating morphologically analysed text
+is a formalism used to disambiguate morphologically analysed text
 % and adding syntactic or dependency labels
@@
 %Authors who have a large hand-annotated corpus can also benefit from our tools.
-With our tools, the grammar writer or user can ask specific questions about their grammar, such as:
+With our tools, the users can diagnose and improve their grammars.
+We can ask specific questions about a given grammar, such as:
 
@@
 
-This technique requires an existing morphological lexicon, but no corpus. 
+Our technique requires an existing morphological lexicon, but no corpus. 
 The lexicon is needed in order to generate realistic ambiguities: e.g. a word can be ambiguous between noun and verb, but not between adjective and punctuation. 

### 2015-10-23 paper/lrec/eval_futurework.tex
eval

@@
 The smallest grammar was Dutch\footnote{\url{https://svn.code.sf.net/p/apertium/svn/languages/apertium-nld/apertium-nld.nld.rlx}}, with 59 rules; second was Spanish\footnote{\url{https://svn.code.sf.net/p/apertium/svn/languages/apertium-spa/apertium-spa.spa.rlx}}with 279 rules, and the largest was Finnish\footnote{\url{https://github.com/flammie/omorfi/blob/master/src/vislcg3/omorfi.cg3}}, with 1185 rules.
-For the smaller grammars, we were able to verify manually that the results are true.
+We left out rule types introduced in CG-3, and only tested the traditional \textsc{remove} and \textsc{select} rules.
 
@@
 
-The first row shows the whole number of rules. We included only \textsc{remove} and \textsc{select} rules.
-
 We tested all three grammars if they contain non-applicable rules:
 for each rule, apply all the preceding rules to the symbolic sentence, and then ask if that last rule can fire.
-This test revealed problems in all grammars.
+This test revealed problems in all grammars, and for the smaller grammars, we were able to verify manually that the results are true.
+
 The Dutch grammar had the following rules in the given order, making the latter impossible to apply:
@@
 We suspect that at least part of is related to the length of the sentences---for a condition that states $n$ or more words to the left or right, we only create a word exactly $n$ words away.
+
+The speed of the check is an obvious issue.
+%All rules is the total time to check every rule against the previous rules, and last rule shows how long it took to check the last rule.
+As indicated by the total time and the time to check the last rule, the size of the SAT problem grows according to how far in the sequence the rule is.
 

### 2015-10-23 paper/lrec/lrec2016_listenmaa_claessen.tex
more text
eval

@@
 \usepackage{graphicx}
-%\usepackage{float}
+\usepackage{slantsc}
 \usepackage[colorlinks=true,citecolor=blue,urlcolor=blue]{hyperref}
@@
 ]{\mainfontfamily}
-%\newfontfamily\scshape[Letters=SmallCaps, Numbers=Uppercase] % {Myriad Pro} % {TeX Gyre Pagella}
+%\newfontfamily\scshape[Letters=SmallCaps, Numbers=Uppercase] % {Myriad Pro} %
 
@@
 
-\title{Analysing Constraint Grammar with a SAT solver}
+\title{Analysing Constraint Grammars with a SAT solver}
 
@@
 \begin{abstract}
-We describe a method for analysing Constraint Grammar by encoding it
-as a SAT problem.
+We describe a method for
+ analysing Constraint Grammars by encoding the
+rules in SAT.
 Our tools can detect internal conflicts or redundancies in a grammar,
-as well as generate examples to demonstrate the effect of some rule or rule set.
+as well as generate examples to demonstrate the effect of some rule or
+rule set.
+This can help users to diagnose and improve their grammars.
 No corpus is required, only a morphological lexicon.

### 2015-10-23 paper/lrec/eval_futurework.tex
some sense found out of the finnish results

@@
 Time: last rule       & 4s              & 7s                & 3min 40s    \\ \hline
-Time: longest rule    & ?s              & ?s                & ?min     \\ \hline
-
+\# variables: last rule & 33915         & 247153            &    \\ \hline
+%Time: longest rule    & ?s              & ?s                & ?min     \\ \hline
 \end{tabular}
@@
 \item[] 
-\begin{verbatim}REMOVE adv IF (1 n|np) ;
-REMOVE adv IF (-1 det) (0 adj) (1 n|np) ;
+\begin{verbatim}REMOVE Adv IF (1 Noun) ;
+REMOVE Adv IF (-1 Det) (0 Adj) (1 Noun) ;
 \end{verbatim}
@@
 The Spanish grammar had an erroneous set definition, where a word was required to have almost all POS tags at the same time. After fixing the set definition, our test didn't reveal other problems.
-The program marked \todo{n} problem rules for the Finnish grammar, but we didn't have time to verify if they were genuinely problematic or if our program is at fault.
-We suspect that at least part of is related to the length of the sentences---for a condition that states $n$ or more words to the left or right, we only create a word exactly $n$ words away.
+The program marked \todo{n} problem rules for the Finnish grammar, but we didn't have time to verify if all of them were genuinely problematic or if our program is at fault.
+Correct errors include \todo{find a real problem and describe shortly}.
+Our program detected tagset errors also in the Finnish grammar.
+The grammar was originally written in 1995, and converted to the Omorfi tagset by \cite{pirinen2015}. 
+For example, the Omorfi tagset marks a possessive suffix with \texttt{Px}, e.g. \texttt{PxPl2} for second person plural, but the grammar used the old convention \texttt{PL2} in the rules.
+As a result, all rules with possessive suffixes were detected as impossible.
 
-The speed of the check is an obvious issue.
+Since we got our tagset from Omorfi, there was no ,
+We suspect that at least part of the result is related to the length of the sentences---for a condition that states $n$ or more words to the left or right, we only create a word exactly $n$ words away.
+This is something to be fixed in the future.
+
+The speed of the check is an obvious issue currently.
+For Finnish and Spanish, each symbolic word has around 2500 variables in the beginning.
+These variables represent tag combinations that come from the Apertium morphological lexicon. 
+The Spanish lexicon includes up to two clitics (e.g. \emph{dámelo}, give.me.it) in the same word form, but none for Dutch but we excluded clitics for the Finnish tag combinations.
+All morphological lexicons are from Apertium.
+
 %All rules is the total time to check every rule against the previous rules, and last rule shows how long it took to check the last rule.

### 2015-10-23 paper/lrec/eval_futurework.tex
amount of variables per symbolic word

@@
 As a result, all rules with possessive suffixes were detected as impossible.
-
-Since we got our tagset from Omorfi, there was no ,
 We suspect that at least part of the result is related to the length of the sentences---for a condition that states $n$ or more words to the left or right, we only create a word exactly $n$ words away.
@@
 The speed of the check is an obvious issue currently.
-For Finnish and Spanish, each symbolic word has around 2500 variables in the beginning.
+The Finnish symbolic word has around 5700 variables in the beginning of the analysis, each representing a possible tag combination.
+Around 3700 come from the Apertium morphological lexicon, excluding clitics, and around 2000 are present in the grammar
+For Spanish and Dutch, the numbers are around 2000 and 200 respectively.
+
 These variables represent tag combinations that come from the Apertium morphological lexicon. 

### 2015-10-23 paper/lrec/eval_futurework.tex
structure

@@
 
-We tested all three grammars if they contain non-applicable rules:
+\paragraph{Accuracy} We tested all three grammars if they contain non-applicable rules:
 for each rule, apply all the preceding rules to the symbolic sentence, and then ask if that last rule can fire.
@@
 As a result, all rules with possessive suffixes were detected as impossible.
-We suspect that at least part of the result is related to the length of the sentences---for a condition that states $n$ or more words to the left or right, we only create a word exactly $n$ words away.
+In addition, we suspect that some of the problems come just from our naive implementation.
+For a condition that states $n$ or more words to the left or right, we only create a word in the sentence exactly $n$ words away. This is something to be fixed in the future.
+
+ at least part of the result is related to the length of the sentences---for a condition that states $n$ or more words to the left or right, we only create a word exactly $n$ words away.
 This is something to be fixed in the future.
 
-The speed of the check is an obvious issue currently.
+\paragraph{Running time} The speed of the check is an obvious issue currently.
 The Finnish symbolic word has around 5700 variables in the beginning of the analysis, each representing a possible tag combination.
-Around 3700 come from the Apertium morphological lexicon, excluding clitics, and around 2000 are present in the grammar
-For Spanish and Dutch, the numbers are around 2000 and 200 respectively.
+Around 3700 come from the Apertium morphological lexicon, excluding clitics, and around 2000 are present in the grammar---many of them include an actual word, both in target (\texttt{REMOVE (``bear'' verb) IF ...}) and in conditions (\texttt{REMOVE ...
+IF (-1 (``bear'' noun))}.
 
-These variables represent tag combinations that come from the Apertium morphological lexicon. 
-The Spanish lexicon includes up to two clitics (e.g. \emph{dámelo}, give.me.it) in the same word form, but none for Dutch but we excluded clitics for the Finnish tag combinations.
-All morphological lexicons are from Apertium.
+For Spanish and Dutch, the numbers are around 2000 and 300 respectively.
+The Spanish lexicon includes up to two clitics (e.g. \emph{dámelo} `give.me.it') in the same word form, but Dutch has none in the lexicon.
 
@@
 
-For future work, we will continue improving the performance, and make sure that the implementation for the basic rules is correct; for instance, fixing the earlier mentioned issue with a non-exact position.
+\paragraph{Future work} We will continue improving the performance, and make sure that the implementation for the basic rules is correct; for instance, fixing the earlier mentioned issue with a non-exact position.
 After this, we want to handle the full expressivity of CG-3, with \textsc{map}, \textsc{add} and \textsc{substitute} rules.

### 2015-10-23 paper/lrec/eval_futurework.tex
number of tag combinations

@@
                       & \textsc{nld}  & \textsc{spa}  & \textsc{fin}  \\ \hline
-Rules in grammar      & 59              & 279               & 1185     \\ \hline
-Conflicts found       & 1               & 1                 & ???    \\ \hline
+\# rules in grammar      & 59              & 279      & 1185     \\ \hline
+\# tag combinations      & 276             & 2191     & 5719    \\ \hline
+\# conflicts found       & 1               & 1        & ???    \\ \hline
 Time: all rules       & 7s              & 3min 22s    & 3h 26min    \\ \hline
-Time: last rule       & 4s              & 7s                & 3min 40s    \\ \hline
+Time: last rule       & 4s              & 7s          & 30min 25s    \\ \hline
+
+% This is with just 2000ish variables for Finnish, above is with 5700
+%Time: last rule       & 4s              & 7s                & 3min 40s    \\ \hline
+%Time: longest rule    & ?s              & ?s                & ?min     \\ \hline
 \# variables: last rule & 33915         & 247153            &    \\ \hline
-%Time: longest rule    & ?s              & ?s                & ?min     \\ \hline
+
 \end{tabular}
-\caption{Results}
+\caption{Detecting rules that cannot be applied}
 \label{table:res}
@@
 
-\paragraph{Accuracy} We tested all three grammars if they contain non-applicable rules:
+\paragraph{Correctness} We tested all three grammars if they contain non-applicable rules:
 for each rule, apply all the preceding rules to the symbolic sentence, and then ask if that last rule can fire.
@@
 
+The Spanish grammar had an erroneous set definition, where a word was required to have almost all POS tags at the same time. After fixing the set definition, our test didn't reveal other problems.
 The Dutch grammar had the following rules in the given order, making the latter impossible to apply:
-
 \begin{itemize}
@@
 
-The Spanish grammar had an erroneous set definition, where a word was required to have almost all POS tags at the same time. After fixing the set definition, our test didn't reveal other problems.
-The program marked \todo{n} problem rules for the Finnish grammar, but we didn't have time to verify if all of them were genuinely problematic or if our program is at fault.
+The program marked \todo{n} problem rules for the Finnish grammar, but we didn't have time to investigate all of them.
 Correct errors include \todo{find a real problem and describe shortly}.
-Our program detected tagset errors also in the Finnish grammar.
+Our program detected also a number of tagset errors in the Finnish grammar.
 The grammar was originally written in 1995, and converted to the Omorfi tagset by \cite{pirinen2015}. 
@@
 In addition, we suspect that some of the problems come just from our naive implementation.
-For a condition that states $n$ or more words to the left or right, we only create a word in the sentence exactly $n$ words away. This is something to be fixed in the future.
+For a condition situated $n$ or more words to the left or right, we only create a word exactly $n$ words away. This is something to be fixed in the future.
 
- at least part of the result is related to the length of the sentences---for a condition that states $n$ or more words to the left or right, we only create a word exactly $n$ words away.
-This is something to be fixed in the future.
+\paragraph{Running time} The speed of the check is an obvious issue, especially for the Finnish grammar. The SAT problem grows the more rules
 
-\paragraph{Running time} The speed of the check is an obvious issue currently.
 The Finnish symbolic word has around 5700 variables in the beginning of the analysis, each representing a possible tag combination.
-Around 3700 come from the Apertium morphological lexicon, excluding clitics, and around 2000 are present in the grammar---many of them include an actual word, both in target (\texttt{REMOVE (``bear'' verb) IF ...}) and in conditions (\texttt{REMOVE ...
-IF (-1 (``bear'' noun))}.
-
+Around 3700 come from the Apertium morphological lexicon, excluding clitics, and around 2000 are present in the grammar---many of them include an actual word, both in target (\texttt{REMOVE ("bear" verb) IF ...}) and in conditions (\texttt{REMOVE ...
+IF (-1 ("bear" noun))}.
 For Spanish and Dutch, the numbers are around 2000 and 300 respectively.
 The Spanish lexicon includes up to two clitics (e.g. \emph{dámelo} `give.me.it') in the same word form, but Dutch has none in the lexicon.
+
+Not all questions are so expensive.
+For instance, checking the tag sets in the Finnish grammar doesn't require the context of the other rules at all,
+ The majority of the correctly detected errors in the Finnish grammar were rule-internal---these errors don't need the context of the other rules at all.
 

### 2015-10-23 paper/lrec/eval_futurework.tex
remove crap

@@
 
-\paragraph{Running time} The speed of the check is an obvious issue, especially for the Finnish grammar. The SAT problem grows the more rules
+\paragraph{Running time} The speed of the check is an obvious issue, especially for the Finnish grammar. 
+A Finnish symbolic word has around 5700 variables in the beginning of the analysis, each representing a possible tag combination.
+Around 3700 come from the Apertium morphological lexicon, excluding clitics, and around 2000 are defined in the grammar, as a target (\texttt{REMOVE ("bear" verb) IF ...}) or as a condition (\texttt{REMOVE ...
+IF (-1 ("bear" noun))}.
+The majority of them are lexical, whereas we don't include anything but morphological tags from the lexicon.
+For Spanish and Dutch, there are around 2000 and 300 tag combinations respectively.
+The Spanish lexicon includes up to two clitics (e.g. \emph{dámelo} `give.me.it') in the same word form, but Dutch has none in the lexicon.
 
-The Finnish symbolic word has around 5700 variables in the beginning of the analysis, each representing a possible tag combination.
-Around 3700 come from the Apertium morphological lexicon, excluding clitics, and around 2000 are present in the grammar---many of them include an actual word, both in target (\texttt{REMOVE ("bear" verb) IF ...}) and in conditions (\texttt{REMOVE ...
-IF (-1 ("bear" noun))}.
-For Spanish and Dutch, the numbers are around 2000 and 300 respectively.
-The Spanish lexicon includes up to two clitics (e.g. \emph{dámelo} `give.me.it') in the same word form, but Dutch has none in the lexicon.
+The SAT problem for checking one rule grows depending on how many rules come before it, how long the context is, and how many tag combinations are there.
+The running time of the last rule gives some indication ...
 
@@
 
-%All rules is the total time to check every rule against the previous rules, and last rule shows how long it took to check the last rule.
-As indicated by the total time and the time to check the last rule, the size of the SAT problem grows according to how far in the sequence the rule is.
+%All rules is the total time to check every rule against the previous rules, and last rule % shows how long it took to check the last rule.
+% As indicated by the total time and the time to check the last rule, the size of the SAT problem grows according to how far in the sequence the rule is.
 
-As seen in table~\ref{table:res}, checking every rule in the Finnish grammar takes currently 3.5 hours.
-A rule takes longer to run if it appears later in the rule sequence: 1000 previous rules is a larger SAT problem than 10 previous rules.
-This can be seen from the time it took to compute the result for the last rule.
+% As seen in table~\ref{table:res}, checking every rule in the Finnish grammar takes currently 3.5 hours.
+% A rule takes longer to run if it appears later in the rule sequence: 1000 previous rules is a larger SAT problem than 10 previous rules.
+% This can be seen from the time it took to compute the result for the last rule.
 For Finnish, the last rule took almost 4 minutes, but if we divide the time with the number of rules, the average time for one rule is just 10 seconds.
-We also measured the running times for the longest rules in the grammar.
+% We also measured the running times for the longest rules in the grammar.
 \todo{find longest rule, change place of that rule, see what matters more. Promise: we can be smarter about SAT things.}

### 2015-10-23 paper/lrec/implementation.tex
stuff
writing

@@
 
-In this section, we describe the implementation of the toolset.
-The SAT encoding we use is introduced in \cite{listenmaa_claessen2015}, along with a working CG compiler.
-Since we want to test the rules in a way that doesn't depend on data, we operate on a \emph{symbolic sentence}: a word sequence which we can modify according to the properties we want to test.
+In this section, we describe the implementation of the tool.
+The SAT-encoding we use is similar to the one introduced in \cite{listenmaa_claessen2015}, with one key difference: in this paper, we operate on {\em symbolic sentences} instead of concrete sentences from a corpus.
+The idea is that the SAT-solver is going to find the concrete sentence for us.
 
-\paragraph{SAT encoding of CG rules}
+\paragraph{Preliminaries}
 
-A single CG rule is a template which creates a concrete clause for each instance where it is applied.
-%Each potential analysis is represented as a variable, and the rules are translated into clauses that operate on those variables.
-Table~\ref{table:vars} shows the rule \texttt{REMOVE v IF (-1 det)} applied to the text \emph{the bear}.
-The analyses of the word forms are represented as variables, and
-the default rule ensures that each word has at least one analysis.
-%This makes \texttt{the\_det} true by default, therefore the implication must hold, and \texttt{bear\_v\_pl} will be removed.
-For more details about the encoding, see \cite{listenmaa_claessen2015}.
+Our analysis operates on one rule $R$ at a time, and is concerned with answering the following question: ``Does there exist an input sentence $S$ that can trigger rule $R$, even after passing all rules $R'$ that came before $R$?''
 
-% table:vars moved to main file for positioning
+Before we can do any analysis any of the rules, we need to find out what the set of all possible readings of a word is.
+We can do this by extracting this information from a lexicon, but there are other ways too.
+In our experiments, this number has ranged from about 300 to about 6000.
 
-\paragraph{Word-internal constraints}
+Furthermore, when we analyze a rule $R$, we need to decide the {\em width} of the rule $R$: How many different words should there be in a sentence that can trigger $R$?
+We denote this value by $w(R)$.
+Most often, $w(R)$ can be easily determined by looking at how far away the rule context indexes in the sentence relative to the target.
+For example, in the rule mentioned in the introduction, the width is 2.
 
-The symbolic sentence consists of symbolic words.
-In the beginning, a symbolic word is a list of all possible tag combinations, and we don't know which ones are true.
-%These tag combinations may be underspecified for lemma or other features---similarly, a CG rule may target any noun, a plural noun or the noun \emph{bear}.
-Given a large morphological lexicon, we can restrict the individual words to contain only realistic ambiguities. 
-For instance, a Spanish morphological lexicon contains the entries \texttt{casa:casar<verb><sg><p3>} and \texttt{casa:casa<noun><sg>}, hence we know that a confusion between a 3rd person singular verb and a singular noun is possible.
-%The lexicon is not likely to contain a wordform that can be analysed both as an adjective and punctuation, so that kind of ambiguous word is never created.
-
-%However, since our method is corpus-free, we don't add any external restrictions as to what words can follow each other.
+If the context contains a \verb!*!, we may need to make an approximation of $w(R)$ which may result in false negatives later on in the analysis.
 
@@
 
-Figure~\ref{fig:example} shows a fragment of of a two-word symbolic sentence, before and after applying the rule \texttt{REMOVE verb IF (-1 det)}\footnote{For simplicity, we don't show the word-internal constraints that come from the lexicon, but only the rule application.}.
-As with real analyses, each tag combination is given a variable, and applying a rule means creating a concrete clause for each instance where a determiner is found preceding a verb.
-In a symbolic sentence, every word has potentially every tag combination, and applying the rules will shape the sentence into something that respects the constraints of the language.
+The SAT-solver comes in when we start looking at the rules and what they do to sentences.
+We start each analysis by creating a so-called {\em symbolic sentence}, which is our representation of the sentence $S$ we are looking for.
+A symbolic sentence is a sequence of {\em symbolic words}, and a symbolic word is a table of all possible readings that a word can have, each reading paired up with a SAT-variable.
 
-\begin{figure}[h]
-\includegraphics[width=0.42\textwidth]{symbolic_sentence.png}
-\caption{Symbolic sentence beginning to take shape \todo{make this less confusing?}}
-\label{fig:example}
-\end{figure}
+The length of the symbolic sentence we create when we analyze a rule $R$ is $w(R)$. For the rule in the introduction, we have $w(R)=2$ and a symbolic sentence may look as follows:
+\begin{center}
+\begin{tabular}{c|c|c}
+word1 & word2 & reading \\
+\hline
+$v_1$ & $w_1$ & det def \\
+$v_2$ & $w_2$ & noun sg \\
+$v_3$ & $w_3$ & noun pl \\
+$v_4$ & $w_4$ & verb sg \\
+$v_5$ & $w_5$ & verb pl \\
+\end{tabular}
+\end{center}
+Here, $v_i$ and $w_j$ are SAT-variables belonging to word1 and word2, respectively. We can see that the possible number of readings here was 5.
 
-The example rule states that the sentence may not have a verb after a determiner.
-The variable \texttt{w2<verb>} is the situation of the second word's verb status before applying the rule.
-We create a new variable \texttt{w2'<verb>}, whose value depends on whether the condition holds (if \texttt{w1<det>} is true); whether it was true on the last round (if \texttt{w2<verb>} is already removed by another rule, \texttt{w2'<verb>} cannot reappear); and whether it is the only analysis left (not depicted in the picture).
-All the analyses not targeted by this rule remain the same.
+The SAT-solver may contain extra constraints about the variables. For example, we would like the input sentence to have at least one reading per word, so we add the following two constraints:
+\begin{center}
+\begin{tabular}{c}
+$v_1 \vee v_2 \vee v_3 \vee v_4 \vee v_5$, \\
+$w_1 \vee w_2 \vee w_3 \vee w_4 \vee w_5$ \\
+\end{tabular}
+\end{center}
+The point of the symbolic sentence is that any solution that is produced by the SAT-solver represents an input sentence that we are looking for.
 
-% \begin{figure}[]
-% \includegraphics[width=0.42\textwidth]{symbolic_sentence.png}
-% \caption{Symbolic sentence beginning to take shape \todo{make this less confusing?}}
-% \label{fig:example}
-% \end{figure}
+\paragraph{Applying a rule}
 
-%The applications of these rules shape our symbolic sentence into containing only things that can survive the rules.
+In order to say something about the rules in the grammar, we need to be able to apply a given rule $R'$ to the symbolic sentence, resulting in a new sentence.
+The way this works is that we investigate the rule, and generate a new sentence based on what changes can be made to the sentence.
 
-The effect of the old rule applications is retained when a new rule is added.
-This means that we can ask a question ``after applying rules $0-i$, is it possible for $i+1$ to apply?'', and the answer is reliable; we cannot find a counterexample by just having a bit more data, because we start from literally all possibilities, and each rule restricts the set of possible sentences.
-%If rule \texttt{r100} would require the first word to be a noun, and causes a conflict because it is not a noun, we know that it is because there is no way for w1 to be a noun.
+For example, if we apply the rule from the introduction to the symbolic sentence above, we generate the following resulting symbolic sentence:
+\begin{center}
+\begin{tabular}{c|c|c}
+word1 & word2 & reading \\
+\hline
+$v_1$ & $w_1$ & det def \\
+$v_2$ & $w_2$ & noun sg \\
+$v_3$ & $w_3$ & noun pl \\
+$v_4$ & $w_4'$ & verb sg \\
+$v_5$ & $w_5'$ & verb pl \\
+\end{tabular}
+\end{center}
+The rule can only affect readings of word2 that have a verb tag, so we create two new variables $w_4'$ and $w_5'$.
+We also have to say what their values are, which we do by adding constraints to the SAT-solver.
+The constraint we add for $w_4'$ is:
+\begin{center}
+\begin{tabular}{c}
+$w_4' \Leftrightarrow [ w_4 \wedge \neg{}(v_1 \wedge (w_1 \vee w_2 \vee w_3)) ]$ \\
+\end{tabular}
+\end{center}
+In other words $w_4'$ can only be true if $w_4$ was true already, and the rule has not triggered. The rule triggers if $v_1$ is true and one of word2's readings is not verb.
 
-%Note that this method is not suitable for language generation: rules in any realistic CG are tailored for real life ambiguities, whereas our symbolic sentence starts with literally all possibilities of tag combinations.
+EXPLAIN THIS BETTER
+
+\paragraph{Putting it all together}
+
+-- JUNK BELOW --
+
+We use a SAT-solver to construct such a sentence $S$, or to prove the non-existence of such a sentence.
+The way we do this is by first constructing a {\em symbolic sentence} $S_0$, a sentence of a particular length where all words in principle can have any possible reading, and where each reading for each word is represented by a SAT variable.
+$S_0$ represents all possible input sentences.
+
+An example
+
+We then proceed by calculating a symbolic representation of all sentences that could exist that have passed applying each rule $R'$ that preceeds the rule $R$ that is under analysis to the symbolic sentence, in the same sequence as they appear in the grammar.
+After each such application, we get a new symbolic sentence that represents all possible sentences that could be the result after applying those rules.
+
+%%% FROM INARI %%%
 

### 2015-10-23 paper/lrec/lrec2016_listenmaa_claessen.tex
remove table

@@
 
-\begin{table*}[]
-\centering
-\begin{tabular}{|l|l|l|l|}
-\hline
-
-\textbf{Analyses}
-& \textbf{Variables}
-& \textbf{Default rule: word has $\geq$1 analysis}                                    & \textbf{Constraint rules}                                                          \\ \hline
-\multicolumn{1}{|l|}{\begin{tabular}[c]{@{}l@{}}``\textless{}the\textgreater''\\$\qquad\quad$"the" det\end{tabular}}     & \multicolumn{1}{l|}{\texttt{the\_det}}                                      & \multicolumn{1}{l|}{\texttt{the\_det}}                    & \multicolumn{1}{l|}{\multirow{2}{*}{\texttt{the\_det $\Rightarrow$ $\neg$bear\_v\_pl}}} \\ \cline{1-3}
-\begin{tabular}[c]{@{}l@{}}"\textless{}bear\textgreater"\\$\qquad\quad$"bear" n sg\\$\qquad\quad$"bear" v pl\end{tabular} & \texttt{\begin{tabular}[c]{@{}l@{}}bear\_n\_sg \\ bear\_v\_pl\end{tabular}} & \multicolumn{1}{l|}{\texttt{bear\_n\_sg $\vee$ bear\_v\_pl}} & \multicolumn{1}{l|}{}                                                              \\ \hline
-\end{tabular}
-\caption{\texttt{REMOVE v IF (-1 det)} applied to \emph{the bear}}
-\label{table:vars}
-\end{table*}
-
 % Introduction & problem description

### 2015-10-24 paper/lrec/problem_desc.tex
eval
add examples of conflicts

@@
 With our tools, the users can diagnose and improve their grammars.
-We can ask specific questions about a given grammar, such as:
+They can ask specific questions about a given grammar, such as:
 
@@
 \item Are there rules that will never fire?
-\item Generate a sequence that triggers rule(s) X but not rule(s) Y
+\item Generate a sequence that triggers rule(s) R1 but not rule(s) R2
 %\item Generate a sequence that is ambiguous but doesn't trigger any rules
 \end{itemize}
+
+Some examples where a rule contradicts another are listed below. R1
+can also be a set of rules: for instance, one rule removes a verb if
+preceded by a determiner, and another if not preceded by a determiner,
+and together these rules remove a verb in all possible cases.
+\begin{enumerate}
+\item If a rule appears twice, the second occurrence will be disabled by the first
+\item R1 selects something in a context, R2 removes it
+\item R1 removes something from R2's context, so R2 can never apply
+\end{enumerate}
 

### 2015-10-24 paper/lrec/implementation.tex
first draft

@@
 
-Our analysis operates on one rule $R$ at a time, and is concerned with answering the following question: ``Does there exist an input sentence $S$ that can trigger rule $R$, even after passing all rules $R'$ that came before $R$?''
+Our analysis operates on one rule $R$, and is concerned with answering the following question: ``Does there exist an input sentence $S$ that can trigger rule $R$, even after passing all rules $R'$ that came before $R$?''
 
@@
 
-In order to say something about the rules in the grammar, we need to be able to apply a given rule $R'$ to the symbolic sentence, resulting in a new sentence.
+In order to say something about the rules in the grammar, we need to be able to apply a given rule $R'$ to a symbolic sentence, resulting in a new symbolic sentence.
 The way this works is that we investigate the rule, and generate a new sentence based on what changes can be made to the sentence.
@@
 \end{center}
-In other words $w_4'$ can only be true if $w_4$ was true already, and the rule has not triggered. The rule triggers if $v_1$ is true and one of word2's readings is not verb.
-
-EXPLAIN THIS BETTER
+In other words, after applying the rule, the reading ``verb sg'' (represented by the variable $w_4'$) can only be in the resulting sentence exactly when (1) ``verb sg'' was a reading of the input sentence (so $w_4$ is true) and (2) the rule has not been triggered (the rule triggers when $v_1$ is true and at least one of the non-verb readings $w_1 \dots w_3$ is true).
+We add a similar constraint for the new variable $w_5'$:
+\begin{center}
+\begin{tabular}{c}
+$w_5' \Leftrightarrow [ w_5 \wedge \neg{}(v_1 \wedge (w_1 \vee w_2 \vee w_3)) ]$ \\
+\end{tabular}
+\end{center}
 
@@
 
--- JUNK BELOW --
+Once we know how to apply one rule $R'$ to a symbolic sentence, we can also apply the sequence of all rules $R$ to a symbolic sentence.
+In this way we end up with a symbolic sentence that represents all sentences that could have been the result after applying all rules $R'$ in the sequence.
+We simply apply each rule to the result of applying the previous rule.
 
-We use a SAT-solver to construct such a sentence $S$, or to prove the non-existence of such a sentence.
-The way we do this is by first constructing a {\em symbolic sentence} $S_0$, a sentence of a particular length where all words in principle can have any possible reading, and where each reading for each word is represented by a SAT variable.
-$S_0$ represents all possible input sentences.
+Finally, we can take a look at the rule $R$ we want to analyze. Here is an example:
+\begin{itemize}
+\item[] \texttt{REMOVE det IF (1 verb) ;}
+\end{itemize}
+If we take the symbolic sentence above as input, we want to ask wether or not it can trigger the rule. We do this by adding the following constraints to the SAT-solver:
+\begin{center}
+\begin{tabular}{c}
+$w_4' \vee w_5'$, \\
+$v_1 \wedge (v_2 \vee v_3 \vee v_4 \vee v_5)$
+\end{tabular}
+\end{center}
+The first constraint says that the context of the rule should apply (the second word should have a reading with a ``verb'' tag).
+The second constraint says that we can remove the tag (the first word should have a reading with a ``det'' tag and there should be at least one other reading).
 
-An example
+If the SAT-solver can find a solution to all constraints, we have found a concrete sentence that satisfies our goal.
+If the SAT-solver cannot find a solution, it means that there are no sentences that can ever trigger rule $R$.
+This means that there is something wrong with the grammar.
 
-We then proceed by calculating a symbolic representation of all sentences that could exist that have passed applying each rule $R'$ that preceeds the rule $R$ that is under analysis to the symbolic sentence, in the same sequence as they appear in the grammar.
-After each such application, we get a new symbolic sentence that represents all possible sentences that could be the result after applying those rules.
+\paragraph{Other questions we can ask}
 
-%%% FROM INARI %%%
-
-\paragraph{What kind of questions we can ask?}
-
-The example so far has been of type ``Can rule Z apply after A-Y?''
-In case a conflict is detected, we can dig into it and refine where the problem is: e.g. is the condition not met, or is it the case that the target is already removed or the only remaining analysis.
-We can find what other rule or rules prevent others from applying, or if the conflict is rule-internal, such as nonexisting tag set or contradicting requirements of a context word (e.g.
+Apart from finding out what rules prevent others from applying, we can also find out if there is a conflict that is rule-internal, such as nonexisting tag set or contradicting requirements of a context word (e.g.
 a word must be unambiguously two different POS).

### 2015-10-24 paper/lrec/implementation.tex
filat lite

@@
 We can do this by extracting this information from a lexicon, but there are other ways too.
-In our experiments, this number has ranged from about 300 to about 6000.
+In our experiments, the number of readings has ranged from about 300 to about 6000.
 
-Furthermore, when we analyze a rule $R$, we need to decide the {\em width} of the rule $R$: How many different words should there be in a sentence that can trigger $R$?
-We denote this value by $w(R)$.
+Furthermore, when we analyze a rule $R$, we need to decide the {\em width} $w(R)$ of the rule $R$: How many different words should there be in a sentence that can trigger $R$?
 Most often, $w(R)$ can be easily determined by looking at how far away the rule context indexes in the sentence relative to the target.
@@
 
-The SAT-solver comes in when we start looking at the rules and what they do to sentences.
 We start each analysis by creating a so-called {\em symbolic sentence}, which is our representation of the sentence $S$ we are looking for.
-A symbolic sentence is a sequence of {\em symbolic words}, and a symbolic word is a table of all possible readings that a word can have, each reading paired up with a SAT-variable.
+A symbolic sentence is a sequence of {\em symbolic words}; a symbolic word is a table of all possible readings that a word can have, where each reading is paired up with a SAT-variable.
 
-The length of the symbolic sentence we create when we analyze a rule $R$ is $w(R)$. For the rule in the introduction, we have $w(R)=2$ and a symbolic sentence may look as follows:
+The number of words in the symbolic sentence we create when we analyze a rule $R$ is $w(R)$. For the rule in the introduction, we have $w(R)=2$ and a symbolic sentence may look as follows:
 \begin{center}
@@
 \end{center}
-Here, $v_i$ and $w_j$ are SAT-variables belonging to word1 and word2, respectively. We can see that the possible number of readings here was 5.
+Here, $v_i$ and $w_j$ are SAT-variables belonging to word1 and word2, respectively. We can also see that the possible number of readings here was 5.
 
-The SAT-solver may contain extra constraints about the variables. For example, we would like the input sentence to have at least one reading per word, so we add the following two constraints:
+The SAT-solver contains extra constraints about the variables. Input sentences should have at least one reading per word, so we add the following two constraints:
 \begin{center}
@@
 \end{center}
-The point of the symbolic sentence is that any solution that is produced by the SAT-solver represents an input sentence that we are looking for.
+Any solution to the constraints found by the SAT-solver can be interpreted as a concrete sentence with $w(R)$ words that each have a set of readings.
 
@@
 
-In order to say something about the rules in the grammar, we need to be able to apply a given rule $R'$ to a symbolic sentence, resulting in a new symbolic sentence.
-The way this works is that we investigate the rule, and generate a new sentence based on what changes can be made to the sentence.
+Next, we need to be able to apply a given rule $R'$ to a symbolic sentence, resulting in a new symbolic sentence.
 
-For example, if we apply the rule from the introduction to the symbolic sentence above, we generate the following resulting symbolic sentence:
+For example, if we apply the rule from the introduction to the symbolic sentence above, the result is the following symbolic sentence:
 \begin{center}
@@
 \end{center}
-The rule can only affect readings of word2 that have a verb tag, so we create two new variables $w_4'$ and $w_5'$.
-We also have to say what their values are, which we do by adding constraints to the SAT-solver.
-The constraint we add for $w_4'$ is:
+The example rule can only affect readings of word2 that have a ``verb'' tag, so we create only two new variables $w_4'$ and $w_5'$ for the result, and reuse the other variables.
+We add the following constraint for $w_4'$ is:
 \begin{center}
@@
 
-Once we know how to apply one rule $R'$ to a symbolic sentence, we can also apply the sequence of all rules $R$ to a symbolic sentence.
-In this way we end up with a symbolic sentence that represents all sentences that could have been the result after applying all rules $R'$ in the sequence.
+Once we know how to apply one rule $R'$ to a symbolic sentence, we can apply all rules preceding the rule $R$ that is under analysis.
 We simply apply each rule to the result of applying the previous rule.
+In this way, we end up with a symbolic sentence that represents all sentences that could be the result of applying all those rules.
 
@@
 \end{itemize}
-If we take the symbolic sentence above as input, we want to ask wether or not it can trigger the rule. We do this by adding the following constraints to the SAT-solver:
+If we take the symbolic sentence above as input, we want to ask wether or not it can trigger the rule $R$. We do this by adding some more constraints to the SAT-solver.
+
+First, the context of the rule should be applicable, meaning that the second word should have a reading with a ``verb'' tag:
 \begin{center}
 \begin{tabular}{c}
-$w_4' \vee w_5'$, \\
+$w_4' \vee w_5'$
+\end{tabular}
+\end{center}
+Second, the rule should be able to remove the ``det'' tag, meaning that the first word should have a reading with a ``det'' tag, and there should be at least one other reading:
+\begin{center}
+\begin{tabular}{c}
 $v_1 \wedge (v_2 \vee v_3 \vee v_4 \vee v_5)$
@@
 \end{center}
-The first constraint says that the context of the rule should apply (the second word should have a reading with a ``verb'' tag).
-The second constraint says that we can remove the tag (the first word should have a reading with a ``det'' tag and there should be at least one other reading).
-
-If the SAT-solver can find a solution to all constraints, we have found a concrete sentence that satisfies our goal.
+If the SAT-solver can find a solution to all constraints generated so far, we have found a concrete sentence that satisfies our goal.
 If the SAT-solver cannot find a solution, it means that there are no sentences that can ever trigger rule $R$.

### 2015-10-24 paper/lrec/eval_futurework.tex
making word choices, orthography, variable naming etc. more homogenous

@@
 
-We tested three grammars to find rules that cannot apply.
-The smallest grammar was Dutch\footnote{\url{https://svn.code.sf.net/p/apertium/svn/languages/apertium-nld/apertium-nld.nld.rlx}}, with 59 rules; second was Spanish\footnote{\url{https://svn.code.sf.net/p/apertium/svn/languages/apertium-spa/apertium-spa.spa.rlx}}with 279 rules, and the largest was Finnish\footnote{\url{https://github.com/flammie/omorfi/blob/master/src/vislcg3/omorfi.cg3}}, with 1185 rules.
-We left out rule types introduced in CG-3, and only tested the traditional \textsc{remove} and \textsc{select} rules.
+We tested three grammars to find rules that cannot apply. The smallest
+grammar was
+Dutch\footnote{\url{https://svn.code.sf.net/p/apertium/svn/languages/apertium-nld/apertium-nld.nld.rlx}},
+with 59 rules; second was
+Spanish\footnote{\url{https://svn.code.sf.net/p/apertium/svn/languages/apertium-spa/apertium-spa.spa.rlx}}with
+279 rules, and the largest was
+Finnish\footnote{\url{https://github.com/flammie/omorfi/blob/master/src/vislcg3/omorfi.cg3}},
+with 1185 rules. We left out \textsc{add}, \textsc{map} and other rule types introduced in CG-3, and only tested the traditional \textsc{remove} and \textsc{select} rules.
 
@@
                       & \textsc{nld}  & \textsc{spa}  & \textsc{fin}  \\ \hline
-\# rules in grammar      & 59              & 279      & 1185     \\ \hline
-\# tag combinations      & 276             & 2191     & 5719    \\ \hline
-\# conflicts found       & 1               & 1        & ???    \\ \hline
-Time: all rules       & 7s              & 3min 22s    & 3h 26min    \\ \hline
-Time: last rule       & 4s              & 7s          & 30min 25s    \\ \hline
+\# rules      & 59              & 279      & 1185     \\ \hline
+\# readings      & 276             & 2191     & 5719    \\ \hline
+\# conflicts       & 1               & 1        & 22    \\ \hline
+\clock{} all rules       & 7s              & 3min 22s    & 3h 26min    \\ \hline
+\clock{} last rule       & 4s              & 7s                & 3min 40s    \\ \hline
+% This is with just 2000ish variables for Finnish, below is with 5700
 
-% This is with just 2000ish variables for Finnish, above is with 5700
-%Time: last rule       & 4s              & 7s                & 3min 40s    \\ \hline
-%Time: longest rule    & ?s              & ?s                & ?min     \\ \hline
-\# variables: last rule & 33915         & 247153            &    \\ \hline
+% \clock{} last rule       & 4s              & 7s          & 30min 25s    \\ \hline
+% \clock{} longest rule    & ?s              & ?s                & ?min     \\ \hline
+%\# variables: last rule & 33915         & 247153            &    \\ \hline
 
@@
 
-\paragraph{Correctness} We tested all three grammars if they contain non-applicable rules:
-for each rule, apply all the preceding rules to the symbolic sentence, and then ask if that last rule can fire.
+\paragraph{Correctness} We tested all three grammars if they contain
+rules that cannot be applied, as described in the previous section.
 This test revealed problems in all grammars, and for the smaller grammars, we were able to verify manually that the results are true.
 
-The Spanish grammar had an erroneous set definition, where a word was required to have almost all POS tags at the same time. After fixing the set definition, our test didn't reveal other problems.
+The Spanish grammar had an erroneous set definition, where a word was
+required to have almost all POS tags at the same time. After fixing
+the set definition, our test did not reveal other problems.
 The Dutch grammar had the following rules in the given order, making the latter impossible to apply:
@@
 
-The program marked \todo{n} problem rules for the Finnish grammar, but we didn't have time to investigate all of them.
-Correct errors include \todo{find a real problem and describe shortly}.
-Our program detected also a number of tagset errors in the Finnish grammar.
-The grammar was originally written in 1995, and converted to the Omorfi tagset by \cite{pirinen2015}. 
+The program found 22 inapplicable rules in the Finnish grammar, but we
+didn't have time to investigate all of them.
+We could verify a number of tagset errors---the grammar was originally written in 1995, and converted to the Omorfi tagset by \cite{pirinen2015}. 
 For example, the Omorfi tagset marks a possessive suffix with \texttt{Px}, e.g. \texttt{PxPl2} for second person plural, but the grammar used the old convention \texttt{PL2} in the rules.
 As a result, all rules with possessive suffixes were detected as impossible.
-In addition, we suspect that some of the problems come just from our naive implementation.
-For a condition situated $n$ or more words to the left or right, we only create a word exactly $n$ words away. This is something to be fixed in the future.
+In addition, we suspect that some of the problems come from our naive
+implementation of rules with * in its context; creating the same
+symbolic sentence for condition \texttt{1*} and \texttt{1}.
+%For a condition situated $n$ or more words to the left or right, we only create a word exactly $n$ words away. This is something to be fixed in the future.
 
-\paragraph{Running time} The speed of the check is an obvious issue, especially for the Finnish grammar. 
-A Finnish symbolic word has around 5700 variables in the beginning of the analysis, each representing a possible tag combination.
-Around 3700 come from the Apertium morphological lexicon, excluding clitics, and around 2000 are defined in the grammar, as a target (\texttt{REMOVE ("bear" verb) IF ...}) or as a condition (\texttt{REMOVE ...
+\paragraph{Running time} Currently the speed of the check is an obvious issue,
+especially for the Finnish grammar. 
+The 3 hours running time is achieved with using 2100 readings for the
+symbolic words; instead of the lexicon, we only used the readings
+specified in the grammar: for example, the rule \texttt{REMOVE ("bear"
+  noun sg) IF (-1 ())
+Using all 6000 readings from the lexicon, the program
+ran the analysis for the last rule in 30 minutes, making the running
+time for 1000+ rules potentially days.
+
+% A Finnish symbolic word has around 5700 variables in the beginning of the analysis, each representing a possible tag combination.
+% Around 3700 come from the Apertium morphological lexicon, excluding clitics, and around 2000 are defined in the grammar, as a target (\texttt{REMOVE ("bear" verb) IF ...}) or as a condition (\texttt{REMOVE ...
 IF (-1 ("bear" noun))}.
 The majority of them are lexical, whereas we don't include anything but morphological tags from the lexicon.
-For Spanish and Dutch, there are around 2000 and 300 tag combinations respectively.
-The Spanish lexicon includes up to two clitics (e.g. \emph{dámelo} `give.me.it') in the same word form, but Dutch has none in the lexicon.
+% For Spanish and Dutch, there are around 2000 and 300 tag combinations respectively.
+% The Spanish lexicon includes up to two clitics (e.g. \emph{dámelo} `give.me.it') in the same word form, but Dutch has none in the lexicon.
 
@@
 % We also measured the running times for the longest rules in the grammar.
-\todo{find longest rule, change place of that rule, see what matters more. Promise: we can be smarter about SAT things.}
+%\todo{find longest rule, change place of that rule, see what matters more. Promise: we can be smarter about SAT things.}
 
-\paragraph{Future work} We will continue improving the performance, and make sure that the implementation for the basic rules is correct; for instance, fixing the earlier mentioned issue with a non-exact position.
+\paragraph{Future work} We will continue improving the performance,
+and make sure that the implementation for the basic rules is correct;
+for instance, fixing the earlier mentioned issue with a non-exact
+position.
+We will also improve usability
 After this, we want to handle the full expressivity of CG-3, with \textsc{map}, \textsc{add} and \textsc{substitute} rules.

### 2015-10-24 paper/lrec/implementation.tex
making word choices, orthography, variable naming etc. more homogenous

@@
 
-Furthermore, when we analyze a rule $R$, we need to decide the {\em width} $w(R)$ of the rule $R$: How many different words should there be in a sentence that can trigger $R$?
+Furthermore, when we analyse a rule $R$, we need to decide the {\em width} $w(R)$ of the rule $R$: How many different words should there be in a sentence that can trigger $R$?
 Most often, $w(R)$ can be easily determined by looking at how far away the rule context indexes in the sentence relative to the target.
@@
 
-The number of words in the symbolic sentence we create when we analyze a rule $R$ is $w(R)$. For the rule in the introduction, we have $w(R)=2$ and a symbolic sentence may look as follows:
+The number of words in the symbolic sentence we create when we analyse a rule $R$ is $w(R)$. For the rule in the introduction, we have $w(R)=2$ and a symbolic sentence may look as follows:
 \begin{center}
@@
 
-Finally, we can take a look at the rule $R$ we want to analyze. Here is an example:
+Finally, we can take a look at the rule $R$ we want to analyse. Here is an example:
 \begin{itemize}
@@
 
-\paragraph{Other questions we can ask}
+% \paragraph{Other questions we can ask}
 
-Apart from finding out what rules prevent others from applying, we can also find out if there is a conflict that is rule-internal, such as nonexisting tag set or contradicting requirements of a context word (e.g.
+% Apart from finding out what rules prevent others from applying, we can also find out if there is a conflict that is rule-internal, such as nonexisting tag set or contradicting requirements of a context word (e.g.
 a word must be unambiguously two different POS).
 
-In addition to analysing a whole grammar, 
-we can construct all kinds of symbolic sentences to test out individual rules. 
-We can set the length, restrict individual words (e.g. ``3rd word must be a noun''), require that it triggers some rule but not other. 
-This functionality can aid the grammar writer in the process, to see if they have missed a case or defined the tagset correctly.
+% In addition to analysing a whole grammar, 
+% we can construct all kinds of symbolic sentences to test out individual rules. 
+% We can set the length, restrict individual words (e.g. ``3rd word must be a noun''), require that it triggers some rule but not other. 
+% This functionality can aid the grammar writer in the process, to see if they have missed a case or defined the tagset correctly.

### 2015-10-24 paper/lrec/problem_desc.tex
changed the example in the intro
making word choices, orthography, variable naming etc. more homogenous

--- 2015-10-23 paper/lrec/problem_desc.tex
@@
 % \citep[CG][]{karlsson1995constraint} 
-is a formalism used to disambiguate morphologically analysed text
-% and adding syntactic or dependency labels
-. 
+is a formalism used to disambiguate morphologically analysed text. 
 A grammar consists of rules that target specific analyses for selection or removal, based on contextual tests. For example, the following rule
 \begin{itemize}
-\item[] \texttt{REMOVE verb IF (-1C det) ;}
+\item[] \texttt{REMOVE verb IF (-1 det) ;}
 \end{itemize}
-removes a verb reading from a word which is preceded by an unambiguously (C) tagged determiner.
+removes a verb reading from a word which is preceded by adeterminer.
+
 Given the following text,
@@
 \item Are there rules that will never fire?
-\item Generate a sequence that triggers rule(s) X but not rule(s) Y
+\item Generate a sequence that triggers rule(s) $R$ but not rule(s) $R'$
 %\item Generate a sequence that is ambiguous but doesn't trigger any rules
+\end{itemize}
+
+Some examples where a rule contradicts another are listed below. R1
+can also be a set of rules: for instance, if one rule removes a verb in
+context $C$, and another in context $\neg C$, together these rules
+remove a verb in all possible cases, disabling any future rule that
+targets verbs.
+\begin{itemize}
+\item If a rule appears twice, the second occurrence will be disabled by the first
+\item $R1$ selects something in a context, $R2$ removes it
+\item $R1$ removes something from $R2$'s context, so $R2$ can never apply
 \end{itemize}
@@
 Constraints for the structure of the sequences come from the CG rules themselves.
-We encode both of these constraints as a satisfiability problem, and use a SAT solver to generate answers to the previous questions.
+We encode both of these constraints as a satisfiability problem, and use a SAT-solver to generate answers to the previous questions.
 
@@
 \cite{lager98} and \cite{lager_nivre01} reconstruct the CG formalism in first-order predicate logic.
-%, and \cite{listenmaa_claessen2015} implement a CG compiler using a SAT solver.
+%, and \cite{listenmaa_claessen2015} implement a CG compiler using a SAT-solver.
 Grammar analysis is a natural use case, due to some key features of the logical reconstruction.

### 2015-10-24 paper/lrec/problem_desc.tex
remove text

@@
 \end{itemize}
-removes a verb reading from a word which is preceded by adeterminer.
-
+removes a verb reading from a word which is preceded by a determiner.
 Given the following text,
@@
 
-As CGs grow larger, it gets harder for the grammar writers and users to keep track of all the rules and their interaction.
-%\cite{voutilainen2004} gives a detailed account about best practices of grammar writing and efficient use of hand-annotated corpora to aid the grammar development.
-%\cite{bick2013tuning} presents a method to optimise hand-written CGs using machine learning techniques.
-%Our tools complement the existing methods in CG analysis and optimisation.
-%While Bick's experiments show improvements in results, the grammar writer is none the wiser why is the grammar better.
-%Authors who have a large hand-annotated corpus can also benefit from our tools.
+As CGs grow larger, it gets harder for the grammar writers and users
+to keep track of all the rules and their interaction.
 With our tools, the users can diagnose and improve their grammars.
-They can ask specific questions about a given grammar, such as:
-
+Some examples of conflicts are listed below:
 \begin{itemize}
-\item Are there rules that contradict each other?
-\item Are there rules that will never fire?
-\item Generate a sequence that triggers rule(s) $R$ but not rule(s) $R'$
-%\item Generate a sequence that is ambiguous but doesn't trigger any rules
+\item If a rule appears twice, the second occurrence will be disabled by the first
+\item $R$ selects something in a context, $R'$ removes it
+\item $R$ removes something from the context of $R'$, so $R'$ can never
+  apply
+\item $R$ has an internal conflict, such as non-existent
+tag combination, or contradicting requirements for a context word
 \end{itemize}
-
-Some examples where a rule contradicts another are listed below. R1
-can also be a set of rules: for instance, if one rule removes a verb in
+$R$ can also be a set of rules: for instance, if one rule removes a verb in
 context $C$, and another in context $\neg C$, together these rules
@@
 targets verbs.
-\begin{itemize}
-\item If a rule appears twice, the second occurrence will be disabled by the first
-\item $R1$ selects something in a context, $R2$ removes it
-\item $R1$ removes something from $R2$'s context, so $R2$ can never apply
-\end{itemize}
 
-Our technique requires an existing morphological lexicon, but no corpus. 
-The lexicon is needed in order to generate realistic ambiguities: e.g. a word can be ambiguous between noun and verb, but not between adjective and punctuation. 
-Constraints for the structure of the sequences come from the CG rules themselves.
-We encode both of these constraints as a satisfiability problem, and use a SAT-solver to generate answers to the previous questions.
+%In addition to analysing the whole grammar,
+Our method can be used to
+test out individual rules. For instance, we can ask the tool to
+generate example sequences that trigger rule(s) $R$ but not rule(s) $R'$. 
+This functionality can help the grammar writer to see
+if they have implemented the rules in the intended way.
+
+% They can ask specific questions about a given grammar, such as:
+% \begin{itemize}
+% \item Are there rules that contradict each other?
+% \item Are there rules that will never fire?
+% \item Generate a sequence that triggers rule(s) $R$ but not rule(s) $R'$
+% %\item Generate a sequence that is ambiguous but doesn't trigger any rules
+% \end{itemize}
+% Our technique requires an existing morphological lexicon, compatible
+% with the tag set used in the grammar, but no corpus. 
+% The lexicon is needed in order to limit the possible tag combinations.
+% Constraints for the structure of the sequences come from the CG rules themselves.
+% We encode both of these constraints as a satisfiability problem, and use a SAT-solver to generate answers to the previous questions.
 

### 2015-10-25 paper/lrec/eval_futurework.tex
more cleanup

@@
 
-We tested three grammars to find rules that cannot apply. The smallest
-grammar was
+We tested three grammars to find rules that cannot apply. 
+The smallest grammar was
 Dutch\footnote{\url{https://svn.code.sf.net/p/apertium/svn/languages/apertium-nld/apertium-nld.nld.rlx}},
 with 59 rules; second was
-Spanish\footnote{\url{https://svn.code.sf.net/p/apertium/svn/languages/apertium-spa/apertium-spa.spa.rlx}}with
-279 rules, and the largest was
+Spanish\footnote{\url{https://svn.code.sf.net/p/apertium/svn/languages/apertium-spa/apertium-spa.spa.rlx}},
+with 279 rules, and the largest was
 Finnish\footnote{\url{https://github.com/flammie/omorfi/blob/master/src/vislcg3/omorfi.cg3}},
-with 1185 rules. We left out \textsc{add}, \textsc{map} and other rule types introduced in CG-3, and only tested the traditional \textsc{remove} and \textsc{select} rules.
+with 1185 rules. We left out \textsc{add}, \textsc{map} and other rule
+types introduced in CG-3, and only tested the traditional
+\textsc{remove} and \textsc{select} rules.
+The results are shown in table~\ref{table:res}.
 
@@
 \# rules      & 59              & 279      & 1185     \\ \hline
-\# readings      & 276             & 2191     & 5719    \\ \hline
-\# conflicts       & 1               & 1        & 22    \\ \hline
+\# readings      & 276             & 2191     & 2012    \\ \hline
+\# conflicts       & 1               & 1        & \todo{n}    \\ \hline
 \clock{} all rules       & 7s              & 3min 22s    & 3h 26min    \\ \hline
@@
 
-\paragraph{Correctness} We tested all three grammars if they contain
-rules that cannot be applied, as described in the previous section.
-This test revealed problems in all grammars, and for the smaller grammars, we were able to verify manually that the results are true.
+\paragraph{Correctness} 
+The test revealed problems in all grammars, and for the smaller
+grammars, we were able to verify manually that the detected rules were
+true negatives. However, for the Finnish grammar, we did not have time
+to investigate all of them; in addition, we suspect that some of
+them are false negatives.
+We did not check for false positives in any of the grammars.
 
@@
 
-The program found 22 inapplicable rules in the Finnish grammar, but we
-didn't have time to investigate all of them.
-We could verify a number of tagset errors---the grammar was originally written in 1995, and converted to the Omorfi tagset by \cite{pirinen2015}. 
-For example, the Omorfi tagset marks a possessive suffix with \texttt{Px}, e.g. \texttt{PxPl2} for second person plural, but the grammar used the old convention \texttt{PL2} in the rules.
-As a result, all rules with possessive suffixes were detected as impossible.
-In addition, we suspect that some of the problems come from our naive
-implementation of rules with * in its context; creating the same
-symbolic sentence for condition \texttt{1*} and \texttt{1}.
+For the Finnish grammar, we could verify a number of tagset errors. 
+The grammar was originally written in 1995, and converted to the
+Omorfi tagset by \cite{pirinen2015}. For example, the Omorfi tagset
+marks a possessive suffix with \texttt{Px}, e.g. \texttt{PxPl2} for second person plural, but the
+grammar used the old convention \texttt{PL2} in the rules.
+As a result, all rules with possessive suffixes were detected as
+impossible.
+
+Because of performance issues, we ignored the readings from the
+lexicon and used only tag combinations that were named in the grammar.
+For example, the rule \texttt{REMOVE (verb sg) IF (-1 det)}
+gives us ``verb sg'' and ``det'', as opposed to fully specified
+readings in the lexicon.
+The shortcut will work most of the time, but it is possible to
+miss some cases: e.g. \texttt{SELECT (verb sg) IF (0 p3)} 
+requires a full ``verb sg p3'' reading, but our method only gives
+``verb sg'' and ``p3'' separately. 
+Some of the false negatives seem to come from this shortcut, or our naive
+implementation of rules with \verb!*! in their context.
+Due to these reasons, we cannot give conclusions about the Finnish
+grammar; we include the results here only to have some preliminary
+data on scaling up to larger grammars.
 %For a condition situated $n$ or more words to the left or right, we only create a word exactly $n$ words away. This is something to be fixed in the future.
 
-\paragraph{Running time} Currently the speed of the check is an obvious issue,
+\paragraph{Running time} The speed of the check is an obvious issue,
 especially for the Finnish grammar. 
-The 3 hours running time is achieved with using 2100 readings for the
-symbolic words; instead of the lexicon, we only used the readings
-specified in the grammar: for example, the rule \texttt{REMOVE ("bear"
-  noun sg) IF (-1 ())
-Using all 6000 readings from the lexicon, the program
-ran the analysis for the last rule in 30 minutes, making the running
-time for 1000+ rules potentially days.
+As mentioned before, we took a shortcut and used 2000 readings from
+the grammar.
+With the full set of 6000 readings from the lexicon, the speed
+was even worse: checking the last rule took 30 minutes, as opposed to
+3 minutes with 2000 readings. 
+%We never finished running the whole rule set with 6000 readings, but it would have taken potentially days.
+However, the code is currently just a proof of concept, and we plan to
+make an effort in optimisation.
 
@@
 
-The SAT problem for checking one rule grows depending on how many rules come before it, how long the context is, and how many tag combinations are there.
+% The SAT problem for checking one rule grows depending on how many rules come before it, how long the context is, and how many tag combinations are there.
 The running time of the last rule gives some indication ...
 
-Not all questions are so expensive.
-For instance, checking the tag sets in the Finnish grammar doesn't require the context of the other rules at all,
- The majority of the correctly detected errors in the Finnish grammar were rule-internal---these errors don't need the context of the other rules at all.
+% Not all questions are so expensive.
+% For instance, checking the tag sets in the Finnish grammar doesn't require the context of the other rules at all,
+%  The majority of the correctly detected errors in the Finnish grammar were rule-internal---these errors don't need the context of the other rules at all.
 
-%All rules is the total time to check every rule against the previous rules, and last rule % shows how long it took to check the last rule.
-% As indicated by the total time and the time to check the last rule, the size of the SAT problem grows according to how far in the sequence the rule is.
+\paragraph{Future work}
 
-% As seen in table~\ref{table:res}, checking every rule in the Finnish grammar takes currently 3.5 hours.
-% A rule takes longer to run if it appears later in the rule sequence: 1000 previous rules is a larger SAT problem than 10 previous rules.
-% This can be seen from the time it took to compute the result for the last rule.
-For Finnish, the last rule took almost 4 minutes, but if we divide the time with the number of rules, the average time for one rule is just 10 seconds.
-% We also measured the running times for the longest rules in the grammar.
-%\todo{find longest rule, change place of that rule, see what matters more. Promise: we can be smarter about SAT things.}
+We will work on the performance, and fix any errors we found in the
+evaluation, such as the issue with non-exact position.
+Another obvious improvement is to restrict possible ambiguities
+in the symbolic words.
+Currently we have only hard-coded some
+very basic assumptions, such as ``word cannot be ambiguous between a
+boundary and a real word'', but we want to extract these restrictions
+automatically.
+For instance, the Spanish lexicon contains the entries
+\texttt{casa:casa<n><sg>}  and  \texttt{casa:casar<v><sg><p3>}, thus
+we know that a word can be ambiguous between a noun and a verb.
+By adding these restrictions to the symbolic words, we hope to
+restrict the search space, and make the SAT problem easier to solve.
 
-\paragraph{Future work} We will continue improving the performance,
-and make sure that the implementation for the basic rules is correct;
-for instance, fixing the earlier mentioned issue with a non-exact
-position.
-We will also improve usability
+%We will also improve usability, by finding out which rules contrad
 After this, we want to handle the full expressivity of CG-3, with \textsc{map}, \textsc{add} and \textsc{substitute} rules.
@@
 in comparison with a corpus-based method or machine learning.
+Finally, if the method proves feasible for Constraint Grammars, we want
+to try applying it to other, more complex grammar formalisms.
 
-As another line of investigation, we want to detect more kinds of conflicts.
-For instance, say we have the following rules:
+% As another line of investigation, we want to detect more kinds of conflicts.
+% For instance, say we have the following rules:
 
-\begin{itemize}
-\item[] \begin{verbatim}
-REMOVE verb IF (-1C det) ;
-REMOVE noun IF (-1C det) ;
-\end{verbatim}
-\end{itemize}
+% \begin{itemize}
+% \item[] \begin{verbatim}
+% REMOVE verb IF (-1 det) ;
+% REMOVE noun IF (-1 det) ;
+% \end{verbatim}
+% \end{itemize}
 
-With our symbolic sentence, these rules will be no problem; to apply the latter, we only need to construct a target that has a realistic ambiguity with a noun; the verb is gone already.
-However, in the context of real life texts, the grammar writer might actually mean the following:
+% With our symbolic sentence, these rules will be no problem; to apply the latter, we only need to construct a target that has a realistic ambiguity with a noun; the verb is gone already.
+% However, in the context of real life texts, the grammar writer might actually mean the following:
 
-\begin{itemize}
-\item[] \begin{verbatim}
-REMOVE verb IF (-1C det) (0 noun) ;
-REMOVE noun IF (-1C det) (0 verb) ;
-\end{verbatim}
-\end{itemize}
+% \begin{itemize}
+% \item[] \begin{verbatim}
+% REMOVE verb IF (-1 det) (0 noun) ;
+% REMOVE noun IF (-1 det) (0 verb) ;
+% \end{verbatim}
+% \end{itemize}
 
-In that case, those rules are contradictory, and we would like to find it out.
-
-Finally, we want to test the approach to other grammar formalisms.
+% In that case, those rules are contradictory, and we would like to find it out.
 

### 2015-10-25 paper/lrec/lrec2016_listenmaa_claessen.tex
fonts, stuff
more cleanup

@@
 \usepackage{fixltx2e}
-%\usepackage{txfonts}
 \usepackage{url}
@@
 \usepackage{cite}
-%\usepackage{graphics}
 \usepackage{graphicx}
-\usepackage{slantsc}
+%\usepackage{slantsc}
 \usepackage[colorlinks=true,citecolor=blue,urlcolor=blue]{hyperref}
@@
 \usepackage{fontspec}
+\usepackage{wasysym}
 
@@
   Ligatures=TeX,
-  Scale=0.95,
+  Scale=0.9,
 ]{\mainfontfamily}
-%\newfontfamily\scshape[Letters=SmallCaps, Numbers=Uppercase] % {Myriad Pro} %
 
-\newcommand{\monofontfamily}{Monaco} % Courier
+\newfontfamily\scshape[Letters=SmallCaps, Numbers=Uppercase]{Palatino
+  Small Caps}
+
+\newcommand{\monofontfamily}{Monaco} % Monaco
 \setmonofont[Scale=0.8]{\monofontfamily}
@@
 \newcommand{\todo}[1]{{\color{cyan}\textbf{[TODO: }#1\textbf{]}}}
-% \usepackage{tikz}
-% \def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) --
-%   (.25,.15) -- cycle;}
-% \newcommand{\Mypm}{\mathbin{\tikz [x=1.4ex,y=1.4ex,line width=.1ex] \draw (0.0,0) -- (1.0,0) (0.5,0.08) -- (0.5,0.92) (0.0,0.5) -- (1.0,0.5);}}%
 
-\title{Analysing Constraint Grammars with a SAT solver}
+\title{Analysing Constraint Grammars with a SAT-solver}
 
@@
 \begin{abstract}
+%\small
 We describe a method for

### 2015-10-25 paper/lrec/problem_desc.tex
more cleanup

@@
 \item If a rule appears twice, the second occurrence will be disabled by the first
-\item $R$ selects something in a context, $R'$ removes it
+\item $R$ selects something in a context, $R'$ removes~it
 \item $R$ removes something from the context of $R'$, so $R'$ can never
@@
 Our method can be used to
-test out individual rules. For instance, we can ask the tool to
+test individual rules. For instance, we can ask the tool to
 generate example sequences that trigger rule(s) $R$ but not rule(s) $R'$. 
-This functionality can help the grammar writer to see
+This can help the grammar writer to see
 if they have implemented the rules in the intended way.

### 2015-10-25 paper/lrec/eval_futurework.tex
remove more words

@@
 
-We tested three grammars to find rules that cannot apply. 
-The smallest grammar was
+We tested three grammars to find rules that cannot apply: 
 Dutch\footnote{\url{https://svn.code.sf.net/p/apertium/svn/languages/apertium-nld/apertium-nld.nld.rlx}},
-with 59 rules; second was
+with 59 rules; 
 Spanish\footnote{\url{https://svn.code.sf.net/p/apertium/svn/languages/apertium-spa/apertium-spa.spa.rlx}},
-with 279 rules, and the largest was
+with 279 rules; and 
 Finnish\footnote{\url{https://github.com/flammie/omorfi/blob/master/src/vislcg3/omorfi.cg3}},
 with 1185 rules. We left out \textsc{add}, \textsc{map} and other rule
-types introduced in CG-3, and only tested the traditional
-\textsc{remove} and \textsc{select} rules.
+types introduced in CG-3, and only tested \textsc{remove} and \textsc{select} rules.
 The results are shown in table~\ref{table:res}.
@@
 grammars, we were able to verify manually that the detected rules were
-true negatives. However, for the Finnish grammar, we did not have time
-to investigate all of them; in addition, we suspect that some of
-them are false negatives.
+true negatives.
+%The results for the Finnish grammar are inconclusive, 
+% For the Finnish grammar, we did not have time to
+% investigate all of them, but among those we could, we found both true and false negatives.
 We did not check for false positives in any of the grammars.
@@
 marks a possessive suffix with \texttt{Px}, e.g. \texttt{PxPl2} for second person plural, but the
-grammar used the old convention \texttt{PL2} in the rules.
+grammar uses the old convention \texttt{PL2} in the rules.
 As a result, all rules with possessive suffixes were detected as
@@
 
-Because of performance issues, we ignored the readings from the
-lexicon and used only tag combinations that were named in the grammar.
-For example, the rule \texttt{REMOVE (verb sg) IF (-1 det)}
-gives us ``verb sg'' and ``det'', as opposed to fully specified
-readings in the lexicon.
-The shortcut will work most of the time, but it is possible to
+Because of performance issues, we ignored the fully specified readings from the
+lexicon and used only tag combinations from the grammar:
+for example, the rule \texttt{REMOVE (verb sg) IF (-1 det)}
+gives us ``verb sg'' and ``det''.
+The shortcut works most of the time, but it is possible to
 miss some cases: e.g. \texttt{SELECT (verb sg) IF (0 p3)} 
-requires a full ``verb sg p3'' reading, but our method only gives
+assumes ``verb sg p3'', but our method only gives
 ``verb sg'' and ``p3'' separately. 
-Some of the false negatives seem to come from this shortcut, or our naive
-implementation of rules with \verb!*! in their context.
-Due to these reasons, we cannot give conclusions about the Finnish
-grammar; we include the results here only to have some preliminary
-data on scaling up to larger grammars.
+Some of the detected rules seem to be false negatives caused by this shortcut. 
+Another suspect is our naive implementation of rules with \verb!*! in
+their context.
+
+Due to these reasons, the results about the Finnish grammar are
+inconclusive. We only include them to show some preliminary data on larger grammars.
 %For a condition situated $n$ or more words to the left or right, we only create a word exactly $n$ words away. This is something to be fixed in the future.
 
-\paragraph{Running time} The speed of the check is an obvious issue,
+\paragraph{Performance} The speed of the check is an obvious issue,
 especially for the Finnish grammar. 
-As mentioned before, we took a shortcut and used 2000 readings from
-the grammar.
+As mentioned before, we used 2000 readings extracted from the grammar.
 With the full set of 6000 readings from the lexicon, the speed
@@
 3 minutes with 2000 readings. 
-%We never finished running the whole rule set with 6000 readings, but it would have taken potentially days.
-However, the code is currently just a proof of concept, and we plan to
-make an effort in optimisation.
+%We never finished running the whole rule set with 6000 readings, but
+%it would have taken potentially days.
+We plan to make an effort in optimising the code; currently it is just
+a naive proof-of-concept implementation.
 
@@
 
-We will work on the performance, and fix any errors we found in the
+We will work on the performance, and fix the errors we found in the
 evaluation, such as the issue with non-exact position.
-Another obvious improvement is to restrict possible ambiguities
-in the symbolic words.
+We also plan to restrict possible ambiguities in the symbolic words. 
 Currently we have only hard-coded some
 very basic assumptions, such as ``word cannot be ambiguous between a
-boundary and a real word'', but we want to extract these restrictions
-automatically.
-For instance, the Spanish lexicon contains the entries
-\texttt{casa:casa<n><sg>}  and  \texttt{casa:casar<v><sg><p3>}, thus
-we know that a word can be ambiguous between a noun and a verb.
+boundary and a real word'', but we want to extract these constraints
+automatically from a lexicon.
+% For instance, the Spanish lexicon contains the entries
+% \texttt{casa:casa<n><sg>}  and  \texttt{casa:casar<v><sg><p3>}, thus
+% we know that a word can be ambiguous between a noun and a verb.
 By adding these restrictions to the symbolic words, we hope to
@@
 %We will also improve usability, by finding out which rules contrad
-After this, we want to handle the full expressivity of CG-3, with \textsc{map}, \textsc{add} and \textsc{substitute} rules.
-At that stage, it would be interesting to test our program with actual grammar writers,
+For longer-term goals, we want to handle the full expressivity of CG-3,
+with \textsc{map}, \textsc{add} and \textsc{substitute} rules, and
+dependency structure.
+As soon as the tools are mature enough, we want to
+evaluate them with actual grammar writers,
 in comparison with a corpus-based method or machine learning.
-Finally, if the method proves feasible for Constraint Grammars, we want
-to try applying it to other, more complex grammar formalisms.
+Finally, if the method proves feasible for CG, we want
+to try applying it to other grammar formalisms.
 

### 2015-10-25 paper/lrec/problem_desc.tex
remove more words

@@
 
-CGs are valuable resources for rule-based NLP, especially for lesser resourced languages.
-They are robust and can be written without large corpora--only morphological analysis is needed.
-The formalism is lightweight and language-independent, and resources can be shared between related languages \cite{bick2006spanish,lene_trond_linda2010}.
+CGs are valuable resources for rule-based NLP, especially for lesser
+resourced languages. They are robust and can be written without large
+corpora---only morphological analysis is needed. The formalism is
+lightweight and language-independent, and resources can be shared
+between related languages \cite{bick2006spanish}.
+%,lene_trond_linda2010}.
 %Mature CGs contain some thousands of rules, but even small CGs are shown to be effective \cite{lene_trond2011}.
@@
 \item Corpus-based methods in manual grammar development \cite{voutilainen2004}
-\item Optimising hand-written grammars \cite{bick2013tuning}
+\item Optimising hand-written CGs~\cite{bick2013tuning}
 \item Encoding CG in logic \cite{lager98,lager_nivre01,listenmaa_claessen2015}
@@
 
-In addition, there is a large body of research on automatically inducing rules, e.g. \cite{inducing_cg1996}, \cite{lindberg_eineborg98ilp}
+In addition, there is a large body of research on automatically inducing rules, e.g. \cite{inducing_cg1996,lindberg_eineborg98ilp}.
 %, \cite{lager01transformation}, \cite{asfrent14} 
-. However, since our work is aimed to aid the process of hand-crafting rules, we omit those works from our discussion.
+However, since our work is aimed to aid the process of hand-crafting rules, we omit those works from our discussion.
 
@@
 
-\cite{bick2013tuning} modifies the grammar automatically, by trying out different orderings of the rules, and changing their contexts to stricter or less strict.
+\cite{bick2013tuning} modifies the grammar automatically, by trying
+out different rule orders and altering the contexts of the rules.
 Bick reports error reduction of 7--15\% compared to the original grammars.
-As a downside, the grammar writer will likely not know why exactly is the tuned grammar performing better.
+As a downside, the grammar writer will likely not know why exactly does the tuned grammar perform better.
 % At a certain point, the grammar gets so big that it is hard to keep track of all the rules and their interactions.

### 2015-10-25 paper/lrec/eval_futurework.tex
nitpicking

@@
 \# readings      & 276             & 2191     & 2012    \\ \hline
-\# conflicts       & 1               & 1        & \todo{n}    \\ \hline
+\# conflicts       & 1               & 1        & 51    \\ \hline
 \clock{} all rules       & 7s              & 3min 22s    & 3h 26min    \\ \hline
@@
 
-\paragraph{Correctness} 
+\paragraph{Accuracy} 
 The test revealed problems in all grammars, and for the smaller
 grammars, we were able to verify manually that the detected rules were
-true negatives.
+true positives.
 %The results for the Finnish grammar are inconclusive, 
@@
 % investigate all of them, but among those we could, we found both true and false negatives.
-We did not check for false positives in any of the grammars.
+We did not check for false negatives in any of the grammars.
 
@@
 
-For the Finnish grammar, we could verify a number of tagset errors. 
+For the Finnish grammar, we did not have time to investigate all the
+results. We could verify approximately \nicefrac{1}{3} as true positive and
+\nicefrac{1}{3} as false positive.
+The verified true positives were all rule-internal conflicts. 
+The following rule requires the target to be genitive and unambiguously nominative: 
+\begin{itemize}
+\item[]\texttt{"oma" SELECT GEN IF [...] (0C NOM) ;}
+\end{itemize}
+We also detected a number of tagset errors, . 
 The grammar was originally written in 1995, and converted to the
-Omorfi tagset by \cite{pirinen2015}. For example, the Omorfi tagset
-marks a possessive suffix with \texttt{Px}, e.g. \texttt{PxPl2} for second person plural, but the
-grammar uses the old convention \texttt{PL2} in the rules.
-As a result, all rules with possessive suffixes were detected as
-impossible.
+Omorfi tagset by \cite{pirinen2015}, but there were 15 rules that used the old convention for possessive suffixes.
 
-Because of performance issues, we ignored the fully specified readings from the
+%  For example, the Omorfi tagset
+% marks a possessive suffix with \texttt{Px}, e.g. \texttt{PxPl2} for second person plural, but the
+% grammar uses the old convention \texttt{PL2} in the rules.
+% As a result, 15 rules with possessive suffixes were detected as
+% impossible.
+To improve performance, we ignored the fully specified readings from the
 lexicon and used only tag combinations from the grammar:
@@
 The shortcut works most of the time, but it is possible to
-miss some cases: e.g. \texttt{SELECT (verb sg) IF (0 p3)} 
-assumes ``verb sg p3'', but our method only gives
-``verb sg'' and ``p3'' separately. 
-Some of the detected rules seem to be false negatives caused by this shortcut. 
-Another suspect is our naive implementation of rules with \verb!*! in
-their context.
+miss some cases: e.g. \texttt{SELECT PRON + REL IF (0 NOM)} 
+assumes ``pron rel nom'', but our  only gives
+``pron rel'' and ``nom'' separately. 
+17 false positives were caused by this shortcut. 
+In addition, 2 false positives were caused by the naive implementation of
+rules with \verb!*! in their context.
 
@@
 %it would have taken potentially days.
-We plan to make an effort in optimising the code; currently it is just
+We plan to optimise the code; currently it is just
 a naive proof-of-concept implementation.
@@
 
-%We will also improve usability, by finding out which rules contrad
-For longer-term goals, we want to handle the full expressivity of CG-3,
+As for longer-term goals, we want to handle the full expressivity of CG-3,
 with \textsc{map}, \textsc{add} and \textsc{substitute} rules, and
-dependency structure.
+dependency structure. This also means finding different kinds of conflicts.
 As soon as the tools are mature enough, we want to

### 2015-10-25 paper/lrec/eval_futurework.tex
small stuff

@@
 
-We tested three grammars to find rules that cannot apply: 
+We tested three grammars to find conflicting rules: 
 Dutch\footnote{\url{https://svn.code.sf.net/p/apertium/svn/languages/apertium-nld/apertium-nld.nld.rlx}},
@@
 \paragraph{Accuracy} 
-The test revealed problems in all grammars, and for the smaller
+The experiments revealed problems in all grammars. For the smaller
 grammars, we were able to verify manually that the detected rules were
@@
 For the Finnish grammar, we did not have time to investigate all the
-results. We could verify approximately \nicefrac{1}{3} as true positive and
-\nicefrac{1}{3} as false positive.
+results. We could verify approximately \nicefrac{1}{3} as true positives and
+\nicefrac{1}{3} as false positives.
 The verified true positives were all rule-internal conflicts. 
-The following rule requires the target to be genitive and unambiguously nominative: 
+For example, the following rule requires the target to be genitive and unambiguously nominative: 
 \begin{itemize}
@@
 \end{itemize}
-We also detected a number of tagset errors, . 
-The grammar was originally written in 1995, and converted to the
+We also detected a number of tagset errors;
+the grammar was originally written in 1995, and converted to the
 Omorfi tagset by \cite{pirinen2015}, but there were 15 rules that used the old convention for possessive suffixes.
@@
 %it would have taken potentially days.
-We plan to optimise the code; currently it is just
-a naive proof-of-concept implementation.
+We believe we can still signifcantly improve the performance of the SAT-encoding. Currently it is just a naive proof-of-concept implementation, and we have not tried to optimize for speed at all.
 
@@
 By adding these restrictions to the symbolic words, we hope to
-restrict the search space, and make the SAT problem easier to solve.
+restrict the search space, and make the SAT problem easier to solve, while at the same time generate more realistic example sentences.
 

### 2015-10-25 paper/lrec/problem_desc.tex
small stuff

@@
 to keep track of all the rules and their interaction.
-With our tools, the users can diagnose and improve their grammars.
-Some examples of conflicts are listed below:
+With our tools, the users can find conflicting rules, diagnose problems and improve their grammars.
+Some examples of such conflicts are listed below:
 \begin{itemize}
@@
 %In addition to analysing the whole grammar,
-Our method can be used to
-test individual rules. For instance, we can ask the tool to
-generate example sequences that trigger rule(s) $R$ but not rule(s) $R'$. 
+%Our method can be used to
+%check individual rules. For instance,
+We can also ask the tool to
+generate example sentences that trigger rule(s) $R$ but not rule(s) $R'$. 
 This can help the grammar writer to see
 if they have implemented the rules in the intended way.
+
+All of this requires a {\em semantic} rather than a {\em syntactic} analysis.
 

### 2015-10-25 paper/lrec/problem_desc.tex
merging two paragraphs; remove the unrelated related work

@@
 
+While rule-internal conflicts can be detected by simpler means, taking
+care of rule interaction requires a {\em semantic} rather than a {\em syntactic} analysis.
 %In addition to analysing the whole grammar,
@@
 if they have implemented the rules in the intended way.
-
-All of this requires a {\em semantic} rather than a {\em syntactic} analysis.
 
@@
 
-In addition, there is a large body of research on automatically inducing rules, e.g. \cite{inducing_cg1996,lindberg_eineborg98ilp}.
-%, \cite{lager01transformation}, \cite{asfrent14} 
-However, since our work is aimed to aid the process of hand-crafting rules, we omit those works from our discussion.
+% In addition, there is a large body of research on automatically inducing rules, e.g. \cite{inducing_cg1996,lindberg_eineborg98ilp}.
+% %, \cite{lager01transformation}, \cite{asfrent14} 
+% However, since our work is aimed to aid the process of hand-crafting rules, we omit those works from our discussion.
 

### 2015-10-25 paper/lrec/eval_futurework.tex
typofixes, minor tweaks

@@
 miss some cases: e.g. \texttt{SELECT PRON + REL IF (0 NOM)} 
-assumes ``pron rel nom'', but our  only gives
+assumes ``pron rel nom'', but our method only gives
 ``pron rel'' and ``nom'' separately. 
@@
 %it would have taken potentially days.
-We believe we can still signifcantly improve the performance of the SAT-encoding. Currently it is just a naive proof-of-concept implementation, and we have not tried to optimize for speed at all.
+We believe we can still signifcantly improve the performance of the SAT-encoding. Currently it is just a naive proof-of-concept implementation, and we have not tried to optimise for speed at all.
 

### 2015-10-25 paper/lrec/implementation.tex
typofixes, minor tweaks

@@
 The example rule can only affect readings of word2 that have a ``verb'' tag, so we create only two new variables $w_4'$ and $w_5'$ for the result, and reuse the other variables.
-We add the following constraint for $w_4'$ is:
+We add the following constraint for $w_4'$:
 \begin{center}
@@
 \end{itemize}
-If we take the symbolic sentence above as input, we want to ask wether or not it can trigger the rule $R$. We do this by adding some more constraints to the SAT-solver.
+If we take the symbolic sentence above as input, we want to ask whether or not it can trigger the rule $R$. We do this by adding some more constraints to the SAT-solver.
 

### 2016-01-24 CG_parse.hs
add support for subreadings in parser, TODO in match function

@@ -77,7 +77,8 @@ main :: IO ()
 main = do args <- getArgs
           case args of
              [file1,file2] -> do readFile file1 >>= return . parseRules True --verbose
-                                 readData file2
+                                 foo <- parseData `fmap` readFile file2 
+                                 print foo
                                  putStrLn "read rules and data"
              _             -> do putStrLn "Usage: CG_parse <rules> <data>"
                                  exitFailure
@@ -273,8 +274,17 @@ transCondSet cs = do
 transCond :: Cond -> State Env CGB.Condition
 transCond c = case c of
 --  CondPos pos ts          -> liftM2 CGB.C (transPosition pos) (transTagSet' True  ts)
-  CondPos pos ts          -> liftM2 CGB.C (transPosition pos) ((,) True `fmap` transTagSet ts)
-  CondNotPos pos ts       -> liftM2 CGB.C (transPosition pos) (transTagSet' False ts)
+  CondPos pos ts          -> do (pos', subr) <- transPosition pos 
+                                ts' <- transTagSet ts
+                                return $ case subr of
+                                  Nothing -> CGB.C pos' (True, ts')
+                                  Just i  -> CGB.C pos' (True, ts') ---TODO map Subreading to TS
+
+  CondNotPos pos ts       -> do (pos', subr) <- transPosition pos 
+                                ts' <- transTagSet ts
+                                return $ case subr of
+                                  Nothing -> CGB.C pos' (True, ts')
+                                  Just i  -> CGB.C pos' (True, ts') --TODO map Subreading to TS
   CondBarrier pos ts bts  -> barrier pos ts bts True  CGB.Barrier
   CondNotBar pos ts bts   -> barrier pos ts bts False CGB.Barrier
   CondCBarrier pos ts bts -> barrier pos ts bts True  CGB.CBarrier
@@ -308,36 +318,31 @@ transCond c = case c of
             CGB.Barrier b i ts -> CGB.Barrier b newI ts 
             CGB.CBarrier b i ts -> CGB.CBarrier b newI ts 
 
-        transTagSet' :: Bool -> TagSet -> State Env (Bool, CGB.TagSet)
-        transTagSet' b ts = 
-          do tags <- transTagSet ts
-             modify $ \env -> env { unnamed = tags : unnamed env }
[…]

### 2016-01-24 paper/lrec/problem_desc.tex
typofixes, minor tweaks
add intro from the paper

@@
 
+% Inari: Added this for my academic writing course, as for general motivation.
+
+Most of NLP research has been since the 90s aimed towards data-driven methods.
+While these approaches have produced good results for languages with large annotated corpora, they are practically unusable for low-resourced languages.
+Even for languages with many resources, \todo{cite Manning (2011)} argues that rule-based methods are needed to push the accuracy of POS taggers from the 97\%s to the near 100\%.
+POS tagging is a fundamental phase which is required in further syntactic or semantic analysis---hence a good POS-tagger is an important resource for any language.
+We believe that Constraint Grammar is a good candidate especially for languages with low resources, and in this paper we describe methods for CG writers and users to analyse and improve their grammars.
+
 Constraint Grammar (CG, \cite{karlsson1995constraint})
-% \citep[CG][]{karlsson1995constraint} 
+% \citep[CG,][]{karlsson1995constraint} 
 is a formalism used to disambiguate morphologically analysed text. 
@@
         "the" det def
-"<bear>"
-        "bear" noun sg
-        "bear" verb pl
+"<wish>"
+        "wish" noun sg
+        "wish" verb pl
 \end{verbatim}
 \end{itemize}
-the rule will match to the word \emph{bear}, and remove the analysis \texttt{verb pl}.
+the rule will match to the word \emph{wish}, and remove the analysis \texttt{verb pl}.
 However, if the target word has only one remaining analysis, then the rule will not apply, even if the condition is met.
@@
 lightweight and language-independent, and resources can be shared
-between related languages \cite{bick2006spanish}.
-%,lene_trond_linda2010}.
+between related languages \cite{bick2006spanish,lene_trond_linda2010}.
 %Mature CGs contain some thousands of rules, but even small CGs are shown to be effective \cite{lene_trond2011}.

### 2016-02-04 paper/lrec/implementation.tex
more stuff to paper

@@
 
+\paragraph{Creating realistic ambiguities}
+
+\todo{List of word forms+readings; list of distinct readings; make into matrix where we see that x and y can be ambiguous but a and z never; smart things with constraints (Koen)}
+
+Now with this, we can detect even these kinds of rules as problematic:
+
+ \begin{itemize}
+ \item[] \begin{verbatim}
+ REMOVE verb IF (-1 det) ;
+ REMOVE noun IF (-1 det) ;
+ \end{verbatim}
+ \end{itemize}
+
+ With our symbolic sentence, these rules will be no problem; to apply the latter, we only need to construct a target that has a realistic ambiguity with a noun; the verb is gone already.
+ However, in the context of real life texts, the grammar writer might actually mean the following:
+
+ \begin{itemize}
+ \item[] \begin{verbatim}
+ REMOVE verb IF (-1 det) (0 noun) ;
+ REMOVE noun IF (-1 det) (0 verb) ;
+ \end{verbatim}
+ \end{itemize}
+
+ In that case, those rules are contradictory, and we would like to find it out.
+The additional constraints will prevent the SAT solver from creating an ambiguity where noun is ambiguous with e.g.
+punctuation.
+
 \paragraph{Putting it all together}

### 2016-02-04 paper/lrec/problem_desc.tex
more stuff to paper

@@
 
-Most of NLP research has been since the 90s aimed towards data-driven methods.
+%Most of NLP research has been since the 90s aimed towards data-driven methods.
 While these approaches have produced good results for languages with large annotated corpora, they are practically unusable for low-resourced languages.
@@
 between related languages \cite{bick2006spanish,lene_trond_linda2010}.
-%Mature CGs contain some thousands of rules, but even small CGs are shown to be effective \cite{lene_trond2011}.
+Mature CGs contain some thousands of rules, but even small CGs are shown to be effective \cite{lene_trond2011}.
 
@@
 to keep track of all the rules and their interaction.
+This is partly by design: CG is a shallow formalism, and individual rules can be written to address any property, 
+%such as "any verb" or "the word form `foo' which is tagged as agent"
+
 With our tools, the users can find conflicting rules, diagnose problems and improve their grammars. 
+We expect two major use cases: 
+one where the grammar writer takes a whole grammar and cleans it up, and
+other where the grammar writer tests new rules as they write them.
+
+%%%%% Motivate testing some rules, not whole grammar
+
+Given the following $n$ rules, a grammar writer may ask some questions while writing a grammar. 
+\begin{itemize}
+\item[] \texttt{SELECT Inf IF (-1 Para OR De) (0C V) ;}
+\item[] \texttt{SELECT Inf IF (-1 Pr) (0C V) ;}
+\item[] \texttt{SELECT Inf IF (-1C Vai) ;}
+\item[] \texttt{SELECT Inf IF (-1C Vbmod) (0C V) ;}
+\item[] \texttt{SELECT Inf IF (-1C Ter/de) ;}
+\item[] \texttt{SELECT Inf IF (-1C Vbmod) (0 Ser) ;}
+\end{itemize}
+
+\begin{itemize}
+\item Are all the rules distinct? (e.g. Para and De are included in Pr)
+\item Could two or more rules be merged (\texttt{IF -1 Pr OR Vai OR Vbmod ...})?
+\item Does the order of the rules matter? What is the best order?
+\end{itemize}
+
+If there
+
+%%%%%
+
 Some examples of such conflicts are listed below:

### 2016-02-04 CG_Symbolic.hs
add mini test for ambiguity class (4 readings)

@@ -113,7 +113,7 @@ testRule verbose ambcls readings (lastrule,rules) = do
  where 
    constrainStuff s form symbword = do
      let mp ind = literal $ fromMaybe true (IM.lookup ind symbword)
-     --putStrLn $ "constrainStuff: " ++ show symbword
+     putStrLn $ "constrainStuff: " ++ show symbword
      constraints s mp [] form
   
 --------------------------------------------------------------------------------

### 2016-02-04 executable/Analyse.hs

@@ -37,10 +37,16 @@ ex_not = concat $ snd $ parseRules False
        "REMOVE:r2 (a) IF (-1 (a)) ;" ++
        "REMOVE:l  (b) IF (-1 (c)) ;" )
 
-kimmo = concat $ snd $ parseRules False
+kimmo_explicit = concat $ snd $ parseRules False
      ( "REMOVE:r1 (a) IF (0 (b)) (-1 (c)) ;" ++
        "REMOVE:r2 (b) IF (0 (a)) (-1 (c)) ;" ++
        "REMOVE:l  (c) IF (-1 (c)) ;" )
+
+kimmo_implicit = concat $ snd $ parseRules False
+     ( "REMOVE:r1 (a) IF  (-1 (c)) ;" ++
+       "REMOVE:r2 (b) IF  (-1 (c)) ;" ++
+       "REMOVE:l  (c) IF (-1 (c)) ;" )
+
 
 main = do
   args <- getArgs
@@ -53,8 +59,13 @@ main = do
 
   case args of 
     ("kimmo":_)
-       -> do let kimmo' = map (ruleToRule' tagmap allinds) kimmo
-             mapM_ (testRule True "" tc) (splits (reverse kimmo'))
+       -> do let kimmo_i' = map (ruleToRule' tagmap allinds) kimmo_implicit
+             putStrLn "testing with implicit kimmo"
+             mapM_ (testRule True "data/abcd-ambiguity-classes" tc) (splits (reverse kimmo_i'))
+
+             let kimmo_e' = map (ruleToRule' tagmap allinds) kimmo_explicit
+             putStrLn "testing with explicit kimmo"
+             mapM_ (testRule True "data/abcd-ambiguity-classes" tc) (splits (reverse kimmo_e'))
     (lang:r)
        -> do let verbose = "v" `elem` r || "d" `elem` r
              let nosub = "nosub" `elem` r

### 2016-02-10 paper/lrec/eval_futurework.tex
more stuff to paper
add stuff about new features

@@
 \# rules      & 59              & 279      & 1185     \\ \hline
-\# readings      & 276             & 2191     & 2012    \\ \hline
-\# conflicts       & 1               & 1        & 51    \\ \hline
+\# readings   & 336             & 2191     & 2012    \\ \hline
+\# conflicts  & 8 (1 + 5 + 2)   & 1        & 51    \\ \hline
 \clock{} all rules       & 7s              & 3min 22s    & 3h 26min    \\ \hline
@@
 \end{tabular}
-\caption{Detecting rules that cannot be applied}
+\caption{}
 \label{table:res}
@@
 the set definition, our test did not reveal other problems.
-The Dutch grammar had the following rules in the given order, making the latter impossible to apply:
+
+The Dutch grammar had two kinds of errors: rule-internal due to nonexistent tags, and rule interaction.
+As for rule-internal conflicts, one was due to a misspelling in the list definition for personal pronouns, which rendered 5 rules ineffective.
+The other was about subreadings: the genitive \emph{s} is analysed as a subreading in the Apertium morphological analyser, but it appeared in 2 rules as the main reading.
+
+There were two genuine conflicts with rule interaction. The first one is shown below:
+
 \begin{itemize}
@@
 \end{itemize}
+
+The first one is more straight-forward: two rules, which both remove an adverb, were in an order where the first has more broad condition: remove adverb if followed by a noun.
+In contrast, the second rule has a stricter requirement: only remove adverb if it is preceded by a determiner; the adverb itself is ambiguous with an adjective, and followed by a noun.
+The proble is that the first rule removes the adverb in all possible cases, and the second will not have any chance to act.
+If the rules were in the opposite order, then there would be no problem.
+
+The second conflict, shown below, is more subtle.
+
+\begin{itemize}
+\item[] 
+\begin{verbatim}SELECT DetPos IF (-1 (vbser pres p3 sg)) (0 ("zijn"i)) (1 Noun);
+SELECT DetPosNotZijn IF (1 Noun);
+\end{verbatim}
+\end{itemize}
+
+On the first glance it seems that there is no conflict: the first rule requires a Det Pos which is at the same time ``zijn'', as seen by the requirement \texttt{(0 ("zijn"i))}.
+However, this has a different semantics in the CG-3 formalism, as well as in SAT-CG.
+The rules match a situation where there is \emph{any} Det Pos and \emph{any} zijn, including the verb reading.
+The correct way would be
 
@@
 
-% As another line of investigation, we want to detect more kinds of conflicts.
-% For instance, say we have the following rules:
-
-% \begin{itemize}
-% \item[] \begin{verbatim}
-% REMOVE verb IF (-1 det) ;
-% REMOVE noun IF (-1 det) ;
-% \end{verbatim}
-% \end{itemize}
-
-% With our symbolic sentence, these rules will be no problem; to apply the latter, we only need to construct a target that has a realistic ambiguity with a noun; the verb is gone already.
-% However, in the context of real life texts, the grammar writer might actually mean the following:
-
-% \begin{itemize}
-% \item[] \begin{verbatim}
-% REMOVE verb IF (-1 det) (0 noun) ;
-% REMOVE noun IF (-1 det) (0 verb) ;
-% \end{verbatim}
-% \end{itemize}
-
-% In that case, those rules are contradictory, and we would like to find it out.
-
 % * Preliminary results

### 2016-02-10 paper/lrec/implementation.tex
add stuff about new features

@@
 
+There is another point to mention: the readings in grammar can be underspecified (e.g. `verb sg''), whereas the lexicon only gives us fully specified (``verb pres p2 sg'') readings. 
+In addition, the sets in the grammar may include word forms or lemmas, but we remove those from the readings that we extract from the lexicon, in order to have 100s-1000s of variables instead of 100 000s.
+A simple solution is to retain the lemmas and word forms only for those entries where it is specified in the grammar, and otherwise leave them out.
+For example, the grammar may contain the following list:
+
+\begin{itemize}
+ \item[] \texttt{LIST ZijnVerb = ("zijn" vbser pres pl) ;}
+\end{itemize}
+
+Clearly there is something special about the verb \emph{zijn}, not just any verb. Looking at the lexicon, we find \emph{zijn} in the following entries:
+
+\begin{itemize}
+ \item[] \texttt{zijn:zijn<det><pos><mfn><pl>}
+ \item[] \texttt{zijn:zijn<det><pos><mfn><sg>}
+ \item[] \texttt{zijn:zijn<vbser><inf>}
+ \item[] \texttt{zijn:zijn<vbser><pres><pl>}
+\end{itemize}
+
+Thus we add special entries for these: in addition to the anonymous \texttt{<det><pos><mfn><pl>} reading, we add \texttt{<"zijn"><det><pos><mfn><pl>}.
+
 Now with this, we can detect even these kinds of rules as problematic:

### 2016-02-11 paper/lrec/lrec2016_listenmaa_claessen.tex
remove more words
nitpicking
typofixes, minor tweaks
update makefile & add Fran to acknowledgements
use LREC official stylesheet

@@
-%
-% File nodalida2015.tex
-%
-% Contact beata.megyesi@lingfil.uu.se
-%
-% Based on the instruction file for EACL 2014
-% which in turn was based on the instruction files for previous 
-% ACL and EACL conferences.
+\documentclass[10pt, a4paper]{article}
+\usepackage{lrec2016}
+\usepackage{multibib}
+\newcites{languageresource}{Language Resources}
+\usepackage{graphicx} % for eps graphics
+\usepackage{epstopdf}
+\usepackage[latin1]{inputenc}
 
-\documentclass[11pt]{article}
-\usepackage{nodalida2015}
-\usepackage{times}
-\usepackage{mathptmx}
-\usepackage{fixltx2e}
-\usepackage{url}
-\usepackage{multirow}
-\usepackage{latexsym}
-\usepackage{cite}
-\usepackage{graphicx}
-%\usepackage{slantsc}
+% Inari's additions
 \usepackage[colorlinks=true,citecolor=blue,urlcolor=blue]{hyperref}
-\usepackage{authordate1-4}
-\usepackage{multirow}
-%\special{papersize=210mm,297mm} % to avoid having to use "-t a4" with dvips 
-%\setlength\titlebox{6.5cm}  % You can expand the title box if you really have to
-
-\usepackage{fontspec}
-\usepackage{wasysym}
-
-\newcommand{\mainfontfamily}{Palatino}
-\setmainfont[
-  Ligatures=TeX,
-  Scale=0.9,
-]{\mainfontfamily}
-
-\newfontfamily\scshape[Letters=SmallCaps, Numbers=Uppercase]{Palatino
-  Small Caps}
-
-\newcommand{\monofontfamily}{Monaco} % Monaco
-\setmonofont[Scale=0.8]{\monofontfamily}
-
 \usepackage{color}
 \newcommand{\todo}[1]{{\color{cyan}\textbf{[TODO: }#1\textbf{]}}}
+\usepackage{units} % for nicefrac
+\usepackage{wasysym} % for clock
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-\author{Inari Listenmaa \and Koen Claessen \\
- Chalmers University of Technology, Gothenburg, Sweden \\
- {\tt \{inari,koen\}@chalmers.se} }
+\name{Inari Listenmaa, Koen Claessen}
 
-\date{\today}
+\address{University of Gothenburg, Chalmers University of Technology \\
+         Gothenburg, Sweden \\
+         inari@chalmers.se, koen@chalmers.se\\}
 
-\begin{document}
-\maketitle
-
-\begin{abstract}
-%\small
-We describe a method for
- analysing Constraint Grammars by encoding the
+\abstract{
+We describe a method for analysing Constraint Grammars by encoding the
 rules in SAT.
@@
 This can help users to diagnose and improve their grammars.
-No corpus is required, only a morphological lexicon.
-%We test our tools with grammars from 100-1000 rules.
-\end{abstract}
+No corpus is required, only a morphological lexicon. \\ 
+\newline 
+\Keywords{Constrain Grammar, SAT, Grammar Analysis} }
+
+\begin{document}
+
+\maketitleabstract
 
@@
 \section*{Acknowledgments}
-We thank Eckhard Bick for the idea to apply SAT to CG analysis.
+We thank Eckhard Bick for the idea to apply SAT to CG analysis, and Francis Tyers for insightful discussions.
 
-\bibliographystyle{acl}
+\section{Bibliographical References}
+\label{main:ref}
+
+\bibliographystyle{lrec2016}
 \bibliography{cg}
 
+% \section{Language Resource References}
+% \label{lr:ref}
+% \bibliographystylelanguageresource{lrec2016}
+% \bibliographylanguageresource{cg}
+
 \end{document}

### 2016-02-11 paper/lrec/problem_desc.tex
more text

@@
 
-Constraint Grammar (CG, \cite{karlsson1995constraint})
+Constraint Grammar (CG, \newcite{karlsson1995constraint})
 % \citep[CG,][]{karlsson1995constraint} 
@@
         "wish" verb pl
+        "wish" verb inf
 \end{verbatim}
 \end{itemize}
-the rule will match to the word \emph{wish}, and remove the analysis \texttt{verb pl}.
+the rule will match to the word \emph{wish}, and remove the analyses \texttt{verb pl} and \texttt{verb inf}.
 However, if the target word has only one remaining analysis, then the rule will not apply, even if the condition is met.
@@
 
-As CGs grow larger, it gets harder for the grammar writers and users
-to keep track of all the rules and their interaction.
-This is partly by design: CG is a shallow formalism, and individual rules can be written to address any property, 
-%such as "any verb" or "the word form `foo' which is tagged as agent"
+By design, CG is a shallow and robust formalism.
+Individual rules can be written to address any property, such as ``verb'', ``copula verb in first person singular'', or ``the word form \emph{sailor} preceded by \emph{drunken} anywhere in the sentence''.
+This makes it possible to treat very particular edge cases without touching the more general rule: we would simply write the narrow rule first (``if noun AND \emph{sailor}''), and use a general rule (``if noun'') as catch-all.
 
-With our tools, the users can find conflicting rules, diagnose problems and improve their grammars. 
+However, this design is not without problems. As CGs grow larger, it gets harder to keep track of all the rules and their interaction.
+Our tools will help grammar writers and users to find conflicting rules, diagnose problems and improve their grammars. 
 We expect two major use cases: 
-one where the grammar writer takes a whole grammar and cleans it up, and
-other where the grammar writer tests new rules as they write them.
+first, to test the effect of new rules while writing a grammar, and second, to take a complete grammar and analyse it as a whole, to find conflicts or dead rules.
 
-%%%%% Motivate testing some rules, not whole grammar
+\begin{figure}[t]
+\begin{center}
 
-Given the following $n$ rules, a grammar writer may ask some questions while writing a grammar. 
 \begin{itemize}
-\item[] \texttt{SELECT Inf IF (-1 Para OR De) (0C V) ;}
-\item[] \texttt{SELECT Inf IF (-1 Pr) (0C V) ;}
-\item[] \texttt{SELECT Inf IF (-1C Vai) ;}
-\item[] \texttt{SELECT Inf IF (-1C Vbmod) (0C V) ;}
-\item[] \texttt{SELECT Inf IF (-1C Ter/de) ;}
-\item[] \texttt{SELECT Inf IF (-1C Vbmod) (0 Ser) ;}
+\item[]\begin{verbatim}SELECT Inf IF (-1 Para OR De) (0C V) ;
+SELECT Inf IF (-1 Prep) (0C V) ;
+SELECT Inf IF (-1C Vai) ;
+SELECT Inf IF (-1C Vbmod) (0C V) ;
+SELECT Inf IF (-1C Ter/de) ;
+SELECT Inf IF (-1C Vbmod) (0 Ser) ;
+\end{verbatim}
 \end{itemize}
 
+\caption{Rules to select infinitive.}
+\end{center}
+
+\label{infrules}
+\end{figure}
+
+Given the rules in figure~\ref{infrules}, a grammar writer may ask the following questions while writing a grammar.
+
 \begin{itemize}
-\item Are all the rules distinct? (e.g. Para and De are included in Pr)
-\item Could two or more rules be merged (\texttt{IF -1 Pr OR Vai OR Vbmod ...})?
-\item Does the order of the rules matter? What is the best order?
+\item Are all the rules distinct? (e.g. \texttt{Para} and \texttt{De} may be included in \texttt{Prep})
+\item Could two or more rules be merged? (e.g. \texttt{SELECT Inf IF -1 Prep OR Vai OR Vbmod ...})
+\item What is the best order for the rules?
+\item Generate a sequence that triggers rule(s) $R$ but not rule(s) $R'$. 
 \end{itemize}
-
-If there
 
@@
 
-Some examples of such conflicts are listed below:
+For the second use case, here are examples of conflicts that our tools will detect.
 \begin{itemize}
@@
 \item $R$ has an internal conflict, such as non-existent
-tag combination, or contradicting requirements for a context word
+tag combination, or contradictory requirements for a context word
 \end{itemize}
@@
 %check individual rules. For instance,
-We can also ask the tool to
-generate example sentences that trigger rule(s) $R$ but not rule(s) $R'$. 
-This can help the grammar writer to see
-if they have implemented the rules in the intended way.
 
@@
 
-% In addition, there is a large body of research on automatically inducing rules, e.g. \cite{inducing_cg1996,lindberg_eineborg98ilp}.
-% %, \cite{lager01transformation}, \cite{asfrent14} 
-% However, since our work is aimed to aid the process of hand-crafting rules, we omit those works from our discussion.
+In addition, there is a large body of research on automatically inducing rules, e.g. \cite{inducing_cg1996,lindberg_eineborg98ilp,lager01transformation,asfrent14} 
+However, since our work is aimed to aid the process of hand-crafting rules, we omit those works from our discussion.
 
@@
 
-\cite{bick2013tuning} modifies the grammar automatically, by trying
+\newcite{bick2013tuning} modifies the grammar automatically, by trying
 out different rule orders and altering the contexts of the rules.
@@
 %The earlier works on logical reconstruction don't envision grammar analysis as one of the use cases,
-\cite{lager98} and \cite{lager_nivre01} reconstruct the CG formalism in first-order predicate logic.
+\newcite{lager98} and \newcite{lager_nivre01} reconstruct the CG formalism in first-order predicate logic.
 %, and \cite{listenmaa_claessen2015} implement a CG compiler using a SAT-solver.

### 2016-02-12 paper/lrec/implementation.tex
right path for cg.bib
add stuff about lexical entries and clitics

--- 2016-02-04 paper/lrec/implementation.tex
@@
 
+\paragraph{Choosing the readings}
+
+Earlier we have shown an example with 5 readings (\texttt{det def}, \texttt{noun sg}, ...).
+In the real implementation, we operate between hundreds and thousands of readings.
+This is very much dependent on language: Dutch, the simplest language, has 336 readings.
+
+We expand an Apertium morphological lexicon, throw away the word forms and take all the analyses. An exception is when the grammar contains rules that target a specific lemma or word form.
+A simple solution is to retain the lemmas and word forms only for those entries where it is specified in the grammar, and otherwise leave them out.
+For example, the grammar may contain the following list:
+
+\begin{itemize}
+ \item[] \texttt{LIST ZijnVerb = ("zijn" vbser pres pl) ;}
+\end{itemize}
+
+Clearly there is something special about the verb \emph{zijn}, not just any verb. Looking at the lexicon, we find \emph{zijn} in the following entries:
+
+\begin{itemize}
+ \item[] \begin{verbatim}zijn:zijn<det><pos><mfn><pl>
+zijn:zijn<det><pos><mfn><sg>
+zijn:zijn<vbser><inf>
+zijn:zijn<vbser><pres><pl>
+\end{verbatim}
+\end{itemize}
+
+Thus we add special entries for these: in addition to the anonymous
+ \texttt{<det><pos><mfn><pl>} reading, we add \texttt{<"zijn"><det><pos><mfn><pl>}.
+
+ However, for languages with more readings, this may not be feasible.
+For instance, the Spanish lexicon has a high number of readings mostly because of the verbs: in addition to the verb forms, there are clitics attached to them.
+The number of verb readings without clitics is 213, and with clitics 1572.
+
+ In our experience,  even ignoring the clitics doubles the amount of readings.
+
+Another note: the readings in grammar can be underspecified (e.g. ``verb sg''), whereas the lexicon only gives us fully specified (``verb pres p2 sg'') readings.
+
 \paragraph{Creating realistic ambiguities}

### 2016-02-12 paper/lrec/implementation.tex
more stuff

@@
 
-Another note: the readings in grammar can be underspecified (e.g. ``verb sg''), whereas the lexicon only gives us fully specified (``verb pres p2 sg'') readings.
+Another note: the readings in grammar can be underspecified (e.g.
+``verb sg''), whereas the lexicon only gives us fully specified (``verb pres p2 sg'') readings.
+\todo{Is this suitable for implementation, or sh0ould it be in eval/discussion/something else?} We tried a version where we took the tag combinations specified in the grammar as readings, and we could insert them into the symbolic sentences as well, but this was not an ideal solution: turns out that the tag lists in the grammars contain often errors (e.g.
+replacing OR with an AND; using a nonexistent tag; using a wrong subreading number), and if we accept those lists as readings, we will generate symbolic sentences that are impossible, and won't discover the bug in the grammar.
+
+However, if we only want to find rule interaction effects, then using the underspecified readings from the grammar makes the task faster, and it will still catch potential interaction errors.
+
+All in all, our solution to hardcode the tag combinations in the readings is feasible for simple morphology, but it can cause problems with more complex morphology.
+One big downside is that in order to implement \textsc{ADD}, \textsc{ADDREADING} and \textsc{MAP}, we need to be prepared for new readings--even if the lexicon gives all readings that exist in the lexicon, the user might give a nonexistent reading, or in the case of MAP, a syntactic tag, which is (by definition) not in the lexicon.
 
 \paragraph{Creating realistic ambiguities}
+
+In the previous paragraph, we have created realistic \emph{readings}:  ``conditional'' can only go in verb readings, or ``comparative'' with adjective readings.
+This is done just by hardcoding those combinations into variables.
+The next step in creating realistic ambiguities is to constrain what readings can go together.
+For instance, the case of \emph{zijn} shows us that a determiner and a verb, in present tense plural or infinitive, is a possible ambiguity.
+In contrast, there is no word form in the lexicon
+that would be ambiguous between an adjective and a comma, hence we don't want to generate such ambiguity in our symbolic sentences.
+
+We solve the problem by looking at the lexicon again.
+We extract all the words and form ambiguity classes \todo{find if that's a term that should be cited/who invented it and where it was first used}: all the words that can be ambiguous with each other.
+However, with thousands of readings, this would
+extract all the  leaves us with so many combinations, that hardcoding wouldn't be efficient. We
 

### 2016-02-12 paper/lrec/eval_futurework.tex
more text

@@
 \item[] 
-\begin{verbatim}SELECT DetPos IF (-1 (vbser pres p3 sg)) (0 ("zijn"i)) (1 Noun);
-SELECT DetPosNotZijn IF (1 Noun);
+\begin{verbatim}SELECT DetPos IF (-1 Verb) (0 "zijn") 
+                 ( 1 Noun) ;
+SELECT DetPosNotZijn IF (1 Noun) ;
 \end{verbatim}
@@
 
-On the first glance it seems that there is no conflict: the first rule requires a Det Pos which is at the same time ``zijn'', as seen by the requirement \texttt{(0 ("zijn"i))}.
-However, this has a different semantics in the CG-3 formalism, as well as in SAT-CG.
-The rules match a situation where there is \emph{any} Det Pos and \emph{any} zijn, including the verb reading.
-The correct way would be
+On the first glance it seems that there is no conflict: the first rule requires a Det Pos which is at the same time ``zijn'', as seen by the requirement \texttt{(0 "zijn")}.
+However, this has a different semantics in VISL CG-3, which is also repeated in our tool.
+The rules match a situation where there is \emph{any} Det Pos and \emph{any} zijn, including the verb.
+The correct way to formulate the first rule would be the following:
+
+\begin{itemize}
+\item[] \texttt{SELECT DetPos + "zijn" IF (-1 Verb) (1 Noun) ;}
+\end{itemize}
 
@@
 
-We will work on the performance, and fix the errors we found in the
-evaluation, such as the issue with non-exact position.
-We also plan to restrict possible ambiguities in the symbolic words. 
-Currently we have only hard-coded some
-very basic assumptions, such as ``word cannot be ambiguous between a
-boundary and a real word'', but we want to extract these constraints
-automatically from a lexicon.
-% For instance, the Spanish lexicon contains the entries
-% \texttt{casa:casa<n><sg>}  and  \texttt{casa:casar<v><sg><p3>}, thus
-% we know that a word can be ambiguous between a noun and a verb.
-By adding these restrictions to the symbolic words, we hope to
-restrict the search space, and make the SAT problem easier to solve, while at the same time generate more realistic example sentences.
+Our solution to hardcode the tag combinations in the readings is feasible for simple morphology, but it can cause problems with more complex morphology.
+One big downside is that in order to implement \textsc{ADD}, \textsc{ADDREADING} and \textsc{MAP}, we need to be prepared for new readings--even if the lexicon gives all readings that exist in the lexicon, the user might give a nonexistent reading, or in the case of MAP, a syntactic tag, which is (by definition) not in the lexicon.
+A more scalable solution would be to make each tag a variable, and ask the question ``can this reading be a noun?
+how about singular?
+how about conditional?'' separately for each tag.
+Then we could lift the restriction of tag combinations into a SAT solver: make SAT clauses that prohibit a comparative to go with a verb, or conditional with a noun.
 

### 2016-02-12 paper/lrec/implementation.tex
more text

--- 2016-02-12 paper/lrec/implementation.tex
@@
 Earlier we have shown an example with 5 readings (\texttt{det def}, \texttt{noun sg}, ...).
-In the real implementation, we operate between hundreds and thousands of readings.
+In a realistic case, we operate between hundreds and thousands of readings.
 This is very much dependent on language: Dutch, the simplest language, has 336 readings.
@@
 A simple solution is to retain the lemmas and word forms only for those entries where it is specified in the grammar, and otherwise leave them out.
-For example, the grammar may contain the following list:
+For example, the Dutch grammar contains the following rule:
 
 \begin{itemize}
- \item[] \texttt{LIST ZijnVerb = ("zijn" vbser pres pl) ;}
+ \item[] \texttt{REMOVE ("zijn"i vbser) IF (-1 (pr)) (1 Noun) ;}
 \end{itemize}
 
-Clearly there is something special about the verb \emph{zijn}, not just any verb. Looking at the lexicon, we find \emph{zijn} in the following entries:
+This hints that there is something special about the verb \emph{zijn}, compared to the other verbs. Looking at the lexicon, we find \emph{zijn} in the following entries:
 
@@
 
-Another note: the readings in grammar can be underspecified (e.g. ``verb sg''), whereas the lexicon only gives us fully specified (``verb pres p2 sg'') readings.
+Another note: the readings in grammar can be underspecified (e.g.
+\texttt{verb sg}), whereas the lexicon only gives us fully specified (\texttt{verb pres p2 sg}) readings.
+We tried a version where we took the tag combinations specified in the grammar as readings, and we could insert them into the symbolic sentences as well, but this was not an ideal solution: turns out that the tag lists in the grammars contain often errors (e.g.
+replacing OR with an AND; using a nonexistent tag; using a wrong level in a subreading), and if we accept those lists as readings, we will generate symbolic sentences that are impossible, and won't discover the bug in the grammar.
+
+However, if we only want to find rule interaction effects, then using the underspecified readings from the grammar makes the task faster, and it will still catch potential interaction errors.
 
@@
 
-\todo{List of word forms+readings; list of distinct readings; make into matrix where we see that x and y can be ambiguous but a and z never; smart things with constraints (Koen)}
+In the previous paragraph, we have created realistic \emph{readings}, by simply hardcoding legal tag combinations into variables.
+The next step in creating realistic ambiguities is to constrain what readings can go together.
+For instance, the case of \emph{zijn} shows us that ``determiner or verb'', is a possible ambiguity.
+In contrast, there is no word form in the lexicon that would be ambiguous between an adjective and a comma, hence we don't want to generate such ambiguity in our symbolic sentences.
+
+We solve the problem creating \emph{ambiguity classes}: groups of analyses that can be ambiguous with each other. 
+We represent the expanded morphological lexicon as a matrix, as seen in figure~\todo{make a nice picture with rows and columns}: word forms on the rows and analyses on the columns.
+Each distinct row forms an ambiguity class.
+For example, the ambiguity class that
+and form clusters based on which readings can be confused with each other. 
+%This is done after the previous step: we simply treat the lemma or word form as another tag. 
+Then we form SAT clauses that allow or prohibit certain combinations. These clauses will interact with the constraints created from the rules, and the end result will be closer to real-life sentences.
+
+Our approach is similar to \todo{cite Cutting, Kupiec et al.}, who use ambiguity classes instead of distinct word forms, in order to reduce the number of parameters in a Hidden Markov Model.
+They take advantage of the fact that they don't have to model ``bear'' and ``wish'' as separate entries, but they can just reduce it to ``word that can be ambiguous between noun and verb'', and use it as a parameter in their HMM.
+We can do a similar thing by saving a list of words with each ambiguity class.
+For example, we map the ambiguity class ``feminine plural noun or a separable verb infinitive'' to the list of word forms \{``uitgaven'', ``toespraken''\}, and then, if we generate such reading for our symbolic word, we can give one of these words as an example word.
 

### 2016-02-16 Matrix.hs
update ambiguity classes after adding , . and () to the lexicon

@@ -32,8 +32,8 @@ main = do
                 let ts = concat tcs
                 let readings = map (toReadingMap tcs) ws 
                 --mapM_ print (take 50 readings)
-                mapM_ (print) $ M.toList $ toGraph ( readings)
---                mapM_ print $ toGraph readings
+--                mapM_ (print) $ M.toList $ toGraph ( readings)
+                mapM_ print $ toGraph readings
 
 --------------------------------------------------------------------------------
 
@@ -43,8 +43,8 @@ toGraph readings = readingMap
   group' acc (a,b) = case M.lookup a acc of
                        Just c  -> M.adjust (b:) a acc 
                        Nothing -> M.insert a [b] acc
-  readingMap = invertMap 99999 $ foldl group' M.empty readings 
-  --readingMap = onlyValues 99999 $ foldl group' M.empty readings 
+  --readingMap = invertMap 99999 $ foldl group' M.empty readings 
+  readingMap = onlyValues 99999 $ foldl group' M.empty readings 
 
 
 --weet:weten<vblex><imp><sg> 

### 2016-02-16 paper/lrec/eval_futurework.tex
more paper

@@
 % investigate all of them, but among those we could, we found both true and false negatives.
-We did not check for false negatives in any of the grammars.
+We did not check for false negatives in any of the grammars. However, the test reported a high number of errors even in the smallest grammars, and some of them are intuitively not that bad; we
 
-The Spanish grammar had an erroneous set definition, where a word was
-required to have almost all POS tags at the same time. After fixing
-the set definition, our test did not reveal other problems.
-
-The Dutch grammar had two kinds of errors: rule-internal due to nonexistent tags, and rule interaction.
+\paragraph{Dutch} The Dutch grammar had two kinds of errors: rule-internal due to nonexistent tags, and rule interaction.
 As for rule-internal conflicts, one was due to a misspelling in the list definition for personal pronouns, which rendered 5 rules ineffective.
@@
 
-For the Finnish grammar, we did not have time to investigate all the
+\paragrahp{Spanish} The Spanish grammar had proportionately the highest number of errors.\footnote{TODO://link/to/error.log} \todo{confirm; I assume this is gonna be so, but let's see Fin & Por}.
+The grammar we ran is like the one found in the Apertium repository, apart from 2 changes: we fixed some typos (capital O for 0) in order to make it compile, and commented out two rules that used regular expressions, because our solution does not handle them properly (see more on Future work/Implementation).
+
+Again, we can classify the conflicts into internal and interaction.
+As an example of internal conflict, there are \todo{n} rules that use \texttt{SET Cog = (np cog)}---the problem is that the tag \texttt{cog} does not exist in the Apertium dictionary for Spanish.
+It is likely that this grammar has been written for an earlier version, where such tag has been in place.
+
+The Spanish grammar has a high number of interaction conflicts, all with the same pattern:
+
+\begin{itemize}
+\item[] 
+\begin{verbatim}REMOVE (a) IF (-1 x) (1 y) ;
+REMOVE (a) IF (-1 x) (1 y) (NOT -1 z);
+\end{verbatim}
+\end{itemize}
+
+This is another, perhaps counter-intuitive instance of more general rule applying before more specific.
+Negated context works as well to make a rule stricter.
+The second rule will never apply, because the first will catch all cases.
+
+In addition, there was a number of definitions that were never used.
+The program doesn't currently point them out to the user, but that is a trivial addition; just like rule-internal conflicts, it doesn't require any semantic analysis.
+Another error, which was not pointed by the program, but discovered
+
+\paragrahp{Finnish} \todo{Rerun the tests & rewrite this whole section!} For the Finnish grammar, we did not have time to investigate all the
 results. We could verify approximately \nicefrac{1}{3} as true positives and
@@
 Then we could lift the restriction of tag combinations into a SAT solver: make SAT clauses that prohibit a comparative to go with a verb, or conditional with a noun.
+Another benefit of this solution is the addition of lemmas and word forms, as well as regular expressions: currently, if we want to add one more lemma for a verb, we need to create as many new variables as there are distinct verb forms---easily hundreds for languages with rich morphology.
+\todo{less rambling: For rules with regular expressions, this would blow up even more: there are rules that only address whether the word form in a condition starts with a lowercase letter, or whether it ends in a certain suffix.
+A realistic, albeit unsatisfactory option is just to treat it as an underspecified reading, and offer that same regular expression to the user when generating a symbolic sentence.
+Another, less feasible, solution is to grep all the word forms or lemmas that match the regular expression in question---in the worst case (e.g.
+lower case), this will match all the entries in the expanded lexicon.}
 

### 2016-02-16 paper/lrec/implementation.tex
more paper

@@
  However, for languages with more readings, this may not be feasible.
-For instance, the Spanish lexicon has a high number of readings mostly because of the verbs: in addition to the verb forms, there are clitics attached to them.
+For instance, Spanish has a high number of readings, not only because of many inflectional forms, but because it is possible to add 1-2 clitics to
+  lexicon has a high number of readings mostly because of the verbs: in addition to the verb forms, there are clitics attached to them.
 The number of verb readings without clitics is 213, and with clitics 1572.

### 2016-02-16 paper/lrec/eval_futurework.tex
fix typos; separate eval and future work in different sections

@@
-\section{Evaluation and Future Work}
+\section{Evaluation}
 \label{sec:eval}
@@
 
-\paragrahp{Spanish} The Spanish grammar had proportionately the highest number of errors.\footnote{TODO://link/to/error.log} \todo{confirm; I assume this is gonna be so, but let's see Fin & Por}.
+\paragraph{Spanish} The Spanish grammar had proportionately the highest number of errors.\footnote{TODO://link/to/error.log} \todo{confirm; I assume this is gonna be so, but let's see Fin and Por}.
 The grammar we ran is like the one found in the Apertium repository, apart from 2 changes: we fixed some typos (capital O for 0) in order to make it compile, and commented out two rules that used regular expressions, because our solution does not handle them properly (see more on Future work/Implementation).
@@
 
-\paragrahp{Finnish} \todo{Rerun the tests & rewrite this whole section!} For the Finnish grammar, we did not have time to investigate all the
+\paragraph{Finnish} \todo{Rerun the tests and rewrite this whole section!} For the Finnish grammar, we did not have time to investigate all the
 results. We could verify approximately \nicefrac{1}{3} as true positives and
@@
 
-\paragraph{Future work}
+\section{Future work}
 

### 2016-02-24 paper/lrec/problem_desc.tex
add text in intro about SAT

@@
 By design, CG is a shallow and robust formalism. 
-Individual rules can be written to address any property, such as ``verb'', ``copula verb in first person singular'', or ``the word form \emph{sailor} preceded by \emph{drunken} anywhere in the sentence''.
-This makes it possible to treat very particular edge cases without touching the more general rule: we would simply write the narrow rule first (``if noun AND \emph{sailor}''), and use a general rule (``if noun'') as catch-all.
+There is no particular hierarchy between lexical, morphological,
+syntactic or even semantic tags: individual rules can be written to address any
+property, such as ``verb'', ``copula verb in first person singular'',
+or ``the word form \emph{sailor}, preceded by \emph{drunken} anywhere in the
+sentence''. This makes it possible to treat very particular edge
+cases without touching the more general rule: we would simply write
+the narrow rule first (``if noun AND \emph{sailor}''), and introduce
+the general rule (``if noun'') later.
 
-However, this design is not without problems. As CGs grow larger, it gets harder to keep track of all the rules and their interaction.
-Our tools will help grammar writers and users to find conflicting rules, diagnose problems and improve their grammars. 
+However, this design is not without problems. As CGs grow larger, it
+gets harder to keep track of all the rules and their interaction.
+Our tools will help grammar writers and users to find conflicting
+rules, diagnose problems and improve their grammars. 
 We expect two major use cases: 
-first, to test the effect of new rules while writing a grammar, and second, to take a complete grammar and analyse it as a whole, to find conflicts or dead rules.
+first, to test the effect of new rules while writing a grammar, and
+second, to take a complete grammar and analyse it as a whole, to find
+conflicts or dead rules.
 
@@
 \begin{itemize}
-\item[]\begin{verbatim}SELECT Inf IF (-1 Para OR De) (0C V) ;
+\item[]
+\begin{verbatim}SELECT Inf IF (-1 Para OR De) (0C V) ;
 SELECT Inf IF (-1 Prep) (0C V) ;
@@
 
-\caption{Rules to select infinitive.}
+\caption{Rules to select infinitive in Portuguese.}
 \end{center}
@@
 While rule-internal conflicts can be detected by simpler means, taking
-care of rule interaction requires a {\em semantic} rather than a {\em syntactic} analysis.
-%In addition to analysing the whole grammar,
-%Our method can be used to
-%check individual rules. For instance,
+care of rule interaction requires a {\em semantic} rather than a {\em
+ syntactic} analysis.
+In order to find effects of rule interaction, we must keep track of
+the possible sentences at each step. After each rule, we have two
+possibilities: the rule fires, or it does not fire. In case the rule does
+not fire, we have again two options: either its conditions are not met,
+or its target is the only remaining analysis.
+
+We express these requirements as a \emph{Boolean satisfiability problem} (SAT).
+SAT problems consist of two components: a set of Boolean variables, and a set
+of clauses on those variables. For instance, let the set be $\{a,b\}$
+and the formulas $\{a \vee b, \neg{}a\}$. A program called \emph{SAT
+  solver} will try to find a solution, where all the variables have a
+value. For this particular problem, the unique solution is $a=False,
+b=True$. It is also possible for a SAT problem to have no solution, or
+multiple solutions.
+
+%these possibilities as SAT variables, and ask for a solution: ``after all
+%rules in $R$ have applied, is there an input that will make $r'$ fire?'' 
+%To our knowledge, there is no other CG tool that would keep track of
+%previously executed rules
 

### 2016-02-24 paper/lrec/implementation.tex
more text

--- 2016-02-12 paper/lrec/implementation.tex
@@
 In this section, we describe the implementation of the tool.
-The SAT-encoding we use is similar to the one introduced in \cite{listenmaa_claessen2015}, with one key difference: in this paper, we operate on {\em symbolic sentences} instead of concrete sentences from a corpus.
+The SAT-encoding we use is similar to the one introduced in \newcite{listenmaa_claessen2015}, with one key difference: in this paper, we operate on {\em symbolic sentences} instead of concrete sentences from a corpus.
 The idea is that the SAT-solver is going to find the concrete sentence for us.
@@
 
-If the context contains a \verb!*!, we may need to make an approximation of $w(R)$ which may result in false negatives later on in the analysis.
+If the context contains a \verb!*!,
+we may need to make an approximation of $w(R)$ which may result in false negatives later on in the analysis.
+%we create sentences that are up to 3 words wider than the context without \verb!*!, and try them all.
 
@@
 
-\paragraph{Choosing the readings}
-
-Earlier we have shown an example with 5 readings (\texttt{det def}, \texttt{noun sg}, ...).
-In a realistic case, we operate between hundreds and thousands of readings.
-This is very much dependent on language: Dutch, the simplest language, has 336 readings.
-
-We expand an Apertium morphological lexicon, throw away the word forms and take all the analyses. An exception is when the grammar contains rules that target a specific lemma or word form.
-A simple solution is to retain the lemmas and word forms only for those entries where it is specified in the grammar, and otherwise leave them out.
-For example, the Dutch grammar contains the following rule:
-
-\begin{itemize}
- \item[] \texttt{REMOVE ("zijn"i vbser) IF (-1 (pr)) (1 Noun) ;}
-\end{itemize}
-
-This hints that there is something special about the verb \emph{zijn}, compared to the other verbs. Looking at the lexicon, we find \emph{zijn} in the following entries:
-
-\begin{itemize}
- \item[] \begin{verbatim}zijn:zijn<det><pos><mfn><pl>
-zijn:zijn<det><pos><mfn><sg>
-zijn:zijn<vbser><inf>
-zijn:zijn<vbser><pres><pl>
-\end{verbatim}
-\end{itemize}
-
-Thus we add special entries for these: in addition to the anonymous
- \texttt{<det><pos><mfn><pl>} reading, we add \texttt{<"zijn"><det><pos><mfn><pl>}.
-
- However, for languages with more readings, this may not be feasible.
-For instance, the Spanish lexicon has a high number of readings mostly because of the verbs: in addition to the verb forms, there are clitics attached to them.
-The number of verb readings without clitics is 213, and with clitics 1572.
-
- In our experience,  even ignoring the clitics doubles the amount of readings.
-
-Another note: the readings in grammar can be underspecified (e.g.
-\texttt{verb sg}), whereas the lexicon only gives us fully specified (\texttt{verb pres p2 sg}) readings.
-We tried a version where we took the tag combinations specified in the grammar as readings, and we could insert them into the symbolic sentences as well, but this was not an ideal solution: turns out that the tag lists in the grammars contain often errors (e.g.
-replacing OR with an AND; using a nonexistent tag; using a wrong level in a subreading), and if we accept those lists as readings, we will generate symbolic sentences that are impossible, and won't discover the bug in the grammar.
-
-However, if we only want to find rule interaction effects, then using the underspecified readings from the grammar makes the task faster, and it will still catch potential interaction errors.
-
-\paragraph{Creating realistic ambiguities}
-
-In the previous paragraph, we have created realistic \emph{readings}, by simply hardcoding legal tag combinations into variables.
-The next step in creating realistic ambiguities is to constrain what readings can go together.
-For instance, the case of \emph{zijn} shows us that ``determiner or verb'', is a possible ambiguity.
-In contrast, there is no word form in the lexicon that would be ambiguous between an adjective and a comma, hence we don't want to generate such ambiguity in our symbolic sentences.
-
-We solve the problem creating \emph{ambiguity classes}: groups of analyses that can be ambiguous with each other. 
-We represent the expanded morphological lexicon as a matrix, as seen in figure~\todo{make a nice picture with rows and columns}: word forms on the rows and analyses on the columns.
-Each distinct row forms an ambiguity class.
-For example, the ambiguity class that
-and form clusters based on which readings can be confused with each other. 
-%This is done after the previous step: we simply treat the lemma or word form as another tag. 
-Then we form SAT clauses that allow or prohibit certain combinations. These clauses will interact with the constraints created from the rules, and the end result will be closer to real-life sentences.
-
-Our approach is similar to \todo{cite Cutting, Kupiec et al.}, who use ambiguity classes instead of distinct word forms, in order to reduce the number of parameters in a Hidden Markov Model.
-They take advantage of the fact that they don't have to model ``bear'' and ``wish'' as separate entries, but they can just reduce it to ``word that can be ambiguous between noun and verb'', and use it as a parameter in their HMM.
-We can do a similar thing by saving a list of words with each ambiguity class.
-For example, we map the ambiguity class ``feminine plural noun or a separable verb infinitive'' to the list of word forms \{``uitgaven'', ``toespraken''\}, and then, if we generate such reading for our symbolic word, we can give one of these words as an example word.
-
-Now with this, we can detect even these kinds of rules as problematic:
-
- \begin{itemize}
- \item[] \begin{verbatim}
- REMOVE verb IF (-1 det) ;
- REMOVE noun IF (-1 det) ;
- \end{verbatim}
- \end{itemize}
-
- With our symbolic sentence, these rules will be no problem; to apply the latter, we only need to construct a target that has a realistic ambiguity with a noun; the verb is gone already.
- However, in the context of real life texts, the grammar writer might actually mean the following:
-
- \begin{itemize}
- \item[] \begin{verbatim}
- REMOVE verb IF (-1 det) (0 noun) ;
- REMOVE noun IF (-1 det) (0 verb) ;
- \end{verbatim}
- \end{itemize}
-
- In that case, those rules are contradictory, and we would like to find it out.
-The additional constraints will prevent the SAT solver from creating an ambiguity where noun is ambiguous with e.g.
-punctuation.
-
 \paragraph{Putting it all together}
@@
 
+\paragraph{Creating realistic readings}
+
+Earlier we have shown an example with 5 readings (``det def'', ``noun sg'', ...). In a realistic case, we operate between hundreds and thousands of readings. 
+%This is very much dependent on language: the simplest language we tested was Dutch, with 336 readings. The most complex was 
+In order to find the set of readings, we expand a morphological lexicon\footnote{We used the lexica from Apertium, found in \url{https://svn.code.sf.net/p/apertium/svn/languages/}.}, ignore the word forms and lemmas, and take all distinct analyses.
+However, many grammar rules target a specific lemma or word form.
+A simple solution is to retain the lemmas and word forms only for those entries where it is specified in the grammar, and otherwise leave them out.
+For example, the Dutch grammar contains the following rule:
+
+\begin{itemize}
+ \item[] \texttt{REMOVE ("zijn" vbser) IF (-1 Prep) (1 Noun) ;}
+\end{itemize}
+
+This hints that there is something special about the verb \emph{zijn}, compared to the other verbs. Looking at the lexicon, we find \emph{zijn} in the following entries:
+
+\begin{itemize}
+ \item[] 
+\begin{verbatim}zijn:zijn<det><pos><mfn><pl>
+zijn:zijn<det><pos><mfn><sg>
+zijn:zijn<vbser><inf>
+zijn:zijn<vbser><pres><pl>
+\end{verbatim}
+\end{itemize}
+
+Thus we add special entries for these: in addition to the anonymous
+ \texttt{<det><pos><mfn><pl>} reading, we add \texttt{<"zijn"><det><pos><mfn><pl>}. 
+The lemma is treated as just another tag.
+
+ However, for languages with more readings, this may not be feasible.
+For instance, Spanish has a high number of readings, not only because of many inflectional forms, but because it is possible to add 1--2 clitics to the verb forms.
+The number of verb readings without clitics is 213, and with clitics 1572.
+With the previously mentioned approach, we would have to double 1572 entries for each verb lemma. Even ignoring the clitics, each verb lemma would still result in 213 new readings.
+
+ In our experience,  even ignoring the clitics doubles the amount of readings.
+
+Another note: the readings in grammar can be underspecified (e.g.
+\texttt{verb sg}), whereas the lexicon only gives us fully specified (\texttt{verb pres p2 sg}) readings.
+We tried a version where we took the tag combinations specified in the grammar as readings, and we could insert them into the symbolic sentences as well, but this was not an ideal solution: turns out that the tag lists in the grammars contain often errors (e.g.
+replacing OR with an AND; using a nonexistent tag; using a wrong level in a subreading), and if we accept those lists as readings, we will generate symbolic sentences that are impossible, and won't discover the bug in the grammar.
+
+However, if we only want to find rule interaction effects, then using the underspecified readings from the grammar makes the task faster, and it will still catch potential interaction errors.
+
+\paragraph{Creating realistic ambiguities}
+
+In the previous section, we have created realistic \emph{readings}, by simply hardcoding legal tag combinations into variables.
+The next step in creating realistic ambiguities is to constrain what readings can go together.
+For instance, the case of \emph{zijn} shows us that ``determiner or verb'', is a possible ambiguity.
+In contrast, there is no word form in the lexicon that would be ambiguous between an adjective and a comma, hence we don't want to generate such ambiguity in our symbolic sentences.
+
+We solve the problem by creating \emph{ambiguity classes}: groups of analyses that can be ambiguous with each other. 
+We represent the expanded morphological lexicon as a matrix, as seen in figure~\todo{make a nice picture with rows and columns}: word forms on the rows and analyses on the columns.
+Each distinct row forms an ambiguity class.
+For example, the ambiguity class \verb![1257,496,16]!
+%TODO better description
+contains masculine plural adjectives and masculine plural past participles.
+Then we form SAT clauses that allow or prohibit certain combinations. These clauses will interact with the constraints created from the rules, and the end result will be closer to real-life sentences.
+
+Our approach is similar to \newcite{cutting_etal92}, who use ambiguity classes instead of distinct word forms, in order to reduce the number of parameters in a Hidden Markov Model.
+They take advantage of the fact that they don't have to model ``bear'' and ``wish'' as separate entries, but they can just reduce it to ``word that can be ambiguous between noun and verb'', and use it as a parameter in their HMM.
+We can do a similar thing by saving a list of words with each ambiguity class.
+For example, we map the ambiguity class ``feminine plural noun or a separable verb infinitive'' to the list of word forms \{``uitgaven'', ``toespraken''\}, and then, if we generate such reading for our symbolic word, we can give one of these words as an example word.
+
+There are two advantages of restricting the ambiguity within words.
+Firstly, we can create more realistic example sentences, which should help the grammar writer.
+Secondly, we can possibly detect some more conflicts. Assume that the grammar contains the following rules:
+
+ \begin{itemize}
+ \item[] 
+\begin{verbatim}
+ REMOVE adj IF (-1 aux) ;
+ REMOVE pp  IF (-1 aux) ;
+ \end{verbatim}
+ \end{itemize}
+
+ With our symbolic sentence, these rules will be no problem; to apply the latter, we only need to construct a target that has a realistic ambiguity with a past participle; the adjective will be gone already.
+However, it could be that past participles (pp) only ever get confused with adjectives---in that case, the above rules would be in conflict with each other.
+ By removing the adjective reading, the first rule selects the past participle reading, making it an instance of ``$R$ selects something in a context, $R'$ removes~it''. 
+The additional constraints will prevent the SAT-solver from creating an ambiguity outside the allowed classes, and such a case would be caught as a conflict.
+
 % \paragraph{Other questions we can ask}

### 2016-02-24 paper/lrec/problem_desc.tex
more text

@@
 POS tagging is a fundamental phase which is required in further syntactic or semantic analysis---hence a good POS-tagger is an important resource for any language.
-We believe that Constraint Grammar is a good candidate especially for languages with low resources, and in this paper we describe methods for CG writers and users to analyse and improve their grammars.
+We believe that Constraint Grammar is a good candidate especially for languages with low resources, and in this paper we describe methods for CG writer and users to analyse and improve their grammars.
 
@@
 
-In addition, there is a large body of research on automatically inducing rules, e.g. \cite{inducing_cg1996,lindberg_eineborg98ilp,lager01transformation,asfrent14} 
+In addition, there is a large body of research on automatically
+inducing rules, e.g. \newcite{inducing_cg1996}, \newcite{lindberg_eineborg98ilp},
+\newcite{lager01transformation} and \newcite{asfrent14}.
 However, since our work is aimed to aid the process of hand-crafting rules, we omit those works from our discussion.

### 2016-02-26 paper/lrec/eval_futurework.tex
better explanatio of DetPos(Not)Zijn example

@@
 \item[] 
-\begin{verbatim}SELECT DetPos IF (-1 Verb) (0 "zijn") 
-                 ( 1 Noun) ;
+\begin{verbatim}
 SELECT DetPosNotZijn IF (1 Noun) ;
+SELECT DetPos IF (-1 (vbser pres p3 sg)) 
+                 (0 "zijn") (1 Noun);
 \end{verbatim}
@@
 
-On the first glance it seems that there is no conflict: the first rule requires a Det Pos which is at the same time ``zijn'', as seen by the requirement \texttt{(0 "zijn")}.
-However, this has a different semantics in VISL CG-3, which is also repeated in our tool.
-The rules match a situation where there is \emph{any} Det Pos and \emph{any} zijn, including the verb.
-The correct way to formulate the first rule would be the following:
+The condition of the second rule cannot be true: there can never be a ``zijn'' in the target, because the previous rule has selected everything that is a possessive determiner and not ``zijn''. 
+%Not firing is not an option: if conditions of the first rule do not hold, then they will not hold for the second either. If the target is the only left, then 
+% "target" in case of SELECT means complement of the mentioned target. so 
+%    some of [adj, det pos zijn, verb zijn, punct, ...]  is true
+%    all of  [det pos f sg, det pos m pl, det pos foo bar] are false.
+% ie. there has been a rule that has removed detPosNotZijn, or they haven't been there in the first place.
+%
+
+Another remark on this formulation of a rule. 
+As per VISL CG-3, the condition \texttt{(0 "zijn")} does not require that the input would have determiner and ``zijn'' in the same reading: just that there is any determiner and any ``zijn'' in the cohort.
+The first example of a rule conflict had a similar requirement, but used correctly: the intended semantics was indeed that the adjective and the adverb are in different readings.
+However, in this case, the correct way to express the target is \texttt{SELECT DetPos + "zijn"}.
+
+Can we catch this additional problem with our tool? The example word constructed by the SAT-solver may as likely be one of the following:
 
 \begin{itemize}
-\item[] \texttt{SELECT DetPos + "zijn" IF (-1 Verb) (1 Noun) ;}
+\item[a.] \begin{verbatim}
+"w2"
+    w2<det><pos><f><sg>
+    w2<vbser><inf><"zijn">
+\end{verbatim}
+
+\item[b.] \begin{verbatim}
+"w2"
+    w2<det><pos><mfn><pl><"zijn">
+\end{verbatim}
 \end{itemize}
+
+In the latter case, the grammar writer would hopefully notice the imprecise definition and change their definition.
+However, this is more of a happy side effect than intended feature---the grammar writer cannot count on our tool to notice all similar cases.
+We could add a feature that alerts in all cases where a condition with 0 is used, or even automatically construct a version of the rule that moves whatever tags in the 0-condition into the target.
 

### 2016-02-26 paper/lrec/eval_futurework.tex
remove DetPos(Not)Zijn example from conflicts

@@
 \begin{tabular}{|l|l|l|l|}
+
 \hline
@@
 \# readings   & 336             & 2191     & 2012    \\ \hline
-\# conflicts  & 8 (1 + 5 + 2)   & 1        & 51    \\ \hline
-\clock{} all rules       & 7s              & 3min 22s    & 3h 26min    \\ \hline
-\clock{} last rule       & 4s              & 7s          & 3min 40s    \\ \hline
-% This is with just 2000ish variables for Finnish, below is with 5700
+\# conflicts  & 7 (1 + 5 + 1)   & 1        & 51    \\ \hline
+\clock{} all rules       & ??s              & ??min ??s    & ?h ??min    \\ \hline
+\clock{} last rule       & ??s              & ??s          & ?min ??s    \\ \hline
 
-% \clock{} last rule       & 4s              & 7s          & 30min 25s    \\ \hline
-% \clock{} longest rule    & ?s              & ?s                & ?min     \\ \hline
 %\# variables: last rule & 33915         & 247153            &    \\ \hline
@@
 \end{tabular}
-\caption{}
+\caption{\todo{make more tables}}
 \label{table:res}
@@
 
-\paragraph{Dutch} The Dutch grammar had two kinds of errors: rule-internal due to nonexistent tags, and rule interaction.
+\paragraph{Dutch} The Dutch grammar had two kinds of errors: rule-internal, and rule interaction.
 As for rule-internal conflicts, one was due to a misspelling in the list definition for personal pronouns, which rendered 5 rules ineffective.
@@
 
-There were two genuine conflicts with rule interaction. The first one is shown below:
+There was one genuine conflict with rule interaction, shown below:
 
@@
 
-The first one is more straight-forward: two rules, which both remove an adverb, were in an order where the first has more broad condition: remove adverb if followed by a noun.
+Two rules, which both remove an adverb, were in an order where the first has more broad condition: remove adverb if followed by a noun.
 In contrast, the second rule has a stricter requirement: only remove adverb if it is preceded by a determiner; the adverb itself is ambiguous with an adjective, and followed by a noun.
@@
 
-The second conflict, shown below, is more subtle.
+%%% Nevermind, this wasn't a conflict, my program was just buggy :-P
+%This is not a conflict, just pointing out an example how the SAT-solver constructs the only possible sentence where this works.
+
+%\begin{itemize}
+%\item[] 
+%\begin{verbatim}
+%SELECT DetPosNotZijn IF (1 Noun) ;
+%SELECT DetPos IF (-1 (vbser pres p3 sg)) 
+%                 (0 "zijn") (1 Noun);
+%\end{verbatim}
+%\end{itemize}
+
+% SELECT DetPosNotZijn IF (1 Noun);
+%If the first rule fires, then the condition of the second rule cannot be true: there can never be a ``zijn'' in the target, because the previous rule has selected everything that is a possessive determiner and not ``zijn''.
+%So, the first rule must not fire. It cannot be because condition does not hold: the condition (1 Noun) is shared with the second rule, and
+
+We also tested rules individually, in a way that a grammar writer might use our tool when writing new rules.
+The following rule was one of them:
 
@@
 \item[] 
-\begin{verbatim}
-SELECT DetPosNotZijn IF (1 Noun) ;
-SELECT DetPos IF (-1 (vbser pres p3 sg)) 
-                 (0 "zijn") (1 Noun);
-\end{verbatim}
+\texttt{SELECT DetPos IF (-1 (vbser pres p3 sg)) (0 "zijn") (1 Noun);}
 \end{itemize}
 
-The condition of the second rule cannot be true: there can never be a ``zijn'' in the target, because the previous rule has selected everything that is a possessive determiner and not ``zijn''. 
-%Not firing is not an option: if conditions of the first rule do not hold, then they will not hold for the second either. If the target is the only left, then 
-% "target" in case of SELECT means complement of the mentioned target. so 
-%    some of [adj, det pos zijn, verb zijn, punct, ...]  is true
-%    all of  [det pos f sg, det pos m pl, det pos foo bar] are false.
-% ie. there has been a rule that has removed detPosNotZijn, or they haven't been there in the first place.
-%
+As per VISL CG-3, the condition \texttt{(0 "zijn")} does not require that the input would have determiner and ``zijn'' in the same reading: just that there is any determiner and any ``zijn'' in the cohort.
 
-Another remark on this formulation of a rule. 
-As per VISL CG-3, the condition \texttt{(0 "zijn")} does not require that the input would have determiner and ``zijn'' in the same reading: just that there is any determiner and any ``zijn'' in the cohort.
-The first example of a rule conflict had a similar requirement, but used correctly: the intended semantics was indeed that the adjective and the adverb are in different readings.
-However, in this case, the correct way to express the target is \texttt{SELECT DetPos + "zijn"}.
+\todo{Find a better example of the imprecise 0 condition --- the ambiguity class thing would eliminate this one!}
 
-Can we catch this additional problem with our tool? The example word constructed by the SAT-solver may as likely be one of the following:
+Can we catch this imprecise formulation with our tool? The example word constructed by the SAT-solver may as likely be one of the following:
 
@@
     w2<det><pos><mfn><pl><"zijn">
+    w2<anything>
 \end{verbatim}
@@
 
+Incidentally, the first example of a rule conflict had a similar requirement, but used correctly: the intended semantics was indeed that the adjective and the adverb are in different readings.
+However, in this case, the correct way to express the target is \texttt{SELECT DetPos + "zijn"}.
+
 In the latter case, the grammar writer would hopefully notice the imprecise definition and change their definition.
 However, this is more of a happy side effect than intended feature---the grammar writer cannot count on our tool to notice all similar cases.
-We could add a feature that alerts in all cases where a condition with 0 is used, or even automatically construct a version of the rule that moves whatever tags in the 0-condition into the target.
+This kind of confusion is found in other rules as well. 
+To catch them more systematically, we could add a feature that alerts in all cases where a condition with 0 is used, and automatically construct a version of the rule that moves whatever tags in the 0-condition into the target, then asks the user which one was meant.
 

### 2016-02-26 paper/lrec/eval_futurework.tex
add some spanish examples

@@
 
-This is another, perhaps counter-intuitive instance of more general rule applying before more specific.
+This is another instance of more general rule applying before more specific.
 Negated context works as well to make a rule stricter.
@@
 The program doesn't currently point them out to the user, but that is a trivial addition; just like rule-internal conflicts, it doesn't require any semantic analysis.
-Another error, which was not pointed by the program, but discovered
+Another error, which was not pointed by the program, but discovered when we tried to extract the readings from the grammar and use them
+
+As with the Dutch grammar, we ran the tool on individual rules and examined the sequences that were generated. None of the following was marked as a conflict, but looking at the output indicated that
+
+\begin{itemize}
+\item[] 
+\begin{verbatim}
+REMOVE Vblex IF (0 "como") (*-1 Vblex BARRIER Cnj_Rel) ;
+REMOVE Vblex IF (0 "como") (*1 Vblex BARRIER Cnj_Rel) ;
+\end{verbatim}
+\end{itemize}
+
+Here is something that the ambiguity class constraint should find out.
+The grammar writer wants to disambiguate the word form ``como'', which can be either a preposition, adverb or the first person singular of the verb ``comer''. 
+However, the rule here addresses the lemma ``como'' instead of the word form, hence the SAT-solver cannot create a solution where \texttt{vblex} and ``como'' are in the same reading.
+The automated 0-to-target conversion could come in handy here too.
+
+Side note: if this rule was written as \texttt{Vblex + "como"}, it would have been caught as a conflicting tag combination.
+``Como'' as a lemma does not occur with \texttt{vblex}; it is only one word form which is ambiguous.
+
+\begin{itemize}
+\item[] 
+\begin{verbatim}
+REMOVE Podar  IF (0 Podar) (1 Adv) (2C Inf OR Enc) ;
+REMOVE Sentar IF (0 Sentar) (1C Inf OR Enc) ;
+\end{verbatim}
+\end{itemize}
+
+The comments make it clear that the first rule is meant to disambiguate between poder and podar, and the second rule between sentir and sentar.
+But the rule does not mention anything about ``poder'' nor ``sentir''; the SAT-solver gave a triggering sequence without
+
+This is actually a shortcoming of our way of handling the lemmas.
+If the lemmas and word forms were integrated to the analyses in a more feasible way, then the ambiguity class constraints would create a sequence where ``podar'' was ambiguous with ``poder'', instead of randomly chosen readings.
+So, even though this was not marked as a conflict, we can say that the weird output of the tool was unnecessary, and by better implementation, the tool would have created more realistic example sequence.
 

### 2016-02-28 paper/lrec/eval_futurework.tex
add more tables; add text about finnish

@@
 \hline
-              & \textsc{nld}  & \textsc{spa}  & \textsc{fin}  \\ \hline
-\# rules      & 59              & 279      & 1185     \\ \hline
-\# readings   & 336             & 2191     & 2012    \\ \hline
-\# conflicts  & 7 (1 + 5 + 1)   & 1        & 51    \\ \hline
-\clock{} all rules       & ??s              & ??min ??s    & ?h ??min    \\ \hline
+                   & \textsc{nld}  & \textsc{spa}lem  & \textsc{spa}nolem \\ \hline
+\# rules           & 59            & 279       & 279     \\ \hline
+\# readings        & 336           & 3905      & 1735    \\ \hline
+\# conflicts       & 7             & n         & m    \\ \hline
+\clock{} all rules & ??s           & + 1h      & 30m-1h   \\ \hline
 \clock{} last rule & ??s           & ??s       & ?min ??s    \\ \hline
+\clock{} all rules 
+without ambiguity 
+classes            & 10ish sec       & ??min ??s    & ?h ??min    \\ \hline
 
@@
 \label{table:res}
+\end{table}
+
+\begin{table}[]
+\centering
+\begin{tabular}{|l|l|l|l|}
+
+\hline
+              & n clitics & m clitics & only underspecified \\ \hline
+\# rules      & 1185      & 1185      & 1185\\ \hline
+\# readings   & many & more & only like <2000   \\ \hline
+\# conflicts  & n   & m        & super many    \\ \hline
+\clock{} all rules       & ??s              & ??min ??s    & ?h ??min    \\ \hline
+\clock{} last rule       & ??s              & ??s          & ?min ??s    \\ \hline
+
+\end{tabular}
+\caption{Special table for Finnish}
+\label{table:resFin}
 \end{table}
@@
 
-\paragraph{Finnish} \todo{Rerun the tests and rewrite this whole section!} For the Finnish grammar, we did not have time to investigate all the
+\paragraph{Finnish} 
+\todo{Rerun the tests and rewrite this whole section!}
+
+The results for the Finnish grammar are shown separately, in table~\ref{table:resFin}.
+We were not able to use the method with ambiguity classes at all---expanding the Finnish morphological lexicon results in \todo{100s of gigabytes?} of word forms, which is simply too big for our method to work.
+For future development, we will see if it is possible to manipulate the finite automata directly, instead of relying on the output in text.
+
+In addition to the previous problem, we had a large number of readings.
+In order to make the test run faster, we tried a few shortcuts.
+First, we restricted the amount of clitics per word form, and even ignored all the specified readings
+
+For the Finnish grammar, we did not have time to investigate all the
 results. We could verify approximately \nicefrac{1}{3} as true positives and

### 2016-02-29 paper/lrec/eval_futurework.tex
add explanation why we didn't evaluate the performance of the original vs. cleaned-up grammar

@@
 types introduced in CG-3, and only tested \textsc{remove} and \textsc{select} rules.
-The results are shown in table~\ref{table:res}.
+The results for Dutch and Spanish are shown in table~\ref{table:res},
+and the results for Finnish in table~\ref{table:resFin}.
+
+A natural follow-up evaluation would be to compare the performance of the
+grammar in the original state, and after removing the conflicts found
+by our tool. Unfortunately, we were not able to perform such
+evaluation, due to the lack of available gold standard corpora for all
+the languages.
 

### 2016-02-29 paper/lrec/lrec2016_listenmaa_claessen.tex
right path for cg.bib
small changes to abstract

@@
 \abstract{
-We describe a method for analysing Constraint Grammars by encoding the
-rules in SAT.
-Our tools can detect internal conflicts or redundancies in a grammar,
+We describe a method for analysing and improving Constraint Grammars (CG).
+% by encoding the rules in SAT.
+Traditional CG engines execute the rules in an imperative and
+sequential manner, which makes it impossible to track interaction
+between rules.
+In contrast, we represent CG rules as logic constraints, which
+can then be tested for contradictions and inconsistencies.
+Our tool can detect internal conflicts or redundancies in a grammar,
 as well as generate examples to demonstrate the effect of some rule or
@@
 \bibliographystyle{lrec2016}
-\bibliography{cg}
+\bibliography{../cg}
 

### 2016-02-29 paper/lrec/problem_desc.tex
small changes to the intro to SAT

@@
 is a formalism used to disambiguate morphologically analysed text. 
-A grammar consists of rules that target specific analyses for selection or removal, based on contextual tests. For example, the following rule
+A grammar consists of rules that target specific readings for selection or removal, based on contextual tests. For example, the following rule
 \begin{itemize}
@@
 \end{itemize}
-removes a verb reading from a word which is preceded by a determiner.
+removes all verb readings from a word which is preceded by a determiner.
 Given the following text,
@@
 \end{itemize}
-the rule will match to the word \emph{wish}, and remove the analyses \texttt{verb pl} and \texttt{verb inf}.
-However, if the target word has only one remaining analysis, then the rule will not apply, even if the condition is met.
+the rule will match to the word \emph{wish}, and remove the readings
+\texttt{"wish" verb pl} and \texttt{"wish" verb inf}.
+However, if the target word has only one remaining reading, then the
+rule will not apply, even if the condition is met.
 
@@
 resourced languages. They are robust and can be written without large
-corpora---only morphological analysis is needed. The formalism is
+corpora---only morphological analyser is needed. The formalism is
 lightweight and language-independent, and resources can be shared
 between related languages \cite{bick2006spanish,lene_trond_linda2010}.
-Mature CGs contain some thousands of rules, but even small CGs are shown to be effective \cite{lene_trond2011}.
+Mature CGs contain some thousands of rules, but even small CGs are
+shown to be effective \cite{lene_trond2011}.
 
@@
 gets harder to keep track of all the rules and their interaction.
-Our tools will help grammar writers and users to find conflicting
+Our tool will help grammar writers and users to find conflicting
 rules, diagnose problems and improve their grammars. 
@@
 
-For the second use case, here are examples of conflicts that our tools will detect.
+For the second use case, here are examples of conflicts that our tool will detect.
 \begin{itemize}
 \item If a rule appears twice, the second occurrence will be disabled by the first
-\item $R$ selects something in a context, $R'$ removes~it
-\item $R$ removes something from the context of $R'$, so $R'$ can never
+\item $R$ selects something in a context, $r$ removes~it
+\item $R$ removes something from the context of $r$, so $r$ can never
   apply
-\item $R$ has an internal conflict, such as non-existent
+\item $r$ has an internal conflict, such as non-existent
 tag combination, or contradictory requirements for a context word
 \end{itemize}
-$R$ can also be a set of rules: for instance, if one rule removes a verb in
+$R$ can be one rule or a set of rules: for instance, if one rule removes a verb in
 context $C$, and another in context $\neg C$, together these rules
@@
  syntactic} analysis.
-In order to find effects of rule interaction, we must keep track of
-the possible sentences at each step. After each rule, we have two
-possibilities: the rule fires, or it does not fire. In case the rule does
-not fire, we have again two options: either its conditions are not met,
-or its target is the only remaining analysis.
+% In order to find effects of rule interaction, we cannot just apply the
+% rules to some actual sentences and hope to find a conflict---corpora
+% can be extremely valuable in the development of CG rules, but there
+% are languages where corpora are not available, or they may contain
+% text from limited domains. 
+We must keep track of all the possible sentences after applying each
+rule. At each step, we have two options: either the rule fires,
+or it does not fire. In case the rule does not fire, we have again two
+options: either its conditions are not met,
+or its target is the only remaining analysis. This results in a complex table
+of interdependent decisions---some determined by being the target of a
+rule, others by being a condition.
 
-We express these requirements as a \emph{Boolean satisfiability problem} (SAT).
+We express these constraints as a \emph{Boolean satisfiability problem} (SAT).
 SAT problems consist of two components: a set of Boolean variables, and a set

### 2016-02-29 paper/lrec/problem_desc.tex
more text in intro & add label for conclusions

@@
 \item Are all the rules distinct? (e.g. \texttt{Para} and \texttt{De} may be included in \texttt{Prep})
-\item Could two or more rules be merged? (e.g. \texttt{SELECT Inf IF -1 Prep OR Vai OR Vbmod ...})
+\item Can two or more rules be merged? (e.g. \texttt{SELECT Inf IF (-1C Prep OR Vai OR Vbmod ...)})
 \item What is the best order for the rules?
-\item Generate a sequence that triggers rule(s) $R$ but not rule(s) $R'$. 
+\item Generate a sequence that triggers rules $R$ but not rules $R'$. 
 \end{itemize}
@@
 \item If a rule appears twice, the second occurrence will be disabled by the first
-\item $R$ selects something in a context, $r$ removes~it
+\item $r$ selects something in a context, $r'$ removes~it
 \item $R$ removes something from the context of $r$, so $r$ can never
@@
 \end{itemize}
-$R$ can be one rule or a set of rules: for instance, if one rule removes a verb in
+$R$ can be a single rule or a set of rules: for instance, if one rule removes a verb in
 context $C$, and another in context $\neg C$, together these rules
@@
 rule. At each step, we have two options: either the rule fires,
-or it does not fire. In case the rule does not fire, we have again two
-options: either its conditions are not met,
-or its target is the only remaining analysis. This results in a complex table
+or it does not fire. In case the rule does not fire, we have two
+reasons why not: one or more of its conditions does not hold, or its
+target is the only remaining analysis. The result is a complex table
 of interdependent decisions---some determined by being the target of a
@@
 We express these constraints as a \emph{Boolean satisfiability problem} (SAT).
-SAT problems consist of two components: a set of Boolean variables, and a set
-of clauses on those variables. For instance, let the set be $\{a,b\}$
-and the formulas $\{a \vee b, \neg{}a\}$. A program called \emph{SAT
-  solver} will try to find a solution, where all the variables have a
-value. For this particular problem, the unique solution is $a=False,
-b=True$. It is also possible for a SAT problem to have no solution, or
-multiple solutions.
+A SAT-problem consists of two components: a set of Boolean variables, and a set
+of clauses on those variables. For instance, let the set ov variables
+be $\{a, b\}$ and the formulas $\{a \vee b, \neg{}a\}$. A program called
+\emph{SAT-solver} will try to find a solution, where all the variables
+are replaced by a Boolean value. For this particular problem, the
+unique solution is $\{a=False, b=True\}$, but it is also possible for a
+SAT-problem to have no solution or multiple solutions.
+In the case of CG analysis, the model we build is a \emph{sentence},
+which starts as potentially anything, and is being shaped by all the
+rules along the way. This is how we can generate possible inputs, as
+well as check if the grammar is internally consistent.
+
+The paper is structured as follows. Section~\ref{sec:prev} relates our
+work to the previous research. 
+Section~\ref{sec:implementation} discusses the implementation more in depth, and
+Section~\ref{sec:eval} presents preliminary results. 
+Section~\ref{sec:conclusion} discusses future work and concludes the paper.
 
@@
 
-%The paper is structured as follows. Section~\ref{sec:prev} relates our work to the previous research. 
 % Section~\ref{sec:implementation} discusses the implementation, 

### 2016-03-01 paper/lrec/problem_desc.tex
add another example of a use case

@@
 \begin{figure}[t]
-\begin{center}
+\centering
 
@@
 \caption{Rules to select infinitive in Portuguese.}
-\end{center}
 
-\label{infrules}
+\label{fig:infrules}
 \end{figure}
 
-Given the rules in figure~\ref{infrules}, a grammar writer may ask the following questions while writing a grammar.
+Given the rules in figures~\ref{fig:infrules}~and~\ref{fig:regroup}, a
+grammar writer may ask the following questions while writing a grammar.
 
@@
 \item Are all the rules distinct? (e.g. \texttt{Para} and \texttt{De} may be included in \texttt{Prep})
-\item Can two or more rules be merged? (e.g. \texttt{SELECT Inf IF (-1C Prep OR Vai OR Vbmod ...)})
+\item Can two or more rules be merged? (e.g. \texttt{SELECT Inf IF
+    (-1C Prep OR Vai OR Vbmod)})
+\item Does my neat rewrite of a messy rule preserve the meaning?
 \item What is the best order for the rules?
-\item Generate a sequence that triggers rules $R$ but not rules $R'$. 
+\item Generate a sequence that triggers rules $R$ but not rules $R'$ 
 \end{itemize}
@@
 Section~\ref{sec:conclusion} discusses future work and concludes the paper.
+
+\begin{figure}[t]
+\centering
+\begin{itemize}
+
+\item[]\texttt{SELECT V + Prs/Imprt + Act + Neg}
+
+\item[\texttt{IF}]
+\begin{verbatim}(*-1C Negv LINK NOT *1 Vfin/PrsPrc/Inf) 
+(NOT 0 N) (NOT 0 Pron) 
+(NOT *1 Neg) (NOT *-1 Neg)
+(NOT 0 Pass) (NOT *-1 Niin) 
+(*-1C Negv LINK NOT *1 CLB?) 
+(*-1C Negv LINK NOT 0 Imprt) ; 
+\end{verbatim}
+
+\item[\texttt{IF}]
+\begin{verbatim}(NOT *-1 Niin OR Neg)  
+(*-1C Negv
+  LINK NOT 0 Imprt
+  LINK NOT *1 Vfin/PrsPrc/Inf OR CLB?) 
+(NOT 0 N OR Pron OR Pass) 
+(NOT *1 Neg) ;
+\end{verbatim}
+\end{itemize}
+
+\caption{Two versions of a condition in Finnish.}
+
+\label{fig:regroup}
+\end{figure}
 

### 2016-03-01 paper/lrec/eval_futurework.tex
more text in intro & add label for conclusions
add another new feature; polish text a bit, actual content still TBD

@@
 \clock{} all rules & ??s           & + 1h      & 30m-1h   \\ \hline
-\clock{} last rule & ??s           & ??s       & ?min ??s    \\ \hline
-\clock{} all rules 
-without ambiguity 
+% \clock{} last rule & ??s           & ??s       & ?min ??s    \\ \hline
+\clock{} no amb. 
 classes            & 10ish sec       & ??min ??s    & ?h ??min    \\ \hline
@@
 
-\paragraph{Spanish} The Spanish grammar had proportionately the highest number of errors.\footnote{TODO://link/to/error.log} \todo{confirm; I assume this is gonna be so, but let's see Fin and Por}.
+\paragraph{Spanish} The Spanish grammar had proportionately the highest number of errors.
+\todo{confirm; I assume this is gonna be so, but let's see Fin and Por}.
 The grammar we ran is like the one found in the Apertium repository, apart from 2 changes: we fixed some typos (capital O for 0) in order to make it compile, and commented out two rules that used regular expressions, because our solution does not handle them properly (see more on Future work/Implementation).
@@
 
-The Spanish grammar has a high number of interaction conflicts, all with the same pattern:
+The Spanish grammar has a high number of interaction conflicts, \todo{n} with the same pattern:
 
@@
 
-This is another instance of more general rule applying before more specific.
-Negated context works as well to make a rule stricter.
-The second rule will never apply, because the first will catch all cases.
+This is another instance of more general rule applying before more
+specific. Negated context works as well to make a rule stricter. The
+second rule will never apply, because the first will catch all
+cases. For a full list of found conflicts, see \todo{\url{actual://link.here/spanish/error.log}}
 
-In addition, there was a number of definitions that were never used.
-The program doesn't currently point them out to the user, but that is a trivial addition; just like rule-internal conflicts, it doesn't require any semantic analysis.
+In addition, there was a number of definitions that were never
+used. The program doesn't currently point them out to the user, but
+that is a trivial addition; just like rule-internal conflicts, it does
+not require any semantic analysis. 
 Another error, which was not pointed by the program, but discovered when we tried to extract the readings from the grammar and use them
 
-As with the Dutch grammar, we ran the tool on individual rules and examined the sequences that were generated. None of the following was marked as a conflict, but looking at the output indicated that
+As with the Dutch grammar, we ran the tool on individual rules and
+examined the sequences that were generated. None of the following was
+marked as a conflict, but looking at the output indicated that there
+are multiple interpretations, such as whether two analyses for a
+context word should be in the same reading or different readings.
+We observed also cases where the grammar writer has specified desired
+behaviour in comments, but the rule does not do what the grammar
+writer intended.
 
-\begin{itemize}
-\item[] 
-\begin{verbatim}
-REMOVE Vblex IF (0 "como") (*-1 Vblex BARRIER Cnj_Rel) ;
-REMOVE Vblex IF (0 "como") (*1 Vblex BARRIER Cnj_Rel) ;
-\end{verbatim}
-\end{itemize}
+% \begin{itemize}
+% \item[] 
+% \begin{verbatim}
+% REMOVE Vblex IF (0 "como") (*-1 Vblex BARRIER Cnj_Rel) ;
+% \end{verbatim}
+% \end{itemize}
 
-Here is something that the ambiguity class constraint should find out.
-The grammar writer wants to disambiguate the word form ``como'', which can be either a preposition, adverb or the first person singular of the verb ``comer''. 
-However, the rule here addresses the lemma ``como'' instead of the word form, hence the SAT-solver cannot create a solution where \texttt{vblex} and ``como'' are in the same reading.
-The automated 0-to-target conversion could come in handy here too.
+% Based on the comments in the file, the grammar writer wants to
+% disambiguate the word form ``como'', which can be either a
+% preposition, adverb or the first person singular of the verb
+% ``comer''. 
+% However, the rule here addresses the \emph{lemma} ``como'' instead of the
+% word form, hence the SAT-solver cannot create a solution where
+% \texttt{vblex} and ``como'' as a lemma are in the same
+% reading. 
+% A precise definition for this rule would target \texttt{vblex "<como>"} 
+% or \texttt{vblex "comer"}, so that it becomes obvious that those two
+% are supposed to be in the same reading.
 
-Side note: if this rule was written as \texttt{Vblex + "como"}, it would have been caught as a conflicting tag combination.
-``Como'' as a lemma does not occur with \texttt{vblex}; it is only one word form which is ambiguous.
+% Currently, as all the word forms are allowed to combine with anything,
+% the ambiguity class constraint does not point this
+% The automated 0-to-target conversion could come in handy here too.
+% However, if the
+
+% TODO An example of the same but in condition. that the ambiguity class constraint, when
+% implemented properly, should find out.
 
@@
 
-For the Finnish grammar, we did not have time to investigate all the
-results. We could verify approximately \nicefrac{1}{3} as true positives and
-\nicefrac{1}{3} as false positives.
-The verified true positives were all rule-internal conflicts. 
-For example, the following rule requires the target to be genitive and unambiguously nominative: 
-\begin{itemize}
-\item[]\texttt{"oma" SELECT GEN IF [...] (0C NOM) ;}
-\end{itemize}
-We also detected a number of tagset errors;
+The Finnish grammar was the largest and most complex of all the
+grammars we have tested.
+
 the grammar was originally written in 1995, and converted to the
@@
 
-%  For example, the Omorfi tagset
-% marks a possessive suffix with \texttt{Px}, e.g. \texttt{PxPl2} for second person plural, but the
-% grammar uses the old convention \texttt{PL2} in the rules.
-% As a result, 15 rules with possessive suffixes were detected as
-% impossible.
 To improve performance, we ignored the fully specified readings from the
@@
 
-\section{Future work}
+\section{Future work and conclusions}
+\label{sec:conclusion}
+
+The evaluation indicates that the tool finds cases that are not
+obvious to the human eye.
+
+One additional feature is to suggest reformattings for a rule. Recall
+figure~\ref{fig:regroup} from the introduction; in that case, the
+original rule was written by the original author, and another
+grammarian thought that the latter form is nicer to read. Doing the
+reverse operation could also be possible. If a rule with long
+disjunctions conflicts, it may be useful to split it into smaller
+conditions, and eliminate one at a time, in order to find the
+reason(s) for the conflict.
 

### 2016-03-02 CG_Symbolic.hs
fix Eq and Ord instances for subreadings; now Subr Wherever foo matches Subr _ foo

@@ -231,12 +231,11 @@ apply s allinds sentence rule = do
 
            condsHold <- orl' s cs
            let trgPos   = mapMaybe   (lu' word) trgIndsList
-
            let otherNeg = map (neg . lu word) (IS.toList otherInds)
 
            someTrgIsTrue <- orl' s trgPos 
            noOtherIsTrue <- andl' s otherNeg
-           onlyTrgLeft <- andl' s [someTrgIsTrue, noOtherIsTrue]
+           onlyTrgLeft <- andl' s [ someTrgIsTrue, noOtherIsTrue ]
            cannotApply <- orl' s [ neg condsHold, onlyTrgLeft ]
 
            newTrgLits <- sequence
@@ -394,6 +393,7 @@ mkTagMap ts tcs = M.fromList $
                    ts `for` \t -> let getInds = IS.fromList . map (1+) . findIndices (elem t)
                                   in (t, getInds tcs) 
 
+                            -- [[Tag]],[[Tag]]
 lookupTag :: TagMap -> WIndSet -> (Trg,Dif) -> (WIndSet,WIndSet)
 lookupTag alltags allinds (trg,dif) = 
   let trgInds = if trg==[[]] then allinds --this means IF (-1 (*))

### 2016-03-02 CG_base.hs

@@ -12,29 +12,45 @@ import Text.Regex
 data Tag = Tag String | Lem String | WF String | Subreading Subpos Tag |
            {-Rgx Regex String | -} EOS | BOS 
 
-data Subpos = FromStart Integer | FromEnd Integer | Wherever deriving (Eq)
+data Subpos = FromStart Integer | FromEnd Integer | Wherever
 
 instance Show Subpos where
   show (FromStart n) = show n
   show (FromEnd   n) = show (-n)
   show Wherever = "*"
 
+instance Eq Subpos where
+  Wherever    == anywhere    = True
+  anywhere    == Wherever    = True
+  FromStart n == FromStart m = n==m
+  FromEnd   n == FromEnd   m = n==m
+  _           == _           = False
+
+
 --because Regex has no Eq instance
 instance Eq Tag where
+  Subreading s t == Subreading s' t' = s==s' && t==t'
   foo == bar = show foo == show bar
 
 -- | Wordform should be first element in an analysis.
 instance Ord Tag where
-  WF word `compare` WF word' = word `compare` word'
-  WF _    `compare` _        = LT
-  _       `compare` WF _     = GT
-  Lem lem `compare` Lem lem' = lem `compare` lem'
-  Lem lem `compare` Tag tag  = GT
-  Tag tag `compare` Lem lem  = LT
+  BOS `compare` _   = GT
+  EOS `compare` BOS = LT
+  EOS `compare` _   = GT 
+  WF word `compare` WF word'  = word `compare` word'
+  WF _    `compare` _         = LT
+  _       `compare` WF _      = GT
+  Lem lem `compare` Lem lem'  = lem `compare` lem'
+  Lem lem `compare` Tag tag   = GT
+  Tag tag `compare` Lem lem   = LT
 
-  Tag tag `compare` Tag tag' = tag `compare` tag'
+  Tag tag `compare` Tag tag'  = tag `compare` tag'
 --  Rgx _ s `compare` Rgx _ s' = s `compare` s'
-  foo     `compare` bar      = show foo `compare` show bar
+  Subreading _ t `compare` 
+       Subreading _ t'        = t `compare` t'
+  Subreading _ t `compare` t' = t `compare` t'
+  t `compare` Subreading _ t' = t `compare` t'
+  foo `compare`bar = error ("compare: " ++ show foo ++ show bar)
 
 -- | Following the conventions of vislcg3
[…]

### 2016-03-02 AmbiguityClass.hs
what is all this uncommitted bullshit here, too lazy to do a specialised commit for all so just -a

@@ -36,7 +36,17 @@ test :: IO ()
 test =
   do let xss  = [[1,2,3],[2,3,4],[1,4],[5],[6,7],[3,7]]
          univ = S.toList . S.fromList . concat $ xss
-     
+
+     print univ
+
+     let acfile = "data/spa/spa-ambiguity-classes-hackylemmas"
+     ambclauses <- readFile acfile
+     let xss'  = map read (lines ambclauses) :: [[Int]]
+         univ' = S.toList . S.fromList . concat $ xss'
+     (univ'==univ') `seq` print univ'
+     print (length univ')
+     print $ [1..1638] \\ univ'
+
      -- compute the formula once
      putStrLn "Computing formula..."
      let p = formula xss
@@ -74,6 +84,12 @@ data Form
   | Or [Form]
   | NotTag Int
  deriving ( Eq, Ord, Show )
+
+formula' :: [[Int]] -> Int ->  Form
+formula' xss univsz = build yss
+ where
+  univ = [1..univsz]
+  yss  = S.toList . S.fromList $ [ univ \\ xs | xs <- xss ]
 
 formula :: [[Int]] -> Form
 formula xss = build yss
@@ -126,8 +142,9 @@ constraints s mp pre (Or [p]) =
 
 constraints s mp pre p =
   do qs <- asOr s mp p
-     --print (take 5 $ pre++qs)
      addClause s (pre ++ qs)
+     --print (take 5 $ pre++qs)
+     --(qs == qs) `seq` putStrLn "constraints: Done."
  where
   asOr s mp (And [])   = do return [true]
   asOr s mp (And [p])  = do asOr s mp p

### 2016-03-02 Matrix.hs

@@ -8,8 +8,8 @@ import qualified Data.IntSet as IS
 import System.Environment ( getArgs ) 
 
 
-import Debug.Trace
---trace str a = a
+--import Debug.Trace
+trace str a = a
 
 type Key = Int
 type ExampleWordForm = String
@@ -30,11 +30,11 @@ main = do
   []    -> error "give a lexicon and tag combinations"
   (w:t:_) -> do ws <- words `fmap` readFile w
                 tcs <- (map parseReadings . words) `fmap` readFile t
-                mapM_ print tcs
+                --mapM_ print tcs
                 let ts = concat tcs
                 let readings = map (toReadingMap tcs) ws 
-                mapM_ (print) $ M.toList $ toGraph ( readings)
-                --mapM_ print $ toGraph readings
+                --mapM_ (print) $ M.toList $ toGraph ( readings)
+                mapM_ print $ toGraph readings
 
 --------------------------------------------------------------------------------
 
@@ -44,8 +44,8 @@ toGraph readings = readingMap
   group' acc (a,b) = case M.lookup a acc of
                        Just c  -> M.adjust (b:) a acc 
                        Nothing -> M.insert a [b] acc
-  readingMap = invertMap 99999 $ foldl group' M.empty readings 
-  --readingMap = onlyValues 99999 $ foldl group' M.empty readings 
+  --readingMap = invertMap 99999 $ foldl group' M.empty readings 
+  readingMap = onlyValues 99999 $ foldl group' M.empty readings 
 
 
 --weet:weten<vblex><imp><sg> 
@@ -57,7 +57,7 @@ toGraph readings = readingMap
 toReadingMap :: [[Tag]] -> String -> (ExampleWordForm, Key)
 toReadingMap tcs str = trace (show (weet,key)) $ (weet, key)
  where
-  (weet, ':':weten_vblex_imp_sg) = break (==':') str
+  (w:eet, ':':weten_vblex_imp_sg) = break (==':') str
   (_weten,         vblex_imp_sg) = break (=='<') weten_vblex_imp_sg
   tags = parseReadings vblex_imp_sg
   key = case elemIndex tags tcs of

### 2016-03-02 executable/Analyse.hs

@@ -117,7 +117,7 @@ main = do
     let tags = nub $ tagsInLex ++ concat readings
 
     --print (length readings, length (filter (not.null) readings))
-    mapM_ print tags
+    --mapM_ print tags
 
     let tagmap = mkTagMap tags readings
     let allinds = IS.fromList [1..length readings]

### 2016-03-02 paper/lrec/eval_futurework.tex
little cleanup

@@
 
-\paragraph{Accuracy} 
 The experiments revealed problems in all grammars. For the smaller
@@
 \item[] 
-\begin{verbatim}REMOVE Adv IF (1 Noun) ;
-REMOVE Adv IF (-1 Det) (0 Adj) (1 Noun) ;
+\begin{verbatim}REMOVE Adv IF (1 N) ;
+REMOVE Adv IF (-1 Det) (0 Adj) (1 N) ;
 \end{verbatim}
@@
 
-As per VISL CG-3, the condition \texttt{(0 "zijn")} does not require that the input would have determiner and ``zijn'' in the same reading: just that there is any determiner and any ``zijn'' in the cohort.
-
-\todo{Find a better example of the imprecise 0 condition --- the ambiguity class thing would eliminate this one!}
-
+As per VISL CG-3, the condition \texttt{(0 "zijn")} does not require
+that the input would have determiner and ``zijn'' in the same reading:
+just that there is a reading with any determiner, and a reading with any ``zijn'' in the cohort. 
 Can we catch this imprecise formulation with our tool? The example word constructed by the SAT-solver may as likely be one of the following:
@@
     w2<det><pos><mfn><pl><"zijn">
-    w2<anything>
 \end{verbatim}
@@
 
-Incidentally, the first example of a rule conflict had a similar requirement, but used correctly: the intended semantics was indeed that the adjective and the adverb are in different readings.
-However, in this case, the correct way to express the target is \texttt{SELECT DetPos + "zijn"}.
+Incidentally, the first example of a rule conflict had a similar
+requirement, but used correctly: the intended semantics was indeed
+that the adjective and the adverb are in different readings. However,
+in this case, the precise way to express the target is \texttt{SELECT DetPos + "zijn"}.
 
-In the latter case, the grammar writer would hopefully notice the imprecise definition and change their definition.
+% In the latter case, the grammar writer would hopefully notice the imprecise definition and change their definition.
 However, this is more of a happy side effect than intended feature---the grammar writer cannot count on our tool to notice all similar cases.
-This kind of confusion is found in other rules as well. 
-To catch them more systematically, we could add a feature that alerts in all cases where a condition with 0 is used, and automatically construct a version of the rule that moves whatever tags in the 0-condition into the target, then asks the user which one was meant.
 
@@
 not require any semantic analysis. 
-Another error, which was not pointed by the program, but discovered when we tried to extract the readings from the grammar and use them
+We also discovered problems in some set definitions, when we tried to
+use the set definitions from the grammar directly as our readings.
+For instance, the following definition requires the word to be all of
+the listed parts of speech at the same time:
+\begin{itemize}
+\item[] 
+\texttt{SET NP\_Member = N + A + Det + PreAdv + Adv + Pron ;}
+\end{itemize}
+
+This particular definition was never used in the rules, so it would not
+have caused a conflict. We noticed it by accident, when the SAT-solver
+offered it in an example sequence that, meant for another rule.
 
@@
 \item[] 
-\begin{verbatim}
-REMOVE Podar  IF (0 Podar) (1 Adv) (2C Inf OR Enc) ;
-REMOVE Sentar IF (0 Sentar) (1C Inf OR Enc) ;
-\end{verbatim}
+\texttt{REMOVE Podar  IF (0 Podar) (1 Adv) (2C Inf OR Enc) ;} \\
+\texttt{REMOVE Sentar IF (0 Sentar) (1C Inf OR Enc) ;}
 \end{itemize}
@@
 
+We found the same kind of formulation in many other rules and grammars.
+To catch them more systematically, we could add a feature that alerts in all cases where a condition with 0 is used, and automatically construct a version of the rule that moves whatever tags in the 0-condition into the target, then asks the user which one was meant.
+
 \paragraph{Finnish}
+
 \todo{Rerun the tests and rewrite this whole section!}

### 2016-03-02 paper/lrec/implementation.tex
add small table about ambiguity classes
little cleanup

@@
 
+\begin{center}
+\begin{tabular}{c|c|c|c|c}
+
+            & n nt sg  & n f pl  & vblex sep inf & det pos mfn  \\ \hline
+uitgaven    & 0        & 1       & 1             & 0    \\ 
+toespraken  & 0        & 1       & 1             & 0    \\ 
+haar        & 1        & 0       & 0             & 1    \\
+
+\end{tabular}
+\end{center}
+
 We solve the problem by creating \emph{ambiguity classes}: groups of analyses that can be ambiguous with each other. 
-We represent the expanded morphological lexicon as a matrix, as seen in figure~\todo{make a nice picture with rows and columns}: word forms on the rows and analyses on the columns.
-Each distinct row forms an ambiguity class.
-For example, the ambiguity class \verb![1257,496,16]!
-%TODO better description
-contains masculine plural adjectives and masculine plural past participles.
+We represent the expanded morphological lexicon as a matrix, as seen
+above: word forms on the rows and analyses on the columns. Each
+distinct row forms an ambiguity class. For example, one class may
+contain words that are ambiguous between plural feminine nouns and
+separable verb infinitives; another contains masculine plural adjectives 
+and masculine plural past participles.
 Then we form SAT clauses that allow or prohibit certain combinations. These clauses will interact with the constraints created from the rules, and the end result will be closer to real-life sentences.

### 2016-03-02 paper/lrec/problem_desc.tex
minor tweaking so that all bullet points are on the same page
little cleanup

@@
     (-1C Prep OR Vai OR Vbmod)})
-\item Does my neat rewrite of a messy rule preserve the meaning?
+\item Can I rewrite a messy rule in a neater way without changing the meaning?
 \item What is the best order for the rules?
-\item Generate a sequence that triggers rules $R$ but not rules $R'$ 
+\item Generate a sequence that triggers $R$ but not $R'$ 
 \end{itemize}
@@
 work to the previous research. 
-Section~\ref{sec:implementation} discusses the implementation more in depth, and
+Section~\ref{sec:implementation} discusses the implementation, and
 Section~\ref{sec:eval} presents preliminary results. 

### 2016-03-02 paper/lrec/eval_futurework.tex
more cleanup

@@
 grammar in the original state, and after removing the conflicts found
-by our tool. Unfortunately, we were not able to perform such
-evaluation, due to the lack of available gold standard corpora for all
-the languages.
+by our tool. Unfortunately, we did not have time to perform such
+evaluation, and in addition, we only have gold standard corpus (20~000) words for Spanish.
 
@@
 % investigate all of them, but among those we could, we found both true and false negatives.
-We did not check for false negatives in any of the grammars. However, the test reported a high number of errors even in the smallest grammars, and some of them are intuitively not that bad; we
+We did not systematically check for false negatives in any of the
+grammars, but we kept track of a number of known tricky cases; mostly
+rules that had been false positives in earlier stages of the
+development, and other rules with similar patterns, as well as rules
+with negations.
+%While this does not guarantee the lack of false negatives, it indicates some degree of reliability. 
+As the tool matures and we add
+new features, a more in-depth analysis will be needed.
 
@@
 %%% Nevermind, this wasn't a conflict, my program was just buggy :-P
-%This is not a conflict, just pointing out an example how the SAT-solver constructs the only possible sentence where this works.
 
@@
 
+We found the same kind of formulation in many other rules and grammars.
+To catch them more systematically, we could add a feature that alerts in all cases where a condition with 0 is used, and automatically construct a version of the rule that moves whatever tags in the 0-condition into the target, then asks the user which one was meant.
+
 % \begin{itemize}
@@
 
-This is actually a shortcoming of our way of handling the lemmas.
-If the lemmas and word forms were integrated to the analyses in a more feasible way, then the ambiguity class constraints would create a sequence where ``podar'' was ambiguous with ``poder'', instead of randomly chosen readings.
+This example shows a shortcoming of our way of handling the lemmas. If
+the lemmas and word forms were integrated to the analyses in a more
+feasible way, then the ambiguity class constraints would create a
+sequence where ``podar'' was ambiguous with ``poder'', instead of randomly chosen readings.
 So, even though this was not marked as a conflict, we can say that the weird output of the tool was unnecessary, and by better implementation, the tool would have created more realistic example sequence.
-
-We found the same kind of formulation in many other rules and grammars.
-To catch them more systematically, we could add a feature that alerts in all cases where a condition with 0 is used, and automatically construct a version of the rule that moves whatever tags in the 0-condition into the target, then asks the user which one was meant.
 
@@
 
-the grammar was originally written in 1995, and converted to the
-Omorfi tagset by \cite{pirinen2015}, but there were 15 rules that used the old convention for possessive suffixes.
-
 To improve performance, we ignored the fully specified readings from the
@@
 The shortcut works most of the time, but it is possible to
-miss some cases: e.g. \texttt{SELECT PRON + REL IF (0 NOM)} 
-assumes ``pron rel nom'', but our method only gives
+miss some cases: e.g. \texttt{SELECT Pron + Rel IF (0 Nom)} 
+may assume ``pron rel nom'' in one reading, but our method only gives
 ``pron rel'' and ``nom'' separately. 

### 2016-03-02 paper/lrec/implementation.tex
move the underspecified readings thing from eval to implementation

@@
 
-Another note: the readings in grammar can be underspecified (e.g.
-\texttt{verb sg}), whereas the lexicon only gives us fully specified (\texttt{verb pres p2 sg}) readings.
-We tried a version where we took the tag combinations specified in the grammar as readings, and we could insert them into the symbolic sentences as well, but this was not an ideal solution: turns out that the tag lists in the grammars contain often errors (e.g.
-replacing OR with an AND; using a nonexistent tag; using a wrong level in a subreading), and if we accept those lists as readings, we will generate symbolic sentences that are impossible, and won't discover the bug in the grammar.
+The readings in grammar can be underspecified: for example, the rule
+\texttt{REMOVE (verb sg) IF (-1 det)} gives us ``verb sg'' and ``det''.
+In contrast, the lexicon only gives us fully specified readings, such
+as \texttt{verb pres p2 sg}. We implemented a version where we took
+the tag combinations specified in the grammar directly as our
+readings, and we could insert them into the symbolic sentences as well.
+The shortcut works most of the time, but it is possible to
+miss some cases: e.g. \texttt{SELECT Pron + Rel IF (0 Nom)} 
+may assume ``pron rel nom'' in one reading, but our method only gives
+``pron rel'' and ``nom'' separately. 
+In addition, we found that the tag lists in the grammars sometimes
+contain errors, such as replacing OR with an AND; using a nonexistent
+tag; using a wrong level in a subreading. If we accept those lists as
+readings, we will generate symbolic sentences that are impossible, and
+won't discover the bug in the grammar.
 

### 2016-03-02 paper/lrec/implementation.tex
change paragraphs to subsections

@@
 
-\paragraph{Preliminaries}
+%\paragraph
+\subsection{Preliminaries}
 
@@
 
-\paragraph{Symbolic sentences}
+%\paragraph
+\subsection{Symbolic sentences}
 
@@
 
-\paragraph{Applying a rule}
+%\paragraph
+\subsection{Applying a rule}
 
@@
 
-\paragraph{Putting it all together}
+%\paragraph
+\subsection{Putting it all together}
 
@@
 
-\paragraph{Creating realistic readings}
+%\paragraph{Creating realistic readings}
+\subsection{Creating realistic readings}
+\label{sec:realistic_readings}
 
@@
 readings, and we could insert them into the symbolic sentences as well.
-The shortcut works most of the time, but it is possible to
+The shortcut works most of the time, but if only take the readings
+from the grammar and ignore the lexicon, it is possible to
 miss some cases: e.g. \texttt{SELECT Pron + Rel IF (0 Nom)} 
-may assume ``pron rel nom'' in one reading, but our method only gives
+may assume ``pron rel nom'' in one reading, but this method only gives
 ``pron rel'' and ``nom'' separately. 
 In addition, we found that the tag lists in the grammars sometimes
-contain errors, such as replacing OR with an AND; using a nonexistent
-tag; using a wrong level in a subreading. If we accept those lists as
-readings, we will generate symbolic sentences that are impossible, and
-won't discover the bug in the grammar.
+contain errors, such as using a nonexistent tag or using a wrong level
+in a subreading. If we accept those lists as readings, we will
+generate symbolic sentences that are impossible, and won't discover
+the bug in the grammar.
 
-However, if we only want to find rule interaction effects, then using the underspecified readings from the grammar makes the task faster, and it will still catch potential interaction errors.
+However, if we only want to find rule interaction effects, then using
+the underspecified readings from the grammar makes the task faster,
+and it will still catch potential interaction errors.
 

### 2016-03-02 paper/lrec/eval_futurework.tex
stuff

@@
 We tested three grammars to find conflicting rules: 
-Dutch\footnote{\url{https://svn.code.sf.net/p/apertium/svn/languages/apertium-nld/apertium-nld.nld.rlx}},
+Dutch\footnote{\scriptsize{\url{https://svn.code.sf.net/p/apertium/svn/languages/apertium-nld/apertium-nld.nld.rlx}}},
 with 59 rules; 
-Spanish\footnote{\url{https://svn.code.sf.net/p/apertium/svn/languages/apertium-spa/apertium-spa.spa.rlx}},
+Spanish\footnote{\scriptsize{\url{https://svn.code.sf.net/p/apertium/svn/languages/apertium-spa/apertium-spa.spa.rlx}}},
 with 279 rules; and 
-Finnish\footnote{\url{https://github.com/flammie/omorfi/blob/master/src/vislcg3/omorfi.cg3}},
+Finnish\footnote{\scriptsize{\url{https://github.com/flammie/apertium-fin/raw/master/apertium-fin.fin.rlx}}},
 with 1185 rules. We left out \textsc{add}, \textsc{map} and other rule
@@
 by our tool. Unfortunately, we did not have time to perform such
-evaluation, and in addition, we only have gold standard corpus (20~000) words for Spanish.
+evaluation, and in addition, we only have gold standard corpus (20~000~words) for Spanish.
 
@@
 \# readings        & 336           & 3905      & 1735    \\ \hline
-\# conflicts       & 7             & n         & m    \\ \hline
-\clock{} all rules & ??s           & + 1h      & 30m-1h   \\ \hline
+\# true positives  & 7             & n         & m    \\ \hline
+\# false positives & 0             & n         & m    \\ \hline
+\clock{} with amb. 
+           classes & ??s           & + 1h      & 30m-1h   \\ \hline
 % \clock{} last rule & ??s           & ??s       & ?min ??s    \\ \hline
@@
 \end{tabular}
-\caption{\todo{make more tables}}
+\caption{Results for Dutch and Spanish grammars.}
 \label{table:res}
-\end{table}
-
-\begin{table}[]
-\centering
-\begin{tabular}{|l|l|l|l|}
-
-\hline
-              & n clitics & m clitics & only underspecified \\ \hline
-\# rules      & 1185      & 1185      & 1185\\ \hline
-\# readings   & many & more & only like <2000   \\ \hline
-\# conflicts  & n   & m        & super many    \\ \hline
-\clock{} all rules       & ??s              & ??min ??s    & ?h ??min    \\ \hline
-\clock{} last rule       & ??s              & ??s          & ?min ??s    \\ \hline
-
-\end{tabular}
-\caption{Special table for Finnish}
-\label{table:resFin}
 \end{table}
@@
 grammars, but we kept track of a number of known tricky cases; mostly
-rules that had been false positives in earlier stages of the
-development, and other rules with similar patterns, as well as rules
-with negations.
+rules with negations and complex set operations.
 %While this does not guarantee the lack of false negatives, it indicates some degree of reliability. 
-As the tool matures and we add
-new features, a more in-depth analysis will be needed.
+As the tool matures and we add new features, a more in-depth analysis
+will be needed.
 
-\paragraph{Dutch} The Dutch grammar had two kinds of errors: rule-internal, and rule interaction.
+\subsection{Dutch} The Dutch grammar had two kinds of errors: rule-internal, and rule interaction.
 As for rule-internal conflicts, one was due to a misspelling in the list definition for personal pronouns, which rendered 5 rules ineffective.
@@
 
+We found the same kind of definitions in many other rules and grammars.
+To catch them more systematically, we could add a feature that alerts in all cases where a condition with 0 is used, and automatically construct a version of the rule that moves whatever tags in the 0-condition into the target, then asks the user which one was meant.
+
 % In the latter case, the grammar writer would hopefully notice the imprecise definition and change their definition.
@@
 
-\paragraph{Spanish} The Spanish grammar had proportionately the highest number of errors.
-\todo{confirm; I assume this is gonna be so, but let's see Fin and Por}.
-The grammar we ran is like the one found in the Apertium repository, apart from 2 changes: we fixed some typos (capital O for 0) in order to make it compile, and commented out two rules that used regular expressions, because our solution does not handle them properly (see more on Future work/Implementation).
+\subsection{Spanish} The Spanish grammar had proportionately the
+highest number of errors. The grammar we ran is like the one
+found in the Apertium repository, apart from two changes: we fixed
+some typos (capital O for 0) in order to make it compile, and
+commented out two rules that used regular expressions, because our
+solution does not handle them properly; see more on Section~\ref{sec:conclusion}.
 
@@
 
-In addition, there was a number of definitions that were never
-used. The program doesn't currently point them out to the user, but
-that is a trivial addition; just like rule-internal conflicts, it does
-not require any semantic analysis. 
-We also discovered problems in some set definitions, when we tried to
-use the set definitions from the grammar directly as our readings.
+In addition, there was a number of set definitions that were never
+used. Since VISL CG-3 already points out unused sets, we did not add such
+feature in our tool. However, we noticed an unexpected benefit when
+we tried to use the set definitions from the grammar directly as our
+readings: this way, we can discover inconsistencies even in
+set definitions that are not used in any rule.
+%If a set definition is impossible, using it in a rule makes the rule have an internal conflict.
 For instance, the following definition requires the word to be all of
-the listed parts of speech at the same time:
+the listed parts of speech at the same time---the grammar writer meant
+OR instead of +:
 \begin{itemize}
@@
 
-This particular definition was never used in the rules, so it would not
-have caused a conflict. We noticed it by accident, when the SAT-solver
-offered it in an example sequence that, meant for another rule.
+If it was used in any rule, that rule would have been marked as
+conflicting. We noticed the error by accident, when the program
+offered the reading \texttt{w2<n><np><adj><det><preadv><adv><prn>}
+in an example sequence meant for another rule.
 
@@
 writer intended.
-
-We found the same kind of formulation in many other rules and grammars.
-To catch them more systematically, we could add a feature that alerts in all cases where a condition with 0 is used, and automatically construct a version of the rule that moves whatever tags in the 0-condition into the target, then asks the user which one was meant.
 
@@
 But the rule does not mention anything about ``poder'' nor ``sentir''; the SAT-solver gave a triggering sequence without
+We found the same kind of formulation in many other rules and grammars.
+To catch them more systematically, we could add a feature that alerts in all cases where a condition with 0 is used, and automatically construct a version of the rule that moves whatever tags in the 0-condition into the target, then asks the user which one was meant.
 
@@
 feasible way, then the ambiguity class constraints would create a
-sequence where ``podar'' was ambiguous with ``poder'', instead of randomly chosen readings.
+sequence where ``podar'' was ambiguous with ``poder'', only in the
+word worm ``<puedo>'' instead of randomly chosen readings.
 So, even though this was not marked as a conflict, we can say that the weird output of the tool was unnecessary, and by better implementation, the tool would have created more realistic example sequence.
 
-\paragraph{Finnish}
+\begin{table*}[t!]
+\centering
+\begin{tabular}{|l|l|l|l|}
 
-\todo{Rerun the tests and rewrite this whole section!}
+\hline
+              & 1 clitic & 2 clitics & readings only from the grammar \\ \hline
+\# readings   & many     & more & \textless{}2000   \\ \hline
+\# true 
+   positives  & n   & m        & o    \\ \hline
+\# false 
+   positives  & p   & q        & super many    \\ \hline
+\clock{}amb.
+      classes & ??s              & ??min ??s    & ?h ??min  \\ \hline
+\clock{}no amb.
+      classes & ??s              & ??s          & ?min ??s  \\ \hline
+
+\end{tabular}
+\caption{Results for Finnish (1185 rules).}
+\label{table:resFin}
+\end{table*}
+
+\subsection{Finnish}
 
@@
 
-In addition to the previous problem, we had a large number of readings.
-In order to make the test run faster, we tried a few shortcuts.
-First, we restricted the amount of clitics per word form, and even ignored all the specified readings
+In addition to the previous problem, we had a large number of
+readings. In order to make the test run faster, we used a few shortcuts.
+shortcuts. 
+Firstly, we restricted the amount of clitics per word form, and even
+ignored all the specified readings.
+Secondly, we added ambiguity class constraints for only
+\emph{intraparadigmatic ambiguities}: homonymous wordforms belonging
+to the same lexeme, in a way that is systematic to the whole
+paradigm. We reduced the morphological lexicon to contain only one
+example of each inflection paradigm, and then expanded it: with all
+the clitics, this resulted in \todo{n} word forms. Then we applied the
+normal method of creating the ambiguity classes, and specified some
+default cases manually, such as punctuation and word boundaries.
+Thirdly, we ignored the fully specified readings from the
+lexicon and used only tag combinations from the grammar, as described
+in Section~\ref{sec:realistic_readings}.
 
-The Finnish grammar was the largest and most complex of all the
-grammars we have tested.
+The results of these three variants are described in
+table~\ref{table:resFin}. As expected, using only readings from the
+grammar resulted in most false positives.
 
-To improve performance, we ignored the fully specified readings from the
-lexicon and used only tag combinations from the grammar:
-for example, the rule \texttt{REMOVE (verb sg) IF (-1 det)}
-gives us ``verb sg'' and ``det''.
-The shortcut works most of the time, but it is possible to
-miss some cases: e.g. \texttt{SELECT Pron + Rel IF (0 Nom)} 
-may assume ``pron rel nom'' in one reading, but our method only gives
-``pron rel'' and ``nom'' separately. 
-17 false positives were caused by this shortcut. 
-In addition, 2 false positives were caused by the naive implementation of
-rules with \verb!*! in their context.
+\subsection{Performance} 
+The running time of the grammars goes from seconds to hours. 
+As can be seen from the times, increasing the number of readings
+matters more than increasing the number of rules.
+In addition, the ambiguity class constraints take a long time.
 
-Due to these reasons, the results about the Finnish grammar are
-inconclusive. We only include them to show some preliminary data on larger grammars.
-%For a condition situated $n$ or more words to the left or right, we only create a word exactly $n$ words away. This is something to be fixed in the future.
+However, many of the use cases do not require running the whole
+grammar. Testing the interaction between 5--10 rules takes just
+seconds, regardless of the language.
 
-\paragraph{Performance} The speed of the check is an obvious issue,
-especially for the Finnish grammar. 
-As mentioned before, we used 2000 readings extracted from the grammar.
-With the full set of 6000 readings from the lexicon, the speed
-was even worse: checking the last rule took 30 minutes, as opposed to
-3 minutes with 2000 readings. 
-%We never finished running the whole rule set with 6000 readings, but
-%it would have taken potentially days.
-We believe we can still signifcantly improve the performance of the SAT-encoding. Currently it is just a naive proof-of-concept implementation, and we have not tried to optimise for speed at all.
+%We believe we can still signifcantly improve the performance of the SAT-encoding. Currently it is just a naive proof-of-concept implementation, and we have not tried to optimise for speed at all.
 
@@
 
-The evaluation indicates that the tool finds cases that are not
-obvious to the human eye.
+The evaluation indicates that the tool finds conflicts and dead rules
+from actual grammars. We did not have a volunteer to test the tool in
+the process of grammar writing, so we cannot conclude whether the
+constructed examples are useful for getting new insights on the rules.
+However, there are still a number of features to improve and add.
 
+\paragraph{Combining morphological and lexical tags}
+
+Our solution to hardcode the tag combinations in the readings is
+feasible for simple morphology, but it can cause problems with more
+complex morphology.
+Another benefit of this solution is the addition of lemmas and word
+forms, as well as regular expressions: currently, if we want to add
+one more lemma for a verb, we need to create as many new variables as
+there are distinct verb forms---easily hundreds for languages with
+rich morphology.
+This may also be feasible to support regular expressions, in a nicer
+way than just spitting out the regex. (MORE ON THAT)
+
+\paragraph{Full expressivity of CG-3}
+
+In order to implement \textsc{substitute},
+\textsc{add}, \textsc{addreading} and \textsc{map}, we need to be
+prepared for new readings--even if the lexicon gives all readings that
+exist in the lexicon, the user might give a nonexistent reading, or in
+the case of MAP, a syntactic tag, which is (by definition) not in the
+lexicon. A more scalable solution would be to make each tag a
+variable, and ask the question ``can this reading be a noun? how about
+singular? how about conditional?'' separately for each tag. Then we
+could lift the restriction of tag combinations into a SAT solver: make
+SAT clauses that prohibit a comparative to go with a verb, or
+conditional with a noun.
+
+% \paragraph{Regular expressions}
+% For rules with regular expressions, this would blow up even more: there are rules that only address whether the word form in a condition starts with a lowercase letter, or whether it ends in a certain suffix.
+A realistic, albeit unsatisfactory option is just to treat it as an underspecified reading, and offer that same regular expression to the user when generating a symbolic sentence.
+Another, less feasible, solution is to grep all the word forms or lemmas that match the regular expression in question---in the worst case (e.g.
+lower case), this will match all the entries in the expanded lexicon.
+
+\paragraph{Support for grammar writers}
 One additional feature is to suggest reformattings for a rule. Recall
@@
 reason(s) for the conflict.
-
-Our solution to hardcode the tag combinations in the readings is feasible for simple morphology, but it can cause problems with more complex morphology.
-One big downside is that in order to implement \textsc{ADD}, \textsc{ADDREADING} and \textsc{MAP}, we need to be prepared for new readings--even if the lexicon gives all readings that exist in the lexicon, the user might give a nonexistent reading, or in the case of MAP, a syntactic tag, which is (by definition) not in the lexicon.
-A more scalable solution would be to make each tag a variable, and ask the question ``can this reading be a noun?
-how about singular?
-how about conditional?'' separately for each tag.
-Then we could lift the restriction of tag combinations into a SAT solver: make SAT clauses that prohibit a comparative to go with a verb, or conditional with a noun.
-Another benefit of this solution is the addition of lemmas and word forms, as well as regular expressions: currently, if we want to add one more lemma for a verb, we need to create as many new variables as there are distinct verb forms---easily hundreds for languages with rich morphology.
-\todo{less rambling: For rules with regular expressions, this would blow up even more: there are rules that only address whether the word form in a condition starts with a lowercase letter, or whether it ends in a certain suffix.
-A realistic, albeit unsatisfactory option is just to treat it as an underspecified reading, and offer that same regular expression to the user when generating a symbolic sentence.
-Another, less feasible, solution is to grep all the word forms or lemmas that match the regular expression in question---in the worst case (e.g.
-lower case), this will match all the entries in the expanded lexicon.}
 
@@
 
+\subsection{Conclusions}
+
+It is nice!
+
 % * Preliminary results

### 2016-03-03 paper/lrec/lrec2016_listenmaa_claessen.tex
abstract

--- 2016-02-11 paper/lrec/lrec2016_listenmaa_claessen.tex
@@
 \abstract{
-We describe a method for analysing Constraint Grammars by encoding the
-rules in SAT.
-Our tools can detect internal conflicts or redundancies in a grammar,
-as well as generate examples to demonstrate the effect of some rule or
-rule set.
-This can help users to diagnose and improve their grammars.
-No corpus is required, only a morphological lexicon. \\ 
+We describe a method for analysing Constraint Grammars (CG) that can detect internal conflicts and redundancies in a given grammar, without the need for a corpus.
+The aim is for grammar writers to be able to automatically diagnose, and then manually improve their grammars.
+Our method works by automatically translating the given grammar into logical constraints that are analysed by a SAT-solver.
+We have evaluated our analysis on a number of non-trivial grammars and found inconsistencies.
 \newline 
-\Keywords{Constrain Grammar, SAT, Grammar Analysis} }
+\Keywords{Constraint Grammar, SAT, Grammar Analysis} }
 

### 2016-03-03 paper/lrec/eval_futurework.tex
change place of paragraph

@@
               & 1 clitic & 2 clitics & readings only from the grammar \\ \hline
-\# readings   & many     & more & \textless{}2000   \\ \hline
+\# readings   & many     & more & lagom   \\ \hline
 \# true 
@@
 \paragraph{Full expressivity of CG-3}
-
-In order to implement \textsc{substitute},
-\textsc{add}, \textsc{addreading} and \textsc{map}, we need to be
-prepared for new readings--even if the lexicon gives all readings that
-exist in the lexicon, the user might give a nonexistent reading, or in
-the case of MAP, a syntactic tag, which is (by definition) not in the
-lexicon. A more scalable solution would be to make each tag a
-variable, and ask the question ``can this reading be a noun? how about
-singular? how about conditional?'' separately for each tag. Then we
-could lift the restriction of tag combinations into a SAT solver: make
-SAT clauses that prohibit a comparative to go with a verb, or
-conditional with a noun.
+As for longer-term goals, we want to handle the full expressivity of CG-3,
+with \textsc{map}, \textsc{add},  \textsc{addreading} and
+\textsc{substitute} rules, as well as dependency structure. 
+This also means finding different kinds of conflicts.
+In order to implement rules that may add new readings, or new tags to
+existing readings, we need to modify our approach in the SAT-encoding.
+Even if the lexicon gives all readings that exist in the lexicon, the
+user might give a nonexistent reading, or in the case of MAP, a
+syntactic tag, which is (by definition) not in the lexicon. A more
+scalable solution would be to make each tag a variable, and ask the
+question ``can this reading be a noun? how about singular? how about
+conditional?'' separately for each tag. Then we could lift the
+restriction of tag combinations into a SAT solver: make SAT clauses
+that prohibit a comparative to go with a verb, or conditional with a noun.
 
@@
 reason(s) for the conflict.
-
-As for longer-term goals, we want to handle the full expressivity of CG-3,
-with \textsc{map}, \textsc{add} and \textsc{substitute} rules, and
-dependency structure. This also means finding different kinds of conflicts.
 As soon as the tools are mature enough, we want to

### 2016-03-03 paper/lrec/problem_desc.tex

@@
 \texttt{"wish" verb pl} and \texttt{"wish" verb inf}.
-However, if the target word has only one remaining reading, then the
+Note that if the target word has only one remaining reading, then the
 rule will not apply, even if the condition is met.
@@
 resourced languages. They are robust and can be written without large
-corpora---only morphological analyser is needed. The formalism is
+corpora---only a morphological analyser is needed. The formalism is
 lightweight and language-independent, and resources can be shared
@@
     (-1C Prep OR Vai OR Vbmod)})
-\item Can I rewrite a messy rule in a neater way without changing the meaning?
+\item Can a messy rule be rewritten in a neater way without changing the meaning?
 \item What is the best order for the rules?
-\item Generate a sequence that triggers $R$ but not $R'$ 
+\item Generate a sentence that triggers a given list of rules $R$ but not $R'$ 
 \end{itemize}
@@
 \begin{itemize}
-\item If a rule appears twice, the second occurrence will be disabled by the first
-\item $r$ selects something in a context, $r'$ removes~it
-\item $R$ removes something from the context of $r$, so $r$ can never
+\item If two equivalent rules $r$ and $r'$ occur in the grammar, the second occurrence will be disabled by the first
+\item Rule $r$ selects something in a context, and $r'$ removes~it
+\item A list of tules $R$ removes something from the context of a rule $r$, so $r$ can never
   apply
-\item $r$ has an internal conflict, such as non-existent
+\item A rule $r$ has an internal conflict, such as non-existent
 tag combination, or contradictory requirements for a context word
 \end{itemize}
-$R$ can be a single rule or a set of rules: for instance, if one rule removes a verb in
+In the above examples, $R$ can be a single rule or a list of rules: for instance, if one rule removes a verb in
 context $C$, and another in context $\neg C$, together these rules
@@
 A SAT-problem consists of two components: a set of Boolean variables, and a set
-of clauses on those variables. For instance, let the set ov variables
-be $\{a, b\}$ and the formulas $\{a \vee b, \neg{}a\}$. A program called
+of clauses on those variables. For instance, let the set of variables
+be $\{a, b\}$ and the formulas $\{a \vee b, \neg{}a\}$. A program called a
 \emph{SAT-solver} will try to find a solution, where all the variables
@@
 SAT-problem to have no solution or multiple solutions.
-In the case of CG analysis, the model we build is a \emph{sentence},
-which starts as potentially anything, and is being shaped by all the
+In the case of CG analysis, the solution we build (called the \emph{model}) represents a \emph{sentence} that starts as potentially anything, and is being shaped by all the
 rules along the way. This is how we can generate possible inputs, as
@@
 The paper is structured as follows. Section~\ref{sec:prev} relates our
-work to the previous research. 
+work to previous work. 
 Section~\ref{sec:implementation} discusses the implementation, and

### 2016-03-03 paper/lrec/implementation.tex
impl

@@
 
-Our analysis operates on one rule $R$, and is concerned with answering the following question: ``Does there exist an input sentence $S$ that can trigger rule $R$, even after passing all rules $R'$ that came before $R$?''
+Our analysis operates on a rule $r$, which is preceded by a list of rules $R$, and is concerned with answering the following question: ``Does there exist an input sentence $S$ that can trigger rule $r$, even after passing all rules $R$ that came before $r$?''
 
@@
 
-Furthermore, when we analyse a rule $R$, we need to decide the {\em width} $w(R)$ of the rule $R$: How many different words should there be in a sentence that can trigger $R$?
-Most often, $w(R)$ can be easily determined by looking at how far away the rule context indexes in the sentence relative to the target.
+Furthermore, when we analyse a rule $r$, we need to decide the {\em width} $w(r)$ of the rule $r$: How many different words should there be in a sentence that can trigger $r$?
+Most often, $w(r)$ can be easily determined by looking at how far away the rule context indexes in the sentence relative to the target.
 For example, in the rule mentioned in the introduction, the width is 2.
@@
 If the context contains a \verb!*!,
-we may need to make an approximation of $w(R)$ which may result in false negatives later on in the analysis.
+we may need to make an approximation of $w(r)$ which may result in false negatives later on in the analysis.
 %we create sentences that are up to 3 words wider than the context without \verb!*!, and try them all.
@@
 
-The number of words in the symbolic sentence we create when we analyse a rule $R$ is $w(R)$. For the rule in the introduction, we have $w(R)=2$ and a symbolic sentence may look as follows:
+The number of words in the symbolic sentence we create when we analyse a rule $r$ is $w(r)$. For the rule in the introduction, we have $w(r)=2$ and a symbolic sentence may look as follows:
 \begin{center}
@@
 \end{center}
-Any solution to the constraints found by the SAT-solver can be interpreted as a concrete sentence with $w(R)$ words that each have a set of readings.
+Any solution to the constraints found by the SAT-solver can be interpreted as a concrete sentence with $w(r)$ words that each have a set of readings.
 
@@
 
-Next, we need to be able to apply a given rule $R'$ to a symbolic sentence, resulting in a new symbolic sentence.
+Next, we need to be able to apply any given rule $r'$ to a symbolic sentence, resulting in a new symbolic sentence.
 
@@
 The example rule can only affect readings of word2 that have a ``verb'' tag, so we create only two new variables $w_4'$ and $w_5'$ for the result, and reuse the other variables.
-We add the following constraint for $w_4'$:
+We must also add the following constraint for $w_4'$:
 \begin{center}
@@
 
-Once we know how to apply one rule $R'$ to a symbolic sentence, we can apply all rules preceding the rule $R$ that is under analysis.
+Once we know how to apply any rule $r'$ to a symbolic sentence, resulting in a new symbolic sentence, we can apply all rules preceding the rule $r$ that is under analysis.
 We simply apply each rule to the result of applying the previous rule.
@@
 
-Finally, we can take a look at the rule $R$ we want to analyse. Here is an example:
+Finally, we can take a look at the rule $r$ we want to analyse. Here is an example:
 \begin{itemize}
@@
 \end{itemize}
-If we take the symbolic sentence above as input, we want to ask whether or not it can trigger the rule $R$. We do this by adding some more constraints to the SAT-solver.
+If we take the symbolic sentence above as input, we want to ask whether or not it can trigger the rule $r$. We do this by adding some more constraints to the SAT-solver.
 
@@
 If the SAT-solver can find a solution to all constraints generated so far, we have found a concrete sentence that satisfies our goal.
-If the SAT-solver cannot find a solution, it means that there are no sentences that can ever trigger rule $R$.
-This means that there is something wrong with the grammar.
+If the SAT-solver cannot find a solution, it means that there are no sentences that can ever trigger rule $r$.
+(This means that there is something wrong with the grammar.)
 
@@
 
-Earlier we have shown an example with 5 readings (``det def'', ``noun sg'', ...). In a realistic case, we operate between hundreds and thousands of readings. 
+Earlier we have shown an example with 5 readings (``det def'', ``noun sg'', ...). In a realistic case, we operate between hundreds and thousands of possible readings. 
 %This is very much dependent on language: the simplest language we tested was Dutch, with 336 readings. The most complex was 
@@
 In the previous section, we have created realistic \emph{readings}, by simply hardcoding legal tag combinations into variables.
-The next step in creating realistic ambiguities is to constrain what readings can go together.
+The next step in creating realistic \emph{ambiguities} is to constrain what readings can go together.
 For instance, the case of \emph{zijn} shows us that ``determiner or verb'', is a possible ambiguity.
@@
 However, it could be that past participles (pp) only ever get confused with adjectives---in that case, the above rules would be in conflict with each other.
- By removing the adjective reading, the first rule selects the past participle reading, making it an instance of ``$R$ selects something in a context, $R'$ removes~it''. 
+ By removing the adjective reading, the first rule selects the past participle reading, making it an instance of ``$r$ selects something in a context, $r'$ removes~it''. 
 The additional constraints will prevent the SAT-solver from creating an ambiguity outside the allowed classes, and such a case would be caught as a conflict.

### 2016-03-06 paper/lrec/implementation.tex
little cleanup

@@
 The number of verb readings without clitics is 213, and with clitics 1572.
-With the previously mentioned approach, we would have to double 1572 entries for each verb lemma. Even ignoring the clitics, each verb lemma would still result in 213 new readings.
+With the previously mentioned approach, we would have to double 1572 entries for each verb lemma. Even ignoring the clitics, each verb lemma still adds 213 new readings.
 
- In our experience,  even ignoring the clitics doubles the amount of readings.
-
-The readings in grammar can be underspecified: for example, the rule
+The readings in a grammar can be underspecified: for example, the rule
 \texttt{REMOVE (verb sg) IF (-1 det)} gives us ``verb sg'' and ``det''.
 In contrast, the lexicon only gives us fully specified readings, such
-as \texttt{verb pres p2 sg}. We implemented a version where we took
+as ``verb pres p2 sg''. We implemented a version where we took
 the tag combinations specified in the grammar directly as our
 readings, and we could insert them into the symbolic sentences as well.
-The shortcut works most of the time, but if only take the readings
+The shortcut works most of the time, but if we only take the readings
 from the grammar and ignore the lexicon, it is possible to
-miss some cases: e.g. \texttt{SELECT Pron + Rel IF (0 Nom)} 
-may assume ``pron rel nom'' in one reading, but this method only gives
+miss some cases: e.g. the rule \texttt{SELECT Pron + Rel IF (0 Nom)} 
+may require ``pron rel nom'' in one reading, but this method only gives
 ``pron rel'' and ``nom'' separately. 
@@
 
-\paragraph{Creating realistic ambiguities}
+\subsection{Creating realistic ambiguities}
 
 In the previous section, we have created realistic \emph{readings}, by simply hardcoding legal tag combinations into variables.
-The next step in creating realistic \emph{ambiguities} is to constrain what readings can go together.
+The next step in creating realistic \emph{ambiguities} is to constrain which readings can go together.
 For instance, the case of \emph{zijn} shows us that ``determiner or verb'', is a possible ambiguity.
-In contrast, there is no word form in the lexicon that would be ambiguous between an adjective and a comma, hence we don't want to generate such ambiguity in our symbolic sentences.
+In contrast, there is no word form in the lexicon that would be ambiguous between an adjective and a comma, hence we do not want to generate such ambiguity in our symbolic sentences.
 
@@
 
-We solve the problem by creating \emph{ambiguity classes}: groups of analyses that can be ambiguous with each other. 
+We solve the problem by creating \emph{ambiguity classes}: groups of readings that can be ambiguous with each other. 
 We represent the expanded morphological lexicon as a matrix, as seen
@@
 They take advantage of the fact that they don't have to model ``bear'' and ``wish'' as separate entries, but they can just reduce it to ``word that can be ambiguous between noun and verb'', and use it as a parameter in their HMM.
-We can do a similar thing by saving a list of words with each ambiguity class.
+%We can do a similar thing by saving a list of words with each ambiguity class.
 For example, we map the ambiguity class ``feminine plural noun or a separable verb infinitive'' to the list of word forms \{``uitgaven'', ``toespraken''\}, and then, if we generate such reading for our symbolic word, we can give one of these words as an example word.
@@
  With our symbolic sentence, these rules will be no problem; to apply the latter, we only need to construct a target that has a realistic ambiguity with a past participle; the adjective will be gone already.
-However, it could be that past participles (pp) only ever get confused with adjectives---in that case, the above rules would be in conflict with each other.
+However, it could be that past participles (pp) only ever get confused with adjectives---in that case, the above rules would contradict each other.
  By removing the adjective reading, the first rule selects the past participle reading, making it an instance of ``$r$ selects something in a context, $r'$ removes~it''. 

### 2016-03-06 paper/lrec/eval_futurework.tex
add more about finnish

@@
 
-Two rules, which both remove an adverb, were in an order where the first has more broad condition: remove adverb if followed by a noun.
-In contrast, the second rule has a stricter requirement: only remove adverb if it is preceded by a determiner; the adverb itself is ambiguous with an adjective, and followed by a noun.
-The proble is that the first rule removes the adverb in all possible cases, and the second will not have any chance to act.
+These two rules both remove an adverb, but the first one has a broader condition:
+remove adverb if followed by a noun.
+In contrast, the second rule has a stricter condition: only remove adverb if it is preceded by a determiner, the adverb itself is ambiguous with an adjective, and followed by a noun.
+The problem is that the first rule removes the adverb in all possible cases, and the second will not have any chance to act.
 If the rules were in the opposite order, then there would be no problem.
@@
 As per VISL CG-3, the condition \texttt{(0 "zijn")} does not require
-that the input would have determiner and ``zijn'' in the same reading:
-just that there is a reading with any determiner, and a reading with any ``zijn'' in the cohort. 
+that the input would have determiner and \emph{zijn} in the same reading:
+just that there is a reading with any determiner, and a reading with any \emph{zijn} in the cohort. 
 Can we catch this imprecise formulation with our tool? The example word constructed by the SAT-solver may as likely be one of the following:
@@
     w2<det><pos><f><sg>
-    w2<vbser><inf><"zijn">
+    zijn<vbser><inf>
 \end{verbatim}
@@
 "w2"
-    w2<det><pos><mfn><pl><"zijn">
+    zijn<det><pos><mfn><pl>
 \end{verbatim}
@@
 We found the same kind of definitions in many other rules and grammars.
-To catch them more systematically, we could add a feature that alerts in all cases where a condition with 0 is used, and automatically construct a version of the rule that moves whatever tags in the 0-condition into the target, then asks the user which one was meant.
+To catch them more systematically, we could add a feature that alerts in all cases where a condition with 0 is used.
+As a possible extension, we could automatically merge the 0-condition into the target reading, then show the user this new version, along with the original, and ask which one was intended.
 
@@
 
-The comments make it clear that the first rule is meant to disambiguate between poder and podar, and the second rule between sentir and sentar.
+The comments make it clear that the first rule is meant to disambiguate between \emph{poder} and \emph{podar}, and the second rule between \emph{sentir} and \emph{sentar}.
 But the rule does not mention anything about ``poder'' nor ``sentir''; the SAT-solver gave a triggering sequence without
-We found the same kind of formulation in many other rules and grammars.
-To catch them more systematically, we could add a feature that alerts in all cases where a condition with 0 is used, and automatically construct a version of the rule that moves whatever tags in the 0-condition into the target, then asks the user which one was meant.
 
@@
 The results for the Finnish grammar are shown separately, in table~\ref{table:resFin}.
+
 We were not able to use the method with ambiguity classes at all---expanding the Finnish morphological lexicon results in \todo{100s of gigabytes?} of word forms, which is simply too big for our method to work.
@@
 In addition to the previous problem, we had a large number of
-readings. In order to make the test run faster, we used a few shortcuts.
-shortcuts. 
+readings to start with: more than 20,000 when all the clitics are added.
+After weeding out the most uncommon combinations, we ended up with sets of 4500--7000 readings.
+The results are shown in table~\ref{table:resFin}, to give an idea how the additional readings affect the accuracy and performance.
+
+The grammar contains around 900 lemmas and word forms.
+The procedure described in Section~\ref{sec:realistic_readings} would not have worked, so we simply took the lemmas and word forms, and added them as single readings, which are allowed to combine with any other reading.
+In addition, we took the underspecified readings when they included a lemma or a word form, such as in \texttt{"aika" SELECT N  (1C VFIN-C) ;}.
+
+ In order to make the test run faster, we used a few shortcuts.
+
 Firstly, we restricted the amount of clitics per word form, and even
 ignored all the specified readings.
-Secondly, we added ambiguity class constraints for only
-\emph{intraparadigmatic ambiguities}: homonymous wordforms belonging
-to the same lexeme, in a way that is systematic to the whole
-paradigm. We reduced the morphological lexicon to contain only one
-example of each inflection paradigm, and then expanded it: with all
-the clitics, this resulted in \todo{n} word forms. Then we applied the
-normal method of creating the ambiguity classes, and specified some
-default cases manually, such as punctuation and word boundaries.
+
+% Secondly, we added ambiguity class constraints for only
+% \emph{intraparadigmatic ambiguities}: homonymous wordforms belonging
+% to the same lexeme, in a way that is systematic to the whole
+% paradigm. We reduced the morphological lexicon to contain only one
+% example of each inflection paradigm, and then expanded it: with all
+% the clitics, this resulted in \todo{n} word forms. Then we applied the
+% normal method of creating the ambiguity classes, and specified some
+% default cases manually, such as punctuation and word boundaries.
 Thirdly, we ignored the fully specified readings from the
@@
 grammar resulted in most false positives.
+
+TODO: using readings from grammar directly, sometimes you can have a typo, like in \emph{lähetii}:
+
+\begin{verbatim}
+"w2"
+        w2<"se"><ela>''''''''
+        w2<"<lähetii>"><vblex>'''''''
+\end{verbatim}
 

### 2016-03-06 paper/lrec/eval_futurework.tex
more finnish

@@
 The comments make it clear that the first rule is meant to disambiguate between \emph{poder} and \emph{podar}, and the second rule between \emph{sentir} and \emph{sentar}.
-But the rule does not mention anything about ``poder'' nor ``sentir''; the SAT-solver gave a triggering sequence without
+But the rule does not mention anything about \emph{poder} nor \emph{sentir}; without the ambiguity class constraints, the SAT-solver could give a triggering sequence with any other reading in the target.
 
-This example shows a shortcoming of our way of handling the lemmas. If
-the lemmas and word forms were integrated to the analyses in a more
-feasible way, then the ambiguity class constraints would create a
-sequence where ``podar'' was ambiguous with ``poder'', only in the
-word worm ``<puedo>'' instead of randomly chosen readings.
+This example shows the 
+shows a shortcoming in our way of handling the lexical conditions. 
+If the lemmas and word forms were integrated to the analyses in a more
+sensible way, then the ambiguity class constraints would create a
+sequence where \emph{podar} was ambiguous with \emph{poder} only in the
+word form \emph{puedo}.
+
+instead of randomly chosen readings.
 So, even though this was not marked as a conflict, we can say that the weird output of the tool was unnecessary, and by better implementation, the tool would have created more realistic example sequence.
@@
 \centering
-\begin{tabular}{|l|l|l|l|}
+\begin{tabular}{|p{2.5cm}|p{3cm}|p{3cm}|p{3cm}|p{3cm}|}
 
 \hline
-              & 1 clitic & 2 clitics & readings only from the grammar \\ \hline
-\# readings   & many     & more & lagom   \\ \hline
+              & 1 clitic + lemmas from grammar & 2 clitics + lemmas from grammar & 1 clitic + all readings 
+                                                from grammar & all readings from grammar \\ \hline
+\# readings   & 5851 (4263 
+                 + 1588) & 9494
+                       (7906 + 1588) & 6657 (4263 + 2394) & 2394  \\ \hline
+\# conflicts  & 214      & 214       & 24                & 24  \\ \hline
+\# interaction & 214      & 214       & 24                & 24  \\ \hline
+
 \# true 
-   positives  & n   & m        & o    \\ \hline
+   positives  & n        & m         & o                & p  \\ \hline
 \# false 
-   positives  & p   & q        & super many    \\ \hline
-\clock{}amb.
-      classes & ??s              & ??min ??s    & ?h ??min  \\ \hline
-\clock{}no amb.
-      classes & ??s              & ??s          & ?min ??s  \\ \hline
+   positives  & p        & q         & r                & s   \\ \hline
+\clock{}all 
+       rules & 12min     & 23min     & 12min &  4min 40s \\ \hline
 
@@
 The results for the Finnish grammar are shown separately, in table~\ref{table:resFin}.
+We encountered a number of difficulties and used a few shortcuts, which we did not need for the other grammars---most importantly, not using the ambiguity class constraints.
+Due to these complications, the results are not directly comparable, but we include Finnish in any case,
+to give an idea how our method scales up: both to grammars with more rules, and to more complex rules.
 
-We were not able to use the method with ambiguity classes at all---expanding the Finnish morphological lexicon results in \todo{100s of gigabytes?} of word forms, which is simply too big for our method to work.
+\paragraph{Challenges with Finnish} The first challenge is the morphological complexity of Finnish.
+There are more than 20,000 readings, when all possible clitic combinations are included.
+After weeding out the most uncommon combinations, we ended up with sets of 4000--8000 readings.
+
+The second challenge comes from the larger size of the grammar. Whereas the Spanish and Dutch had only tens of word forms or lemmas, the Finnish grammar specifies around 900 of them.
+Due to both of these factors, the procedure described in Section~\ref{sec:realistic_readings} would have exploded the number of readings, so we simply took the lemmas and word forms, and added them as single readings.
+In cases where they were combined with another tag in the grammar, we took that reading directly: for instance, we took both \emph{aika} and ``\emph{aika} n'' from \texttt{SELECT "aika" + N},
+but nothing from \texttt{SELECT Pron + Sg}. This method gave us 1588 additional readings.
+
+The third were not able to create ambiguity class constraints---expanding the Finnish morphological lexicon results in \todo{100s of gigabytes?} of word forms, which is simply too big for our method to work.
 For future development, we will see if it is possible to manipulate the finite automata directly, instead of relying on the output in text.
 
-In addition to the previous problem, we had a large number of
-readings to start with: more than 20,000 when all the clitics are added.
-After weeding out the most uncommon combinations, we ended up with sets of 4500--7000 readings.
-The results are shown in table~\ref{table:resFin}, to give an idea how the additional readings affect the accuracy and performance.
+\paragraph{Results}
+The results are shown in table~\ref{table:resFin}.
+In the first column, we included only possessive suffixes. In the second column, we included question clitics as well.
+In the third column, we included all the tag combinations specified in the grammar, and in the fourth, we took only those, ignoring the morphological lexicon completely.
 
-The grammar contains around 900 lemmas and word forms.
-The procedure described in Section~\ref{sec:realistic_readings} would not have worked, so we simply took the lemmas and word forms, and added them as single readings, which are allowed to combine with any other reading.
-In addition, we took the underspecified readings when they included a lemma or a word form, such as in \texttt{"aika" SELECT N  (1C VFIN-C) ;}.
+For the first two columns, there are a high number of internal conflicts. 
+These are almost all due to nonexisting tags.
+The grammar was written in 1995, and updated by \newcite{pirinen2015}; but this high number of internal conflicts indicates that possibly something has gone wrong in the conversion, or in our expansion of the morphological lexicon.
+As for accuracy, adding the question clitics did not change anything: they were already included in some of the 1588 sets with word forms or lemmas, and that was enough for the SAT-solver to find models with question clitics.
+We left the result in the table just to demonstrate the change in speed.
 
- In order to make the test run faster, we used a few shortcuts.
+%Example reading that adds question clitic to our arsenal: \texttt{SELECT "joka" + Pron + Q}
 
-Firstly, we restricted the amount of clitics per word form, and even
-ignored all the specified readings.
+The second two columns are playing with the full set of readings from the grammar. For both of these, the number of reported conflicts was only 24.
+
+To show some examples of true positives:
+
+%SELECT ("oma" gen) IF (NOT 1 n|adj|vblex|vaux|pron|cs|cc|adp|post|pr|intj|num|abbr) (0 LINK *-1 "olla"|"voida"|"saattaa") (0 n|adj|vblex|vaux|pron|cs|cc|adp|post|pr|intj|num|abbr) (0C nom)
+
+\begin{itemize}
+\item[] \todo{Put back the rule with 0 gen and 0C nom!!! that was actually nice and legit}
+
+\item[]\texttt{SELECT ADV (-1C AD) (NOT 0 PP) (NOT 1 A) ;}
+\end{itemize}
+
+This is an internal conflict, which is not trivial to see: the problem is that the definition of \texttt{PP} includes \texttt{adv} among others.
+With the sets expanded, this rule becomes \texttt{SELECT adv IF (NOT 0 pp|adv|adp|po|pr) ...}.
+
+The following two examples are interaction conflicts:
+
+\begin{itemize}
+\item[]\begin{verbatim}
+REMOVE A (0 DER) ; 
+REMOVE N (0 DER) ; 
+REMOVE A/N (0 DER) ; 
+\end{verbatim}
+\end{itemize}
+
+In this case, two rules together remove the target of the third, leaving no way for there to be adjective or noun.
+
+\begin{itemize}
+\item[]\begin{verbatim}
+SELECT:r1 .. IF (-1 Comma/N/Pron/Q) ;
+SELECT:r2 .. IF (-2 ..) (-1 Comma) ;
+\end{verbatim}
+\end{itemize}
+
+The rules above have been simplified to show only the relevant part.
+The conflict lies in the fact that \texttt{Comma} is a subset of \texttt{Comma/N/Pron/Q}:
+there is no way to trigger $r2$ without placing a comma in position -1, and thereby triggering $r1$.
+
+% All in all, using readings from grammar directly may be useful in order to find just the , sometimes you can have a typo, like in \emph{l\"{a}hetii}:
+
+% \begin{verbatim}
+% "w2"
+%         w2<"se"><ela>
+%         w2<"<lähetii>"><vblex>
+% \end{verbatim}
+
+% The second part concerns the tools in use, specifically the Apertium morphological lexicon.
+We noticed, only after running the experiments, that the expansion of the lexicon does not work properly: for instance, some nominal forms of the verbs were reported as nouns instead of verbs.
+% We reported the errors to the developers of the lexicon, but as there was no time to fix them, we used the readings that we got from expanding the lexicon, and manually added either noun, verb or adjective tag where it was missing.
+
+% kosimisiin:kosia<n><pl><ill>
+% kosimattomuudekseen:kosia<sg><tra><pxsp3>
 
@@
 % default cases manually, such as punctuation and word boundaries.
-Thirdly, we ignored the fully specified readings from the
-lexicon and used only tag combinations from the grammar, as described
-in Section~\ref{sec:realistic_readings}.
-
-The results of these three variants are described in
-table~\ref{table:resFin}. As expected, using only readings from the
-grammar resulted in most false positives.
-
-TODO: using readings from grammar directly, sometimes you can have a typo, like in \emph{lähetii}:
-
-\begin{verbatim}
-"w2"
-        w2<"se"><ela>''''''''
-        w2<"<lähetii>"><vblex>'''''''
-\end{verbatim}
 

### 2016-03-07 paper/lrec/eval_futurework.tex
conclusions

@@
 
-\section{Future work and conclusions}
+\section{Conclusions and Future Work}
 \label{sec:conclusion}
 
-The evaluation indicates that the tool finds conflicts and dead rules
-from actual grammars. We did not have a volunteer to test the tool in
+We set out to design and implement an automatic analysis of constraint grammars that can find problematic rules and rule combinations, without the need for a corpus.
+Our evaluation indicates that the tool indeed finds non-trivial conflicts and dead rules
+from actual grammars.
+
+We did not have a volunteer to test the tool in
 the process of grammar writing, so we cannot conclude whether the
 constructed examples are useful for getting new insights on the rules.
-However, there are still a number of features to improve and add.
+
+Also, there are still a number of features to improve and add.
 
@@
 complex morphology.
-Another benefit of this solution is the addition of lemmas and word
-forms, as well as regular expressions: currently, if we want to add
-one more lemma for a verb, we need to create as many new variables as
-there are distinct verb forms---easily hundreds for languages with
-rich morphology.
+We are currently working on adding the concepts of lemmas and word
+forms directly to the representation of the possible readings.
+Currently, if we want to add one more lemma for a verb, we need to
+create as many new variables as there are distinct verb forms---easily
+hundreds for languages with rich morphology. (INARI: Do you really mean to say 'verb forms' here?)
+
 This may also be feasible to support regular expressions, in a nicer
-way than just spitting out the regex. (MORE ON THAT)
+way than just spitting out the regex. (INARI: You said MORE ON THAT here, good because I don't understand what you want to say here :)
 
@@
 user might give a nonexistent reading, or in the case of MAP, a
-syntactic tag, which is (by definition) not in the lexicon. A more
-scalable solution would be to make each tag a variable, and ask the
-question ``can this reading be a noun? how about singular? how about
-conditional?'' separately for each tag. Then we could lift the
-restriction of tag combinations into a SAT solver: make SAT clauses
-that prohibit a comparative to go with a verb, or conditional with a noun.
+syntactic tag, which is (by definition) not in the lexicon. We may need to move
+to a more scalable solution.
+% would be to make each tag a variable, and ask the
+%question ``can this reading be a noun? how about singular? how about
+%conditional?'' separately for each tag. Then we could lift the
+%restriction of tag combinations into a SAT solver: make SAT clauses
+%that prohibit a comparative to go with a verb, or conditional with a noun.
 
@@
 reason(s) for the conflict.
-As soon as the tools are mature enough, we want to
-evaluate them with actual grammar writers,
+Our next step is to evaluate our tools together with actual grammar writers,
 in comparison with a corpus-based method or machine learning.
-Finally, if the method proves feasible for CG, we want
-to try applying it to other grammar formalisms.
 
-\subsection{Conclusions}
+\paragraph{Other grammar formalisms}
+Finally, we would like to investigate logic-based methods for analysing other grammar
+formalisms, for example Grammatical Framework \cite{gf}. (INARI: can you add a reference here?)
 
-It is nice!
+%\subsection{Conclusions}
+
+%It is nice!
 

### 2016-03-07 paper/lrec/eval_futurework.tex
mention two versions of spanish + ambiguity class constraint generation takes time

--- 2016-03-06 paper/lrec/eval_futurework.tex
@@
 solution does not handle them properly; see more on Section~\ref{sec:conclusion}.
+
+We include two versions of the Spanish grammar in table~\ref{table:res}: in column 2, we added the lemmas and word forms as described in Section~\ref{sec:realistic_readings}, and in column 3, we just added each word form and lemma as individual readings, allowed to combine with any other reading.
+\todo{check if true: This did not affect the amount of conflicts detected, but we found some differences in the generated output; examples will be shown later in this section.}
 
@@
 grammar. Testing the interaction between 5--10 rules takes just
-seconds, regardless of the language.
+seconds in all languages, if the ambiguity class constraints are not included. 
+A downside in the ambiguity classes is that generating them takes a long time, 
+and while 5-minute overhead may be acceptable when checking the full grammar,
+it is hardly so when analysing just a handful of rules.
 

### 2016-03-07 paper/lrec/eval_futurework.tex
change order of paragraphs in future work

@@
 
-\section{Future work and conclusions}
+\section{Conclusions and Future Work}
 \label{sec:conclusion}
 
-The evaluation indicates that the tool finds conflicts and dead rules
-from actual grammars. We did not have a volunteer to test the tool in
+We set out to design and implement an automatic analysis of constraint grammars that can find problematic rules and rule combinations, without the need for a corpus.
+Our evaluation indicates that the tool indeed finds non-trivial conflicts and dead rules
+from actual grammars.
+
+We did not have a volunteer to test the tool in
 the process of grammar writing, so we cannot conclude whether the
 constructed examples are useful for getting new insights on the rules.
-However, there are still a number of features to improve and add.
+In any case, there are still a number of features to improve and add.
 
@@
 complex morphology.
+A more scalable solution would be to make each tag a variable, and ask the
+question ``can this reading be a noun? how about singular? how about
+conditional?'' separately for each tag. Then we could lift the
+restriction of tag combinations into a SAT solver: make SAT clauses
+that prohibit a comparative to go with a verb, or conditional with a noun.
+
 Another benefit of this solution is the addition of lemmas and word
@@
 rich morphology.
+
+We are currently working on adding the concepts of lemmas and word
+forms directly to the representation of the possible readings.
+Currently, if we want to add one more lemma for a verb, we need to
+create as many new variables as there are distinct word forms---easily
+hundreds for languages with rich morphology.
+
 This may also be feasible to support regular expressions, in a nicer
-way than just spitting out the regex. (MORE ON THAT)
+way than just spitting out the regex. (INARI: You said MORE ON THAT here, good because I don't understand what you want to say here :)
 
@@
 user might give a nonexistent reading, or in the case of MAP, a
-syntactic tag, which is (by definition) not in the lexicon. A more
-scalable solution would be to make each tag a variable, and ask the
-question ``can this reading be a noun? how about singular? how about
-conditional?'' separately for each tag. Then we could lift the
-restriction of tag combinations into a SAT solver: make SAT clauses
-that prohibit a comparative to go with a verb, or conditional with a noun.
+syntactic tag, which is (by definition) not in the lexicon. We may need to move
+to a more scalable solution.
 

### 2016-03-07 paper/lrec/eval_futurework.tex
small cleanup

@@
 matters more than increasing the number of rules.
-In addition, the ambiguity class constraints take a long time.
+In addition, the ambiguity class constraints increase the time.
 
@@
 that prohibit a comparative to go with a verb, or conditional with a noun.
-
-Another benefit of this solution is the addition of lemmas and word
-forms, as well as regular expressions: currently, if we want to add
-one more lemma for a verb, we need to create as many new variables as
-there are distinct verb forms---easily hundreds for languages with
-rich morphology.
 

### 2016-03-08 paper/lrec/eval_futurework.tex
more numbers and stuff on spanish

@@
 \# true positives  & 7             & n         & m    \\ \hline
-\# false positives & 0             & n         & m    \\ \hline
+\# internal/interaction & 6 + 1    & n         & + 25    \\ \hline
+
+\# false positives & 0             & n         & 5    \\ \hline
+\# internal/interaction & 0        & n         & 5 +   \\ \hline
+
 \clock{} with amb. 
-           classes & ??s           & + 1h      & 30m-1h   \\ \hline
+           classes & ??s           & 5h 42m    &  11m25  \\ \hline
 % \clock{} last rule & ??s           & ??s       & ?min ??s    \\ \hline
 \clock{} no amb. 
-           classes & 10ish sec       & ??min ??s    & ?h ??min    \\ \hline
+           classes & 10ish sec       & 14m 16s    & ?h ??min    \\ \hline
 
@@
 We include two versions of the Spanish grammar in table~\ref{table:res}: in column 2, we added the lemmas and word forms as described in Section~\ref{sec:realistic_readings}, and in column 3, we just added each word form and lemma as individual readings, allowed to combine with any other reading.
-\todo{check if true: This did not affect the amount of conflicts detected, but we found some differences in the generated output; examples will be shown later in this section.}
+This latter version ran much faster, but also resulted in 5 false positives, all for rules that require a lemma in a condition.
+In addition, there were differences in the generated output; examples will be shown later in this section.
 
 Again, we can classify the conflicts into internal and interaction.
-As an example of internal conflict, there are \todo{n} rules that use \texttt{SET Cog = (np cog)}---the problem is that the tag \texttt{cog} does not exist in the Apertium dictionary for Spanish.
-It is likely that this grammar has been written for an earlier version, where such tag has been in place.
+As an example of internal conflict, there are two rules that use \texttt{SET Cog = (np cog)}---the problem is that the tag \texttt{cog} does not exist in the Apertium dictionary for Spanish.
+As another example, four rules require a context word tagged as NP with explicit number, but the lexicon does not indicate any number with NPs.
+It is likely that this grammar has been written for an earlier version, where such tags have been in place.
 
-The Spanish grammar has a high number of interaction conflicts, \todo{n} with the same pattern:
+As for the 25 interaction conflicts, there were only 9 distinct rules that rendered 25 other rules ineffective.
+In fact, we can reduce these 9 rules further into 3 different groups: 4 + 4 + 1, where the groups of 4 rules are variants of otherwise identical rule, with  combinations feminine singular, masculine singular, feminine plural and masculine plular.
+An example of such conflict is below (gender and number omitted for readability):
 
@@
 \item[] 
-\begin{verbatim}REMOVE (a) IF (-1 x) (1 y) ;
-REMOVE (a) IF (-1 x) (1 y) (NOT -1 z);
+\begin{verbatim}
+# NOM ADJ ADJ
+SELECT A OR PP IF (-2 N) 
+ (-1 Adj_PP) (0 Adj_PP) (NOT 0 Det);
+
+# NOM ADJ ADJ ADJ
+SELECT A OR PP IF (-3 N) (-2 N) 
+ (-1 Adj_PP) (0 Adj_PP) (NOT 0 Det);
 \end{verbatim}
@@
 
-This is another instance of more general rule applying before more
-specific. Negated context works as well to make a rule stricter. The
-second rule will never apply, because the first will catch all
-cases. For a full list of found conflicts, see \todo{\url{actual://link.here/spanish/error.log}}
+For a full list of found conflicts, see \todo{\url{actual://link.here/spanish/error.log}}
 
@@
 Due to both of these factors, the procedure described in Section~\ref{sec:realistic_readings} would have exploded the number of readings, so we simply took the lemmas and word forms, and added them as single readings.
-In cases where they were combined with another tag in the grammar, we took that reading directly: for instance, we took both \emph{aika} and ``\emph{aika} n'' from \texttt{SELECT "aika" + N},
-but nothing from \texttt{SELECT Pron + Sg}. This method gave us 1588 additional readings.
+In cases where they were combined with another tag in the grammar, we took that reading directly: for instance, we included both \emph{aika} and ``\emph{aika} n'' from the rule \texttt{SELECT "aika" + N},
+but nothing from the rule \texttt{SELECT Pron + Sg}. This method gave us 1588 additional readings.
 
-The third were not able to create ambiguity class constraints---expanding the Finnish morphological lexicon results in \todo{100s of gigabytes?} of word forms, which is simply too big for our method to work.
+Finally, we were not able to create ambiguity class constraints---expanding the Finnish morphological lexicon results in \todo{100s of gigabytes?} of word forms, which is simply too big for our method to work.
 For future development, we will see if it is possible to manipulate the finite automata directly, instead of relying on the output in text.
@@
 \begin{itemize}
-\item[] \todo{Put back the rule with 0 gen and 0C nom!!! that was actually nice and legit}
-
+\item[] \item[]\texttt{SELECT "oma" + Gen IF .. (0C Nom) ;}
 \item[]\texttt{SELECT ADV (-1C AD) (NOT 0 PP) (NOT 1 A) ;}
@@
 
-This is an internal conflict, which is not trivial to see: the problem is that the definition of \texttt{PP} includes \texttt{adv} among others.
+Both of these are internal conflicts, which may not be trivial to see: 
+the first rule requires the target to be genitive and unambiguously nominative: 
+the problem in the first rule is that the target is required to be  the problem is that the definition of \texttt{PP} includes \texttt{adv} among others.
 With the sets expanded, this rule becomes \texttt{SELECT adv IF (NOT 0 pp|adv|adp|po|pr) ...}.
@@
 \item[]\begin{verbatim}
-SELECT:r1 .. IF (-1 Comma/N/Pron/Q) ;
-SELECT:r2 .. IF (-2 ..) (-1 Comma) ;
+SELECT .. IF (-1 Comma/N/Pron/Q) ;
+SELECT .. IF (-2 ..) (-1 Comma) ;
 \end{verbatim}
@@
 The conflict lies in the fact that \texttt{Comma} is a subset of \texttt{Comma/N/Pron/Q}:
-there is no way to trigger $r2$ without placing a comma in position -1, and thereby triggering $r1$.
+there is no way to trigger the second without placing a comma in position -1, and thereby triggering the first rule.
 
@@
 \subsection{Performance} 
-The running time of the grammars goes from seconds to hours. 
+The running time of the grammars ranges from seconds to hours. 
 As can be seen from the times, increasing the number of readings
@@
 it is hardly so when analysing just a handful of rules.
+
+%Spanish lemmahack: time to create the formulas 5m32.214s
+
+%Spanish no lemmahack: time to create the formulas  62m6.049s
 

### 2016-03-08 paper/lrec/eval_futurework.tex
more on finnish

@@
 \centering
-\begin{tabular}{|l|l|l|l|}
+\begin{tabular}{|p{2.80cm}|l|l|l|}
 
@@
 \# readings        & 336           & 3905      & 1735    \\ \hline
-\# true positives  & 7             & n         & m    \\ \hline
-\# internal/interaction & 6 + 1    & n         & + 25    \\ \hline
-
-\# false positives & 0             & n         & 5    \\ \hline
-\# internal/interaction & 0        & n         & 5 +   \\ \hline
+\# true positives 
+    \small{(internal + interaction)}
+                   & 7 (6 + 1)     & n         & m ( + 25)   \\ \hline
+\# false positives 
+ \small{(internal + interaction)}
+                   & 0        & n         & 5 +   \\ \hline
 
 \clock{} with amb. 
-           classes & ??s           & 5h 42m    &  11m25  \\ \hline
-% \clock{} last rule & ??s           & ??s       & ?min ??s    \\ \hline
+           classes & ??s           & 5h 42m    &  ??m ?s  \\ \hline
 \clock{} no amb. 
-           classes & 10ish sec       & 14m 16s    & ?h ??min    \\ \hline
+           classes & 10ish sec       & ?m ?s    & 15m 57s    \\ \hline
 
@@
 \begin{itemize}
-\item[] 
-\texttt{REMOVE Podar  IF (0 Podar) (1 Adv) (2C Inf OR Enc) ;} \\
-\texttt{REMOVE Sentar IF (0 Sentar) (1C Inf OR Enc) ;}
+\item[]\begin{verbatim}
+REMOVE Podar  IF (0 Podar) (..) ;
+REMOVE Sentar IF (0 Sentar) (..) ;
+
+\end{verbatim}
 \end{itemize}
@@
 \hline
-              & 1 clitic + lemmas from grammar & 2 clitics + lemmas from grammar & 1 clitic + all readings 
+              & 1 clitic + 
+                lemmas from 
+                 grammar & 2 clitics 
+                           + lemmas from 
+                              grammar & 1 clitic +
+                                          all readings 
                                           from grammar    & all readings from grammar \\ \hline
@@
                        (7906 + 1588) & 6657 (4263 + 2394) & 2394  \\ \hline
-\# conflicts  & 214      & 214       & 24                & 24  \\ \hline
-\# interaction & 214      & 214       & 24                & 24  \\ \hline
+\# conflicts  & 214      & 214       & 22                & 22  \\ \hline
+\# internal/interaction & 211 + 3    & 211 + 3           & 19 + 3    & 19 + 3  \\ \hline
 
@@
 
-The second two columns are playing with the full set of readings from the grammar. For both of these, the number of reported conflicts was only 24.
+The second two columns are playing with the full set of readings from the grammar. For both of these, the number of reported conflicts was only 22.
+Given the preliminary nature of the results, we did not do a full analysis of all the 214 reported conflicts.
+We analysed all the 22, and
 
@@
 \begin{itemize}
-\item[] \item[]\texttt{SELECT "oma" + Gen IF .. (0C Nom) ;}
-\item[]\texttt{SELECT ADV (-1C AD) (NOT 0 PP) (NOT 1 A) ;}
+\item[]\begin{verbatim}
+"oma" SELECT Gen IF (..) (0C Nom) ;
+
+SELECT Adv IF (NOT 0 PP) (..) ;
+\end{verbatim}
 \end{itemize}
 
-Both of these are internal conflicts, which may not be trivial to see: 
-the first rule requires the target to be genitive and unambiguously nominative: 
-the problem in the first rule is that the target is required to be  the problem is that the definition of \texttt{PP} includes \texttt{adv} among others.
-With the sets expanded, this rule becomes \texttt{SELECT adv IF (NOT 0 pp|adv|adp|po|pr) ...}.
+Both of these are internal conflicts, which may not be trivial to see.
+The first rule requires the target to be genitive and unambiguously nominative; however, these two tags cannot combine in the same reading.
+As for the second rule, the definition of \texttt{PP} includes \texttt{adv} among others---with the sets expanded, this rule becomes \texttt{SELECT adv IF (NOT 0 pp|adv|adp|po|pr) ...}.
 
@@
 
-In this case, two rules together remove the target of the third, leaving no way for there to be adjective or noun.
+The first two rules together remove the target of the third, leaving no way for there to be adjective or noun.
 
@@
 A downside in the ambiguity classes is that generating them takes a long time, 
-and while 5-minute overhead may be acceptable when checking the full grammar,
+and while the overhead may be acceptable when checking the full grammar,
 it is hardly so when analysing just a handful of rules.

### 2016-03-08 paper/lrec/eval_futurework.tex
trying to explain the podar/sentar example nicer; maybe I should just remove it?

@@
 As for rule-internal conflicts, one was due to a misspelling in the list definition for personal pronouns, which rendered 5 rules ineffective.
-The other was about subreadings: the genitive \emph{s} is analysed as a subreading in the Apertium morphological analyser, but it appeared in 2 rules as the main reading.
+The other was about subreadings: the genitive \emph{s} is analysed as a subreading in the Apertium morphological analyser, but it appeared in one rule as the main reading.
 
@@
 
-These two rules both remove an adverb, but the first one has a broader condition:
-remove adverb if followed by a noun.
+% These two rules both remove an adverb, but the first one has a broader condition:
+% remove adverb if followed by a noun.
 In contrast, the second rule has a stricter condition: only remove adverb if it is preceded by a determiner, the adverb itself is ambiguous with an adjective, and followed by a noun.
-The problem is that the first rule removes the adverb in all possible cases, and the second will not have any chance to act.
-If the rules were in the opposite order, then there would be no problem.
+The problem is that that
+The first rule has a broader condition, 
+it removes the adverb in all possible cases 
+ removes the adverb in all possible cases, and the second will not have any chance to act. If the rules were in the opposite order, then there would be no problem.
 
@@
 As for the 25 interaction conflicts, there were only 9 distinct rules that rendered 25 other rules ineffective.
-In fact, we can reduce these 9 rules further into 3 different groups: 4 + 4 + 1, where the groups of 4 rules are variants of otherwise identical rule, with  combinations feminine singular, masculine singular, feminine plural and masculine plular.
+In fact, we can reduce these 9 rules further into 3 different groups: 4 + 4 + 1, where the groups of 4 rules are variants of otherwise identical rule, each with different gender and number.
 An example of such conflict is below (gender and number omitted for readability):
@@
 
-For a full list of found conflicts, see \todo{\url{actual://link.here/spanish/error.log}}
+For a full list of found conflicts, see error log\footnote{\url{https://github.com/inariksit/cgsat/raw/master/data/spa/conflicts.log}}.
 
@@
 The comments make it clear that the first rule is meant to disambiguate between \emph{poder} and \emph{podar}, and the second rule between \emph{sentir} and \emph{sentar}.
-But the rule does not mention anything about \emph{poder} nor \emph{sentir}; without the ambiguity class constraints, the SAT-solver could give a triggering sequence with any other reading in the target.
+But the rule does not mention anything about \emph{poder} nor \emph{sentir}.
+Even with the ambiguity class constraints, the SAT-solver only created an ambiguity where \emph{sentar} in 1st person plural is ambiguous with an anonymous 1st person plural reading.
+This does not reflect the reality, where these targets are only ambiguous with certain verbs, and in certain conjugated forms.
 
-This example shows the 
-shows a shortcoming in our way of handling the lexical conditions. 
-If the lemmas and word forms were integrated to the analyses in a more
-sensible way, then the ambiguity class constraints would create a
-sequence where \emph{podar} was ambiguous with \emph{poder} only in the
-word form \emph{puedo}.
+Another question is whether this level of detail is necessary. 
+After all, the grammar will be used to disambiguate real life texts, where \emph{podar} will likely not have any other ambiguities. 
+If our handling of lexical conditions was better, this sequence would, in fact, only be ambiguous with \emph{poder} in the right conjugation form.
 
-instead of randomly chosen readings.
+% shows a shortcoming in our way of handling the lexical conditions. 
+% If the lemmas and word forms were integrated to the analyses in a more
+% sensible way, then the ambiguity class constraints would create a
+% sequence where \emph{podar} was ambiguous with \emph{poder} only in the
+% word form \emph{puedo}.
+
+% instead of randomly chosen readings.
 So, even though this was not marked as a conflict, we can say that the weird output of the tool was unnecessary, and by better implementation, the tool would have created more realistic example sequence.

### 2016-03-08 paper/lrec/eval_futurework.tex
stuff

@@
 \centering
-\begin{tabular}{|p{2.80cm}|l|l|l|}
+\begin{tabular}{|p{2.8125cm}|l|l|l|}
 
@@
 \# readings        & 336           & 3905      & 1735    \\ \hline
-\# true positives 
-    \small{(internal + interaction)}
-                   & 7 (6 + 1)     & n         & m ( + 25)   \\ \hline
-\# false positives 
- \small{(internal + interaction)}
-                   & 0        & n         & 5 +   \\ \hline
+\# conflicts found \small{(internal + interaction)}
+                   & 7 (6 + 1)     & 44 (20 + 24)     & 42 (17 + 25) \\ \hline
+% \# false positives 
+%  \small{(internal + interaction)}
+%                    & 0             & 1 (1 + 0)       & 1 (1 + 0)  \\ \hline
 
 \clock{} with amb. 
-           classes & ??s           & 5h 42m    &  ??m ?s  \\ \hline
+           classes & ??s           & 1h 46m    &  Xm Ys  \\ \hline
+
 \clock{} no amb. 
-           classes & 10ish sec       & ?m ?s    & 15m 57s    \\ \hline
-
-%\# variables: last rule & 33915         & 247153            &    \\ \hline
+           classes & XXs        & 44m       & 16m      \\ \hline
 
@@
 rules with negations and complex set operations.
+
 %While this does not guarantee the lack of false negatives, it indicates some degree of reliability. 
@@
 We include two versions of the Spanish grammar in table~\ref{table:res}: in column 2, we added the lemmas and word forms as described in Section~\ref{sec:realistic_readings}, and in column 3, we just added each word form and lemma as individual readings, allowed to combine with any other reading.
-This latter version ran much faster, but also resulted in 5 false positives, all for rules that require a lemma in a condition.
-In addition, there were differences in the generated output; examples will be shown later in this section.
+This latter version ran much faster, but failed to detect two internal conflicts, both for rules that target a lemma in a condition.
+More interesting observation, the first version failed to detect a true interaction conflict, because it marked the two rules as internal conflicts.
+%In addition, there were differences in the generated output; examples will be shown later in this section.
 
-Again, we can classify the conflicts into internal and interaction.
 As an example of internal conflict, there are two rules that use \texttt{SET Cog = (np cog)}---the problem is that the tag \texttt{cog} does not exist in the Apertium dictionary for Spanish.

### 2016-03-08 paper/lrec/eval_futurework.tex
found nicer example for the imprecise rules part ^^ now spanish is pretty much done

@@
 \centering
-\begin{tabular}{|p{2.8125cm}|l|l|l|}
+%\begin{tabular}{|p{2.84cm}|p{1.2cm}|p{1.2cm}|p{1.25cm}|}
+\begin{tabular}{|p{2.84cm}|p{1cm}|p{1.15cm}|p{1.55cm}|}
 
 \hline
-                   & \textsc{nld}  & \textsc{spa}lem  & \textsc{spa}nolem \\ \hline
+                   & \textsc{nld}  & \textsc{spa}  & \textsc{spa}~\small{sep.~lem.} \\ \hline
 \# rules           & 59            & 279       & 279     \\ \hline
 \# readings        & 336           & 3905      & 1735    \\ \hline
-\# conflicts found \small{(internal + interaction)}
-                   & 7 (6 + 1)     & 44 (20 + 24)     & 42 (17 + 25) \\ \hline
+\# conflicts found \small{(internal~+~interaction)}
+                   & 7 \small{(6~+~1)}     & 44 \small{(20~+~24)}     & 43 \small{(18~+~25)} \\ \hline
 % \# false positives 
@@
 \clock{} with amb. 
-           classes & ??s           & 1h 46m    &  Xm Ys  \\ \hline
+           classes & ??s           & 1h 46m    &  23 min  \\ \hline
 
 \clock{} no amb. 
-           classes & XXs        & 44m       & 16m      \\ \hline
+           classes & XXs        & 44 min       & 16 min     \\ \hline
 
@@
 highest number of errors. The grammar we ran is like the one
-found in the Apertium repository, apart from two changes: we fixed
-some typos (capital O for 0) in order to make it compile, and
+found in the Apertium repository (linked on the previous page), 
+apart from two changes: we fixed some typos (capital O for 0) in order to make it compile, and
 commented out two rules that used regular expressions, because our
@@
 
-We include two versions of the Spanish grammar in table~\ref{table:res}: in column 2, we added the lemmas and word forms as described in Section~\ref{sec:realistic_readings}, and in column 3, we just added each word form and lemma as individual readings, allowed to combine with any other reading.
+We include two versions of the Spanish grammar in table~\ref{table:res}: in column \textsc{spa}, we added the lemmas and word forms as described in Section~\ref{sec:realistic_readings}, and in column \textsc{spa}~{\small sep.~lem.}, we just added each word form and lemma as individual readings, allowed to combine with any other reading.
 This latter version ran much faster, but failed to detect two internal conflicts, both for rules that target a lemma in a condition.
+% SELECT:pr_cnjadv_3 cnjadv IF (0 "después de") 
 More interesting observation, the first version failed to detect a true interaction conflict, because it marked the two rules as internal conflicts.
@@
 For a full list of found conflicts, see error log\footnote{\url{https://github.com/inariksit/cgsat/raw/master/data/spa/conflicts.log}}. 
+%It also contains a list of conflicts which were only found by one of the variants.
 
-In addition, there was a number of set definitions that were never
+In addition, the grammar contained a number of set definitions that were never
 used. Since VISL CG-3 already points out unused sets, we did not add such
@@
 For instance, the following definition requires the word to be all of
-the listed parts of speech at the same time---the grammar writer meant
+the listed parts of speech at the same time---most likely, the grammar writer meant 
 OR instead of +:
@@
 \item[]\begin{verbatim}
-REMOVE Podar  IF (0 Podar) (..) ;
 REMOVE Sentar IF (0 Sentar) (..) ;
-
+SELECT PP     IF (0 "estado") (..) ;
 \end{verbatim}
@@
 
-The comments make it clear that the first rule is meant to disambiguate between \emph{poder} and \emph{podar}, and the second rule between \emph{sentir} and \emph{sentar}.
-But the rule does not mention anything about \emph{poder} nor \emph{sentir}.
+The comments make it clear that the first rule is meant to disambiguate between \emph{sentar} and \emph{sentir}, but the rule does not mention anything about \emph{sentir}.
 Even with the ambiguity class constraints, the SAT-solver only created an ambiguity where \emph{sentar} in 1st person plural is ambiguous with an anonymous 1st person plural reading.
-This does not reflect the reality, where these targets are only ambiguous with certain verbs, and in certain conjugated forms.
+This does not reflect the reality, where the target is only ambiguous with certain verbs, and in certain conjugated forms.
+
+The second case is potentially more dangerous. 
+The word form $estado_{W}$ 
+can be either a noun ($estado_{L}$, `state'), or the past participle of the verb $estar_{L}$. 
+The condition, however, addresses the lemma of the noun, $estado_L$, whereas the lemma of the PP is $estar_{L}$.
+This means that, in theory, there can be a case where the condition to select the PP is already removed.
+As for now, the lexicon does not contain other ambiguities with the word form $estado_{W}$, but we could conceive of a scenario where someone adds e.g.
+a proper noun $Estado_{W,L}$ to the lexicon.
+Then, if some rule removes the lemma $estado_{L}$, the rule to select PP will not be able to trigger.
 
 Another question is whether this level of detail is necessary. 
-After all, the grammar will be used to disambiguate real life texts, where \emph{podar} will likely not have any other ambiguities. 
-If our handling of lexical conditions was better, this sequence would, in fact, only be ambiguous with \emph{poder} in the right conjugation form.
+After all, the grammar will be used to disambiguate real life texts, where \emph{sentar} will likely not have any other ambiguities, and aside from our contrived example, neither will \emph{estado}. 
+In fact, we are planning to change how we handle the lexical forms; with those changes, it will become clear whether the imprecision will result in potential errors, given the current lexicon.
 

### 2016-03-08 paper/lrec/eval_futurework.tex
cleanup of finnish section

@@
 Due to these complications, the results are not directly comparable, but we include Finnish in any case,
-to give an idea how our method scales up: both to grammars with more rules, and to more complex rules.
+to give an idea how our method scales up: both to more rules, and more complex rules.
 
@@
 In the first column, we included only possessive suffixes. In the second column, we included question clitics as well.
+Both of these readings include the 1588 lemmas and word forms from the grammar.
 In the third column, we included all the tag combinations specified in the grammar, and in the fourth, we took only those, ignoring the morphological lexicon completely.
 
-For the first two columns, there are a high number of internal conflicts. 
+For the first two variants, there are a high number of internal conflicts. 
 These are almost all due to nonexisting tags.
-The grammar was written in 1995, and updated by \newcite{pirinen2015}; but this high number of internal conflicts indicates that possibly something has gone wrong in the conversion, or in our expansion of the morphological lexicon.
+The grammar was written in 1995, and updated by \newcite{pirinen2015}; this high number of internal conflicts indicates that possibly something has gone wrong in the conversion, or in our expansion of the morphological lexicon.
 As for accuracy, adding the question clitics did not change anything: they were already included in some of the 1588 sets with word forms or lemmas, and that was enough for the SAT-solver to find models with question clitics.
@@
 
-The second two columns are playing with the full set of readings from the grammar. For both of these, the number of reported conflicts was only 22.
+The second two variants are playing with the full set of readings from the grammar. For both of these, the number of reported conflicts was only 22.
 Given the preliminary nature of the results, we did not do a full analysis of all the 214 reported conflicts.

### 2016-03-08 paper/lrec/eval_futurework.tex
I'm starting to believe that this will be a paper

@@
 
-For a full list of found conflicts, see error log\footnote{\url{https://github.com/inariksit/cgsat/raw/master/data/spa/conflicts.log}}. 
-%It also contains a list of conflicts which were only found by one of the variants.
+For a full list of found conflicts, see the log\footnote{\url{https://github.com/inariksit/cgsat/raw/master/data/spa/conflicts.log}} of the running our program. 
+In addition, the log contains a list of conflicts which were only found by one of the variants.
 
@@
 
-Finally, we were not able to create ambiguity class constraints---expanding the Finnish morphological lexicon results in \todo{100s of gigabytes?} of word forms, which is simply too big for our method to work.
+Finally, we were not able to create ambiguity class constraints---expanding the Finnish morphological lexicon results in 100s of gigabytes of word forms, which is simply too big for our method to work.
 For future development, we will see if it is possible to manipulate the finite automata directly, instead of relying on the output in text.
@@
 Both of these readings include the 1588 lemmas and word forms from the grammar.
-In the third column, we included all the tag combinations specified in the grammar, and in the fourth, we took only those, ignoring the morphological lexicon completely.
+In the third column, we included all the tag combinations specified in the grammar, and in the fourth, we took only those, ignoring the morphological lexicon.
 
-For the first two variants, there are a high number of internal conflicts. 
+The first two variants contain a high number of internal conflicts. 
 These are almost all due to nonexisting tags.
-The grammar was written in 1995, and updated by \newcite{pirinen2015}; this high number of internal conflicts indicates that possibly something has gone wrong in the conversion, or in our expansion of the morphological lexicon.
+The grammar was written in 1995, and updated by \newcite{pirinen2015}; such a high number of internal conflicts indicates that possibly something has gone wrong in the conversion, or in our expansion of the morphological lexicon.
 As for accuracy, adding the question clitics did not change anything: they were already included in some of the 1588 sets with word forms or lemmas, and that was enough for the SAT-solver to find models with question clitics.
@@
 Given the preliminary nature of the results, we did not do a full analysis of all the 214 reported conflicts.
-We analysed all the 22, and
+Out of the 22, we found 17 of them as true conflicts,
+ but 5 seemed to be caused by our handling of rules with \verb!*!. 
+All of these 5 rules contained a LINK and multiple \verb!*!s. On a positive note, our naive handling of the \verb!*! seems to cover the simplest cases.
 
-To show some examples of true positives:
+Some examples of true positives are shown in the following.
 
@@
 \item[]\begin{verbatim}
-REMOVE A (0 DER) ; 
-REMOVE N (0 DER) ; 
-REMOVE A/N (0 DER) ; 
+REMOVE A (0 Der) ; 
+REMOVE N (0 Der) ; 
+REMOVE A/N (0 Der) ; 
 \end{verbatim}

### 2016-03-08 paper/lrec/implementation.tex
I'm starting to believe that this will be a paper

@@
 
-If the context contains a \verb!*!,
+If the context contains a \verb!*! (context word can be anywhere),
 we may need to make an approximation of $w(r)$ which may result in false negatives later on in the analysis.
@@
 Thus we add special entries for these: in addition to the anonymous
- \texttt{<det><pos><mfn><pl>} reading, we add \texttt{<"zijn"><det><pos><mfn><pl>}. 
+``det pos mfn pl'' reading, we add ``\emph{zijn} det pos mfn pl''. 
 The lemma is treated as just another tag.
@@
 Our approach is similar to \newcite{cutting_etal92}, who use ambiguity classes instead of distinct word forms, in order to reduce the number of parameters in a Hidden Markov Model.
-They take advantage of the fact that they don't have to model ``bear'' and ``wish'' as separate entries, but they can just reduce it to ``word that can be ambiguous between noun and verb'', and use it as a parameter in their HMM.
+They take advantage of the fact that they don't have to model ``bear'' and ``wish'' as separate entries, but they can just reduce it to ``word that can be ambiguous between noun and verb'', and use that as a parameter in their HMM.
 %We can do a similar thing by saving a list of words with each ambiguity class.

### 2016-03-08 paper/lrec/eval_futurework.tex
cleanup & more examples to future work

@@
 In contrast, the second rule has a stricter condition: only remove adverb if it is preceded by a determiner, the adverb itself is ambiguous with an adjective, and followed by a noun.
-The problem is that that
-The first rule has a broader condition, 
-it removes the adverb in all possible cases 
- removes the adverb in all possible cases, and the second will not have any chance to act. If the rules were in the opposite order, then there would be no problem.
-
-%%% Nevermind, this wasn't a conflict, my program was just buggy :-P
-
-%\begin{itemize}
-%\item[] 
-%\begin{verbatim}
-%SELECT DetPosNotZijn IF (1 Noun) ;
-%SELECT DetPos IF (-1 (vbser pres p3 sg)) 
-%                 (0 "zijn") (1 Noun);
-%\end{verbatim}
-%\end{itemize}
-
-% SELECT DetPosNotZijn IF (1 Noun);
-%If the first rule fires, then the condition of the second rule cannot be true: there can never be a ``zijn'' in the target, because the previous rule has selected everything that is a possessive determiner and not ``zijn''.
-%So, the first rule must not fire. It cannot be because condition does not hold: the condition (1 Noun) is shared with the second rule, and
+The problem is that the first rule has a broader condition than the second.
+The first rule removes the adverb in all possible cases where the second would, and the second will not have any chance to act. 
+If the rules were in the opposite order, then there would be no problem.
 
@@
 apart from two changes: we fixed some typos (capital O for 0) in order to make it compile, and
-commented out two rules that used regular expressions, because our
-solution does not handle them properly; see more on Section~\ref{sec:conclusion}.
+commented out two rules that used regular expressions, because we did not implement the support for them yet.
 
 We include two versions of the Spanish grammar in table~\ref{table:res}: in column \textsc{spa}, we added the lemmas and word forms as described in Section~\ref{sec:realistic_readings}, and in column \textsc{spa}~{\small sep.~lem.}, we just added each word form and lemma as individual readings, allowed to combine with any other reading.
-This latter version ran much faster, but failed to detect two internal conflicts, both for rules that target a lemma in a condition.
+This latter version ran much faster, but failed to detect an internal conflict for a rule that targets a lemma in a condition.
 % SELECT:pr_cnjadv_3 cnjadv IF (0 "después de") 
-More interesting observation, the first version failed to detect a true interaction conflict, because it marked the two rules as internal conflicts.
-%In addition, there were differences in the generated output; examples will be shown later in this section.
+Surprisingly, the first version failed to detect a true interaction conflict, because it marked the two rules both as internal conflicts.
 
@@
 \centering
-\begin{tabular}{|p{2.5cm}|p{3cm}|p{3cm}|p{3cm}|p{3cm}|}
+\begin{tabular}{|p{3.15cm}|p{2.55cm}|p{2.55cm}|p{2.7cm}|p{2.1cm}|}
 
@@
                                           from grammar    & all readings from grammar \\ \hline
-\# readings   & 5851 (4263 
-                 + 1588) & 9494
-                       (7906 + 1588) & 6657 (4263 + 2394) & 2394  \\ \hline
-\# conflicts  & 214      & 214       & 22                & 22  \\ \hline
-\# internal/interaction & 211 + 3    & 211 + 3           & 19 + 3    & 19 + 3  \\ \hline
+\# readings   & 5851 (4263~+~1588)
+                       & 9494
+                       (7906~+~1588) & 6657 (4263~+~2394) & 2394  \\ \hline
+\#~conflicts  & 214    & 214         & 22                 & 22 \\
+(internal + 
+ interaction) & (211~+~3) & (211~+~3) & (19~+~3)        &  (19~+~3)  \\ \hline
 
-\# true 
-   positives  & n        & m         & o                & p  \\ \hline
-\# false 
-   positives  & p        & q         & r                & s   \\ \hline
-\clock{}all 
-       rules & 12min     & 23min     & 12min &  4min 40s \\ \hline
+\clock{} all 
+       rules & Xh 12min     & Xh 23min   & Xh 12min &  Xh 4min \\ \hline
 
@@
 The conflict lies in the fact that \texttt{Comma} is a subset of \texttt{Comma/N/Pron/Q}:
-there is no way to trigger the second without placing a comma in position -1, and thereby triggering the first rule.
-
-% All in all, using readings from grammar directly may be useful in order to find just the , sometimes you can have a typo, like in \emph{l\"{a}hetii}:
-
-% \begin{verbatim}
-% "w2"
-%         w2<"se"><ela>
-%         w2<"<lähetii>"><vblex>
-% \end{verbatim}
-
-% The second part concerns the tools in use, specifically the Apertium morphological lexicon.
-We noticed, only after running the experiments, that the expansion of the lexicon does not work properly: for instance, some nominal forms of the verbs were reported as nouns instead of verbs.
-% We reported the errors to the developers of the lexicon, but as there was no time to fix them, we used the readings that we got from expanding the lexicon, and manually added either noun, verb or adjective tag where it was missing.
-
-% kosimisiin:kosia<n><pl><ill>
-% kosimattomuudekseen:kosia<sg><tra><pxsp3>
-
-% Secondly, we added ambiguity class constraints for only
-% \emph{intraparadigmatic ambiguities}: homonymous wordforms belonging
-% to the same lexeme, in a way that is systematic to the whole
-% paradigm. We reduced the morphological lexicon to contain only one
-% example of each inflection paradigm, and then expanded it: with all
-% the clitics, this resulted in \todo{n} word forms. Then we applied the
-% normal method of creating the ambiguity classes, and specified some
-% default cases manually, such as punctuation and word boundaries.
+there is no way to trigger the second rule without placing a comma in position -1, and thereby triggering the first rule.
 
@@
 
-This may also be feasible to support regular expressions, in a nicer
-way than just spitting out the regex. (INARI: You said MORE ON THAT here, good because I don't understand what you want to say here :)
-
 \paragraph{Full expressivity of CG-3}
@@
 \paragraph{Support for grammar writers}
-One additional feature is to suggest reformattings for a rule. Recall
+As mentioned earlier, we could support additional checks for common issues,
+such as conditions that concern the target word.
+Another possible feature is to suggest reformattings for a rule. Recall
 figure~\ref{fig:regroup} from the introduction; in that case, the

### 2016-03-08 paper/lrec/eval_futurework.tex
add GF reference

@@
 \clock{} all 
-       rules & Xh 12min     & Xh 23min   & Xh 12min &  Xh 4min \\ \hline
+       rules & Xh Ymin     & Xh Ymin   & Xh Ymin &  Xh Ymin \\ \hline
 
@@
 Finally, we would like to investigate logic-based methods for analysing other grammar
-formalisms, for example Grammatical Framework \cite{gf}. (INARI: can you add a reference here?)
+formalisms, for example Grammatical Framework \cite{ranta2010gfbook}.
 

### 2016-03-08 paper/lrec/eval_futurework.tex
update finnish running time

@@
 \clock{} all 
-       rules & Xh Ymin     & Xh Ymin   & Xh Ymin &  Xh Ymin \\ \hline
+       rules & 2h 2min     & Xh Ymin   & Xh Ymin &  Xh Ymin \\ \hline
 
@@
 The first rule requires the target to be genitive and unambiguously nominative; however, these two tags cannot combine in the same reading.
-As for the second rule, the definition of \texttt{PP} includes \texttt{adv} among others---with the sets expanded, this rule becomes \texttt{SELECT adv IF (NOT 0 pp|adv|adp|po|pr) ...}.
+As for the second rule, the definition of \texttt{PP} includes \texttt{adv} among others---with the sets expanded, this rule becomes \texttt{SELECT adv IF (NOT 0 pp|adv|adp|po|pr) (...)}.
 
@@
 
+This is the same pattern we have already seen before, but with a set of rules as the reason for conflict.
 The first two rules together remove the target of the third, leaving no way for there to be adjective or noun.

### 2016-03-09 paper/lrec/eval_futurework.tex
grmbl redo the numbers for spanish

@@
 types introduced in CG-3, and only tested \textsc{remove} and \textsc{select} rules.
-The results for Dutch and Spanish are shown in table~\ref{table:res},
-and the results for Finnish in table~\ref{table:resFin}.
+The results for Dutch and Spanish are shown in Table~\ref{table:res},
+and the results for Finnish in Table~\ref{table:resFin}.
 
@@
 \centering
-%\begin{tabular}{|p{2.84cm}|p{1.2cm}|p{1.2cm}|p{1.25cm}|}
 \begin{tabular}{|p{2.84cm}|p{1cm}|p{1.15cm}|p{1.55cm}|}
+%\begin{tabular}{|p{2.9cm}|p{0.95cm}|p{1.15cm}|p{1.55cm}|}
 
@@
 \# readings        & 336           & 3905      & 1735    \\ \hline
-\# conflicts found \small{(internal~+~interaction)}
-                   & 7 \small{(6~+~1)}     & 44 \small{(20~+~24)}     & 43 \small{(18~+~25)} \\ \hline
+\# conflicts found 
+ with amb. classes & 7             & 45        & X      \\ 
+{\small (internal~+~interaction)}
+                   & {\small
+                      (6~+~1)}     & {\small 
+                                    (21~+~24)} & {\small (X~+~24)} \\ \hline
+
+\# conflicts found 
+ no amb. classes   & 7             & X        & 43      \\ 
+{\small (internal~+~interaction)}
+                   & {\small
+                      (6~+~1)}     & {\small 
+                                    (18~+~25)} & {\small (18~+~25)} \\ \hline
 % \# false positives 
@@
 \clock{} with amb. 
-           classes & ??s           & 1h 46m    &  23 min  \\ \hline
+           classes & 7 s        & 1h 46m    &  23 min  \\ \hline
 
 \clock{} no amb. 
-           classes & XXs        & 44 min       & 16 min     \\ \hline
+           classes & 3 s        & 44 min       & 16 min     \\ \hline
 
@@
 grammars, we were able to verify manually that the detected rules were
-true positives.
+true positives, save for one case, where we took a shortcut in the Spanish grammar.
 %The results for the Finnish grammar are inconclusive, 
@@
 just that there is a reading with any determiner, and a reading with any \emph{zijn} in the cohort. 
-Can we catch this imprecise formulation with our tool? The example word constructed by the SAT-solver may as likely be one of the following:
+Can we catch this imprecise formulation with our tool?
+The SAT-solver will not mark it as a conflict (which is the correct behaviour).
+But if we ask it to generate an example sequence, the target word may be any of the following options:
 
@@
 
-We include two versions of the Spanish grammar in table~\ref{table:res}: in column \textsc{spa}, we added the lemmas and word forms as described in Section~\ref{sec:realistic_readings}, and in column \textsc{spa}~{\small sep.~lem.}, we just added each word form and lemma as individual readings, allowed to combine with any other reading.
-This latter version ran much faster, but failed to detect an internal conflict for a rule that targets a lemma in a condition.
-% SELECT:pr_cnjadv_3 cnjadv IF (0 "después de") 
-Surprisingly, the first version failed to detect a true interaction conflict, because it marked the two rules both as internal conflicts.
+We include two versions of the Spanish grammar in Table~\ref{table:res}: in column \textsc{spa}, we added the lemmas and word forms as described in Section~\ref{sec:realistic_readings}, and in column \textsc{spa}~{\small sep.~lem.}, we just added each word form and lemma as individual readings, allowed to combine with any other reading.
+This latter version ran much faster, but failed to detect an internal conflict for one rule, and reported a false positive for another. The numbers look the same
+% FAIL TO DETECT: SELECT:pr_cnjadv_3 cnjadv IF (0 "después de") 
+% FALSE POSITIVE: SELECT:este_1 det IF (0 "este" + (m sg)|(m sp)|(mf sg)|(mf sp))
+
+Interestingly, the version $with$ ambiguity classes fails to detect a true interaction conflict, because it---correctly---found an interaction conflict in one of the rules.
 
@@
 
-For a full list of found conflicts, see the log\footnote{\url{https://github.com/inariksit/cgsat/raw/master/data/spa/conflicts.log}} of the running our program. 
+For a full list of found conflicts, see the log\footnote{\url{https://github.com/inariksit/cgsat/raw/master/data/spa/conflicts.log}} of running our program. 
 In addition, the log contains a list of conflicts which were only found by one of the variants.
 
-In addition, the grammar contained a number of set definitions that were never
+In addition, the grammar contains a number of set definitions that were never
 used. Since VISL CG-3 already points out unused sets, we did not add such
@@
 conflicting. We noticed the error by accident, when the program
-offered the reading \texttt{w2<n><np><adj><det><preadv><adv><prn>}
+offered the reading \texttt{w2<n><adj><det><preadv><adv><prn>}
 in an example sequence meant for another rule.
@@
 REMOVE Sentar IF (0 Sentar) (..) ;
+
 SELECT PP IF (0 "estado") (..) ;
@@
 As for now, the lexicon does not contain other ambiguities with the word form $estado_{W}$, but we could conceive of a scenario where someone adds e.g.
-a proper noun $Estado_{W,L}$ to the lexicon.
+a proper noun $Estado_{L}$ to the lexicon.
 Then, if some rule removes the lemma $estado_{L}$, the rule to select PP will not be able to trigger.
@@
 \clock{} all 
-       rules & 2h 2min     & Xh Ymin   & Xh Ymin &  Xh Ymin \\ \hline
+       rules & 2h 2min     & 4h 27min    & Xh Ymin &  Xh Ymin \\ \hline
 
@@
 
-The results for the Finnish grammar are shown separately, in table~\ref{table:resFin}.
+The results for the Finnish grammar are shown separately, in Table~\ref{table:resFin}.
 We encountered a number of difficulties and used a few shortcuts, which we did not need for the other grammars---most importantly, not using the ambiguity class constraints.
@@
 \paragraph{Results}
-The results are shown in table~\ref{table:resFin}.
+The results are shown in Table~\ref{table:resFin}.
 In the first column, we included only possessive suffixes. In the second column, we included question clitics as well.
@@
 Out of the 22, we found 17 of them as true conflicts,
- but 5 seemed to be caused by our handling of rules with \verb!*!. 
-All of these 5 rules contained a LINK and multiple \verb!*!s. On a positive note, our naive handling of the \verb!*! seems to cover the simplest cases.
+ but 5 seemed to be caused by our handling of rules with \verb!*!: all of these 5 rules contain a LINK and multiple \verb!*!s.
+On a positive note, our naive handling of the \verb!*!
+seems to cover the simplest cases.
 
@@
 Another possible feature is to suggest reformattings for a rule. Recall
-figure~\ref{fig:regroup} from the introduction; in that case, the
+Figure~\ref{fig:regroup} from the introduction; in that case, the
 original rule was written by the original author, and another

### 2016-03-09 paper/lrec/eval_futurework.tex
more reporting on the newly found spanish error

@@
 \# conflicts found 
- no amb. classes   & 7             & X        & 43      \\ 
+ w/o~amb. classes   & 7             & X        & 43      \\ 
 {\small (internal~+~interaction)}
@@
 
-\clock{} no amb. 
+\clock{} w/o amb. 
            classes & 3 s        & 44 min       & 16 min     \\ \hline
@@
 We include two versions of the Spanish grammar in Table~\ref{table:res}: in column \textsc{spa}, we added the lemmas and word forms as described in Section~\ref{sec:realistic_readings}, and in column \textsc{spa}~{\small sep.~lem.}, we just added each word form and lemma as individual readings, allowed to combine with any other reading.
-This latter version ran much faster, but failed to detect an internal conflict for one rule, and reported a false positive for another. The numbers look the same
+This latter version ran much faster, but failed to detect an internal conflict for one rule, and reported a false positive for another. 
 % FAIL TO DETECT: SELECT:pr_cnjadv_3 cnjadv IF (0 "después de") 
@@
 
-Interestingly, the version $with$ ambiguity classes fails to detect a true interaction conflict, because it---correctly---found an interaction conflict in one of the rules.
+When we added ambiguity class constraints, we found three more true conflicts, all internal.
+Interestingly, the version \emph{with} ambiguity classes fails to detect an interaction conflict, which the simpler version reports, because one of the rules was first detected as internal conflict.
 
@@
 It is likely that this grammar has been written for an earlier version, where such tags have been in place.
+One of the conflicts that was only caught by the ambiguity class constraints had the condition \texttt{IF (1 Comma) (..) (1 CnjCoo)}.
+The additional constraints correctly prevent the SAT-solver to make comma ambiguous with anything else.
 

### 2016-03-09 paper/lrec/eval_futurework.tex
small tweaks

--- 2016-03-09 paper/lrec/eval_futurework.tex
@@
 \hline
-                   & \textsc{nld}  & \textsc{spa}  & \textsc{spa}~\small{sep.~lem.} \\ \hline
+                   & \textsc{nld}  & \textsc{spa}  & \textsc{spa}~\textsuperscript{sep.~lem.} \\ \hline
 \# rules           & 59            & 279       & 279     \\ \hline
 \# readings        & 336           & 3905      & 1735    \\ \hline
-\# conflicts found 
- with amb. classes & 7             & 45        & X      \\ 
+\# true~positives~\textsuperscript{AC}% with~amb.~classes
+                   & 7             & 45        & 44      \\ 
 {\small (internal~+~interaction)}
@@
                       (6~+~1)}     & {\small 
-                                    (21~+~24)} & {\small (X~+~24)} \\ \hline
+                                    (21~+~24)} & {\small (20~+~24)} \\ \hline
 
-\# conflicts found 
- no amb. classes   & 7             & X        & 43      \\ 
+\# true~positives~\textsuperscript{no AC}  & 7             & 43        & 42      \\ 
 {\small (internal~+~interaction)}
@@
                       (6~+~1)}     & {\small 
-                                    (18~+~25)} & {\small (18~+~25)} \\ \hline
-% \# false positives 
-%  \small{(internal + interaction)}
-%                    & 0             & 1 (1 + 0)       & 1 (1 + 0)  \\ \hline
+                                    (18~+~25)} & {\small (17~+~25)} \\ \hline
+\# false positives 
+                   & 0             & 0        & 1  \\ \hline
 
@@
 The experiments revealed problems in all grammars. For the smaller
-grammars, we were able to verify manually that the detected rules were
-true positives, save for one case, where we took a shortcut in the Spanish grammar.
+grammars, we were able to verify manually that nearly all detected conflicts were
+true positives---we found one false positive and one false negative,
+when using a shortcut for the Spanish grammar. 
 %The results for the Finnish grammar are inconclusive, 
@@
 rules with negations and complex set operations.
-
-%While this does not guarantee the lack of false negatives, it indicates some degree of reliability. 
 As the tool matures and we add new features, a more in-depth analysis
@@
 In contrast, the second rule has a stricter condition: only remove adverb if it is preceded by a determiner, the adverb itself is ambiguous with an adjective, and followed by a noun.
-The problem is that the first rule has a broader condition than the second.
-The first rule removes the adverb in all possible cases where the second would, and the second will not have any chance to act. 
+These two rules share a target: both remove an adverb.
+The problem is that the first rule has a broader condition than the second, hence the second will not have any chance to act. 
 If the rules were in the opposite order, then there would be no problem.
@@
 that the input would have determiner and \emph{zijn} in the same reading:
-just that there is a reading with any determiner, and a reading with any \emph{zijn} in the cohort. 
+just that there is a reading with any determiner, and a reading with any \emph{zijn} among all the readings.
 Can we catch this imprecise formulation with our tool?
 The SAT-solver will not mark it as a conflict (which is the correct behaviour).
-But if we ask it to generate an example sequence, the target word may be any of the following options:
+But if we ask it to generate an example sequence, the target word may be either of the following options:
 
@@
 
-We include two versions of the Spanish grammar in Table~\ref{table:res}: in column \textsc{spa}, we added the lemmas and word forms as described in Section~\ref{sec:realistic_readings}, and in column \textsc{spa}~{\small sep.~lem.}, we just added each word form and lemma as individual readings, allowed to combine with any other reading.
-This latter version ran much faster, but failed to detect an internal conflict for one rule, and reported a false positive for another. The numbers look the same
+We include two versions of the Spanish grammar in Table~\ref{table:res}: in column \textsc{spa}, we added the lemmas and word forms as described in Section~\ref{sec:realistic_readings}, and in column \textsc{spa}\textsuperscript{sep.~lem.}, we just added each word form and lemma as individual readings, allowed to combine with any other reading.
+This latter version ran much faster, but failed to detect an internal conflict for one rule, and reported a false positive for another. 
 % FAIL TO DETECT: SELECT:pr_cnjadv_3 cnjadv IF (0 "después de") 
@@
 
-Interestingly, the version $with$ ambiguity classes fails to detect a true interaction conflict, because it---correctly---found an interaction conflict in one of the rules.
+When we added ambiguity class constraints, we found three more internal conflicts.
+Interestingly, the version with ambiguity classes fails to detect an interaction conflict, which the simpler version reports, because one of the rules is first detected as an internal conflict. 
+We think that neither of these versions is a false positive or negative; it is just a matter of priority. Sometimes we prefer to know that the rule cannot apply, given the current lexicon. 
+However, we may know that the lexicon is about to be updated, and would rather learn about all potential interaction conflicts.
 
-As an example of internal conflict, there are two rules that use \texttt{SET Cog = (np cog)}---the problem is that the tag \texttt{cog} does not exist in the Apertium dictionary for Spanish.
+As an example of internal conflict, there are two rules that use \texttt{SET Cog = (np cog)}: the problem is that the tag ``cog'' does not exist in the lexicon.
 As another example, four rules require a context word tagged as NP with explicit number, but the lexicon does not indicate any number with NPs.
 It is likely that this grammar has been written for an earlier version, where such tags have been in place.
+One of the conflicts that was only caught by the ambiguity class constraints had the condition \texttt{IF (1 Comma) (..) (1 CnjCoo)}.
+The additional constraints correctly prevent a comma from being ambiguous with anything else.
 
@@
 
-For a full list of found conflicts, see the log\footnote{\url{https://github.com/inariksit/cgsat/raw/master/data/spa/conflicts.log}} of running our program. 
-In addition, the log contains a list of conflicts which were only found by one of the variants.
+For a full list of found conflicts, see the annotated log\footnote{\url{https://github.com/inariksit/cgsat/raw/master/data/spa/conflicts.log}} of running our program.
 
@@
 set definitions that are not used in any rule.
-%If a set definition is impossible, using it in a rule makes the rule have an internal conflict.
 For instance, the following definition requires the word to be all of
@@
 writer intended.
-
-% \begin{itemize}
-% \item[] 
-% \begin{verbatim}
-% REMOVE Vblex IF (0 "como") (*-1 Vblex BARRIER Cnj_Rel) ;
-% \end{verbatim}
-% \end{itemize}
-
-% Based on the comments in the file, the grammar writer wants to
-% disambiguate the word form ``como'', which can be either a
-% preposition, adverb or the first person singular of the verb
-% ``comer''. 
-% However, the rule here addresses the \emph{lemma} ``como'' instead of the
-% word form, hence the SAT-solver cannot create a solution where
-% \texttt{vblex} and ``como'' as a lemma are in the same
-% reading. 
-% A precise definition for this rule would target \texttt{vblex "<como>"} 
-% or \texttt{vblex "comer"}, so that it becomes obvious that those two
-% are supposed to be in the same reading.
-
-% Currently, as all the word forms are allowed to combine with anything,
-% the ambiguity class constraint does not point this
-% The automated 0-to-target conversion could come in handy here too.
-% However, if the
-
-% TODO An example of the same but in condition. that the ambiguity class constraint, when
-% implemented properly, should find out.
 
@@
 Another question is whether this level of detail is necessary. 
-After all, the grammar will be used to disambiguate real life texts, where \emph{sentar} will likely not have any other ambiguities, and aside from our contrived example, neither will \emph{estado}. 
+After all, the grammar will be used to disambiguate real life texts, where neither \emph{sentar} nor \emph{estado} are likely to have any other ambiguities.
 In fact, we are planning to change how we handle the lexical forms; with those changes, it will become clear whether the imprecision will result in potential errors, given the current lexicon.
-
-% shows a shortcoming in our way of handling the lexical conditions. 
-% If the lemmas and word forms were integrated to the analyses in a more
-% sensible way, then the ambiguity class constraints would create a
-% sequence where \emph{podar} was ambiguous with \emph{poder} only in the
-% word form \emph{puedo}.
-
-% instead of randomly chosen readings.
-So, even though this was not marked as a conflict, we can say that the weird output of the tool was unnecessary, and by better implementation, the tool would have created more realistic example sequence.
 
@@
 \clock{} all 
-       rules & 2h 2min     & 4h 27min    & Xh Ymin &  Xh Ymin \\ \hline
+       rules & Xh Ymin     & 8h 28min    & 4h 27min &  2h 29min \\ \hline
 

### 2016-03-09 paper/lrec/lrec2016_listenmaa_claessen.tex
add anonymous reviewers to acknowledgments

@@
 The aim is for grammar writers to be able to automatically diagnose, and then manually improve their grammars.
-Our method works by automatically translating the given grammar into logical constraints that are analysed by a SAT-solver.
+Our method works by translating the given grammar into logical constraints that are analysed by a SAT-solver.
 We have evaluated our analysis on a number of non-trivial grammars and found inconsistencies.
@@
 \section*{Acknowledgments}
-We thank Eckhard Bick for the idea to apply SAT to CG analysis, and Francis Tyers for insightful discussions.
+We thank Eckhard Bick for the idea to apply SAT to CG analysis, and Francis Tyers for insightful discussions, as well as our anonymous reviewers for comments and suggestions.
 

### 2016-03-10 paper/lrec/eval_futurework.tex
last day fixes

@@
 
-\subsection{Dutch} The Dutch grammar had two kinds of errors: rule-internal, and rule interaction.
+\subsection{Dutch} The Dutch grammar had two kinds of errors: rule-internal and rule interaction.
 As for rule-internal conflicts, one was due to a misspelling in the list definition for personal pronouns, which rendered 5 rules ineffective.
-The other was about subreadings: the genitive \emph{s} is analysed as a subreading in the Apertium morphological analyser, but it appeared in one rule as the main reading.
+The other was about subreadings: the genitive \emph{s} is analysed as a subreading in the Apertium morphological lexicon, but it appeared in one rule as the main reading.
 
@@
 \begin{itemize}
-\item[] 
+\item[\textsc{d$_1$.}] 
 \begin{verbatim}REMOVE Adv IF (1 N) ;
@@
 \begin{itemize}
-\item[] 
+\item[\textsc{d$_2$.}] 
 \texttt{SELECT DetPos IF (-1 (vbser pres p3 sg)) (0 "zijn") (1 Noun);}
@@
 As per VISL CG-3, the condition \texttt{(0 "zijn")} does not require
-that the input would have determiner and \emph{zijn} in the same reading:
-just that there is a reading with any determiner, and a reading with any \emph{zijn} among all the readings.
+ \emph{zijn} to be in the same reading with the target \emph{DetPos}.
+It just means that at index 0,
+there is a reading with any possessive determiner, and a reading with any \emph{zijn}.
+However, the intended action is to select a ``det pos \emph{zijn}'' all in one reading;
+this is expressed as \texttt{SELECT DetPos + "zijn"}.
+In contrast, the 0-condition in example~\textsc{d$_1$} is used correctly:
+the adjective and the adverb are supposed to be in different readings.
+
 Can we catch this imprecise formulation with our tool?
@@
 
-Incidentally, the first example of a rule conflict had a similar
-requirement, but used correctly: the intended semantics was indeed
-that the adjective and the adverb are in different readings. However,
-in this case, the precise way to express the target is \texttt{SELECT DetPos + "zijn"}.
-
-We found the same kind of definitions in many other rules and grammars.
+We found the same kind of definition in many other rules and grammars.
 To catch them more systematically, we could add a feature that alerts in all cases where a condition with 0 is used.
@@
 \begin{itemize}
-\item[] 
+\item[\textsc{s$_{1}$.}] 
 \begin{verbatim}
@@
 \begin{itemize}
-\item[] 
+\item[\textsc{s$_2$.}] 
 \texttt{SET NP\_Member = N + A + Det + PreAdv + Adv + Pron ;}
@@
 \begin{itemize}
-\item[]\begin{verbatim}
+\item[\textsc{s$_3$.}] \begin{verbatim}
 REMOVE Sentar IF (0 Sentar) (..) ;
@@
 \clock{} all 
-       rules & Xh Ymin     & 8h 28min    & 4h 27min &  2h 29min \\ \hline
+       rules (approx.) & \~{}4h 30min     & \~{}9h 30min    & \~{}7h 45min &  \~{}2h 30min \\ \hline
 
@@
 As for accuracy, adding the question clitics did not change anything: they were already included in some of the 1588 sets with word forms or lemmas, and that was enough for the SAT-solver to find models with question clitics.
-We left the result in the table just to demonstrate the change in speed.
+We left the result in the table just to demonstrate the change in the running time.
 
@@
 \begin{itemize}
-\item[]\begin{verbatim}
+\item[\textsc{f$_1$}.]\begin{verbatim}
 "oma" SELECT Gen IF (..) (0C Nom) ;
@@
 \begin{itemize}
-\item[]\begin{verbatim}
+\item[\textsc{f$_2$}.]\begin{verbatim}
 REMOVE A (0 Der) ; 
@@
 \begin{itemize}
-\item[]\begin{verbatim}
+\item[\textsc{f$_3$.}]\begin{verbatim}
 SELECT .. IF (-1 Comma/N/Pron/Q) ;
@@
 In addition, the ambiguity class constraints increase the time.
+The times in the Finnish table are not entirely comparable with each other: we were forced to run the tests in smaller batches, and it is possible that there are different overheads, unrelated to the size of the SAT-problem, from testing 50 or 500 rules at a time.
 
@@
 it is hardly so when analysing just a handful of rules. 
+We are working on an option to store and reuse the ambiguity class constraints.
 

### 2016-03-10 paper/lrec/implementation.tex
last day fixes

@@
 We can do this by extracting this information from a lexicon, but there are other ways too.
-In our experiments, the number of readings has ranged from about 300 to about 6000.
+In our experiments, the number of readings has ranged from about 300 to about 9000.
 
@@
 If the context contains a \verb!*! (context word can be anywhere),
-we may need to make an approximation of $w(r)$ which may result in false negatives later on in the analysis.
+we may need to make an approximation of $w(r)$, which may result in false positives or negatives later on in the analysis.
 %we create sentences that are up to 3 words wider than the context without \verb!*!, and try them all.
@@
 The number of verb readings without clitics is 213, and with clitics 1572.
-With the previously mentioned approach, we would have to double 1572 entries for each verb lemma. Even ignoring the clitics, each verb lemma still adds 213 new readings.
+With the previously mentioned approach, we would have to duplicate 1572 entries for each verb lemma. Even ignoring the clitics, each verb lemma still adds 213 new readings.
 
@@
 ``pron rel'' and ``nom'' separately.
+
 In addition, we found that the tag lists in the grammars sometimes
@@
 in a subreading. If we accept those lists as readings, we will
-generate symbolic sentences that are impossible, and won't discover
+generate symbolic sentences that are impossible, and not discover
 the bug in the grammar.
+However, if we are primarily interested in rule interaction, then using
+the underspecified readings from the grammar may be an adequate solution.
 
-However, if we only want to find rule interaction effects, then using
-the underspecified readings from the grammar makes the task faster,
-and it will still catch potential interaction errors.
+%In fact, it may even catch interaction conflicts between rules which have an internal conflict
 
@@
 The next step in creating realistic \emph{ambiguities} is to constrain which readings can go together.
-For instance, the case of \emph{zijn} shows us that ``determiner or verb'', is a possible ambiguity.
+For instance, the case of \emph{zijn} shows us that ``determiner or verb'' is a possible ambiguity.
 In contrast, there is no word form in the lexicon that would be ambiguous between an adjective and a comma, hence we do not want to generate such ambiguity in our symbolic sentences.
@@
 and masculine plural past participles.
-Then we form SAT clauses that allow or prohibit certain combinations. These clauses will interact with the constraints created from the rules, and the end result will be closer to real-life sentences.
+Then we form SAT-clauses that allow or prohibit certain combinations. These clauses will interact with the constraints created from the rules, and the end result will be closer to real-life sentences.
 

### 2016-03-10 paper/lrec/eval_futurework.tex
more last-minute changes

@@
 The SAT-solver will not mark it as a conflict (which is the correct behaviour).
-But if we ask it to generate an example sequence, the target word may be either of the following options:
+But if we ask it to generate an example sequence, the target word may be either of the following options.
+Seeing interpretation a) could then direct the grammar writer to modify the rule.
 
 \begin{itemize}
-\item[a.] \begin{verbatim}
+\item[a)] \begin{verbatim}
 "w2"
@@
 
-\item[b.] \begin{verbatim}
+\item[b)] \begin{verbatim}
 "w2"
@@
 commented out two rules that used regular expressions, because we did not implement the support for them yet.
+For a full list of found conflicts, see the annotated log of running our program in \url{
+https://github.com/inariksit/cgsat/blob/master/data/spa/conflicts.log}.
 
@@
 One of the conflicts that was only caught by the ambiguity class constraints had the condition \texttt{IF (1 Comma) (..) (1 CnjCoo)}.
-The additional constraints correctly prevent a comma from being ambiguous with anything else.
+The additional constraints correctly prevent commas from being ambiguous with anything else.
 
@@
 \end{itemize}
-
-For a full list of found conflicts, see the annotated log\footnote{\url{https://github.com/inariksit/cgsat/raw/master/data/spa/conflicts.log}} of running our program.
 

### 2016-03-10 paper/lrec/eval_futurework.tex
changes to future work

@@
 
-The first two variants contain a high number of internal conflicts. 
+The first two variants reported a high number of internal conflicts. 
 These are almost all due to nonexisting tags.
@@
 The running time of the grammars ranges from seconds to hours. 
-As can be seen from the times, increasing the number of readings
-matters more than increasing the number of rules.
-In addition, the ambiguity class constraints increase the time.
-The times in the Finnish table are not entirely comparable with each other: we were forced to run the tests in smaller batches, and it is possible that there are different overheads, unrelated to the size of the SAT-problem, from testing 50 or 500 rules at a time.
+Note that the times in the Finnish table are not entirely comparable with each other: we were forced to run the tests in smaller batches, and it is possible that there are different overheads, unrelated to the size of the SAT-problem, from testing 50 or 500 rules at a time.
+Despite the inaccuracies, we can see that increasing the number of readings 
+and adding the ambiguity class constraints
+slow the program down significantly.
+%In order to understand the effect of each feature more precisely, we should run our tests on grammars that differ on only one of the variables.
 
@@
 complex morphology.
-A more scalable solution would be to make each tag a variable, and ask the
-question ``can this reading be a noun? how about singular? how about
-conditional?'' separately for each tag. Then we could lift the
-restriction of tag combinations into a SAT solver: make SAT clauses
-that prohibit a comparative to go with a verb, or conditional with a noun.
+Currently, if we add one new lemma to the set of readings, we need to
+create as many new variables as there are inflectional forms
+for that lemma. %easily hundreds for languages with rich morphology.
 
@@
 forms directly to the representation of the possible readings.
-Currently, if we want to add one more lemma for a verb, we need to
-create as many new variables as there are distinct word forms---easily
-hundreds for languages with rich morphology.
+A possible solution would be to make each tag a variable, and ask the
+question ``can this reading be a noun? singular? 
+conditional?'' separately for each tag. Then we could lift the
+restriction of tag combinations into the SAT side: make SAT-clauses
+that prohibit a comparative to go with a verb, or conditional with a noun.
+Alternatively, we can still hardcode the set of morphological readings, and 
+only use SAT-clauses to restrict which lexical form can go with which morphological analysis.
 

### 2016-11-02 AmbiguityClass.hs
remove submodule

@@ -10,27 +10,27 @@ import SAT.Named
 --------------------------------------------------------------------------------
 -- example use
 
--- main :: IO ()
--- main =
---   do s <- readFile "spa-ambiguity-classes"
---      let ls   = lines s
---          xss  = map read ls :: [[Int]]
---          univ = S.toList . S.fromList . concat $ xss
+ --main :: IO ()
+ --main =
+ --  do s <- readFile "spa-ambiguity-classes"
+ --     let ls   = lines s
+ --         xss  = map read ls :: [[Int]]
+ --         univ = S.toList . S.fromList . concat $ xss
      
---      -- compute the formula once
---      putStrLn "Computing formula..."
---      let p = formula xss
---      (p == p) `seq` putStrLn "Done."
+ --     -- compute the formula once
+ --     putStrLn "Computing formula..."
+ --     let p = formula xss
+ --     (p == p) `seq` putStrLn "Done."
 
---      -- reuse it many times
---      putStrLn "Generating constraints..."
---      s  <- newSolver
---      qs <- sequence [ newLit s | y <- univ ]
---      let tab  = M.fromList (univ `zip` qs)
---          mp y = fromJust (M.lookup y tab)
---      constraints s mp [] p
---      deleteSolver s
---      putStrLn "Done.
+ --     -- reuse it many times
+ --     putStrLn "Generating constraints..."
+ --     s  <- newSolver
+ --     qs <- sequence [ newLit s | y <- univ ]
+ --     let tab  = M.fromList (univ `zip` qs)
+ --         mp y = fromJust (M.lookup y tab)
+ --     constraints s mp [] p
+ --     deleteSolver s
+ --     putStrLn "Done."
 
 test :: IO ()
 test =

### 2016-11-02 CG_SAT_old.hs

@@ -151,9 +151,9 @@ applyRule :: Solver
           -> [Token]      -- ^ Text to disambiguate
           -> Rule         -- ^ Rule to apply 
                           -- returns a list of triples:
-          -> IO [(Rule,   -- ^ * the rule with the clauses it has generated
-                [Clause], -- ^ * list of clauses
-                [Lit])]   -- ^ * debug help output; helper variables created for conditions
+          -> IO [(Rule,   -- the rule with the clauses it has generated
+                [Clause], --  list of clauses
+                [Lit])]   --  debug help output; helper variables created for conditions
 applyRule s toks rule = do
   case rule of 
     (Remove _name target conds) 

### 2016-11-02 RuleInduction.hs

@@ -1,6 +1,6 @@
 module Main where
 
-import Prelude hiding ( nub )
+import Prelude hiding ( nub, Word )
 import Data.Char
 import Data.List hiding ( nub )
 import qualified Data.Set as S
@@ -10,8 +10,8 @@ import Data.Map( Map )
 import System.IO
 
 file, gold :: FilePath
-file = "data/es.tagged.ambiguous"
-gold = "data/es.tagged"
+file = "data/spa/20k.tagged.ambiguous"
+gold = "data/spa/20k.tagged"
 
 evens (x:y:xs) = x : evens xs
 evens xs       = xs

### 2016-11-02 executable/Disambiguate.hs

@@ -45,8 +45,8 @@ main = do
       rules <- concat `fmap` map reverse `fmap` readRules grfile
       text <- readData txtfile
 
-      --let sortedText = sortBy (\a b -> length a `compare` length b) text
-      --mapM_ (\x -> putStr $ (show (length x) ++ "\n\t" ++ show x++ "\n")) (reverse $ sortedText)
+      let sortedText = sortBy (\a b -> length a `compare` length b) text
+      mapM_ (\x -> putStr $ (show (length x) ++ "\n\t" ++ show x++ "\n")) (take 5 $ reverse $ sortedText)
 
       allrds <- readReadings rdsfile 
       --let allrds = concat $ concat text

### 2016-11-02 test/Test.hs

@@ -46,15 +46,15 @@ main = do
     [r,"reverse"]   -> rev r spa20k spa20kgold spaPre
     ["enreverse"]   -> rev engr entext engold enpre
     ["opticomp",n]  -> optiComp spaSmall spa20k (read n)
-    ("gold":"orig":o) -> gold' ("par" `elem` o) spaFull spa20k spa20kgold
+    ("gold":"orig":o) -> gold' ("max" `elem` o) spaFull spa20k spa20kgold
     ["gold","opti"] -> opti spaSmall spa20k 
     ["gold","obs"]  -> optiBySz spaSmall spa20k
     ["gold"]        -> gold spaSmall spa20k spa20kgold
-    ["gold","par"]  -> goldPar spaSmall spa20k spa20kgold
+    ["gold","max"]  -> goldPar spaSmall spa20k spa20kgold
     ["goldrus"]     -> gold rusgr rustext undefined
     ["engold"]      -> gold engr entext engold
     ["nldgold"]     -> gold nldgr nldtext nldgold
-    (r:"gold":o)    -> gold' ("par" `elem` o) r spa20k spa20kgold --specify grammar
+    (r:"gold":o)    -> gold' ("max" `elem` o) r spa20k spa20kgold --specify grammar
 
     (r:d:o) -> do rules <- readRules r
                   text <- readData d
@@ -75,7 +75,8 @@ main = do
                     prAll "" resSAT resVISL text verbose
                     putStrLn ""
 
-    _          -> putStrLn "usage: ./test <rules> <data> (or something else, check the source code test/Test.hs >__>)"
+    _          -> do putStrLn "usage: one of the following"
+                     mapM_ putStrLn ["<rules> [gold | <data>] [par run sec]"]
 
 
 goldPar rl dt g = gold' True rl dt g

### 2016-12-23 old/december2016/CG_base.hs
restructuring

@@ -97,16 +97,16 @@ data TagSet =
   deriving (Eq,Ord)
 
 instance Show TagSet where
-  show (TS tags) = showReadinget tags
+  show (TS tags) = showReadings tags
   show (Or ts1 ts2) = show ts1 ++ "|" ++ show ts2
   show (Diff ts1 ts2) = show ts1 ++ " - " ++ show ts2 
 --  show (Cart ts1 ts2) = show $ map concat $ sequence [(toTags' ts1), (toTags' ts2)] 
   show (Cart ts1 ts2) = show ts1 ++ " + " ++ show ts2
   show All = "(*)"
 
-showReadinget :: [[Tag]] -> String
-showReadinget [[x]] = show x 
-showReadinget xs    = intercalate "|" $ map show' xs
+showReadings :: [[Tag]] -> String
+showReadings [[x]] = show x 
+showReadings xs    = intercalate "|" $ map show' xs
   where show' [y] = show y 
         show' ys  = "(" ++ unwords (map show ys) ++ ")"
 
@@ -245,21 +245,25 @@ instance Show Position where
   show (AtLeast Careful i) = "*" ++ show i ++ "C"
   show (Exactly NotCareful i) = show i
   show (AtLeast NotCareful i) = "*" ++ show i
-  show (LINK pos1 linkedpos) = show pos1 ++ " LINK " ++ show linkedpos
-  show p                 = let (a,b) = showPosTuple p in a++b
+  show (LINK pos1 linkedpos)  = show pos1 ++ " LINK " ++ show linkedpos
+  show p = let (a,b) = showPosTuple p in a++b
 
 showPosTuple :: Position -> (String, String)
-showPosTuple (Barrier Careful  NotCareful i ts) = ("*" ++ show i ++ "C ", " BARRIER " ++ show ts)
-showPosTuple (Barrier Careful  Careful  i ts) = ("*" ++ show i ++ "C ", " CBARRIER " ++ show ts)
-showPosTuple (Barrier NotCareful NotCareful i ts) = ("*" ++ show i ++ " ", " BARRIER " ++ show ts)
-showPosTuple (Barrier NotCareful Careful  i ts) = ("*" ++ show i ++ " ", " CBARRIER " ++ show ts)
+showPosTuple (Barrier Careful    NotCareful i ts) = ( "*" ++ show i ++ "C " 
+                                                    , " BARRIER " ++ show ts)
+showPosTuple (Barrier Careful    Careful  i ts)   = ( "*" ++ show i ++ "C "
[…]

### 2016-12-23 old/december2016/CG_parse.hs

@@ -23,6 +23,7 @@ import Control.Applicative
 import Control.Monad.State.Lazy
 import Data.Either
 import Data.List
+import Data.Maybe
 import Debug.Trace
 import Text.Regex ( mkRegex )
 
@@ -160,42 +161,37 @@ strip n = drop n . reverse . drop n . reverse
 transSetDecl :: SetDecl -> State Env (String, CGB.TagSet)
 transSetDecl (Set setname tagset) = 
   case setname of
-    BOS          -> return (">>>", CGB.TS bos)
-    EOS          -> return ("<<<", CGB.TS eos)
-    (SetName (UIdent name)) -> do 
-      ts <- transTagSet tagset
-      return (name, ts)
-    (SetMeta (UIdent name)) -> do 
-      ts <- transTagSet tagset
-      return ("<" ++ name ++ ">", ts)
-    (SetSynt (UIdent name)) -> do 
-      ts <- transTagSet tagset
-      return ("@" ++ name, ts)
+    BOS -> return (transSetName BOS, CGB.TS bos)
+    EOS -> return (transSetName EOS, CGB.TS eos)
+    nm  -> (,) (transSetName nm) `fmap` transTagSet tagset
 
 transSetDecl (List setname tags) = 
   case setname of
-    BOS          -> return (">>>", CGB.TS bos)
-    EOS          -> return ("<<<", CGB.TS eos)
-    (SetName (UIdent name)) -> do 
-      tl <- mapM transTag tags
-      let tl' = concatMap toTagsLIST tl
-      return (name, CGB.TS tl')
-    (SetMeta nm) -> transSetDecl $ List (SetName nm) tags
-    (SetSynt nm) -> transSetDecl $ List (SetName nm) tags
+    BOS -> return (transSetName BOS, CGB.TS bos)
+    EOS -> return (transSetName EOS, CGB.TS eos)
+    nm  -> do taglist <- mapM transTag tags
+              let tl' = concatMap toTagsLIST taglist
+              return (transSetName nm, CGB.TS tl')
 
-    --(SetMeta (UIdent name)) 
-    --  tl <- mapM transTag tags
-    --  let tl' = concatMap toTagsLIST tl
-    --  return (name, CGB.TS tl')
-    --(SetSynt (UIdent name)) -> do 
-    --  tl <- mapM transTag tags
-    --  let tl' = concatMap toTagsLIST tl
-    --  return (name, CGB.TS tl')      
[…]

### 2016-12-23 doc/lrec/eval_futurework.tex
renaming

--- 2016-03-10 paper/lrec/eval_futurework.tex
@@
 As for longer-term goals, we want to handle the full expressivity of CG-3,
-with \textsc{map}, \textsc{add},  \textsc{addreading} and
+with \textsc{map}, \textsc{add},  \textsc{append} and
 \textsc{substitute} rules, as well as dependency structure. 
@@
 Even if the lexicon gives all readings that exist in the lexicon, the
-user might give a nonexistent reading, or in the case of MAP, a
+user might give a nonexistent reading, or in the case of {\sc map}, a
 syntactic tag, which is (by definition) not in the lexicon. We may need to move

### 2016-12-23 doc/nodalida/_region_.tex
renaming

--- 2015-03-30 paper/nodalida2015.tex
@@
-%
+\message{ !name(nodalida2015.tex)}%
 % File nodalida2015.tex
@@
 \usepackage{tikz}
-\def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) --
-  (.25,.15) -- cycle;}
-\newcommand{\Mypm}{\mathbin{\tikz [x=1.4ex,y=1.4ex,line width=.1ex] \draw (0.0,0) -- (1.0,0) (0.5,0.08) -- (0.5,0.92) (0.0,0.5) -- (1.0,0.5);}}%
-
+\def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;}
 \title{Constraint Grammar as a SAT problem}
 
-% \author{Inari Listenmaa \and Koen Claessen \\
-%  Chalmers University of Technology, Gothenburg, Sweden \\
-%  {\tt \{inari,koen\}@chalmers.se} }
+\author{Inari Listenmaa \and Koen Claessen \\
+  Chalmers University of Technology, Gothenburg, Sweden \\
+  {\tt \{inari,koen\}@chalmers.se} }
 
@@
 \begin{document}
+
+\message{ !name(nodalida2015.tex) !offset(-3) }
+
 \maketitle
@@
 \section{Introduction}
+%Constraint Grammar (CG) was first introduced by \cite{karlsson1995constraint}. 
+%It disambiguates output by morphological analyser, 
+%and has been used for many tasks in computational linguistics, such as POS tagging,
+%surface syntax and machine translation.
+%The current state-of-the-art parser VISL CG-3 is implemented (...);
+%in addition there are a number of finite-state based implementations \cite{nemeskey14}.
 
@@
 as a Boolean satisfiability (SAT) problem.
-This is attractive from several reasons: formal logic is
-well-studied, and serves as an abstract language to reason about the
-properties of CG. Constraint rules encoded in logic capture richer
-dependencies between the tags than standard CG.
-
-Applying logic to reductionist grammars has been explored earlier by \cite{lager98,lager_nivre01}, but it was never adopted for use.
-% ; logic programming was too slow to be used for tagging or parsing. 
-Since those works, SAT solving techniques have improved significantly, and 
-they are used in domains such as microprocessor design and computational 
-biology---these problems easily match or exceed CG in complexity. 
-% Is this true for grammars with thousands of rules?
-
-Due to these advances, we were able to revisit the idea and develop it further. 
-% Whereas \cite{lager98} provides a proof-of-concept description, 
-Missing from the previous logic-based works, 
-we solve eventual rule conflicts by finding a solution that 
-discards the least number of rules.
-% allows the maximal number of rules to be applied.
-We test our implementation by parsing texts in the order of 10,000s--100,000s
-words, using grammars with hundreds of rules.
-
+Applying logic to reductionist grammars has been explored earlier by 
+\cite{lager98,lager_nivre01}, but it was never adopted for use---logic programming was simply 
+too slow to be used for tagging or parsing. 
+However, modern SAT solvers are fast and used for many applications where constraints must be
+satisfied, such as software and hardware verification.
 %SAT solving is based on a technique called unit propagation:
@@
 %solution, where all variables in the set have a True or False value.
+
+It is attractive to model CG in terms of logic:
+these techniques are well-studied, and they capture more dependencies between the tags than the standard CG.
 
@@
 and at least one analysis for each word is true---the notion of ``last'' is not applicable.
-With these constraints, we get two solutions: 
-\begin{enumerate}
-\item []
-\begin{verbatim}
-"<la>"
-        "el" det def f sg
-"<casa>"
-        "casa" n f sg
-\end{verbatim}
-\item [and]
-\begin{verbatim}
-"<la>"
-        "lo" prn p3 f sg
-"<casa>"
-        "casar" v pri p3 sg
-        "casar" v imp p2 sg
-\end{verbatim}
-\end{enumerate} 
-% \texttt{det, n} and \texttt{prn, v pri p3, v imp p2}. 
+With these constraints, we get two solutions: \texttt{det, n} and \texttt{prn, v pri p3, v imp p2}. 
 The interaction of the rules regarding \emph{la}  disambiguates the POS of \emph{casa} 
@@
 \paragraph{Rules are unordered.}
+%A typical CG rule sequence starts with rules that are either very common, or have a complex %condition to fulfil.
+%Stronger rules appear towards the end, to be applied if none of the weaker rules has. 
+%For a rule set such as following, this order is crucial:
+%\begin{itemize}
+%\item [] \texttt{SELECT Year IF (-1 Seasons OR Months) ;} \\
+ %            \texttt{REMOVE Year ;}
+%\end{itemize}
 
@@
 \end{itemize}
+
+%combined with the following
+%\begin{itemize}
+%\item Maximise number of true analyses---Without this heuristic, both solutions to our example %\emph{la casa} are as likely.
+With this heuristic,  the example would be disambiguated as \texttt{prn v}, because there happens to be two analyses with \texttt{v}, and choosing that, 3 analyses are true, since there is no rule to disambiguate between imperative and indicative.
+% We want analyses to be true by default, but still disambiguate as much as possible.
+%We have a rule to maximise the number of true analyses; then we take a solution that has the least. 
+%\item Minimise number of true analyses---To combat the previous, we want each solution to contain as many true analyses as possible, to ensure that analyses untouched by rules are kept true.
+%After having multiple solutions which all ensure , we pick the one that disambiguates the most.
+%\end{itemize}
+
 The chosen heuristic produced the best results with examples in the order of tens or hundreds of rules.
-Larger rule sets are yet to be tested.
+Larger rule sets (thousands of rules) are yet to be tested.
 
@@
 \paragraph{Rules disambiguate more.}
-Considering our example phrase and rules, standard CG implementation
-can only remove readings for the word which is tagged as \texttt{prn} or
-\texttt{det}. The SAT-based implementation interprets the rules as
-``determiner and verb together are illegal'', and is free to take action that concerns also the word in the condition \texttt{(1 n)} or \texttt{(1 v)}.
-This means that the rules are logically flipped: \texttt{REMOVE prn IF
-  (1 n)} translates into the same logical formula as  \texttt{REMOVE n
-  IF (-1 prn)}. A rule with more conditions can be written as many
-rules, each condition taking its turn to be in position to be removed
-or selected. \\
+% Rules interact with each other and disambiguate more than same rules in the standard implementation. 
+In standard CG, the rules only tell to remove readings for the word which is tagged as 
+\texttt{prn} or \texttt{det}. The SAT-based implementation interprets the rules as ``noun and pronoun together are illegal'', and is free to take action that concerns also the following word.
+This means that the rules are logically flipped: compare the rules below to the ones in the example.
 
-% \begin{itemize}
-% \item [] \texttt{REMOVE n IF (-1 prn) ;} \\
-%          \texttt{REMOVE v IF (-1 det) ;}
-% \end{itemize}
+\begin{itemize}
+\item [] \texttt{REMOVE n IF (-1 prn) ;} \\
+             \texttt{REMOVE v IF (-1 det) ;}
+\end{itemize}
 
-% SAT-based approach gives identical results with both sets of rules,
-% whereas the standard CG would remove one reading from \emph{casa} and leave \emph{la} ambiguous.
-% Testing this property with more complex rules and larger rule sets remains to be done.
+SAT-based approach gives identical results with both sets of rules,
+whereas the standard CG would remove one reading from \emph{casa} and leave \emph{la} ambiguous.
+Testing this property with more complex rules and larger rule sets remains to be done.
 
@@
 
-\noindent These three features might change the way that rules are written. 
-We predict that less rules are needed; whether this holds in
-the order of thousands of rules remains to be tested.
-Getting rid of ordering could ease the task of the grammar writer,
-removing the responsibility to estimate the best sequence of rule applications.
-However, the implementation should not be too different from the
-existing ones---our preliminary evaluation against VISL CG-3 shows
-promising results in this regard.
-\pagebreak
+These three features are likely to change the way that rules are written. \todo{more speculation, examples}
 
@@
 
-The worst-case complexity of SAT is exponential, whereas the standard
-implementations of CG are polynomial, but with advances in SAT solving
-techniques, the performance in the average case is much more feasible than in the previous works done in 90s--00s.
+The worst-case complexity of SAT is exponential, whereas the standard implementations 
+of CG are polynomial, but with advances in SAT solving techniques, the performance 
+in the average case is much more feasible than in the previous works done in 90s--00s.
 We used the open-source SAT solver MiniSat \cite{een04sat}.
+%We do foreach sentence: foreach rule; 
+%whereas VISL-CG3 has foreach rule: foreach sentence.
+%the size of the SAT problem is dependent of the number of rules
 
-Table~\ref{table:time} shows the execution time compared to VISL CG-3. 
-The number of rules in the grammar is more significant than the word
-count; this holds also for the performance of VISL CG-3.
-From the SAT solving side, maximisation is the most costly operation.
-Parsing Don Quijote (384,155 words) with the rule set of 261 rules,
-the function was called 147,253 times, and with 19 rules, 132,255 times.
-%However, with both rule sets, half of the sentences in Don Quijote
-%needed less than 6 calls of maximise, and 75 \% of the sentences
-%needed less than 14.
-The difference in the execution times suggests that there are other
-reasons for the worse performance---this is to be expected,
-SATCG is just a first proof-of-concept implementation with no optimisations.
+Table~\ref{table:time} shows the execution time compared to VISL-CG3. 
+The number of rules in the grammar is 
+Maximisation is the most costly operation
+Also, some of the time is explained simply by
+SATCG is just a first proof-of-concept implementation and can be optimised.
 
@@
   \centering
-  \begin{tabular}{|l|c|c|c|}
+  \begin{tabular}{|l|l|c|c|}
      \hline
-   \textbf{\# words} & \textbf{\# rules} &  \textbf{SATCG} & \textbf{VISL CG-3} \\ \hline
+   \textbf{\# words} & \textbf{\# rules} &  \textbf{SATCG} & \textbf{VISL-CG3} \\ \hline
+                124,493  & 99    & 55.7s   & 4.8s \\  \hline	                         
                 384,155  & 19    &  13.7s  & 4.2s\\ %\hline
-                124,493  & 99   &  55.7s   & 4.8s \\ % \hline
                 384,155  & 130  &  1m14.1s & 6.1s \\ %\hline
@@
 
-\paragraph{Performance against VISL CG-3}
+\paragraph{Performance against VISL-CG3}
 
@@
 We only took select and remove rules (no substitute/iff), after which the constraint grammar had 261 rules.
-With this setup, we took the text of the tagged corpus with all ambiguities, and ran both VISL CG-3 and our implementation with the same grammar. 
-Treating the chosen tagset as the gold standard, the disambiguation by
-SATCG achieves 75.09 \% correct result and VISL CG-3 75.80 \%.
-Similar patterns were observed with small (\textless{}20) rule sets written by
-the authors; depending on the subset, SATCG and VISL CG-3 had
-a difference of at most $\Mypm$ 1.5 \%. 
-Introducing rules one by one up to 19, the
-performance improved in a very similar rate, with less than 0.5 \%
-difference between the systems at each new rule.
-
-% Additional tests could include plugging SATCG into Apertium
-% translation pipeline, and comparing the translation quality.
+With this setup, we took the text of the tagged corpus with all ambiguities, and ran both VISL-CG3 and our implementation with the same grammar. 
+Treating the chosen tagset as the gold standard, the disambiguation by SATCG achieves 75.09 \% correct result and VISL-CG3 75.80 \%.
 
@@
 %      \hline
-%     \textbf{\# rules} & \textbf{SATCG} & \textbf{VISL CG-3} \\
+%     \textbf{\# rules} & \textbf{SATCG} & \textbf{VISL-CG3} \\
 %     \hline 
@@
 % 	16 &  83.197 & 83.030 \\ \hline  % 16 out of 17/18, no diff
-% 	15 &  83.188 & 83.022\\ \hline  % 15 out of 16/17. Tie with 14 for SATCG, chose to retain the rule that gave better results for VISL CG-3.
+% 	15 &  83.188 & 83.022\\ \hline  % 15 out of 16/17. Tie with 14 for SATCG, chose to retain the rule that gave better results for VISL-CG3.
 % 	14 &  83.180 & 83.005 \\ \hline  % Tie with 15 for SATCG.
@@
 % \end{table}
-% In this experiment, we took 19 rules written by the authors, which give 83.15 \% correct results with SATCG and 83.33 \% with VISL CG-3.
-% Then we removed rules one by one; after getting a score for $n$ rules, we took the combination of $n-1$ rules for which SATCG gives the best score, and we ran the rule set with VISL CG-3 for comparison.
+% In this experiment, we took 19 rules written by the authors, which give 83.15 \% correct results with SATCG and 83.33 \% with VISL-CG3.
+% Then we removed rules one by one; after getting a score for $n$ rules, we took the combination of $n-1$ rules for which SATCG gives the best score, and we ran the rule set with VISL-CG3 for comparison.
 Table~\ref{table:results_remove} shows the percentages compared to the gold standard for both systems.
@@
 \end{document}
+
+\message{ !name(nodalida2015.tex) !offset(-290) }
